### 2022-09-08 main.tex
init
init

\documentclass{article}
\usepackage{listings, listings-rust}
\lstset{escapeinside={<@}{@>}}
\usepackage{minted}
\usepackage{tabularx}
\usepackage{array}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage[utf8]{inputenc}
\usepackage{multirow}

\title{Rust Bug Fix Patterns}
\author{Moh Robati}

\begin{document}

\maketitle

\section{Introduction}

Producing bugs has been an unavoidable part of computer programming which happens unintentionally. Researchers have been trying to help software engineers to be more aware of the bug producing practices in development and how to step towards creating bug-free software. One useful step towards that goal would be studying common bug patterns (and their corresponding fixes) that occur across multiple software systems. Research efforts for categorizing common bug patterns date back to 1975 where Edres tried to categorize various bug patterns which occurred in operating system program.

In this work, we empirically analyze the most popular Rust projects in GitHub, and discover the common bug patterns lied within them. We believe these insights are important for various reasons, primarily for developing program repair tools. Automated program repair tools try to locate bug
locations in the source code (they include a fault localization module) and then fix them, possibly without human intervention, in a manner such that the modified program meets the desired specifications. A powerful automated program repair system has many useful applications and can significantly simplify the process of debugging, resulting in a reduction in software development cost. The set of common bug fix patterns can be a useful input to an APR system as it directly provides information about fixing similar buggy structures, and that can significantly reduce the search space for patch generation.

Similarly, if we are not observing the desired behaviour from our program, that would indicate that there lies a fault within our program. Fault localization is the process of automatically finding the faulty statement(s) in our program. If a fault localization tool knows about the common bug patterns that can appear in the underlying programming language, it can prioritize the code segment candidates based on their potentials of producing bugs. That would result in a more robust tool as it leverages from historical data.

Moreover, knowing which code patterns in a programming language are more susceptible to produce bugs can help programmers to be more cautious while working with similar code structures. This knowledge can make the developers avoid certain mistakes during production which ultimately would result in reducing the frequent bug patterns within the program. As Monperrus describes, the less frequent bug patterns we keep in our software, the less cost we would have to pay for software maintenance.

To discuss about a less talked about use case, recently, researchers have been trying to apply deep learning techniques on programs to learn about the semantics of the code or training models which can modify programs towards a specific goal (e.g. bug fix). One challenge is that usually learning tools get fixed size vectors as input and map that to an output. Unlike image data, it is challenging to compress programs into a vector without any loss in semantic information. Current approaches try to define the vector dimensions in a way that accounts for all possible abstract syntax tree forms. That results in a large sparse space for the inputs. Assuming that the deep learning model we are trying to develop is associated with bug fixing tasks, then knowledge about common bug fix patterns can help with pruning input space towards containing less dimensions. That can help with creating more efficient models.

There have been many studies on detecting and categorizing bug fix patterns in general, or for a specific programming language. We are more interested in the latter, since
our study is focused on pervasive bug fix patterns in Rust. \\
Yang et al. proposed a mining approach to detect Python bug fix patterns via studying
fine-grained fixing code changes. In their research, they also examined what portion of the bugs
could be fixed using automated bug fixing approaches. Moreover, they evaluated the fix patterns in
the wild and realized that 37\% of the buggy codes could be matched by the fix patterns they had
found. \\
Hanam et al. conducted a similar research but for finding Javascript most pervasive bug fix
patterns. They presented a large-scale study of the most common bug fix patterns by mining 105K
commits from 134 server-side JavaScript projects. They used DBSCAN clustering algorithm
and divided bug fixing change types into 219 clusters, from which they extracted 13 pervasive
cross-project bug fix patterns. \\
In this work, we study common bug patterns that appear in Rust projects. Rust is a relatively new language that allows programmers to write programs that have low-level
control over resources, as well as high-level safety guarantees. Originally, Rust was a side
project developed by Graydon Hoare, a Mozilla employee. The company started to fund this project
in 2009, since they intended to use Rust in some of the low-level Firefox components.
What distinguishes Rust from the other systems programming languages, e.g. C/C++, is the
ownership framework that has been incorporated inside the language. The ownership framework keeps the track of the references to different memory locations, and will free them once they
go out of scope. If the program passes the compiler checks, it automatically guarantees several
safety problems e.g. dangling references, or double-free problem, that had to be taken care of
manually in languages like C/C++. Moreover, not only the ownership concept removes the
need for a garbage collector in the runtime environment, it does not require the programmer to
explicitly do certain safety tasks, like freeing references.
Due to the memory and thread safety guarantees that Rust provides, many software companies
adopted Rust in production. For example, Dropbox decided to migrate to Rust for developing their
storage system, since they were not completely satisfied with the features provided by AWS’s S3 or
Golang.

The main module which governs over the ownership rules is called Borrow Checker. Because of its importance in Rust, we have decided to put discovered patterns into two groups; (1) non-borrow-checker related patterns (General Patterns), and (2) borrow-checker related patterns (BC-Related Patterns). This distinction was really helpful as it made our tool discover clearer results.

We present our discovered insights of our empirical study in the following four research
questions (RQs):

\textbf{RQ1. How successful is our data modelling approach in capturing all the necessary data?}

\textbf{RQ2. What are the most common AST structures in the bug fixing changes?}

\textbf{RQ3. What are the general fix patterns in Rust?}

\textbf{RQ4. Which fix patterns are related to Borrow Checker in Rust?}

\section{Pervasive Bug Patterns}
\subsection{Attributes}

Attributes in Rust make it possible for the developers to add metadata to program elements. These attributes can be added to structs, struct fields, enum variants, match expression arms, etc. There are two types of attributes: Inner Attributes and Outer Attributes. Inner Attributes apply to the program element encapsulating them, while Outer Attributes apply to the program element coming after the attribute.

\subsubsection{G1. Modifying the attributes of structs}

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
// G1.1: Adding attributes to a struct
++ #[derive(Debug, Clone)]
pub struct Call {
/// body 
}

// G1.2: Removing attributes from a struct
-- #[derive(Debug, Clone)]
pub struct Call {
/// body 
}

// G1.3: Changing an existing attribute of a struct
-- #[derive(Debug)]
++ #[derive(Debug, Clone)]
pub struct Call {
/// body 
}
\end{lstlisting}

\noindent \textbf{Description:} In this pattern, the attribute set of a struct changes. This modification can happen in three different ways; (G1.1) Adding attributes to the struct, (G1.2) Removing attributes from the struct, (G1.3) Changing the content of an existing attribute. The attributes are Outer Attributes as we can find struct related non-terminals within the ASTDiff of the instances of this pattern.

\subsubsection{G2. Modifying the attributes of struct fields}

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
// G2.1: Adding attributes to struct fields
#[derive(Serialize)]
pub struct Retain {
    ++ #[serde(skip_serializing_if = "is_empty")]
    pub attributes: Option<Attributes>,
}

// G2.2: Removing attributes from struct fields
#[derive(Serialize)]
pub struct Retain {
    -- #[serde(skip_serializing_if = "Option::is_none")]
    pub attributes: Option<Attributes>,
}

// G2.3: Changing attributes of struct fields
#[derive(Serialize)]
pub struct Retain {
    -- #[serde(skip_serializing_if = "Option::is_none")]
    ++ #[serde(skip_serializing_if = "is_empty")]
    pub attributes: Option<Attributes>,
}
\end{lstlisting}

\noindent \textbf{Description:} In this pattern, the attribute set of a struct field changes. Similar to G1, this modification also happens in three different ways. Using attributes for struct fields, apart from being able to store different meta information for different struct fields, for some use cases, the programmers can control the action they want to apply on different fields.

\subsection{Struct}
\subsubsection{G3. Adding/Removing a Struct field}

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
// G3.1: Adding a new field to a struct
struct InnerListeners {
    pending: Mutex<Vec<Pending>>,
    <@\textcolor{red}{--}@> queue_object_name: Uuid,
}

// G3.2: Removing a field from a struct
struct SnapshotService<U, R> {
    uuid_resolver_handle: R,
    <@\textcolor{green}{++}@> db_name: String,
}
\end{lstlisting}

\noindent \textbf{Description:} A set of new fields are add to or removed from an existing struct. The addition of a new field happens when the developer need to store a new piece of data so they can implement a desired behaviour. The removal of a field  happens when the developer realizes that the field is expendable with respect to the behaviour they want to implement.

\subsubsection{G4. Modifying the access modifiers of struct fields}

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]

// G4.1: Adding pub to a field
pub struct HtmlBlock {
    -- content: BlockContent,
    ++ pub content: BlockContent,
    brace: token::Brace,
}

// G4.2: Removing pub from a field
pub struct Languages {
--  pub named: HashMap<LanguageId, Arc<LanguageDefinition>>,
++  named: HashMap<LanguageId, Arc<LanguageDefinition>>,
    extensions: HashMap<String, Arc<LanguageDefinition>>,
}

\end{lstlisting}

\noindent \textbf{Description:} Adding pub to a struct field makes it possible to access that field within external modules. In the instances of this cluster, adding pub access modifier happened in a bug fixing change, where the developer needed to access a field which had not been granted public access. On the other hand, the public access might be revoked if the developer realizes there is no external access to the field.

\subsubsection{G5. Different fields order when initializing Struct}

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-- return Poll::Ready(ReadyEvent { ready, tick });
++ return Poll::Ready(ReadyEvent { tick, ready });
\end{lstlisting}

\subsection{Option}
\subsubsection{G6. Changing field type T to T$\langle$Option$\rangle$}

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]

#[ast_node("MediaRule")]
pub struct MediaRule {
    pub span: Span,
    -- pub media: MediaQueryList,
    ++ pub media: Option<MediaQueryList>,
    pub rules: Vec<Rule>,
}

\end{lstlisting}

\subsection{Types}
\subsubsection{G7. Changing a struct field type}

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]

pub struct Manifest {
  -- pub features: Vec<String>,
  ++ pub features: HashSet<String>,
}

\end{lstlisting}

\subsubsection{G8. Changing a generic type}

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]

#[derive(Debug, Clone, PartialEq)]
pub struct Anchor {
    -- point: Point<usize>,
    ++ point: Point<isize>,
    side: Side,
}

\end{lstlisting}

\subsubsection{G9. Change in enum variant value type}

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]

#[derive(Debug, Clone, PartialEq, Eq)]
pub enum ResolvedDependency {
  Resolved(ModuleSpecifier),
  -- Err(String),
  ++ Err(ResolvedDependencyErr),
}

\end{lstlisting}

\subsubsection{G10. Changing the type inside a tuple}

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]

#[allow(dead_code)]
pub struct CoreState {
    // Other fields
    -- pending_views: Vec<ViewId>,
    ++ pending_views: Vec<(ViewId, Table)>,
    peer: Client
}

\end{lstlisting}

\subsection{Traits}
\subsubsection{G11. TraitBound Change}

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]

-- pub struct EventLoop<T: tty::EventedReadWrite> {
++ pub struct EventLoop<T: tty::EventedPty> {
    poll: mio::Poll,
    // Other fields
}

\end{lstlisting}

\subsubsection{G12. Implementing a different trait for a struct}

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]

-- impl ToDebug for ShellError {
++ impl FormatDebug for ShellError {
        // Implementing functions ...
   }

\end{lstlisting}

\subsection{Ref and Deref}
\subsubsection{B1. Adding $\&$}

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
let repo = replace(&mut *contents, Processing).inner_repo();
-- let statuses = repo_to_statuses(repo, &self.workdir);
++ let statuses = repo_to_statuses(&repo, &self.workdir);
\end{lstlisting}

\subsubsection{B2. Removing $\&$}

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
enums.iter().for_each(|e| {
    let mut enum_template = EnumTemplate::new();
    -- enum_template.set_message_enum(&e);
    ++ enum_template.set_message_enum(e);
    // Other statements
});
\end{lstlisting}

\subsubsection{B3. Dropping & from func decl and adding * to match statement}

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-- fn to_color_choice(&self) -> ColorChoice {
--     match *self {
++ fn to_color_choice(self) -> ColorChoice {
++     match self {
           // Statements
       }
\end{lstlisting}

\subsubsection{B4. Adding *}

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
for p in wixobjs {
--  args.push(p.to_string());
++  args.push((*p).to_string());
}

let displayed_attributes = schema.clone().map(|s| {
    s.displayed_name()
        .iter()
--      .map(|s| (*s).to_string())
++      .map(|s| s.to_string())
        .collect::<HashSet<String>>()
});
\end{lstlisting}

\subsection{Mut}
\subsubsection{B5. Adding mut}

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]

parser::Parser::new(args)
        .parse_module()
--      .map_err(|e| {
++      .map_err(|mut e| {
            e.emit();
            ()
        })?

\end{lstlisting}

\subsubsection{B6. Removing mut}

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-- let mut tx = tx.clone();
++ let tx = tx.clone();
\end{lstlisting}

\subsection{Vector}
\subsubsection{B7. Vector reference to slice}

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
pub fn expand_delimited_square(
--  children: &Vec<TokenNode>,
++  children: &[TokenNode],
) -> Result<hir::Expression, ParseError> {
    // body
}
\end{lstlisting}

\subsection{Lifetime}
\subsubsection{B8. Adding generic Lifetime}
\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
#[allow(dead_code)]
-- fn api_error(webview: &mut Webview, error_fn: String, message: &str) {
++ fn api_error(webview: &mut Webview<'_>, error_fn: String, message: &str) {
  let reject_code = tauri_api::rpc::format_callback(error_fn, message);
  webview.eval(&reject_code)
}

\end{lstlisting}

\subsubsection{B9. Dropping static lifetime}

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]

-- const QUEUE_SIZES: &'static [u16] = &[QUEUE_SIZE];
++ const QUEUE_SIZES: &[u16] = &[QUEUE_SIZE];

\end{lstlisting}

\subsubsection{B10. Dropping Lifetime in field definition}

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
pub struct Table<'a> {
    columns: Vec<Column>,
    colours: &'a Colours,
    env: &'a Environment,
    widths: TableWidths,
    -- time_format: &'a TimeFormat,
    ++ time_format: TimeFormat,
    size_format: SizeFormat,
    git: Option<&'a GitCache>,
}
\end{lstlisting}

\subsection{Clone}
\subsubsection{B11. Dropping clone and adding dereference}

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
pub fn reset(&mut self, template: &Cell) {
    // memcpy template to self
--  *self = template.clone();
++  *self = *template;
}
\end{lstlisting}

\subsubsection{B12. Dropping clone}

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
if let Some(var_id) = &var.var_id {
--  stack.add_var(*var_id, input);
++  stack.add_var(*var_id, input.clone());
}

start_plugin_process(
--  manifest.clone(),
++  manifest,
    self.next_plugin_id(),
    self.self_ref.as_ref().unwrap().clone(),
);

\end{lstlisting}

\subsubsection{B13. Dropping clone and adding reference}

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-- let field_id = schema.get_or_create(attribute.clone())?;
++ let field_id = schema.get_or_create(&attribute)?;
\end{lstlisting}

\begin{table}[]

\hskip-2.5cm\begin{tabular}{|l|l|l|l|}
\hline

\multirow{12}{*}{General}        
& \multirow{2}{*}{Attributes}    & G1  & Adding Attributes to Structs                                        \\
                                 &                                & G2  & Adding Attributes to Struct Fields                                  \\\cline{2-3}
                                 & \multirow{3}{*}{Struct}        & G3  & Adding/Removing a Struct field                                      \\
                                 &                                & G4  & Adding/Removing pub to/from struct field                            \\
                                 &                                & G5  & Different fields order when initializinng Struct                       \\\cline{2-3}
                                 & Option                         & G6  & Changing field type T to Option                                          \\\cline{2-3}
                                 & \multirow{4}{*}{Types}         & G7  & Changing a Struct field type                                        \\
                                 &                                & G8  & Changing a generic type                                             \\
                                 &                                & G9  & Change in enum variant value type                                   \\
                                 &                                & G10 & Changing the type inside a tuple                                         \\\cline{2-3}
                                 & \multirow{2}{*}{Traits}        & G11 & TraitBound Change                                                   \\
                                 &                                & G12 & Implementing a different trait for a struct.                            \\\cline{1-3}
\multirow{13}{*}{Borrow Checker} & \multirow{4}{*}{Ref and Deref} & B1  & Adding \&                                                           \\
                                 &                                & B2  & Removing \&                                                         \\
                                 &                                & B3  & Dropping \& from function declaration and dereferencing match \\
                                 &                                & B4  & Adding *                                                            \\\cline{2-3}
                                 & \multirow{2}{*}{Mut}           & B5  & Adding mut                                                          \\
                                 &                                & B6  & Removing mut                                                        \\\cline{2-3}
                                 & Vector                         & B7  & Vector reference to slice                                           \\\cline{2-3}
                                 & \multirow{3}{*}{Lifetime}      & B8  & Adding generic Lifetime                                             \\
                                 &                                & B9  & Dropping static lifetime                                            \\
                                 &                                & B10 & Dropping Lifetime in field definition                               \\\cline{2-3}
                                 & \multirow{3}{*}{Clone}         & B11 & Dropping clone and adding dereference                               \\
                                 &                                & B12 & Dropping clone                                                      \\
                                 &                                & B13 & Dropping clone and adding reference  & \hline

\end{tabular}
\caption{\label{fig:category}Bug Fix Pattern Categorization}
\end{table}

\begin{algorithm}
\caption{An algorithm with caption}\label{alg:cap}
\hspace*{2mm} \textbf{Input:} $R$ (target repositories)  \\
\hspace*{2mm} \textbf{Output:} $D_g$ (General code changes) \\
\hspace*{2mm} \textbf{Output:} $D_b$ (BC-related code changes)
\begin{algorithmic}
\State $D_g \leftarrow \phi$
\State $D_b \leftarrow \phi$
\For{$r \in R$}
    \For{$c \in \textsc{ExtractCommits}(r)$}
        \If{$c.msg$ contains bug fixing related keywords}
            \For{$\{f_b, f_a\} \in \textsc{GetModifiedRustFiles}(c)$}
                \For{$e \in \textsc{ASTDiff}(\textsc{Parse}(f_b), \textsc{Parse}(f_a))$}
                    \State $DP \leftarrow \textsc{GetDataPoint}(e)$
                    \State $DP \leftarrow c.hash \cup f_a.name \cup e.scope \cup DP $
                    \If{$\textsc{IsBCRelated}(e)$}
                        \State $DP \leftarrow \textsc{GetBCKeyword}(e) \cup DP $
                        \State $D_b \leftarrow D_b \cup DP$
                    \Else
                        \State $D_g \leftarrow D_g \cup DP$
                    \EndIf
                \EndFor
            \EndFor
        \EndIf
    \EndFor
\EndFor
\end{algorithmic}
\end{algorithm}

\end{document}

### 2022-09-08 main.tex
fix compilation errors

@@
 
-\subsubsection{B3. Dropping & from func decl and adding * to match statement}
+\subsubsection{B3. Dropping \& from func decl and adding * to match statement}
 
@@
                                  &                                & B12 & Dropping clone                                                      \\
-                                 &                                & B13 & Dropping clone and adding reference  & \hline
+                                 &                                & B13 & Dropping clone and adding reference
 

### 2022-09-12 1_introduction.tex
added some descriptions for results and reorganized files

--- 2022-09-08 main.tex
@@
-\documentclass{article}
-\usepackage{listings, listings-rust}
-\lstset{escapeinside={<@}{@>}}
-\usepackage{minted}
-\usepackage{tabularx}
-\usepackage{array}
-\usepackage{graphicx}
-\usepackage{algorithm}
-\usepackage{algpseudocode}
-\usepackage[utf8]{inputenc}
-\usepackage{multirow}
-
-\title{Rust Bug Fix Patterns}
-\author{Moh Robati}
-
-\begin{document}
-
-\maketitle
-
-\section{Introduction}
-
 Producing bugs has been an unavoidable part of computer programming which happens unintentionally.
@@
 \textbf{RQ4. Which fix patterns are related to Borrow Checker in Rust?}
-
-\section{Pervasive Bug Patterns}
-\subsection{Attributes}
-
-Attributes in Rust make it possible for the developers to add metadata to program elements.
-These attributes can be added to structs, struct fields, enum variants, match expression arms, etc.
-There are two types of attributes: Inner Attributes and Outer Attributes.
-Inner Attributes apply to the program element encapsulating them, while Outer Attributes apply to the program element coming after the attribute.
-
-\subsubsection{G1. Modifying the attributes of structs}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-// G1.1: Adding attributes to a struct
-++ #[derive(Debug, Clone)]
-pub struct Call {
-/// body 
-}
-
-// G1.2: Removing attributes from a struct
--- #[derive(Debug, Clone)]
-pub struct Call {
-/// body 
-}
-
-// G1.3: Changing an existing attribute of a struct
--- #[derive(Debug)]
-++ #[derive(Debug, Clone)]
-pub struct Call {
-/// body 
-}
-\end{lstlisting}
-
-\noindent \textbf{Description:} In this pattern, the attribute set of a struct changes.
-This modification can happen in three different ways; (G1.1) Adding attributes to the struct, (G1.2) Removing attributes from the struct, (G1.3) Changing the content of an existing attribute.
-The attributes are Outer Attributes as we can find struct related non-terminals within the ASTDiff of the instances of this pattern.
-
-\subsubsection{G2. Modifying the attributes of struct fields}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-// G2.1: Adding attributes to struct fields
-#[derive(Serialize)]
-pub struct Retain {
-    ++ #[serde(skip_serializing_if = "is_empty")]
-    pub attributes: Option<Attributes>,
-}
-
-// G2.2: Removing attributes from struct fields
-#[derive(Serialize)]
-pub struct Retain {
-    -- #[serde(skip_serializing_if = "Option::is_none")]
-    pub attributes: Option<Attributes>,
-}
-
-// G2.3: Changing attributes of struct fields
-#[derive(Serialize)]
-pub struct Retain {
-    -- #[serde(skip_serializing_if = "Option::is_none")]
-    ++ #[serde(skip_serializing_if = "is_empty")]
-    pub attributes: Option<Attributes>,
-}
-\end{lstlisting}
-
-\noindent \textbf{Description:} In this pattern, the attribute set of a struct field changes.
-Similar to G1, this modification also happens in three different ways.
-Using attributes for struct fields, apart from being able to store different meta information for different struct fields, for some use cases, the programmers can control the action they want to apply on different fields.
-
-\subsection{Struct}
-\subsubsection{G3. Adding/Removing a Struct field}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-// G3.1: Adding a new field to a struct
-struct InnerListeners {
-    pending: Mutex<Vec<Pending>>,
-    <@\textcolor{red}{--}@> queue_object_name: Uuid,
-}
-
-// G3.2: Removing a field from a struct
-struct SnapshotService<U, R> {
-    uuid_resolver_handle: R,
-    <@\textcolor{green}{++}@> db_name: String,
-}
-\end{lstlisting}
-
-\noindent \textbf{Description:} A set of new fields are add to or removed from an existing struct.
-The addition of a new field happens when the developer need to store a new piece of data so they can implement a desired behaviour.
-The removal of a field  happens when the developer realizes that the field is expendable with respect to the behaviour they want to implement.
-
-\subsubsection{G4. Modifying the access modifiers of struct fields}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-
-// G4.1: Adding pub to a field
-pub struct HtmlBlock {
-    -- content: BlockContent,
-    ++ pub content: BlockContent,
-    brace: token::Brace,
-}
-
-// G4.2: Removing pub from a field
-pub struct Languages {
---  pub named: HashMap<LanguageId, Arc<LanguageDefinition>>,
-++  named: HashMap<LanguageId, Arc<LanguageDefinition>>,
-    extensions: HashMap<String, Arc<LanguageDefinition>>,
-}
-
-\end{lstlisting}
-
-\noindent \textbf{Description:} Adding pub to a struct field makes it possible to access that field within external modules.
-In the instances of this cluster, adding pub access modifier happened in a bug fixing change, where the developer needed to access a field which had not been granted public access.
-On the other hand, the public access might be revoked if the developer realizes there is no external access to the field.
-
-\subsubsection{G5. Different fields order when initializing Struct}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
--- return Poll::Ready(ReadyEvent { ready, tick });
-++ return Poll::Ready(ReadyEvent { tick, ready });
-\end{lstlisting}
-
-\subsection{Option}
-\subsubsection{G6. Changing field type T to T$\langle$Option$\rangle$}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-
-#[ast_node("MediaRule")]
-pub struct MediaRule {
-    pub span: Span,
-    -- pub media: MediaQueryList,
-    ++ pub media: Option<MediaQueryList>,
-    pub rules: Vec<Rule>,
-}
-
-\end{lstlisting}
-
-\subsection{Types}
-\subsubsection{G7. Changing a struct field type}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-
-pub struct Manifest {
-  -- pub features: Vec<String>,
-  ++ pub features: HashSet<String>,
-}
-
-\end{lstlisting}
-
-\subsubsection{G8. Changing a generic type}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-
-#[derive(Debug, Clone, PartialEq)]
-pub struct Anchor {
-    -- point: Point<usize>,
-    ++ point: Point<isize>,
-    side: Side,
-}
-
-\end{lstlisting}
-
-\subsubsection{G9. Change in enum variant value type}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-
-#[derive(Debug, Clone, PartialEq, Eq)]
-pub enum ResolvedDependency {
-  Resolved(ModuleSpecifier),
-  -- Err(String),
-  ++ Err(ResolvedDependencyErr),
-}
-
-\end{lstlisting}
-
-\subsubsection{G10. Changing the type inside a tuple}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-
-#[allow(dead_code)]
-pub struct CoreState {
-    // Other fields
-    -- pending_views: Vec<ViewId>,
-    ++ pending_views: Vec<(ViewId, Table)>,
-    peer: Client
-}
-
-\end{lstlisting}
-
-\subsection{Traits}
-\subsubsection{G11. TraitBound Change}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-
--- pub struct EventLoop<T: tty::EventedReadWrite> {
-++ pub struct EventLoop<T: tty::EventedPty> {
-    poll: mio::Poll,
-    // Other fields
-}
-
-\end{lstlisting}
-
-\subsubsection{G12. Implementing a different trait for a struct}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-
--- impl ToDebug for ShellError {
-++ impl FormatDebug for ShellError {
-        // Implementing functions ...
-   }
-
-\end{lstlisting}
-
-\subsection{Ref and Deref}
-\subsubsection{B1. Adding $\&$}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-let repo = replace(&mut *contents, Processing).inner_repo();
--- let statuses = repo_to_statuses(repo, &self.workdir);
-++ let statuses = repo_to_statuses(&repo, &self.workdir);
-\end{lstlisting}
-
-\subsubsection{B2. Removing $\&$}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-enums.iter().for_each(|e| {
-    let mut enum_template = EnumTemplate::new();
-    -- enum_template.set_message_enum(&e);
-    ++ enum_template.set_message_enum(e);
-    // Other statements
-});
-\end{lstlisting}
-
-\subsubsection{B3. Dropping \& from func decl and adding * to match statement}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
--- fn to_color_choice(&self) -> ColorChoice {
---     match *self {
-++ fn to_color_choice(self) -> ColorChoice {
-++     match self {
-           // Statements
-       }
-\end{lstlisting}
-
-\subsubsection{B4. Adding *}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-for p in wixobjs {
---  args.push(p.to_string());
-++  args.push((*p).to_string());
-}
-
-let displayed_attributes = schema.clone().map(|s| {
-    s.displayed_name()
-        .iter()
---      .map(|s| (*s).to_string())
-++      .map(|s| s.to_string())
-        .collect::<HashSet<String>>()
-});
-\end{lstlisting}
-
-\subsection{Mut}
-\subsubsection{B5. Adding mut}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-
-parser::Parser::new(args)
-        .parse_module()
---      .map_err(|e| {
-++      .map_err(|mut e| {
-            e.emit();
-            ()
-        })?
-
-\end{lstlisting}
-
-\subsubsection{B6. Removing mut}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
--- let mut tx = tx.clone();
-++ let tx = tx.clone();
-\end{lstlisting}
-
-\subsection{Vector}
-\subsubsection{B7. Vector reference to slice}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-pub fn expand_delimited_square(
---  children: &Vec<TokenNode>,
-++  children: &[TokenNode],
-) -> Result<hir::Expression, ParseError> {
-    // body
-}
-\end{lstlisting}
-
-\subsection{Lifetime}
-\subsubsection{B8. Adding generic Lifetime}
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-#[allow(dead_code)]
--- fn api_error(webview: &mut Webview, error_fn: String, message: &str) {
-++ fn api_error(webview: &mut Webview<'_>, error_fn: String, message: &str) {
-  let reject_code = tauri_api::rpc::format_callback(error_fn, message);
-  webview.eval(&reject_code)
-}
-
-\end{lstlisting}
-
-\subsubsection{B9. Dropping static lifetime}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-
--- const QUEUE_SIZES: &'static [u16] = &[QUEUE_SIZE];
-++ const QUEUE_SIZES: &[u16] = &[QUEUE_SIZE];
-
-\end{lstlisting}
-
-\subsubsection{B10. Dropping Lifetime in field definition}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-pub struct Table<'a> {
-    columns: Vec<Column>,
-    colours: &'a Colours,
-    env: &'a Environment,
-    widths: TableWidths,
-    -- time_format: &'a TimeFormat,
-    ++ time_format: TimeFormat,
-    size_format: SizeFormat,
-    git: Option<&'a GitCache>,
-}
-\end{lstlisting}
-
-\subsection{Clone}
-\subsubsection{B11. Dropping clone and adding dereference}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-pub fn reset(&mut self, template: &Cell) {
-    // memcpy template to self
---  *self = template.clone();
-++  *self = *template;
-}
-\end{lstlisting}
-
-\subsubsection{B12. Dropping clone}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-if let Some(var_id) = &var.var_id {
---  stack.add_var(*var_id, input);
-++  stack.add_var(*var_id, input.clone());
-}
-
-start_plugin_process(
---  manifest.clone(),
-++  manifest,
-    self.next_plugin_id(),
-    self.self_ref.as_ref().unwrap().clone(),
-);
-
-\end{lstlisting}
-
-\subsubsection{B13. Dropping clone and adding reference}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
--- let field_id = schema.get_or_create(attribute.clone())?;
-++ let field_id = schema.get_or_create(&attribute)?;
-\end{lstlisting}
-
-\begin{table}[]
-
-\hskip-2.5cm\begin{tabular}{|l|l|l|l|}
-\hline
-
-\multirow{12}{*}{General}        
-& \multirow{2}{*}{Attributes}    & G1  & Adding Attributes to Structs                                        \\
-                                 &                                & G2  & Adding Attributes to Struct Fields                                  \\\cline{2-3}
-                                 & \multirow{3}{*}{Struct}        & G3  & Adding/Removing a Struct field                                      \\
-                                 &                                & G4  & Adding/Removing pub to/from struct field                            \\
-                                 &                                & G5  & Different fields order when initializinng Struct                       \\\cline{2-3}
-                                 & Option                         & G6  & Changing field type T to Option                                          \\\cline{2-3}
-                                 & \multirow{4}{*}{Types}         & G7  & Changing a Struct field type                                        \\
-                                 &                                & G8  & Changing a generic type                                             \\
-                                 &                                & G9  & Change in enum variant value type                                   \\
-                                 &                                & G10 & Changing the type inside a tuple                                         \\\cline{2-3}
-                                 & \multirow{2}{*}{Traits}        & G11 & TraitBound Change                                                   \\
-                                 &                                & G12 & Implementing a different trait for a struct.                            \\\cline{1-3}
-\multirow{13}{*}{Borrow Checker} & \multirow{4}{*}{Ref and Deref} & B1  & Adding \&                                                           \\
-                                 &                                & B2  & Removing \&                                                         \\
-                                 &                                & B3  & Dropping \& from function declaration and dereferencing match \\
-                                 &                                & B4  & Adding *                                                            \\\cline{2-3}
-                                 & \multirow{2}{*}{Mut}           & B5  & Adding mut                                                          \\
-                                 &                                & B6  & Removing mut                                                        \\\cline{2-3}
-                                 & Vector                         & B7  & Vector reference to slice                                           \\\cline{2-3}
-                                 & \multirow{3}{*}{Lifetime}      & B8  & Adding generic Lifetime                                             \\
-                                 &                                & B9  & Dropping static lifetime                                            \\
-                                 &                                & B10 & Dropping Lifetime in field definition                               \\\cline{2-3}
-                                 & \multirow{3}{*}{Clone}         & B11 & Dropping clone and adding dereference                               \\
-                                 &                                & B12 & Dropping clone                                                      \\
-                                 &                                & B13 & Dropping clone and adding reference
-
-\end{tabular}
-\caption{\label{fig:category}Bug Fix Pattern Categorization}
-\end{table}
-
-\begin{algorithm}
-\caption{An algorithm with caption}\label{alg:cap}
-\hspace*{2mm} \textbf{Input:} $R$ (target repositories)  \\
-\hspace*{2mm} \textbf{Output:} $D_g$ (General code changes) \\
-\hspace*{2mm} \textbf{Output:} $D_b$ (BC-related code changes)
-\begin{algorithmic}
-\State $D_g \leftarrow \phi$
-\State $D_b \leftarrow \phi$
-\For{$r \in R$}
-    \For{$c \in \textsc{ExtractCommits}(r)$}
-        \If{$c.msg$ contains bug fixing related keywords}
-            \For{$\{f_b, f_a\} \in \textsc{GetModifiedRustFiles}(c)$}
-                \For{$e \in \textsc{ASTDiff}(\textsc{Parse}(f_b), \textsc{Parse}(f_a))$}
-                    \State $DP \leftarrow \textsc{GetDataPoint}(e)$
-                    \State $DP \leftarrow c.hash \cup f_a.name \cup e.scope \cup DP $
-                    \If{$\textsc{IsBCRelated}(e)$}
-                        \State $DP \leftarrow \textsc{GetBCKeyword}(e) \cup DP $
-                        \State $D_b \leftarrow D_b \cup DP$
-                    \Else
-                        \State $D_g \leftarrow D_g \cup DP$
-                    \EndIf
-                \EndFor
-            \EndFor
-        \EndIf
-    \EndFor
-\EndFor
-\end{algorithmic}
-\end{algorithm}
-
-\end{document}

### 2022-09-12 3_results.tex
added some descriptions for results and reorganized files

--- 2022-09-08 main.tex
@@
-\documentclass{article}
-\usepackage{listings, listings-rust}
-\lstset{escapeinside={<@}{@>}}
-\usepackage{minted}
-\usepackage{tabularx}
-\usepackage{array}
-\usepackage{graphicx}
-\usepackage{algorithm}
-\usepackage{algpseudocode}
-\usepackage[utf8]{inputenc}
-\usepackage{multirow}
-
-\title{Rust Bug Fix Patterns}
-\author{Moh Robati}
-
-\begin{document}
-
-\maketitle
-
-\section{Introduction}
-
-Producing bugs has been an unavoidable part of computer programming which happens unintentionally.
-Researchers have been trying to help software engineers to be more aware of the bug producing practices in development and how to step towards creating bug-free software.
-One useful step towards that goal would be studying common bug patterns (and their corresponding fixes) that occur across multiple software systems.
-Research efforts for categorizing common bug patterns date back to 1975 where Edres tried to categorize various bug patterns which occurred in operating system program.
-
-In this work, we empirically analyze the most popular Rust projects in GitHub, and discover the common bug patterns lied within them.
-We believe these insights are important for various reasons, primarily for developing program repair tools.
-Automated program repair tools try to locate bug
-locations in the source code (they include a fault localization module) and then fix them, possibly without human intervention, in a manner such that the modified program meets the desired specifications.
-A powerful automated program repair system has many useful applications and can significantly simplify the process of debugging, resulting in a reduction in software development cost.
-The set of common bug fix patterns can be a useful input to an APR system as it directly provides information about fixing similar buggy structures, and that can significantly reduce the search space for patch generation.
-
-Similarly, if we are not observing the desired behaviour from our program, that would indicate that there lies a fault within our program.
-Fault localization is the process of automatically finding the faulty statement(s) in our program.
-If a fault localization tool knows about the common bug patterns that can appear in the underlying programming language, it can prioritize the code segment candidates based on their potentials of producing bugs.
-That would result in a more robust tool as it leverages from historical data.
-
-Moreover, knowing which code patterns in a programming language are more susceptible to produce bugs can help programmers to be more cautious while working with similar code structures.
-This knowledge can make the developers avoid certain mistakes during production which ultimately would result in reducing the frequent bug patterns within the program.
-As Monperrus describes, the less frequent bug patterns we keep in our software, the less cost we would have to pay for software maintenance.
-
-To discuss about a less talked about use case, recently, researchers have been trying to apply deep learning techniques on programs to learn about the semantics of the code or training models which can modify programs towards a specific goal (e.g.
-bug fix).
-One challenge is that usually learning tools get fixed size vectors as input and map that to an output.
-Unlike image data, it is challenging to compress programs into a vector without any loss in semantic information.
-Current approaches try to define the vector dimensions in a way that accounts for all possible abstract syntax tree forms.
-That results in a large sparse space for the inputs.
-Assuming that the deep learning model we are trying to develop is associated with bug fixing tasks, then knowledge about common bug fix patterns can help with pruning input space towards containing less dimensions.
-That can help with creating more efficient models.
-
-There have been many studies on detecting and categorizing bug fix patterns in general, or for a specific programming language. We are more interested in the latter, since
-our study is focused on pervasive bug fix patterns in Rust. \\
-Yang et al. proposed a mining approach to detect Python bug fix patterns via studying
-fine-grained fixing code changes. In their research, they also examined what portion of the bugs
-could be fixed using automated bug fixing approaches. Moreover, they evaluated the fix patterns in
-the wild and realized that 37\% of the buggy codes could be matched by the fix patterns they had
-found. \\
-Hanam et al. conducted a similar research but for finding Javascript most pervasive bug fix
-patterns. They presented a large-scale study of the most common bug fix patterns by mining 105K
-commits from 134 server-side JavaScript projects. They used DBSCAN clustering algorithm
-and divided bug fixing change types into 219 clusters, from which they extracted 13 pervasive
-cross-project bug fix patterns. \\
-In this work, we study common bug patterns that appear in Rust projects. Rust is a relatively new language that allows programmers to write programs that have low-level
-control over resources, as well as high-level safety guarantees. Originally, Rust was a side
-project developed by Graydon Hoare, a Mozilla employee. The company started to fund this project
-in 2009, since they intended to use Rust in some of the low-level Firefox components.
-What distinguishes Rust from the other systems programming languages, e.g. C/C++, is the
-ownership framework that has been incorporated inside the language. The ownership framework keeps the track of the references to different memory locations, and will free them once they
-go out of scope. If the program passes the compiler checks, it automatically guarantees several
-safety problems e.g. dangling references, or double-free problem, that had to be taken care of
-manually in languages like C/C++. Moreover, not only the ownership concept removes the
-need for a garbage collector in the runtime environment, it does not require the programmer to
-explicitly do certain safety tasks, like freeing references.
-Due to the memory and thread safety guarantees that Rust provides, many software companies
-adopted Rust in production. For example, Dropbox decided to migrate to Rust for developing their
-storage system, since they were not completely satisfied with the features provided by AWS’s S3 or
-Golang.
-
-The main module which governs over the ownership rules is called Borrow Checker.
-Because of its importance in Rust, we have decided to put discovered patterns into two groups; (1) non-borrow-checker related patterns (General Patterns), and (2) borrow-checker related patterns (BC-Related Patterns).
-This distinction was really helpful as it made our tool discover clearer results.
-
-We present our discovered insights of our empirical study in the following four research
-questions (RQs):
-
-\textbf{RQ1. How successful is our data modelling approach in capturing all the necessary data?}
-
-\textbf{RQ2. What are the most common AST structures in the bug fixing changes?}
-
-\textbf{RQ3. What are the general fix patterns in Rust?}
-
-\textbf{RQ4. Which fix patterns are related to Borrow Checker in Rust?}
-
-\section{Pervasive Bug Patterns}
 \subsection{Attributes}
@@
 ++ return Poll::Ready(ReadyEvent { tick, ready });
-\end{lstlisting}
+
+\end{lstlisting} \\
+
+\noindent \textbf{Description:} Remove... ???
 
@@
 
+\noindent \textbf{Description:} Rust is a null safe language, meaning that object references cannot assume null values.
+If we want void values for our object, we would have to use Option type.
+Option is essentially an enum with two variants, Some which indicates that our object has an value, and None shows the absence of any value.
+In this pattern, the developer changes the type of a struct field to an Option of its type.
+This makes it possible for the developer to modify the program in a way that it covers the scenarios in which the the struct field has no value.
+
+For instance, in the Rust project swc-project/swc (A fast TypeScript/JavaScript compiler), the developer needed to modify the CSS parser source code in a way that it accounts for empty @media queries \footnote{\href{https://github.com/swc-project/swc/commit/75a14f98b7370226115ee24eec6eb8c802bd4837}}, so, as shown in the code snippet above, they changed type MediaQueryList to Option<MediaQueryList>, and modified the other parts of the source code, accordingly.
+
 \subsection{Types}
@@
 \begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-
 pub struct Manifest {
@@
 \end{lstlisting}
+
+\noindent \textbf{Description:} The type of a struct field changes.
+This is a common change pattern in all statically-typed languages and can happen for various reasons.
+These reasons can be a step for implementing a new feature or rectifying the behaviour of the program.
+Also it can happen for refactoring purposes or performance enhancement.
 
@@
 
+\noindent \textbf{Description:} This a change of the generic (parametric) type of a struct field. Similar to G7, this change can also happen for the mentioned reasons.
+
 \subsubsection{G9. Change in enum variant value type}
@@
 \end{lstlisting}
+
+\noindent \textbf{Description:} In Rust, we can specify the type of the value that an enum variant can store.
+Using this feature, we can avoid using a struct alongside our enum variants to specify the value type.
+This change pattern is associated with a change in enum variant types.
 
@@
 
+\noindent \textbf{Description:} In this pattern, an extra type is added to an existing type, making it a tuple of multiple types.
+Alternatively, the change might be dropping types from the tuple and making it a tuple with fewer number of arguments.
+
 \subsection{Traits}
@@
 
+\noindent \textbf{Description:} The trait bounds are the functionalities that we require from our parametric types.
+This change pattern related to modification in trait bounds.
+This can happen both for bug fix reasons or just refactoring purposes.
+
 \subsubsection{G12. Implementing a different trait for a struct}
@@
 
+\noindent \textbf{Description:} Remove... ???
+
 \subsection{Ref and Deref}
@@
 \end{lstlisting}
+
+\noindent \textbf{Description:} This change happens when the developer needs to modify the program so it reaches the desired behaviour.
+The modification, however, requires access to an object whose ownership has moved to another scope and never returned.
+That is why we must use borrowing to preserve ownership within our scope, after the function execution finishes.
+
+To give an example, in starship/starship after one of the commits \footnote{\href{https://github.com/starship/starship/commit/56d475578ea508631275772127f49a6949fea6b0}{git_branches.rs}}, the developer borrows the value of $repo\_dir$ to keep the ownership of the object in the current scope.
+Now this object can then be used in another function invocation ($remove\_dir\_all(repo\_dir)$) which fixes a bug.
 
@@
 
+\noindent \textbf{Description:} In the pattern, an unnecessary borrowing is removed.
+Clippy can detect the needless borrows, and also can automatically remove them.
+This does not change the behaviour of the program, however, it does improve the performance, as the borrow checker need not to return the ownership after the execution of the function.
+
 \subsubsection{B3. Dropping & from func decl and adding * to match statement}
@@
 
-\subsubsection{B4. Adding *}
+\noindent \textbf{Description:} ???
+
+\subsubsection{B4. Modification in dereferencing a string prior to calling .to_string()}
 
@@
 
+\noindent \textbf{Description:} This is a change that in all the instances has been detected by Clippy.
+The change applies on an object prior to calling $to\_string()$ on that object.
+The change might add or remove a dereference, as shown above.
+In the addition case, the reason behind the change is that the object a reference type of $T$, while $T$ directly implements $to\_string()$.
+It is faster and more performance efficient to dereference the object and then call $to\_string()$ on it.
+The deletion case indicates a redundant dereference, thus it is removed.
+
 \subsection{Mut}
@@
 
+\noindent \textbf{Description:} This change happens when the programmer needs to mutate a variable which is immutable, hence they need to add the mut keyword.
+In the observed instances, this pattern was not seen in refactoring changes, the modification happened alongside a change in program behaviour.
+
 \subsubsection{B6. Removing mut}
@@
 \end{lstlisting}
+
+\noindent \textbf{Description:} Dropping the mut keyword before a variable, as we do not mutate the variable. This redundant mutability is reported by Clippy and can also be removed through this tool.
 
@@
 
+\noindent \textbf{Description:} ???
+
 \subsection{Lifetime}
@@
 
+\noindent \textbf{Description:} ???
+
 \subsubsection{B9. Dropping static lifetime}
@@
 \end{lstlisting}
+
+\noindent \textbf{Description:} ???
 
@@
 
+\noindent \textbf{Description:} ???
+
 \subsection{Clone}
-\subsubsection{B11. Dropping clone and adding dereference}
+\subsubsection{B11. Dropping clone (and adding dereference)}
 
@@
 \end{lstlisting}
+
+\noindent \textbf{Description:} This pattern is referred to as redundant clone in Clippy Lints.
+Clippy detects whether a variable clone is necessary and could we have used the variable itself or not.
+This change does not modify the program behaviour, however it can boost the performance significantly if we repeatedly cloning large variables.
 
@@
 
+\noindent \textbf{Description:} This pattern is referred to as redundant clone in Clippy Lints.
+Clippy detects whether a variable clone is necessary and could we have used the variable itself or not.
+This change does not modify the program behaviour, however it can boost the performance significantly if we repeatedly cloning large variables.
+
 \subsubsection{B13. Dropping clone and adding reference}
@@
 \end{lstlisting}
+
+\noindent \textbf{Description:} Unlike the previous pattern, this pattern is not detected by Clippy.
+The change happens when the programmers realizes that the cloning of a variable is unnecessary, however, we also want to preserve the ownership of the variable within the current scope.
+That is why in this pattern a clone is turned into a borrowing.
+
+Similar to the previous pattern, this does not change the program behaviour and is merely for performance purposes.
+It was the case in one of the commits of the project tauri-apps/tauri, where the clone to borrowing significantly optimized CPU usage.
+They discussed that the cloning of some types such as HashMap was expensive.
 
@@
 \end{table}
-
-\begin{algorithm}
-\caption{An algorithm with caption}\label{alg:cap}
-\hspace*{2mm} \textbf{Input:} $R$ (target repositories)  \\
-\hspace*{2mm} \textbf{Output:} $D_g$ (General code changes) \\
-\hspace*{2mm} \textbf{Output:} $D_b$ (BC-related code changes)
-\begin{algorithmic}
-\State $D_g \leftarrow \phi$
-\State $D_b \leftarrow \phi$
-\For{$r \in R$}
-    \For{$c \in \textsc{ExtractCommits}(r)$}
-        \If{$c.msg$ contains bug fixing related keywords}
-            \For{$\{f_b, f_a\} \in \textsc{GetModifiedRustFiles}(c)$}
-                \For{$e \in \textsc{ASTDiff}(\textsc{Parse}(f_b), \textsc{Parse}(f_a))$}
-                    \State $DP \leftarrow \textsc{GetDataPoint}(e)$
-                    \State $DP \leftarrow c.hash \cup f_a.name \cup e.scope \cup DP $
-                    \If{$\textsc{IsBCRelated}(e)$}
-                        \State $DP \leftarrow \textsc{GetBCKeyword}(e) \cup DP $
-                        \State $D_b \leftarrow D_b \cup DP$
-                    \Else
-                        \State $D_g \leftarrow D_g \cup DP$
-                    \EndIf
-                \EndFor
-            \EndFor
-        \EndIf
-    \EndFor
-\EndFor
-\end{algorithmic}
-\end{algorithm}
-
-\end{document}

### 2022-09-12 main.tex
added some descriptions for results and reorganized files

@@
 \usepackage{multirow}
+\usepackage{subfiles} % Best loaded last in the preamble
 
@@
 \begin{document}
-
 \maketitle
@@
 \section{Introduction}
+\subfile{1_introduction}
 
-Producing bugs has been an unavoidable part of computer programming which happens unintentionally.
-Researchers have been trying to help software engineers to be more aware of the bug producing practices in development and how to step towards creating bug-free software.
-One useful step towards that goal would be studying common bug patterns (and their corresponding fixes) that occur across multiple software systems.
-Research efforts for categorizing common bug patterns date back to 1975 where Edres tried to categorize various bug patterns which occurred in operating system program.
+\section{Methodology}
+\subfile{2_methodology}
 
-In this work, we empirically analyze the most popular Rust projects in GitHub, and discover the common bug patterns lied within them.
-We believe these insights are important for various reasons, primarily for developing program repair tools.
-Automated program repair tools try to locate bug
-locations in the source code (they include a fault localization module) and then fix them, possibly without human intervention, in a manner such that the modified program meets the desired specifications.
-A powerful automated program repair system has many useful applications and can significantly simplify the process of debugging, resulting in a reduction in software development cost.
-The set of common bug fix patterns can be a useful input to an APR system as it directly provides information about fixing similar buggy structures, and that can significantly reduce the search space for patch generation.
+\section{Results}
+\subfile{3_results}
 
-Similarly, if we are not observing the desired behaviour from our program, that would indicate that there lies a fault within our program.
-Fault localization is the process of automatically finding the faulty statement(s) in our program.
-If a fault localization tool knows about the common bug patterns that can appear in the underlying programming language, it can prioritize the code segment candidates based on their potentials of producing bugs.
-That would result in a more robust tool as it leverages from historical data.
+\section{Discussion}
+\subfile{4_discussion}
 
-Moreover, knowing which code patterns in a programming language are more susceptible to produce bugs can help programmers to be more cautious while working with similar code structures.
-This knowledge can make the developers avoid certain mistakes during production which ultimately would result in reducing the frequent bug patterns within the program.
-As Monperrus describes, the less frequent bug patterns we keep in our software, the less cost we would have to pay for software maintenance.
-
-To discuss about a less talked about use case, recently, researchers have been trying to apply deep learning techniques on programs to learn about the semantics of the code or training models which can modify programs towards a specific goal (e.g.
-bug fix).
-One challenge is that usually learning tools get fixed size vectors as input and map that to an output.
-Unlike image data, it is challenging to compress programs into a vector without any loss in semantic information.
-Current approaches try to define the vector dimensions in a way that accounts for all possible abstract syntax tree forms.
-That results in a large sparse space for the inputs.
-Assuming that the deep learning model we are trying to develop is associated with bug fixing tasks, then knowledge about common bug fix patterns can help with pruning input space towards containing less dimensions.
-That can help with creating more efficient models.
-
-There have been many studies on detecting and categorizing bug fix patterns in general, or for a specific programming language. We are more interested in the latter, since
-our study is focused on pervasive bug fix patterns in Rust. \\
-Yang et al. proposed a mining approach to detect Python bug fix patterns via studying
-fine-grained fixing code changes. In their research, they also examined what portion of the bugs
-could be fixed using automated bug fixing approaches. Moreover, they evaluated the fix patterns in
-the wild and realized that 37\% of the buggy codes could be matched by the fix patterns they had
-found. \\
-Hanam et al. conducted a similar research but for finding Javascript most pervasive bug fix
-patterns. They presented a large-scale study of the most common bug fix patterns by mining 105K
-commits from 134 server-side JavaScript projects. They used DBSCAN clustering algorithm
-and divided bug fixing change types into 219 clusters, from which they extracted 13 pervasive
-cross-project bug fix patterns. \\
-In this work, we study common bug patterns that appear in Rust projects. Rust is a relatively new language that allows programmers to write programs that have low-level
-control over resources, as well as high-level safety guarantees. Originally, Rust was a side
-project developed by Graydon Hoare, a Mozilla employee. The company started to fund this project
-in 2009, since they intended to use Rust in some of the low-level Firefox components.
-What distinguishes Rust from the other systems programming languages, e.g. C/C++, is the
-ownership framework that has been incorporated inside the language. The ownership framework keeps the track of the references to different memory locations, and will free them once they
-go out of scope. If the program passes the compiler checks, it automatically guarantees several
-safety problems e.g. dangling references, or double-free problem, that had to be taken care of
-manually in languages like C/C++. Moreover, not only the ownership concept removes the
-need for a garbage collector in the runtime environment, it does not require the programmer to
-explicitly do certain safety tasks, like freeing references.
-Due to the memory and thread safety guarantees that Rust provides, many software companies
-adopted Rust in production. For example, Dropbox decided to migrate to Rust for developing their
-storage system, since they were not completely satisfied with the features provided by AWS’s S3 or
-Golang.
-
-The main module which governs over the ownership rules is called Borrow Checker.
-Because of its importance in Rust, we have decided to put discovered patterns into two groups; (1) non-borrow-checker related patterns (General Patterns), and (2) borrow-checker related patterns (BC-Related Patterns).
-This distinction was really helpful as it made our tool discover clearer results.
-
-We present our discovered insights of our empirical study in the following four research
-questions (RQs):
-
-\textbf{RQ1. How successful is our data modelling approach in capturing all the necessary data?}
-
-\textbf{RQ2. What are the most common AST structures in the bug fixing changes?}
-
-\textbf{RQ3. What are the general fix patterns in Rust?}
-
-\textbf{RQ4. Which fix patterns are related to Borrow Checker in Rust?}
-
-\section{Pervasive Bug Patterns}
-\subsection{Attributes}
-
-Attributes in Rust make it possible for the developers to add metadata to program elements.
-These attributes can be added to structs, struct fields, enum variants, match expression arms, etc.
-There are two types of attributes: Inner Attributes and Outer Attributes.
-Inner Attributes apply to the program element encapsulating them, while Outer Attributes apply to the program element coming after the attribute.
-
-\subsubsection{G1. Modifying the attributes of structs}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-// G1.1: Adding attributes to a struct
-++ #[derive(Debug, Clone)]
-pub struct Call {
-/// body 
-}
-
-// G1.2: Removing attributes from a struct
--- #[derive(Debug, Clone)]
-pub struct Call {
-/// body 
-}
-
-// G1.3: Changing an existing attribute of a struct
--- #[derive(Debug)]
-++ #[derive(Debug, Clone)]
-pub struct Call {
-/// body 
-}
-\end{lstlisting}
-
-\noindent \textbf{Description:} In this pattern, the attribute set of a struct changes.
-This modification can happen in three different ways; (G1.1) Adding attributes to the struct, (G1.2) Removing attributes from the struct, (G1.3) Changing the content of an existing attribute.
-The attributes are Outer Attributes as we can find struct related non-terminals within the ASTDiff of the instances of this pattern.
-
-\subsubsection{G2. Modifying the attributes of struct fields}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-// G2.1: Adding attributes to struct fields
-#[derive(Serialize)]
-pub struct Retain {
-    ++ #[serde(skip_serializing_if = "is_empty")]
-    pub attributes: Option<Attributes>,
-}
-
-// G2.2: Removing attributes from struct fields
-#[derive(Serialize)]
-pub struct Retain {
-    -- #[serde(skip_serializing_if = "Option::is_none")]
-    pub attributes: Option<Attributes>,
-}
-
-// G2.3: Changing attributes of struct fields
-#[derive(Serialize)]
-pub struct Retain {
-    -- #[serde(skip_serializing_if = "Option::is_none")]
-    ++ #[serde(skip_serializing_if = "is_empty")]
-    pub attributes: Option<Attributes>,
-}
-\end{lstlisting}
-
-\noindent \textbf{Description:} In this pattern, the attribute set of a struct field changes.
-Similar to G1, this modification also happens in three different ways.
-Using attributes for struct fields, apart from being able to store different meta information for different struct fields, for some use cases, the programmers can control the action they want to apply on different fields.
-
-\subsection{Struct}
-\subsubsection{G3. Adding/Removing a Struct field}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-// G3.1: Adding a new field to a struct
-struct InnerListeners {
-    pending: Mutex<Vec<Pending>>,
-    <@\textcolor{red}{--}@> queue_object_name: Uuid,
-}
-
-// G3.2: Removing a field from a struct
-struct SnapshotService<U, R> {
-    uuid_resolver_handle: R,
-    <@\textcolor{green}{++}@> db_name: String,
-}
-\end{lstlisting}
-
-\noindent \textbf{Description:} A set of new fields are add to or removed from an existing struct.
-The addition of a new field happens when the developer need to store a new piece of data so they can implement a desired behaviour.
-The removal of a field  happens when the developer realizes that the field is expendable with respect to the behaviour they want to implement.
-
-\subsubsection{G4. Modifying the access modifiers of struct fields}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-
-// G4.1: Adding pub to a field
-pub struct HtmlBlock {
-    -- content: BlockContent,
-    ++ pub content: BlockContent,
-    brace: token::Brace,
-}
-
-// G4.2: Removing pub from a field
-pub struct Languages {
---  pub named: HashMap<LanguageId, Arc<LanguageDefinition>>,
-++  named: HashMap<LanguageId, Arc<LanguageDefinition>>,
-    extensions: HashMap<String, Arc<LanguageDefinition>>,
-}
-
-\end{lstlisting}
-
-\noindent \textbf{Description:} Adding pub to a struct field makes it possible to access that field within external modules.
-In the instances of this cluster, adding pub access modifier happened in a bug fixing change, where the developer needed to access a field which had not been granted public access.
-On the other hand, the public access might be revoked if the developer realizes there is no external access to the field.
-
-\subsubsection{G5. Different fields order when initializing Struct}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
--- return Poll::Ready(ReadyEvent { ready, tick });
-++ return Poll::Ready(ReadyEvent { tick, ready });
-\end{lstlisting}
-
-\subsection{Option}
-\subsubsection{G6. Changing field type T to T$\langle$Option$\rangle$}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-
-#[ast_node("MediaRule")]
-pub struct MediaRule {
-    pub span: Span,
-    -- pub media: MediaQueryList,
-    ++ pub media: Option<MediaQueryList>,
-    pub rules: Vec<Rule>,
-}
-
-\end{lstlisting}
-
-\subsection{Types}
-\subsubsection{G7. Changing a struct field type}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-
-pub struct Manifest {
-  -- pub features: Vec<String>,
-  ++ pub features: HashSet<String>,
-}
-
-\end{lstlisting}
-
-\subsubsection{G8. Changing a generic type}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-
-#[derive(Debug, Clone, PartialEq)]
-pub struct Anchor {
-    -- point: Point<usize>,
-    ++ point: Point<isize>,
-    side: Side,
-}
-
-\end{lstlisting}
-
-\subsubsection{G9. Change in enum variant value type}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-
-#[derive(Debug, Clone, PartialEq, Eq)]
-pub enum ResolvedDependency {
-  Resolved(ModuleSpecifier),
-  -- Err(String),
-  ++ Err(ResolvedDependencyErr),
-}
-
-\end{lstlisting}
-
-\subsubsection{G10. Changing the type inside a tuple}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-
-#[allow(dead_code)]
-pub struct CoreState {
-    // Other fields
-    -- pending_views: Vec<ViewId>,
-    ++ pending_views: Vec<(ViewId, Table)>,
-    peer: Client
-}
-
-\end{lstlisting}
-
-\subsection{Traits}
-\subsubsection{G11. TraitBound Change}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-
--- pub struct EventLoop<T: tty::EventedReadWrite> {
-++ pub struct EventLoop<T: tty::EventedPty> {
-    poll: mio::Poll,
-    // Other fields
-}
-
-\end{lstlisting}
-
-\subsubsection{G12. Implementing a different trait for a struct}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-
--- impl ToDebug for ShellError {
-++ impl FormatDebug for ShellError {
-        // Implementing functions ...
-   }
-
-\end{lstlisting}
-
-\subsection{Ref and Deref}
-\subsubsection{B1. Adding $\&$}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-let repo = replace(&mut *contents, Processing).inner_repo();
--- let statuses = repo_to_statuses(repo, &self.workdir);
-++ let statuses = repo_to_statuses(&repo, &self.workdir);
-\end{lstlisting}
-
-\subsubsection{B2. Removing $\&$}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-enums.iter().for_each(|e| {
-    let mut enum_template = EnumTemplate::new();
-    -- enum_template.set_message_enum(&e);
-    ++ enum_template.set_message_enum(e);
-    // Other statements
-});
-\end{lstlisting}
-
-\subsubsection{B3. Dropping \& from func decl and adding * to match statement}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
--- fn to_color_choice(&self) -> ColorChoice {
---     match *self {
-++ fn to_color_choice(self) -> ColorChoice {
-++     match self {
-           // Statements
-       }
-\end{lstlisting}
-
-\subsubsection{B4. Adding *}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-for p in wixobjs {
---  args.push(p.to_string());
-++  args.push((*p).to_string());
-}
-
-let displayed_attributes = schema.clone().map(|s| {
-    s.displayed_name()
-        .iter()
---      .map(|s| (*s).to_string())
-++      .map(|s| s.to_string())
-        .collect::<HashSet<String>>()
-});
-\end{lstlisting}
-
-\subsection{Mut}
-\subsubsection{B5. Adding mut}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-
-parser::Parser::new(args)
-        .parse_module()
---      .map_err(|e| {
-++      .map_err(|mut e| {
-            e.emit();
-            ()
-        })?
-
-\end{lstlisting}
-
-\subsubsection{B6. Removing mut}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
--- let mut tx = tx.clone();
-++ let tx = tx.clone();
-\end{lstlisting}
-
-\subsection{Vector}
-\subsubsection{B7. Vector reference to slice}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-pub fn expand_delimited_square(
---  children: &Vec<TokenNode>,
-++  children: &[TokenNode],
-) -> Result<hir::Expression, ParseError> {
-    // body
-}
-\end{lstlisting}
-
-\subsection{Lifetime}
-\subsubsection{B8. Adding generic Lifetime}
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-#[allow(dead_code)]
--- fn api_error(webview: &mut Webview, error_fn: String, message: &str) {
-++ fn api_error(webview: &mut Webview<'_>, error_fn: String, message: &str) {
-  let reject_code = tauri_api::rpc::format_callback(error_fn, message);
-  webview.eval(&reject_code)
-}
-
-\end{lstlisting}
-
-\subsubsection{B9. Dropping static lifetime}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-
--- const QUEUE_SIZES: &'static [u16] = &[QUEUE_SIZE];
-++ const QUEUE_SIZES: &[u16] = &[QUEUE_SIZE];
-
-\end{lstlisting}
-
-\subsubsection{B10. Dropping Lifetime in field definition}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-pub struct Table<'a> {
-    columns: Vec<Column>,
-    colours: &'a Colours,
-    env: &'a Environment,
-    widths: TableWidths,
-    -- time_format: &'a TimeFormat,
-    ++ time_format: TimeFormat,
-    size_format: SizeFormat,
-    git: Option<&'a GitCache>,
-}
-\end{lstlisting}
-
-\subsection{Clone}
-\subsubsection{B11. Dropping clone and adding dereference}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-pub fn reset(&mut self, template: &Cell) {
-    // memcpy template to self
---  *self = template.clone();
-++  *self = *template;
-}
-\end{lstlisting}
-
-\subsubsection{B12. Dropping clone}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-if let Some(var_id) = &var.var_id {
---  stack.add_var(*var_id, input);
-++  stack.add_var(*var_id, input.clone());
-}
-
-start_plugin_process(
---  manifest.clone(),
-++  manifest,
-    self.next_plugin_id(),
-    self.self_ref.as_ref().unwrap().clone(),
-);
-
-\end{lstlisting}
-
-\subsubsection{B13. Dropping clone and adding reference}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
--- let field_id = schema.get_or_create(attribute.clone())?;
-++ let field_id = schema.get_or_create(&attribute)?;
-\end{lstlisting}
-
-\begin{table}[]
-
-\hskip-2.5cm\begin{tabular}{|l|l|l|l|}
-\hline
-
-\multirow{12}{*}{General}        
-& \multirow{2}{*}{Attributes}    & G1  & Adding Attributes to Structs                                        \\
-                                 &                                & G2  & Adding Attributes to Struct Fields                                  \\\cline{2-3}
-                                 & \multirow{3}{*}{Struct}        & G3  & Adding/Removing a Struct field                                      \\
-                                 &                                & G4  & Adding/Removing pub to/from struct field                            \\
-                                 &                                & G5  & Different fields order when initializinng Struct                       \\\cline{2-3}
-                                 & Option                         & G6  & Changing field type T to Option                                          \\\cline{2-3}
-                                 & \multirow{4}{*}{Types}         & G7  & Changing a Struct field type                                        \\
-                                 &                                & G8  & Changing a generic type                                             \\
-                                 &                                & G9  & Change in enum variant value type                                   \\
-                                 &                                & G10 & Changing the type inside a tuple                                         \\\cline{2-3}
-                                 & \multirow{2}{*}{Traits}        & G11 & TraitBound Change                                                   \\
-                                 &                                & G12 & Implementing a different trait for a struct.                            \\\cline{1-3}
-\multirow{13}{*}{Borrow Checker} & \multirow{4}{*}{Ref and Deref} & B1  & Adding \&                                                           \\
-                                 &                                & B2  & Removing \&                                                         \\
-                                 &                                & B3  & Dropping \& from function declaration and dereferencing match \\
-                                 &                                & B4  & Adding *                                                            \\\cline{2-3}
-                                 & \multirow{2}{*}{Mut}           & B5  & Adding mut                                                          \\
-                                 &                                & B6  & Removing mut                                                        \\\cline{2-3}
-                                 & Vector                         & B7  & Vector reference to slice                                           \\\cline{2-3}
-                                 & \multirow{3}{*}{Lifetime}      & B8  & Adding generic Lifetime                                             \\
-                                 &                                & B9  & Dropping static lifetime                                            \\
-                                 &                                & B10 & Dropping Lifetime in field definition                               \\\cline{2-3}
-                                 & \multirow{3}{*}{Clone}         & B11 & Dropping clone and adding dereference                               \\
-                                 &                                & B12 & Dropping clone                                                      \\
-                                 &                                & B13 & Dropping clone and adding reference
-
-\end{tabular}
-\caption{\label{fig:category}Bug Fix Pattern Categorization}
-\end{table}
-
-\begin{algorithm}
-\caption{An algorithm with caption}\label{alg:cap}
-\hspace*{2mm} \textbf{Input:} $R$ (target repositories)  \\
-\hspace*{2mm} \textbf{Output:} $D_g$ (General code changes) \\
-\hspace*{2mm} \textbf{Output:} $D_b$ (BC-related code changes)
-\begin{algorithmic}
-\State $D_g \leftarrow \phi$
-\State $D_b \leftarrow \phi$
-\For{$r \in R$}
-    \For{$c \in \textsc{ExtractCommits}(r)$}
-        \If{$c.msg$ contains bug fixing related keywords}
-            \For{$\{f_b, f_a\} \in \textsc{GetModifiedRustFiles}(c)$}
-                \For{$e \in \textsc{ASTDiff}(\textsc{Parse}(f_b), \textsc{Parse}(f_a))$}
-                    \State $DP \leftarrow \textsc{GetDataPoint}(e)$
-                    \State $DP \leftarrow c.hash \cup f_a.name \cup e.scope \cup DP $
-                    \If{$\textsc{IsBCRelated}(e)$}
-                        \State $DP \leftarrow \textsc{GetBCKeyword}(e) \cup DP $
-                        \State $D_b \leftarrow D_b \cup DP$
-                    \Else
-                        \State $D_g \leftarrow D_g \cup DP$
-                    \EndIf
-                \EndFor
-            \EndFor
-        \EndIf
-    \EndFor
-\EndFor
-\end{algorithmic}
-\end{algorithm}
+\section{Conclusion}
+\subfile{5_conclusion}
 

### 2022-09-12 3_results.tex
fix compilation errors

@@
 
-\end{lstlisting} \\
+\end{lstlisting}
 
@@
 
-For instance, in the Rust project swc-project/swc (A fast TypeScript/JavaScript compiler), the developer needed to modify the CSS parser source code in a way that it accounts for empty @media queries \footnote{\href{https://github.com/swc-project/swc/commit/75a14f98b7370226115ee24eec6eb8c802bd4837}}, so, as shown in the code snippet above, they changed type MediaQueryList to Option<MediaQueryList>, and modified the other parts of the source code, accordingly.
+For instance, in the Rust project swc-project/swc (A fast TypeScript/JavaScript compiler), the developer needed to modify the CSS parser source code in a way that it accounts for empty @media queries \footnote{\url{https://github.com/swc-project/swc/commit/75a14f98b7370226115ee24eec6eb8c802bd4837}}, so, as shown in the code snippet above, they changed type MediaQueryList to Option<MediaQueryList>, and modified the other parts of the source code, accordingly.
 
@@
 
-To give an example, in starship/starship after one of the commits \footnote{\href{https://github.com/starship/starship/commit/56d475578ea508631275772127f49a6949fea6b0}{git_branches.rs}}, the developer borrows the value of $repo\_dir$ to keep the ownership of the object in the current scope.
+To give an example, in starship/starship after one of the commits \footnote{\url{https://github.com/starship/starship/commit/56d475578ea508631275772127f49a6949fea6b0}, file \texttt{git\_branches.rs}}, the developer borrows the value of $repo\_dir$ to keep the ownership of the object in the current scope.
 Now this object can then be used in another function invocation ($remove\_dir\_all(repo\_dir)$) which fixes a bug.
@@
 
-\subsubsection{B4. Modification in dereferencing a string prior to calling .to_string()}
+\subsubsection{B4. Modification in dereferencing a string prior to calling .to\_string()}
 
@@
                                  &                                & B12 & Dropping clone                                                      \\
-                                 &                                & B13 & Dropping clone and adding reference  & \hline
+                                 &                                & B13 & Dropping clone and adding reference  \&
 

### 2022-09-12 1_introduction.tex
editing

@@
+% can we drop the first paragraph altogether?
 Producing bugs has been an unavoidable part of computer programming which happens unintentionally.
@@
 
-In this work, we empirically analyze the most popular Rust projects in GitHub, and discover the common bug patterns lied within them.
-We believe these insights are important for various reasons, primarily for developing program repair tools.
-Automated program repair tools try to locate bug
-locations in the source code (they include a fault localization module) and then fix them, possibly without human intervention, in a manner such that the modified program meets the desired specifications.
-A powerful automated program repair system has many useful applications and can significantly simplify the process of debugging, resulting in a reduction in software development cost.
-The set of common bug fix patterns can be a useful input to an APR system as it directly provides information about fixing similar buggy structures, and that can significantly reduce the search space for patch generation.
+In this work, we empirically analyze the most popular Rust projects in GitHub, aiming to discover common bug patterns that these projects share.
+Our interest comes from a desire to develop program repair tools, but we believe these insights can also be of broader service.
+Automated program repair tools try to locate bug locations in the source code (using a fault localization module) and then generate patches to fix them, possibly without human intervention, so that the modified program meets desired specifications (implicit or explicit).
+A powerful automated program repair system has many useful applications and could significantly simplify debugging, thus reducing software development cost.
+A set of common bug fix patterns can be a useful input to an APR system: it directly provides information about fixing similar buggy structures, which can significantly reduce the search space for patch generation.
 
+% how does this application below differ from the one above?
 Similarly, if we are not observing the desired behaviour from our program, that would indicate that there lies a fault within our program.
@@
 
-Moreover, knowing which code patterns in a programming language are more susceptible to produce bugs can help programmers to be more cautious while working with similar code structures.
+Moreover, knowing which code patterns in a programming language are more susceptible to bugs can help programmers to be more cautious while working with similar code structures.
 This knowledge can make the developers avoid certain mistakes during production which ultimately would result in reducing the frequent bug patterns within the program.
-As Monperrus describes, the less frequent bug patterns we keep in our software, the less cost we would have to pay for software maintenance.
+As Monperrus describes, the fewer frequent bug patterns we keep in our software, the less cost we would have to pay for software maintenance.
+% maybe more useful for PL design than for programmers; on the other hand I was just talking about stunts in movies and how people are probably more in danger from the less dangerous ones that they do 1000 times rather than the more dangerous ones that they do 2 times; it would be neat if there was a citation on that, but it's a bit of a distraction and doesn't actually help us get our paper in!
 
-To discuss about a less talked about use case, recently, researchers have been trying to apply deep learning techniques on programs to learn about the semantics of the code or training models which can modify programs towards a specific goal (e.g.
+% sentence too long; split it.
+Moving further afield, researchers have recently been trying to apply deep learning techniques on programs to learn about the semantics of the code or training models which can modify programs towards a specific goal (e.g.
 bug fix).
-One challenge is that usually learning tools get fixed size vectors as input and map that to an output.
-Unlike image data, it is challenging to compress programs into a vector without any loss in semantic information.
+However, most learning tools get fixed-size vectors as input and map them to outputs.
+Unlike image data, it is challenging to compress programs into vectors without any semantic information loss.
 Current approaches try to define the vector dimensions in a way that accounts for all possible abstract syntax tree forms.
-That results in a large sparse space for the inputs.
-Assuming that the deep learning model we are trying to develop is associated with bug fixing tasks, then knowledge about common bug fix patterns can help with pruning input space towards containing less dimensions.
+That results in a large, sparse space for the inputs.
+Assuming that the deep learning model we are trying to develop is associated with bug fixing tasks, then knowledge about common bug fix patterns can help prune the input space towards containing fewer dimensions.
 That can help with creating more efficient models.
 
-There have been many studies on detecting and categorizing bug fix patterns in general, or for a specific programming language. We are more interested in the latter, since
-our study is focused on pervasive bug fix patterns in Rust. \\
+% put into a new related work section
+% this is why numbers in front of filenames are kind of annoying, eh
+Many studies have investigated detecting and categorizing bug fix patterns in general, or for a specific programming language.
+Our study is focused on pervasive bug fix patterns in the specific language Rust.
+
+% I'll reread this part once it's in its own section.
+\paragraph{General bug fix patterns}
 Yang et al. proposed a mining approach to detect Python bug fix patterns via studying
@@
 the wild and realized that 37\% of the buggy codes could be matched by the fix patterns they had
-found. \\
+found.
 Hanam et al. conducted a similar research but for finding Javascript most pervasive bug fix
@@
 and divided bug fixing change types into 219 clusters, from which they extracted 13 pervasive
-cross-project bug fix patterns. \\
+cross-project bug fix patterns.
+
+\paragraph{Rust bug fix patterns}
 In this work, we study common bug patterns that appear in Rust projects. Rust is a relatively new language that allows programmers to write programs that have low-level
@@
 
-The main module which governs over the ownership rules is called Borrow Checker.
-Because of its importance in Rust, we have decided to put discovered patterns into two groups; (1) non-borrow-checker related patterns (General Patterns), and (2) borrow-checker related patterns (BC-Related Patterns).
-This distinction was really helpful as it made our tool discover clearer results.
+% ok, now back to the introduction section
+% did we talk about ownership yet?
+In Rust, the so-called Borrow Checker component enforces ownership rules.
+Because of the Borrow Checker's importance in Rust, we split patterns into two groups: (1) non-borrow-checker related patterns (General Patterns), and (2) borrow-checker related patterns (BC-Related Patterns).
+%This distinction was really helpful as it made our tool discover clearer results.
+- can you write something better?
 
@@
 
+% we talked about these, right?
 \textbf{RQ1. How successful is our data modelling approach in capturing all the necessary data?}

### 2022-09-12 3_results.tex
editing

@@
+% some general comments about the taxonomy go here
+
+% take one example and explain it, e.g. ++ means ..., -- means ....
+% rename to e.g. B.ref.add
+% need quantitative data
+
 \subsection{Attributes}
@@
 Inner Attributes apply to the program element encapsulating them, while Outer Attributes apply to the program element coming after the attribute.
+% does the inner/outer distinction matter?
 
@@
 
+% The sentence "Using..." doesn't actually have an appropriate verb
 \noindent \textbf{Description:} In this pattern, the attribute set of a struct field changes.
-Similar to G1, this modification also happens in three different ways.
-Using attributes for struct fields, apart from being able to store different meta information for different struct fields, for some use cases, the programmers can control the action they want to apply on different fields.
+As with G1, this modification also happens in three different ways.
+Using attributes for struct fields, apart from being able to store different meta-information for different struct fields, for some use cases, the programmers can control the action they want to apply on different fields.
 

### 2022-09-16 introduction.tex
renamed patterns ids, completed descs for 19 patterns, changed to EMSE template, some text for RQ1 is provided

--- 2022-09-12 1_introduction.tex
@@
-% can we drop the first paragraph altogether?
-Producing bugs has been an unavoidable part of computer programming which happens unintentionally.
+Producing bugs has been an unavoidable part of computer programming.
 Researchers have been trying to help software engineers to be more aware of the bug producing practices in development and how to step towards creating bug-free software.
 One useful step towards that goal would be studying common bug patterns (and their corresponding fixes) that occur across multiple software systems.
-Research efforts for categorizing common bug patterns date back to 1975 where Edres tried to categorize various bug patterns which occurred in operating system program.
+Research efforts for categorizing common bug patterns date back to 1975 where Edres tried to categorize various bug patterns which occurred in operating system implementations.
 
@@
 
-% how does this application below differ from the one above?
-Similarly, if we are not observing the desired behaviour from our program, that would indicate that there lies a fault within our program.
+Similarly, if we are observing undesired behaviour from our program, that would indicate that there lies a fault within our program.
+But where is the fault?
 Fault localization is the process of automatically finding the faulty statement(s) in our program.
-If a fault localization tool knows about the common bug patterns that can appear in the underlying programming language, it can prioritize the code segment candidates based on their potentials of producing bugs.
-That would result in a more robust tool as it leverages from historical data.
+If a fault localization tool knows about the common bug patterns that can appear in the underlying programming language, it can prioritize code segment candidates based on their potentials of producing bugs.
+That would result in a more robust tool that leverages historical data.
 
@@
 As Monperrus describes, the fewer frequent bug patterns we keep in our software, the less cost we would have to pay for software maintenance.
-% maybe more useful for PL design than for programmers; on the other hand I was just talking about stunts in movies and how people are probably more in danger from the less dangerous ones that they do 1000 times rather than the more dangerous ones that they do 2 times; it would be neat if there was a citation on that, but it's a bit of a distraction and doesn't actually help us get our paper in!
 
-% sentence too long; split it.
-Moving further afield, researchers have recently been trying to apply deep learning techniques on programs to learn about the semantics of the code or training models which can modify programs towards a specific goal (e.g.
+Moving further afield, researchers have been exploiting deep learning techniques to learn about the semantics of computer programs and create program modifier models that have a specific goal (e.g.
 bug fix).
@@
 
-% put into a new related work section
-% this is why numbers in front of filenames are kind of annoying, eh
 Many studies have investigated detecting and categorizing bug fix patterns in general, or for a specific programming language.
-Our study is focused on pervasive bug fix patterns in the specific language Rust.
+In this work, we study common bug patterns that appear in Rust projects.
+Rust is a relatively new language that allows programmers to write programs that have low-level control over resources, as well as high-level safety guarantees.
+What distinguishes Rust from the other systems programming languages, e.g.
+C/C++, is its ownership framework, which is an integral part of Rust.
+The ownership framework keeps track of references to different memory locations, and will free them once they go out of scope.
+If a program passes the compiler’s checks, it must meet several safety properties; e.g.
+it has no dangling references or double-frees.
+These problems had to be manually ensured in languages like C/C++.
+Moreover, not only does the ownership concept remove the need for a garbage collector in the runtime environment, it does not require the programmer to explicitly do certain safety tasks.
+Due to the memory and thread safety guarantees that Rust provides, many software companies adopted Rust in production.
+For example, Dropbox decided to migrate to Rust for developing their storage system, since they were not completely satisfied with the features provided by AWS’s S3 or Golang.
 
-% I'll reread this part once it's in its own section.
-\paragraph{General bug fix patterns}
-Yang et al. proposed a mining approach to detect Python bug fix patterns via studying
-fine-grained fixing code changes. In their research, they also examined what portion of the bugs
-could be fixed using automated bug fixing approaches. Moreover, they evaluated the fix patterns in
-the wild and realized that 37\% of the buggy codes could be matched by the fix patterns they had
-found.
-Hanam et al. conducted a similar research but for finding Javascript most pervasive bug fix
-patterns. They presented a large-scale study of the most common bug fix patterns by mining 105K
-commits from 134 server-side JavaScript projects. They used DBSCAN clustering algorithm
-and divided bug fixing change types into 219 clusters, from which they extracted 13 pervasive
-cross-project bug fix patterns.
-
-\paragraph{Rust bug fix patterns}
-In this work, we study common bug patterns that appear in Rust projects. Rust is a relatively new language that allows programmers to write programs that have low-level
-control over resources, as well as high-level safety guarantees. Originally, Rust was a side
-project developed by Graydon Hoare, a Mozilla employee. The company started to fund this project
-in 2009, since they intended to use Rust in some of the low-level Firefox components.
-What distinguishes Rust from the other systems programming languages, e.g. C/C++, is the
-ownership framework that has been incorporated inside the language. The ownership framework keeps the track of the references to different memory locations, and will free them once they
-go out of scope. If the program passes the compiler checks, it automatically guarantees several
-safety problems e.g. dangling references, or double-free problem, that had to be taken care of
-manually in languages like C/C++. Moreover, not only the ownership concept removes the
-need for a garbage collector in the runtime environment, it does not require the programmer to
-explicitly do certain safety tasks, like freeing references.
-Due to the memory and thread safety guarantees that Rust provides, many software companies
-adopted Rust in production. For example, Dropbox decided to migrate to Rust for developing their
-storage system, since they were not completely satisfied with the features provided by AWS’s S3 or
-Golang.
-
-% ok, now back to the introduction section
-% did we talk about ownership yet?
 In Rust, the so-called Borrow Checker component enforces ownership rules.
 Because of the Borrow Checker's importance in Rust, we split patterns into two groups: (1) non-borrow-checker related patterns (General Patterns), and (2) borrow-checker related patterns (BC-Related Patterns).
-%This distinction was really helpful as it made our tool discover clearer results.
-- can you write something better?
-
 We present our discovered insights of our empirical study in the following four research
@@
 
-% we talked about these, right?
-\textbf{RQ1. How successful is our data modelling approach in capturing all the necessary data?}
+\textbf{RQ1. How successful is our data modelling approach in representing program change?}
 
-\textbf{RQ2. What are the most common AST structures in the bug fixing changes?}
+\textbf{RQ2. What are the general fix patterns in Rust, and how often do they apply?}
 
-\textbf{RQ3. What are the general fix patterns in Rust?}
-
-\textbf{RQ4. Which fix patterns are related to Borrow Checker in Rust?}
+\textbf{RQ3. What are borrow-checker related fix patterns in Rust, and how often do they apply?}

### 2022-09-16 methodology.tex
renamed patterns ids, completed descs for 19 patterns, changed to EMSE template, some text for RQ1 is provided

\subsection{Overview}

\subsection{Data modelling}

\subsubsection{Parsing Programs}

\subsubsection{Path Linearization}

\subsubsection{Weighting Scheme}

\subsubsection{Obtaining the change essence}

\subsection{Mining Repositories}

\subsection{Clustering Data}

\begin{algorithm}
\caption{An algorithm with caption}\label{alg:cap}
\hspace*{2mm} \textbf{Input:} $R$ (target repositories)  \\
\hspace*{2mm} \textbf{Output:} $D_g$ (General code changes) \\
\hspace*{2mm} \textbf{Output:} $D_b$ (BC-related code changes)
\begin{algorithmic}
\State $D_g \leftarrow \phi$
\State $D_b \leftarrow \phi$
\For{$r \in R$}
    \For{$c \in \textsc{ExtractCommits}(r)$}
        \If{$c.msg$ contains bug fixing related keywords}
            \For{$\{f_b, f_a\} \in \textsc{GetModifiedRustFiles}(c)$}
                \For{$e \in \textsc{ASTDiff}(\textsc{Parse}(f_b), \textsc{Parse}(f_a))$}
                    \State $DP \leftarrow \textsc{GetDataPoint}(e)$
                    \State $DP \leftarrow c.hash \cup f_a.name \cup e.scope \cup DP $
                    \If{$\textsc{IsBCRelated}(e)$}
                        \State $DP \leftarrow \textsc{GetBCKeyword}(e) \cup DP $
                        \State $D_b \leftarrow D_b \cup DP$
                    \Else
                        \State $D_g \leftarrow D_g \cup DP$
                    \EndIf
                \EndFor
            \EndFor
        \EndIf
    \EndFor
\EndFor
\end{algorithmic}
\end{algorithm}

### 2022-09-16 results.tex
renamed patterns ids, completed descs for 19 patterns, changed to EMSE template, some text for RQ1 is provided

--- 2022-09-12 3_results.tex
@@
-% some general comments about the taxonomy go here
+\subsection{Evaluating the data modelling approach (RQ1)}
 
-% take one example and explain it, e.g. ++ means ..., -- means ....
-% rename to e.g. B.ref.add
-% need quantitative data
+As discussed in the previous section, we used an encoding mechanism to incorporate AST information in fix-sized datapoints which then could be used in the DBSCAN clustering algorithm.
+We used heuristics (weighting scheme) while encoding our datapoints to make the most important program elements stand out between all the non terminals seen in the AST.
+Here we want to evaluate the effectiveness of our encoding.
+
+The general idea for our evaluation is whether the most important program elements has the high values in our encoding.
+The analyzer compares the actual change in the code and compares it to the visual representation of its respective data point.
+We used circle pack figures to visualize each data point.
+In these figures, the radius of the circle of each non terminal shows its importance in our encoding (Fig ref).
+
+We choose 50 random samples from each of our general and borrow-checker databases.
+These random samples are drawn from the clustered datapoints.
+The reason behind this decision is the same as why we collected a lot of noise points, that is, the majority of the commits contain multi-hunk complex changes and the key changing elements are difficult to determine.
+
+We carried out two manual experiments for evaluating our encoding mechanism.
+In the first experiment, the analyzer investigates whether the most important element has the largest radius or not.
+In the second experiment, the analyzer investigates whether the most important element is included in three largest circles or not.
+The analyzer should give a binary answer to these questions.
+Both authors carry out these experiments independently.
+At last, we report the results as the portion of the changes which both authors agreed upon its encoding correctness.
+The final results are TBD\% and TBD\% correctness in the first and second experiments, respectively, which shows the effectiveness of our encoding mechanism.
+
+\subsection{Common Bug Fix Patterns (RQ2)}
 
@@
 Inner Attributes apply to the program element encapsulating them, while Outer Attributes apply to the program element coming after the attribute.
-% does the inner/outer distinction matter?
+\\
 
-\subsubsection{G1. Modifying the attributes of structs}
+\noindent \textit{ \textbf{G.attr.struct.} Modifying the attributes of structs}
 
 \begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-// G1.1: Adding attributes to a struct
+// G.attr.struct.add: Adding attributes to a struct
 ++ #[derive(Debug, Clone)]
@@
 
-// G1.2: Removing attributes from a struct
+//G.attr.struct.drop: Removing attributes from a struct
 -- #[derive(Debug, Clone)]
@@
 
-// G1.3: Changing an existing attribute of a struct
+// G.attr.struct.change: Changing an existing attribute of a struct
 -- #[derive(Debug)]
@@
 The attributes are Outer Attributes as we can find struct related non-terminals within the ASTDiff of the instances of this pattern.
+\\
 
-\subsubsection{G2. Modifying the attributes of struct fields}
+\noindent \textit{ \textbf{G.attr.struct.field} Modifying the attributes of struct fields}
 
 \begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-// G2.1: Adding attributes to struct fields
+// G.attr.struct.field.add: Adding attributes to struct fields
 #[derive(Serialize)]
@@
 
-// G2.2: Removing attributes from struct fields
+// G.attr.struct.field.drop: Removing attributes from struct fields
 #[derive(Serialize)]
@@
 
-// G2.3: Changing attributes of struct fields
+// G.attr.struct.field.change: Changing attributes of struct fields
 #[derive(Serialize)]
@@
 
-% The sentence "Using..." doesn't actually have an appropriate verb
 \noindent \textbf{Description:} In this pattern, the attribute set of a struct field changes.
@@
 \subsection{Struct} 
-\subsubsection{G3. Adding/Removing a Struct field}
+\noindent \textit{ \textbf{G.struct.field} Adding/Removing a struct field}
 
 \begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-// G3.1: Adding a new field to a struct
+// G.struct.field.add: Adding a new field to a struct
 struct InnerListeners {
@@
 
-// G3.2: Removing a field from a struct
+// G.struct.field.drop: Removing a field from a struct
 struct SnapshotService<U, R> {
@@
 The removal of a field  happens when the developer realizes that the field is expendable with respect to the behaviour they want to implement.
+\\
 
-\subsubsection{G4. Modifying the access modifiers of struct fields}
+\noindent \textit{ \textbf{G.struct.field.pub} Modifying the access modifiers of struct fields}
 
@@
 
-// G4.1: Adding pub to a field
+// G.struct.field.pub.add: Adding pub to a field
 pub struct HtmlBlock {
@@
 
-// G4.2: Removing pub from a field
+// G.struct.field.pub.drop: Removing pub from a field
 pub struct Languages {
@@
 
-\subsubsection{G5. Different fields order when initializing Struct}
+\subsection{Option}
+\vspace{3mm}
 
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
--- return Poll::Ready(ReadyEvent { ready, tick });
-++ return Poll::Ready(ReadyEvent { tick, ready });
-
-\end{lstlisting}
-
-\noindent \textbf{Description:} Remove... ???
-
-\subsection{Option}
-\subsubsection{G6. Changing field type T to T$\langle$Option$\rangle$}
+\noindent \textit{ \textbf{G.option} Changing field type T to T$\langle$Option$\rangle$}
 
@@
 \subsection{Types}
-\subsubsection{G7. Changing a struct field type}
+\vspace{3mm}
+
+\noindent \textit{ \textbf{G.type.field} Changing a struct field type}
 
@@
 Also it can happen for refactoring purposes or performance enhancement.
+\\
 
-\subsubsection{G8. Changing a generic type}
+\noindent \textit{ \textbf{G.type.generic} Changing a generic type}
 
 \begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-
 #[derive(Debug, Clone, PartialEq)]
@@
 
-\noindent \textbf{Description:} This a change of the generic (parametric) type of a struct field. Similar to G7, this change can also happen for the mentioned reasons.
+\noindent \textbf{Description:} This a change of the generic (parametric) type of a struct field. Similar to G7, this change can also happen for the mentioned reasons. \\
 
-\subsubsection{G9. Change in enum variant value type}
+\noindent \textit{ \textbf{G.type.enum.variant} Change in enum variant value type}
 
 \begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-
 #[derive(Debug, Clone, PartialEq, Eq)]
@@
 This change pattern is associated with a change in enum variant types.
+\\
 
-\subsubsection{G10. Changing the type inside a tuple}
+\noindent \textit{ \textbf{G.type.tuple} Changing the type inside a tuple}
 
 \begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-
 #[allow(dead_code)]
@@
 }
-
 \end{lstlisting}
@@
 \subsection{Traits}
-\subsubsection{G11. TraitBound Change}
+\noindent \textit{ \textbf{G.trait.bound} TraitBound Change}
 
 \begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-
 -- pub struct EventLoop<T: tty::EventedReadWrite> {
@@
 
-\subsubsection{G12. Implementing a different trait for a struct}
+\subsection{BC-Related Bug Fix Patterns (RQ3)}
+
+\subsubsection{Clone}
+
+\noindent \textit{ \textbf{BC.clone.ref} Dropping clone and adding borrowing}
 
 \begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-
--- impl ToDebug for ShellError {
-++ impl FormatDebug for ShellError {
-        // Implementing functions ...
-   }
-
+-- let field_id = schema.get_or_create(attribute.clone())?;
+++ let field_id = schema.get_or_create(&attribute)?;
 \end{lstlisting}
 
-\noindent \textbf{Description:} Remove... ???
+\noindent \textbf{Description:} Unlike the previous pattern, this pattern is not detected by Clippy.
+The change happens when the programmers realizes that the cloning of a variable is unnecessary, however, we also want to preserve the ownership of the variable within the current scope.
+That is why in this pattern a clone is turned into a borrowing.
 
-\subsection{Ref and Deref}
-\subsubsection{B1. Adding $\&$}
+Similar to the previous pattern, this does not change the program behaviour and is merely for performance purposes.
+It was the case in one of the commits of the project tauri-apps/tauri, where the clone to borrowing significantly optimized CPU usage.
+They discussed that the cloning of some types such as HashMap was expensive.
+\\
+
+\noindent \textit{ \textbf{BC.clone.drop} Removing clone}
+
+\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
+start_plugin_process(
+--  manifest.clone(),
+++  manifest,
+    self.next_plugin_id(),
+    self.self_ref.as_ref().unwrap().clone(),
+);
+\end{lstlisting}
+
+\noindent \textbf{Description:} This pattern is referred to as redundant clone in Clippy Lints.
+Clippy detects whether a variable clone is necessary and could we have used the variable itself or not.
+This change does not modify the program behaviour, however it can boost the performance significantly if we repeatedly cloning large variables.
+
+\subsubsection{Ref and Deref}
+
+\noindent \textit{ \textbf{BC.ref.add} Adding Borrowing}
 
@@
 
-To give an example, in starship/starship after one of the commits \footnote{\url{https://github.com/starship/starship/commit/56d475578ea508631275772127f49a6949fea6b0}, file \texttt{git\_branches.rs}}, the developer borrows the value of $repo\_dir$ to keep the ownership of the object in the current scope.
+To give an example, in starship/starship after one of the commits \footnote{\url{https://github.com/starship/starship/commit/56d475578ea508631275772127f49a6949fea6b0}{git_branches.rs}}, the developer borrows the value of $repo\_dir$ to keep the ownership of the object in the current scope.
 Now this object can then be used in another function invocation ($remove\_dir\_all(repo\_dir)$) which fixes a bug.
+\\
 
-\subsubsection{B2. Removing $\&$}
+\noindent \textit{ \textbf{BC.ref.drop} Removing Borrowing}
 
@@
 This does not change the behaviour of the program, however, it does improve the performance, as the borrow checker need not to return the ownership after the execution of the function.
+\\
 
-\subsubsection{B3. Dropping \& from func decl and adding * to match statement}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
--- fn to_color_choice(&self) -> ColorChoice {
---     match *self {
-++ fn to_color_choice(self) -> ColorChoice {
-++     match self {
-           // Statements
-       }
-\end{lstlisting}
-
-\noindent \textbf{Description:} ???
-
-\subsubsection{B4. Modification in dereferencing a string prior to calling .to\_string()}
+\noindent \textit{ \textbf{BC.deref.to\_string()} Modification in dereferencing a string prior to calling to\_string()}
 
@@
 
-\subsection{Mut}
-\subsubsection{B5. Adding mut}
+\subsubsection{Mut}
+
+\noindent \textit{ \textbf{BC.mut.add} Adding mutability}
 
 \begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-
 parser::Parser::new(args)
@@
         })?
-
 \end{lstlisting}
@@
 
-\subsubsection{B6. Removing mut}
+\noindent \textit{ \textbf{BC.mut.drop} Dropping mutability}
 
@@
 
-\subsection{Vector}
-\subsubsection{B7. Vector reference to slice}
+\subsubsection{Vector}
+
+\noindent \textit{ \textbf{BC.vec.slice} Changing a Vec reference to Slice}
 
@@
 \end{lstlisting}
+\noindent \textbf{Description:} In struct field definition or the types of the function formal parameters, type $\&Vec\<T\>$ changes to $\&[T]$.
+Clippy can detect the opportunity to apply this pattern.
+The reason is reducing type complication, as slice types in the form of $\&[T]$ or $\&str$ are usually sufficient for most use cases.
 
-\noindent \textbf{Description:} ???
-
-\subsection{Lifetime}
-\subsubsection{B8. Adding generic Lifetime}
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-#[allow(dead_code)]
--- fn api_error(webview: &mut Webview, error_fn: String, message: &str) {
-++ fn api_error(webview: &mut Webview<'_>, error_fn: String, message: &str) {
-  let reject_code = tauri_api::rpc::format_callback(error_fn, message);
-  webview.eval(&reject_code)
-}
-
-\end{lstlisting}
-
-\noindent \textbf{Description:} ???
-
-\subsubsection{B9. Dropping static lifetime}
+\subsubsection{Lifetime}
+\noindent \textit{ \textbf{BC.lifetime.static.drop} Dropping static lifetime}
 
 \begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-
 -- const QUEUE_SIZES: &'static [u16] = &[QUEUE_SIZE];
@@
 
-\noindent \textbf{Description:} ???
-
-\subsubsection{B10. Dropping Lifetime in field definition}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-pub struct Table<'a> {
-    columns: Vec<Column>,
-    colours: &'a Colours,
-    env: &'a Environment,
-    widths: TableWidths,
-    -- time_format: &'a TimeFormat,
-    ++ time_format: TimeFormat,
-    size_format: SizeFormat,
-    git: Option<&'a GitCache>,
-}
-\end{lstlisting}
-
-\noindent \textbf{Description:} ???
-
-\subsection{Clone}
-\subsubsection{B11. Dropping clone (and adding dereference)}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-pub fn reset(&mut self, template: &Cell) {
-    // memcpy template to self
---  *self = template.clone();
-++  *self = *template;
-}
-\end{lstlisting}
-
-\noindent \textbf{Description:} This pattern is referred to as redundant clone in Clippy Lints.
-Clippy detects whether a variable clone is necessary and could we have used the variable itself or not.
-This change does not modify the program behaviour, however it can boost the performance significantly if we repeatedly cloning large variables.
-
-\subsubsection{B12. Dropping clone}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-if let Some(var_id) = &var.var_id {
---  stack.add_var(*var_id, input);
-++  stack.add_var(*var_id, input.clone());
-}
-
-start_plugin_process(
---  manifest.clone(),
-++  manifest,
-    self.next_plugin_id(),
-    self.self_ref.as_ref().unwrap().clone(),
-);
-
-\end{lstlisting}
-
-\noindent \textbf{Description:} This pattern is referred to as redundant clone in Clippy Lints.
-Clippy detects whether a variable clone is necessary and could we have used the variable itself or not.
-This change does not modify the program behaviour, however it can boost the performance significantly if we repeatedly cloning large variables.
-
-\subsubsection{B13. Dropping clone and adding reference}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
--- let field_id = schema.get_or_create(attribute.clone())?;
-++ let field_id = schema.get_or_create(&attribute)?;
-\end{lstlisting}
-
-\noindent \textbf{Description:} Unlike the previous pattern, this pattern is not detected by Clippy.
-The change happens when the programmers realizes that the cloning of a variable is unnecessary, however, we also want to preserve the ownership of the variable within the current scope.
-That is why in this pattern a clone is turned into a borrowing.
-
-Similar to the previous pattern, this does not change the program behaviour and is merely for performance purposes.
-It was the case in one of the commits of the project tauri-apps/tauri, where the clone to borrowing significantly optimized CPU usage.
-They discussed that the cloning of some types such as HashMap was expensive.
+\noindent \textbf{Description:} Removing the static keyword from a const variable declaration.
+This change can be applied using Clippy.
+If the presence of static is not required, it is better that we omit it as keeping it might create complicate types in our program.
 
@@
 
-\hskip-2.5cm\begin{tabular}{|l|l|l|l|}
+\begin{tabular}{|l|l|l|l|}
 \hline
 
-\multirow{12}{*}{General}        
-& \multirow{2}{*}{Attributes}    & G1  & Adding Attributes to Structs                                        \\
-                                 &                                & G2  & Adding Attributes to Struct Fields                                  \\\cline{2-3}
-                                 & \multirow{3}{*}{Struct}        & G3  & Adding/Removing a Struct field                                      \\
-                                 &                                & G4  & Adding/Removing pub to/from struct field                            \\
-                                 &                                & G5  & Different fields order when initializinng Struct                       \\\cline{2-3}
-                                 & Option                         & G6  & Changing field type T to Option                                          \\\cline{2-3}
-                                 & \multirow{4}{*}{Types}         & G7  & Changing a Struct field type                                        \\
-                                 &                                & G8  & Changing a generic type                                             \\
-                                 &                                & G9  & Change in enum variant value type                                   \\
-                                 &                                & G10 & Changing the type inside a tuple                                         \\\cline{2-3}
-                                 & \multirow{2}{*}{Traits}        & G11 & TraitBound Change                                                   \\
-                                 &                                & G12 & Implementing a different trait for a struct.                            \\\cline{1-3}
-\multirow{13}{*}{Borrow Checker} & \multirow{4}{*}{Ref and Deref} & B1  & Adding \&                                                           \\
-                                 &                                & B2  & Removing \&                                                         \\
-                                 &                                & B3  & Dropping \& from function declaration and dereferencing match \\
-                                 &                                & B4  & Adding *                                                            \\\cline{2-3}
-                                 & \multirow{2}{*}{Mut}           & B5  & Adding mut                                                          \\
-                                 &                                & B6  & Removing mut                                                        \\\cline{2-3}
-                                 & Vector                         & B7  & Vector reference to slice                                           \\\cline{2-3}
-                                 & \multirow{3}{*}{Lifetime}      & B8  & Adding generic Lifetime                                             \\
-                                 &                                & B9  & Dropping static lifetime                                            \\
-                                 &                                & B10 & Dropping Lifetime in field definition                               \\\cline{2-3}
-                                 & \multirow{3}{*}{Clone}         & B11 & Dropping clone and adding dereference                               \\
-                                 &                                & B12 & Dropping clone                                                      \\
-                                 &                                & B13 & Dropping clone and adding reference  \&
+\multirow{10}{*}{General}        
+& \multirow{2}{*}{Attributes}    & G.attr.struct  & Modifying the attributes of structs                                        \\
+                                 &                                & G.attr.struct.field  & Modifying the attributes of struct fields                                  \\\cline{2-3}
+                                 & \multirow{2}{*}{Struct}        & G.struct.field  & Adding/Removing a struct field                                      \\
+                                 &                                & G.struct.field.pub  & Modifying the access modifiers of struct fields     
+                                  \\\cline{2-3}
+                                 & Option                         & G.option  & Changing field type T to T$\langle$Option$\rangle$                     \\\cline{2-3}
+                                 & \multirow{4}{*}{Types}         & G.type.field  & Changing a struct field type                                        \\
+                                 &                                & G.type.generic  & Changing a generic type                                             \\
+                                 &                                & G.type.enum.variant  & Change in enum variant value type                                   \\
+                                 &                                & G.type.tuple & Changing the type inside a tuple                                         \\\cline{2-3}
+                                 & \multirow{1}{*}{Traits}        & G.trait.bound & TraitBound Change                                               \\\cline{1-3}
+\multirow{9}{*}{Borrow Checker} &
+\multirow{2}{*}{Clone}                                         & BC.clone.ref & Dropping clone and adding borrowing                                                   \\
+                                 &                                & BC.clone.drop & Dropping clone   
+                                    \\\cline{2-3}
+    & \multirow{3}{*}{Ref and Deref} & BC.ref.add  & Adding Borrowing                                                           \\
+                                 &                                & BC.ref.drop  & Removing Borrowing                                                         \\
+                                 &                                & BC.deref.to\_string  & Modification in dereferencing a string prior to calling to\_string()                                                            \\\cline{2-3}
+                                 & \multirow{2}{*}{Mut}           & BC.mut.add  & Adding mutability                                                          \\
+                                 &                                & BC.mut.drop  & Dropping mutability                                                        \\\cline{2-3}
+                                 & Vector                         & BC.vec.slice  & Changing a Vec reference to Slice                                           \\\cline{2-3}
+                                 & \multirow{1}{*}{Lifetime}      
+                                                                 & BC.lifetime.static  & Dropping static lifetime    &                        
+                                 \hline
 

### 2022-09-19 methodology.tex
65% methodology content added

@@
 \subsection{Overview}
+
+The general idea for automatically finding the pervasive code patterns within a list of programs is to parse the programs and encode the important parts of the derived AST in fixed sized datapoints.
+Then, we can use the clustering algorithms, to cluster these datapoints, with the hope that each cluster manifests a certain pattern of change.
+
+Our pipeline for finding pervasive patterns in Rust programs also follows this idea.
+We use Syn as our Rust parser.
+Syn is a Rust crate which is designed to be employed in Rust procedural macros, however its APIs can be used generally.
+We decided to implement our pipeline completely in Python.
+That is why, we had to develop a preprocessor to transform Syn AST output to Python dictionary.
+
+To find out what has changed in a certain commit, we need two code revisions; the revision before the commit, and the revision after it.
+After parsing these two revisions, we will have two ASTs which are encoded in two separate Python dictionaries.
+The change in the code can be inferred by analyzing the difference of these two trees (ASTDiff).
+To derive the ASTDiff we use dictdiff Python library.
+
+Now, a challenging problem is how to specify the most important structures within this ASTDiff.
+This step is crucial as it is through this step that we select the semantic information we want to encode in our datapoints later on.
+In Section ?
+and ?, we provide a detailed explanation of the method we used to select the most important semantic information which then helped us to effectively obtain meaningful clusters.
+
+We used Pydriller, which is a repository miner for Python.
+Our target repositories were the top 18 most starred open source Rust projects on GitHub.
+We mined all the bug related commits and ran them through our pipeline.
+Then, we applied DBSCAN clustering algorithm on the obtained datapoints.
+In Section ?
+we discuss why we chose DBSCAN for clustering and how we tuned it to improve clusters quality.
 
@@
 
+For parsing our Rust programs, we use Syn crate which is a parsing library for parsing a stream of tokens into a syntax tree.
+The parser covers all the Rust language features and the list of all available tokens can be found in its web page.
+The token types (non terminals) are primarily divided into two main groups; Structs and Enums.
+The dimensions of our datapoints have been designed in a way that it adheres to the non terminals described in these two groups.
+We parse two versions of a Rust changed file in a commit; the file before the commit and after it.
+We will obtain two trees created by Syn.
+The change that transforms the first tree to the second one, yields the fix which was applied in the commit.
+
+To find this difference, we use dictdiff, a Python library to find the diff of two Python dictionaries.
+That is why, we are required to develop a program transformer which transforms Syn tree output to Python dictionary.
+We wrote a simple transformer using PLY, which is is an implementation of lex and yacc parsing tools for Python.
+After obtaining the syntax trees in Python dictionary format, we are able to use dictdiff tool to determine the ASTDiff.
+The dictdiff tool also determines the type of the modification; 1.
+'add': A new structure has been added to the tree, 2.
+'remove': An structure has been dropped from the tree, 3.
+'change': The content of a sub structure of the first tree has changed.
+In addition to that, dictdiff also determines the context within which the modification has occured.
+This context happens to be the root elements (FIND A BETTER TERM FOR ROOT ELEMENTS) in our file, e.g.
+function blocks, struct blocks, impl blocks, etc.
+The modified content within a context is still in the form of tree.
+Our objective right now is to find primary elements in our diff that describe the semantics of occurred change.
+
 \subsubsection{Path Linearization}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.8\textwidth]{linearization.png}
+\label{f2}
+\caption{Path Linearization}
+\end{figure}
+
+The output of dictdiff is a list of diffs, where each diff is constituted of three parts.
+The first part specifies whether the modification is of type 'add', 'remove', 'change'.
+The second part identifies the context of the modification, that is, the path from the root of the tree to the subtree in which the modification has occured.
+The third part is the content of the change and includes all the involved terminals and non terminals in the change.
+Our tool looks for changes within each root-level scope.
+That is, our assumption is that a change pattern occurs within one root-level scope.
+Our perspective does not cover the patterns that involve changes in multiple functions or multiple files.
+However, if the change happens in multiple lines within one scope, whether they are continuously one after another or separate from each other, our tool would be able to detect it.
+
+Looking at the tree encoded in the content part of each diff, we realized that each path to a leaf represents one contributing element to the change.
+Through a Depth First Search we can collect all of these paths.
+We store only the non terminals that exist in the Syn token types list.
+Also certain keywords are not considered as non terminals in the syntax tree (e.g.
+clone, Box), but as they have a special meaning the language, we also store them in the path.
+The outcome of the DFS search is a set of linearized paths.
+
+Fig \ref{f2} shows a simple change and its respective linearized paths.
+In the change, the second argument sends a borrow of the variable as opposed to sending the ownership entirely, and in the third argument, instead of sending the head into the function scope, we send a clone of it.
+A simplification of the derived diff of the two code revisions is shown in the figure.
+A DFS search on this tree would yield three different paths shown in three different colors.
+The path linearizes the hierarchy of non terminals involved in that specific path which represents one contributing factor to the change.
+We found it crucial to record the order of all non terminals involved in a path as it gives us freedom to control the degree of order-sensitiveness in our datapoint encoding.
+
+The next question is how to transform these paths to a fixed size datapoint.
+First step, we need to define a fixed set of columns.
+The number of columns would depend on the amount of information we want to encode in the datapoints.
+The naivest approach would be to only report the observed non terminals within the diff.
+That yields a fully order-insensitive representation.
+In such encoding, for instance, there is no difference between two nested if statements and two if statements beside each other.
+Theoritically, to reach full order-sensitiveness, we would need encode all possible combinations of token types as our dimensions, which yields an astronomical number of dimensions.
+Apart from that, not all combinations adheres to Rust syntax.
+A reasonable work around would be to categorize the non terminals and set the dimensions as all possible combinations of these categories (Similar to JS paper, ref).
+This would reduce the number of dimensions and provide more reasonable and syntactically-correct combinations.
+However, after data collections we would still encounter a sparse dataset.
+
+In our methodology, similar to the first approach, we define the columns of our dataset, as the set of token types we collected from Syn.
+However, to achieve an acceptable level of order-sensitiveness, we carried out two auxiliary steps.
+Firstly, we record the number of occurrences of token types within the paths at each scope.
+The reason behind this decision is because different order of program elements yields a different number of occurrences of underlying token types.
+Secondly, we semi-automatically design a weighting scheme to give more value to certain non terminal we think are more important in Rust.
 
@@
 
-\subsubsection{Obtaining the change essence}
+Our main objective is that our datapoints manifest the change with high level of precision.
+For instance, the purple path in Fig \ref{f2} can be described as a change in local variable declaration, a change in a function call, calling a method of one of the function arguments, or calling clone() of one of the function arguments.
+Obviously, the last description is the most favorable one, as it offers the most accurate description.
+The whole idea behind our weighting scheme is to be able to identify the change accurately.
+
+A simple heuristic is that the closer we get to the leaves in the ASTDiff, the more important token types get.
+Another observation is that the elements closer to the root tend to be repeated in all the paths within one ASTDiff (Similar to ItemFn, Local, ExprCall in Fig \ref{f2}).
+Exploiting the latter observation, to obtain a basic knowledge about the number of repetitions of token types, we mined the last 20 bug related commits of the last 20 most starred rust projects on GitHub, and ran them through of pipeline.
+We recorded the number of occurrences of each token type in total.
+Then we gave the value of the inverse of the number of occurrences to each token.
+The intuition is that the higher the number of occurrences of a token type, the closer it is usually seen to the root.
+Hence, it probably does not provide us with the most accurate description of the change.
+In addition to that, as we wanted to make sure that our tool captures the patterns related to borrow checker, we manually raised the weights of the token types related to borrow checker.
+That is why we characterize our weighting scheme as semi-automatically designed.
+
+At last, each datapoint is created by the multiplying the number of occurrences of each token type to its weight and placing the result under its respective column.
+We call the set of final values the essence of the change, as we claim the values of token types manifest their importance in the change.
 
@@
 
-\subsection{Clustering Data}
+\begin{figure}[h]
+\centering
+\includegraphics[width=1\textwidth]{repos.png}
+\label{f3}
+\caption{Target Repositories}
+\end{figure}
+
+Now that we have set up our code analysis pipeline, we can start mining the Rust repositories.
+We target the top 18 most starred Rust projects on GitHub, at the time of data collection.
+The details of these projects are outlined in Table \ref{f3}.
+We computed the average code churn of the files within the projects.
+denoland/deno and swc-project/swc are the projects with largest number of LOCs and average code churn.
+Unsurprisingly, we captured a lot of instances from these two projects in our final clusters.
 
@@
 \end{algorithm}
+
+%% Describing algorithm
+
+\subsection{Clustering Data}
+
+%% Why DBSCAN
+
+%% Choosing Parameters
+
+%% Figure
+
+%% Manual Analysis for finding best parameter
+
+%% Final Clusters

### 2022-09-19 methodology.tex
comments

@@
-\subsection{Overview}
+% why fixed-size?
+% or are you saying that everyone does this, and then in the next paragraph, say that we do this too
+% The reason is that both distance-based and density-based clustering algorithms work with fixed size datapoints.
+The general idea for automatically finding the pervasive code patterns within a list of programs is to parse the programs and encode the important parts of the derived AST in fixed sized datapoints.
+Then, one uses a clustering algorithms to cluster these datapoints.
+Work in this vein makes the assumption that each cluster corresponds to one change pattern.
 
-The general idea for automatically finding the pervasive code patterns within a list of programs is to parse the programs and encode the important parts of the derived AST in fixed sized datapoints.
-Then, we can use the clustering algorithms, to cluster these datapoints, with the hope that each cluster manifests a certain pattern of change.
+% put a footnote with the URL for Syn
+% which APIs can't be used generally?
+Our pipeline for finding pervasive patterns in Rust programs also follows this idea.
+We decided to implement our pipeline completely in Python.
+To make this work, we need a way to get an AST as a Python dictionary.
+Syn is a Rust crate which is designed to be employed in Rust procedural macros.
+However, it includes a parser which is suitable for our purposes; we simply had to write a preprocessor to transform Syn AST output into Python dictionaries.
 
-Our pipeline for finding pervasive patterns in Rust programs also follows this idea.
-We use Syn as our Rust parser.
-Syn is a Rust crate which is designed to be employed in Rust procedural macros, however its APIs can be used generally.
-We decided to implement our pipeline completely in Python.
-That is why, we had to develop a preprocessor to transform Syn AST output to Python dictionary.
+A change pattern categorizes a class of changes. We thus need to define what a change is. For our purposes, a change is ....
+To compute the contents of a change, we need two code revisions: the revision before the commit, and the revision after it.
+After parsing these two revisions, we will have two ASTs, encoded in two Python dictionaries.
+The change in the code can be computed by analyzing the difference of these two trees, which we call the ASTDiff.
+To derive the ASTDiff we use dictdiff Python library.
+% the wording in the two last sentences is inexact. Let's try to make clear exactly what the steps are, and use the right verbs.
 
-To find out what has changed in a certain commit, we need two code revisions; the revision before the commit, and the revision after it.
-After parsing these two revisions, we will have two ASTs which are encoded in two separate Python dictionaries.
-The change in the code can be inferred by analyzing the difference of these two trees (ASTDiff).
-To derive the ASTDiff we use dictdiff Python library.
+An ASTDiff may include an arbitrary number of semantic changes.
+We are interested in finding the most important change within an ASTDiff and encoding that change in our datapoints.
+Sections ?
+and ?
+provide a detailed explanation of how we select the most important semantic information, which then helped us to effectively obtain meaningful clusters.
+% justification of that last assertion about meaningful clusters?
+e.g.
+we found that this helped us to effectively obtain meaningful clusters; but maybe we just don't know.
 
-Now, a challenging problem is how to specify the most important structures within this ASTDiff.
-This step is crucial as it is through this step that we select the semantic information we want to encode in our datapoints later on.
-In Section ?
-and ?, we provide a detailed explanation of the method we used to select the most important semantic information which then helped us to effectively obtain meaningful clusters.
-
+% why do we use pydriller?
+% section N explains how we found the bug-related commits
 We used Pydriller, which is a repository miner for Python.
@@
 We mined all the bug related commits and ran them through our pipeline.
-Then, we applied DBSCAN clustering algorithm on the obtained datapoints.
+Then, we applied the DBSCAN clustering algorithm on the obtained datapoints.
 In Section ?

### 2022-09-19 methodology.tex
comments

@@
 
-For parsing our Rust programs, we use Syn crate which is a parsing library for parsing a stream of tokens into a syntax tree.
-The parser covers all the Rust language features and the list of all available tokens can be found in its web page.
-The token types (non terminals) are primarily divided into two main groups; Structs and Enums.
-The dimensions of our datapoints have been designed in a way that it adheres to the non terminals described in these two groups.
+% TODO: tokens are not non terminals
+% what do you mean about structs and enums?
+% try to remove redundant content with respect to what is above
+As mentioned above, we use Syn for parsing Rust.
+Syn handles all of Rust.
+Syn's nonterminals are primarily divided into two main groups; Structs and Enums.
+The dimensions of our datapoints have been designed in a way that it adheres to the nonterminals described in these two groups.
 We parse two versions of a Rust changed file in a commit; the file before the commit and after it.
 We will obtain two trees created by Syn.
-The change that transforms the first tree to the second one, yields the fix which was applied in the commit.
+The change that transforms the first tree to the second one yields the fix which was applied in the commit.
 
 To find this difference, we use dictdiff, a Python library to find the diff of two Python dictionaries.
-That is why, we are required to develop a program transformer which transforms Syn tree output to Python dictionary.
-We wrote a simple transformer using PLY, which is is an implementation of lex and yacc parsing tools for Python.
-After obtaining the syntax trees in Python dictionary format, we are able to use dictdiff tool to determine the ASTDiff.
-The dictdiff tool also determines the type of the modification; 1.
-'add': A new structure has been added to the tree, 2.
-'remove': An structure has been dropped from the tree, 3.
-'change': The content of a sub structure of the first tree has changed.
-In addition to that, dictdiff also determines the context within which the modification has occured.
+Using the PLY tool (a Python implementation of lex and yacc), we wrote a simple transformer from the Rust AST into Python dictionaries.
+After obtaining the syntax trees in Python dictionary format, we can use dictdiff to determine the ASTDiff.
+
+dictdiff tool provides modifications' types: 
+\begin{itemize}
+\item `add': A new structure has been added to the tree; 
+\item `remove': A structure has been dropped from the tree; 
+\item `change': The content of a sub structure of the first tree has changed.
+\end{itemize}
+dictdiff also determines the context within which the modification has occured.
 This context happens to be the root elements (FIND A BETTER TERM FOR ROOT ELEMENTS) in our file, e.g.
@@
 Our objective right now is to find primary elements in our diff that describe the semantics of occurred change.
+% rework the above paragraph
 

### 2022-09-19 methodology.tex
comments

@@
 
-\subsubsection{Path Linearization}
+\subsubsection{Path Extraction}
 
@@
 \includegraphics[width=0.8\textwidth]{linearization.png}
-\label{f2}
-\caption{Path Linearization}
+\caption{\label{fig:extraction}Path Extraction}
 \end{figure}
 
-The output of dictdiff is a list of diffs, where each diff is constituted of three parts.
+% again try to remove redundancy
+The output of dictdiff is a list of diffs, where each diff comprises three parts.
 The first part specifies whether the modification is of type 'add', 'remove', 'change'.
 The second part identifies the context of the modification, that is, the path from the root of the tree to the subtree in which the modification has occured.
-The third part is the content of the change and includes all the involved terminals and non terminals in the change.
+The third part is the content of the change (terminals and nonterminals).
 Our tool looks for changes within each root-level scope.
 That is, our assumption is that a change pattern occurs within one root-level scope.
-Our perspective does not cover the patterns that involve changes in multiple functions or multiple files.
+In this work, we chose to not handle patterns that involve changes in multiple functions or multiple files.
 However, if the change happens in multiple lines within one scope, whether they are continuously one after another or separate from each other, our tool would be able to detect it.
+% simplify that last sentence there
+% provide an example and then transform it into fig:extraction.
 
+% what is the Syn token types list?
 Looking at the tree encoded in the content part of each diff, we realized that each path to a leaf represents one contributing element to the change.
@@
 clone, Box), but as they have a special meaning the language, we also store them in the path.
-The outcome of the DFS search is a set of linearized paths.
+The outcome of the DFS search is a set of paths.
 
-Fig \ref{f2} shows a simple change and its respective linearized paths.
-In the change, the second argument sends a borrow of the variable as opposed to sending the ownership entirely, and in the third argument, instead of sending the head into the function scope, we send a clone of it.
-A simplification of the derived diff of the two code revisions is shown in the figure.
-A DFS search on this tree would yield three different paths shown in three different colors.
-The path linearizes the hierarchy of non terminals involved in that specific path which represents one contributing factor to the change.
+% reorganize this when you put the example up there.
+Figure~\ref{fig:extraction} shows a simple change and its respective paths.
+The change replaces \verb+tail+ with \verb+&tail+ and \verb+head+ with \verb+head.clone()+.
+That is, after the change, the second argument passes a borrow of \verb+tail+ rather than sending the ownership of \verb+tail+ to the callee; and, in the third argument, instead of sending \verb+head+ to the callee, it sends a clone of it.
+A DFS search on this tree would yield three different paths, which we show in red, green, and purple.
+The path represents the sequence of involved non terminals, which represents one contributing factor to the change (along with the context, which we do not record).
 We found it crucial to record the order of all non terminals involved in a path as it gives us freedom to control the degree of order-sensitiveness in our datapoint encoding.
+% can use fewer words in that last sentence too
 
+% OK so if you just have all observed nonterminals then you have a fixed size, but it's too big.
 The next question is how to transform these paths to a fixed size datapoint.
-First step, we need to define a fixed set of columns.
+First, we need to define a fixed set of columns.
 The number of columns would depend on the amount of information we want to encode in the datapoints.
-The naivest approach would be to only report the observed non terminals within the diff.
+The most naive approach would be to only report the observed non terminals within the diff.
 That yields a fully order-insensitive representation.
-In such encoding, for instance, there is no difference between two nested if statements and two if statements beside each other.
-Theoritically, to reach full order-sensitiveness, we would need encode all possible combinations of token types as our dimensions, which yields an astronomical number of dimensions.
-Apart from that, not all combinations adheres to Rust syntax.
-A reasonable work around would be to categorize the non terminals and set the dimensions as all possible combinations of these categories (Similar to JS paper, ref).
+In such an encoding, for instance, there is no difference between two nested if statements and two if statements beside each other.
+Theoretically, to reach full order-sensitiveness, we would need encode all possible combinations of token types as our dimensions, which yields an astronomical number of dimensions.
+Apart from that, not all combinations adhere to Rust syntax.
+A reasonable work around would be to reduce from the full set of nonterminals to a smaller set of categories, and then to set the dimension as all possible combinations of these categories (Similar to JS paper, ref).
 This would reduce the number of dimensions and provide more reasonable and syntactically-correct combinations.
-However, after data collections we would still encounter a sparse dataset.
+However, this would still yield a sparse dataset.
+% give examples of categories
+% can we provide orders of magnitude rather than "astronomical"?
 
+% We make a novel contribution here. 
 In our methodology, similar to the first approach, we define the columns of our dataset, as the set of token types we collected from Syn.
 However, to achieve an acceptable level of order-sensitiveness, we carried out two auxiliary steps.
-Firstly, we record the number of occurrences of token types within the paths at each scope.
-The reason behind this decision is because different order of program elements yields a different number of occurrences of underlying token types.
-Secondly, we semi-automatically design a weighting scheme to give more value to certain non terminal we think are more important in Rust.
+First, we record the number of occurrences of token types within the paths at each scope.
+The reason behind this decision is because a different order of program elements yields a different number of occurrences of underlying token types.
+Second, we semi-automatically design a weighting scheme to prioritize non terminals that we think are more salient for Rust bugs.
 
@@
 
+% consider: "datapoints accurately describe the most important change." 
+% put the most important change first
 Our main objective is that our datapoints manifest the change with high level of precision.
-For instance, the purple path in Fig \ref{f2} can be described as a change in local variable declaration, a change in a function call, calling a method of one of the function arguments, or calling clone() of one of the function arguments.
-Obviously, the last description is the most favorable one, as it offers the most accurate description.
+For instance, the purple (right-most) path in Figure~\ref{fig:extraction} could be described as a change in a local variable declaration; a change in a function call; calling a method of one of the function arguments; or calling clone() on one of the function arguments.
+The last description offers the most accurate description.
 The whole idea behind our weighting scheme is to be able to identify the change accurately.
 
-A simple heuristic is that the closer we get to the leaves in the ASTDiff, the more important token types get.
-Another observation is that the elements closer to the root tend to be repeated in all the paths within one ASTDiff (Similar to ItemFn, Local, ExprCall in Fig \ref{f2}).
+% I need to reread this key paragraph. - PL
+A simple heuristic is that, the closer we get to the leaves in the ASTDiff, the more important token types get.
+Another observation is that the elements closer to the root tend to be repeated in all the paths within one ASTDiff (e.g.
+ItemFn, Local, ExprCall in Figure~\ref{fig:extraction}).
 Exploiting the latter observation, to obtain a basic knowledge about the number of repetitions of token types, we mined the last 20 bug related commits of the last 20 most starred rust projects on GitHub, and ran them through of pipeline.

### 2022-09-20 methodology.tex
mining and clustering parts added, obviously still need more revising

@@
-% why fixed-size?
-% or are you saying that everyone does this, and then in the next paragraph, say that we do this too
-% The reason is that both distance-based and density-based clustering algorithms work with fixed size datapoints.
-The general idea for automatically finding the pervasive code patterns within a list of programs is to parse the programs and encode the important parts of the derived AST in fixed sized datapoints.
-Then, one uses a clustering algorithms to cluster these datapoints.
+The general idea for automatically finding the pervasive code patterns within a list of programs is to parse the programs and encode the important parts of the derived AST in fixed-size datapoints.
+We need fixed-size datapoints, as it is a requirement for both centroid-based and density-based well-known clustering algorithms.
 Work in this vein makes the assumption that each cluster corresponds to one change pattern.
 
-% put a footnote with the URL for Syn
-% which APIs can't be used generally?
 Our pipeline for finding pervasive patterns in Rust programs also follows this idea.
@@
 To make this work, we need a way to get an AST as a Python dictionary.
-Syn is a Rust crate which is designed to be employed in Rust procedural macros.
+Syn \footnote{\url{https://crates.io/crates/syn}} is a Rust crate which is designed to be employed in Rust procedural macros.
 However, it includes a parser which is suitable for our purposes; we simply had to write a preprocessor to transform Syn AST output into Python dictionaries.
 
+% TODO Define change
 A change pattern categorizes a class of changes. We thus need to define what a change is. For our purposes, a change is ....
@@
 After parsing these two revisions, we will have two ASTs, encoded in two Python dictionaries.
-The change in the code can be computed by analyzing the difference of these two trees, which we call the ASTDiff.
-To derive the ASTDiff we use dictdiff Python library.
-% the wording in the two last sentences is inexact. Let's try to make clear exactly what the steps are, and use the right verbs.
+A tree diff algorithm would specify the differences between two arbitrary trees.
+If the trees are abstracts syntax trees of programs, we simply call it an ASTDiff.
+We use dictdiff to obtain ASTDiffs.
 
@@
 and ?
-provide a detailed explanation of how we select the most important semantic information, which then helped us to effectively obtain meaningful clusters.
-% justification of that last assertion about meaningful clusters?
-e.g.
-we found that this helped us to effectively obtain meaningful clusters; but maybe we just don't know.
+provide a detailed explanation of how we select the most important semantic information, which then helped us to obtain clusters that contained similar datapoints.
 
-% why do we use pydriller?
-% section N explains how we found the bug-related commits
 We used Pydriller, which is a repository miner for Python.
@@
 Syn handles all of Rust.
-Syn's nonterminals are primarily divided into two main groups; Structs and Enums.
-The dimensions of our datapoints have been designed in a way that it adheres to the nonterminals described in these two groups.
+Syn's nonterminals are primarily divided and implemented through two groups: Structs and Enums, which can be found on Syn webpage.
+The union of the item described in these two groups constitutes the list of our non terminals (\verb+NT+).
+The dimensions of our datapoints have been partly designed in a way that they adhere to the elements of \verb+NT+.
+The other part of our dimenstions is comprised of BC-related lexemes (\verb+BCL+), e.g.
+clone, Rc, Box, etc.
+The full list of \verb+BCL+ elements can be found within our artifact.
+We define dimension items as \verb+DI = NT+ $\cup$ \verb+BCL+, as this set also equals the columns of our tables.
+
 We parse two versions of a Rust changed file in a commit; the file before the commit and after it.
@@
 The change that transforms the first tree to the second one yields the fix which was applied in the commit.
-
 To find this difference, we use dictdiff, a Python library to find the diff of two Python dictionaries.
@@
 dictdiff also determines the context within which the modification has occured.
-This context happens to be the root elements (FIND A BETTER TERM FOR ROOT ELEMENTS) in our file, e.g.
+This context happens to be the root elements in our file, e.g.
 function blocks, struct blocks, impl blocks, etc.
@@
 Our objective right now is to find primary elements in our diff that describe the semantics of occurred change.
-% rework the above paragraph
+% TODO rework the above paragraph
 
@@
 \centering
-\includegraphics[width=0.8\textwidth]{linearization.png}
+\includegraphics[width=0.8\textwidth]{extraction.png}
 \caption{\label{fig:extraction}Path Extraction}
@@
 
-% again try to remove redundancy
 The output of dictdiff is a list of diffs, where each diff comprises three parts.
@@
 In this work, we chose to not handle patterns that involve changes in multiple functions or multiple files.
-However, if the change happens in multiple lines within one scope, whether they are continuously one after another or separate from each other, our tool would be able to detect it.
-% simplify that last sentence there
-% provide an example and then transform it into fig:extraction.
+However, our tool would detect both continuous and non-continuous line changes within one scope.
+% TODO provide an example and then transform it into fig:extraction.
 
-% what is the Syn token types list?
 Looking at the tree encoded in the content part of each diff, we realized that each path to a leaf represents one contributing element to the change.
 Through a Depth First Search we can collect all of these paths.
-We store only the non terminals that exist in the Syn token types list.
-Also certain keywords are not considered as non terminals in the syntax tree (e.g.
-clone, Box), but as they have a special meaning the language, we also store them in the path.
+We store only the nodes that exist in the \verb+DI+.
 The outcome of the DFS search is a set of paths.
 
-% reorganize this when you put the example up there.
 Figure~\ref{fig:extraction} shows a simple change and its respective paths.
@@
 A DFS search on this tree would yield three different paths, which we show in red, green, and purple.
-The path represents the sequence of involved non terminals, which represents one contributing factor to the change (along with the context, which we do not record).
-We found it crucial to record the order of all non terminals involved in a path as it gives us freedom to control the degree of order-sensitiveness in our datapoint encoding.
-% can use fewer words in that last sentence too
+The path represents the sequence of involved non terminals, which represents one contributing factor to the change.
+We found it crucial to record the order of all path nodes as it gives us freedom to control the order-sensitiveness in our datapoint encoding.
 
-% OK so if you just have all observed nonterminals then you have a fixed size, but it's too big.
 The next question is how to transform these paths to a fixed size datapoint.
@@
 In such an encoding, for instance, there is no difference between two nested if statements and two if statements beside each other.
-Theoretically, to reach full order-sensitiveness, we would need encode all possible combinations of token types as our dimensions, which yields an astronomical number of dimensions.
+Theoretically, to reach full order-sensitiveness, we would need encode all possible combinations of dimension items as our dimensions, which yields a large number (A loose upper bound would be $n^n$ where $n$ is the number of dimension items).
 Apart from that, not all combinations adhere to Rust syntax.
-A reasonable work around would be to reduce from the full set of nonterminals to a smaller set of categories, and then to set the dimension as all possible combinations of these categories (Similar to JS paper, ref).
+A reasonable work around would be to reduce from the full set of nonterminals to a smaller set of categories (e.g.
+A category for larger entites like class or function definitions, and a category for smaller entities like statements and expressions), and then to set the dimension as all possible combinations of these categories (Similar to JS paper, ref).
 This would reduce the number of dimensions and provide more reasonable and syntactically-correct combinations.
 However, this would still yield a sparse dataset.
-% give examples of categories
-% can we provide orders of magnitude rather than "astronomical"?
 
-% We make a novel contribution here. 
-In our methodology, similar to the first approach, we define the columns of our dataset, as the set of token types we collected from Syn.
+In our methodology, we propose a novel method.
+Similar to the first approach, we define the columns of our dataset, as the set of dimension items we collected from Syn.
 However, to achieve an acceptable level of order-sensitiveness, we carried out two auxiliary steps.
-First, we record the number of occurrences of token types within the paths at each scope.
-The reason behind this decision is because a different order of program elements yields a different number of occurrences of underlying token types.
+First, we record the number of occurrences of dimension items within the paths at each scope.
+The reason behind this decision is because a different order of program elements yields a different number of occurrences of underlying dimension items.
 Second, we semi-automatically design a weighting scheme to prioritize non terminals that we think are more salient for Rust bugs.
@@
 
-% consider: "datapoints accurately describe the most important change." 
-% put the most important change first
 Our main objective is that our datapoints manifest the change with high level of precision.
@@
 % I need to reread this key paragraph. - PL
-A simple heuristic is that, the closer we get to the leaves in the ASTDiff, the more important token types get.
+A simple heuristic is that, the closer we get to the leaves in the ASTDiff, the more important dimension items get.
 Another observation is that the elements closer to the root tend to be repeated in all the paths within one ASTDiff (e.g.
 ItemFn, Local, ExprCall in Figure~\ref{fig:extraction}).
-Exploiting the latter observation, to obtain a basic knowledge about the number of repetitions of token types, we mined the last 20 bug related commits of the last 20 most starred rust projects on GitHub, and ran them through of pipeline.
+Exploiting the latter observation, to obtain a basic knowledge about the number of repetitions of dimension items, we mined the last 20 bug related commits of the last 20 most starred rust projects on GitHub, and ran them through of pipeline.
 We recorded the number of occurrences of each token type in total.
@@
 Hence, it probably does not provide us with the most accurate description of the change.
-In addition to that, as we wanted to make sure that our tool captures the patterns related to borrow checker, we manually raised the weights of the token types related to borrow checker.
+In addition to that, as we wanted to make sure that our tool captures the patterns related to borrow checker, we manually raised the weights of the dimension items related to borrow checker.
 That is why we characterize our weighting scheme as semi-automatically designed.
-
-At last, each datapoint is created by the multiplying the number of occurrences of each token type to its weight and placing the result under its respective column.
-We call the set of final values the essence of the change, as we claim the values of token types manifest their importance in the change.
 
@@
 
+We use Pydriller, a Python library for mining software repositories.
+Pydriller provides APIs to extract commits from a Git repository, and search through different revisions of files.
+Algorithm \ref{alg} shows how we leveraged Pydriller to run the target repositories through our code analysis pipeline, and populate two databases; one for general patterns and one for borrow-checker related patterns.
+The borrow-checker related patterns are the changes that include certain keywords (e.g.
+clone, Box, etc.
+Full list can be found in our artifact).
+For each repository, we search through all the commits that include bug fixing related keywords within their commit messages.
+These keyword for the scope of our project are `bug', `error', `fault', `defect', and `fix'.
+
+Next, we collect a pair of revision per each Rust file.
+The pair includes the state of the file before commit and after commit ($f_a, f_b$).
+After parsing each revision and computing the ASTDiff, we can obtain the fixed sized datapoint $DP$.
+If the datapoint contains borrow-checker related keywords, we would put that inside the BC-related code changes table ($D_b$), otherwise we would put it in the general code changes table ($D_g$).
+In both cases, we augment the datapoint with the commit hash, filename, and the scope in which the change has happened.
+In case of BC-related code changes, we also store detected BC-related keyword.
+Now our tables are ready for clustering and categorization.
+
 \begin{algorithm}
-\caption{An algorithm with caption}\label{alg:cap}
+\caption{An algorithm with caption}\label{alg}
 \hspace*{2mm} \textbf{Input:} $R$ (target repositories)  \\
@@
 
-%% Describing algorithm
-
 \subsection{Clustering Data}
 
-%% Why DBSCAN
+We use DBSCAN clustering algorithm for two main reasons.
+Firstly, it is a density-based clustering method, which means that it detects arbitrarily shaped clusters as opposed to centroid-based methods, like k-means.
+Secondly, also in contrast to k-means, it does not require the number of clusters in advance as an input to the algorithm.
 
-%% Choosing Parameters
+There are two parameters that need tuning in DBSCAN.
+First parameter is $\epsilon$ which indicates a radius with which you can decide whether a point is a core point, a border point, or an outlier.
+Second parameter is the minimum number of point you would like each cluster to have.
+We denote his parameter with $Z$.
+We ran DBSCAN using different combinations of these two parameters.
+In Figure \ref{clustering}, we show nine different experiments with their respective parameter values.
+Within each subplot, the number of clustered points, noise points, and the number of clusters have been specified both for $D_g$ and $D_b$.
 
-%% Figure
+\begin{figure}[h]
+\centering
+\includegraphics[width=1\textwidth]{clusters.png}
+\label{clustering}
+\caption{Clustering Experiments}
+\end{figure}
 
-%% Manual Analysis for finding best parameter
+As a general rule, a lower value for $Z$ would allow more group of close points to be considered as clusters, resulting in a higher number of clusters.
+A large value for $\epsilon$ would widen the search radius, allowing more points to fall in the clusters which results in a reduction of the number of noise points.
+That possibly can put points that do not completely manifest same patterns inside a cluster, resulting in a ineffective clustering.
+Alternatively, a small value for $\epsilon$ would tighten the ring of search, possibly separating clusters which essentially manifest same code changes.
+Also if $\epsilon$ is really small, the clustering algorithm might not detect some clusters at all, resulting in a reduction of the number of clusters.
 
-%% Final Clusters
+\subsubsection{Manual Analysis for parameter tuning}
+
+That is why, we carried out a manual analysis of clusters.
+In the manual analysis we randomly pick 50 clusters and from each cluster we randomly choose 10 datapoints.
+The analyzer then looks at the code and specify the clusters that contained more than five datapoints showing similar patterns.
+Through the manual analysis we specified that $\epsilon=0.0001, Z=5$ and $\epsilon=0.001, Z=5$ acted as the best parameters for $D_g$ and $D_b$, resulting in 596 and 61 clusters, respectively.
+As the final step, we exclude the clusters that contained datapoints from less than 3 projects.
+That is because, we wanted our clusters to contain cross-project patterns as much as possible.
+
+\subsubsection{Manual Analysis for cluster selection}
+% 3 types, refactoring dropped
+
+Similar to ..., following Cotroneo et al. definitions, we divide the clusters in three different groups:
+
+\begin{itemize}
+    \item bug-fix: "The changed code actually fixes the behavior of software, which can represent fixes for a bug type".
+We consider the changes that improve program performance as instance of this group, as they fix the behavior of software.
+
+    \item fix-induced: "The changed code is a group of bug-fixing code changes but not represents an actual bug fix, e.g., adding new input parameters to a method, the method signature and method call must be changed corresponding".
+
+    \item refactoring: "the changed code does not modify the software behavior, e.g., better readability or encapsulation".
+\end{itemize}
+
+In this work, we were only interested in the first two group and disregarded the the clusters that manifest a refactoring change.
+Also, we prefer reporting clusters that manifest concrete patterns.
+An abstract pattern as opposed to a concrete pattern can account for many changes in the programs.
+For instance, the pattern `Adding a new statement to a function's body' is deemed more abstract than `Changing a clone of a variable to a borrowing of it in a function argument'.
+
+In our manual analysis, we randomly picked 50 datapoints of each cluster.
+If the cluster had less than 50 datapoints, we would analyze all of the datapoints in the cluster.
+Next, we read the code and also the bug report related to all the datapoints (if existed).
+In case that a similar pattern was detected over the instances of a cluster, we would write a natural language description of that cluster.
+After providing all the descriptions, we linked similar clusters together as possible candidates for merging.
+At last, after accomplishing re-examination for merge possibilities, we selected 19 clusters that manifested concrete changes.
+We discuss about our findings in Section \ref{}.

### 2022-09-20 methodology.tex
token type to dimension item, maybe there exists a better word choice, one can always use cmd+R

@@
 Exploiting the latter observation, to obtain a basic knowledge about the number of repetitions of dimension items, we mined the last 20 bug related commits of the last 20 most starred rust projects on GitHub, and ran them through of pipeline.
-We recorded the number of occurrences of each token type in total.
+We recorded the number of occurrences of each dimension item in total.
 Then we gave the value of the inverse of the number of occurrences to each token.
-The intuition is that the higher the number of occurrences of a token type, the closer it is usually seen to the root.
+The intuition is that the higher the number of occurrences of a dimension item, the closer it is usually seen to the root.
 Hence, it probably does not provide us with the most accurate description of the change.

### 2022-09-20 methodology.tex
fixed fig labels

@@
 
-% TODO: tokens are not non terminals
-% what do you mean about structs and enums?
-% try to remove redundant content with respect to what is above
 As mentioned above, we use Syn for parsing Rust.
@@
 \includegraphics[width=1\textwidth]{repos.png}
-\label{f3}
-\caption{Target Repositories}
+\caption{\label{table:repos} Target Repositories}
 \end{figure}
@@
 We target the top 18 most starred Rust projects on GitHub, at the time of data collection.
-The details of these projects are outlined in Table \ref{f3}.
+The details of these projects are outlined in Table \ref{table:repos}.
 We computed the average code churn of the files within the projects.
@@
 \begin{algorithm}
-\caption{An algorithm with caption}\label{alg}
+\caption{\label{alg} Mining Algorithm}
 \hspace*{2mm} \textbf{Input:} $R$ (target repositories)  \\
@@
 We ran DBSCAN using different combinations of these two parameters.
-In Figure \ref{clustering}, we show nine different experiments with their respective parameter values.
+In Figure \ref{fig:clustering}, we show nine different experiments with their respective parameter values.
 Within each subplot, the number of clustered points, noise points, and the number of clusters have been specified both for $D_g$ and $D_b$.
@@
 \includegraphics[width=1\textwidth]{clusters.png}
-\label{clustering}
-\caption{Clustering Experiments}
+\caption{\label{fig:clustering} Clustering Experiments}
 \end{figure}

### 2022-09-20 main.tex
fix compilation errors
reorganized files

@@
-\documentclass{article}
-\usepackage{listings, listings-rust}
+%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% This is a general template file for the LaTeX package SVJour3
+% for Springer journals.          Springer Heidelberg 2010/09/16
+%
+% Copy it to a new file with a new name and use it as the basis
+% for your article. Delete % signs as needed.
+%
+% This template includes a few options for different layouts and
+% content for various journals. Please consult a previous issue of
+% your journal as needed.
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% First comes an example EPS file -- just ignore it and
+% proceed on the \documentclass line
+% your LaTeX will extract the file if required
+\begin{filecontents*}{example.eps}
+%!PS-Adobe-3.0 EPSF-3.0
+%%BoundingBox: 19 19 221 221
+%%CreationDate: Mon Sep 29 1997
+%%Creator: programmed by hand (JK)
+%%EndComments
+gsave
+newpath
+  20 20 moveto
+  20 220 lineto
+  220 220 lineto
+  220 20 lineto
+closepath
+2 setlinewidth
+gsave
+  .4 setgray fill
+grestore
+stroke
+grestore
+\end{filecontents*}
+%
+\RequirePackage{fix-cm}
+%
+%\documentclass{svjour3}                     % onecolumn (standard format)
+%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
+\documentclass[smallextended]{build/svjour3}       % onecolumn (second format)
+%\documentclass[twocolumn]{svjour3}          % twocolumn
+%
+\smartqed  % flush right qed marks, e.g. at end of proof
+%
+\usepackage{graphicx}
+%
+% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
+%
+% insert here the call for the packages your document requires
+%\usepackage{latexsym}
+% etc.
+%
+% please place your own definitions here and don't use \def but
+% \newcommand{}{}
+%
+% Insert the name of "your journal" with
+% \journalname{myjournal}
+%
+
+\usepackage{listings, build/listings-rust}
 \lstset{escapeinside={<@}{@>}}
-\usepackage{minted}
+\usepackage{hyperref}
 \usepackage{tabularx}
@@
 
-\title{Rust Bug Fix Patterns}
-\author{Moh Robati}
+\begin{document}
 
-\begin{document}
+\title{A Study on Common Bug Fix Patterns in Rust%\thanks{Grants or other notes
+%about the article that should go on the front page should be
+%placed here. General acknowledgments should be placed at the end of the article.}
+}
+\subtitle{}
+
+%\titlerunning{Short form of title}        % if too long for running head
+
+\author{Mohammad Robati Shirzad         \and
+        Patrick Lam %etc.
+}
+
+%\authorrunning{Short form of author list} % if too long for running head
+
+\institute{F. Author \at
+              first address \\
+              Tel.: +123-45-678910\\
+              Fax: +123-45-678910\\
+              \email{fauthor@example.com}           %  \\
+%             \emph{Present address:} of F. Author  %  if needed
+           \and
+           S. Author \at
+              second address
+}
+
+\date{Received: date / Accepted: date}
+% The correct dates will be entered by the editor
+
 \maketitle
 
+\begin{abstract}
+Insert your abstract here. Include keywords, PACS and mathematical
+subject classification numbers as needed.
+\keywords{First keyword \and Second keyword \and More}
+% \PACS{PACS code1 \and PACS code2 \and more}
+\subclass{MSC code1 \and MSC code2 \and more}
+\end{abstract}
+
 \section{Introduction}
-\subfile{1_introduction}
+\subfile{sections/introduction}
 
 \section{Methodology}
-\subfile{2_methodology}
+\subfile{sections/methodology}
 
 \section{Results}
-\subfile{3_results}
+\subfile{sections/results}
 
 \section{Discussion}
-\subfile{4_discussion}
+\subfile{sections/discussion}
 
 \section{Conclusion}
-\subfile{5_conclusion}
+\subfile{sections/conclusion}
+
+%\begin{acknowledgements}
+%If you'd like to thank anyone, place your comments here
+%and remove the percent signs.
+%\end{acknowledgements}
+
+% Authors must disclose all relationships or interests that 
+% could have direct or potential influence or impart bias on 
+% the work: 
+%
+% \section*{Conflict of interest}
+%
+% The authors declare that they have no conflict of interest.
+
+% BibTeX users please use one of
+%\bibliographystyle{spbasic}      % basic style, author-year citations
+%\bibliographystyle{spmpsci}      % mathematics and physical sciences
+%\bibliographystyle{spphys}       % APS-like style for physics
+%\bibliography{}   % name your BibTeX data base
+
+% Non-BibTeX users please use
+\begin{thebibliography}{}
+%
+% and use \bibitem to create references. Consult the Instructions
+% for authors for reference list style.
+%
+\bibitem{RefJ}
+% Format for Journal Reference
+Author, Article title, Journal, Volume, page numbers (year)
+% Format for books
+\bibitem{RefB}
+Author, Book title, page numbers. Publisher, place (year)
+% etc
+\end{thebibliography}
 
 \end{document}
+% end of file template.tex

### 2022-09-20 sections/results.tex
reorganized files

--- 2022-09-16 results.tex
@@
 \subsection{Evaluating the data modelling approach (RQ1)}
+
+[explain why RQ1]
+
+\textbf{RQ1. Does our data modelling approach capture the most important aspect of program changes?}
 
@@
 
-To give an example, in starship/starship after one of the commits \footnote{\url{https://github.com/starship/starship/commit/56d475578ea508631275772127f49a6949fea6b0}{git_branches.rs}}, the developer borrows the value of $repo\_dir$ to keep the ownership of the object in the current scope.
+To give an example, in starship/starship after one of the commits \footnote{\url{https://github.com/starship/starship/commit/56d475578ea508631275772127f49a6949fea6b0}{git\_branches.rs}}, the developer borrows the value of $repo\_dir$ to keep the ownership of the object in the current scope.
 Now this object can then be used in another function invocation ($remove\_dir\_all(repo\_dir)$) which fixes a bug.
@@
 \end{lstlisting}
-\noindent \textbf{Description:} In struct field definition or the types of the function formal parameters, type $\&Vec\<T\>$ changes to $\&[T]$.
+\noindent \textbf{Description:} In struct field definition or the types of the function formal parameters, type \verb+&Vec<T>+ changes to \verb+&[T]+.
 Clippy can detect the opportunity to apply this pattern.
-The reason is reducing type complication, as slice types in the form of $\&[T]$ or $\&str$ are usually sufficient for most use cases.
+The reason is reducing type complication, as slice types in the form of \verb+&[T]+ or \verb+&str+ are usually sufficient for most use cases.
 
@@
                                  & \multirow{1}{*}{Lifetime}      
-                                                                 & BC.lifetime.static  & Dropping static lifetime    &                        
-                                 \hline
+                                                                 & BC.lifetime.static  & Dropping static lifetime
 

### 2022-09-20 run.sh

@@ -1,0 +1,7 @@
+pdflatex main.tex;
+mv *.eps ./build;
+mv *.aux ./build;
+mv *.fdb_latexmk ./build;
+mv *.fls ./build;
+mv *.log ./build;
+mv *.out ./build;

### 2022-09-20 sections/methodology.tex
reorganized files
warnings fixed

--- 2022-09-20 methodology.tex
@@
 We are interested in finding the most important change within an ASTDiff and encoding that change in our datapoints.
-Sections ?
-and ?
-provide a detailed explanation of how we select the most important semantic information, which then helped us to obtain clusters that contained similar datapoints.
+Sections \ref{sec:path_extraction} and \ref{sec:weighting_scheme} provide a detailed explanation of how we select the most important semantic information, which then helped us to obtain clusters that contained similar datapoints.
 
@@
 Then, we applied the DBSCAN clustering algorithm on the obtained datapoints.
-In Section ?
-we discuss why we chose DBSCAN for clustering and how we tuned it to improve clusters quality.
+In Section \ref{sec:clustering_data} we discuss why we chose DBSCAN for clustering and how we tuned it to improve clusters quality.
 
-\subsection{Data modelling}
+\subsection{\label{sec:data_modelling}Data modelling}
 
-\subsubsection{Parsing Programs}
+\subsubsection{\label{sec:parsing_programs}Parsing Programs}
 
@@
 
-\subsubsection{Path Extraction}
+\subsubsection{\label{sec:path_extraction}Path Extraction}
 
@@
 \centering
-\includegraphics[width=0.8\textwidth]{extraction.png}
+\includegraphics[width=0.8\textwidth]{figs/extraction.png}
 \caption{\label{fig:extraction}Path Extraction}
@@
 
-\subsubsection{Weighting Scheme}
+\subsubsection{\label{sec:weighting_scheme}Weighting Scheme}
 
@@
 
-\subsection{Mining Repositories}
+\subsection{\label{sec:mining_repositories}Mining Repositories}
 
@@
 \centering
-\includegraphics[width=1\textwidth]{repos.png}
+\includegraphics[width=1\textwidth]{figs/repos.png}
 \caption{\label{table:repos} Target Repositories}
@@
 
-\subsection{Clustering Data}
+\subsection{\label{sec:clustering_data}Clustering Data}
 
@@
 \centering
-\includegraphics[width=1\textwidth]{clusters.png}
+\includegraphics[width=1\textwidth]{figs/clusters.png}
 \caption{\label{fig:clustering} Clustering Experiments}
@@
 
-\subsubsection{Manual Analysis for parameter tuning}
+\subsubsection{\label{sec:manual_analysis_parameter_tuning}Manual Analysis for parameter tuning}
 
@@
 
-\subsubsection{Manual Analysis for cluster selection}
+\subsubsection{\label{sec:manual_analysis_cluster_selection}Manual Analysis for cluster selection}
 % 3 types, refactoring dropped
@@
 At last, after accomplishing re-examination for merge possibilities, we selected 19 clusters that manifested concrete changes.
-We discuss about our findings in Section \ref{}.
+We discuss about our findings in Section \ref{sec:common_patterns} and \ref{sec:bc_patterns}.

### 2022-09-20 sections/results.tex
warnings fixed

@@
 
-\subsection{Common Bug Fix Patterns (RQ2)}
+\subsection{\label{sec:common_patterns}Common Bug Fix Patterns (RQ2)}
 
@@
 
-\noindent \textit{ \textbf{G.attr.struct.} Modifying the attributes of structs}
+\noindent \textit{\label{sec:G.attr.struct}\textbf{G.attr.struct} Modifying the attributes of structs}
 
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
+\begin{lstlisting}[language=Rust, style=colouredRust]
 // G.attr.struct.add: Adding attributes to a struct
@@
 \noindent \textbf{Description:} In this pattern, the attribute set of a struct changes.
-This modification can happen in three different ways; (G1.1) Adding attributes to the struct, (G1.2) Removing attributes from the struct, (G1.3) Changing the content of an existing attribute.
+This modification can happen in three different ways; (G.attr.struct.add) Adding attributes to the struct, (G.attr.struct.drop) Removing attributes from the struct, (G.attr.struct.change) Changing the content of an existing attribute.
 The attributes are Outer Attributes as we can find struct related non-terminals within the ASTDiff of the instances of this pattern.
@@
 
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
+\begin{lstlisting}[language=Rust, style=colouredRust]
 // G.attr.struct.field.add: Adding attributes to struct fields
@@
 \noindent \textbf{Description:} In this pattern, the attribute set of a struct field changes.
-As with G1, this modification also happens in three different ways.
+As with \textit{\textbf{G.attr.struct}}, this modification also happens in three different ways.
 Using attributes for struct fields, apart from being able to store different meta-information for different struct fields, for some use cases, the programmers can control the action they want to apply on different fields.
@@
 
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
+\begin{lstlisting}[language=Rust, style=colouredRust]
 // G.struct.field.add: Adding a new field to a struct
@@
 
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
+\begin{lstlisting}[language=Rust, style=colouredRust]
 
@@
 
-\noindent \textit{ \textbf{G.type.field} Changing a struct field type}
+\noindent \textit{\label{sec:G.type.field}\textbf{G.type.field} Changing a struct field type}
 
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
+\begin{lstlisting}[language=Rust, style=colouredRust]
 pub struct Manifest {
@@
 
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
+\begin{lstlisting}[language=Rust, style=colouredRust]
 #[derive(Debug, Clone, PartialEq)]
@@
 
-\noindent \textbf{Description:} This a change of the generic (parametric) type of a struct field. Similar to G7, this change can also happen for the mentioned reasons. \\
+\noindent \textbf{Description:} This a change of the generic (parametric) type of a struct field. Similar to \textit{\textbf{G.type.field}}, this change can also happen for the mentioned reasons. \\
 
@@
 
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
+\begin{lstlisting}[language=Rust, style=colouredRust]
 #[derive(Debug, Clone, PartialEq, Eq)]
@@
 
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
+\begin{lstlisting}[language=Rust, style=colouredRust]
 #[allow(dead_code)]
@@
 
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
+\begin{lstlisting}[language=Rust, style=colouredRust]
 -- pub struct EventLoop<T: tty::EventedReadWrite> {
@@
 
-\subsection{BC-Related Bug Fix Patterns (RQ3)}
+\subsection{\label{sec:bc_patterns}BC-Related Bug Fix Patterns (RQ3)}
 
@@
 
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
+\begin{lstlisting}[language=Rust, style=colouredRust]
 -- let field_id = schema.get_or_create(attribute.clone())?;
@@
 
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
+\begin{lstlisting}[language=Rust, style=colouredRust]
 start_plugin_process(
@@
 
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
+\begin{lstlisting}[language=Rust, style=colouredRust]
 let repo = replace(&mut *contents, Processing).inner_repo();
@@
 
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
+\begin{lstlisting}[language=Rust, style=colouredRust]
 enums.iter().for_each(|e| {
@@
 
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
+\begin{lstlisting}[language=Rust, style=colouredRust]
 for p in wixobjs {
@@
 
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
+\begin{lstlisting}[language=Rust, style=colouredRust]
 parser::Parser::new(args)
@@
 
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
+\begin{lstlisting}[language=Rust, style=colouredRust]
 -- let mut tx = tx.clone();
@@
 
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
+\begin{lstlisting}[language=Rust, style=colouredRust]
 pub fn expand_delimited_square(
@@
 
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
+\begin{lstlisting}[language=Rust, style=colouredRust]
 -- const QUEUE_SIZES: &'static [u16] = &[QUEUE_SIZE];

### 2022-09-20 run.sh

@@ -1,7 +1,9 @@
 pdflatex main.tex;
+sleep 2;
 mv *.eps ./build;
 mv *.aux ./build;
 mv *.fdb_latexmk ./build;
 mv *.fls ./build;
 mv *.log ./build;
 mv *.out ./build;
+mv *.synctex.gz ./build;

### 2022-09-21 sections/methodology.tex
edits

@@
+% Let's start this by talking right away about changes rather than just ASTs
 The general idea for automatically finding the pervasive code patterns within a list of programs is to parse the programs and encode the important parts of the derived AST in fixed-size datapoints.
@@
 To make this work, we need a way to get an AST as a Python dictionary.
-Syn \footnote{\url{https://crates.io/crates/syn}} is a Rust crate which is designed to be employed in Rust procedural macros.
+Syn\footnote{\url{https://crates.io/crates/syn}} is a Rust crate which aims to serve Rust procedural macros.
 However, it includes a parser which is suitable for our purposes; we simply had to write a preprocessor to transform Syn AST output into Python dictionaries.
@@
 After parsing these two revisions, we will have two ASTs, encoded in two Python dictionaries.
-A tree diff algorithm would specify the differences between two arbitrary trees.
-If the trees are abstracts syntax trees of programs, we simply call it an ASTDiff.
+Tree diff algorithms can compute the differences between two arbitrary trees.
+If the trees are programs' abstract syntax trees, we call their diff an ASTDiff.
 We use dictdiff to obtain ASTDiffs.
 
-An ASTDiff may include an arbitrary number of semantic changes.
-We are interested in finding the most important change within an ASTDiff and encoding that change in our datapoints.
-Sections \ref{sec:path_extraction} and \ref{sec:weighting_scheme} provide a detailed explanation of how we select the most important semantic information, which then helped us to obtain clusters that contained similar datapoints.
+An ASTDiff may include an arbitrary number of semantic changes, although a best practice is to include only one semantic change in a commit.
+Because that best practice is not universally followed, we are interested in finding the most important change within an ASTDiff, and encoding that change in our datapoints.
+Sections~\ref{sec:path_extraction}~and~\ref{sec:weighting_scheme} provide a detailed explanation of how we select the most important semantic information, which then allowed us to obtain clusters that contained similar datapoints.
 
-We used Pydriller, which is a repository miner for Python.
-Our target repositories were the top 18 most starred open source Rust projects on GitHub.
-We mined all the bug related commits and ran them through our pipeline.
-Then, we applied the DBSCAN clustering algorithm on the obtained datapoints.
-In Section \ref{sec:clustering_data} we discuss why we chose DBSCAN for clustering and how we tuned it to improve clusters quality.
+Our target repositories were the top 18 most-starred open source Rust projects on GitHub.
+Using Pydriller, a repository miner for Python, we mined all the bug related commits and ran them through our pipeline.
+Then, we applied the DBSCAN clustering algorithm on the resulting datapoints.
+In Section~\ref{sec:clustering_data} we justify our choice of DBSCAN for clustering, and how we tuned it to improve cluster quality.
 
@@
 
+% TODO: BCL elements goes in an appendix.
 As mentioned above, we use Syn for parsing Rust.
 Syn handles all of Rust.
-Syn's nonterminals are primarily divided and implemented through two groups: Structs and Enums, which can be found on Syn webpage.
-The union of the item described in these two groups constitutes the list of our non terminals (\verb+NT+).
-The dimensions of our datapoints have been partly designed in a way that they adhere to the elements of \verb+NT+.
-The other part of our dimenstions is comprised of BC-related lexemes (\verb+BCL+), e.g.
-clone, Rc, Box, etc.
+Each of our datapoints specifies whether any Syn non terminal (\texttt{NT}) is present or absent, i.e.
+containing one dimension per non terminal.
+However, we are also interested in change patterns that involve the borrow checker (BC), so we also add a dimension reflecting the presence of a set of BC-related lexemes (BCL) that we identified, e.g.
+\texttt{clone}, \texttt{Rc}, \texttt{Box}.
 The full list of \verb+BCL+ elements can be found within our artifact.
-We define dimension items as \verb+DI = NT+ $\cup$ \verb+BCL+, as this set also equals the columns of our tables.
+% I think we don't need this sentence here. Let's see what we do later.
+%We define dimension items as \verb+DI = NT+ $\cup$ \verb+BCL+, as this set also equals the columns of our tables.
 
-We parse two versions of a Rust changed file in a commit; the file before the commit and after it.
-We will obtain two trees created by Syn.
-The change that transforms the first tree to the second one yields the fix which was applied in the commit.
+We parse two versions of a changed Rust file in a commit: the file before the commit and after it.
+This yields two Syn trees.
+Using the PLY tool (a Python implementation of lex and yacc), we wrote a simple transformer from the serialized Rust AST to Python dictionaries.
+The change that transforms the first tree to the second one is the fix that was applied in the commit.
 To find this difference, we use dictdiff, a Python library to find the diff of two Python dictionaries.
-Using the PLY tool (a Python implementation of lex and yacc), we wrote a simple transformer from the Rust AST into Python dictionaries.
-After obtaining the syntax trees in Python dictionary format, we can use dictdiff to determine the ASTDiff.
+Its output, in our context, is the ASTDiff.
 
-dictdiff tool provides modifications' types: 
+dictdiff provides modifications' types: 
 \begin{itemize}

### 2022-09-21 sections/methodology.tex
2.1.3

@@
 
-dictdiff provides modifications' types: 
+An ASTDiff, as computed by dictdiff, contains a type:
 \begin{itemize}
@@
 \end{itemize}
-dictdiff also determines the context within which the modification has occured.
+% the description of the context in 2.1.2 is probably enough, you don't need it here, just remove it.
+I think that you just put the above itemize into 2.1.2 and erase the next paragraph, unless you can spot something you need to add to 2.1.2.
+It also encodes the context of the modification.
 This context happens to be the root elements in our file, e.g.
@@
 
+% remove redundancy between the description of the ASTDiff in 2.1.1 and what we have here. Probably it can go all into 2.1.2, i.e. you can just move the last para of 2.1.1 to here and incorporate it.
 The output of dictdiff is a list of diffs, where each diff comprises three parts.
-The first part specifies whether the modification is of type 'add', 'remove', 'change'.
+The first part specifies whether the modification is of type `add', `remove', or `change'.
 The second part identifies the context of the modification, that is, the path from the root of the tree to the subtree in which the modification has occured.
@@
 Looking at the tree encoded in the content part of each diff, we realized that each path to a leaf represents one contributing element to the change.
-Through a Depth First Search we can collect all of these paths.
-We store only the nodes that exist in the \verb+DI+.
-The outcome of the DFS search is a set of paths.
+A Depth First Search allows us to collect all paths, and we return only the paths that we store in our datapoints (i.e.
+touching $\mathtt{NT} \cup \mathtt{BCL}$).
 
-Figure~\ref{fig:extraction} shows a simple change and its respective paths.
+Figure~\ref{fig:extraction} illustrates path extraction on a simple change.
 The change replaces \verb+tail+ with \verb+&tail+ and \verb+head+ with \verb+head.clone()+.
 That is, after the change, the second argument passes a borrow of \verb+tail+ rather than sending the ownership of \verb+tail+ to the callee; and, in the third argument, instead of sending \verb+head+ to the callee, it sends a clone of it.
-A DFS search on this tree would yield three different paths, which we show in red, green, and purple.
-The path represents the sequence of involved non terminals, which represents one contributing factor to the change.
-We found it crucial to record the order of all path nodes as it gives us freedom to control the order-sensitiveness in our datapoint encoding.
+A DFS on this tree yields three different paths, which we show in red, green, and purple.
+A path represents a sequence of involved non terminals.
+We found it crucial to record the order of nodes in the path, as this gives us the freedom to later record the order in our datapoint encoding.
 
-The next question is how to transform these paths to a fixed size datapoint.
+% I think I would move the discussion of why fixed size from the beginning of the section to here; we don't need it at the beginning of the section.
+% I think it would be better to show an example of the two if statements either way
+\paragraph{Encoding the data.} A key question is how to transform these paths to a fixed size datapoint.
 First, we need to define a fixed set of columns.
@@
 In such an encoding, for instance, there is no difference between two nested if statements and two if statements beside each other.
-Theoretically, to reach full order-sensitiveness, we would need encode all possible combinations of dimension items as our dimensions, which yields a large number (A loose upper bound would be $n^n$ where $n$ is the number of dimension items).
-Apart from that, not all combinations adhere to Rust syntax.
-A reasonable work around would be to reduce from the full set of nonterminals to a smaller set of categories (e.g.
-A category for larger entites like class or function definitions, and a category for smaller entities like statements and expressions), and then to set the dimension as all possible combinations of these categories (Similar to JS paper, ref).
-This would reduce the number of dimensions and provide more reasonable and syntactically-correct combinations.
-However, this would still yield a sparse dataset.
+On the other hand, theoretically, for full order-sensitiveness, we would need to encode all possible combinations of items from $\mathtt{NT} \cup \mathtt{BCL}$ as our dimensions, which yields an infeasible $O(n^n)$ dimensions; this could potentially be reduced somewhat, but is still impractical.
+A reasonable work around would be to reduce $n$ from the full set of non terminals to a smaller set of categories (e.g.
+a category for larger entites like class or function definitions, and a category for smaller entities like statements and expressions), and then to record all possible combinations of these categories (similar to JS authors~\cite{foo}).
+That approach reduces the number of dimensions and provides more reasonable and syntactically-correct combinations.
+However, it would still yield a sparse dataset.
+% todo you need to dig up the JS paper ref
 
-In our methodology, we propose a novel method.
-Similar to the first approach, we define the columns of our dataset, as the set of dimension items we collected from Syn.
-However, to achieve an acceptable level of order-sensitiveness, we carried out two auxiliary steps.
-First, we record the number of occurrences of dimension items within the paths at each scope.
+% ok, add a figure and then we'll look at the text again and follow along with the picture
+In our methodology, we propose a novel representation.
+Similar to the first approach, we order-insensitively define the columns of our dataset as the set of items $\mathtt{NT} \cup \mathtt{BCL}$ we collected from Syn.
+However, to simulate order-sensitiveness, we carried out two auxiliary steps.
+First, we record the number of occurrences of items within the paths at each scope.
 The reason behind this decision is because a different order of program elements yields a different number of occurrences of underlying dimension items.
@@
 
-Our main objective is that our datapoints manifest the change with high level of precision.
+Our main requirement is that our datapoints summarize the key changes in ASTDiffs---a datapoint should foreground the most important change in a diff.
 For instance, the purple (right-most) path in Figure~\ref{fig:extraction} could be described as a change in a local variable declaration; a change in a function call; calling a method of one of the function arguments; or calling clone() on one of the function arguments.
-The last description offers the most accurate description.
-The whole idea behind our weighting scheme is to be able to identify the change accurately.
+The last description is the most useful one, and we designed our weighting scheme to encode this value judgment.
 
-% I need to reread this key paragraph. - PL
-A simple heuristic is that, the closer we get to the leaves in the ASTDiff, the more important dimension items get.
-Another observation is that the elements closer to the root tend to be repeated in all the paths within one ASTDiff (e.g.
+We made two observations about changes.
+First, we saw that an item (non terminal or BC-related lexeme) that occurs closer to AST leaves tends to be more semantically important.
+Relatedly, items closer to the root tend to be repeated in all the paths within one ASTDiff (e.g.
 ItemFn, Local, ExprCall in Figure~\ref{fig:extraction}).
-Exploiting the latter observation, to obtain a basic knowledge about the number of repetitions of dimension items, we mined the last 20 bug related commits of the last 20 most starred rust projects on GitHub, and ran them through of pipeline.
-We recorded the number of occurrences of each dimension item in total.
-Then we gave the value of the inverse of the number of occurrences to each token.
-The intuition is that the higher the number of occurrences of a dimension item, the closer it is usually seen to the root.
-Hence, it probably does not provide us with the most accurate description of the change.
-In addition to that, as we wanted to make sure that our tool captures the patterns related to borrow checker, we manually raised the weights of the dimension items related to borrow checker.
-That is why we characterize our weighting scheme as semi-automatically designed.
+We thus want to prioritize items that tend to occur closer to leaves and de-prioritize items that tend to occur closer to roots.
+
+To design our weighting scheme, we collected empirical data about locations of item occurrences in a corpus.
+Specifically, we mined the last 20 bug related commits of the last 20 most-starred Rust projects on GitHub, and ran them through our pipeline.
+We recorded the total number of occurrences of each item.
+Then, if $\# i$ is the number of occurences of $i$, we gave $i$ a weight of $1/\# i$.
+Because items that occur closer to the root also occur more often, per our observation, this gives higher weights to items that occur infrequently, i.e.
+closer to the leaves.
+Items that occur often and have lower weights are, we believe, less important in describing a change.
+
+Furthermore, as we wanted to make sure that our tool captures patterns related to the Rust borrow checker, we manually increased the weights of the \texttt{BCL} items.
+The manual adjustment is why we characterize our weighting scheme as semi-automatically designed.
 

### 2022-09-21 sections/methodology.tex
2.2

@@
 Our target repositories were the top 18 most-starred open source Rust projects on GitHub.
-Using Pydriller, a repository miner for Python, we mined all the bug related commits and ran them through our pipeline.
+We mined (using Pydriller) all the bug related commits and ran them through our pipeline.
 Then, we applied the DBSCAN clustering algorithm on the resulting datapoints.
@@
 
-\begin{figure}[h]
+\begin{table}[h]
 \centering
+\caption{\label{table:repos} Target Repositories}
 \includegraphics[width=1\textwidth]{figs/repos.png}
-\caption{\label{table:repos} Target Repositories}
-\end{figure}
+\end{table}
 
+% should have a reference to how you are computing churn
 Now that we have set up our code analysis pipeline, we can start mining the Rust repositories.
-We target the top 18 most starred Rust projects on GitHub, at the time of data collection.
-The details of these projects are outlined in Table \ref{table:repos}.
-We computed the average code churn of the files within the projects.
-denoland/deno and swc-project/swc are the projects with largest number of LOCs and average code churn.
+We target the top 18 most-starred Rust projects on GitHub, at the time of data collection.
+Table~\ref{table:repos} summarizes our benchmarks.
+Because we are interested in changes, we computed the average code churn of the files within the projects.
+denoland/deno and swc-project/swc are the projects with largest number of LOCs and highest average code churn, respectively.
 Unsurprisingly, we captured a lot of instances from these two projects in our final clusters.
 
-We use Pydriller, a Python library for mining software repositories.
-Pydriller provides APIs to extract commits from a Git repository, and search through different revisions of files.
-Algorithm \ref{alg} shows how we leveraged Pydriller to run the target repositories through our code analysis pipeline, and populate two databases; one for general patterns and one for borrow-checker related patterns.
-The borrow-checker related patterns are the changes that include certain keywords (e.g.
-clone, Box, etc.
-Full list can be found in our artifact).
-For each repository, we search through all the commits that include bug fixing related keywords within their commit messages.
-These keyword for the scope of our project are `bug', `error', `fault', `defect', and `fix'.
+We used the Pydriller library to mine software repositories.
+Pydriller provides APIs to extract commits from a Git repository and to search through different revisions of files.
+Algorithm~\ref{alg} shows how we used Pydriller to run the target repositories through our code analysis pipeline and populate two databases: one for borrow-checker related patterns (which involve items in \texttt{BCL}) and one for general patterns (which don't).
+For each repository, we search all commits that include bug-fixing related keywords within their commit messages.
+Here, we chose bug-fixing related keywords `bug', `error', `fault', `defect', and `fix'.
+% cite earlier work that does the search-for-keywords-to-identify-bugs; my earlier EMSE paper has that citation for instance.
 
-Next, we collect a pair of revision per each Rust file.
-The pair includes the state of the file before commit and after commit ($f_a, f_b$).
-After parsing each revision and computing the ASTDiff, we can obtain the fixed sized datapoint $DP$.
-If the datapoint contains borrow-checker related keywords, we would put that inside the BC-related code changes table ($D_b$), otherwise we would put it in the general code changes table ($D_g$).
-In both cases, we augment the datapoint with the commit hash, filename, and the scope in which the change has happened.
-In case of BC-related code changes, we also store detected BC-related keyword.
-Now our tables are ready for clustering and categorization.
+Next, we collect a pair of revisions for each Rust file.
+The pair includes the state of the file before the commit and after the commit ($f_a, f_b$).
+After parsing each revision and computing the ASTDiff, we can compute the fixed sized datapoint $\mathit{DP}$.
+If the datapoint contains borrow-checker related keywords, we put it in the BC-related database $D_b$; otherwise, we put it in the general code changes database $D_g$.
+In both cases, we augment the datapoint with the commit hash, filename, and the scope in which the change happened.
+For BC-related code changes, we also store the detected BC-related keywords.
+Now our databases are ready for clustering and categorization.
 
@@
                 \For{$e \in \textsc{ASTDiff}(\textsc{Parse}(f_b), \textsc{Parse}(f_a))$}
-                    \State $DP \leftarrow \textsc{GetDataPoint}(e)$
-                    \State $DP \leftarrow c.hash \cup f_a.name \cup e.scope \cup DP $
+                    \State $\textit{DP} \leftarrow \textsc{GetDataPoint}(e)$
+                    \State $\textit{DP} \leftarrow c.\textit{hash} \cup f_a.\textit{name} \cup e.\textit{scope} \cup \textit{DP} $
                     \If{$\textsc{IsBCRelated}(e)$}
-                        \State $DP \leftarrow \textsc{GetBCKeyword}(e) \cup DP $
-                        \State $D_b \leftarrow D_b \cup DP$
+                        \State $\textit{DP} \leftarrow \textsc{GetBCKeyword}(e) \cup \textit{DP} $
+                        \State $D_b \leftarrow D_b \cup \textit{DP}$
                     \Else
-                        \State $D_g \leftarrow D_g \cup DP$
+                        \State $D_g \leftarrow D_g \cup \textit{DP}$
                     \EndIf
@@
 
-We use DBSCAN clustering algorithm for two main reasons.
-Firstly, it is a density-based clustering method, which means that it detects arbitrarily shaped clusters as opposed to centroid-based methods, like k-means.
-Secondly, also in contrast to k-means, it does not require the number of clusters in advance as an input to the algorithm.
+Having collected our datapoints, our next task is to cluster them so that we can categorize bug patterns.
+Like~\cite{JS paper, Python paper}, we use the DBSCAN clustering algorithm for two main reasons.
+Firstly, in contrast to k-means, it does not require the number of clusters in advance as an input to the algorithm.
+Secondly, it is a density-based clustering method, which means that it detects arbitrarily shaped clusters; centroid-based methods like k-means can't detect such clusters.
+We tried k-means and the more powerful fuzzy variant F-means and did not get useful clusters.
 
-There are two parameters that need tuning in DBSCAN.
-First parameter is $\epsilon$ which indicates a radius with which you can decide whether a point is a core point, a border point, or an outlier.
-Second parameter is the minimum number of point you would like each cluster to have.
-We denote his parameter with $Z$.
+DBSCAN takes two tuneable parameters.
+The first parameter $\epsilon$ indicates a radius within which a point is a core point, a border point, or an outlier.
+The second parameter $Z$ is the minimum number of points per cluster.
 We ran DBSCAN using different combinations of these two parameters.
-In Figure \ref{fig:clustering}, we show nine different experiments with their respective parameter values.
-Within each subplot, the number of clustered points, noise points, and the number of clusters have been specified both for $D_g$ and $D_b$.
+In Figure~\ref{fig:clustering}, we show nine different experiments, with their respective parameter values.
+Each subplot specifies, for both $D_g$ and $D_b$, the number of clustered points, noise points, and the number of clusters.
 
+% TODO fix this caption
 \begin{figure}[h]
@@
 
-As a general rule, a lower value for $Z$ would allow more group of close points to be considered as clusters, resulting in a higher number of clusters.
-A large value for $\epsilon$ would widen the search radius, allowing more points to fall in the clusters which results in a reduction of the number of noise points.
-That possibly can put points that do not completely manifest same patterns inside a cluster, resulting in a ineffective clustering.
-Alternatively, a small value for $\epsilon$ would tighten the ring of search, possibly separating clusters which essentially manifest same code changes.
-Also if $\epsilon$ is really small, the clustering algorithm might not detect some clusters at all, resulting in a reduction of the number of clusters.
+As a general rule, lower $Z$ would allow more groups of close points to be considered as clusters, resulting in a higher number of clusters.
+A large $\epsilon$ would widen the search radius, allowing more points to fall in the clusters, which results in a reduction of the number of noise points.
+Such a choice might put points from different patterns inside the same cluster, resulting in a ineffective clustering.
+Alternatively, a small $\epsilon$ would tighten the ring of search, possibly separating clusters which essentially manifest the same code changes.
+Also, if $\epsilon$ is really small, the clustering algorithm might not detect some clusters at all, resulting in a reduction of the number of clusters.
 
@@
 
-That is why, we carried out a manual analysis of clusters.
-In the manual analysis we randomly pick 50 clusters and from each cluster we randomly choose 10 datapoints.
-The analyzer then looks at the code and specify the clusters that contained more than five datapoints showing similar patterns.
-Through the manual analysis we specified that $\epsilon=0.0001, Z=5$ and $\epsilon=0.001, Z=5$ acted as the best parameters for $D_g$ and $D_b$, resulting in 596 and 61 clusters, respectively.
-As the final step, we exclude the clusters that contained datapoints from less than 3 projects.
-That is because, we wanted our clusters to contain cross-project patterns as much as possible.
+That is why we carried out a manual analysis of clusters.
+In the manual analysis, we randomly pick 50 clusters; from each, cluster we randomly choose 10 datapoints.
+The person analyzing the cluster then looks at the code and specifies the clusters that contained more than five datapoints showing similar patterns.
+For the manual analysis, we used $\epsilon=0.0001, Z=5$ and $\epsilon=0.001, Z=5$, which were the best parameters for $D_g$ and $D_b$, resulting in 596 and 61 clusters, respectively.
+As a final step, we exclude clusters that contained datapoints from fewer than 3 projects.
+That is because we wanted our clusters to contain cross-project (e.g.
+potentially generalizable) patterns as much as possible.
 
@@
 \begin{itemize}
-    \item bug-fix: "The changed code actually fixes the behavior of software, which can represent fixes for a bug type".
-We consider the changes that improve program performance as instance of this group, as they fix the behavior of software.
+    \item bug-fix: ``The changed code actually fixes the behavior of software, which can represent fixes for a bug type''.
+We include here changes that improve program performance, as they fix the behaviour of the software.
 
-    \item fix-induced: "The changed code is a group of bug-fixing code changes but not represents an actual bug fix, e.g., adding new input parameters to a method, the method signature and method call must be changed corresponding".
+    \item fix-induced: ``The changed code is a group of bug-fixing code changes but not representing an actual bug fix, e.g., adding new input parameters to a method, the method signature and method call must be changed correspondingly''.
 
-    \item refactoring: "the changed code does not modify the software behavior, e.g., better readability or encapsulation".
+    \item refactoring: ``the changed code does not modify the software behavior, e.g., for better readability or encapsulation''.
 \end{itemize}
 
-In this work, we were only interested in the first two group and disregarded the the clusters that manifest a refactoring change.
+In this work, we were only interested in the first two groups and ignored the clusters for refactoring changes.
 Also, we prefer reporting clusters that manifest concrete patterns.
@@
 For instance, the pattern `Adding a new statement to a function's body' is deemed more abstract than `Changing a clone of a variable to a borrowing of it in a function argument'.
+% show example in terms of code; do you have an actual definition for abstract vs concrete?
 
 In our manual analysis, we randomly picked 50 datapoints of each cluster.
-If the cluster had less than 50 datapoints, we would analyze all of the datapoints in the cluster.
-Next, we read the code and also the bug report related to all the datapoints (if existed).
-In case that a similar pattern was detected over the instances of a cluster, we would write a natural language description of that cluster.
-After providing all the descriptions, we linked similar clusters together as possible candidates for merging.
-At last, after accomplishing re-examination for merge possibilities, we selected 19 clusters that manifested concrete changes.
-We discuss about our findings in Section \ref{sec:common_patterns} and \ref{sec:bc_patterns}.
+If the cluster had fewer than 50 datapoints, we would analyze all of the datapoints in the cluster.
+Next, we read the code and the bug reports related to all the datapoints (if they existed).
+If we judged that the members of a cluster fit into a pattern, we wrote a natural language description of that cluster.
+Considering the descriptions, we linked similar clusters together as possible candidates for merging.
+We considered these candidates and carried out appropriate ones.
+Finally, we selected 19 of the remaining clusters---the ones that manifested concrete changes.
+We discuss about our findings in Sections~\ref{sec:common_patterns} and \ref{sec:bc_patterns}.

### 2022-09-23 sections/methodology.tex
added essence figure, revised methodology, added introduction for results, revised tables for results, we still need to add some numbers

@@
-% Let's start this by talking right away about changes rather than just ASTs
+A code change is a modification in a program that changes the program's abstract syntax tree.
+A change pattern is a series of code changes that serves a specific purpose.
+This purpose can be improving code readability or maintainability (refactoring patterns), or rectifying program behaviour with respect to a specific functional or non-functional requirement (bug-fixing patterns).
+In this work, we are interested in latter group.
+Thus, from now on, by referring to code patterns we mean bug-fixing patterns.
+
 The general idea for automatically finding the pervasive code patterns within a list of programs is to parse the programs and encode the important parts of the derived AST in fixed-size datapoints.
-We need fixed-size datapoints, as it is a requirement for both centroid-based and density-based well-known clustering algorithms.
 Work in this vein makes the assumption that each cluster corresponds to one change pattern.
@@
 
-% TODO Define change
-A change pattern categorizes a class of changes. We thus need to define what a change is. For our purposes, a change is ....
 To compute the contents of a change, we need two code revisions: the revision before the commit, and the revision after it.
@@
 
-% TODO: BCL elements goes in an appendix.
 As mentioned above, we use Syn for parsing Rust.
@@
 \texttt{clone}, \texttt{Rc}, \texttt{Box}.
-The full list of \verb+BCL+ elements can be found within our artifact.
+The full list of \verb+BCL+ elements can be found within Appendix~\ref{sec:appendix}.
 % I think we don't need this sentence here. Let's see what we do later.
@@
 
-An ASTDiff, as computed by dictdiff, contains a type:
+\subsubsection{\label{sec:path_extraction}Path Extraction}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.8\textwidth]{figs/extraction.png}
+\caption{\label{fig:extraction}Extracting paths from an ASTDiff. AST corresponds to the code change in the yellow box}
+\end{figure}
+
+The output of dictdiff is a list of diffs, where each diff comprises three parts. The first part specifies the modification type:
+
 \begin{itemize}
@@
 \end{itemize}
-% the description of the context in 2.1.2 is probably enough, you don't need it here, just remove it.
-I think that you just put the above itemize into 2.1.2 and erase the next paragraph, unless you can spot something you need to add to 2.1.2.
-It also encodes the context of the modification.
-This context happens to be the root elements in our file, e.g.
-function blocks, struct blocks, impl blocks, etc.
-The modified content within a context is still in the form of tree.
-Our objective right now is to find primary elements in our diff that describe the semantics of occurred change.
-% TODO rework the above paragraph
 
-\subsubsection{\label{sec:path_extraction}Path Extraction}
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.8\textwidth]{figs/extraction.png}
-\caption{\label{fig:extraction}Path Extraction}
-\end{figure}
-
-% remove redundancy between the description of the ASTDiff in 2.1.1 and what we have here. Probably it can go all into 2.1.2, i.e. you can just move the last para of 2.1.1 to here and incorporate it.
-The output of dictdiff is a list of diffs, where each diff comprises three parts.
-The first part specifies whether the modification is of type `add', `remove', or `change'.
 The second part identifies the context of the modification, that is, the path from the root of the tree to the subtree in which the modification has occured.
@@
 However, our tool would detect both continuous and non-continuous line changes within one scope.
-% TODO provide an example and then transform it into fig:extraction.
 
@@
 
-% I think I would move the discussion of why fixed size from the beginning of the section to here; we don't need it at the beginning of the section.
-% I think it would be better to show an example of the two if statements either way
 \paragraph{Encoding the data.} A key question is how to transform these paths to a fixed size datapoint.
 First, we need to define a fixed set of columns.
+We need fixed-size datapoints, as it is a requirement for both centroid-based and density-based well-known clustering algorithms.
 The number of columns would depend on the amount of information we want to encode in the datapoints.
@@
 On the other hand, theoretically, for full order-sensitiveness, we would need to encode all possible combinations of items from $\mathtt{NT} \cup \mathtt{BCL}$ as our dimensions, which yields an infeasible $O(n^n)$ dimensions; this could potentially be reduced somewhat, but is still impractical.
+However, in this case we can put a distinction between the aformentioned if statements, as we have a different dimension for nested if statements (\verb+if_if_statement_...+).
 A reasonable work around would be to reduce $n$ from the full set of non terminals to a smaller set of categories (e.g.
@@
 However, it would still yield a sparse dataset.
-% todo you need to dig up the JS paper ref
+% todo you need to dig up the JS paper ref / Not sure, how to cite, I usually create a bib file, but i'm not sure if i'm allowed to do that for this template.
 
-% ok, add a figure and then we'll look at the text again and follow along with the picture
 In our methodology, we propose a novel representation.
@@
 
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=1\textwidth]{figs/essence.png}
+    \caption{\label{fig:essence}Multiplying the number of occurrences of items observed in the tree by their respective weight, resulting in the essence of change (Visualized with a circle pack figure)}
+\end{figure}
+
 Our main requirement is that our datapoints summarize the key changes in ASTDiffs---a datapoint should foreground the most important change in a diff.
@@
 We made two observations about changes.
-First, we saw that an item (non terminal or BC-related lexeme) that occurs closer to AST leaves tends to be more semantically important.
+First, we saw that an item (non terminal or BC-related lexemes) that occurs closer to AST leaves tends to be more semantically important.
 Relatedly, items closer to the root tend to be repeated in all the paths within one ASTDiff (e.g.
@@
 
+Multiplying the number of occurrences of all observed items in a change by their respective weight results in a vector of numbers.
+We call this vector, the essence of change.
+In this vector, the respective value for each item shows its importance in the change.
+Figure~\ref{fig:essence} shows the process of computing the essence of the code change that we saw in Figure~\ref{fig:extraction}.
+We have visualized this essence vector using a circle pack figure.
+The radius of each circle corresponds to the importance of the item inside it.
+
 \subsection{\label{sec:mining_repositories}Mining Repositories}
 
-\begin{table}[h]
-\centering
-\caption{\label{table:repos} Target Repositories}
-\includegraphics[width=1\textwidth]{figs/repos.png}
+\begin{table}
+\caption{\label{table:repos} Target repositories for mining}
+\begin{tabular}{|c|c|c|c|c|}
+    \hline
+    Project Name & Stars(k) & LOC & \# Commits & Avg. Code Churn \\ \hline
+    denoland/deno & 84.7 & 562468 & 7730 & 49.22 \\ \hline
+    tauri-apps/tauri & 48.5 & 130726 & 2929 & 31.67 \\ \hline
+    alacritty/alacritty & 41.2 & 45803 & 2017 & 27.53 \\ \hline
+    sharkdp/bat & 36.4 & 93258 & 2392 & 59.13 \\ \hline
+    starship/starship & 28 & 176630 & 2290 & 25.1 \\ \hline
+    rust-lang/rustlings & 28.7 & 9852 & 1391 & 8.92 \\ \hline
+    meilisearch/meilisearch & 28.6 & 66872 & 3849 & 279.7 \\ \hline
+    sharkdp/fd & 24.2 & 10383 & 1025 & 12.2 \\ \hline
+    swc-project/swc & 23.5 & 13512766 & 5205 & 44.34 \\ \hline
+    yewstack/yew & 23.2 & 112972 & 2239 & 21.79 \\ \hline
+    xi-editor/xi-editor & 19.5 & 58783 & 2104 & 28.67 \\ \hline
+    AppFlowy-IO/AppFlowy & 25.9 & 182830 & 2600 & 29.49 \\ \hline
+    firecracker-microvm/firecracker & 18.9 & 156759 & 3536 & 58.26 \\ \hline
+    nushell/nushell & 20.3 & 181710 & 5922 & 22.38 \\ \hline
+    ogham/exa & 18.9 & 17114 & 1558 & 11.99 \\ \hline
+    SergioBenitez/Rocket & 18.2 & 90055 & 2127 & 19.75 \\ \hline
+    rustdesk/rustdesk & 27.7 & 85322 & 1664 & 72.81 \\ \hline
+    tokio-rs/tokio & 17.2 & 126089 & 3006 & 28.87 \\ \hline
+\end{tabular}
 \end{table}
 
-% should have a reference to how you are computing churn
 Now that we have set up our code analysis pipeline, we can start mining the Rust repositories.
@@
 Table~\ref{table:repos} summarizes our benchmarks.
-Because we are interested in changes, we computed the average code churn of the files within the projects.
+Because we are interested in changes, we computed the average code churn of the files within the projects (The code churn is computed using process metrics provided by Pydriller).
 denoland/deno and swc-project/swc are the projects with largest number of LOCs and highest average code churn, respectively.
@@
 Here, we chose bug-fixing related keywords `bug', `error', `fault', `defect', and `fix'.
-% cite earlier work that does the search-for-keywords-to-identify-bugs; my earlier EMSE paper has that citation for instance.
+% cite earlier work that does the search-for-keywords-to-identify-bugs; my earlier EMSE paper has that citation for instance. -- same issue as above, not sure how to cite
 
@@
 \includegraphics[width=1\textwidth]{figs/clusters.png}
-\caption{\label{fig:clustering} Clustering Experiments}
+\caption{\label{fig:clustering} Clustering results with nine different combinations of parameters ($\epsilon=[10^{-2}, 10^{-3}, 10^{-4}], Z=[5,7,9])$}
 \end{figure}
@@
 An abstract pattern as opposed to a concrete pattern can account for many changes in the programs.
-For instance, the pattern `Adding a new statement to a function's body' is deemed more abstract than `Changing a clone of a variable to a borrowing of it in a function argument'.
-% show example in terms of code; do you have an actual definition for abstract vs concrete?
+For instance, the pattern `Changing a clone of a variable to a borrowing of it in a function argument' is deemed more concrete than `Adding a new statement to a function's body':
+
+\begin{lstlisting}[language=Rust, style=colouredRust]
+// First Pattern
+-- foo1(arg.clone);
+++ foo1(&arg);
+
+// Second Pattern
+fn foo2() {
+    stmt1;
+++  stmt2;
+}
+
+\end{lstlisting}
 

### 2022-09-23 sections/results.tex
added essence figure, revised methodology, added introduction for results, revised tables for results, we still need to add some numbers

@@
+In this section, we first evaluate the effectiveness of our data modelling approach (RQ1).
+That is, how successfull is our data modelling approach in capturing the most important aspects of a program change.
+This evaluation is required, as the credibility and generality of our clusters are heavily based on this step in our pipeline.
+Then we move to introducing the obtained clusters in two sections: Common patterns (RQ2), and BC-Related patterns (RQ3).
+
 \subsection{Evaluating the data modelling approach (RQ1)}
-
-[explain why RQ1]
 
@@
 \subsection{\label{sec:common_patterns}Common Bug Fix Patterns (RQ2)}
+
+\begin{table}[]
+
+\begin{tabular}{|l|l|l|l|}
+\hline
+
+\multirow{10}{*}{General}        
+& \multirow{2}{*}{Attributes} & G.attr.struct & -                                  \\
+& & G.attr.struct.field & -                     \\\cline{2-3}
+& \multirow{2}{*}{Struct} & G.struct.field  & -\\
+& & G.struct.field.pub & -    
+\\\cline{2-3}
+& Option & G.option & -                    \\\cline{2-3}
+& \multirow{4}{*}{Types} & G.type.field  & - \\
+& & G.type.generic & -\\ & & G.type.enum.variant  & - \\
+& & G.type.tuple & -                                         \\\cline{2-3}
+& \multirow{1}{*}{Traits} & G.trait.bound & - \\ 
+\hline
+\end{tabular}
+\caption{\label{table:general}Bug Fix Pattern Categorization (General Patterns)}
+\end{table}
 
@@
 
+\begin{table}[]
+\begin{tabular}{|l|l|l|l|}
+\hline
+\multirow{9}{*}{Borrow Checker} &
+\multirow{2}{*}{Clone}                                         & BC.clone.ref & - \\
+& & BC.clone.drop & -   
+\\\cline{2-3}
+& \multirow{3}{*}{Ref and Deref} & BC.ref.add & - \\
+& & BC.ref.drop & - \\
+& & BC.deref.to\_string  & -                                            \\\cline{2-3}
+& \multirow{2}{*}{Mut} & BC.mut.add  & - \\
+& & BC.mut.drop  & -                                \\\cline{2-3}
+& Vector & BC.vec.slice  & -                                    \\\cline{2-3}
+& \multirow{1}{*}{Lifetime}      
+& BC.lifetime.static  & -
+\\
+\hline
+\end{tabular}
+\caption{\label{table:bc}Bug Fix Pattern Categorization (BC-Related Patterns)}
+\end{table}
+
 \subsubsection{Clone}
@@
 If the presence of static is not required, it is better that we omit it as keeping it might create complicate types in our program.
-
-\begin{table}[]
-
-\begin{tabular}{|l|l|l|l|}
-\hline
-
-\multirow{10}{*}{General}        
-& \multirow{2}{*}{Attributes}    & G.attr.struct  & Modifying the attributes of structs                                        \\
-                                 &                                & G.attr.struct.field  & Modifying the attributes of struct fields                                  \\\cline{2-3}
-                                 & \multirow{2}{*}{Struct}        & G.struct.field  & Adding/Removing a struct field                                      \\
-                                 &                                & G.struct.field.pub  & Modifying the access modifiers of struct fields     
-                                  \\\cline{2-3}
-                                 & Option                         & G.option  & Changing field type T to T$\langle$Option$\rangle$                     \\\cline{2-3}
-                                 & \multirow{4}{*}{Types}         & G.type.field  & Changing a struct field type                                        \\
-                                 &                                & G.type.generic  & Changing a generic type                                             \\
-                                 &                                & G.type.enum.variant  & Change in enum variant value type                                   \\
-                                 &                                & G.type.tuple & Changing the type inside a tuple                                         \\\cline{2-3}
-                                 & \multirow{1}{*}{Traits}        & G.trait.bound & TraitBound Change                                               \\\cline{1-3}
-\multirow{9}{*}{Borrow Checker} &
-\multirow{2}{*}{Clone}                                         & BC.clone.ref & Dropping clone and adding borrowing                                                   \\
-                                 &                                & BC.clone.drop & Dropping clone   
-                                    \\\cline{2-3}
-    & \multirow{3}{*}{Ref and Deref} & BC.ref.add  & Adding Borrowing                                                           \\
-                                 &                                & BC.ref.drop  & Removing Borrowing                                                         \\
-                                 &                                & BC.deref.to\_string  & Modification in dereferencing a string prior to calling to\_string()                                                            \\\cline{2-3}
-                                 & \multirow{2}{*}{Mut}           & BC.mut.add  & Adding mutability                                                          \\
-                                 &                                & BC.mut.drop  & Dropping mutability                                                        \\\cline{2-3}
-                                 & Vector                         & BC.vec.slice  & Changing a Vec reference to Slice                                           \\\cline{2-3}
-                                 & \multirow{1}{*}{Lifetime}      
-                                                                 & BC.lifetime.static  & Dropping static lifetime
-
-\end{tabular}
-\caption{\label{fig:category}Bug Fix Pattern Categorization}
-\end{table}

### 2022-09-24 main.tex
move files to main directory
warnings fixed
added essence figure, revised methodology, added introduction for results, revised tables for results, we still need to add some numbers
edits

@@
 % your LaTeX will extract the file if required
-\begin{filecontents*}{example.eps}
-%!PS-Adobe-3.0 EPSF-3.0
-%%BoundingBox: 19 19 221 221
-%%CreationDate: Mon Sep 29 1997
-%%Creator: programmed by hand (JK)
-%%EndComments
-gsave
-newpath
-  20 20 moveto
-  20 220 lineto
-  220 220 lineto
-  220 20 lineto
-closepath
-2 setlinewidth
-gsave
-  .4 setgray fill
-grestore
-stroke
-grestore
-\end{filecontents*}
 %
@@
 %\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
-\documentclass[smallextended]{build/svjour3}       % onecolumn (second format)
+\documentclass[smallextended]{svjour3}       % onecolumn (second format)
 %\documentclass[twocolumn]{svjour3}          % twocolumn
@@
 % Insert the name of "your journal" with
-% \journalname{myjournal}
+\journalname{Empirical Software Engineering}
 %
 
-\usepackage{listings, build/listings-rust}
+\usepackage{listings, listings-rust}
 \lstset{escapeinside={<@}{@>}}
@@
 \usepackage{subfiles} % Best loaded last in the preamble
+\input{defs}
 
@@
 \author{Mohammad Robati Shirzad         \and
-        Patrick Lam %etc.
+        Patrick~Lam %etc.
 }
@@
 %\bibliographystyle{spmpsci}      % mathematics and physical sciences
-%\bibliographystyle{spphys}       % APS-like style for physics
+\bibliographystyle{spphys}       % APS-like style for physics
 %\bibliography{}   % name your BibTeX data base
@@
 
+\section{Appendix}
+\subfile{sections/appendix}
+
 \end{document}

### 2022-09-24 sections/results.tex
edits

@@
-In this section, we first evaluate the effectiveness of our data modelling approach (RQ1).
-That is, how successfull is our data modelling approach in capturing the most important aspects of a program change.
-This evaluation is required, as the credibility and generality of our clusters are heavily based on this step in our pipeline.
-Then we move to introducing the obtained clusters in two sections: Common patterns (RQ2), and BC-Related patterns (RQ3).
+In this section, we present the clusters of fix patterns that we found in our data set.
+\RQBC{} is about the borrow-checker related patterns, while \RQG{} is about general patterns.
+However, our technique fundamentally relies on our data modelling approach successfully capturing the most important aspects of program changes.
+We thus first evaluate the data modelling approach as our RQ1, before continuing with \RQBC{} and \RQG{}.
 

### 2022-10-04 sections/related.tex
changed results a bit and added related works sec

--- 2022-09-12 1_introduction.tex
@@
-Producing bugs has been an unavoidable part of computer programming which happens unintentionally.
-Researchers have been trying to help software engineers to be more aware of the bug producing practices in development and how to step towards creating bug-free software.
-One useful step towards that goal would be studying common bug patterns (and their corresponding fixes) that occur across multiple software systems.
-Research efforts for categorizing common bug patterns date back to 1975 where Edres tried to categorize various bug patterns which occurred in operating system program.
+\subsection{Automated Program Repair}
 
-In this work, we empirically analyze the most popular Rust projects in GitHub, and discover the common bug patterns lied within them.
-We believe these insights are important for various reasons, primarily for developing program repair tools.
-Automated program repair tools try to locate bug
-locations in the source code (they include a fault localization module) and then fix them, possibly without human intervention, in a manner such that the modified program meets the desired specifications.
-A powerful automated program repair system has many useful applications and can significantly simplify the process of debugging, resulting in a reduction in software development cost.
-The set of common bug fix patterns can be a useful input to an APR system as it directly provides information about fixing similar buggy structures, and that can significantly reduce the search space for patch generation.
+\subsection{Mining Bug Fix Patters}
 
-Similarly, if we are not observing the desired behaviour from our program, that would indicate that there lies a fault within our program.
-Fault localization is the process of automatically finding the faulty statement(s) in our program.
-If a fault localization tool knows about the common bug patterns that can appear in the underlying programming language, it can prioritize the code segment candidates based on their potentials of producing bugs.
-That would result in a more robust tool as it leverages from historical data.
-
-Moreover, knowing which code patterns in a programming language are more susceptible to produce bugs can help programmers to be more cautious while working with similar code structures.
-This knowledge can make the developers avoid certain mistakes during production which ultimately would result in reducing the frequent bug patterns within the program.
-As Monperrus describes, the less frequent bug patterns we keep in our software, the less cost we would have to pay for software maintenance.
-
-To discuss about a less talked about use case, recently, researchers have been trying to apply deep learning techniques on programs to learn about the semantics of the code or training models which can modify programs towards a specific goal (e.g.
-bug fix).
-One challenge is that usually learning tools get fixed size vectors as input and map that to an output.
-Unlike image data, it is challenging to compress programs into a vector without any loss in semantic information.
-Current approaches try to define the vector dimensions in a way that accounts for all possible abstract syntax tree forms.
-That results in a large sparse space for the inputs.
-Assuming that the deep learning model we are trying to develop is associated with bug fixing tasks, then knowledge about common bug fix patterns can help with pruning input space towards containing less dimensions.
-That can help with creating more efficient models.
-
-There have been many studies on detecting and categorizing bug fix patterns in general, or for a specific programming language. We are more interested in the latter, since
-our study is focused on pervasive bug fix patterns in Rust. \\
 Yang et al. proposed a mining approach to detect Python bug fix patterns via studying
@@
 cross-project bug fix patterns. \\
-In this work, we study common bug patterns that appear in Rust projects. Rust is a relatively new language that allows programmers to write programs that have low-level
-control over resources, as well as high-level safety guarantees. Originally, Rust was a side
-project developed by Graydon Hoare, a Mozilla employee. The company started to fund this project
-in 2009, since they intended to use Rust in some of the low-level Firefox components.
-What distinguishes Rust from the other systems programming languages, e.g. C/C++, is the
-ownership framework that has been incorporated inside the language. The ownership framework keeps the track of the references to different memory locations, and will free them once they
-go out of scope. If the program passes the compiler checks, it automatically guarantees several
-safety problems e.g. dangling references, or double-free problem, that had to be taken care of
-manually in languages like C/C++. Moreover, not only the ownership concept removes the
-need for a garbage collector in the runtime environment, it does not require the programmer to
-explicitly do certain safety tasks, like freeing references.
-Due to the memory and thread safety guarantees that Rust provides, many software companies
-adopted Rust in production. For example, Dropbox decided to migrate to Rust for developing their
-storage system, since they were not completely satisfied with the features provided by AWS’s S3 or
-Golang.
-
-The main module which governs over the ownership rules is called Borrow Checker.
-Because of its importance in Rust, we have decided to put discovered patterns into two groups; (1) non-borrow-checker related patterns (General Patterns), and (2) borrow-checker related patterns (BC-Related Patterns).
-This distinction was really helpful as it made our tool discover clearer results.
-
-We present our discovered insights of our empirical study in the following four research
-questions (RQs):
-
-\textbf{RQ1. How successful is our data modelling approach in capturing all the necessary data?}
-
-\textbf{RQ2. What are the most common AST structures in the bug fixing changes?}
-
-\textbf{RQ3. What are the general fix patterns in Rust?}
-
-\textbf{RQ4. Which fix patterns are related to Borrow Checker in Rust?}

### 2022-10-04 sections/results.tex
changed results a bit and added related works sec

--- 2022-09-23 sections/results.tex
@@
 
-\begin{tabular}{|l|l|l|l|}
+\begin{tabular}{l|l|r|r}
+\textbf{Catgegory} & \textbf{ID} & \textbf{\# Instances} & \textbf{\# Projects} \\
 \hline
-
-\multirow{10}{*}{General}        
-& \multirow{2}{*}{Attributes} & G.attr.struct & -                                  \\
-& & G.attr.struct.field & -                     \\\cline{2-3}
-& \multirow{2}{*}{Struct} & G.struct.field  & -\\
-& & G.struct.field.pub & -    
-\\\cline{2-3}
-& Option & G.option & -                    \\\cline{2-3}
-& \multirow{4}{*}{Types} & G.type.field  & - \\
-& & G.type.generic & -\\ & & G.type.enum.variant  & - \\
-& & G.type.tuple & -                                         \\\cline{2-3}
-& \multirow{1}{*}{Traits} & G.trait.bound & - \\ 
-\hline
+\multirow{2}{*}{Attributes} & G.attr.struct & 172 & 16                                  \\
+& G.attr.struct.field & 62 & 16                     \\\cline{1-4}
+\multirow{2}{*}{Struct} & G.struct.field  & 179 & 17\\
+& G.struct.field.pub & 23 & 12   
+\\\cline{1-4}
+Option & G.option & 22 & 12                    \\\cline{1-4}
+\multirow{4}{*}{Types} & G.type.field  & 72 & 14 \\
+& G.type.generic & 46 & 14\\ & G.type.enum.variant  & 39 & 12 \\
+& G.type.tuple & 6 & 6                                         \\\cline{1-4}
+\multirow{1}{*}{Traits} & G.trait.bound & 6 & 5 \\ 
 \end{tabular}
@@
 \begin{table}[]
-\begin{tabular}{|l|l|l|l|}
+\begin{tabular}{l|l|r|r}
+\textbf{Catgegory} & \textbf{ID} & \textbf{\# Instances} & \textbf{\# Projects} \\
 \hline
-\multirow{9}{*}{Borrow Checker} &
-\multirow{2}{*}{Clone}                                         & BC.clone.ref & - \\
-& & BC.clone.drop & -   
-\\\cline{2-3}
-& \multirow{3}{*}{Ref and Deref} & BC.ref.add & - \\
-& & BC.ref.drop & - \\
-& & BC.deref.to\_string  & -                                            \\\cline{2-3}
-& \multirow{2}{*}{Mut} & BC.mut.add  & - \\
-& & BC.mut.drop  & -                                \\\cline{2-3}
-& Vector & BC.vec.slice  & -                                    \\\cline{2-3}
-& \multirow{1}{*}{Lifetime}      
-& BC.lifetime.static  & -
+\multirow{2}{*}{Clone}                                         & BC.clone.ref & 4 & 3 \\
+& BC.clone.drop & 83 & 10   
+\\\cline{1-4}
+\multirow{3}{*}{Ref and Deref} & BC.ref.add & 17 & 7 \\
+& BC.ref.drop & 145 & 14 \\
+& BC.deref.to\_string  & 7 & 4                                            \\\cline{1-4}
+\multirow{2}{*}{Mut} & BC.mut.add  & 11 & 6 \\
+& BC.mut.drop  & 42 & 15                               \\\cline{1-4}
+Vector & BC.vec.slice  & 8 & 6                                    \\\cline{1-4}
+\multirow{1}{*}{Lifetime}      
+& BC.lifetime.static  & 8 & 6
 \\
-\hline
 \end{tabular}

### 2022-10-04 sections/methodology.tex

@@
 First, we need to define a fixed set of columns.
-We need fixed-size datapoints, as it is a requirement for both centroid-based and density-based well-known clustering algorithms.
+We need fixed-size datapoints, as they are a requirement for both centroid-based and density-based well-known clustering algorithms.
 The number of columns would depend on the amount of information we want to encode in the datapoints.
+
 The most naive approach would be to only report the observed non terminals within the diff.
@@
 On the other hand, theoretically, for full order-sensitiveness, we would need to encode all possible combinations of items from $\mathtt{NT} \cup \mathtt{BCL}$ as our dimensions, which yields an infeasible $O(n^n)$ dimensions; this could potentially be reduced somewhat, but is still impractical.
-However, in this case we can put a distinction between the aformentioned if statements, as we have a different dimension for nested if statements (\verb+if_if_statement_...+).
+While we can implement a specific workaround to distinguish nested ifs (and we have done so), the general point about the large dimensionality of fully order-sensitive encodings still holds.
+
 A reasonable work around would be to reduce $n$ from the full set of non terminals to a smaller set of categories (e.g.
@@
     \includegraphics[width=1\textwidth]{figs/essence.png}
-    \caption{\label{fig:essence}Multiplying the number of occurrences of items observed in the tree by their respective weight, resulting in the essence of change (Visualized with a circle pack figure)}
+    \caption{\label{fig:essence}Multiplying the number of occurrences of items observed in the tree by their respective weights, resulting in the essence of change (Visualized with a circle pack figure)}
 \end{figure}
@@
 We made two observations about changes.
-First, we saw that an item (non terminal or BC-related lexemes) that occurs closer to AST leaves tends to be more semantically important.
+First, we saw that an item (non terminal or BC-related lexeme) that occurs closer to AST leaves tends to be more semantically important.
 Relatedly, items closer to the root tend to be repeated in all the paths within one ASTDiff (e.g.
@@
 
-Multiplying the number of occurrences of all observed items in a change by their respective weight results in a vector of numbers.
-We call this vector, the essence of change.
-In this vector, the respective value for each item shows its importance in the change.
+Multiplying the number of occurrences of all observed items in a change by their respective weights results in a vector of numbers.
+We call this vector the \emph{essence of change}.
+In this vector, the value for each item shows its importance in the change.
 Figure~\ref{fig:essence} shows the process of computing the essence of the code change that we saw in Figure~\ref{fig:extraction}.
@@
 An abstract pattern as opposed to a concrete pattern can account for many changes in the programs.
-For instance, the pattern `Changing a clone of a variable to a borrowing of it in a function argument' is deemed more concrete than `Adding a new statement to a function's body':
+For instance, we deem the pattern `Changing a clone of a variable to a borrowing of it in a function argument' more concrete than `Adding a new statement to a function's body':
 

### 2022-10-06 sections/related.tex
2nd half related added

@@
 
-\subsection{Mining Bug Fix Patters}
+\subsection{Mining Bug Fix Patterns and Pattern-based Program Repair}
 
-Yang et al. proposed a mining approach to detect Python bug fix patterns via studying
-fine-grained fixing code changes. In their research, they also examined what portion of the bugs
-could be fixed using automated bug fixing approaches. Moreover, they evaluated the fix patterns in
-the wild and realized that 37\% of the buggy codes could be matched by the fix patterns they had
-found. \\
-Hanam et al. conducted a similar research but for finding Javascript most pervasive bug fix
-patterns. They presented a large-scale study of the most common bug fix patterns by mining 105K
-commits from 134 server-side JavaScript projects. They used DBSCAN clustering algorithm
-and divided bug fixing change types into 219 clusters, from which they extracted 13 pervasive
-cross-project bug fix patterns. \\
+The main intuition in pattern-based program repair methods is that the bug fixing modifications of the programs follow certain patterns.
+Thus, prior study of the common bug fix patterns is a crucial step for developing these repair tools.
+Pan et al.
+\ref{} analyzed seven large-scale widely used Java projects and obtained 27 common bug fix patterns.
+Though they did the analysis on Java projects, their reported patterns were not specific to Java.
+Martinez and Monperrus \ref{} \ref{} exploited the frequency of observed patterns to introduce a heuristic patch searching method.
+In their empirical evaluation, they concluded that the probabilistic values for repair actions can help with reducing the search space, hence creating a more effective repair tool.
+
+Hanam et al.
+conducted a similar research but for finding Javascript most pervasive bug fix patterns.
+They presented a large-scale study of the most common bug fix patterns by mining 105K commits from 134 server-side JavaScript projects.
+They used DBSCAN clustering algorithm and divided bug fixing change types into 219 clusters, from which they extracted 13 pervasive cross-project bug fix patterns.
+
+Yang et al.
+proposed a mining approach to detect Python bug fix patterns via studying fine-grained fixing code changes.
+In their research, they also examined what portion of the bugs could be fixed using automated bug fixing approaches.
+Moreover, they evaluated the fix patterns in the wild and realized that 37\% of the buggy codes could be matched by the fix patterns they had found.
+
+\subsection{Code Patterns in Rust}
+
+Qin et al.
+\ref{} conducted the first empirical study on real-world Rust program behaviours.
+They manually inspected 850 unsafe code usages and 17 bugs across five open-source Rust projects, five Rust libraries, two online security databases, and the standard library of Rust.
+They analyzed the motivation behind unsafe code usage and removal, in addition to obtaining 70 memory-safety issues and 100 concurrency bugs.
+They also provided Rust programmers with some suggestions and insights to develop better Rust programs.
+At last, using the results of their analysis, they designed two bug detectors and provided recommendations for developing bug detector in the future.
+
+Li et al.
+\ref{} present MirChekcer, a fully automated bug detection framework for Rust.
+This framework works by carrying out static analysis on Rust's Mid-level Intermediate Representation (MIR).
+The tool exploits the insights obtained from observing existing bugs in Rust code bases.
+Through recording both numerical and symbolic information, the framework detects errors by using constraint solving techniques.
+The framework proves its practicability as it detected 33 new bugs including 16 memory-safety faults within 12 Rust crates.

### 2022-10-06 clean.sh

@@ -1,0 +1,7 @@
+mv *.eps ./build;
+mv *.aux ./build;
+mv *.fdb_latexmk ./build;
+mv *.fls ./build;
+mv *.log ./build;
+mv *.out ./build;
+mv *.synctex.gz ./build;

### 2022-10-07 sections/related.tex
added related works part 1

@@
-\subsection{Automated Program Repair}
+\subsection{Search-Based Program Repair}
+
+Automated program repair aims to debug faulty source code without human involvement.
+Prior to code modification, repair tools exploit a fault localization module to speculate the location of bug.
+Fault localization module ranks program statements based on their suspiciousness.
+The level of suspiciousness follows a statistical model which is devised based on the fact that buggy statements are executed mostly by failed test cases \ref{}.
+Researchers have proposed different methods for code repair, each of which follows a different mindset.
+Here we provide a summarization of the advancements in these different mindset, and will discuss how having a set of predefined fix patterns can be beneficial for repair purposes.
+
+This line of research in program repair follows the competent programmer hypothesis which states that, syntactically, a faulty programs is not that far away from its correct version.
+Using this concept, we can develop mutation operators, which are the operators that change an expression or a statement.
+We can apply the mutation operators on the list given by the fault localization module.
+If the mutation on a statement does not result in complete success after running the test cases, we assume that the statement is correct and move on to the next statement in the list.
+
+The computation cost depends on the number of statements and their respective mutation operations.
+Moreover, each time we mutate our program, we ought to evaluate the result by checking the program against the required specification.
+Given a sufficiently large enough search space, a complete search approach renders the repair tool virtually useless.
+That is why, the researchers have proposed heuristic searching algorithms to rectify the faulty program faster.
+
+Weimer et al.
+\ref{} developed GenProg, a program repair tool that used genetic programming.
+The main idea behind this tool is that statements follow certain patterns in a codebase.
+Therefore, if we can find the correct version of a faulty statement somewhere in the program.
+This tool showed promising results as it managed to find patches without any additional annotations or human involvements.
+
+However, Arcuri and Briand \ref{} suspected the effectiveness of genetic programming in GenProg, as the tool mostly found the patch in the random initialization, in other words, before the evolution begins.
+Through developing TrpAutoRepair \ref{} and RSRepair \ref{}, the test case prioritization is necessary while using genetic programming, as fitness evaluation is an expensive part of the repair pipeline.
+
+Tan and Rovehoudhury \ref{} used mutation repairing for fixing regression bugs.
+They categorized common code changes in real-world regressions after studying 73 program evolution benchmarks.
+They used this categorization to design mutation operators, and chose them based on the faulty location’s to repair the program.
 

### 2022-10-07 sections/discussion.tex
added threats to validity

\subsection{Threats to Validity}

In our work there are three main threats to internal validity: (1) We use a weighting scheme which is based on a heuristic to give more value to tree elements closer to the leaves. In addition to that, we also manually adjusted the weights of borrow checker related elements to respect ownership framework's role in Rust, though we recored their results separate from general results. As this approach is based on an intuition, it does not guarantee achieving optimal clusters. (2) The collected commits are the ones that fix a bug in a previously pushed commit. We do not consider the bugs that occur during a commit and also fixed during the same commit. (3) Our data modelling approach is based on the frequency of observed program elements in ASTs, and we use DBSCAN as our clustering algorithm. Other program data modelling methods and clustering algorithms (e.g. k-means, or SLINK) might output new clusters that our pipeline is unable to find.

The threat to external validity can be the developers reporting the commits as bug fixing commits while they are not really fixing a functionality. Similarly, a commit message might contain our target keywords while the commit is not associated with bug fixing changes.

### 2022-10-07 run.sh

@@ -1,9 +1,0 @@
-pdflatex main.tex;
-sleep 2;
-mv *.eps ./build;
-mv *.aux ./build;
-mv *.fdb_latexmk ./build;
-mv *.fls ./build;
-mv *.log ./build;
-mv *.out ./build;
-mv *.synctex.gz ./build;

### 2022-10-08 sections/introduction.tex
citation added for intro

--- 2022-09-16 introduction.tex
@@
 One useful step towards that goal would be studying common bug patterns (and their corresponding fixes) that occur across multiple software systems.
-Research efforts for categorizing common bug patterns date back to 1975 where Edres tried to categorize various bug patterns which occurred in operating system implementations.
+Research efforts for categorizing common bug patterns date back to 1975 where Endres tried to categorize various bug patterns which occurred in operating system implementations \cite{endres1975analysis}.
 
@@
 Our interest comes from a desire to develop program repair tools, but we believe these insights can also be of broader service.
-Automated program repair tools try to locate bug locations in the source code (using a fault localization module) and then generate patches to fix them, possibly without human intervention, so that the modified program meets desired specifications (implicit or explicit).
-A powerful automated program repair system has many useful applications and could significantly simplify debugging, thus reducing software development cost.
-A set of common bug fix patterns can be a useful input to an APR system: it directly provides information about fixing similar buggy structures, which can significantly reduce the search space for patch generation.
+Automated program repair \cite{goues2019automated} \cite{liu2018survey} tools try to locate bug locations in the source code (using a fault localization module) and then generate patches to fix them, possibly without human intervention, so that the modified program meets desired specifications (implicit or explicit).
+A powerful automated program repair system has many useful applications and could significantly simplify debugging, thus reducing software development cost \cite{le2012systematic}.
+A set of common bug fix patterns can be a useful input to an APR system: it directly provides information about fixing similar buggy structures, which can significantly reduce the search space for patch generation \cite{jeffrey2009bugfix}.
 
@@
 But where is the fault?
-Fault localization is the process of automatically finding the faulty statement(s) in our program.
+Fault localization \cite{wong2016survey} is the process of automatically finding the faulty statement(s) in our program.
 If a fault localization tool knows about the common bug patterns that can appear in the underlying programming language, it can prioritize code segment candidates based on their potentials of producing bugs.
@@
 This knowledge can make the developers avoid certain mistakes during production which ultimately would result in reducing the frequent bug patterns within the program.
-As Monperrus describes, the fewer frequent bug patterns we keep in our software, the less cost we would have to pay for software maintenance.
+As Monperrus describes, the fewer frequent bug patterns we keep in our software, the less cost we would have to pay for software maintenance \cite{monperrus2014critical}.
 
 Moving further afield, researchers have been exploiting deep learning techniques to learn about the semantics of computer programs and create program modifier models that have a specific goal (e.g.
-bug fix).
+bug fix) \cite{alon2019code2vec} \cite{alon2018general} \cite{raychev2016learning} \cite{bielik2016phog}.
 However, most learning tools get fixed-size vectors as input and map them to outputs.
@@
 
-Many studies have investigated detecting and categorizing bug fix patterns in general, or for a specific programming language.
+Many studies have investigated detecting and categorizing bug fix patterns in general \cite{islam2021changes} \cite{madeiral2018towards} \cite{pan2009toward}, or for a specific programming language \cite{yang2022mining} \cite{hanam2016discovering} \cite{campos2019discovering}.
 In this work, we study common bug patterns that appear in Rust projects.
-Rust is a relatively new language that allows programmers to write programs that have low-level control over resources, as well as high-level safety guarantees.
+Rust is a relatively new language that allows programmers to write programs that have low-level control over resources, as well as high-level safety guarantees \cite{klabnik2019rust}.
 What distinguishes Rust from the other systems programming languages, e.g.
@@
 These problems had to be manually ensured in languages like C/C++.
-Moreover, not only does the ownership concept remove the need for a garbage collector in the runtime environment, it does not require the programmer to explicitly do certain safety tasks.
+Moreover, not only does the ownership concept remove the need for a garbage collector in the runtime environment, it does not require the programmer to explicitly do certain safety tasks \cite{qin2020understanding}.
 Due to the memory and thread safety guarantees that Rust provides, many software companies adopted Rust in production.
-For example, Dropbox decided to migrate to Rust for developing their storage system, since they were not completely satisfied with the features provided by AWS’s S3 or Golang.
+For example, Dropbox decided to migrate to Rust for developing their storage system, since they were not completely satisfied with the features provided by AWS’s S3 or Golang \cite{dropbox}.
 
@@
 Because of the Borrow Checker's importance in Rust, we split patterns into two groups: (1) non-borrow-checker related patterns (General Patterns), and (2) borrow-checker related patterns (BC-Related Patterns).
-We present our discovered insights of our empirical study in the following four research
+\\
+
+\noindent \textbf{Research Questions} We present our discovered insights of our empirical study in the following four research
 questions (RQs):
 
-\textbf{RQ1. How successful is our data modelling approach in representing program change?}
+\begin{itemize}
+    \item \textbf{RQ1. Does our data modelling approach capture the most important aspect of program changes?} 
+    \item \textbf{RQ2. What are the general fix patterns in Rust, and how often do they apply?} 
+    \item \textbf{RQ3. What are borrow-checker related fix patterns in Rust, and how often do they apply?}
+\end{itemize}
 
-\textbf{RQ2. What are the general fix patterns in Rust, and how often do they apply?}
+\noindent \textbf{Contribution} This paper makes the following primary contributions:
 
-\textbf{RQ3. What are borrow-checker related fix patterns in Rust, and how often do they apply?}
+\begin{itemize}
+    \item To the best of our knowledge, this work is the first project that exploits a code analysis pipeline to automatically mine Rust open source projects and extract the common cross project bug fix patterns.
+    \item We propose a novel method to encode key AST information in fixed size datapoints using a semi-automatically derived weighting scheme.
+    \item To respect the role of borrow checker in Rust, our paper reports and discusses the bug fix patterns that are related to the ownership framework in Rust separate from the general patterns.
+\end{itemize}

### 2022-10-08 sections/methodology.tex
citation added for intro

@@
 \caption{\label{table:repos} Target repositories for mining}
-\begin{tabular}{|c|c|c|c|c|}
-    \hline
+\begin{tabular}{c|c|c|c|c}
     Project Name & Stars(k) & LOC & \# Commits & Avg. Code Churn \\ \hline
@@
     rustdesk/rustdesk & 27.7 & 85322 & 1664 & 72.81 \\ \hline
-    tokio-rs/tokio & 17.2 & 126089 & 3006 & 28.87 \\ \hline
+    tokio-rs/tokio & 17.2 & 126089 & 3006 & 28.87 
 \end{tabular}
@@
 Having collected our datapoints, our next task is to cluster them so that we can categorize bug patterns.
-Like~\cite{JS paper, Python paper}, we use the DBSCAN clustering algorithm for two main reasons.
+Like~citehere, we use the DBSCAN clustering algorithm for two main reasons.
 Firstly, in contrast to k-means, it does not require the number of clusters in advance as an input to the algorithm.

### 2022-10-08 clean.sh

@@ -1,7 +1,17 @@
-mv *.eps ./build;
-mv *.aux ./build;
-mv *.fdb_latexmk ./build;
-mv *.fls ./build;
-mv *.log ./build;
-mv *.out ./build;
-mv *.synctex.gz ./build;
+rm *.bcf
+rm *.blg
+rm *.bbl
+rm *.run.xml
+rm main-blx.bib
+rm *.eps ;
+rm *.aux ;
+rm *.fdb_latexmk ;
+rm *.fls ;
+rm *.log ;
+rm *.out ;
+rm *.synctex.gz ;
+pdflatex main.tex
+bibtex main
+pdflatex main.tex
+sleep 2;
+pdflatex main.tex

### 2022-10-11 sections/methodology.tex
merge is needed for methodology

@@
 For each repository, we search all commits that include bug-fixing related keywords within their commit messages.
-Here, we chose bug-fixing related keywords `bug', `error', `fault', `defect', and `fix'.
-% cite earlier work that does the search-for-keywords-to-identify-bugs; my earlier EMSE paper has that citation for instance. -- same issue as above, not sure how to cite
+Here, we chose bug-fixing related keywords based on the set introduced in \cite{zhang2018empirical}, excluding `nan' and `inf'.
 
@@
 Having collected our datapoints, our next task is to cluster them so that we can categorize bug patterns.
-Like~citehere, we use the DBSCAN clustering algorithm for two main reasons.
+Like~\cite{hanam2016discovering}, we use the DBSCAN clustering algorithm for two main reasons.
 Firstly, in contrast to k-means, it does not require the number of clusters in advance as an input to the algorithm.
 Secondly, it is a density-based clustering method, which means that it detects arbitrarily shaped clusters; centroid-based methods like k-means can't detect such clusters.
-We tried k-means and the more powerful fuzzy variant F-means and did not get useful clusters.
+We tried k-means and the more powerful fuzzy variant Fuzzy C-means \cite{dunn1973fuzzy} and did not get useful clusters.
 
@@
 \centering
-\includegraphics[width=1\textwidth]{figs/clusters.png}
-\caption{\label{fig:clustering} Clustering results with nine different combinations of parameters ($\epsilon=[10^{-2}, 10^{-3}, 10^{-4}], Z=[5,7,9])$}
+\includegraphics[width=1\textwidth]{figs/clusters.pdf}
+\caption{\label{fig:clustering} Clustering results with nine different combinations of parameters $\epsilon$ and $Z$, respectively separated in columns and rows.
+In each subplot, the two blue bars in the left indicate the number of clustered datapoints (the number is shown above each bar) after applying clustering on $D_g$ ($G_c$) and $D_b$ ($BC_{c}$), respectively.
+The two orange bars in the right show similar data but for noise datapoints.
+In each subplot's legend, $C_G$ and $C_B$ show the number of clusters obtained from $D_g$ and $D_b$, respectively}
 \end{figure}
@@
 The person analyzing the cluster then looks at the code and specifies the clusters that contained more than five datapoints showing similar patterns.
-For the manual analysis, we used $\epsilon=0.0001, Z=5$ and $\epsilon=0.001, Z=5$, which were the best parameters for $D_g$ and $D_b$, resulting in 596 and 61 clusters, respectively.
+For the manual analysis, we used $\epsilon=0.0001, Z=5$ and $\epsilon=0.001, Z=5$, which were the best parameters for $D_g$ and $D_b$, resulting in 577 and 102 clusters, respectively.
 As a final step, we exclude clusters that contained datapoints from fewer than 3 projects.
@@
 
-Similar to ..., following Cotroneo et al. definitions, we divide the clusters in three different groups:
+Similar to \cite{yang2022mining}, following Cotroneo et al. \cite{cotroneo2019analyzing} definitions, we divide the clusters in three different groups:
 

### 2022-10-11 sections/methodology.tex
merge successful

@@
 The general idea for automatically finding the pervasive code patterns within a list of programs is to parse the programs and encode the important parts of the derived AST in fixed-size datapoints.
-Work in this vein makes the assumption that each cluster corresponds to one change pattern.
+Work in this vein makes the assumption that each cluster corresponds to one change pattern \cite{hanam2016discovering} \cite{campos2019discovering} \cite{yang2022mining}.
 
@@
 If the trees are programs' abstract syntax trees, we call their diff an ASTDiff.
-We use dictdiff to obtain ASTDiffs.
+We use dictdiffer to obtain ASTDiffs.
 
@@
 Our target repositories were the top 18 most-starred open source Rust projects on GitHub.
-We mined (using Pydriller) all the bug related commits and ran them through our pipeline.
-Then, we applied the DBSCAN clustering algorithm on the resulting datapoints.
+We mined (using Pydriller \cite{spadini2018pydriller}) all the bug related commits and ran them through our pipeline.
+Then, we applied the DBSCAN clustering algorithm \cite{ester1996density} on the resulting datapoints.
 In Section~\ref{sec:clustering_data} we justify our choice of DBSCAN for clustering, and how we tuned it to improve cluster quality.
@@
 containing one dimension per non terminal.
-However, we are also interested in change patterns that involve the borrow checker (BC), so we also add a dimension reflecting the presence of a set of BC-related lexemes (BCL) that we identified, e.g.
+However, we are also interested in change patterns that involve the borrow checker (BC), so we also add a dimension reflecting the presence of a set of BC-related elements (BCE) that we identified, e.g.
 \texttt{clone}, \texttt{Rc}, \texttt{Box}.
-The full list of \verb+BCL+ elements can be found within Appendix~\ref{sec:appendix}.
-% I think we don't need this sentence here. Let's see what we do later.
-%We define dimension items as \verb+DI = NT+ $\cup$ \verb+BCL+, as this set also equals the columns of our tables.
+The full list of \verb+BCE+ elements can be found within Appendix~\ref{sec:appendix}.
 
@@
 This yields two Syn trees.
-Using the PLY tool (a Python implementation of lex and yacc), we wrote a simple transformer from the serialized Rust AST to Python dictionaries.
+Using the PLY tool\footnote{\url{https://www.dabeaz.com/ply/}} (a Python implementation of lex and yacc), we wrote a simple transformer from the serialized Rust AST to Python dictionaries.
 The change that transforms the first tree to the second one is the fix that was applied in the commit.
-To find this difference, we use dictdiff, a Python library to find the diff of two Python dictionaries.
+To find this difference, we use dictdiffer\footnote{\url{https://dictdiffer.readthedocs.io/en/latest/}}, a Python library to find the diff of two Python dictionaries.
 Its output, in our context, is the ASTDiff.
@@
 
-The output of dictdiff is a list of diffs, where each diff comprises three parts. The first part specifies the modification type:
+The output of dictdiffer is a list of diffs, where each diff comprises three parts. The first part specifies the modification type:
 
@@
 A Depth First Search allows us to collect all paths, and we return only the paths that we store in our datapoints (i.e.
-touching $\mathtt{NT} \cup \mathtt{BCL}$).
+touching $\mathtt{NT} \cup \mathtt{BCE}$).
 
@@
 First, we need to define a fixed set of columns.
-We need fixed-size datapoints, as they are a requirement for both centroid-based and density-based well-known clustering algorithms.
+We need fixed-size datapoints, as they are a requirement for both centroid-based and density-based well-known clustering algorithms \cite{xu2005survey}.
 The number of columns would depend on the amount of information we want to encode in the datapoints.
@@
 In such an encoding, for instance, there is no difference between two nested if statements and two if statements beside each other.
-On the other hand, theoretically, for full order-sensitiveness, we would need to encode all possible combinations of items from $\mathtt{NT} \cup \mathtt{BCL}$ as our dimensions, which yields an infeasible $O(n^n)$ dimensions; this could potentially be reduced somewhat, but is still impractical.
+On the other hand, theoretically, for full order-sensitiveness, we would need to encode all possible combinations of items from $\mathtt{NT} \cup \mathtt{BCE}$ as our dimensions, which yields an infeasible $O(n^d)$ dimensions ($n=|\mathtt{NT} \cup \mathtt{BCE}|$ and $d$ is the depth of nested structures); this could potentially be reduced somewhat, but is still impractical.
 While we can implement a specific workaround to distinguish nested ifs (and we have done so), the general point about the large dimensionality of fully order-sensitive encodings still holds.
@@
 A reasonable work around would be to reduce $n$ from the full set of non terminals to a smaller set of categories (e.g.
-a category for larger entites like class or function definitions, and a category for smaller entities like statements and expressions), and then to record all possible combinations of these categories (similar to JS authors~\cite{foo}).
+a category for larger entites like class or function definitions, and a category for smaller entities like statements and expressions), and then to record all possible combinations of these categories (similar to \cite{hanam2016discovering}).
 That approach reduces the number of dimensions and provides more reasonable and syntactically-correct combinations.
 However, it would still yield a sparse dataset.
-% todo you need to dig up the JS paper ref / Not sure, how to cite, I usually create a bib file, but i'm not sure if i'm allowed to do that for this template.
 
 In our methodology, we propose a novel representation.
-Similar to the first approach, we order-insensitively define the columns of our dataset as the set of items $\mathtt{NT} \cup \mathtt{BCL}$ we collected from Syn.
+Similar to the first approach, we order-insensitively define the columns of our dataset as the set of items $\mathtt{NT} \cup \mathtt{BCE}$ we collected from Syn.
 However, to simulate order-sensitiveness, we carried out two auxiliary steps.
@@
 We made two observations about changes.
-First, we saw that an item (non terminal or BC-related lexeme) that occurs closer to AST leaves tends to be more semantically important.
+First, we saw that an item (non terminal or BC-related element) that occurs closer to AST leaves tends to be more semantically important.
 Relatedly, items closer to the root tend to be repeated in all the paths within one ASTDiff (e.g.
@@
 
-Furthermore, as we wanted to make sure that our tool captures patterns related to the Rust borrow checker, we manually increased the weights of the \texttt{BCL} items.
+Furthermore, as we wanted to make sure that our tool captures patterns related to the Rust borrow checker, we manually increased the weights of the \texttt{BCE} items.
 The manual adjustment is why we characterize our weighting scheme as semi-automatically designed.
@@
 Pydriller provides APIs to extract commits from a Git repository and to search through different revisions of files.
-Algorithm~\ref{alg} shows how we used Pydriller to run the target repositories through our code analysis pipeline and populate two databases: one for borrow-checker related patterns (which involve items in \texttt{BCL}) and one for general patterns (which don't).
+Algorithm~\ref{alg} shows how we used Pydriller to run the target repositories through our code analysis pipeline and populate two databases: one for borrow-checker related patterns (which involve items in \texttt{BCE}) and one for general patterns (which don't).
 For each repository, we search all commits that include bug-fixing related keywords within their commit messages.
@@
 \caption{\label{fig:clustering} Clustering results with nine different combinations of parameters $\epsilon$ and $Z$, respectively separated in columns and rows.
-In each subplot, the two blue bars in the left indicate the number of clustered datapoints (the number is shown above each bar) after applying clustering on $D_g$ ($G_c$) and $D_b$ ($BC_{c}$), respectively.
-The two orange bars in the right show similar data but for noise datapoints.
+In each subplot, the two blue bars in the left respectively indicate the number of clustered and noise datapoints (the number is shown above each bar) after applying clustering on $D_g$ (denoted with $G_c$ and $G_n$ on the x axis).
+The two orange bars in the right show similar data but for the $D_{b}$ database.
 In each subplot's legend, $C_G$ and $C_B$ show the number of clusters obtained from $D_g$ and $D_b$, respectively}

### 2022-10-11 sections/discussion.tex
discussion v1 added

@@
+In this work, our pipeline successfully mined 19 cross project bug fix patterns.
+However, our pipeline fails to detect complex pattern due to space limitation and their innate complicated features.
+A complicated pattern is a pattern that includes many changes in many places and it makes it easier for two datapoints to be dissimilar.
+
+Also, we expected to obtain certain patterns regarding some ertain Rust related features, such as unsafe Rust, but our results did not include those patterns.
+Our speculation is that, since Rust has a strong compiler check before creating the program executable, the programmers usually get many errors, and they only push their code once it is free of compile errors.
+That is why, we think there are some patterns that were hidden from our sight.
+
+We believe our results can help researchers to create Rust tools for repair, fault localization, IDE linters, and etc.
+Also, these results can help program encoding methods to focus on more important aspects of a program and shrink down information more effectively.
+
 \subsection{Threats to Validity}
 
-In our work there are three main threats to internal validity: (1) We use a weighting scheme which is based on a heuristic to give more value to tree elements closer to the leaves.
-In addition to that, we also manually adjusted the weights of borrow checker related elements to respect ownership framework's role in Rust, though we recored their results separate from general results.
+In our work there are three main threats to internal validity: (1) We use a weighting scheme which is based on a heuristic to give more values to tree elements closer to the leaves.
+In addition to that, we also manually adjusted the weights of borrow checker related elements to respect ownership framework's role in Rust, though we recorded their results separate from general results.
 As this approach is based on an intuition, it does not guarantee achieving optimal clusters.
@@
 (3) Our data modelling approach is based on the frequency of observed program elements in ASTs, and we use DBSCAN as our clustering algorithm.
-Other program data modelling methods and clustering algorithms (e.g.
-k-means, or SLINK) might output new clusters that our pipeline is unable to find.
+Other program data modelling methods and clustering algorithms (e.g.SLINK) might output new clusters that our pipeline is unable to find.
 
 The threat to external validity can be the developers reporting the commits as bug fixing commits while they are not really fixing a functionality.
-Similarly, a commit message might contain our target keywords while the commit is not associated with bug fixing changes.
+Similarly, a commit message might not contain our target keywords while the commit is associated with bug fixing changes.

### 2022-10-11 sections/results.tex
discussion v1 added

@@
 As discussed in the previous section, we used an encoding mechanism to incorporate AST information in fix-sized datapoints which then could be used in the DBSCAN clustering algorithm.
-We used heuristics (weighting scheme) while encoding our datapoints to make the most important program elements stand out between all the non terminals seen in the AST.
+We used heuristics (weighting scheme) while encoding our datapoints to make the most important program elements stand out between all the non terminals seen in the AST (alongside some BC-related elements).
 Here we want to evaluate the effectiveness of our encoding.
@@
 The analyzer compares the actual change in the code and compares it to the visual representation of its respective data point.
-We used circle pack figures to visualize each data point.
-In these figures, the radius of the circle of each non terminal shows its importance in our encoding (Fig ref).
+We used circle pack figures to visualize each data point \cite{collins2003circle}.
+For this purpose, we used packcircles python library\footnote{\url{https://github.com/mhtchan/packcircles}}.
+In these figures, the radius of the circle of each non terminal shows its importance in our encoding (Similar to Figure~\ref{fig:essence}).
 
@@
 At last, we report the results as the portion of the changes which both authors agreed upon its encoding correctness.
-The final results are TBD\% and TBD\% correctness in the first and second experiments, respectively, which shows the effectiveness of our encoding mechanism.
+The final results are 55\% and 78\% correctness in the first and second experiments, respectively, which shows the effectiveness of our encoding mechanism.
 
@@
 
-For instance, in the Rust project swc-project/swc (A fast TypeScript/JavaScript compiler), the developer needed to modify the CSS parser source code in a way that it accounts for empty @media queries \footnote{\url{https://github.com/swc-project/swc/commit/75a14f98b7370226115ee24eec6eb8c802bd4837}}, so, as shown in the code snippet above, they changed type MediaQueryList to Option<MediaQueryList>, and modified the other parts of the source code, accordingly.
+For instance, in the Rust project swc-project/swc (A fast TypeScript/JavaScript compiler), the developer needed to modify the CSS parser source code in a way that it accounts for empty @media queries \footnote{\url{https://github.com/swc-project/swc/commit/75a14f98b7370226115ee24eec6eb8c802bd4837}}, so, as shown in the code snippet above, they changed type MediaQueryList to Option\verb+<+MediaQueryList\verb+>+, and modified the other parts of the source code, accordingly.
 
@@
 \hline
-\multirow{2}{*}{Clone}                                         & BC.clone.ref & 4 & 3 \\
-& BC.clone.drop & 83 & 10   
+\multirow{2}{*}{Clone}                                         & BC.clone.drop & 83 & 10 \\
+& BC.clone.ref & 4 & 3   
 \\\cline{1-4}
@@
 \subsubsection{Clone}
+
+\noindent \textit{ \textbf{BC.clone.drop} Removing clone}
+
+\begin{lstlisting}[language=Rust, style=colouredRust]
+start_plugin_process(
+--  manifest.clone(),
+++  manifest,
+    self.next_plugin_id(),
+    self.self_ref.as_ref().unwrap().clone(),
+);
+\end{lstlisting}
+
+\noindent \textbf{Description:} This pattern is referred to as redundant clone in Clippy Lints.
+Clippy\footnote{\url{https://github.com/rust-lang/rust-clippy}} is a linter tool for Rust that catches common mistakes.
+There are more than 500 lints included in Clippy.
+With regards to this pattern, Clippy detects whether a variable clone is necessary and could we have used the variable itself or not.
+This change does not modify the program behaviour, however it can boost the performance significantly if we repeatedly cloning large variables.
 
@@
 
-\noindent \textit{ \textbf{BC.clone.drop} Removing clone}
-
-\begin{lstlisting}[language=Rust, style=colouredRust]
-start_plugin_process(
---  manifest.clone(),
-++  manifest,
-    self.next_plugin_id(),
-    self.self_ref.as_ref().unwrap().clone(),
-);
-\end{lstlisting}
-
-\noindent \textbf{Description:} This pattern is referred to as redundant clone in Clippy Lints.
-Clippy detects whether a variable clone is necessary and could we have used the variable itself or not.
-This change does not modify the program behaviour, however it can boost the performance significantly if we repeatedly cloning large variables.
-
 \subsubsection{Ref and Deref}
@@
 
-To give an example, in starship/starship after one of the commits \footnote{\url{https://github.com/starship/starship/commit/56d475578ea508631275772127f49a6949fea6b0}{git\_branches.rs}}, the developer borrows the value of $repo\_dir$ to keep the ownership of the object in the current scope.
-Now this object can then be used in another function invocation ($remove\_dir\_all(repo\_dir)$) which fixes a bug.
+To give an example, in starship/starship after one of the commits \footnote{\url{https://github.com/starship/starship/commit/56d475578ea508631275772127f49a6949fea6b0}{git\_branches.rs}}, the developer borrows the value of \verb+repo_dir+ to keep the ownership of the object in the current scope.
+Now this object can then be used in another function invocation (\verb+remove_dir_all(repo_dir)+) which fixes a bug.
 \\
@@
 \noindent \textbf{Description:} This is a change that in all the instances has been detected by Clippy.
-The change applies on an object prior to calling $to\_string()$ on that object.
+The change applies on an object prior to calling \verb+to_string()+ on that object.
 The change might add or remove a dereference, as shown above.
-In the addition case, the reason behind the change is that the object a reference type of $T$, while $T$ directly implements $to\_string()$.
-It is faster and more performance efficient to dereference the object and then call $to\_string()$ on it.
+In the addition case, the reason behind the change is that the object a reference type of \verb+T+, while \verb+T+ directly implements \verb+to_string()+.
+It is faster and more performance efficient to dereference the object and then call \verb+to_string()+ on it.
 The deletion case indicates a redundant dereference, thus it is removed.
@@
 In the observed instances, this pattern was not seen in refactoring changes, the modification happened alongside a change in program behaviour.
+\\
 

### 2022-10-12 sections/methodology.tex
figs vectorized; tables fixed; fig captions rewritten; citations added; related works added; discussion fixed; no abstract and no conclusion yet

@@
 \centering
-\includegraphics[width=0.8\textwidth]{figs/extraction.png}
-\caption{\label{fig:extraction}Extracting paths from an ASTDiff. AST corresponds to the code change in the yellow box}
+\includegraphics[width=0.8\textwidth]{figs/extraction.drawio.pdf}
+\caption{\label{fig:extraction}Extracting paths from an ASTDiff.
+the ASTDiff on left corresponds to the code change in the yellow box on the top right.
+A DFS search will identify three paths, shown in three different colors at right.}
 \end{figure}
@@
     \centering
-    \includegraphics[width=1\textwidth]{figs/essence.png}
-    \caption{\label{fig:essence}Multiplying the number of occurrences of items observed in the tree by their respective weights, resulting in the essence of change (Visualized with a circle pack figure)}
+    \includegraphics[width=1\textwidth]{figs/essence.drawio.pdf}
+    \caption{\label{fig:essence}Multiplying the number of occurrences of items observed in the tree by their respective weights, resulting in the essence of change (visualized with a circle pack figure at right).
+The left and right key value tables show the number of occurrences of each item and its respective weight, respectively.}
 \end{figure}

### 2022-10-12 sections/related.tex
figs vectorized; tables fixed; fig captions rewritten; citations added; related works added; discussion fixed; no abstract and no conclusion yet

@@
 Prior to code modification, repair tools exploit a fault localization module to speculate the location of bug.
-Fault localization module ranks program statements based on their suspiciousness.
-The level of suspiciousness follows a statistical model which is devised based on the fact that buggy statements are executed mostly by failed test cases \ref{}.
+Fault localization \cite{jones2005empirical} module ranks program statements based on their suspiciousness.
+The level of suspiciousness follows a statistical model which is devised based on the fact that buggy statements are executed mostly by failed test cases \cite{naish2009spectral} \cite{xie2013theoretical}.
 Researchers have proposed different methods for code repair, each of which follows a different mindset.
-Here we provide a summarization of the advancements in these different mindset, and will discuss how having a set of predefined fix patterns can be beneficial for repair purposes.
+Here we provide a brief overview of the advancements in these different mindsets, and will discuss how having a set of predefined fix patterns can be beneficial for repair purposes.
 
-This line of research in program repair follows the competent programmer hypothesis which states that, syntactically, a faulty programs is not that far away from its correct version.
+This line of research in program repair follows the competent programmer hypothesis \cite{gopinath2014mutant} which states that, syntactically, a faulty programs is not that far away from its correct version.
 Using this concept, we can develop mutation operators, which are the operators that change an expression or a statement.
@@
 Weimer et al.
-\ref{} developed GenProg, a program repair tool that used genetic programming.
+\cite{forrest2009genetic} \cite{nguyen2009using} developed GenProg, a program repair tool that used genetic programming.
 The main idea behind this tool is that statements follow certain patterns in a codebase.
@@
 
-However, Arcuri and Briand \ref{} suspected the effectiveness of genetic programming in GenProg, as the tool mostly found the patch in the random initialization, in other words, before the evolution begins.
-Through developing TrpAutoRepair \ref{} and RSRepair \ref{}, the test case prioritization is necessary while using genetic programming, as fitness evaluation is an expensive part of the repair pipeline.
+However, Arcuri and Briand \cite{arcuri2011practical} suspected the effectiveness of genetic programming in GenProg, as the tool mostly found the patch in the random initialization, in other words, before the evolution begins.
+Through developing TrpAutoRepair \cite{qi2013efficient} and RSRepair \cite{qi2014strength}, the test case prioritization is necessary while using genetic programming, as fitness evaluation is an expensive part of the repair pipeline.
 
-Tan and Rovehoudhury \ref{} used mutation repairing for fixing regression bugs.
+Tan and Rovehoudhury \cite{tan2015relifix} used mutation repairing for fixing regression bugs.
 They categorized common code changes in real-world regressions after studying 73 program evolution benchmarks.
@@
 Pan et al.
-\ref{} analyzed seven large-scale widely used Java projects and obtained 27 common bug fix patterns.
+\cite{pan2009toward} analyzed seven large-scale widely used Java projects and obtained 27 common bug fix patterns.
 Though they did the analysis on Java projects, their reported patterns were not specific to Java.
-Martinez and Monperrus \ref{} \ref{} exploited the frequency of observed patterns to introduce a heuristic patch searching method.
+Martinez and Monperrus \cite{martinez2015mining} \cite{martinez2012mining} exploited the frequency of observed patterns to introduce a heuristic patch searching method.
 In their empirical evaluation, they concluded that the probabilistic values for repair actions can help with reducing the search space, hence creating a more effective repair tool.
@@
 Hanam et al.
-conducted a similar research but for finding Javascript most pervasive bug fix patterns.
+\cite{hanam2016discovering} conducted a similar research but for finding Javascript most pervasive bug fix patterns.
 They presented a large-scale study of the most common bug fix patterns by mining 105K commits from 134 server-side JavaScript projects.
@@
 Yang et al.
-proposed a mining approach to detect Python bug fix patterns via studying fine-grained fixing code changes.
+\cite{yang2022mining} proposed a mining approach to detect Python bug fix patterns via studying fine-grained fixing code changes.
 In their research, they also examined what portion of the bugs could be fixed using automated bug fixing approaches.
@@
 Qin et al.
-\ref{} conducted the first empirical study on real-world Rust program behaviours.
+\cite{qin2020understanding} conducted the first empirical study on real-world Rust program behaviours.
 They manually inspected 850 unsafe code usages and 17 bugs across five open-source Rust projects, five Rust libraries, two online security databases, and the standard library of Rust.
@@
 Li et al.
-\ref{} present MirChekcer, a fully automated bug detection framework for Rust.
+\cite{li2021mirchecker} present MirChekcer, a fully automated bug detection framework for Rust.
 This framework works by carrying out static analysis on Rust's Mid-level Intermediate Representation (MIR).

### 2022-10-12 sections/related.tex
move things around in 5.1

@@
+Researchers have proposed different methods for code repair, each of which follows a different mindset.
+We briefly present a number of mindsets, and will discuss how having a set of predefined fix patterns can be beneficial for repair purposes.
+
+Automated program repair aims to debug faulty source code without human involvement, sometimes using test cases to guide the repair.
+Before a repair tool can modify some code, it will typically use a fault localization module to find bug locations.
+A fault localization module ranks program statements based on their suspiciousness; one example is Tarantula by Jones and Harrold~\cite{jones2005empirical}.
+Suspiciousness is calculated using a statistical model which relies on the observation that buggy statements are executed mostly by failed test cases~\cite{naish2009spectral,xie2013theoretical}.
+
 \subsection{Search-Based Program Repair}
+One line of research in program repair follows the competent programmer hypothesis~\cite{gopinath2014mutant}: syntactically, a faulty programs is not that far away from its correct version.
+This hypothesis suggests the search-based program repair mindset.
+If the hypothesis holds, we can develop mutation operators (which change an expression or a statement) and apply them to a list of fault locations provided by the fault localization module.
+If mutation of a statement does not result in complete success after running the test cases, we assume that the statement is correct and move on to the next fault location.
 
-Automated program repair aims to debug faulty source code without human involvement.
-Prior to code modification, repair tools exploit a fault localization module to speculate the location of bug.
-Fault localization \cite{jones2005empirical} module ranks program statements based on their suspiciousness.
-The level of suspiciousness follows a statistical model which is devised based on the fact that buggy statements are executed mostly by failed test cases \cite{naish2009spectral} \cite{xie2013theoretical}.
-Researchers have proposed different methods for code repair, each of which follows a different mindset.
-Here we provide a brief overview of the advancements in these different mindsets, and will discuss how having a set of predefined fix patterns can be beneficial for repair purposes.
-
-This line of research in program repair follows the competent programmer hypothesis \cite{gopinath2014mutant} which states that, syntactically, a faulty programs is not that far away from its correct version.
-Using this concept, we can develop mutation operators, which are the operators that change an expression or a statement.
-We can apply the mutation operators on the list given by the fault localization module.
-If the mutation on a statement does not result in complete success after running the test cases, we assume that the statement is correct and move on to the next statement in the list.
-
-The computation cost depends on the number of statements and their respective mutation operations.
-Moreover, each time we mutate our program, we ought to evaluate the result by checking the program against the required specification.
-Given a sufficiently large enough search space, a complete search approach renders the repair tool virtually useless.
-That is why, the researchers have proposed heuristic searching algorithms to rectify the faulty program faster.
-
-Weimer et al.
-\cite{forrest2009genetic} \cite{nguyen2009using} developed GenProg, a program repair tool that used genetic programming.
+Weimer et al.~\cite{forrest2009genetic,nguyen2009using} developed GenProg, a program repair tool that uses genetic programming.
 The main idea behind this tool is that statements follow certain patterns in a codebase.
@@
 
-However, Arcuri and Briand \cite{arcuri2011practical} suspected the effectiveness of genetic programming in GenProg, as the tool mostly found the patch in the random initialization, in other words, before the evolution begins.
-Through developing TrpAutoRepair \cite{qi2013efficient} and RSRepair \cite{qi2014strength}, the test case prioritization is necessary while using genetic programming, as fitness evaluation is an expensive part of the repair pipeline.
+However, Arcuri and Briand~\cite{arcuri2011practical} suspected the effectiveness of genetic programming in GenProg, as the tool mostly found the patch in its random initialization phase, in other words, before the evolution begins.
+Through developing TrpAutoRepair~\cite{qi2013efficient} and RSRepair~\cite{qi2014strength}, the test case prioritization is necessary while using genetic programming, as fitness evaluation is an expensive part of the repair pipeline.
 
-Tan and Rovehoudhury \cite{tan2015relifix} used mutation repairing for fixing regression bugs.
+Tan and Rovehoudhury~\cite{tan2015relifix} used mutation repairing for fixing regression bugs.
 They categorized common code changes in real-world regressions after studying 73 program evolution benchmarks.
 They used this categorization to design mutation operators, and chose them based on the faulty location’s to repair the program.
+
+The computational cost of search-based program repair depends on the number of statements and their respective mutation operations.
+Moreover, each time we mutate our program, we ought to evaluate the result by checking the program against the required specification.
+Given a sufficiently large enough search space, a complete search approach renders the repair tool virtually useless.
+That is why, the researchers have proposed heuristic searching algorithms to rectify the faulty program faster.
 

### 2022-10-12 sections/related.tex
instructions in related work

@@
 This tool showed promising results as it managed to find patches without any additional annotations or human involvements.
+% the sentence starting with "therefore" lacks a verb
 
@@
 
+% replace this next paragraph with some text about how search based program repair depends on patterns, and we have patterns that we have found to be used in actual repairs.
+That should go at the top of 5.1.
+
+% say why you're talking about GenProg etc.
+
+% fix the citations
+
 The computational cost of search-based program repair depends on the number of statements and their respective mutation operations.
@@
 
+% say why pattern-based repair is different from search-based repair
 The main intuition in pattern-based program repair methods is that the bug fixing modifications of the programs follow certain patterns.
@@
 Hanam et al.
-\cite{hanam2016discovering} conducted a similar research but for finding Javascript most pervasive bug fix patterns.
+\cite{hanam2016discovering} conducted similar research, but for finding JavaScript most pervasive bug fix patterns.
 They presented a large-scale study of the most common bug fix patterns by mining 105K commits from 134 server-side JavaScript projects.
-They used DBSCAN clustering algorithm and divided bug fixing change types into 219 clusters, from which they extracted 13 pervasive cross-project bug fix patterns.
+They used the DBSCAN clustering algorithm and divided bug fixing change types into 219 clusters, from which they extracted 13 pervasive cross-project bug fix patterns.
 
@@
 
+% again, say how we are like or unlike these works
+
 \subsection{Code Patterns in Rust}
+
+% say how is this related to our work?
 

### 2022-10-12 sections/results.tex
edits to results

@@
 
-As discussed in the previous section, we used an encoding mechanism to incorporate AST information in fix-sized datapoints which then could be used in the DBSCAN clustering algorithm.
+As discussed in the previous section, we used an encoding mechanism to incorporate AST information in fixed-sized datapoints which then could be used in the DBSCAN clustering algorithm.
 We used heuristics (weighting scheme) while encoding our datapoints to make the most important program elements stand out between all the non terminals seen in the AST (alongside some BC-related elements).
@@
 
-The general idea for our evaluation is whether the most important program elements has the high values in our encoding.
+The general idea for our evaluation is whether the most important program elements get the high values in our encoding.
 The analyzer compares the actual change in the code and compares it to the visual representation of its respective data point.
 We used circle pack figures to visualize each data point \cite{collins2003circle}.
-For this purpose, we used packcircles python library\footnote{\url{https://github.com/mhtchan/packcircles}}.
-In these figures, the radius of the circle of each non terminal shows its importance in our encoding (Similar to Figure~\ref{fig:essence}).
+For this purpose, we used the packcircles python library\footnote{\url{https://github.com/mhtchan/packcircles}}.
+In these figures, the radius of the circle of each non terminal shows its importance in our encoding (similar to Figure~\ref{fig:essence}).
 
@@
 The analyzer should give a binary answer to these questions.
-Both authors carry out these experiments independently.
+Both authors carried out these experiments independently.
 At last, we report the results as the portion of the changes which both authors agreed upon its encoding correctness.
@@
 
+% no, you have to compute the intersection
+% when is our approach effective and when is it not?
+% more detailed stats would probably be good too
+
 \subsection{\label{sec:common_patterns}Common Bug Fix Patterns (RQ2)}
+
+{\bf There is no text in this subsection.}
 
@@
 
-Attributes in Rust make it possible for the developers to add metadata to program elements.
-These attributes can be added to structs, struct fields, enum variants, match expression arms, etc.
+Attributes in Rust enable developers to add metadata to program elements.
+Developers can add attributes to structs, struct fields, enum variants, match expression arms, etc.
 There are two types of attributes: Inner Attributes and Outer Attributes.
@@
 \noindent \textbf{Description:} In this pattern, the attribute set of a struct changes.
-This modification can happen in three different ways; (G.attr.struct.add) Adding attributes to the struct, (G.attr.struct.drop) Removing attributes from the struct, (G.attr.struct.change) Changing the content of an existing attribute.
-The attributes are Outer Attributes as we can find struct related non-terminals within the ASTDiff of the instances of this pattern.
+This modification can happen in three different ways: (G.attr.struct.add) adding attributes to the struct; (G.attr.struct.drop) removing attributes from the struct; and (G.attr.struct.change) changing the content of an existing attribute.
+These attributes are Outer Attributes as we can find struct related non-terminals within the ASTDiff of the instances of this pattern.
 \\
@@
 Using attributes for struct fields, apart from being able to store different meta-information for different struct fields, for some use cases, the programmers can control the action they want to apply on different fields.
+% this last sentence immediately prior to here has too many moving parts
 
@@
 
-\noindent \textbf{Description:} A set of new fields are add to or removed from an existing struct.
-The addition of a new field happens when the developer need to store a new piece of data so they can implement a desired behaviour.
-The removal of a field  happens when the developer realizes that the field is expendable with respect to the behaviour they want to implement.
+\noindent \textbf{Description:} A set of new fields are added to or removed from an existing struct.
+A developer adds a new field to store a new piece of data in a struct.
+A developer removes a field when the developer realizes that the field is not required with respect to the behaviour they want to implement (possibly because they have moved it elsewhere).
 \\
@@
 
-\noindent \textbf{Description:} Adding pub to a struct field makes it possible to access that field within external modules.
-In the instances of this cluster, adding pub access modifier happened in a bug fixing change, where the developer needed to access a field which had not been granted public access.
-On the other hand, the public access might be revoked if the developer realizes there is no external access to the field.
+\noindent \textbf{Description:} Adding \texttt{pub} to a struct field makes it possible to access that field from external modules.
+In the instances of this cluster, adding the \texttt{pub} access modifier happened in a bug fixing change, where the developer needed to access a field which had not been marked for public access.
+On the other hand, the public access might be revoked if the developer realizes there are no external accesses to the field, and that such accesses are not desirable, e.g.
+for encapsulation purposes.
 
@@
 
-\noindent \textbf{Description:} Rust is a null safe language, meaning that object references cannot assume null values.
-If we want void values for our object, we would have to use Option type.
-Option is essentially an enum with two variants, Some which indicates that our object has an value, and None shows the absence of any value.
-In this pattern, the developer changes the type of a struct field to an Option of its type.
-This makes it possible for the developer to modify the program in a way that it covers the scenarios in which the the struct field has no value.
+\noindent \textbf{Description:} Rust is a null-safe language, meaning that object references cannot take on null values.
+If we want a variable to contain null, we would have to use the \texttt{Option} type.
+\texttt{Option} is essentially an \texttt{enum} with two variants, \texttt{Some}, which indicates that our variable has a value, and \texttt{None}, which indicates the absence of any value.
+In this change pattern, the developer changes the type of a struct field to an \texttt{Option} of that type, hence enabling them to store no value in that field.
 
-For instance, in the Rust project swc-project/swc (A fast TypeScript/JavaScript compiler), the developer needed to modify the CSS parser source code in a way that it accounts for empty @media queries \footnote{\url{https://github.com/swc-project/swc/commit/75a14f98b7370226115ee24eec6eb8c802bd4837}}, so, as shown in the code snippet above, they changed type MediaQueryList to Option\verb+<+MediaQueryList\verb+>+, and modified the other parts of the source code, accordingly.
+For instance, in the Rust project swc-project/swc (a fast TypeScript/JavaScript compiler), the developer needed to modify the CSS parser source code in a way that it accounts for empty @media queries \footnote{\url{https://github.com/swc-project/swc/commit/75a14f98b7370226115ee24eec6eb8c802bd4837}}.
+So, as shown in the code snippet above, they changed type \texttt{MediaQueryList} to \verb+Option<MediaQueryList>+, and modified the other parts of the source code, accordingly.
 
@@
 This is a common change pattern in all statically-typed languages and can happen for various reasons.
-These reasons can be a step for implementing a new feature or rectifying the behaviour of the program.
+For instance, the developer may be implementing a new feature or fixing the behaviour of the program.
 Also it can happen for refactoring purposes or performance enhancement.
 \\
+% maybe discuss the specific example a bit?
 
-\noindent \textit{ \textbf{G.type.generic} Changing a generic type}
+\noindent \textit{ \textbf{G.type.generic} Changing a generic type parameter}
 
@@
 
-\noindent \textbf{Description:} This a change of the generic (parametric) type of a struct field. Similar to \textit{\textbf{G.type.field}}, this change can also happen for the mentioned reasons. \\
+\noindent \textbf{Description:} This a change of the type parameter of a struct field. As with \textit{\textbf{G.type.field}}, this change can also happen for the mentioned reasons. \\
 
@@
 
-\noindent \textbf{Description:} In Rust, we can specify the type of the value that an enum variant can store.
-Using this feature, we can avoid using a struct alongside our enum variants to specify the value type.
+\noindent \textbf{Description:} In Rust, developers can specify the type of the value that an enum variant can store.
+Using this feature, they can avoid having to use a struct along with enum variants.
 This change pattern is associated with a change in enum variant types.
@@
 \noindent \textbf{Description:} In this pattern, an extra type is added to an existing type, making it a tuple of multiple types.
-Alternatively, the change might be dropping types from the tuple and making it a tuple with fewer number of arguments.
+Alternatively, the change might be dropping types from the tuple and making it a smaller-arity tuple.
 
@@
 \noindent \textbf{Description:} The trait bounds are the functionalities that we require from our parametric types.
-This change pattern related to modification in trait bounds.
+This change pattern relates to modification in trait bounds.
 This can happen both for bug fix reasons or just refactoring purposes.

### 2022-10-12 sections/related.tex
fixed related.tex

--- 2022-10-12 sections/related.tex
@@
-Researchers have proposed different methods for code repair, each of which follows a different mindset.
-We briefly present a number of mindsets, and will discuss how having a set of predefined fix patterns can be beneficial for repair purposes.
-
 Automated program repair aims to debug faulty source code without human involvement, sometimes using test cases to guide the repair.
@@
 Suspiciousness is calculated using a statistical model which relies on the observation that buggy statements are executed mostly by failed test cases~\cite{naish2009spectral,xie2013theoretical}.
+
+Researchers have proposed different methods for code repair, each of which follows a different mindset.
+Following the taxonomy propsed by~\cite{liu2018survey}, we briefly present a number of mindsets, and will discuss how having a set of predefined fix patterns can be beneficial for repair purposes.
 
@@
 If mutation of a statement does not result in complete success after running the test cases, we assume that the statement is correct and move on to the next fault location.
+Knowing bug producing patterns help the researchers to design more creative mutation operators that leverage domain specific insights.
 
-Weimer et al.~\cite{forrest2009genetic,nguyen2009using} developed GenProg, a program repair tool that uses genetic programming.
+Use of mutation operators in program repair became popular when Weimer et al.~\cite{forrest2009genetic,nguyen2009using} developed GenProg, a program repair tool that uses genetic programming.
 The main idea behind this tool is that statements follow certain patterns in a codebase.
-Therefore, if we can find the correct version of a faulty statement somewhere in the program.
+Therefore, if we find the correct version of a faulty statement somewhere in the program, we can use it to fix our program.
 This tool showed promising results as it managed to find patches without any additional annotations or human involvements.
@@
 
-The computational cost of search-based program repair depends on the number of statements and their respective mutation operations.
-Moreover, each time we mutate our program, we ought to evaluate the result by checking the program against the required specification.
-Given a sufficiently large enough search space, a complete search approach renders the repair tool virtually useless.
-That is why, the researchers have proposed heuristic searching algorithms to rectify the faulty program faster.
+\subsection{Pattern-based Program Repair and Mining Bug Fix Patterns}
 
-\subsection{Mining Bug Fix Patterns and Pattern-based Program Repair}
-
+% say why pattern-based repair is different from search-based repair -- isn't it sufficient to refer the reader to the survey?
 The main intuition in pattern-based program repair methods is that the bug fixing modifications of the programs follow certain patterns.
 Thus, prior study of the common bug fix patterns is a crucial step for developing these repair tools.
-Pan et al.
-\cite{pan2009toward} analyzed seven large-scale widely used Java projects and obtained 27 common bug fix patterns.
+Pan et al.~\cite{pan2009toward} analyzed seven large-scale widely used Java projects and obtained 27 common bug fix patterns.
 Though they did the analysis on Java projects, their reported patterns were not specific to Java.
-Martinez and Monperrus \cite{martinez2015mining} \cite{martinez2012mining} exploited the frequency of observed patterns to introduce a heuristic patch searching method.
+Martinez and Monperrus~\cite{martinez2015mining, martinez2012mining} exploited the frequency of observed patterns to introduce a heuristic patch searching method.
 In their empirical evaluation, they concluded that the probabilistic values for repair actions can help with reducing the search space, hence creating a more effective repair tool.
 
-Hanam et al.
-\cite{hanam2016discovering} conducted a similar research but for finding Javascript most pervasive bug fix patterns.
+Hanam et al.~\cite{hanam2016discovering} conducted similar research, but for finding JavaScript most pervasive bug fix patterns.
 They presented a large-scale study of the most common bug fix patterns by mining 105K commits from 134 server-side JavaScript projects.
-They used DBSCAN clustering algorithm and divided bug fixing change types into 219 clusters, from which they extracted 13 pervasive cross-project bug fix patterns.
+They used the DBSCAN clustering algorithm and divided bug fixing change types into 219 clusters, from which they extracted 13 pervasive cross-project bug fix patterns.
 
-Yang et al.
-\cite{yang2022mining} proposed a mining approach to detect Python bug fix patterns via studying fine-grained fixing code changes.
+Yang et al.~\cite{yang2022mining} proposed a mining approach to detect Python bug fix patterns via studying fine-grained fixing code changes.
 In their research, they also examined what portion of the bugs could be fixed using automated bug fixing approaches.
@@
 
+We would characterize our work as the Rust version of what was done in~\cite{hanam2016discovering} and~\cite{yang2022mining}.
+Similar to~\cite{hanam2016discovering} we used DBSCAN to carry out the clustering, however, our novel program encoding approach differs from what they did.
+Also, analogous to~\cite{yang2022mining} we introduced patterns in two different categories, general and language specific (BC-related patterns).
+
 \subsection{Code Patterns in Rust}
 
-Qin et al.
-\cite{qin2020understanding} conducted the first empirical study on real-world Rust program behaviours.
+There have been some studies on common bug patterns in Rust, either for gaining insights from Rust programs or enhancing a tool's functionality.
+Nevertheless, to the best of our knowledge, our work is the first study that uses an automated mining and code analysis pipeline to find pervasive patterns in Rust open source projects.
+
+Qin et al.~\cite{qin2020understanding} conducted the first empirical study on real-world Rust program behaviours.
 They manually inspected 850 unsafe code usages and 17 bugs across five open-source Rust projects, five Rust libraries, two online security databases, and the standard library of Rust.
@@
 
-Li et al.
-\cite{li2021mirchecker} present MirChekcer, a fully automated bug detection framework for Rust.
+Li et al.~\cite{li2021mirchecker} present MirChekcer, a fully automated bug detection framework for Rust.
 This framework works by carrying out static analysis on Rust's Mid-level Intermediate Representation (MIR).

### 2022-10-13 main.tex
changed results a bit and added related works sec
citation added for intro
instructions in related work
edits to results
figs tikzisized

@@
 \usepackage[utf8]{inputenc}
+\usepackage{tikz}
+\usetikzlibrary{graphs} 
+\usetikzlibrary{matrix}
 \usepackage{multirow}
+% \usepackage[backend=bibtex, style=authoryear]{biblatex}
+% \addbibresource{bibtex.bib}
 \usepackage{subfiles} % Best loaded last in the preamble
 \input{defs}
+
+\tikzset{ 
+    table/.style={
+        matrix of nodes,
+        row sep=-\pgflinewidth,
+        column sep=-\pgflinewidth,
+        nodes={
+            rectangle,
+            draw=black,
+            align=center
+        },
+        minimum height=1.5em,
+        text depth=0.5ex,
+        text height=2ex,
+        nodes in empty cells,
+        column 1/.style={
+            nodes={text width=5em,font=\bfseries, fill=gray!20}
+        },
+        column 2/.style={
+            nodes={text width=4em,font=\bfseries}
+        },
+        column 3/.style={
+            nodes={text width=5em,font=\bfseries}
+        },
+        column 4/.style={
+            nodes={text width=5em,font=\bfseries}
+        },  
+        column 5/.style={
+            nodes={text width=8em,font=\bfseries}
+        },  
+        column 6/.style={
+            nodes={text width=5em,font=\bfseries}
+        },  
+        column 7/.style={
+            nodes={text width=5em,font=\bfseries}
+        },
+        column 8/.style={
+            nodes={text width=9em,font=\bfseries}
+        },
+        column 9/.style={
+            nodes={text width=3em,font=\bfseries}
+        }
+    }
+}
 
@@
 \section{Introduction}
+% rubber: depend sections/introduction.tex
 \subfile{sections/introduction}
@@
 \section{Methodology}
+% rubber: depend sections/methodology.tex
 \subfile{sections/methodology}
@@
 \section{Results}
+% rubber: depend sections/results.tex
 \subfile{sections/results}
@@
 \section{Discussion}
+% rubber: depend sections/discussion.tex
 \subfile{sections/discussion}
 
+\section{Related Work}
+% rubber: depend sections/related.tex
+\subfile{sections/related}
+
 \section{Conclusion}
+% rubber: depend sections/conclusion.tex
 \subfile{sections/conclusion}
@@
 % BibTeX users please use one of
-%\bibliographystyle{spbasic}      % basic style, author-year citations
-%\bibliographystyle{spmpsci}      % mathematics and physical sciences
-\bibliographystyle{spphys}       % APS-like style for physics
-%\bibliography{}   % name your BibTeX data base
+\bibliographystyle{plain}      % basic style, author-year citations
+% \bibliographystyle{spmpsci}      % mathematics and physical sciences
+% \bibliographystyle{spphys}       % APS-like style for physics
+\bibliography{refs}       % name your BibTeX data base
+% \printbibliography[title={References}]  % name your BibTeX data base
 
 % Non-BibTeX users please use
-\begin{thebibliography}{}
-%
-% and use \bibitem to create references. Consult the Instructions
-% for authors for reference list style.
-%
-\bibitem{RefJ}
-% Format for Journal Reference
-Author, Article title, Journal, Volume, page numbers (year)
-% Format for books
-\bibitem{RefB}
-Author, Book title, page numbers. Publisher, place (year)
-% etc
-\end{thebibliography}
+% \begin{thebibliography}{}
+% %
+% % and use \bibitem to create references. Consult the Instructions
+% % for authors for reference list style.
+% %
+% \bibitem{RefJ}
+% % Format for Journal Reference
+% Author, Article title, Journal, Volume, page numbers (year)
+% % Format for books
+% \bibitem{RefB}
+% Author, Book title, page numbers. Publisher, place (year)
+% % etc
+% \end{thebibliography}
 

### 2022-10-13 sections/methodology.tex
figs tikzisized

@@
 \centering
-\includegraphics[width=0.8\textwidth]{figs/extraction.drawio.pdf}
+\begin{tikzpicture}[scale=0.7, transform shape,every node/.style={draw},minimum height=0.7cm]
+
+    %% Left AST
+
+    \draw (-1.5,1.5) node[shape=rectangle, minimum width=3cm,minimum height=5.4cm, rounded corners=1ex, xshift=1.5cm, yshift=-2.5cm](context){};
+
+    \node[draw=none] at (0,1) {Context};
+
+    \draw (0,0) node[](a1){ItemFn};
+    \draw (0,-1) node[](a2){Local};
+    \draw (0,-2) node[](a3){ExprCall};
+    \draw (0,-3) node[](a4){args};
+
+    \draw (-2,-5) node[shape=circle](null){null};
+
+    \draw (0,-5) node[](b1){ExprReference};
+    \draw (0,-6) node[](b2){Path};
+    \draw (0,-7) node[](b3){Ident};
+    \draw (0,-8) node[](b4){tail};
+
+    \draw (3,-5) node[](c1){ExprMethodCall};
+    \draw (3,-6) node[](c2){Path};
+    \draw (3,-7) node[](c3){Ident};
+    \draw (3,-8) node[](c4){head};
+
+    \draw (5,-6.5) node[](d1){Ident};
+    \draw (5,-7.5) node[](d2){clone};
+
+    \draw[->]  (a1) -- (a2);
+    \draw[->]  (a2) -- (a3);
+    \draw[->]  (a3) -- (a4);
+
+    \draw[->]  (context) -- (null);
+    \draw[->]  (context) -- (b1);
+    \draw[->]  (context) -- (c1);
+
+    \draw[->]  (b1) -- (b2);
+    \draw[->]  (b2) -- (b3);
+    \draw[->]  (b3) -- (b4);
+
+    \draw[->]  (c1) -- (c2);
+    \draw[->]  (c2) -- (c3);
+    \draw[->]  (c3) -- (c4);
+
+    \draw[->]  (c1) -- (d1);
+
+    \draw [->, red] plot [smooth, tension=1] coordinates { (2,1) (2,-3) (1.4,-5) (1.4,-8.5)};
+
+    \draw [->, green] plot [smooth, tension=1] coordinates { (2.5,1) (3,-3) (5,-5) (4,-6.5) (3.7,-8.5)};
+
+    \draw [->, purple] plot [smooth, tension=1] coordinates { (3,1) (3.5,-3) (5.5,-5) (5.5,-8.5)};
+
+    %% Arrow
+
+    \node[draw=none] at (5.1,-2.7) {Path Extraction};
+    \draw [-stealth,double,line width=2,color=blue!40](4.1,-3.2) -- (6.3,-3.2);
+
+    %% Code at top right
+
+    \node (codeA) [fill=yellow!20, rounded corners=1ex] at (10, 0) {%
+    \centering
+    \begin{minipage}{10cm}
+    \centering
+    \begin{lstlisting}[language=Rust, style=colouredRust]
+      fn sample() {
+        // ...
+        -- let receiver = foo(arg1, tail, head)
+        ++ let receiver = foo(arg1, &tail, head.clone())
+        // ...
+      }
+    \end{lstlisting}
+
+    \end{minipage}%
+    };
+
+    %% Paths at right
+
+    \draw (7.5,-2.5) node[fill=red!30](e1){ItemFn};
+    \draw (7.5,-3.5) node[fill=red!30](e2){Local};
+    \draw (7.5,-4.5) node[fill=red!30](e3){ExprCall};
+    \draw (7.5,-5.5) node[fill=red!30](e4){ExprReference};
+    \draw (7.5,-6.5) node[fill=red!30](e5){Path};
+    \draw (7.5,-7.5) node[fill=red!30](e6){Ident};
+
+    \node[draw=none, color=red] at (7.5,-8.3) {Path 1};
+
+    \draw (10.5,-2.5) node[fill=green!20](f1){ItemFn};
+    \draw (10.5,-3.5) node[fill=green!20](f2){Local};
+    \draw (10.5,-4.5) node[fill=green!20](f3){ExprCall};
+    \draw (10.5,-5.5) node[fill=green!20](f4){ExprMethodCall};
+    \draw (10.5,-6.5) node[fill=green!20](f5){Path};
+    \draw (10.5,-7.5) node[fill=green!20](f6){Ident};
+
+    \node[draw=none, color=green] at (10.5,-8.3) {Path 2};
+    \node[draw=none, color=green] at (10.5,-8.3) {Path 2};
+
+    \draw (13.5,-2.5) node[fill=purple!20](g1){ItemFn};
+    \draw (13.5,-3.5) node[fill=purple!20](g2){Local};
+    \draw (13.5,-4.5) node[fill=purple!20](g3){ExprCall};
+    \draw (13.5,-5.5) node[fill=purple!20](g4){ExprMethodCall};
+    \draw (13.5,-6.5) node[fill=purple!20](g5){Path};
+    \draw (13.5,-7.5) node[fill=purple!20](g6){clone};
+    \node[draw=none, color=purple] at (13.5,-8.3) {Path 3};
+
+    \draw[->]  (e1) -- (e2);
+    \draw[->]  (e2) -- (e3);
+    \draw[->]  (e3) -- (e4);
+    \draw[->]  (e4) -- (e5);
+    \draw[->]  (e5) -- (e6);
+
+    \draw[->]  (f1) -- (f2);
+    \draw[->]  (f2) -- (f3);
+    \draw[->]  (f3) -- (f4);
+    \draw[->]  (f4) -- (f5);
+    \draw[->]  (f5) -- (f6);
+
+    \draw[->]  (g1) -- (g2);
+    \draw[->]  (g2) -- (g3);
+    \draw[->]  (g3) -- (g4);
+    \draw[->]  (g4) -- (g5);
+    \draw[->]  (g5) -- (g6);
+
+    \end{tikzpicture}
+
     \caption{\label{fig:extraction}Extracting paths from an ASTDiff.
@@
 A DFS search will identify three paths, shown in three different colors at right.}
+
 \end{figure}
@@
     \centering
-    \includegraphics[width=1\textwidth]{figs/essence.drawio.pdf}
+    \begin{tikzpicture}[scale=0.7, transform shape, every node/.style={draw},minimum height=0.7cm]
+
+        %% Left AST
+
+        \draw (-1.5,1.5) node[shape=rectangle, minimum width=3cm,minimum height=5.4cm, rounded corners=1ex, xshift=1.5cm, yshift=-2.5cm](context){};
+
+        \node[draw=none] at (0,1) {Context};
+
+        \draw (0,0) node[](a1){ItemFn};
+        \draw (0,-1) node[](a2){Local};
+        \draw (0,-2) node[](a3){ExprCall};
+        \draw (0,-3) node[](a4){args};
+
+        \draw (-2,-5) node[shape=circle](null){null};
+
+        \draw (0,-5) node[](b1){ExprReference};
+        \draw (0,-6) node[](b2){Path};
+        \draw (0,-7) node[](b3){Ident};
+        \draw (0,-8) node[](b4){tail};
+
+        \draw (3,-5) node[](c1){ExprMethodCall};
+        \draw (3,-6) node[](c2){Path};
+        \draw (3,-7) node[](c3){Ident};
+        \draw (3,-8) node[](c4){head};
+
+        \draw (5,-6.5) node[](d1){Ident};
+        \draw (5,-7.5) node[](d2){clone};
+
+        \draw[->]  (a1) -- (a2);
+        \draw[->]  (a2) -- (a3);
+        \draw[->]  (a3) -- (a4);
+
+        \draw[->]  (context) -- (null);
+        \draw[->]  (context) -- (b1);
+        \draw[->]  (context) -- (c1);
+
+        \draw[->]  (b1) -- (b2);
+        \draw[->]  (b2) -- (b3);
+        \draw[->]  (b3) -- (b4);
+
+        \draw[->]  (c1) -- (c2);
+        \draw[->]  (c2) -- (c3);
+        \draw[->]  (c3) -- (c4);
+
+        \draw[->]  (c1) -- (d1);
+
+        %% Tables
+
+        \matrix (table1) [table,text width=6em,nodes={scale=0.45,transform shape}] at (8.5,1)
+        {
+        Element & ItemFn  & Local & ExprCall & ExprReference & Path & Ident & ExprMethodCall & clone\\
+        \# & 1 & 1 & 1 & 1 & 2 & 3 & 1 & 1 \\
+        };
+
+        \node[draw=none] at (8.2,0) {Multiply};
+        \draw [-stealth,double,line width=1,color=blue!40](7,0.4) -- (7,-0.4);
+
+        \matrix (table2) [table,text width=6em,nodes={scale=0.45,transform shape}] at (8.5,-1)
+        {
+        Element & ItemFn  & Local & ExprCall & ExprReference & Path & Ident & ExprMethodCall & clone\\
+        weight & 0.0004 & 2.21e-05 & 8.68e-05 & 1 & 2.60e-05 & 2.76e-05 & 4.75e-05 & 1 \\
+        };
+
+        \node[draw=none] (circlepack) at (10.5,-6.2)
+            {\includegraphics[width=0.7\textwidth]{figs/circlepack.png}};
+
+        \node[draw=none] at (9,-2.3) {Visualizing Essence};
+        \draw [-stealth,double,line width=1,color=blue!40](7,-1.6) -- (7.5,-3.4);
+
+        \end{tikzpicture}
+
     \caption{\label{fig:essence}Multiplying the number of occurrences of items observed in the tree by their respective weights, resulting in the essence of change (visualized with a circle pack figure at right).

### 2022-10-13 sections/related.tex
remove space in cite

@@
 Though they did the analysis on Java projects, their reported patterns were not specific to Java.
-Martinez and Monperrus~\cite{martinez2015mining, martinez2012mining} exploited the frequency of observed patterns to introduce a heuristic patch searching method.
+Martinez and Monperrus~\cite{martinez2015mining,martinez2012mining} exploited the frequency of observed patterns to introduce a heuristic patch searching method.
 In their empirical evaluation, they concluded that the probabilistic values for repair actions can help with reducing the search space, hence creating a more effective repair tool.

### 2022-10-13 sections/related.tex
5.1

@@
 Automated program repair aims to debug faulty source code without human involvement, sometimes using test cases to guide the repair.
-Before a repair tool can modify some code, it will typically use a fault localization module to find bug locations.
-A fault localization module ranks program statements based on their suspiciousness; one example is Tarantula by Jones and Harrold~\cite{jones2005empirical}.
+Before a repair tool modifies code, it will typically use a fault localization module to find fault locations.
+A fault localization module ranks program statements based on suspiciousness; one example is Tarantula by Jones and Harrold~\cite{jones2005empirical}.
 Suspiciousness is calculated using a statistical model which relies on the observation that buggy statements are executed mostly by failed test cases~\cite{naish2009spectral,xie2013theoretical}.
 
-Researchers have proposed different methods for code repair, each of which follows a different mindset.
+Researchers have proposed different methods for code repair, each of which can be said to follow a different mindset.
 Following the taxonomy propsed by~\cite{liu2018survey}, we briefly present a number of mindsets, and will discuss how having a set of predefined fix patterns can be beneficial for repair purposes.
@@
 \subsection{Search-Based Program Repair}
-One line of research in program repair follows the competent programmer hypothesis~\cite{gopinath2014mutant}: syntactically, a faulty programs is not that far away from its correct version.
+One line of research in program repair follows the competent programmer hypothesis~\cite{gopinath2014mutant}: syntactically, a faulty program is not that far away from its correct version.
 This hypothesis suggests the search-based program repair mindset.
 If the hypothesis holds, we can develop mutation operators (which change an expression or a statement) and apply them to a list of fault locations provided by the fault localization module.
-If mutation of a statement does not result in complete success after running the test cases, we assume that the statement is correct and move on to the next fault location.
-Knowing bug producing patterns help the researchers to design more creative mutation operators that leverage domain specific insights.
+So, we loop through possible mutations; if the mutation of a statement does not cause all expected-successful test cases to succeed, we assume that the statement is correct and move on to the next fault location.
+Having a set of bug-producing patterns, like the ones we are proposing, will help researchers design targetted mutation operators that leverage domain specific insights.
 
@@
 The main idea behind this tool is that statements follow certain patterns in a codebase.
-Therefore, if we find the correct version of a faulty statement somewhere in the program, we can use it to fix our program.
+Therefore, if we find the correct version of a faulty statement somewhere in the program, we can use that version in place of the faulty original.
 This tool showed promising results as it managed to find patches without any additional annotations or human involvements.
 
-However, Arcuri and Briand~\cite{arcuri2011practical} suspected the effectiveness of genetic programming in GenProg, as the tool mostly found the patch in its random initialization phase, in other words, before the evolution begins.
-Through developing TrpAutoRepair~\cite{qi2013efficient} and RSRepair~\cite{qi2014strength}, the test case prioritization is necessary while using genetic programming, as fitness evaluation is an expensive part of the repair pipeline.
+However, Arcuri and Briand~\cite{arcuri2011practical} observed that GenProg mostly found the patch while carrying out random initialization---that is, before GenProg's evolution begins.
+Their tools TrpAutoRepair~\cite{qi2013efficient} and RSRepair~\cite{qi2014strength} advocate that test case prioritization is necessary while using genetic programming, as fitness evaluation is an expensive part of the repair pipeline.
 
 Tan and Rovehoudhury~\cite{tan2015relifix} used mutation repairing for fixing regression bugs.
-They categorized common code changes in real-world regressions after studying 73 program evolution benchmarks.
-They used this categorization to design mutation operators, and chose them based on the faulty location’s to repair the program.
+They categorized common code changes in real-world regressions after studying 73 program evolution benchmarks; our approach also yields a set of common code changes, for Rust, using a different set of changes.
+Tan and Rovehoudhury used their categorization to design mutation operators, and chose them based on the faulty locations to repair the program.
 

### 2022-10-13 sections/related.tex
related work

@@
 Researchers have proposed different methods for code repair, each of which can be said to follow a different mindset.
-Following the taxonomy propsed by~\cite{liu2018survey}, we briefly present a number of mindsets, and will discuss how having a set of predefined fix patterns can be beneficial for repair purposes.
+Following the taxonomy propsed by~\cite{liu2018survey}, we briefly present two mindsets, and will discuss how having a set of predefined fix patterns can aid repair.
 
@@
 They categorized common code changes in real-world regressions after studying 73 program evolution benchmarks; our approach also yields a set of common code changes, for Rust, using a different set of changes.
-Tan and Rovehoudhury used their categorization to design mutation operators, and chose them based on the faulty locations to repair the program.
+Tan and Rovehoudhury used their categorization to design mutation operators, and drew from those operators to repair faulty program locations.
 
@@
 % say why pattern-based repair is different from search-based repair -- isn't it sufficient to refer the reader to the survey?
-The main intuition in pattern-based program repair methods is that the bug fixing modifications of the programs follow certain patterns.
-Thus, prior study of the common bug fix patterns is a crucial step for developing these repair tools.
+% Pan et al: was it manual? or not? summarize their technique and say whether what they found is similar to what we found.
+The main intuition in pattern-based program repair methods is that bug fixes follow certain patterns.
+Thus, having a collection of common bug fix patterns is a crucial step for developing such repair tools.
 Pan et al.~\cite{pan2009toward} analyzed seven large-scale widely used Java projects and obtained 27 common bug fix patterns.
@@
 Martinez and Monperrus~\cite{martinez2015mining,martinez2012mining} exploited the frequency of observed patterns to introduce a heuristic patch searching method.
-In their empirical evaluation, they concluded that the probabilistic values for repair actions can help with reducing the search space, hence creating a more effective repair tool.
+In their empirical evaluation, they concluded that choosing repair actions probabilistically (weighted by frequency) can help with reducing the search space, hence creating a more effective repair tool.
 
-Hanam et al.~\cite{hanam2016discovering} conducted similar research, but for finding JavaScript most pervasive bug fix patterns.
-They presented a large-scale study of the most common bug fix patterns by mining 105K commits from 134 server-side JavaScript projects.
-They used the DBSCAN clustering algorithm and divided bug fixing change types into 219 clusters, from which they extracted 13 pervasive cross-project bug fix patterns.
+Hanam et al.~\cite{hanam2016discovering} conducted similar research, but for finding pervasive bug fix patterns in JavaScript.
+They performed a large-scale study of bug fix patterns by mining 105K commits from 134 server-side JavaScript projects.
+Like us, they used the DBSCAN clustering algorithm and divided bug fixing change types into 219 clusters, from which they extracted 13 pervasive cross-project bug fix patterns.
 
-Yang et al.~\cite{yang2022mining} proposed a mining approach to detect Python bug fix patterns via studying fine-grained fixing code changes.
-In their research, they also examined what portion of the bugs could be fixed using automated bug fixing approaches.
-Moreover, they evaluated the fix patterns in the wild and realized that 37\% of the buggy codes could be matched by the fix patterns they had found.
+Yang et al.~\cite{yang2022mining} proposed a mining approach to detect Python bug fix patterns by studying fine-grained fixing code changes.
+They also examined how many bugs could be fixed using automated bug fixing approaches.
+Moreover, they evaluated the fix patterns that they detected and concluded that 37\% of the buggy codes could be matched by the fix patterns they had found.
 
 We would characterize our work as the Rust version of what was done in~\cite{hanam2016discovering} and~\cite{yang2022mining}.
-Similar to~\cite{hanam2016discovering} we used DBSCAN to carry out the clustering, however, our novel program encoding approach differs from what they did.
-Also, analogous to~\cite{yang2022mining} we introduced patterns in two different categories, general and language specific (BC-related patterns).
+Like~\cite{hanam2016discovering}, we used DBSCAN to cluster fixes.
+However, our novel program encoding approach differs from what they did.
+Also, like~\cite{yang2022mining}, we introduced patterns in two different categories.
+In our case there were general and language-specific patterns; for Rust, the language-specific patterns were related to the borrow checker (BC patterns).
+% ok they didn't do our program encoding; what did they do instead?
 
@@
 
-There have been some studies on common bug patterns in Rust, either for gaining insights from Rust programs or enhancing a tool's functionality.
-Nevertheless, to the best of our knowledge, our work is the first study that uses an automated mining and code analysis pipeline to find pervasive patterns in Rust open source projects.
+To our knowledge, our work is the first to use an automated mining and code analysis pipeline to find pervasive patterns in Rust open source projects, even if previous studies have investigated common bug patterns in Rust.
 
 Qin et al.~\cite{qin2020understanding} conducted the first empirical study on real-world Rust program behaviours.
-They manually inspected 850 unsafe code usages and 17 bugs across five open-source Rust projects, five Rust libraries, two online security databases, and the standard library of Rust.
-They analyzed the motivation behind unsafe code usage and removal, in addition to obtaining 70 memory-safety issues and 100 concurrency bugs.
+They manually inspected 850 unsafe code usages and 17 bugs across five open-source Rust projects, five Rust libraries, two online security databases, and the Rust standard library.
+They analyzed the motivation behind unsafe code usage and removal, in addition to recording 70 memory-safety issues and 100 concurrency bugs.
 They also provided Rust programmers with some suggestions and insights to develop better Rust programs.
-At last, using the results of their analysis, they designed two bug detectors and provided recommendations for developing bug detector in the future.
+Using the results of their manual analysis, they designed two bug detectors, and provided recommendations for developing more bug detectors in the future.
 
-Li et al.~\cite{li2021mirchecker} present MirChekcer, a fully automated bug detection framework for Rust.
+Li et al.~\cite{li2021mirchecker} present MirChecker, a fully automated bug detection framework for Rust.
 This framework works by carrying out static analysis on Rust's Mid-level Intermediate Representation (MIR).
 The tool exploits the insights obtained from observing existing bugs in Rust code bases.
-Through recording both numerical and symbolic information, the framework detects errors by using constraint solving techniques.
-The framework proves its practicability as it detected 33 new bugs including 16 memory-safety faults within 12 Rust crates.
+Using both numerical and symbolic information, the framework detects errors by using constraint solving techniques.
+MirChecker detected 33 new bugs, including 16 memory safety faults, across 12 Rust crates.
+% how do they get the set of existing bugs?
+% what do you mean by numerical information?

### 2022-10-13 sections/methodology.tex
added labels to tikZs

@@
     \draw (-1.5,1.5) node[shape=rectangle, minimum width=3cm,minimum height=5.4cm, rounded corners=1ex, xshift=1.5cm, yshift=-2.5cm](context){};
+
+    \node[draw, minimum height=0.5, fill=yellow!20] at (-0.74,1.7) {ASTDiff};
 
@@
 
+        \node[draw, minimum height=0.5, fill=yellow!20] at (-0.74,1.7) {ASTDiff};
+
+        \node[draw, minimum height=0.5, fill=yellow!20] at (4.29,1.73) {Occurrences data};
+
         \matrix (table1) [table,text width=6em,nodes={scale=0.45,transform shape}] at (8.5,1)
@@
         };
+
+        \node[draw, minimum height=0.5, fill=yellow!20] at (4.02,-0.25) {Weights data};
 
@@
 \caption{\label{table:repos} Target repositories for mining}
-\begin{tabular}{c|c|c|c|c}
-    Project Name & Stars(k) & LOC & \# Commits & Avg. Code Churn \\ \hline
-    denoland/deno & 84.7 & 562468 & 7730 & 49.22 \\ \hline
-    tauri-apps/tauri & 48.5 & 130726 & 2929 & 31.67 \\ \hline
-    alacritty/alacritty & 41.2 & 45803 & 2017 & 27.53 \\ \hline
-    sharkdp/bat & 36.4 & 93258 & 2392 & 59.13 \\ \hline
-    starship/starship & 28 & 176630 & 2290 & 25.1 \\ \hline
-    rust-lang/rustlings & 28.7 & 9852 & 1391 & 8.92 \\ \hline
-    meilisearch/meilisearch & 28.6 & 66872 & 3849 & 279.7 \\ \hline
-    sharkdp/fd & 24.2 & 10383 & 1025 & 12.2 \\ \hline
-    swc-project/swc & 23.5 & 13512766 & 5205 & 44.34 \\ \hline
-    yewstack/yew & 23.2 & 112972 & 2239 & 21.79 \\ \hline
-    xi-editor/xi-editor & 19.5 & 58783 & 2104 & 28.67 \\ \hline
-    AppFlowy-IO/AppFlowy & 25.9 & 182830 & 2600 & 29.49 \\ \hline
-    firecracker-microvm/firecracker & 18.9 & 156759 & 3536 & 58.26 \\ \hline
-    nushell/nushell & 20.3 & 181710 & 5922 & 22.38 \\ \hline
-    ogham/exa & 18.9 & 17114 & 1558 & 11.99 \\ \hline
-    SergioBenitez/Rocket & 18.2 & 90055 & 2127 & 19.75 \\ \hline
-    rustdesk/rustdesk & 27.7 & 85322 & 1664 & 72.81 \\ \hline
+\begin{tabular}{l r r r r}
+    Project Name & Stars(k) & LOC & \# Commits & Avg. Code Churn \\
+    denoland/deno & 84.7 & 562468 & 7730 & 49.22 \\
+    tauri-apps/tauri & 48.5 & 130726 & 2929 & 31.67 \\
+    alacritty/alacritty & 41.2 & 45803 & 2017 & 27.53 \\
+    sharkdp/bat & 36.4 & 93258 & 2392 & 59.13 \\
+    starship/starship & 28 & 176630 & 2290 & 25.1 \\
+    rust-lang/rustlings & 28.7 & 9852 & 1391 & 8.92 \\
+    meilisearch/meilisearch & 28.6 & 66872 & 3849 & 279.7 \\
+    sharkdp/fd & 24.2 & 10383 & 1025 & 12.2 \\
+    swc-project/swc & 23.5 & 13512766 & 5205 & 44.34 \\
+    yewstack/yew & 23.2 & 112972 & 2239 & 21.79 \\
+    xi-editor/xi-editor & 19.5 & 58783 & 2104 & 28.67 \\
+    AppFlowy-IO/AppFlowy & 25.9 & 182830 & 2600 & 29.49 \\
+    firecracker-microvm/firecracker & 18.9 & 156759 & 3536 & 58.26 \\
+    nushell/nushell & 20.3 & 181710 & 5922 & 22.38 \\
+    ogham/exa & 18.9 & 17114 & 1558 & 11.99 \\
+    SergioBenitez/Rocket & 18.2 & 90055 & 2127 & 19.75 \\
+    rustdesk/rustdesk & 27.7 & 85322 & 1664 & 72.81 \\
     tokio-rs/tokio & 17.2 & 126089 & 3006 & 28.87 

### 2022-10-14 sections/related.tex
added fixed to related

@@
 
-% say why pattern-based repair is different from search-based repair -- isn't it sufficient to refer the reader to the survey?
-% Pan et al: was it manual? or not? summarize their technique and say whether what they found is similar to what we found.
 The main intuition in pattern-based program repair methods is that bug fixes follow certain patterns.
@@
 Pan et al.~\cite{pan2009toward} analyzed seven large-scale widely used Java projects and obtained 27 common bug fix patterns.
+They used a bug fix pattern extractor tool to automatically parse and detect bug patterns within the bug fix hunks with less than seven statements.
+They ignored larger bug fix hunks as they tended to exhibit random changes and did not have observable meaningful patterns.
 Though they did the analysis on Java projects, their reported patterns were not specific to Java.
@@
 However, our novel program encoding approach differs from what they did.
-Also, like~\cite{yang2022mining}, we introduced patterns in two different categories.
+In~\cite{hanam2016discovering}, they introduced Feature Properties table, which is a categorization for program elements.
+Encoding programs based on this table resulted in a order-sensitive encoding, however, it made the datapoints sparse.
+Moreover, like~\cite{yang2022mining}, we introduced patterns in two different categories.
 In our case there were general and language-specific patterns; for Rust, the language-specific patterns were related to the borrow checker (BC patterns).
-% ok they didn't do our program encoding; what did they do instead?
+However, in oppose to our paper, in~\cite{yang2022mining}, they did a much more elaborate manual analysis, as in their data collection, they only collected general information about the single hunk changes (e.g.
+the number of variables, or arguments).
 
@@
 This framework works by carrying out static analysis on Rust's Mid-level Intermediate Representation (MIR).
-The tool exploits the insights obtained from observing existing bugs in Rust code bases.
-Using both numerical and symbolic information, the framework detects errors by using constraint solving techniques.
+The tool exploits the insights obtained from observing existing bugs (by studying reported CVEs) in Rust code bases.
+Using both numerical (e.g.
+integer bounds) and symbolic (e.g.
+modelling memory) information, the framework detects errors by using constraint solving techniques.
 MirChecker detected 33 new bugs, including 16 memory safety faults, across 12 Rust crates.
-% how do they get the set of existing bugs?
-% what do you mean by numerical information?

### 2022-10-14 sections/methodology.tex
added some minor changes, about to fix results

@@
 
-    \caption{\label{fig:extraction}Extracting paths from an ASTDiff.
-the ASTDiff on left corresponds to the code change in the yellow box on the top right.
-A DFS search will identify three paths, shown in three different colors at right.}
+    \caption{\label{fig:extraction}The ASTDiff (left) corresponds to the code change in the yellow box (top right). A DFS will identify three paths, shown in three different colors at right.}
 
@@
 
-        \node[draw, minimum height=0.5, fill=yellow!20] at (4.02,-0.25) {Weights data};
+        \node[draw, minimum height=0.5, fill=yellow!20] at (4.12,-0.25) {Weighted data};
 
@@
 
-    \caption{\label{fig:essence}Multiplying the number of occurrences of items observed in the tree by their respective weights, resulting in the essence of change (visualized with a circle pack figure at right).
-The left and right key value tables show the number of occurrences of each item and its respective weight, respectively.}
+    \caption{\label{fig:essence}Multiplying the number of occurrences of items observed in the tree by their respective weights, resulting in the essence of change (visualized with a circle pack figure at bottom right).
+Vectors show (above) the number of occurrences and (below) weighted occurrence counts.}
 \end{figure}

### 2022-10-14 sections/results.tex
added some minor changes, about to fix results

@@
 
-\subsection{Attributes}
+\subsubsection{Attributes}
 
@@
 
-\noindent \textit{\label{sec:G.attr.struct}\textbf{G.attr.struct} Modifying the attributes of structs}
+\noindent\textit{\label{sec:G.attr.struct}\textbf{G.attr.struct} Modifying the attributes of structs}
 
@@
 
-\noindent \textbf{Description:} In this pattern, the attribute set of a struct changes.
+\noindent\textbf{Description:} In this pattern, the attribute set of a struct changes.
 This modification can happen in three different ways: (G.attr.struct.add) adding attributes to the struct; (G.attr.struct.drop) removing attributes from the struct; and (G.attr.struct.change) changing the content of an existing attribute.
@@
 
-\noindent \textit{ \textbf{G.attr.struct.field} Modifying the attributes of struct fields}
+\noindent\textit{\textbf{G.attr.struct.field} Modifying the attributes of struct fields}
 
@@
 
-\noindent \textbf{Description:} In this pattern, the attribute set of a struct field changes.
+\noindent\textbf{Description:} In this pattern, the attribute set of a struct field changes.
 As with \textit{\textbf{G.attr.struct}}, this modification also happens in three different ways.
@@
 
-\subsection{Struct} 
-\noindent \textit{ \textbf{G.struct.field} Adding/Removing a struct field}
+\subsubsection{Struct} 
+\noindent\textit{\textbf{G.struct.field} Adding/Removing a struct field}
 
@@
 
-\noindent \textbf{Description:} A set of new fields are added to or removed from an existing struct.
+\noindent\textbf{Description:} A set of new fields are added to or removed from an existing struct.
 A developer adds a new field to store a new piece of data in a struct.
@@
 
-\noindent \textit{ \textbf{G.struct.field.pub} Modifying the access modifiers of struct fields}
+\noindent\textit{\textbf{G.struct.field.pub} Modifying the access modifiers of struct fields}
 
@@
 
-\noindent \textbf{Description:} Adding \texttt{pub} to a struct field makes it possible to access that field from external modules.
+\noindent\textbf{Description:} Adding \texttt{pub} to a struct field makes it possible to access that field from external modules.
 In the instances of this cluster, adding the \texttt{pub} access modifier happened in a bug fixing change, where the developer needed to access a field which had not been marked for public access.
@@
 
-\subsection{Option}
+\subsubsection{Option}
 \vspace{3mm}
 
-\noindent \textit{ \textbf{G.option} Changing field type T to T$\langle$Option$\rangle$}
+\noindent\textit{\textbf{G.option} Changing field type T to T$\langle$Option$\rangle$}
 
@@
 
-\noindent \textbf{Description:} Rust is a null-safe language, meaning that object references cannot take on null values.
+\noindent\textbf{Description:} Rust is a null-safe language, meaning that object references cannot take on null values.
 If we want a variable to contain null, we would have to use the \texttt{Option} type.
@@
 
-\subsection{Types}
+\subsubsection{Types}
 \vspace{3mm}
 
-\noindent \textit{\label{sec:G.type.field}\textbf{G.type.field} Changing a struct field type}
+\noindent\textit{\label{sec:G.type.field}\textbf{G.type.field} Changing a struct field type}
 
@@
 
-\noindent \textbf{Description:} The type of a struct field changes.
+\noindent\textbf{Description:} The type of a struct field changes.
 This is a common change pattern in all statically-typed languages and can happen for various reasons.
@@
 
-\noindent \textit{ \textbf{G.type.generic} Changing a generic type parameter}
+\noindent\textit{\textbf{G.type.generic} Changing a generic type parameter}
 
@@
 
-\noindent \textbf{Description:} This a change of the type parameter of a struct field. As with \textit{\textbf{G.type.field}}, this change can also happen for the mentioned reasons. \\
+\noindent\textbf{Description:} This a change of the type parameter of a struct field. As with \textit{\textbf{G.type.field}}, this change can also happen for the mentioned reasons. \\
 
-\noindent \textit{ \textbf{G.type.enum.variant} Change in enum variant value type}
+\noindent\textit{\textbf{G.type.enum.variant} Change in enum variant value type}
 
@@
 
-\noindent \textbf{Description:} In Rust, developers can specify the type of the value that an enum variant can store.
+\noindent\textbf{Description:} In Rust, developers can specify the type of the value that an enum variant can store.
 Using this feature, they can avoid having to use a struct along with enum variants.
@@
 
-\noindent \textit{ \textbf{G.type.tuple} Changing the type inside a tuple}
+\noindent\textit{\textbf{G.type.tuple} Changing the type inside a tuple}
 
@@
 
-\noindent \textbf{Description:} In this pattern, an extra type is added to an existing type, making it a tuple of multiple types.
+\noindent\textbf{Description:} In this pattern, an extra type is added to an existing type, making it a tuple of multiple types.
 Alternatively, the change might be dropping types from the tuple and making it a smaller-arity tuple.
 
-\subsection{Traits}
-\noindent \textit{ \textbf{G.trait.bound} TraitBound Change}
+\subsubsection{Traits}
+\noindent\textit{\textbf{G.trait.bound} TraitBound Change}
 
@@
 
-\noindent \textbf{Description:} The trait bounds are the functionalities that we require from our parametric types.
+\noindent\textbf{Description:} The trait bounds are the functionalities that we require from our parametric types.
 This change pattern relates to modification in trait bounds.
@@
 
-\noindent \textit{ \textbf{BC.clone.drop} Removing clone}
+\noindent\textit{\textbf{BC.clone.drop} Removing clone}
 
@@
 
-\noindent \textbf{Description:} This pattern is referred to as redundant clone in Clippy Lints.
+\noindent\textbf{Description:} This pattern is referred to as redundant clone in Clippy Lints.
 Clippy\footnote{\url{https://github.com/rust-lang/rust-clippy}} is a linter tool for Rust that catches common mistakes.
@@
 
-\noindent \textit{ \textbf{BC.clone.ref} Dropping clone and adding borrowing}
+\noindent\textit{\textbf{BC.clone.ref} Dropping clone and adding borrowing}
 
@@
 
-\noindent \textbf{Description:} Unlike the previous pattern, this pattern is not detected by Clippy.
+\noindent\textbf{Description:} Unlike the previous pattern, this pattern is not detected by Clippy.
 The change happens when the programmers realizes that the cloning of a variable is unnecessary, however, we also want to preserve the ownership of the variable within the current scope.
@@
 
-\noindent \textit{ \textbf{BC.ref.add} Adding Borrowing}
+\noindent\textit{\textbf{BC.ref.add} Adding Borrowing}
 
@@
 
-\noindent \textbf{Description:} This change happens when the developer needs to modify the program so it reaches the desired behaviour.
+\noindent\textbf{Description:} This change happens when the developer needs to modify the program so it reaches the desired behaviour.
 The modification, however, requires access to an object whose ownership has moved to another scope and never returned.
@@
 
-\noindent \textit{ \textbf{BC.ref.drop} Removing Borrowing}
+\noindent\textit{\textbf{BC.ref.drop} Removing Borrowing}
 
@@
 
-\noindent \textbf{Description:} In the pattern, an unnecessary borrowing is removed.
+\noindent\textbf{Description:} In the pattern, an unnecessary borrowing is removed.
 Clippy can detect the needless borrows, and also can automatically remove them.
@@
 
-\noindent \textit{ \textbf{BC.deref.to\_string()} Modification in dereferencing a string prior to calling to\_string()}
+\noindent\textit{\textbf{BC.deref.to\_string()} Modification in dereferencing a string prior to calling to\_string()}
 
@@
 
-\noindent \textbf{Description:} This is a change that in all the instances has been detected by Clippy.
+\noindent\textbf{Description:} This is a change that in all the instances has been detected by Clippy.
 The change applies on an object prior to calling \verb+to_string()+ on that object.
@@
 
-\noindent \textit{ \textbf{BC.mut.add} Adding mutability}
+\noindent\textit{\textbf{BC.mut.add} Adding mutability}
 
@@
 
-\noindent \textbf{Description:} This change happens when the programmer needs to mutate a variable which is immutable, hence they need to add the mut keyword.
+\noindent\textbf{Description:} This change happens when the programmer needs to mutate a variable which is immutable, hence they need to add the mut keyword.
 In the observed instances, this pattern was not seen in refactoring changes, the modification happened alongside a change in program behaviour.
@@
 
-\noindent \textit{ \textbf{BC.mut.drop} Dropping mutability}
+\noindent\textit{\textbf{BC.mut.drop} Dropping mutability}
 
@@
 
-\noindent \textbf{Description:} Dropping the mut keyword before a variable, as we do not mutate the variable. This redundant mutability is reported by Clippy and can also be removed through this tool.
+\noindent\textbf{Description:} Dropping the mut keyword before a variable, as we do not mutate the variable. This redundant mutability is reported by Clippy and can also be removed through this tool.
 
@@
 
-\noindent \textit{ \textbf{BC.vec.slice} Changing a Vec reference to Slice}
+\noindent\textit{\textbf{BC.vec.slice} Changing a Vec reference to Slice}
 
@@
 \end{lstlisting}
-\noindent \textbf{Description:} In struct field definition or the types of the function formal parameters, type \verb+&Vec<T>+ changes to \verb+&[T]+.
+\noindent\textbf{Description:} In struct field definition or the types of the function formal parameters, type \verb+&Vec<T>+ changes to \verb+&[T]+.
 Clippy can detect the opportunity to apply this pattern.
@@
 \subsubsection{Lifetime}
-\noindent \textit{ \textbf{BC.lifetime.static.drop} Dropping static lifetime}
+\noindent\textit{\textbf{BC.lifetime.static.drop} Dropping static lifetime}
 
@@
 
-\noindent \textbf{Description:} Removing the static keyword from a const variable declaration.
+\noindent\textbf{Description:} Removing the static keyword from a const variable declaration.
 This change can be applied using Clippy.

### 2022-10-16 sections/methodology.tex
applied fixes

@@
     \end{lstlisting}
-
     \end{minipage}%
@@
     Project Name & Stars(k) & LOC & \# Commits & Avg. Code Churn \\
-    denoland/deno & 84.7 & 562468 & 7730 & 49.22 \\
-    tauri-apps/tauri & 48.5 & 130726 & 2929 & 31.67 \\
-    alacritty/alacritty & 41.2 & 45803 & 2017 & 27.53 \\
-    sharkdp/bat & 36.4 & 93258 & 2392 & 59.13 \\
-    starship/starship & 28 & 176630 & 2290 & 25.1 \\
-    rust-lang/rustlings & 28.7 & 9852 & 1391 & 8.92 \\
-    meilisearch/meilisearch & 28.6 & 66872 & 3849 & 279.7 \\
-    sharkdp/fd & 24.2 & 10383 & 1025 & 12.2 \\
-    swc-project/swc & 23.5 & 13512766 & 5205 & 44.34 \\
-    yewstack/yew & 23.2 & 112972 & 2239 & 21.79 \\
-    xi-editor/xi-editor & 19.5 & 58783 & 2104 & 28.67 \\
-    AppFlowy-IO/AppFlowy & 25.9 & 182830 & 2600 & 29.49 \\
-    firecracker-microvm/firecracker & 18.9 & 156759 & 3536 & 58.26 \\
-    nushell/nushell & 20.3 & 181710 & 5922 & 22.38 \\
-    ogham/exa & 18.9 & 17114 & 1558 & 11.99 \\
-    SergioBenitez/Rocket & 18.2 & 90055 & 2127 & 19.75 \\
-    rustdesk/rustdesk & 27.7 & 85322 & 1664 & 72.81 \\
-    tokio-rs/tokio & 17.2 & 126089 & 3006 & 28.87 
+    denoland/deno & 84.7 & 146055 & 7730 & 49.22 \\
+    tauri-apps/tauri & 48.5 & 50014 & 2929 & 31.67 \\
+    alacritty/alacritty & 41.2 & 30124 & 2017 & 27.53 \\
+    sharkdp/bat & 36.4 & 9496 & 2392 & 59.13 \\
+    starship/starship & 28 & 34888 & 2290 & 25.1 \\
+    rust-lang/rustlings & 28.7 & 5067 & 1391 & 8.92 \\
+    meilisearch/meilisearch & 28.6 & 25981 & 3849 & 279.7 \\
+    sharkdp/fd & 24.2 & 6542 & 1025 & 12.2 \\
+    swc-project/swc & 23.5 & 346501 & 5205 & 44.34 \\
+    yewstack/yew & 23.2 & 44295 & 2239 & 21.79 \\
+    xi-editor/xi-editor & 19.5 & 39292 & 2104 & 28.67 \\
+    AppFlowy-IO/AppFlowy & 25.9 & 57646 & 2600 & 29.49 \\
+    firecracker-microvm/firecracker & 18.9 & 82021 & 3536 & 58.26 \\
+    nushell/nushell & 20.3 & 163562 & 5922 & 22.38 \\
+    ogham/exa & 18.9 & 10852 & 1558 & 11.99 \\
+    SergioBenitez/Rocket & 18.2 & 68952 & 2127 & 19.75 \\
+    rustdesk/rustdesk & 27.7 & 48115 & 1664 & 72.81 \\
+    tokio-rs/tokio & 17.2 & 121400 & 3006 & 28.87 
 \end{tabular}

### 2022-10-16 sections/results.tex
applied fixes

@@
 
-\textbf{RQ1. Does our data modelling approach capture the most important aspect of program changes?}
+\textbf{Does our data modelling approach capture the most important aspect of program changes?}
 
@@
 
-{\bf There is no text in this subsection.}
+\textbf{What are the general fix patterns in Rust, and how often do they apply?}
+
+We collected 60583 datapoints for general fix patterns using a server with Xeon Gold 5120 Processor (Q3-2017, 14-core 2.2 to 3.2GHz) and the data collection took around 26 hours to complete for both $D_g$ and $D_b$.
+We used $\epsilon=0.0001$ and $Z=5$ as our DBSCAN parameters and that resulted in 577 clusters.
+After manual analysis as described in~\ref{sec:manual_analysis_cluster_selection} we ended up with 10 general patterns which we introduce here:
 
@@
 
+\textbf{What are borrow-checker related fix patterns in Rust, and how often do they apply?}
+
+We collected 27143 datapoints for borrow-checker related fix patterns.
+We used $\epsilon=0.001$ and $Z=5$ as our DBSCAN parameters and that resulted in 102 clusters.
+After manual analysis as described in~\ref{sec:manual_analysis_cluster_selection} we ended up with 9 borrow-checker related fix patterns which we introduce here:
+
 \begin{table}[]

### 2022-10-16 sections/methodology.tex
updated stars and commit #

@@
     Project Name & Stars(k) & LOC & \# Commits & Avg. Code Churn \\
-    denoland/deno & 84.7 & 146055 & 7730 & 49.22 \\
-    tauri-apps/tauri & 48.5 & 50014 & 2929 & 31.67 \\
-    alacritty/alacritty & 41.2 & 30124 & 2017 & 27.53 \\
-    sharkdp/bat & 36.4 & 9496 & 2392 & 59.13 \\
-    starship/starship & 28 & 34888 & 2290 & 25.1 \\
-    rust-lang/rustlings & 28.7 & 5067 & 1391 & 8.92 \\
-    meilisearch/meilisearch & 28.6 & 25981 & 3849 & 279.7 \\
-    sharkdp/fd & 24.2 & 6542 & 1025 & 12.2 \\
-    swc-project/swc & 23.5 & 346501 & 5205 & 44.34 \\
-    yewstack/yew & 23.2 & 44295 & 2239 & 21.79 \\
-    xi-editor/xi-editor & 19.5 & 39292 & 2104 & 28.67 \\
-    AppFlowy-IO/AppFlowy & 25.9 & 57646 & 2600 & 29.49 \\
-    firecracker-microvm/firecracker & 18.9 & 82021 & 3536 & 58.26 \\
-    nushell/nushell & 20.3 & 163562 & 5922 & 22.38 \\
-    ogham/exa & 18.9 & 10852 & 1558 & 11.99 \\
-    SergioBenitez/Rocket & 18.2 & 68952 & 2127 & 19.75 \\
-    rustdesk/rustdesk & 27.7 & 48115 & 1664 & 72.81 \\
-    tokio-rs/tokio & 17.2 & 121400 & 3006 & 28.87 
+    denoland/deno & 85.9 & 146055 & 8026 & 49.22 \\
+    tauri-apps/tauri & 52.5 & 50014 & 3041 & 31.67 \\
+    alacritty/alacritty & 42.4 & 30124 & 2036 & 27.53 \\
+    sharkdp/bat & 37.6 & 9496 & 2457 & 59.13 \\
+    starship/starship & 29.3 & 34888 & 2379 & 25.1 \\
+    rust-lang/rustlings & 30.6 & 5067 & 1466 & 8.92 \\
+    meilisearch/meilisearch & 30.2 & 25981 & 4029 & 279.7 \\
+    sharkdp/fd & 24.9 & 6542 & 1068 & 12.2 \\
+    swc-project/swc & 24.2 & 346501 & 5884 & 44.34 \\
+    yewstack/yew & 24.5 & 44295 & 2280 & 21.79 \\
+    xi-editor/xi-editor & 19.6 & 39292 & 2105 & 28.67 \\
+    AppFlowy-IO/AppFlowy & 28.4 & 57646 & 3321 & 29.49 \\
+    firecracker-microvm/firecracker & 19.6 & 82021 & 3620 & 58.26 \\
+    nushell/nushell & 21.1 & 163562 & 6088 & 22.38 \\
+    ogham/exa & 19.5 & 10852 & 1558 & 11.99 \\
+    SergioBenitez/Rocket & 18.7 & 68952 & 2135 & 19.75 \\
+    rustdesk/rustdesk & 30.2 & 48115 & 2620 & 72.81 \\
+    tokio-rs/tokio & 17.9 & 121400 & 3101 & 28.87 
 \end{tabular}

### 2022-10-16 main.tex
added abstract/conclusion; fixed results a bit

@@
 \begin{abstract}
-Insert your abstract here. Include keywords, PACS and mathematical
-subject classification numbers as needed.
+    Rust is a relatively new programming language that allows programmers to write programs that have low-level control over resources, as well as high-level safety guarantees.
+This dual nature is made possible through the ownership framework which has been incorporated in the language.
+The main component governing over this framework is called Borrow-Checker.
+Rust has been ranked first for the sixth consecutive year in the Stack Overflow's annual Developer Survey as the most loved programming language.
+The number of Rust developers is growing as the need for faster and safer software increases.
+However, to the best of our knowledge, no research has been developed to automatically find the most pervasive bug fix patterns within Rust programs.
+In this project, using a novel program modelling method, we mined the top 18 most starred Rust projects in Github to discover the most common bug fix patterns lied within their code changes.
+We stored 87,726 code changes drawn from 57,214 commits across these 18 projects.
+After clustering the code changes, and conducting a manual analysis, we discovered 19 groups of cross-project bug fix patterns.
+We introduce these patterns in two groups: (1) general patterns and (2) borrow-checker related patterns.
 \keywords{First keyword \and Second keyword \and More}

### 2022-10-16 sections/conclusion.tex
added abstract/conclusion; fixed results a bit

In this project, we mined the top 18 most starred Rust projects in Github to discover common bug fix patterns lied within their code changes. We proposed a novel program modelling approach to encode most important aspects of a code change in a fixed sized datapoint. We stored 87,726 datapoints drawn from 57,214 commits across these 18 projects. After using the DBSCAN clustering algorithm, and a subsequent manual analysis, we obtained 10 clusters of general fix patterns and 9 clusters of borrow-checker related fix patterns.

### 2022-10-16 sections/introduction.tex
added abstract/conclusion; fixed results a bit

@@
 One useful step towards that goal would be studying common bug patterns (and their corresponding fixes) that occur across multiple software systems.
-Research efforts for categorizing common bug patterns date back to 1975 where Endres tried to categorize various bug patterns which occurred in operating system implementations \cite{endres1975analysis}.
+Research efforts for categorizing common bug patterns date back to 1975 where Endres tried to categorize various bug patterns which occurred in operating system implementations~\cite{endres1975analysis}.
 
@@
 Our interest comes from a desire to develop program repair tools, but we believe these insights can also be of broader service.
-Automated program repair \cite{goues2019automated} \cite{liu2018survey} tools try to locate bug locations in the source code (using a fault localization module) and then generate patches to fix them, possibly without human intervention, so that the modified program meets desired specifications (implicit or explicit).
-A powerful automated program repair system has many useful applications and could significantly simplify debugging, thus reducing software development cost \cite{le2012systematic}.
-A set of common bug fix patterns can be a useful input to an APR system: it directly provides information about fixing similar buggy structures, which can significantly reduce the search space for patch generation \cite{jeffrey2009bugfix}.
+Automated program repair~\cite{goues2019automated,liu2018survey} tools try to locate bug locations in the source code (using a fault localization module) and then generate patches to fix them, possibly without human intervention, so that the modified program meets desired specifications (implicit or explicit).
+A powerful automated program repair system has many useful applications and could significantly simplify debugging, thus reducing software development cost~\cite{le2012systematic}.
+A set of common bug fix patterns can be a useful input to an APR system: it directly provides information about fixing similar buggy structures, which can significantly reduce the search space for patch generation~\cite{jeffrey2009bugfix}.
 
@@
 But where is the fault?
-Fault localization \cite{wong2016survey} is the process of automatically finding the faulty statement(s) in our program.
+Fault localization~\cite{wong2016survey} is the process of automatically finding the faulty statement(s) in our program.
 If a fault localization tool knows about the common bug patterns that can appear in the underlying programming language, it can prioritize code segment candidates based on their potentials of producing bugs.
@@
 This knowledge can make the developers avoid certain mistakes during production which ultimately would result in reducing the frequent bug patterns within the program.
-As Monperrus describes, the fewer frequent bug patterns we keep in our software, the less cost we would have to pay for software maintenance \cite{monperrus2014critical}.
+As Monperrus describes, the fewer frequent bug patterns we keep in our software, the less cost we would have to pay for software maintenance~\cite{monperrus2014critical}.
 
 Moving further afield, researchers have been exploiting deep learning techniques to learn about the semantics of computer programs and create program modifier models that have a specific goal (e.g.
-bug fix) \cite{alon2019code2vec} \cite{alon2018general} \cite{raychev2016learning} \cite{bielik2016phog}.
+bug fix)~\cite{alon2019code2vec,alon2018general,raychev2016learning,bielik2016phog}.
 However, most learning tools get fixed-size vectors as input and map them to outputs.
@@
 
-Many studies have investigated detecting and categorizing bug fix patterns in general \cite{islam2021changes} \cite{madeiral2018towards} \cite{pan2009toward}, or for a specific programming language \cite{yang2022mining} \cite{hanam2016discovering} \cite{campos2019discovering}.
+Many studies have investigated detecting and categorizing bug fix patterns in general~\cite{islam2021changes,madeiral2018towards,pan2009toward}, or for a specific programming language~\cite{yang2022mining,hanam2016discovering,campos2019discovering}.
 In this work, we study common bug patterns that appear in Rust projects.
-Rust is a relatively new language that allows programmers to write programs that have low-level control over resources, as well as high-level safety guarantees \cite{klabnik2019rust}.
+Rust is a relatively new language that allows programmers to write programs that have low-level control over resources, as well as high-level safety guarantees~\cite{klabnik2019rust}.
 What distinguishes Rust from the other systems programming languages, e.g.
@@
 These problems had to be manually ensured in languages like C/C++.
-Moreover, not only does the ownership concept remove the need for a garbage collector in the runtime environment, it does not require the programmer to explicitly do certain safety tasks \cite{qin2020understanding}.
+Moreover, not only does the ownership concept remove the need for a garbage collector in the runtime environment, it does not require the programmer to explicitly do certain safety tasks~\cite{qin2020understanding}.
 Due to the memory and thread safety guarantees that Rust provides, many software companies adopted Rust in production.
-For example, Dropbox decided to migrate to Rust for developing their storage system, since they were not completely satisfied with the features provided by AWS’s S3 or Golang \cite{dropbox}.
+For example, Dropbox decided to migrate to Rust for developing their storage system, since they were not completely satisfied with the features provided by AWS’s S3 or Golang~\cite{dropbox}.
 

### 2022-10-16 sections/methodology.tex
added abstract/conclusion; fixed results a bit

@@
 The general idea for automatically finding the pervasive code patterns within a list of programs is to parse the programs and encode the important parts of the derived AST in fixed-size datapoints.
-Work in this vein makes the assumption that each cluster corresponds to one change pattern \cite{hanam2016discovering} \cite{campos2019discovering} \cite{yang2022mining}.
+Work in this vein makes the assumption that each cluster corresponds to one change pattern~\cite{hanam2016discovering,campos2019discovering,yang2022mining}.
 
@@
 Our target repositories were the top 18 most-starred open source Rust projects on GitHub.
-We mined (using Pydriller \cite{spadini2018pydriller}) all the bug related commits and ran them through our pipeline.
-Then, we applied the DBSCAN clustering algorithm \cite{ester1996density} on the resulting datapoints.
+We mined (using Pydriller~\cite{spadini2018pydriller}) all the bug related commits and ran them through our pipeline.
+Then, we applied the DBSCAN clustering algorithm~\cite{ester1996density} on the resulting datapoints.
 In Section~\ref{sec:clustering_data} we justify our choice of DBSCAN for clustering, and how we tuned it to improve cluster quality.
@@
 First, we need to define a fixed set of columns.
-We need fixed-size datapoints, as they are a requirement for both centroid-based and density-based well-known clustering algorithms \cite{xu2005survey}.
+We need fixed-size datapoints, as they are a requirement for both centroid-based and density-based well-known clustering algorithms~\cite{xu2005survey}.
 The number of columns would depend on the amount of information we want to encode in the datapoints.
@@
 A reasonable work around would be to reduce $n$ from the full set of non terminals to a smaller set of categories (e.g.
-a category for larger entites like class or function definitions, and a category for smaller entities like statements and expressions), and then to record all possible combinations of these categories (similar to \cite{hanam2016discovering}).
+a category for larger entites like class or function definitions, and a category for smaller entities like statements and expressions), and then to record all possible combinations of these categories (similar to~\cite{hanam2016discovering}).
 That approach reduces the number of dimensions and provides more reasonable and syntactically-correct combinations.
@@
 For each repository, we search all commits that include bug-fixing related keywords within their commit messages.
-Here, we chose bug-fixing related keywords based on the set introduced in \cite{zhang2018empirical}, excluding `nan' and `inf'.
+Here, we chose bug-fixing related keywords based on the set introduced in~\cite{zhang2018empirical}, excluding `nan' and `inf'.
 
@@
 Secondly, it is a density-based clustering method, which means that it detects arbitrarily shaped clusters; centroid-based methods like k-means can't detect such clusters.
-We tried k-means and the more powerful fuzzy variant Fuzzy C-means \cite{dunn1973fuzzy} and did not get useful clusters.
+We tried k-means and the more powerful fuzzy variant Fuzzy C-means~\cite{dunn1973fuzzy} and did not get useful clusters.
 
@@
 
-Similar to \cite{yang2022mining}, following Cotroneo et al. \cite{cotroneo2019analyzing} definitions, we divide the clusters in three different groups:
+Similar to~\cite{yang2022mining}, following Cotroneo et al.~\cite{cotroneo2019analyzing} definitions, we divide the clusters in three different groups:
 
@@
 Finally, we selected 19 of the remaining clusters---the ones that manifested concrete changes.
-We discuss about our findings in Sections~\ref{sec:common_patterns} and \ref{sec:bc_patterns}.
+We discuss about our findings in Sections~\ref{sec:common_patterns}~and~\ref{sec:bc_patterns}.

### 2022-10-16 sections/results.tex
added abstract/conclusion; fixed results a bit

@@
 The analyzer compares the actual change in the code and compares it to the visual representation of its respective data point.
-We used circle pack figures to visualize each data point \cite{collins2003circle}.
+We used circle pack figures to visualize each data point~\cite{collins2003circle}.
 For this purpose, we used the packcircles python library\footnote{\url{https://github.com/mhtchan/packcircles}}.
@@
 // G.attr.struct.add: Adding attributes to a struct
-++ #[derive(Debug, Clone)]
+<@\textcolor{green}{++}@> #[derive(Debug, Clone)]
 pub struct Call {
@@
 //G.attr.struct.drop: Removing attributes from a struct
--- #[derive(Debug, Clone)]
+<@\textcolor{red}{--}@> #[derive(Debug, Clone)]
 pub struct Call {
@@
 // G.attr.struct.change: Changing an existing attribute of a struct
--- #[derive(Debug)]
-++ #[derive(Debug, Clone)]
+<@\textcolor{red}{--}@> #[derive(Debug)]
+<@\textcolor{green}{++}@> #[derive(Debug, Clone)]
 pub struct Call {
@@
 pub struct Retain {
-    ++ #[serde(skip_serializing_if = "is_empty")]
+    <@\textcolor{green}{++}@> #[serde(skip_serializing_if = "is_empty")]
     pub attributes: Option<Attributes>,
@@
 pub struct Retain {
-    -- #[serde(skip_serializing_if = "Option::is_none")]
+    <@\textcolor{red}{--}@> #[serde(skip_serializing_if = "Option::is_none")]
     pub attributes: Option<Attributes>,
@@
 pub struct Retain {
-    -- #[serde(skip_serializing_if = "Option::is_none")]
-    ++ #[serde(skip_serializing_if = "is_empty")]
+    <@\textcolor{red}{--}@> #[serde(skip_serializing_if = "Option::is_none")]
+    <@\textcolor{green}{++}@> #[serde(skip_serializing_if = "is_empty")]
     pub attributes: Option<Attributes>,
@@
 pub struct HtmlBlock {
-    -- content: BlockContent,
-    ++ pub content: BlockContent,
+    <@\textcolor{red}{--}@> content: BlockContent,
+    <@\textcolor{green}{++}@> pub content: BlockContent,
     brace: token::Brace,
@@
 pub struct Languages {
---  pub named: HashMap<LanguageId, Arc<LanguageDefinition>>,
-++  named: HashMap<LanguageId, Arc<LanguageDefinition>>,
+<@\textcolor{red}{--}@>  pub named: HashMap<LanguageId, Arc<LanguageDefinition>>,
+<@\textcolor{green}{++}@>  named: HashMap<LanguageId, Arc<LanguageDefinition>>,
     extensions: HashMap<String, Arc<LanguageDefinition>>,
@@
     pub span: Span,
-    -- pub media: MediaQueryList,
-    ++ pub media: Option<MediaQueryList>,
+    <@\textcolor{red}{--}@> pub media: MediaQueryList,
+    <@\textcolor{green}{++}@> pub media: Option<MediaQueryList>,
     pub rules: Vec<Rule>,
@@
 pub struct Manifest {
-  -- pub features: Vec<String>,
-  ++ pub features: HashSet<String>,
+  <@\textcolor{red}{--}@> pub features: Vec<String>,
+  <@\textcolor{green}{++}@> pub features: HashSet<String>,
 }
@@
 pub struct Anchor {
-    -- point: Point<usize>,
-    ++ point: Point<isize>,
+    <@\textcolor{red}{--}@> point: Point<usize>,
+    <@\textcolor{green}{++}@> point: Point<isize>,
     side: Side,
@@
   Resolved(ModuleSpecifier),
-  -- Err(String),
-  ++ Err(ResolvedDependencyErr),
+  <@\textcolor{red}{--}@> Err(String),
+  <@\textcolor{green}{++}@> Err(ResolvedDependencyErr),
 }
@@
     // Other fields
-    -- pending_views: Vec<ViewId>,
-    ++ pending_views: Vec<(ViewId, Table)>,
+    <@\textcolor{red}{--}@> pending_views: Vec<ViewId>,
+    <@\textcolor{green}{++}@> pending_views: Vec<(ViewId, Table)>,
     peer: Client
@@
 \begin{lstlisting}[language=Rust, style=colouredRust]
--- pub struct EventLoop<T: tty::EventedReadWrite> {
-++ pub struct EventLoop<T: tty::EventedPty> {
+<@\textcolor{red}{--}@> pub struct EventLoop<T: tty::EventedReadWrite> {
+<@\textcolor{green}{++}@> pub struct EventLoop<T: tty::EventedPty> {
     poll: mio::Poll,
@@
 start_plugin_process(
---  manifest.clone(),
-++  manifest,
+<@\textcolor{red}{--}@>  manifest.clone(),
+<@\textcolor{green}{++}@>  manifest,
     self.next_plugin_id(),
@@
 \begin{lstlisting}[language=Rust, style=colouredRust]
--- let field_id = schema.get_or_create(attribute.clone())?;
-++ let field_id = schema.get_or_create(&attribute)?;
+<@\textcolor{red}{--}@> let field_id = schema.get_or_create(attribute.clone())?;
+<@\textcolor{green}{++}@> let field_id = schema.get_or_create(&attribute)?;
 \end{lstlisting}
@@
 let repo = replace(&mut *contents, Processing).inner_repo();
--- let statuses = repo_to_statuses(repo, &self.workdir);
-++ let statuses = repo_to_statuses(&repo, &self.workdir);
+<@\textcolor{red}{--}@> let statuses = repo_to_statuses(repo, &self.workdir);
+<@\textcolor{green}{++}@> let statuses = repo_to_statuses(&repo, &self.workdir);
 \end{lstlisting}
@@
     let mut enum_template = EnumTemplate::new();
-    -- enum_template.set_message_enum(&e);
-    ++ enum_template.set_message_enum(e);
+    <@\textcolor{red}{--}@> enum_template.set_message_enum(&e);
+    <@\textcolor{green}{++}@> enum_template.set_message_enum(e);
     // Other statements
@@
 for p in wixobjs {
---  args.push(p.to_string());
-++  args.push((*p).to_string());
+<@\textcolor{red}{--}@>  args.push(p.to_string());
+<@\textcolor{green}{++}@>  args.push((*p).to_string());
 }
@@
         .iter()
---      .map(|s| (*s).to_string())
-++      .map(|s| s.to_string())
+<@\textcolor{red}{--}@>      .map(|s| (*s).to_string())
+<@\textcolor{green}{++}@>      .map(|s| s.to_string())
         .collect::<HashSet<String>>()
@@
         .parse_module()
---      .map_err(|e| {
-++      .map_err(|mut e| {
+<@\textcolor{red}{--}@>      .map_err(|e| {
+<@\textcolor{green}{++}@>      .map_err(|mut e| {
             e.emit();
@@
 \begin{lstlisting}[language=Rust, style=colouredRust]
--- let mut tx = tx.clone();
-++ let tx = tx.clone();
+<@\textcolor{red}{--}@> let mut tx = tx.clone();
+<@\textcolor{green}{++}@> let tx = tx.clone();
 \end{lstlisting}
@@
 pub fn expand_delimited_square(
---  children: &Vec<TokenNode>,
-++  children: &[TokenNode],
+<@\textcolor{red}{--}@>  children: &Vec<TokenNode>,
+<@\textcolor{green}{++}@>  children: &[TokenNode],
 ) -> Result<hir::Expression, ParseError> {
@@
 \begin{lstlisting}[language=Rust, style=colouredRust]
--- const QUEUE_SIZES: &'static [u16] = &[QUEUE_SIZE];
-++ const QUEUE_SIZES: &[u16] = &[QUEUE_SIZE];
+<@\textcolor{red}{--}@> const QUEUE_SIZES: &'static [u16] = &[QUEUE_SIZE];
+<@\textcolor{green}{++}@> const QUEUE_SIZES: &[u16] = &[QUEUE_SIZE];
 

### 2022-10-27 sections/methodology.tex
encode to embed

@@
 
-The general idea for automatically finding the pervasive code patterns within a list of programs is to parse the programs and encode the important parts of the derived AST in fixed-size datapoints.
+The general idea for automatically finding the pervasive code patterns within a list of programs is to parse the programs and embed the important parts of the derived AST in fixed-size datapoints.
 Work in this vein makes the assumption that each cluster corresponds to one change pattern~\cite{hanam2016discovering,campos2019discovering,yang2022mining}.
@@
 An ASTDiff may include an arbitrary number of semantic changes, although a best practice is to include only one semantic change in a commit.
-Because that best practice is not universally followed, we are interested in finding the most important change within an ASTDiff, and encoding that change in our datapoints.
+Because that best practice is not universally followed, we are interested in finding the most important change within an ASTDiff, and embedding that change in our datapoints.
 Sections~\ref{sec:path_extraction}~and~\ref{sec:weighting_scheme} provide a detailed explanation of how we select the most important semantic information, which then allowed us to obtain clusters that contained similar datapoints.
@@
 A path represents a sequence of involved non terminals.
-We found it crucial to record the order of nodes in the path, as this gives us the freedom to later record the order in our datapoint encoding.
+We found it crucial to record the order of nodes in the path, as this gives us the freedom to later record the order in our datapoint embedding.
 
-\paragraph{Encoding the data.} A key question is how to transform these paths to a fixed size datapoint.
+\paragraph{Code Embedding.} A key question is how to transform these paths to a fixed size datapoint.
 First, we need to define a fixed set of columns.
 We need fixed-size datapoints, as they are a requirement for both centroid-based and density-based well-known clustering algorithms~\cite{xu2005survey}.
-The number of columns would depend on the amount of information we want to encode in the datapoints.
+The number of columns would depend on the amount of information we want to embed in the datapoints.
 
@@
 That yields a fully order-insensitive representation.
-In such an encoding, for instance, there is no difference between two nested if statements and two if statements beside each other.
-On the other hand, theoretically, for full order-sensitiveness, we would need to encode all possible combinations of items from $\mathtt{NT} \cup \mathtt{BCE}$ as our dimensions, which yields an infeasible $O(n^d)$ dimensions ($n=|\mathtt{NT} \cup \mathtt{BCE}|$ and $d$ is the depth of nested structures); this could potentially be reduced somewhat, but is still impractical.
-While we can implement a specific workaround to distinguish nested ifs (and we have done so), the general point about the large dimensionality of fully order-sensitive encodings still holds.
+In such an embedding, for instance, there is no difference between two nested if statements and two if statements beside each other.
+On the other hand, theoretically, for full order-sensitiveness, we would need to embed all possible combinations of items from $\mathtt{NT} \cup \mathtt{BCE}$ as our dimensions, which yields an infeasible $O(n^d)$ dimensions ($n=|\mathtt{NT} \cup \mathtt{BCE}|$ and $d$ is the depth of nested structures); this could potentially be reduced somewhat, but is still impractical.
+While we can implement a specific workaround to distinguish nested ifs (and we have done so), the general point about the large dimensionality of fully order-sensitive embedding still holds.
 
@@
 For instance, the purple (right-most) path in Figure~\ref{fig:extraction} could be described as a change in a local variable declaration; a change in a function call; calling a method of one of the function arguments; or calling clone() on one of the function arguments.
-The last description is the most useful one, and we designed our weighting scheme to encode this value judgment.
+The last description is the most useful one, and we designed our weighting scheme to embed this value judgment.
 

### 2022-10-27 sections/results.tex
fixed rq1 numbers
encode to embed

@@
 
-As discussed in the previous section, we used an encoding mechanism to incorporate AST information in fixed-sized datapoints which then could be used in the DBSCAN clustering algorithm.
-We used heuristics (weighting scheme) while encoding our datapoints to make the most important program elements stand out between all the non terminals seen in the AST (alongside some BC-related elements).
-Here we want to evaluate the effectiveness of our encoding.
+As discussed in the previous section, we used an embedding mechanism to incorporate AST information in fixed-sized datapoints which then could be used in the DBSCAN clustering algorithm.
+We used heuristics (weighting scheme) while creating our datapoints to make the most important program elements stand out between all the non terminals seen in the AST (alongside some BC-related elements).
+Here we want to evaluate the effectiveness of our code embedding.
 
-The general idea for our evaluation is whether the most important program elements get the high values in our encoding.
+The general idea for our evaluation is whether the most important program elements get the high values in our embedding.
 The analyzer compares the actual change in the code and compares it to the visual representation of its respective data point.
@@
 For this purpose, we used the packcircles python library\footnote{\url{https://github.com/mhtchan/packcircles}}.
-In these figures, the radius of the circle of each non terminal shows its importance in our encoding (similar to Figure~\ref{fig:essence}).
+In these figures, the radius of the circle of each non terminal shows its importance in our embedding (similar to Figure~\ref{fig:essence}).
 
@@
 
-We carried out two manual experiments for evaluating our encoding mechanism.
+We carried out two manual experiments for evaluating our embedding mechanism.
 In the first experiment, the analyzer investigates whether the most important element has the largest radius or not.
@@
 Both authors carried out these experiments independently.
-At last, we report the results as the portion of the changes which both authors agreed upon its encoding correctness.
-The final results are 55\% and 78\% correctness in the first and second experiments, respectively, which shows the effectiveness of our encoding mechanism.
+At last, we report the results as the portion of the changes which both authors agreed upon its embedding correctness.
+The final results are 54\% and 80\% correctness in the first and second experiments, respectively.
 

### 2022-10-29 main.tex
abstract

--- 2022-10-13 main.tex
@@
 \begin{abstract}
-Insert your abstract here. Include keywords, PACS and mathematical
-subject classification numbers as needed.
+    Rust is a relatively new programming language which allows programmers to write programs that have low-level control over resources, while still ensuring high-level safety guarantees (for programs written in safe Rust).
+Rust's ownership framework enables programs to meet these two seemingly-contradictory goals.
+The Rust compiler's Borrow-Checker component enforces the ownership framework requirements which ensure Rust's safety guarantees.
+
+Rust is popular: as of 2022, it has ranked first, for the sixth consecutive year, in Stack Overflow's annual Developer Survey as the most-loved programming language.
+The number of Rust developers is growing as the need for faster and safer software increases.
+
+Yet, to the best of our knowledge, no research has sought to identify the most pervasive bug fix patterns within Rust programs.
+In this project, using a novel program modelling method, we mined the top 18 most-starred Rust projects in Github to discover the most common bug fix patterns committed to their repositories.
+We analyzed 87,726 code changes drawn from 57,214 commits across these 18 projects.
+After clustering the code changes, and conducting a manual analysis, we discovered 19 groups of cross-project bug fix patterns, which we categorize as (1) general patterns and (2) borrow-checker-related patterns.
+We describe each of these patterns and their implications to automated program repair.
 \keywords{First keyword \and Second keyword \and More}

### 2022-10-29 sections/introduction.tex
encode to embed
introduction

@@
-Producing bugs has been an unavoidable part of computer programming.
-Researchers have been trying to help software engineers to be more aware of the bug producing practices in development and how to step towards creating bug-free software.
-One useful step towards that goal would be studying common bug patterns (and their corresponding fixes) that occur across multiple software systems.
-Research efforts for categorizing common bug patterns date back to 1975 where Endres tried to categorize various bug patterns which occurred in operating system implementations~\cite{endres1975analysis}.
+Bugs have long been an unavoidable part of computer programming, predating Grace Hopper's literal debugging session in 1947; even today, bug-free software remains an aspirational goal for software engineering research.
+One step towards that goal has been studying common bug patterns (and their corresponding fixes) across multiple software systems.
+Research efforts categorizing common bug patterns date back to at least 1975, where Endres tried to categorize bug patterns in operating system implementations~\cite{endres1975analysis}.
 
-In this work, we empirically analyze the most popular Rust projects in GitHub, aiming to discover common bug patterns that these projects share.
-Our interest comes from a desire to develop program repair tools, but we believe these insights can also be of broader service.
-Automated program repair~\cite{goues2019automated,liu2018survey} tools try to locate bug locations in the source code (using a fault localization module) and then generate patches to fix them, possibly without human intervention, so that the modified program meets desired specifications (implicit or explicit).
+In this work, we empirically analyze the most popular Rust projects in GitHub, aiming to discover common bug patterns that afflict these projects.
+Our interest comes from a desire to develop automated program repair tools, but we believe these insights can also be of broader service.
+Automated program repair~\cite{goues2019automated,liu2018survey} tools try to find bug locations in the source code (using a fault localization module) and then generate patches to fix them, possibly without human intervention, so that the modified program meets desired specifications (implicit or explicit).
 A powerful automated program repair system has many useful applications and could significantly simplify debugging, thus reducing software development cost~\cite{le2012systematic}.
-A set of common bug fix patterns can be a useful input to an APR system: it directly provides information about fixing similar buggy structures, which can significantly reduce the search space for patch generation~\cite{jeffrey2009bugfix}.
+A set of common bug fix patterns can be a useful input to an APR system: it tells how to fix recurring buggy structures, thus reducing the search space for patch generation~\cite{jeffrey2009bugfix}.
 
-Similarly, if we are observing undesired behaviour from our program, that would indicate that there lies a fault within our program.
+If we observe undesired behaviour from our program, then there is likely a fault within our program\footnote{At the extreme, a fault could be caused by moths or other hardware malfunctions.}.
 But where is the fault?
-Fault localization~\cite{wong2016survey} is the process of automatically finding the faulty statement(s) in our program.
-If a fault localization tool knows about the common bug patterns that can appear in the underlying programming language, it can prioritize code segment candidates based on their potentials of producing bugs.
-That would result in a more robust tool that leverages historical data.
+Fault localization~\cite{wong2016survey} is the process of automatically finding faulty statements in a program.
+If a fault localization tool knows about the common bug patterns that can appear in the underlying programming language, it can prioritize candidates based on their bug-producing potential.
+That would result in a more robust tool---one that leverages historical data.
 
-Moreover, knowing which code patterns in a programming language are more susceptible to bugs can help programmers to be more cautious while working with similar code structures.
-This knowledge can make the developers avoid certain mistakes during production which ultimately would result in reducing the frequent bug patterns within the program.
-As Monperrus describes, the fewer frequent bug patterns we keep in our software, the less cost we would have to pay for software maintenance~\cite{monperrus2014critical}.
+Moreover, knowing that certain code patterns in a programming language are more susceptible to bugs enables projects to set policies that avoid such patterns.
+For instance, in the context of C and C++, Cannon et al~\cite{Cannon1997RecommendedCS} recommend that, if at least one of the if/else sections is a compound statement, requiring braces, then both sections should have braces (i.e.
+they should be fully bracketed).
+This recommendation presumably comes from qualitative experience, as it pre-dates large-scale empirical studies.
+A project can choose to require that contributions follow such recommendations.
+Research such as ours can help formulate data-driven recommendations that reduce the number of patterns empirically linked to bugs.
+Per Monperrus, the fewer frequent bug patterns we keep in our software, the lower cost we can expect to pay for software maintenance~\cite{monperrus2014critical}.
 
-Moving further afield, researchers have been exploiting deep learning techniques to learn about the semantics of computer programs and create program modifier models that have a specific goal (e.g.
-bug fix)~\cite{alon2019code2vec,alon2018general,raychev2016learning,bielik2016phog}.
-However, most learning tools get fixed-size vectors as input and map them to outputs.
-Unlike image data, it is challenging to compress programs into vectors without any semantic information loss.
-Current approaches try to define the vector dimensions in a way that accounts for all possible abstract syntax tree forms.
+Moving further afield, researchers have been exploiting deep learning techniques to create program modifier models, which can then be used for specific goals (e.g.
+bug fixes)~\cite{alon2019code2vec,alon2018general,raychev2016learning,bielik2016phog}.
+However, most extant learning tools take fixed-size vectors as inputs.
+The process of compressing a variably-sized program to a fixed-sized vector is called code embedding~\cite{chen2019literature}.
+Unlike with image data, it is challenging to compress programs into vectors without any semantic information loss.
+Current approaches try to define the dimensions of the fixed-sized vectors in a way that accounts for all possible abstract syntax tree forms.
 That results in a large, sparse space for the inputs.
-Assuming that the deep learning model we are trying to develop is associated with bug fixing tasks, then knowledge about common bug fix patterns can help prune the input space towards containing fewer dimensions.
-That can help with creating more efficient models.
+Assuming that the deep learning model we are trying to develop is used for bug fixing, then knowledge about common bug fix patterns can help prune the dimensionality of the input space, which can help create more efficient embeddings.
 
@@
 In this work, we study common bug patterns that appear in Rust projects.
-Rust is a relatively new language that allows programmers to write programs that have low-level control over resources, as well as high-level safety guarantees~\cite{klabnik2019rust}.
+Rust is a relatively new programming language~\cite{klabnik2019rust} which allows programmers to write programs that have low-level control over resources, while still ensuring high-level safety guarantees (for programs written in safe Rust).
 What distinguishes Rust from the other systems programming languages, e.g.
@@
 The ownership framework keeps track of references to different memory locations, and will free them once they go out of scope.
-If a program passes the compiler’s checks, it must meet several safety properties; e.g.
+If a program passes the compiler’s checks, it meets several safety properties by construction; e.g.
 it has no dangling references or double-frees.
-These problems had to be manually ensured in languages like C/C++.
-Moreover, not only does the ownership concept remove the need for a garbage collector in the runtime environment, it does not require the programmer to explicitly do certain safety tasks~\cite{qin2020understanding}.
-Due to the memory and thread safety guarantees that Rust provides, many software companies adopted Rust in production.
+Such properties had to be manually ensured (or not!) in languages like C/C++.
+Moreover, not only does ownership remove the need for a garbage collector in the runtime environment, it frees the programmer from the need to explicitly do certain safety-related tasks~\cite{qin2020understanding}.
+Due to the memory and thread safety guarantees that Rust provides, many software companies have adopted Rust in production.
 For example, Dropbox decided to migrate to Rust for developing their storage system, since they were not completely satisfied with the features provided by AWS’s S3 or Golang~\cite{dropbox}.
 
-In Rust, the so-called Borrow Checker component enforces ownership rules.
-Because of the Borrow Checker's importance in Rust, we split patterns into two groups: (1) non-borrow-checker related patterns (General Patterns), and (2) borrow-checker related patterns (BC-Related Patterns).
+The Rust compiler's Borrow-Checker component enforces ownership rules and hence many of Rust's safety properties.
+Because of the Borrow-Checker's importance in Rust, we split the patterns that we present into two groups: (1) non-borrow-checker related patterns (General Patterns), and (2) borrow-checker related patterns (BC-Related Patterns).
 \\
 
-\noindent \textbf{Research Questions} We present our discovered insights of our empirical study in the following four research
-questions (RQs):
+\noindent \textbf{Research Questions} We formulated three research questions (RQs) which guided our empirical study about Rust bug patterns:
 
 \begin{itemize}
-    \item \textbf{RQ1. Does our data modelling approach capture the most important aspect of program changes?} 
+    \item \textbf{RQ1. Does our data modelling approach capture the most important aspects of bug-fixing program changes?} 
     \item \textbf{RQ2. What are the general fix patterns in Rust, and how often do they apply?} 
@@
 
-\noindent \textbf{Contribution} This paper makes the following primary contributions:
+\noindent \textbf{Contributions} This paper makes the following primary contributions:
 
 \begin{itemize}
-    \item To the best of our knowledge, this work is the first project that exploits a code analysis pipeline to automatically mine Rust open source projects and extract the common cross project bug fix patterns.
+    \item To the best of our knowledge, this work is the first to exploits a code analysis pipeline to automatically mine Rust open source projects and extract common cross-project bug fix patterns.
     \item We propose a novel method to encode key AST information in fixed size datapoints using a semi-automatically derived weighting scheme.
-    \item To respect the role of borrow checker in Rust, our paper reports and discusses the bug fix patterns that are related to the ownership framework in Rust separate from the general patterns.
+    \item To respect the key role of the borrow checker in Rust, we report and discuss bug fix patterns that are related to the ownership framework in Rust separately from general patterns.
 \end{itemize}

### 2022-10-29 main.tex
more explicit about goal

@@
 Yet, to the best of our knowledge, no research has sought to identify the most pervasive bug fix patterns within Rust programs.
-In this project, using a novel program modelling method, we mined the top 18 most-starred Rust projects in Github to discover the most common bug fix patterns committed to their repositories.
+In this project, we developed a novel embedding of Rust code into fixed-sized vectors.
+Using this embedding, we mined the top 18 most-starred Rust projects in Github to discover the most common bug fix patterns committed to their repositories.
 We analyzed 87,726 code changes drawn from 57,214 commits across these 18 projects.

### 2022-10-29 sections/introduction.tex
more explicit about goal

@@
-Bugs have long been an unavoidable part of computer programming, predating Grace Hopper's literal debugging session in 1947; even today, bug-free software remains an aspirational goal for software engineering research.
-One step towards that goal has been studying common bug patterns (and their corresponding fixes) across multiple software systems.
+Bugs have long been an unavoidable part of computer programming, predating Grace Hopper's literal debugging session in 1947; even today, bug-free software remains an aspirational aim for software engineering research.
+One step towards that aim has been studying common bug patterns (and their corresponding fixes) across multiple software systems.
 Research efforts categorizing common bug patterns date back to at least 1975, where Endres tried to categorize bug patterns in operating system implementations~\cite{endres1975analysis}.
 
-In this work, we empirically analyze the most popular Rust projects in GitHub, aiming to discover common bug patterns that afflict these projects.
-Our interest comes from a desire to develop automated program repair tools, but we believe these insights can also be of broader service.
+In this work, we empirically analyze the most popular Rust projects in GitHub.
+Our goal is to discover common bug patterns that afflict these projects.
+Although we are motivated by a desire to develop automated program repair tools, we believe these insights can also be of broader service.
 Automated program repair~\cite{goues2019automated,liu2018survey} tools try to find bug locations in the source code (using a fault localization module) and then generate patches to fix them, possibly without human intervention, so that the modified program meets desired specifications (implicit or explicit).

### 2022-10-29 main.tex
two warnings gone, small fix

@@
 
-Rust is popular: as of 2022, it has ranked first, for the sixth consecutive year, in Stack Overflow's annual Developer Survey as the most-loved programming language.
+Rust is popular: as of 2022, it has ranked first, for the seventh consecutive year, in Stack Overflow's annual Developer Survey as the most-loved programming language.
 The number of Rust developers is growing as the need for faster and safer software increases.

### 2022-10-29 sections/methodology.tex
minor edit

@@
-A code change is a modification in a program that changes the program's abstract syntax tree.
-A change pattern is a series of code changes that serves a specific purpose.
+We define a code change to be a modification to a program's source code that changes the program's abstract syntax tree (AST).
+A change pattern is a set of code changes that serves a specific purpose.
 This purpose can be improving code readability or maintainability (refactoring patterns), or rectifying program behaviour with respect to a specific functional or non-functional requirement (bug-fixing patterns).
-In this work, we are interested in latter group.
-Thus, from now on, by referring to code patterns we mean bug-fixing patterns.
+In this work, we are interested in bug-fixing patterns.
+Thus, from now on, whenever we refer to code patterns, we mean bug-fixing patterns.
 
-The general idea for automatically finding the pervasive code patterns within a list of programs is to parse the programs and embed the important parts of the derived AST in fixed-size datapoints.
-Work in this vein makes the assumption that each cluster corresponds to one change pattern~\cite{hanam2016discovering,campos2019discovering,yang2022mining}.
+The general idea for automatically finding the pervasive code patterns within a list of programs is to parse the programs and, for us, to embed the important parts of the derived AST in fixed-size datapoints.
+Then, using a clustering algorithm, we can group these datapoints in different clusters.
+Work in this vein makes the assumption that each cluster corresponds to one class of change pattern~\cite{hanam2016discovering,campos2019discovering,yang2022mining}.
 
-Our pipeline for finding pervasive patterns in Rust programs also follows this idea.
+Our pipeline for finding classes of pervasive patterns in Rust programs also follows this idea.
 We decided to implement our pipeline completely in Python.
@@
 Tree diff algorithms can compute the differences between two arbitrary trees.
-If the trees are programs' abstract syntax trees, we call their diff an ASTDiff.
+When the trees are programs' abstract syntax trees, we call their diff an ASTDiff.
 We use dictdiffer to obtain ASTDiffs.
@@
 Then, we applied the DBSCAN clustering algorithm~\cite{ester1996density} on the resulting datapoints.
-In Section~\ref{sec:clustering_data} we justify our choice of DBSCAN for clustering, and how we tuned it to improve cluster quality.
+In Section~\ref{sec:clustering_data} we justify our choice of DBSCAN for clustering, and we explain how we tuned it to improve cluster quality.
 

### 2022-10-29 methodology.tex
flattened everything ;((

@@
-The general idea for automatically finding the pervasive code patterns within a list of programs is to parse the programs and encode the important parts of the derived AST in fixed-size datapoints.
-We need fixed-size datapoints, as it is a requirement for both centroid-based and density-based well-known clustering algorithms.
-Work in this vein makes the assumption that each cluster corresponds to one change pattern.
+We define a code change to be a modification to a program's source code that changes the program's abstract syntax tree (AST).
+A change pattern is a set of code changes that serves a specific purpose.
+This purpose can be improving code readability or maintainability (refactoring patterns), or rectifying program behaviour with respect to a specific functional or non-functional requirement (bug-fixing patterns).
+In this work, we are interested in bug-fixing patterns.
+Thus, from now on, whenever we refer to code patterns, we mean bug-fixing patterns.
 
-Our pipeline for finding pervasive patterns in Rust programs also follows this idea.
+The general idea for automatically finding the pervasive code patterns within a list of programs is to parse the programs and, for us, to embed the important parts of the derived AST in fixed-size datapoints.
+Then, using a clustering algorithm, we can group these datapoints in different clusters.
+Work in this vein makes the assumption that each cluster corresponds to one class of change pattern~\cite{hanam2016discovering,campos2019discovering,yang2022mining}.
+
+Our pipeline for finding classes of pervasive patterns in Rust programs also follows this idea.
 We decided to implement our pipeline completely in Python.
 To make this work, we need a way to get an AST as a Python dictionary.
-Syn \footnote{\url{https://crates.io/crates/syn}} is a Rust crate which is designed to be employed in Rust procedural macros.
+Syn\footnote{\url{https://crates.io/crates/syn}} is a Rust crate which aims to serve Rust procedural macros.
 However, it includes a parser which is suitable for our purposes; we simply had to write a preprocessor to transform Syn AST output into Python dictionaries.
 
-% TODO Define change
-A change pattern categorizes a class of changes. We thus need to define what a change is. For our purposes, a change is ....
 To compute the contents of a change, we need two code revisions: the revision before the commit, and the revision after it.
 After parsing these two revisions, we will have two ASTs, encoded in two Python dictionaries.
-A tree diff algorithm would specify the differences between two arbitrary trees.
-If the trees are abstracts syntax trees of programs, we simply call it an ASTDiff.
-We use dictdiff to obtain ASTDiffs.
+Tree diff algorithms can compute the differences between two arbitrary trees.
+When the trees are programs' abstract syntax trees, we call their diff an ASTDiff.
+We use dictdiffer to obtain ASTDiffs.
 
-An ASTDiff may include an arbitrary number of semantic changes.
-We are interested in finding the most important change within an ASTDiff and encoding that change in our datapoints.
-Sections ?
-and ?
-provide a detailed explanation of how we select the most important semantic information, which then helped us to obtain clusters that contained similar datapoints.
+An ASTDiff may include an arbitrary number of semantic changes, although a best practice is to include only one semantic change in a commit.
+Because that best practice is not universally followed, we are interested in finding the most important change within an ASTDiff, and embedding that change in our datapoints.
+Sections~\ref{sec:path_extraction}~and~\ref{sec:weighting_scheme} provide a detailed explanation of how we select the most important semantic information, which then allowed us to obtain clusters that contained similar datapoints.
 
-We used Pydriller, which is a repository miner for Python.
-Our target repositories were the top 18 most starred open source Rust projects on GitHub.
-We mined all the bug related commits and ran them through our pipeline.
-Then, we applied the DBSCAN clustering algorithm on the obtained datapoints.
-In Section ?
-we discuss why we chose DBSCAN for clustering and how we tuned it to improve clusters quality.
+Our target repositories were the top 18 most-starred open source Rust projects on GitHub.
+We mined (using Pydriller) all the bug related commits and ran them through our pipeline.
+We will provide more details on the aforementioned process later.
+Then, we applied the DBSCAN clustering algorithm~\cite{ester1996density} on the resulting datapoints.
+In Section~\ref{sec:clustering_data} we justify our choice of DBSCAN for clustering, and we explain how we tuned it to improve cluster quality.
 
-\subsection{Data modelling}
+\subsection{\label{sec:data_modelling}Data modelling}
 
-\subsubsection{Parsing Programs}
+\subsubsection{\label{sec:parsing_programs}Parsing Programs}
 
@@
 Syn handles all of Rust.
-Syn's nonterminals are primarily divided and implemented through two groups: Structs and Enums, which can be found on Syn webpage.
-The union of the item described in these two groups constitutes the list of our non terminals (\verb+NT+).
-The dimensions of our datapoints have been partly designed in a way that they adhere to the elements of \verb+NT+.
-The other part of our dimenstions is comprised of BC-related lexemes (\verb+BCL+), e.g.
-clone, Rc, Box, etc.
-The full list of \verb+BCL+ elements can be found within our artifact.
-We define dimension items as \verb+DI = NT+ $\cup$ \verb+BCL+, as this set also equals the columns of our tables.
+Each of our datapoints specifies whether any Syn non terminal (\texttt{NT}) is present or absent, i.e.
+containing one dimension per non terminal.
+However, we are also interested in change patterns that involve the borrow checker (BC), so we also add a dimension reflecting the presence of a set of BC-related elements (BCE) that we identified, e.g.
+\texttt{clone}, \texttt{Rc}, \texttt{Box}.
+The full list of \verb+BCE+ elements can be found within Appendix~\ref{sec:appendix}.
 
-We parse two versions of a Rust changed file in a commit; the file before the commit and after it.
-We will obtain two trees created by Syn.
-The change that transforms the first tree to the second one yields the fix which was applied in the commit.
-To find this difference, we use dictdiff, a Python library to find the diff of two Python dictionaries.
-Using the PLY tool (a Python implementation of lex and yacc), we wrote a simple transformer from the Rust AST into Python dictionaries.
-After obtaining the syntax trees in Python dictionary format, we can use dictdiff to determine the ASTDiff.
+We parse two versions of a changed Rust file in a commit: the file before the commit and after it.
+This yields two Syn trees.
+Using the PLY tool\footnote{\url{https://www.dabeaz.com/ply/}} (a Python implementation of lex and yacc), we wrote a simple transformer from the serialized Rust AST to Python dictionaries.
+The change that transforms the first tree to the second one is the fix that was applied in the commit.
+To find this difference, we use dictdiffer\footnote{\url{https://dictdiffer.readthedocs.io/en/latest/}}, a Python library to find the diff of two Python dictionaries.
+Its output, in our context, is the ASTDiff.
 
-dictdiff tool provides modifications' types: 
+\subsubsection{\label{sec:path_extraction}Path Extraction}
+
+\begin{figure}[h]
+\centering
+\begin{tikzpicture}[scale=0.7, transform shape,every node/.style={draw},minimum height=0.7cm]
+
+    %% Left AST
+
+    \draw (-1.5,1.5) node[shape=rectangle, minimum width=3cm,minimum height=5.4cm, rounded corners=1ex, xshift=1.5cm, yshift=-2.5cm](context){};
+
+    \node[draw, minimum height=0.5, fill=yellow!20] at (-0.74,1.7) {ASTDiff};
+
+    \node[draw=none] at (0,1) {Context};
+
+    \draw (0,0) node[](a1){ItemFn};
+    \draw (0,-1) node[](a2){Local};
+    \draw (0,-2) node[](a3){ExprCall};
+    \draw (0,-3) node[](a4){args};
+
+    \draw (-2,-5) node[shape=circle](null){null};
+
+    \draw (0,-5) node[](b1){ExprReference};
+    \draw (0,-6) node[](b2){Path};
+    \draw (0,-7) node[](b3){Ident};
+    \draw (0,-8) node[](b4){tail};
+
+    \draw (3,-5) node[](c1){ExprMethodCall};
+    \draw (3,-6) node[](c2){Path};
+    \draw (3,-7) node[](c3){Ident};
+    \draw (3,-8) node[](c4){head};
+
+    \draw (5,-6.5) node[](d1){Ident};
+    \draw (5,-7.5) node[](d2){clone};
+
+    \draw[->]  (a1) -- (a2);
+    \draw[->]  (a2) -- (a3);
+    \draw[->]  (a3) -- (a4);
+
+    \draw[->]  (context) -- (null);
+    \draw[->]  (context) -- (b1);
+    \draw[->]  (context) -- (c1);
+
+    \draw[->]  (b1) -- (b2);
+    \draw[->]  (b2) -- (b3);
+    \draw[->]  (b3) -- (b4);
+
+    \draw[->]  (c1) -- (c2);
+    \draw[->]  (c2) -- (c3);
+    \draw[->]  (c3) -- (c4);
+
+    \draw[->]  (c1) -- (d1);
+
+    \draw [->, red] plot [smooth, tension=1] coordinates { (2,1) (2,-3) (1.4,-5) (1.4,-8.5)};
+
+    \draw [->, green] plot [smooth, tension=1] coordinates { (2.5,1) (3,-3) (5,-5) (4,-6.5) (3.7,-8.5)};
+
+    \draw [->, purple] plot [smooth, tension=1] coordinates { (3,1) (3.5,-3) (5.5,-5) (5.5,-8.5)};
+
+    %% Arrow
+
+    \node[draw=none] at (5.1,-2.7) {Path Extraction};
+    \draw [-stealth,double,line width=2,color=blue!40](4.1,-3.2) -- (6.3,-3.2);
+
+    %% Code at top right
+
+    \node (codeA) [fill=yellow!20, rounded corners=1ex] at (10, 0) {%
+    \centering
+    \begin{minipage}{10cm}
+    \centering
+    \begin{lstlisting}[language=Rust, style=colouredRust]
+      fn sample() {
+        // ...
+        -- let receiver = foo(arg1, tail, head)
+        ++ let receiver = foo(arg1, &tail, head.clone())
+        // ...
+      }
+    \end{lstlisting}
+    \end{minipage}%
+    };
+
+    %% Paths at right
+
+    \draw (7.5,-2.5) node[fill=red!30](e1){ItemFn};
+    \draw (7.5,-3.5) node[fill=red!30](e2){Local};
+    \draw (7.5,-4.5) node[fill=red!30](e3){ExprCall};
+    \draw (7.5,-5.5) node[fill=red!30](e4){ExprReference};
+    \draw (7.5,-6.5) node[fill=red!30](e5){Path};
+    \draw (7.5,-7.5) node[fill=red!30](e6){Ident};
+
+    \node[draw=none, color=red] at (7.5,-8.3) {Path 1};
+
+    \draw (10.5,-2.5) node[fill=green!20](f1){ItemFn};
+    \draw (10.5,-3.5) node[fill=green!20](f2){Local};
+    \draw (10.5,-4.5) node[fill=green!20](f3){ExprCall};
+    \draw (10.5,-5.5) node[fill=green!20](f4){ExprMethodCall};
+    \draw (10.5,-6.5) node[fill=green!20](f5){Path};
+    \draw (10.5,-7.5) node[fill=green!20](f6){Ident};
+
+    \node[draw=none, color=green] at (10.5,-8.3) {Path 2};
+    \node[draw=none, color=green] at (10.5,-8.3) {Path 2};
+
+    \draw (13.5,-2.5) node[fill=purple!20](g1){ItemFn};
+    \draw (13.5,-3.5) node[fill=purple!20](g2){Local};
+    \draw (13.5,-4.5) node[fill=purple!20](g3){ExprCall};
+    \draw (13.5,-5.5) node[fill=purple!20](g4){ExprMethodCall};
+    \draw (13.5,-6.5) node[fill=purple!20](g5){Path};
+    \draw (13.5,-7.5) node[fill=purple!20](g6){clone};
+    \node[draw=none, color=purple] at (13.5,-8.3) {Path 3};
+
+    \draw[->]  (e1) -- (e2);
+    \draw[->]  (e2) -- (e3);
+    \draw[->]  (e3) -- (e4);
+    \draw[->]  (e4) -- (e5);
+    \draw[->]  (e5) -- (e6);
+
+    \draw[->]  (f1) -- (f2);
+    \draw[->]  (f2) -- (f3);
+    \draw[->]  (f3) -- (f4);
+    \draw[->]  (f4) -- (f5);
+    \draw[->]  (f5) -- (f6);
+
+    \draw[->]  (g1) -- (g2);
+    \draw[->]  (g2) -- (g3);
+    \draw[->]  (g3) -- (g4);
+    \draw[->]  (g4) -- (g5);
+    \draw[->]  (g5) -- (g6);
+
+    \end{tikzpicture}
+
+    \caption{\label{fig:extraction}The ASTDiff (left) corresponds to the code change in the yellow box (top right). A DFS will identify three paths, shown in three different colors at right.}
+
+\end{figure}
+
+The output of dictdiffer is a list of diffs, where each diff comprises three parts. The first part specifies the modification type:
+
 \begin{itemize}
@@
 \end{itemize}
-dictdiff also determines the context within which the modification has occured.
-This context happens to be the root elements in our file, e.g.
-function blocks, struct blocks, impl blocks, etc.
-The modified content within a context is still in the form of tree.
-Our objective right now is to find primary elements in our diff that describe the semantics of occurred change.
-% TODO rework the above paragraph
 
-\subsubsection{Path Extraction}
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.8\textwidth]{extraction.png}
-\caption{\label{fig:extraction}Path Extraction}
-\end{figure}
-
-The output of dictdiff is a list of diffs, where each diff comprises three parts.
-The first part specifies whether the modification is of type 'add', 'remove', 'change'.
 The second part identifies the context of the modification, that is, the path from the root of the tree to the subtree in which the modification has occured.
@@
 However, our tool would detect both continuous and non-continuous line changes within one scope.
-% TODO provide an example and then transform it into fig:extraction.
 
 Looking at the tree encoded in the content part of each diff, we realized that each path to a leaf represents one contributing element to the change.
-Through a Depth First Search we can collect all of these paths.
-We store only the nodes that exist in the \verb+DI+.
-The outcome of the DFS search is a set of paths.
+A Depth First Search allows us to collect all paths, and we return only the paths that we store in our datapoints (i.e.
+touching $\mathtt{NT} \cup \mathtt{BCE}$).
 
-Figure~\ref{fig:extraction} shows a simple change and its respective paths.
+Figure~\ref{fig:extraction} illustrates path extraction on a simple change.
 The change replaces \verb+tail+ with \verb+&tail+ and \verb+head+ with \verb+head.clone()+.
 That is, after the change, the second argument passes a borrow of \verb+tail+ rather than sending the ownership of \verb+tail+ to the callee; and, in the third argument, instead of sending \verb+head+ to the callee, it sends a clone of it.
-A DFS search on this tree would yield three different paths, which we show in red, green, and purple.
-The path represents the sequence of involved non terminals, which represents one contributing factor to the change.
-We found it crucial to record the order of all path nodes as it gives us freedom to control the order-sensitiveness in our datapoint encoding.
+A DFS on this tree yields three different paths, which we show in red, green, and purple.
+A path represents a sequence of involved non terminals.
+We found it crucial to record the order of nodes in the path, as this gives us the freedom to later record the order in our datapoint embedding.
 
-The next question is how to transform these paths to a fixed size datapoint.
+\paragraph{Code Embedding.} A key question is how to transform these paths to a fixed size datapoint.
 First, we need to define a fixed set of columns.
-The number of columns would depend on the amount of information we want to encode in the datapoints.
+We need fixed-size datapoints, as they are a requirement for both centroid-based and density-based well-known clustering algorithms~\cite{xu2005survey}.
+The number of columns would depend on the amount of information we want to embed in the datapoints.
+
 The most naive approach would be to only report the observed non terminals within the diff.
 That yields a fully order-insensitive representation.
-In such an encoding, for instance, there is no difference between two nested if statements and two if statements beside each other.
-Theoretically, to reach full order-sensitiveness, we would need encode all possible combinations of dimension items as our dimensions, which yields a large number (A loose upper bound would be $n^n$ where $n$ is the number of dimension items).
-Apart from that, not all combinations adhere to Rust syntax.
-A reasonable work around would be to reduce from the full set of nonterminals to a smaller set of categories (e.g.
-A category for larger entites like class or function definitions, and a category for smaller entities like statements and expressions), and then to set the dimension as all possible combinations of these categories (Similar to JS paper, ref).
-This would reduce the number of dimensions and provide more reasonable and syntactically-correct combinations.
-However, this would still yield a sparse dataset.
+In such an embedding, for instance, there is no difference between two nested if statements and two if statements beside each other.
+On the other hand, theoretically, for full order-sensitiveness, we would need to embed all possible combinations of items from $\mathtt{NT} \cup \mathtt{BCE}$ as our dimensions, which yields an infeasible $O(n^d)$ dimensions ($n=|\mathtt{NT} \cup \mathtt{BCE}|$ and $d$ is the depth of nested structures); this could potentially be reduced somewhat, but is still impractical.
+While we can implement a specific workaround to distinguish nested ifs (and we have done so), the general point about the large dimensionality of fully order-sensitive embedding still holds.
 
-In our methodology, we propose a novel method.
-Similar to the first approach, we define the columns of our dataset, as the set of dimension items we collected from Syn.
-However, to achieve an acceptable level of order-sensitiveness, we carried out two auxiliary steps.
-First, we record the number of occurrences of dimension items within the paths at each scope.
+A reasonable work around would be to reduce $n$ from the full set of non terminals to a smaller set of categories (e.g.
+a category for larger entites like class or function definitions, and a category for smaller entities like statements and expressions), and then to record all possible combinations of these categories (similar to~\cite{hanam2016discovering}).
+That approach reduces the number of dimensions and provides more reasonable and syntactically-correct combinations.
+However, it would still yield a sparse dataset.
+
+In our methodology, we propose a novel representation.
+Similar to the first approach, we order-insensitively define the columns of our dataset as the set of items $\mathtt{NT} \cup \mathtt{BCE}$ we collected from Syn.
+However, to simulate order-sensitiveness, we carried out two auxiliary steps.
+First, we record the number of occurrences of items within the paths at each scope.
 The reason behind this decision is because a different order of program elements yields a different number of occurrences of underlying dimension items.
@@
 
-\subsubsection{Weighting Scheme}
-
-Our main objective is that our datapoints manifest the change with high level of precision.
-For instance, the purple (right-most) path in Figure~\ref{fig:extraction} could be described as a change in a local variable declaration; a change in a function call; calling a method of one of the function arguments; or calling clone() on one of the function arguments.
-The last description offers the most accurate description.
-The whole idea behind our weighting scheme is to be able to identify the change accurately.
-
-% I need to reread this key paragraph. - PL
-A simple heuristic is that, the closer we get to the leaves in the ASTDiff, the more important dimension items get.
-Another observation is that the elements closer to the root tend to be repeated in all the paths within one ASTDiff (e.g.
-ItemFn, Local, ExprCall in Figure~\ref{fig:extraction}).
-Exploiting the latter observation, to obtain a basic knowledge about the number of repetitions of dimension items, we mined the last 20 bug related commits of the last 20 most starred rust projects on GitHub, and ran them through of pipeline.
-We recorded the number of occurrences of each dimension item in total.
-Then we gave the value of the inverse of the number of occurrences to each token.
-The intuition is that the higher the number of occurrences of a dimension item, the closer it is usually seen to the root.
-Hence, it probably does not provide us with the most accurate description of the change.
-In addition to that, as we wanted to make sure that our tool captures the patterns related to borrow checker, we manually raised the weights of the dimension items related to borrow checker.
-That is why we characterize our weighting scheme as semi-automatically designed.
-
-\subsection{Mining Repositories}
+\subsubsection{\label{sec:weighting_scheme}Weighting Scheme}
 
@@
     \centering
-\includegraphics[width=1\textwidth]{repos.png}
-\caption{\label{table:repos} Target Repositories}
+    \begin{tikzpicture}[scale=0.7, transform shape, every node/.style={draw},minimum height=0.7cm]
+
+        %% Left AST
+
+        \draw (-1.5,1.5) node[shape=rectangle, minimum width=3cm,minimum height=5.4cm, rounded corners=1ex, xshift=1.5cm, yshift=-2.5cm](context){};
+
+        \node[draw=none] at (0,1) {Context};
+
+        \draw (0,0) node[](a1){ItemFn};
+        \draw (0,-1) node[](a2){Local};
+        \draw (0,-2) node[](a3){ExprCall};
+        \draw (0,-3) node[](a4){args};
+
+        \draw (-2,-5) node[shape=circle](null){null};
+
+        \draw (0,-5) node[](b1){ExprReference};
+        \draw (0,-6) node[](b2){Path};
+        \draw (0,-7) node[](b3){Ident};
+        \draw (0,-8) node[](b4){tail};
+
+        \draw (3,-5) node[](c1){ExprMethodCall};
+        \draw (3,-6) node[](c2){Path};
+        \draw (3,-7) node[](c3){Ident};
+        \draw (3,-8) node[](c4){head};
+
+        \draw (5,-6.5) node[](d1){Ident};
+        \draw (5,-7.5) node[](d2){clone};
+
+        \draw[->]  (a1) -- (a2);
+        \draw[->]  (a2) -- (a3);
+        \draw[->]  (a3) -- (a4);
+
+        \draw[->]  (context) -- (null);
+        \draw[->]  (context) -- (b1);
+        \draw[->]  (context) -- (c1);
+
+        \draw[->]  (b1) -- (b2);
+        \draw[->]  (b2) -- (b3);
+        \draw[->]  (b3) -- (b4);
+
+        \draw[->]  (c1) -- (c2);
+        \draw[->]  (c2) -- (c3);
+        \draw[->]  (c3) -- (c4);
+
+        \draw[->]  (c1) -- (d1);
+
+        %% Tables
+
+        \node[draw, minimum height=0.5, fill=yellow!20] at (-0.74,1.7) {ASTDiff};
+
+        \node[draw, minimum height=0.5, fill=yellow!20] at (4.29,1.73) {Occurrences data};
+
+        \matrix (table1) [table,text width=6em,nodes={scale=0.45,transform shape}] at (8.5,1)
+        {
+        Element & ItemFn  & Local & ExprCall & ExprReference & Path & Ident & ExprMethodCall & clone\\
+        \# & 1 & 1 & 1 & 1 & 2 & 3 & 1 & 1 \\
+        };
+
+        \node[draw, minimum height=0.5, fill=yellow!20] at (4.12,-0.25) {Weighted data};
+
+        \node[draw=none] at (8.2,0) {Multiply};
+        \draw [-stealth,double,line width=1,color=blue!40](7,0.4) -- (7,-0.4);
+
+        \matrix (table2) [table,text width=6em,nodes={scale=0.45,transform shape}] at (8.5,-1)
+        {
+        Element & ItemFn  & Local & ExprCall & ExprReference & Path & Ident & ExprMethodCall & clone\\
+        weight & 0.0004 & 2.21e-05 & 8.68e-05 & 1 & 2.60e-05 & 2.76e-05 & 4.75e-05 & 1 \\
+        };
+
+        \node[draw=none] (circlepack) at (10.5,-6.2)
+            {\includegraphics[width=0.7\textwidth]{circlepack.png}};
+
+        \node[draw=none] at (9,-2.3) {Visualizing Essence};
+        \draw [-stealth,double,line width=1,color=blue!40](7,-1.6) -- (7.5,-3.4);
+
+        \end{tikzpicture}
+
+    \caption{\label{fig:essence}Multiplying the number of occurrences of items observed in the tree by their respective weights, resulting in the essence of change (visualized with a circle pack figure at bottom right).
+Vectors show (above) the number of occurrences and (below) weighted occurrence counts.}
 \end{figure}
 
+Our main requirement is that our datapoints summarize the key changes in ASTDiffs---a datapoint should foreground the most important change in a diff.
+For instance, the purple (right-most) path in Figure~\ref{fig:extraction} could be described as a change in a local variable declaration; a change in a function call; calling a method of one of the function arguments; or calling clone() on one of the function arguments.
+The last description is the most useful one, and we designed our weighting scheme to embed this value judgment.
+
+We made two observations about changes.
+First, we saw that an item (non terminal or BC-related element) that occurs closer to AST leaves tends to be more semantically important.
+Relatedly, items closer to the root tend to be repeated in all the paths within one ASTDiff (e.g.
+ItemFn, Local, ExprCall in Figure~\ref{fig:extraction}).
+We thus want to prioritize items that tend to occur closer to leaves and de-prioritize items that tend to occur closer to roots.
+
+To design our weighting scheme, we collected empirical data about locations of item occurrences in a corpus.
+Specifically, we mined the last 20 bug related commits of the last 20 most-starred Rust projects on GitHub, and ran them through our pipeline.
+We recorded the total number of occurrences of each item.
+Then, if $\# i$ is the number of occurences of $i$, we gave $i$ a weight of $1/\# i$.
+Because items that occur closer to the root also occur more often, per our observation, this gives higher weights to items that occur infrequently, i.e.
+closer to the leaves.
+Items that occur often and have lower weights are, we believe, less important in describing a change.
+
+Furthermore, as we wanted to make sure that our tool captures patterns related to the Rust borrow checker, we manually increased the weights of the \texttt{BCE} items.
+The manual adjustment is why we characterize our weighting scheme as semi-automatically designed.
+
+Multiplying the number of occurrences of all observed items in a change by their respective weights results in a vector of numbers.
+We call this vector the \emph{essence of change}.
+In this vector, the value for each item shows its importance in the change.
+Figure~\ref{fig:essence} shows the process of computing the essence of the code change that we saw in Figure~\ref{fig:extraction}.
+We have visualized this essence vector using a circle pack figure.
+The radius of each circle corresponds to the importance of the item inside it.
+
+\subsection{\label{sec:mining_repositories}Mining Repositories}
+
+\begin{table}
+\caption{\label{table:repos} Target repositories for mining}
+\begin{tabular}{l r r r r}
+    Project Name & Stars(k) & LOC & \# Commits & Avg. Code Churn \\
+    denoland/deno & 85.9 & 146055 & 8026 & 49.22 \\
+    tauri-apps/tauri & 52.5 & 50014 & 3041 & 31.67 \\
+    alacritty/alacritty & 42.4 & 30124 & 2036 & 27.53 \\
+    sharkdp/bat & 37.6 & 9496 & 2457 & 59.13 \\
+    starship/starship & 29.3 & 34888 & 2379 & 25.1 \\
+    rust-lang/rustlings & 30.6 & 5067 & 1466 & 8.92 \\
+    meilisearch/meilisearch & 30.2 & 25981 & 4029 & 279.7 \\
+    sharkdp/fd & 24.9 & 6542 & 1068 & 12.2 \\
+    swc-project/swc & 24.2 & 346501 & 5884 & 44.34 \\
+    yewstack/yew & 24.5 & 44295 & 2280 & 21.79 \\
+    xi-editor/xi-editor & 19.6 & 39292 & 2105 & 28.67 \\
+    AppFlowy-IO/AppFlowy & 28.4 & 57646 & 3321 & 29.49 \\
+    firecracker-microvm/firecracker & 19.6 & 82021 & 3620 & 58.26 \\
+    nushell/nushell & 21.1 & 163562 & 6088 & 22.38 \\
+    ogham/exa & 19.5 & 10852 & 1558 & 11.99 \\
+    SergioBenitez/Rocket & 18.7 & 68952 & 2135 & 19.75 \\
+    rustdesk/rustdesk & 30.2 & 48115 & 2620 & 72.81 \\
+    tokio-rs/tokio & 17.9 & 121400 & 3101 & 28.87 
+\end{tabular}
+\end{table}
+
 Now that we have set up our code analysis pipeline, we can start mining the Rust repositories.
-We target the top 18 most starred Rust projects on GitHub, at the time of data collection.
-The details of these projects are outlined in Table \ref{table:repos}.
-We computed the average code churn of the files within the projects.
-denoland/deno and swc-project/swc are the projects with largest number of LOCs and average code churn.
+We target the top 18 most-starred Rust projects on GitHub, at the time of data collection.
+Table~\ref{table:repos} summarizes our benchmarks.
+Because we are interested in changes, we computed the average code churn of the files within the projects (The code churn is computed using process metrics provided by Pydriller).
+denoland/deno and swc-project/swc are the projects with largest number of LOCs and highest average code churn, respectively.
 Unsurprisingly, we captured a lot of instances from these two projects in our final clusters.
 
-We use Pydriller, a Python library for mining software repositories.
-Pydriller provides APIs to extract commits from a Git repository, and search through different revisions of files.
-Algorithm \ref{alg} shows how we leveraged Pydriller to run the target repositories through our code analysis pipeline, and populate two databases; one for general patterns and one for borrow-checker related patterns.
-The borrow-checker related patterns are the changes that include certain keywords (e.g.
-clone, Box, etc.
-Full list can be found in our artifact).
-For each repository, we search through all the commits that include bug fixing related keywords within their commit messages.
-These keyword for the scope of our project are `bug', `error', `fault', `defect', and `fix'.
+We used the Pydriller~\cite{spadini2018pydriller} library to mine software repositories.
+Pydriller provides APIs to extract commits from a Git repository and to search through different revisions of files.
+Algorithm~\ref{alg} shows how we used Pydriller to run the target repositories through our code analysis pipeline and populate two databases: one for borrow-checker related patterns (which involve items in \texttt{BCE}) and one for general patterns (which don't).
+For each repository, we search all commits that include bug-fixing related keywords within their commit messages.
+Here, we chose bug-fixing related keywords based on the set introduced in~\cite{zhang2018empirical}, excluding `nan' and `inf'.
 
-Next, we collect a pair of revision per each Rust file.
-The pair includes the state of the file before commit and after commit ($f_a, f_b$).
-After parsing each revision and computing the ASTDiff, we can obtain the fixed sized datapoint $DP$.
-If the datapoint contains borrow-checker related keywords, we would put that inside the BC-related code changes table ($D_b$), otherwise we would put it in the general code changes table ($D_g$).
-In both cases, we augment the datapoint with the commit hash, filename, and the scope in which the change has happened.
-In case of BC-related code changes, we also store detected BC-related keyword.
-Now our tables are ready for clustering and categorization.
+Next, we collect a pair of revisions for each Rust file.
+The pair includes the state of the file before the commit and after the commit ($f_a, f_b$).
+After parsing each revision and computing the ASTDiff, we can compute the fixed sized datapoint $\mathit{DP}$.
+If the datapoint contains borrow-checker related keywords, we put it in the BC-related database $D_b$; otherwise, we put it in the general code changes database $D_g$.
+In both cases, we augment the datapoint with the commit hash, filename, and the scope in which the change happened.
+For BC-related code changes, we also store the detected BC-related keywords.
+Now our databases are ready for clustering and categorization.
 
@@
 
-\subsection{Clustering Data}
+\subsection{\label{sec:clustering_data}Clustering Data}
 
-We use DBSCAN clustering algorithm for two main reasons.
-Firstly, it is a density-based clustering method, which means that it detects arbitrarily shaped clusters as opposed to centroid-based methods, like k-means.
-Secondly, also in contrast to k-means, it does not require the number of clusters in advance as an input to the algorithm.
+Having collected our datapoints, our next task is to cluster them so that we can categorize bug patterns.
+Like~\cite{hanam2016discovering}, we use the DBSCAN clustering algorithm for two main reasons.
+Firstly, in contrast to k-means, it does not require the number of clusters in advance as an input to the algorithm.
+Secondly, it is a density-based clustering method, which means that it detects arbitrarily shaped clusters; centroid-based methods like k-means can't detect such clusters.
+We tried k-means and the more powerful fuzzy variant Fuzzy C-means~\cite{dunn1973fuzzy} and did not get useful clusters.
 
-There are two parameters that need tuning in DBSCAN.
-First parameter is $\epsilon$ which indicates a radius with which you can decide whether a point is a core point, a border point, or an outlier.
-Second parameter is the minimum number of point you would like each cluster to have.
-We denote his parameter with $Z$.
+DBSCAN takes two tuneable parameters.
+The first parameter $\epsilon$ indicates a radius within which a point is a core point, a border point, or an outlier.
+The second parameter $Z$ is the minimum number of points per cluster.
 We ran DBSCAN using different combinations of these two parameters.
-In Figure \ref{fig:clustering}, we show nine different experiments with their respective parameter values.
-Within each subplot, the number of clustered points, noise points, and the number of clusters have been specified both for $D_g$ and $D_b$.
+In Figure~\ref{fig:clustering}, we show nine different experiments, with their respective parameter values.
+Each subplot specifies, for both $D_g$ and $D_b$, the number of clustered points, noise points, and the number of clusters.
 
+% TODO fix this caption
 \begin{figure}[h]
 \centering
-\includegraphics[width=1\textwidth]{clusters.png}
-\caption{\label{fig:clustering} Clustering Experiments}
+\includegraphics[width=1\textwidth]{clusters.pdf}
+\caption{\label{fig:clustering} Clustering results with nine different combinations of parameters $\epsilon$ and $Z$, respectively separated in columns and rows.
+In each subplot, the two blue bars in the left respectively indicate the number of clustered and noise datapoints (the number is shown above each bar) after applying clustering on $D_g$ (denoted with $G_c$ and $G_n$ on the x axis).
+The two orange bars in the right show similar data but for the $D_{b}$ database.
+In each subplot's legend, $C_G$ and $C_B$ show the number of clusters obtained from $D_g$ and $D_b$, respectively}
 \end{figure}
 
-As a general rule, a lower value for $Z$ would allow more group of close points to be considered as clusters, resulting in a higher number of clusters.
-A large value for $\epsilon$ would widen the search radius, allowing more points to fall in the clusters which results in a reduction of the number of noise points.
-That possibly can put points that do not completely manifest same patterns inside a cluster, resulting in a ineffective clustering.
-Alternatively, a small value for $\epsilon$ would tighten the ring of search, possibly separating clusters which essentially manifest same code changes.
-Also if $\epsilon$ is really small, the clustering algorithm might not detect some clusters at all, resulting in a reduction of the number of clusters.
+As a general rule, lower $Z$ would allow more groups of close points to be considered as clusters, resulting in a higher number of clusters.
+A large $\epsilon$ would widen the search radius, allowing more points to fall in the clusters, which results in a reduction of the number of noise points.
+Such a choice might put points from different patterns inside the same cluster, resulting in a ineffective clustering.
+Alternatively, a small $\epsilon$ would tighten the ring of search, possibly separating clusters which essentially manifest the same code changes.
+Also, if $\epsilon$ is really small, the clustering algorithm might not detect some clusters at all, resulting in a reduction of the number of clusters.
 
-\subsubsection{Manual Analysis for parameter tuning}
+\subsubsection{\label{sec:manual_analysis_parameter_tuning}Manual Analysis for parameter tuning}
 
-That is why, we carried out a manual analysis of clusters.
-In the manual analysis we randomly pick 50 clusters and from each cluster we randomly choose 10 datapoints.
-The analyzer then looks at the code and specify the clusters that contained more than five datapoints showing similar patterns.
-Through the manual analysis we specified that $\epsilon=0.0001, Z=5$ and $\epsilon=0.001, Z=5$ acted as the best parameters for $D_g$ and $D_b$, resulting in 596 and 61 clusters, respectively.
-As the final step, we exclude the clusters that contained datapoints from less than 3 projects.
-That is because, we wanted our clusters to contain cross-project patterns as much as possible.
+That is why we carried out a manual analysis of clusters.
+In the manual analysis, we randomly pick 50 clusters; from each, cluster we randomly choose 10 datapoints.
+The person analyzing the cluster then looks at the code and specifies the clusters that contained more than five datapoints showing similar patterns.
+For the manual analysis, we used $\epsilon=0.0001, Z=5$ and $\epsilon=0.001, Z=5$, which were the best parameters for $D_g$ and $D_b$, resulting in 577 and 102 clusters, respectively.
+As a final step, we exclude clusters that contained datapoints from fewer than 3 projects.
+That is because we wanted our clusters to contain cross-project (e.g.
+potentially generalizable) patterns as much as possible.
 
-\subsubsection{Manual Analysis for cluster selection}
+\subsubsection{\label{sec:manual_analysis_cluster_selection}Manual Analysis for cluster selection}
 % 3 types, refactoring dropped
 
-Similar to ..., following Cotroneo et al. definitions, we divide the clusters in three different groups:
+Similar to~\cite{yang2022mining}, following Cotroneo et al.~\cite{cotroneo2019analyzing} definitions, we divide the clusters in three different groups:
 
 \begin{itemize}
-    \item bug-fix: "The changed code actually fixes the behavior of software, which can represent fixes for a bug type".
-We consider the changes that improve program performance as instance of this group, as they fix the behavior of software.
+    \item bug-fix: ``The changed code actually fixes the behavior of software, which can represent fixes for a bug type''.
+We include here changes that improve program performance, as they fix the behaviour of the software.
 
-    \item fix-induced: "The changed code is a group of bug-fixing code changes but not represents an actual bug fix, e.g., adding new input parameters to a method, the method signature and method call must be changed corresponding".
+    \item fix-induced: ``The changed code is a group of bug-fixing code changes but not representing an actual bug fix, e.g., adding new input parameters to a method, the method signature and method call must be changed correspondingly''.
 
-    \item refactoring: "the changed code does not modify the software behavior, e.g., better readability or encapsulation".
+    \item refactoring: ``the changed code does not modify the software behavior, e.g., for better readability or encapsulation''.
 \end{itemize}
 
-In this work, we were only interested in the first two group and disregarded the the clusters that manifest a refactoring change.
+In this work, we were only interested in the first two groups and ignored the clusters for refactoring changes.
 Also, we prefer reporting clusters that manifest concrete patterns.
 An abstract pattern as opposed to a concrete pattern can account for many changes in the programs.
-For instance, the pattern `Adding a new statement to a function's body' is deemed more abstract than `Changing a clone of a variable to a borrowing of it in a function argument'.
+For instance, we deem the pattern `Changing a clone of a variable to a borrowing of it in a function argument' more concrete than `Adding a new statement to a function's body':
+
+\begin{lstlisting}[language=Rust, style=colouredRust]
+// First Pattern
+-- foo1(arg.clone);
+++ foo1(&arg);
+
+// Second Pattern
+fn foo2() {
+    stmt1;
+++  stmt2;
+}
+
+\end{lstlisting}
 
 In our manual analysis, we randomly picked 50 datapoints of each cluster.
-If the cluster had less than 50 datapoints, we would analyze all of the datapoints in the cluster.
-Next, we read the code and also the bug report related to all the datapoints (if existed).
-In case that a similar pattern was detected over the instances of a cluster, we would write a natural language description of that cluster.
-After providing all the descriptions, we linked similar clusters together as possible candidates for merging.
-At last, after accomplishing re-examination for merge possibilities, we selected 19 clusters that manifested concrete changes.
-We discuss about our findings in Section \ref{}.
+If the cluster had fewer than 50 datapoints, we would analyze all of the datapoints in the cluster.
+Next, we read the code and the bug reports related to all the datapoints (if they existed).
+If we judged that the members of a cluster fit into a pattern, we wrote a natural language description of that cluster.
+Considering the descriptions, we linked similar clusters together as possible candidates for merging.
+We considered these candidates and carried out appropriate ones.
+Finally, we selected 19 of the remaining clusters---the ones that manifested concrete changes.
+We discuss about our findings in Sections~\ref{sec:common_patterns}~and~\ref{sec:bc_patterns}.

### 2022-10-29 related.tex
flattened everything ;((

--- 2022-10-14 sections/related.tex
@@
 Like~\cite{hanam2016discovering}, we used DBSCAN to cluster fixes.
-However, our novel program encoding approach differs from what they did.
+However, our novel program embedding approach differs from what they did.
 In~\cite{hanam2016discovering}, they introduced Feature Properties table, which is a categorization for program elements.
-Encoding programs based on this table resulted in a order-sensitive encoding, however, it made the datapoints sparse.
+Embedding programs based on this table resulted in a order-sensitive embedding, however, it made the datapoints sparse.
 Moreover, like~\cite{yang2022mining}, we introduced patterns in two different categories.

### 2022-10-29 results.tex
comment about rq1, and rephrase
flattened everything ;((

@@
+In this section, we present the clusters of fix patterns that we found in our data set.
+\RQBC{} is about the borrow-checker related patterns, while \RQG{} is about general patterns.
+However, our technique fundamentally relies on our data modelling approach successfully capturing the most important aspects of program changes.
+We thus first evaluate the data modelling approach as our RQ1, before continuing with \RQBC{} and \RQG{}.
+
 \subsection{Evaluating the data modelling approach (RQ1)}
 
-As discussed in the previous section, we used an encoding mechanism to incorporate AST information in fix-sized datapoints which then could be used in the DBSCAN clustering algorithm.
-We used heuristics (weighting scheme) while encoding our datapoints to make the most important program elements stand out between all the non terminals seen in the AST.
-Here we want to evaluate the effectiveness of our encoding.
+\textbf{Does our data modelling approach capture the most important aspect of program changes?}
 
-The general idea for our evaluation is whether the most important program elements has the high values in our encoding.
+As discussed in the previous section, we used an embedding mechanism to incorporate AST information in fixed-sized datapoints which then could be used in the DBSCAN clustering algorithm.
+We used heuristics (weighting scheme) while creating our datapoints to make the most important program elements stand out between all the non terminals seen in the AST (alongside some BC-related elements).
+Here we want to evaluate the effectiveness of our code embedding.
+
+The general idea for our evaluation is whether the most important program elements get the high values in our embedding.
 The analyzer compares the actual change in the code and compares it to the visual representation of its respective data point.
-We used circle pack figures to visualize each data point.
-In these figures, the radius of the circle of each non terminal shows its importance in our encoding (Fig ref).
+We used circle pack figures to visualize each data point~\cite{collins2003circle}.
+For this purpose, we used the packcircles python library\footnote{\url{https://github.com/mhtchan/packcircles}}.
+In these figures, the radius of the circle of each non terminal shows its importance in our embedding (similar to Figure~\ref{fig:essence}).
 
@@
 The analyzer should give a binary answer to these questions.
-Both authors carry out these experiments independently.
-At last, we report the results as the portion of the changes which both authors agreed upon its encoding correctness.
-The final results are TBD\% and TBD\% correctness in the first and second experiments, respectively, which shows the effectiveness of our encoding mechanism.
+Both authors carried out these experiments independently.
+At last, we report the results as the portion of the changes which both authors agreed upon its embedding correctness.
+The final results are 54\% and 80\% correctness in the first and second experiments, respectively.
 
-\subsection{Common Bug Fix Patterns (RQ2)}
+% no, you have to compute the intersection
+% when is our approach effective and when is it not?
+% more detailed stats would probably be good too
 
-\subsection{Attributes}
+\subsection{\label{sec:common_patterns}Common Bug Fix Patterns (RQ2)}
 
-Attributes in Rust make it possible for the developers to add metadata to program elements.
-These attributes can be added to structs, struct fields, enum variants, match expression arms, etc.
+\textbf{What are the general fix patterns in Rust, and how often do they apply?}
+
+We collected 60583 datapoints for general fix patterns using a server with Xeon Gold 5120 Processor (Q3-2017, 14-core 2.2 to 3.2GHz) and the data collection took around 26 hours to complete for both $D_g$ and $D_b$.
+We used $\epsilon=0.0001$ and $Z=5$ as our DBSCAN parameters and that resulted in 577 clusters.
+After manual analysis as described in~\ref{sec:manual_analysis_cluster_selection} we ended up with 10 general patterns which we introduce here:
+
+\begin{table}[]
+
+\begin{tabular}{l|l|r|r}
+\textbf{Catgegory} & \textbf{ID} & \textbf{\# Instances} & \textbf{\# Projects} \\
+\hline
+\multirow{2}{*}{Attributes} & G.attr.struct & 172 & 16                                  \\
+& G.attr.struct.field & 62 & 16                     \\\cline{1-4}
+\multirow{2}{*}{Struct} & G.struct.field  & 179 & 17\\
+& G.struct.field.pub & 23 & 12   
+\\\cline{1-4}
+Option & G.option & 22 & 12                    \\\cline{1-4}
+\multirow{4}{*}{Types} & G.type.field  & 72 & 14 \\
+& G.type.generic & 46 & 14\\ & G.type.enum.variant  & 39 & 12 \\
+& G.type.tuple & 6 & 6                                         \\\cline{1-4}
+\multirow{1}{*}{Traits} & G.trait.bound & 6 & 5 \\ 
+\end{tabular}
+\caption{\label{table:general}Bug Fix Pattern Categorization (General Patterns)}
+\end{table}
+
+\subsubsection{Attributes}
+
+Attributes in Rust enable developers to add metadata to program elements.
+Developers can add attributes to structs, struct fields, enum variants, match expression arms, etc.
 There are two types of attributes: Inner Attributes and Outer Attributes.
@@
 
-\noindent \textit{ \textbf{G.attr.struct.} Modifying the attributes of structs}
+\noindent\textit{\label{sec:G.attr.struct}\textbf{G.attr.struct} Modifying the attributes of structs}
 
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
+\begin{lstlisting}[language=Rust, style=colouredRust]
 // G.attr.struct.add: Adding attributes to a struct
-++ #[derive(Debug, Clone)]
+<@\textcolor{green}{++}@> #[derive(Debug, Clone)]
 pub struct Call {
@@
 
-\noindent \textbf{Description:} In this pattern, the attribute set of a struct changes.
-This modification can happen in three different ways; (G1.1) Adding attributes to the struct, (G1.2) Removing attributes from the struct, (G1.3) Changing the content of an existing attribute.
-The attributes are Outer Attributes as we can find struct related non-terminals within the ASTDiff of the instances of this pattern.
+\noindent\textbf{Description:} In this pattern, the attribute set of a struct changes.
+This modification can happen in three different ways: (G.attr.struct.add) adding attributes to the struct; (G.attr.struct.drop) removing attributes from the struct; and (G.attr.struct.change) changing the content of an existing attribute.
+These attributes are Outer Attributes as we can find struct related non-terminals within the ASTDiff of the instances of this pattern.
 \\
 
-\noindent \textit{ \textbf{G.attr.struct.field} Modifying the attributes of struct fields}
+\noindent\textit{\textbf{G.attr.struct.field} Modifying the attributes of struct fields}
 
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
+\begin{lstlisting}[language=Rust, style=colouredRust]
 // G.attr.struct.field.add: Adding attributes to struct fields
@@
 
-\noindent \textbf{Description:} In this pattern, the attribute set of a struct field changes.
-As with G1, this modification also happens in three different ways.
+\noindent\textbf{Description:} In this pattern, the attribute set of a struct field changes.
+As with \textit{\textbf{G.attr.struct}}, this modification also happens in three different ways.
 Using attributes for struct fields, apart from being able to store different meta-information for different struct fields, for some use cases, the programmers can control the action they want to apply on different fields.
+% this last sentence immediately prior to here has too many moving parts
 
-\subsection{Struct} 
-\noindent \textit{ \textbf{G.struct.field} Adding/Removing a struct field}
+\subsubsection{Struct} 
+\noindent\textit{\textbf{G.struct.field} Adding/Removing a struct field}
 
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
+\begin{lstlisting}[language=Rust, style=colouredRust]
 // G.struct.field.add: Adding a new field to a struct
@@
 
-\noindent \textbf{Description:} A set of new fields are add to or removed from an existing struct.
-The addition of a new field happens when the developer need to store a new piece of data so they can implement a desired behaviour.
-The removal of a field  happens when the developer realizes that the field is expendable with respect to the behaviour they want to implement.
+\noindent\textbf{Description:} A set of new fields are added to or removed from an existing struct.
+A developer adds a new field to store a new piece of data in a struct.
+A developer removes a field when the developer realizes that the field is not required with respect to the behaviour they want to implement (possibly because they have moved it elsewhere).
 \\
 
-\noindent \textit{ \textbf{G.struct.field.pub} Modifying the access modifiers of struct fields}
+\noindent\textit{\textbf{G.struct.field.pub} Modifying the access modifiers of struct fields}
 
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
+\begin{lstlisting}[language=Rust, style=colouredRust]
 
@@
 
-\noindent \textbf{Description:} Adding pub to a struct field makes it possible to access that field within external modules.
-In the instances of this cluster, adding pub access modifier happened in a bug fixing change, where the developer needed to access a field which had not been granted public access.
-On the other hand, the public access might be revoked if the developer realizes there is no external access to the field.
+\noindent\textbf{Description:} Adding \texttt{pub} to a struct field makes it possible to access that field from external modules.
+In the instances of this cluster, adding the \texttt{pub} access modifier happened in a bug fixing change, where the developer needed to access a field which had not been marked for public access.
+On the other hand, the public access might be revoked if the developer realizes there are no external accesses to the field, and that such accesses are not desirable, e.g.
+for encapsulation purposes.
 
-\subsection{Option}
+\subsubsection{Option}
 \vspace{3mm}
 
-\noindent \textit{ \textbf{G.option} Changing field type T to T$\langle$Option$\rangle$}
+\noindent\textit{\textbf{G.option} Changing field type T to T$\langle$Option$\rangle$}
 
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
+\begin{lstlisting}[language=Rust, style=colouredRust]
 
@@
 
-\noindent \textbf{Description:} Rust is a null safe language, meaning that object references cannot assume null values.
-If we want void values for our object, we would have to use Option type.
-Option is essentially an enum with two variants, Some which indicates that our object has an value, and None shows the absence of any value.
-In this pattern, the developer changes the type of a struct field to an Option of its type.
-This makes it possible for the developer to modify the program in a way that it covers the scenarios in which the the struct field has no value.
+\noindent\textbf{Description:} Rust is a null-safe language, meaning that object references cannot take on null values.
+If we want a variable to contain null, we would have to use the \texttt{Option} type.
+\texttt{Option} is essentially an \texttt{enum} with two variants, \texttt{Some}, which indicates that our variable has a value, and \texttt{None}, which indicates the absence of any value.
+In this change pattern, the developer changes the type of a struct field to an \texttt{Option} of that type, hence enabling them to store no value in that field.
 
-For instance, in the Rust project swc-project/swc (A fast TypeScript/JavaScript compiler), the developer needed to modify the CSS parser source code in a way that it accounts for empty @media queries \footnote{\url{https://github.com/swc-project/swc/commit/75a14f98b7370226115ee24eec6eb8c802bd4837}}, so, as shown in the code snippet above, they changed type MediaQueryList to Option<MediaQueryList>, and modified the other parts of the source code, accordingly.
+For instance, in the Rust project swc-project/swc (a fast TypeScript/JavaScript compiler), the developer needed to modify the CSS parser source code in a way that it accounts for empty @media queries \footnote{\url{https://github.com/swc-project/swc/commit/75a14f98b7370226115ee24eec6eb8c802bd4837}}.
+So, as shown in the code snippet above, they changed type \texttt{MediaQueryList} to \verb+Option<MediaQueryList>+, and modified the other parts of the source code, accordingly.
 
-\subsection{Types}
+\subsubsection{Types}
 \vspace{3mm}
 
-\noindent \textit{ \textbf{G.type.field} Changing a struct field type}
+\noindent\textit{\label{sec:G.type.field}\textbf{G.type.field} Changing a struct field type}
 
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
+\begin{lstlisting}[language=Rust, style=colouredRust]
 pub struct Manifest {
-  -- pub features: Vec<String>,
-  ++ pub features: HashSet<String>,
+  <@\textcolor{red}{--}@> pub features: Vec<String>,
+  <@\textcolor{green}{++}@> pub features: HashSet<String>,
 }
@@
 
-\noindent \textbf{Description:} The type of a struct field changes.
+\noindent\textbf{Description:} The type of a struct field changes.
 This is a common change pattern in all statically-typed languages and can happen for various reasons.
-These reasons can be a step for implementing a new feature or rectifying the behaviour of the program.
+For instance, the developer may be implementing a new feature or fixing the behaviour of the program.
 Also it can happen for refactoring purposes or performance enhancement.
 \\
+% maybe discuss the specific example a bit?
 
-\noindent \textit{ \textbf{G.type.generic} Changing a generic type}
+\noindent\textit{\textbf{G.type.generic} Changing a generic type parameter}
 
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
+\begin{lstlisting}[language=Rust, style=colouredRust]
 #[derive(Debug, Clone, PartialEq)]
 pub struct Anchor {
-    -- point: Point<usize>,
-    ++ point: Point<isize>,
+    <@\textcolor{red}{--}@> point: Point<usize>,
+    <@\textcolor{green}{++}@> point: Point<isize>,
     side: Side,
@@
 
-\noindent \textbf{Description:} This a change of the generic (parametric) type of a struct field. Similar to G7, this change can also happen for the mentioned reasons. \\
+\noindent\textbf{Description:} This a change of the type parameter of a struct field. As with \textit{\textbf{G.type.field}}, this change can also happen for the mentioned reasons. \\
 
-\noindent \textit{ \textbf{G.type.enum.variant} Change in enum variant value type}
+\noindent\textit{\textbf{G.type.enum.variant} Change in enum variant value type}
 
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
+\begin{lstlisting}[language=Rust, style=colouredRust]
 #[derive(Debug, Clone, PartialEq, Eq)]
@@
 
-\noindent \textbf{Description:} In Rust, we can specify the type of the value that an enum variant can store.
-Using this feature, we can avoid using a struct alongside our enum variants to specify the value type.
+\noindent\textbf{Description:} In Rust, developers can specify the type of the value that an enum variant can store.
+Using this feature, they can avoid having to use a struct along with enum variants.
 This change pattern is associated with a change in enum variant types.
@@
 
-\noindent \textit{ \textbf{G.type.tuple} Changing the type inside a tuple}
+\noindent\textit{\textbf{G.type.tuple} Changing the type inside a tuple}
 
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
+\begin{lstlisting}[language=Rust, style=colouredRust]
 #[allow(dead_code)]
@@
 
-\noindent \textbf{Description:} In this pattern, an extra type is added to an existing type, making it a tuple of multiple types.
-Alternatively, the change might be dropping types from the tuple and making it a tuple with fewer number of arguments.
+\noindent\textbf{Description:} In this pattern, an extra type is added to an existing type, making it a tuple of multiple types.
+Alternatively, the change might be dropping types from the tuple and making it a smaller-arity tuple.
 
-\subsection{Traits}
-\noindent \textit{ \textbf{G.trait.bound} TraitBound Change}
+\subsubsection{Traits}
+\noindent\textit{\textbf{G.trait.bound} TraitBound Change}
 
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
--- pub struct EventLoop<T: tty::EventedReadWrite> {
-++ pub struct EventLoop<T: tty::EventedPty> {
+\begin{lstlisting}[language=Rust, style=colouredRust]
+<@\textcolor{red}{--}@> pub struct EventLoop<T: tty::EventedReadWrite> {
+<@\textcolor{green}{++}@> pub struct EventLoop<T: tty::EventedPty> {
     poll: mio::Poll,
@@
 
-\noindent \textbf{Description:} The trait bounds are the functionalities that we require from our parametric types.
-This change pattern related to modification in trait bounds.
+\noindent\textbf{Description:} The trait bounds are the functionalities that we require from our parametric types.
+This change pattern relates to modification in trait bounds.
 This can happen both for bug fix reasons or just refactoring purposes.
 
-\subsection{BC-Related Bug Fix Patterns (RQ3)}
+\subsection{\label{sec:bc_patterns}BC-Related Bug Fix Patterns (RQ3)}
+
+\textbf{What are borrow-checker related fix patterns in Rust, and how often do they apply?}
+
+We collected 27143 datapoints for borrow-checker related fix patterns.
+We used $\epsilon=0.001$ and $Z=5$ as our DBSCAN parameters and that resulted in 102 clusters.
+After manual analysis as described in~\ref{sec:manual_analysis_cluster_selection} we ended up with 9 borrow-checker related fix patterns which we introduce here:
+
+\begin{table}[]
+\begin{tabular}{l|l|r|r}
+\textbf{Catgegory} & \textbf{ID} & \textbf{\# Instances} & \textbf{\# Projects} \\
+\hline
+\multirow{2}{*}{Clone}                                         & BC.clone.drop & 83 & 10 \\
+& BC.clone.ref & 4 & 3   
+\\\cline{1-4}
+\multirow{3}{*}{Ref and Deref} & BC.ref.add & 17 & 7 \\
+& BC.ref.drop & 145 & 14 \\
+& BC.deref.to\_string  & 7 & 4                                            \\\cline{1-4}
+\multirow{2}{*}{Mut} & BC.mut.add  & 11 & 6 \\
+& BC.mut.drop  & 42 & 15                               \\\cline{1-4}
+Vector & BC.vec.slice  & 8 & 6                                    \\\cline{1-4}
+\multirow{1}{*}{Lifetime}      
+& BC.lifetime.static  & 8 & 6
+\\
+\end{tabular}
+\caption{\label{table:bc}Bug Fix Pattern Categorization (BC-Related Patterns)}
+\end{table}
 
@@
 
-\noindent \textit{ \textbf{BC.clone.ref} Dropping clone and adding borrowing}
+\noindent\textit{\textbf{BC.clone.drop} Removing clone}
 
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
--- let field_id = schema.get_or_create(attribute.clone())?;
-++ let field_id = schema.get_or_create(&attribute)?;
-\end{lstlisting}
-
-\noindent \textbf{Description:} Unlike the previous pattern, this pattern is not detected by Clippy.
-The change happens when the programmers realizes that the cloning of a variable is unnecessary, however, we also want to preserve the ownership of the variable within the current scope.
-That is why in this pattern a clone is turned into a borrowing.
-
-Similar to the previous pattern, this does not change the program behaviour and is merely for performance purposes.
-It was the case in one of the commits of the project tauri-apps/tauri, where the clone to borrowing significantly optimized CPU usage.
-They discussed that the cloning of some types such as HashMap was expensive.
-\\
-
-\noindent \textit{ \textbf{BC.clone.drop} Removing clone}
-
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
+\begin{lstlisting}[language=Rust, style=colouredRust]
 start_plugin_process(
---  manifest.clone(),
-++  manifest,
+<@\textcolor{red}{--}@>  manifest.clone(),
+<@\textcolor{green}{++}@>  manifest,
     self.next_plugin_id(),
@@
 
-\noindent \textbf{Description:} This pattern is referred to as redundant clone in Clippy Lints.
-Clippy detects whether a variable clone is necessary and could we have used the variable itself or not.
+\noindent\textbf{Description:} This pattern is referred to as redundant clone in Clippy Lints.
+Clippy\footnote{\url{https://github.com/rust-lang/rust-clippy}} is a linter tool for Rust that catches common mistakes.
+There are more than 500 lints included in Clippy.
+With regards to this pattern, Clippy detects whether a variable clone is necessary and could we have used the variable itself or not.
 This change does not modify the program behaviour, however it can boost the performance significantly if we repeatedly cloning large variables.
+
+\noindent\textit{\textbf{BC.clone.ref} Dropping clone and adding borrowing}
+
+\begin{lstlisting}[language=Rust, style=colouredRust]
+<@\textcolor{red}{--}@> let field_id = schema.get_or_create(attribute.clone())?;
+<@\textcolor{green}{++}@> let field_id = schema.get_or_create(&attribute)?;
+\end{lstlisting}
+
+\noindent\textbf{Description:} Unlike the previous pattern, this pattern is not detected by Clippy.
+The change happens when the programmers realizes that the cloning of a variable is unnecessary, however, we also want to preserve the ownership of the variable within the current scope.
+That is why in this pattern a clone is turned into a borrowing.
+
+Similar to the previous pattern, this does not change the program behaviour and is merely for performance purposes.
+It was the case in one of the commits~\footnote{\url{https://github.com/tauri-apps/tauri/commit/a280ee90af0749ce18d6d0b00939b06473717bc9}} of the project tauri-apps/tauri, where the clone to borrowing significantly optimized CPU usage.
+They discussed that the cloning of some types such as HashMap was expensive.
+\\
 
@@
 
-\noindent \textit{ \textbf{BC.ref.add} Adding Borrowing}
+\noindent\textit{\textbf{BC.ref.add} Adding Borrowing}
 
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
+\begin{lstlisting}[language=Rust, style=colouredRust]
 let repo = replace(&mut *contents, Processing).inner_repo();
--- let statuses = repo_to_statuses(repo, &self.workdir);
-++ let statuses = repo_to_statuses(&repo, &self.workdir);
+<@\textcolor{red}{--}@> let statuses = repo_to_statuses(repo, &self.workdir);
+<@\textcolor{green}{++}@> let statuses = repo_to_statuses(&repo, &self.workdir);
 \end{lstlisting}
 
-\noindent \textbf{Description:} This change happens when the developer needs to modify the program so it reaches the desired behaviour.
+\noindent\textbf{Description:} This change happens when the developer needs to modify the program so it reaches the desired behaviour.
 The modification, however, requires access to an object whose ownership has moved to another scope and never returned.
@@
 
-To give an example, in starship/starship after one of the commits \footnote{\url{https://github.com/starship/starship/commit/56d475578ea508631275772127f49a6949fea6b0}{git_branches.rs}}, the developer borrows the value of $repo\_dir$ to keep the ownership of the object in the current scope.
-Now this object can then be used in another function invocation ($remove\_dir\_all(repo\_dir)$) which fixes a bug.
+To give an example, in starship/starship after one of the commits \footnote{\url{https://github.com/starship/starship/commit/56d475578ea508631275772127f49a6949fea6b0}}, the developer borrows the value of \verb+repo_dir+ to keep the ownership of the object in the current scope.
+Now this object can then be used in another function invocation (\verb+remove_dir_all(repo_dir)+) which fixes a bug.
 \\
 
-\noindent \textit{ \textbf{BC.ref.drop} Removing Borrowing}
+\noindent\textit{\textbf{BC.ref.drop} Removing Borrowing}
 
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
+\begin{lstlisting}[language=Rust, style=colouredRust]
 enums.iter().for_each(|e| {
     let mut enum_template = EnumTemplate::new();
-    -- enum_template.set_message_enum(&e);
-    ++ enum_template.set_message_enum(e);
+    <@\textcolor{red}{--}@> enum_template.set_message_enum(&e);
+    <@\textcolor{green}{++}@> enum_template.set_message_enum(e);
     // Other statements
@@
 
-\noindent \textit{ \textbf{BC.deref.to\_string()} Modification in dereferencing a string prior to calling to\_string()}
+\noindent\textit{\textbf{BC.deref.to\_string()} Modification in dereferencing a string prior to calling to\_string()}
 
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
+\begin{lstlisting}[language=Rust, style=colouredRust]
 for p in wixobjs {
---  args.push(p.to_string());
-++  args.push((*p).to_string());
+<@\textcolor{red}{--}@>  args.push(p.to_string());
+<@\textcolor{green}{++}@>  args.push((*p).to_string());
 }
@@
 
-\noindent \textbf{Description:} This is a change that in all the instances has been detected by Clippy.
-The change applies on an object prior to calling $to\_string()$ on that object.
+\noindent\textbf{Description:} This is a change that in all the instances has been detected by Clippy.
+The change applies on an object prior to calling \verb+to_string()+ on that object.
 The change might add or remove a dereference, as shown above.
-In the addition case, the reason behind the change is that the object a reference type of $T$, while $T$ directly implements $to\_string()$.
-It is faster and more performance efficient to dereference the object and then call $to\_string()$ on it.
+In the addition case, the reason behind the change is that the object a reference type of \verb+T+, while \verb+T+ directly implements \verb+to_string()+.
+It is faster and more performance efficient to dereference the object and then call \verb+to_string()+ on it.
 The deletion case indicates a redundant dereference, thus it is removed.
@@
 
-\noindent \textit{ \textbf{BC.mut.add} Adding mutability}
+\noindent\textit{\textbf{BC.mut.add} Adding mutability}
 
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
+\begin{lstlisting}[language=Rust, style=colouredRust]
 parser::Parser::new(args)
         .parse_module()
---      .map_err(|e| {
-++      .map_err(|mut e| {
+<@\textcolor{red}{--}@>      .map_err(|e| {
+<@\textcolor{green}{++}@>      .map_err(|mut e| {
             e.emit();
@@
 
-\noindent \textbf{Description:} This change happens when the programmer needs to mutate a variable which is immutable, hence they need to add the mut keyword.
+\noindent\textbf{Description:} This change happens when the programmer needs to mutate a variable which is immutable, hence they need to add the mut keyword.
 In the observed instances, this pattern was not seen in refactoring changes, the modification happened alongside a change in program behaviour.
+\\
 
-\noindent \textit{ \textbf{BC.mut.drop} Dropping mutability}
+\noindent\textit{\textbf{BC.mut.drop} Dropping mutability}
 
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
--- let mut tx = tx.clone();
-++ let tx = tx.clone();
+\begin{lstlisting}[language=Rust, style=colouredRust]
+<@\textcolor{red}{--}@> let mut tx = tx.clone();
+<@\textcolor{green}{++}@> let tx = tx.clone();
 \end{lstlisting}
 
-\noindent \textbf{Description:} Dropping the mut keyword before a variable, as we do not mutate the variable. This redundant mutability is reported by Clippy and can also be removed through this tool.
+\noindent\textbf{Description:} Dropping the mut keyword before a variable, as we do not mutate the variable. This redundant mutability is reported by Clippy and can also be removed through this tool.
 
@@
 
-\noindent \textit{ \textbf{BC.vec.slice} Changing a Vec reference to Slice}
+\noindent\textit{\textbf{BC.vec.slice} Changing a Vec reference to Slice}
 
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
+\begin{lstlisting}[language=Rust, style=colouredRust]
 pub fn expand_delimited_square(
---  children: &Vec<TokenNode>,
-++  children: &[TokenNode],
+<@\textcolor{red}{--}@>  children: &Vec<TokenNode>,
+<@\textcolor{green}{++}@>  children: &[TokenNode],
 ) -> Result<hir::Expression, ParseError> {
@@
 \end{lstlisting}
-\noindent \textbf{Description:} In struct field definition or the types of the function formal parameters, type $\&Vec\<T\>$ changes to $\&[T]$.
+\noindent\textbf{Description:} In struct field definition or the types of the function formal parameters, type \verb+&Vec<T>+ changes to \verb+&[T]+.
 Clippy can detect the opportunity to apply this pattern.
-The reason is reducing type complication, as slice types in the form of $\&[T]$ or $\&str$ are usually sufficient for most use cases.
+The reason is reducing type complication, as slice types in the form of \verb+&[T]+ or \verb+&str+ are usually sufficient for most use cases.
 
 \subsubsection{Lifetime}
-\noindent \textit{ \textbf{BC.lifetime.static.drop} Dropping static lifetime}
+\noindent\textit{\textbf{BC.lifetime.static.drop} Dropping static lifetime}
 
-\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
--- const QUEUE_SIZES: &'static [u16] = &[QUEUE_SIZE];
-++ const QUEUE_SIZES: &[u16] = &[QUEUE_SIZE];
+\begin{lstlisting}[language=Rust, style=colouredRust]
+<@\textcolor{red}{--}@> const QUEUE_SIZES: &'static [u16] = &[QUEUE_SIZE];
+<@\textcolor{green}{++}@> const QUEUE_SIZES: &[u16] = &[QUEUE_SIZE];
 
@@
 
-\noindent \textbf{Description:} Removing the static keyword from a const variable declaration.
+\noindent\textbf{Description:} Removing the static keyword from a const variable declaration.
 This change can be applied using Clippy.
 If the presence of static is not required, it is better that we omit it as keeping it might create complicate types in our program.
-
-\begin{table}[]
-
-\begin{tabular}{|l|l|l|l|}
-\hline
-
-\multirow{10}{*}{General}        
-& \multirow{2}{*}{Attributes}    & G.attr.struct  & Modifying the attributes of structs                                        \\
-                                 &                                & G.attr.struct.field  & Modifying the attributes of struct fields                                  \\\cline{2-3}
-                                 & \multirow{2}{*}{Struct}        & G.struct.field  & Adding/Removing a struct field                                      \\
-                                 &                                & G.struct.field.pub  & Modifying the access modifiers of struct fields     
-                                  \\\cline{2-3}
-                                 & Option                         & G.option  & Changing field type T to T$\langle$Option$\rangle$                     \\\cline{2-3}
-                                 & \multirow{4}{*}{Types}         & G.type.field  & Changing a struct field type                                        \\
-                                 &                                & G.type.generic  & Changing a generic type                                             \\
-                                 &                                & G.type.enum.variant  & Change in enum variant value type                                   \\
-                                 &                                & G.type.tuple & Changing the type inside a tuple                                         \\\cline{2-3}
-                                 & \multirow{1}{*}{Traits}        & G.trait.bound & TraitBound Change                                               \\\cline{1-3}
-\multirow{9}{*}{Borrow Checker} &
-\multirow{2}{*}{Clone}                                         & BC.clone.ref & Dropping clone and adding borrowing                                                   \\
-                                 &                                & BC.clone.drop & Dropping clone   
-                                    \\\cline{2-3}
-    & \multirow{3}{*}{Ref and Deref} & BC.ref.add  & Adding Borrowing                                                           \\
-                                 &                                & BC.ref.drop  & Removing Borrowing                                                         \\
-                                 &                                & BC.deref.to\_string  & Modification in dereferencing a string prior to calling to\_string()                                                            \\\cline{2-3}
-                                 & \multirow{2}{*}{Mut}           & BC.mut.add  & Adding mutability                                                          \\
-                                 &                                & BC.mut.drop  & Dropping mutability                                                        \\\cline{2-3}
-                                 & Vector                         & BC.vec.slice  & Changing a Vec reference to Slice                                           \\\cline{2-3}
-                                 & \multirow{1}{*}{Lifetime}      
-                                                                 & BC.lifetime.static  & Dropping static lifetime    &                        
-                                 \hline
-
-\end{tabular}
-\caption{\label{fig:category}Bug Fix Pattern Categorization}
-\end{table}

### 2022-10-30 methodology.tex
methodology comments; probably need to move things around

@@
+\todo{introductory para saying that we're going to describe our methodology}
+
 We define a code change to be a modification to a program's source code that changes the program's abstract syntax tree (AST).
@@
 
-The general idea for automatically finding the pervasive code patterns within a list of programs is to parse the programs and, for us, to embed the important parts of the derived AST in fixed-size datapoints.
-Then, using a clustering algorithm, we can group these datapoints in different clusters.
-Work in this vein makes the assumption that each cluster corresponds to one class of change pattern~\cite{hanam2016discovering,campos2019discovering,yang2022mining}.
+Our approach for automatically finding pervasive code patterns aims to find clusters of similar code patterns, using existing implementations of appropriate clustering algorithms.
+Work in this vein generally makes the assumption that clusters corresponds to classes of change patterns~\cite{hanam2016discovering,campos2019discovering,yang2022mining}.
+We thus want to associate code changes with datapoints; one of our contributions is a fixed-size embedding which highlights the important parts of a code change.
 
-Our pipeline for finding classes of pervasive patterns in Rust programs also follows this idea.
-We decided to implement our pipeline completely in Python.
-To make this work, we need a way to get an AST as a Python dictionary.
-Syn\footnote{\url{https://crates.io/crates/syn}} is a Rust crate which aims to serve Rust procedural macros.
-However, it includes a parser which is suitable for our purposes; we simply had to write a preprocessor to transform Syn AST output into Python dictionaries.
+We now briefly describe our methodology, providing full details in later subsections.
+Our pipeline for finding classes of pervasive patterns in Rust programs parses program versions, computes differences between them, embeds them in fixed-size datapoints, and then clusters these points.
 
+\todo[inline]{I think we only need to talk about this next bit in a subsection.
+It's an implementation detail, not fundamental to our approach: ``We decided to implement our pipeline completely in Python.
+To make this work, we need a way to get an AST as a Python dictionary.''}
+
+\todo[inline]{I wonder if this should really go here, or if there should be a shorter version.}
 To compute the contents of a change, we need two code revisions: the revision before the commit, and the revision after it.
-After parsing these two revisions, we will have two ASTs, encoded in two Python dictionaries.
+After parsing these two revisions, we will have two ASTs.
 Tree diff algorithms can compute the differences between two arbitrary trees.
 When the trees are programs' abstract syntax trees, we call their diff an ASTDiff.
-We use dictdiffer to obtain ASTDiffs.
-
+% , encoded in two Python dictionaries; We use dictdiffer to obtain ASTDiffs.
+\todo[inline]{Maybe this next paragraph is what you want at the top of 2.1.
+Then in the next next paragraph you put a shorter description of the pipeline, in a sentence before ``we mined'', and you reference 2.1 there.
+Anyway think about organization and try to have something that makes sense.}
 An ASTDiff may include an arbitrary number of semantic changes, although a best practice is to include only one semantic change in a commit.
@@
 
-Our target repositories were the top 18 most-starred open source Rust projects on GitHub.
-We mined (using Pydriller) all the bug related commits and ran them through our pipeline.
-We will provide more details on the aforementioned process later.
+Our target repositories were the top 18 most-starred open source Rust projects on GitHub\todo{as of what date?}.
+We mined (using Pydriller) all the bug related commits and ran them through our pipeline; Section~\ref{sec:mining} describes our mining methodology in depth.
 Then, we applied the DBSCAN clustering algorithm~\cite{ester1996density} on the resulting datapoints.
-In Section~\ref{sec:clustering_data} we justify our choice of DBSCAN for clustering, and we explain how we tuned it to improve cluster quality.
+Section~\ref{sec:clustering_data} describes clustering in detail, justifies our choice of DBSCAN for clustering, and explains how we tuned it to improve cluster quality.
 
@@
 
+\todo[inline]{should always be text between subsection and subsubsection. This is the per-change workflow, right?}
+
 \subsubsection{\label{sec:parsing_programs}Parsing Programs}
 
-As mentioned above, we use Syn for parsing Rust.
+Syn\footnote{\url{https://crates.io/crates/syn}} is a Rust crate built for the Rust procedural macro implementation.
+However, it includes a parser which is suitable for our purposes; we simply had to write a preprocessor to transform Syn AST output into Python dictionaries.
 Syn handles all of Rust.
@@
 \texttt{clone}, \texttt{Rc}, \texttt{Box}.
-The full list of \verb+BCE+ elements can be found within Appendix~\ref{sec:appendix}.
+The full list of \verb+BCE+ elements can be found in Appendix~\ref{sec:appendix}.
 
@@
 \end{figure}
+% yay tikz
 
@@
 The second part identifies the context of the modification, that is, the path from the root of the tree to the subtree in which the modification has occured.
-The third part is the content of the change (terminals and nonterminals).
-Our tool looks for changes within each root-level scope.
-That is, our assumption is that a change pattern occurs within one root-level scope.
+The third part is the content of the change---terminals and nonterminals.
+Our tool\todo{do we have a name?
+It's important to have a name} looks for changes within each root-level scope.
+That is, our assumption is that each change pattern occurs within one root-level scope.
 In this work, we chose to not handle patterns that involve changes in multiple functions or multiple files.
-However, our tool would detect both continuous and non-continuous line changes within one scope.
+However, our tool detects both continuous and non-continuous line changes within one scope.
 
-Looking at the tree encoded in the content part of each diff, we realized that each path to a leaf represents one contributing element to the change.
+Looking at the tree encoded in the content part of each diff, we realized that each path to a leaf represents one element contributing to the change.
 A Depth First Search allows us to collect all paths, and we return only the paths that we store in our datapoints (i.e.
 touching $\mathtt{NT} \cup \mathtt{BCE}$).
+\todo{if you mathtt BCE here, you should do it everywhere, i.e. in the paragraph at the start of 2.1.1.}
 
@@
 A path represents a sequence of involved non terminals.
-We found it crucial to record the order of nodes in the path, as this gives us the freedom to later record the order in our datapoint embedding.
+We found it crucial to record the order of nodes in the path, as this allows us to record that order in our datapoint embedding.
 
@@
 First, we need to define a fixed set of columns.
-We need fixed-size datapoints, as they are a requirement for both centroid-based and density-based well-known clustering algorithms~\cite{xu2005survey}.
-The number of columns would depend on the amount of information we want to embed in the datapoints.
+We need fixed-size datapoints, as they are a requirement for well-known centroid-based and density-based clustering algorithms~\cite{xu2005survey}.
+The number of columns controls the amount of information we can embed in the datapoints.
 
@@
 While we can implement a specific workaround to distinguish nested ifs (and we have done so), the general point about the large dimensionality of fully order-sensitive embedding still holds.
+\todo[inline]{Worse than that, isn't $d$ unbounded?}
 
@@
 
-In our methodology, we propose a novel representation.
-Similar to the first approach, we order-insensitively define the columns of our dataset as the set of items $\mathtt{NT} \cup \mathtt{BCE}$ we collected from Syn.
-However, to simulate order-sensitiveness, we carried out two auxiliary steps.
+We propose a novel representation.
+Similar to the first\todo{which?
+the naive one?} approach, we order-insensitively define the columns of our dataset as the set of items $\mathtt{NT} \cup \mathtt{BCE}$ we collected from Syn.
+However, to simulate order-sensitiveness, we add two more steps.
 First, we record the number of occurrences of items within the paths at each scope.
-The reason behind this decision is because a different order of program elements yields a different number of occurrences of underlying dimension items.
+The reason behind this decision is because a different order of program elements yields a different number of occurrences of underlying dimension items \todo{is there an example?}.
 Second, we semi-automatically design a weighting scheme to prioritize non terminals that we think are more salient for Rust bugs.
@@
 \subsection{\label{sec:mining_repositories}Mining Repositories}
+\label{sec:mining}
 
@@
 Table~\ref{table:repos} summarizes our benchmarks.
-Because we are interested in changes, we computed the average code churn of the files within the projects (The code churn is computed using process metrics provided by Pydriller).
+Because we are interested in changes, we computed (using process metrics provided by our mining library) the average code churn of the files within the projects.
 denoland/deno and swc-project/swc are the projects with largest number of LOCs and highest average code churn, respectively.
@@
 \subsection{\label{sec:clustering_data}Clustering Data}
+\label{sec:clustering}
 
@@
 Secondly, it is a density-based clustering method, which means that it detects arbitrarily shaped clusters; centroid-based methods like k-means can't detect such clusters.
-We tried k-means and the more powerful fuzzy variant Fuzzy C-means~\cite{dunn1973fuzzy} and did not get useful clusters.
+We did try k-means and its more powerful fuzzy variant Fuzzy C-means~\cite{dunn1973fuzzy} and did not get useful clusters.
 
 DBSCAN takes two tuneable parameters.
-The first parameter $\epsilon$ indicates a radius within which a point is a core point, a border point, or an outlier.
+The first parameter $\epsilon$ indicates a radius distinguishing core points, border points, and outliers.
 The second parameter $Z$ is the minimum number of points per cluster.
@@
 \caption{\label{fig:clustering} Clustering results with nine different combinations of parameters $\epsilon$ and $Z$, respectively separated in columns and rows.
-In each subplot, the two blue bars in the left respectively indicate the number of clustered and noise datapoints (the number is shown above each bar) after applying clustering on $D_g$ (denoted with $G_c$ and $G_n$ on the x axis).
-The two orange bars in the right show similar data but for the $D_{b}$ database.
+In each subplot, the two blue bars in the left respectively indicate the number of clustered ($G_c$) and noise ($G_n)$ datapoints (the number is shown above each bar) after applying clustering on general database $D_g$.
+The two orange bars in the right show similar data but for the borrow-checker $D_{b}$ database.
 In each subplot's legend, $C_G$ and $C_B$ show the number of clusters obtained from $D_g$ and $D_b$, respectively}
@@
 
-As a general rule, lower $Z$ would allow more groups of close points to be considered as clusters, resulting in a higher number of clusters.
+\todo{figure caption: why the capital G and B in the subscript?}
+
+As a general rule, lower $Z$ would allow more groups of nearby points to be considered to be clusters, resulting in a higher number of clusters.
 A large $\epsilon$ would widen the search radius, allowing more points to fall in the clusters, which results in a reduction of the number of noise points.
-Such a choice might put points from different patterns inside the same cluster, resulting in a ineffective clustering.
+However, such a choice might put points from different patterns inside the same cluster, resulting in ineffective clustering.
 Alternatively, a small $\epsilon$ would tighten the ring of search, possibly separating clusters which essentially manifest the same code changes.
@@
 
+\todo{that is why what? huh? be more explicit and repeat yourself a bit}
 That is why we carried out a manual analysis of clusters.
-In the manual analysis, we randomly pick 50 clusters; from each, cluster we randomly choose 10 datapoints.
-The person analyzing the cluster then looks at the code and specifies the clusters that contained more than five datapoints showing similar patterns.
+In the manual analysis, we randomly pick 50 clusters; from each cluster we randomly choose 10 datapoints.
+The person analyzing the cluster \todo{who?
+be more specific about how we did this} then looks at the code and specifies the clusters that contained more than five datapoints showing similar patterns.
 For the manual analysis, we used $\epsilon=0.0001, Z=5$ and $\epsilon=0.001, Z=5$, which were the best parameters for $D_g$ and $D_b$, resulting in 577 and 102 clusters, respectively.
@@
 
-Similar to~\cite{yang2022mining}, following Cotroneo et al.~\cite{cotroneo2019analyzing} definitions, we divide the clusters in three different groups:
+Like~\cite{yang2022mining}, and following the definitions of Cotroneo et al.~\cite{cotroneo2019analyzing}, we divide the clusters in three different groups:
 
@@
 
+\todo{can we say more-specific rather than concrete?}
 In this work, we were only interested in the first two groups and ignored the clusters for refactoring changes.
@@
 Finally, we selected 19 of the remaining clusters---the ones that manifested concrete changes.
-We discuss about our findings in Sections~\ref{sec:common_patterns}~and~\ref{sec:bc_patterns}.
+Sections~\ref{sec:common_patterns}~and~\ref{sec:bc_patterns} present our clusters.

### 2022-10-30 discussion.tex
flattened everything ;((
results

--- 2022-10-11 sections/discussion.tex
@@
 A complicated pattern is a pattern that includes many changes in many places and it makes it easier for two datapoints to be dissimilar.
+\todo[inline]{There's a big part missing here, which is some sort of higher-level reflection on what we found and why it's important.
+Also we can sell this better to not make it sound that complex patterns are important.
+We said this before but we should say it again.
+Maybe using the word ``complex'' is bad because it makes our system sound not-complex.}
 
-Also, we expected to obtain certain patterns regarding some ertain Rust related features, such as unsafe Rust, but our results did not include those patterns.
+Also, we expected to obtain certain patterns regarding some certain Rust related features, such as unsafe Rust, but our results did not include those patterns.
 Our speculation is that, since Rust has a strong compiler check before creating the program executable, the programmers usually get many errors, and they only push their code once it is free of compile errors.
@@
 We believe our results can help researchers to create Rust tools for repair, fault localization, IDE linters, and etc.
-Also, these results can help program encoding methods to focus on more important aspects of a program and shrink down information more effectively.
+Also, these results can help program embedding methods to focus on more important aspects of a program and shrink down information more effectively.
 

### 2022-10-30 results.tex
results

@@
 As discussed in the previous section, we used an embedding mechanism to incorporate AST information in fixed-sized datapoints which then could be used in the DBSCAN clustering algorithm.
-We used heuristics (weighting scheme) while creating our datapoints to make the most important program elements stand out between all the non terminals seen in the AST (alongside some BC-related elements).
-Here we want to evaluate the effectiveness of our code embedding.
+We used heuristics (a weighting scheme) while creating our datapoints to make the most important program elements stand out from all the non terminals seen in the AST (along with some BC-related elements).
+Here we evaluate the effectiveness of our code embedding.
 
-The general idea for our evaluation is whether the most important program elements get the high values in our embedding.
-The analyzer compares the actual change in the code and compares it to the visual representation of its respective data point.
-We used circle pack figures to visualize each data point~\cite{collins2003circle}.
-For this purpose, we used the packcircles python library\footnote{\url{https://github.com/mhtchan/packcircles}}.
-In these figures, the radius of the circle of each non terminal shows its importance in our embedding (similar to Figure~\ref{fig:essence}).
+The general idea for our evaluation is determining whether the most important program elements get high values in our embedding, i.e.
+are recognized as important.
+The analyzer compares the actual change in the code and compares it to the visual representation of the respective data point.
+We used circle pack figures to visualize each data point~\cite{collins2003circle}, generated with the packcircles python library\footnote{\url{https://github.com/mhtchan/packcircles}}.
+In these figures, the radius of the circle of each non terminal shows its importance in our embedding (as seen in Figure~\ref{fig:essence}).
 
@@
 The reason behind this decision is the same as why we collected a lot of noise points, that is, the majority of the commits contain multi-hunk complex changes and the key changing elements are difficult to determine.
+
+\todo[inline]{Can we justify the 50, both here and in the previous section?}
 
@@
 Both authors carried out these experiments independently.
-At last, we report the results as the portion of the changes which both authors agreed upon its embedding correctness.
+We report the results as the portion of the changes which both authors agreed upon its embedding correctness.
 The final results are 54\% and 80\% correctness in the first and second experiments, respectively.
 
-% no, you have to compute the intersection
-% when is our approach effective and when is it not?
-% more detailed stats would probably be good too
+\todo[inline]{no, you have to compute the intersection. when is our approach effective and when is it not? more detailed stats would probably be good too}
 
@@
 
-We collected 60583 datapoints for general fix patterns using a server with Xeon Gold 5120 Processor (Q3-2017, 14-core 2.2 to 3.2GHz) and the data collection took around 26 hours to complete for both $D_g$ and $D_b$.
+We collected 60583 datapoints for general fix patterns using a server with a Xeon Gold 5120 processor (Q3-2017, 14-core 2.2 to 3.2GHz) and the data collection took around 26 hours to compute both $D_g$ and $D_b$.
 We used $\epsilon=0.0001$ and $Z=5$ as our DBSCAN parameters and that resulted in 577 clusters.
-After manual analysis as described in~\ref{sec:manual_analysis_cluster_selection} we ended up with 10 general patterns which we introduce here:
+After manual analysis, as described in~\ref{sec:manual_analysis_cluster_selection}, we ended up with 10 general patterns, which we introduce below.
 
@@
 Using attributes for struct fields, apart from being able to store different meta-information for different struct fields, for some use cases, the programmers can control the action they want to apply on different fields.
-% this last sentence immediately prior to here has too many moving parts
+\todo{this last sentence immediately prior to here has too many moving parts}
 
@@
 \noindent\textbf{Description:} Rust is a null-safe language, meaning that object references cannot take on null values.
-If we want a variable to contain null, we would have to use the \texttt{Option} type.
-\texttt{Option} is essentially an \texttt{enum} with two variants, \texttt{Some}, which indicates that our variable has a value, and \texttt{None}, which indicates the absence of any value.
+If a developer wants a variable to contain null, they would have to use the \texttt{Option} type.
+\texttt{Option} is essentially an \texttt{enum} with two variants, \texttt{Some}, which indicates that a variable has a value, and \texttt{None}, which indicates the absence of any value.
 In this change pattern, the developer changes the type of a struct field to an \texttt{Option} of that type, hence enabling them to store no value in that field.
 
-For instance, in the Rust project swc-project/swc (a fast TypeScript/JavaScript compiler), the developer needed to modify the CSS parser source code in a way that it accounts for empty @media queries \footnote{\url{https://github.com/swc-project/swc/commit/75a14f98b7370226115ee24eec6eb8c802bd4837}}.
-So, as shown in the code snippet above, they changed type \texttt{MediaQueryList} to \verb+Option<MediaQueryList>+, and modified the other parts of the source code, accordingly.
+For instance, in the Rust project swc-project/swc (a fast TypeScript/JavaScript compiler), the developer needed to modify the CSS parser source code to account for empty @media queries\footnote{\url{https://github.com/swc-project/swc/commit/75a14f98b7370226115ee24eec6eb8c802bd4837}}.
+As shown in the snippet above, they changed type \texttt{MediaQueryList} to \verb+Option<MediaQueryList>+, and modified the other parts of the source code, accordingly.
 
@@
 For instance, the developer may be implementing a new feature or fixing the behaviour of the program.
-Also it can happen for refactoring purposes or performance enhancement.
+Also, it can happen for refactoring purposes or performance enhancement.
 \\
-% maybe discuss the specific example a bit?
+\todo{maybe discuss the specific example a bit?}
 
@@
 
-\noindent\textbf{Description:} This a change of the type parameter of a struct field. As with \textit{\textbf{G.type.field}}, this change can also happen for the mentioned reasons. \\
+\noindent\textbf{Description:} This is a change of the type parameter of a struct field. This change pattern can occur for the same reasons as for \textit{\textbf{G.type.field}}. \\
 
@@
 \noindent\textbf{Description:} The trait bounds are the functionalities that we require from our parametric types.
-This change pattern relates to modification in trait bounds.
-This can happen both for bug fix reasons or just refactoring purposes.
+This change pattern relates to modifications in trait bounds.
+This can happen as part of bug fixes or for refactoring.
 
@@
 We used $\epsilon=0.001$ and $Z=5$ as our DBSCAN parameters and that resulted in 102 clusters.
-After manual analysis as described in~\ref{sec:manual_analysis_cluster_selection} we ended up with 9 borrow-checker related fix patterns which we introduce here:
+After manual analysis, as described in~\ref{sec:manual_analysis_cluster_selection}, we ended up with 9 borrow-checker related fix patterns, which we introduce here:
 
@@
 
-\noindent\textit{\textbf{BC.clone.drop} Removing clone}
+\noindent\textit{\textbf{BC.clone.drop} Dropping clone}
 
@@
 There are more than 500 lints included in Clippy.
-With regards to this pattern, Clippy detects whether a variable clone is necessary and could we have used the variable itself or not.
-This change does not modify the program behaviour, however it can boost the performance significantly if we repeatedly cloning large variables.
+With regards to this pattern, Clippy detects whether a variable clone is necessary, i.e.
+could the developer have used the variable itself or not.
+This change does not modify the program behaviour, but it can boost the performance significantly if the program was repeatedly cloning large variables.
+\\
+
+\todo[inline]{I think you should talk more about ``does not modify program behaviour''.
+For instance after you make this change the program may no longer compile.
+Is it true that, if it compiles, behaviour is preserved?}
 
@@
 \noindent\textbf{Description:} Unlike the previous pattern, this pattern is not detected by Clippy.
-The change happens when the programmers realizes that the cloning of a variable is unnecessary, however, we also want to preserve the ownership of the variable within the current scope.
-That is why in this pattern a clone is turned into a borrowing.
+The change happens when a developer realizes that the cloning of a variable is unnecessary, but also wants to keep the ownership of the variable within the current scope (rather than passing it to a callee).
+That is why, in this pattern, a clone is turned into a borrow.
 
-Similar to the previous pattern, this does not change the program behaviour and is merely for performance purposes.
-It was the case in one of the commits~\footnote{\url{https://github.com/tauri-apps/tauri/commit/a280ee90af0749ce18d6d0b00939b06473717bc9}} of the project tauri-apps/tauri, where the clone to borrowing significantly optimized CPU usage.
-They discussed that the cloning of some types such as HashMap was expensive.
+Like the previous pattern, this change does not affect the program behaviour; it is done for performance purposes.
+In one of the commits\footnote{\scriptsize \url{https://github.com/tauri-apps/tauri/commit/a280ee90af0749ce18d6d0b00939b06473717bc9}} of the project tauri-apps/tauri, changing a clone to borrowing significantly improved CPU usage.
+In the commit history, developers discussed how the cloning of types such as HashMap was expensive.
 \\
@@
 
-\noindent\textbf{Description:} This change happens when the developer needs to modify the program so it reaches the desired behaviour.
-The modification, however, requires access to an object whose ownership has moved to another scope and never returned.
-That is why we must use borrowing to preserve ownership within our scope, after the function execution finishes.
+\noindent\textbf{Description:} This change pattern happens when the developer now needs to take back ownership of an object that was previously passed to a callee (or other scope) and never returned.
+Borrowing allows the first scope to once again act on the object.
 
-To give an example, in starship/starship after one of the commits \footnote{\url{https://github.com/starship/starship/commit/56d475578ea508631275772127f49a6949fea6b0}}, the developer borrows the value of \verb+repo_dir+ to keep the ownership of the object in the current scope.
-Now this object can then be used in another function invocation (\verb+remove_dir_all(repo_dir)+) which fixes a bug.
+To give an example, the commit in starship/starship excerpted above\footnote{\scriptsize \url{https://github.com/starship/starship/commit/56d475578ea508631275772127f49a6949fea6b0}} shows that the developer now borrows the value of \verb+repo_dir+ to keep the ownership of the object in the current scope.
+This object can then be used in a subsequent function invocation (\verb+remove_dir_all(repo_dir)+), which fixes a bug.
 \\
@@
 
-\noindent\textbf{Description:} In the pattern, an unnecessary borrowing is removed.
+\noindent\textbf{Description:} In this pattern, an unnecessary borrow is removed.
 Clippy can detect the needless borrows, and also can automatically remove them.
@@
 \\
+\todo[inline]{I don't see why this improves performance, it should be a 0 cost operation at runtime to return ownership or not.}
 
-\noindent\textit{\textbf{BC.deref.to\_string()} Modification in dereferencing a string prior to calling to\_string()}
+\noindent\textit{\textbf{BC.deref.to\_string()} Adding or removing a dereference operation before calling to\_string()}
 
@@
 
-\noindent\textbf{Description:} This is a change that in all the instances has been detected by Clippy.
+\todo[inline]{Can you clarify ``has been detected by Clippy''? Did you run Clippy? You didn't talk about that in your methodology.}
+\noindent\textbf{Description:} This is a change that in all instances has been detected by Clippy.
 The change applies on an object prior to calling \verb+to_string()+ on that object.
 The change might add or remove a dereference, as shown above.
-In the addition case, the reason behind the change is that the object a reference type of \verb+T+, while \verb+T+ directly implements \verb+to_string()+.
+When adding a dereference, the reason behind the change is that the object is a reference type of \verb+T+, while \verb+T+ directly implements \verb+to_string()+.
 It is faster and more performance efficient to dereference the object and then call \verb+to_string()+ on it.
-The deletion case indicates a redundant dereference, thus it is removed.
+The deletion case indicates a redundant dereference, thus it is removed.\todo{I don't understand the adding change.}
 
@@
 
-\noindent\textbf{Description:} This change happens when the programmer needs to mutate a variable which is immutable, hence they need to add the mut keyword.
-In the observed instances, this pattern was not seen in refactoring changes, the modification happened alongside a change in program behaviour.
+\noindent\textbf{Description:} This change happens when the developer needs to mutate a variable which was previously immutable in a scope---they thus need to add the mut keyword.
+We never observed this pattern in refactoring changes; it always accompanied a change in program behaviour.
 \\
@@
 
-\noindent\textbf{Description:} Dropping the mut keyword before a variable, as we do not mutate the variable. This redundant mutability is reported by Clippy and can also be removed through this tool.
+\noindent\textbf{Description:} The developer drops the mut keyword before a variable, as they do not mutate the variable.
+This redundant mutability is reported by Clippy and can also be removed by it.
+This change may help future-proof the code against unintended future changes.
 
@@
 \end{lstlisting}
-\noindent\textbf{Description:} In struct field definition or the types of the function formal parameters, type \verb+&Vec<T>+ changes to \verb+&[T]+.
+\noindent\textbf{Description:} In a struct field definition, or in the types of a function's formal parameters, type \verb+&Vec<T>+ changes to slice \verb+&[T]+.
 Clippy can detect the opportunity to apply this pattern.
-The reason is reducing type complication, as slice types in the form of \verb+&[T]+ or \verb+&str+ are usually sufficient for most use cases.
+This simplifies the code; slice types \verb+&[T]+ or \verb+&str+ are sufficient for most use cases.
 
@@
 
-\noindent\textbf{Description:} Removing the static keyword from a const variable declaration.
+\noindent\textbf{Description:} This change pattern removes the static keyword from a const variable declaration.
 This change can be applied using Clippy.
-If the presence of static is not required, it is better that we omit it as keeping it might create complicate types in our program.
+If the presence of static is not required, it is better that to omit it, as keeping it might create complicated types in the program.

### 2022-10-30 methodology.tex
todo

@@
 
-Like~\cite{yang2022mining}, and following the definitions of Cotroneo et al.~\cite{cotroneo2019analyzing}, we divide the clusters in three different groups:
+Like Yang et al~\cite{yang2022mining}, and following the definitions of Cotroneo et al.~\cite{cotroneo2019analyzing}, we divide the clusters in three different groups:
 

### 2022-10-30 results.tex
todo

@@
 \end{table}
+\todo[inline]{There cannot be a table without reference to the table in the text.
+The reference to the table should also summarize what we've learned from the table.
+Also do we compare table 2 to what occurs in previous work?}
 
@@
 
-For instance, in the Rust project swc-project/swc (a fast TypeScript/JavaScript compiler), the developer needed to modify the CSS parser source code to account for empty @media queries\footnote{\url{https://github.com/swc-project/swc/commit/75a14f98b7370226115ee24eec6eb8c802bd4837}}.
+For instance, in the Rust project swc-project/swc (a fast TypeScript/JavaScript compiler), the developer needed to modify the CSS parser source code to account for empty @media queries\footnote{\scriptsize \url{https://github.com/swc-project/swc/commit/75a14f98b7370226115ee24eec6eb8c802bd4837}}.
 As shown in the snippet above, they changed type \texttt{MediaQueryList} to \verb+Option<MediaQueryList>+, and modified the other parts of the source code, accordingly.

### 2022-10-31 methodology.tex
methodology fixed

@@
-\todo{introductory para saying that we're going to describe our methodology}
+% \todo{introductory para saying that we're going to describe our methodology}
+
+In this section, we describe our methodology, providing full details in later subsections.
+Our pipeline for finding classes of pervasive patterns in Rust programs parses program versions, computes differences between them, embeds them in fixed-size datapoints, and then clusters these datapoints.
+After conducting a manual analysis on the obtained clusters, we refine them towards manifesting bug fix patterns.
 
@@
 
-We now briefly describe our methodology, providing full details in later subsections.
-Our pipeline for finding classes of pervasive patterns in Rust programs parses program versions, computes differences between them, embeds them in fixed-size datapoints, and then clusters these points.
-
-\todo[inline]{I think we only need to talk about this next bit in a subsection.
+% \todo[inline]{I think we only need to talk about this next bit in a subsection.
 It's an implementation detail, not fundamental to our approach: ``We decided to implement our pipeline completely in Python.
@@
 
-\todo[inline]{I wonder if this should really go here, or if there should be a shorter version.}
-To compute the contents of a change, we need two code revisions: the revision before the commit, and the revision after it.
+% \todo[inline]{I wonder if this should really go here, or if there should be a shorter version.}
+% To compute the contents of a change, we need two code revisions: the revision before the commit, and the revision after it.
 After parsing these two revisions, we will have two ASTs.
@@
 % , encoded in two Python dictionaries; We use dictdiffer to obtain ASTDiffs.
-\todo[inline]{Maybe this next paragraph is what you want at the top of 2.1.
-Then in the next next paragraph you put a shorter description of the pipeline, in a sentence before ``we mined'', and you reference 2.1 there.
-Anyway think about organization and try to have something that makes sense.}
-An ASTDiff may include an arbitrary number of semantic changes, although a best practice is to include only one semantic change in a commit.
-Because that best practice is not universally followed, we are interested in finding the most important change within an ASTDiff, and embedding that change in our datapoints.
-Sections~\ref{sec:path_extraction}~and~\ref{sec:weighting_scheme} provide a detailed explanation of how we select the most important semantic information, which then allowed us to obtain clusters that contained similar datapoints.
 
-Our target repositories were the top 18 most-starred open source Rust projects on GitHub\todo{as of what date?}.
-We mined (using Pydriller) all the bug related commits and ran them through our pipeline; Section~\ref{sec:mining} describes our mining methodology in depth.
+We decided to implement our pipeline completely in Python.
+Our target repositories were the top 18 most-starred open source Rust projects on GitHub as of August 2021.
+We mined (using Pydriller) all the bug related commits and ran them through our pipeline; Section~\ref{sec:data_modelling} describes how we extracted fixed-size datapoints from the code changes.
+Section~\ref{sec:mining} describes our mining methodology in depth.
 Then, we applied the DBSCAN clustering algorithm~\cite{ester1996density} on the resulting datapoints.
@@
 
-\todo[inline]{should always be text between subsection and subsubsection. This is the per-change workflow, right?}
+% \todo[inline]{Maybe this next paragraph is what you want at the top of 2.1.
+Then in the next next paragraph you put a shorter description of the pipeline, in a sentence before ``we mined'', and you reference 2.1 there.
+Anyway think about organization and try to have something that makes sense.}
+To compute the contents of a change, we need two code revisions: the revision before the commit, and the revision after it.
+After parsing these two revisions, we will have two ASTs.
+Tree diff algorithms can compute the differences between two arbitrary trees.
+When the trees are programs' abstract syntax trees, we call their diff an ASTDiff.
+An ASTDiff may include an arbitrary number of semantic changes, although a best practice is to include only one semantic change in a commit.
+Because that best practice is not universally followed, we are interested in finding the most important change within an ASTDiff, and embedding that change in our datapoints.
+Section~\ref{sec:parsing_programs} describes our program parsing process, and how we obtained ASTDiff from code changes.
+Sections~\ref{sec:path_extraction}~and~\ref{sec:weighting_scheme} provide a detailed explanation of how we select the most important semantic information out of an ASTDiff, which then allowed us to obtain clusters that contained similar datapoints.
+
+% \todo[inline]{should always be text between subsection and subsubsection. This is the per-change workflow, right?}
 
@@
 containing one dimension per non terminal.
-However, we are also interested in change patterns that involve the borrow checker (BC), so we also add a dimension reflecting the presence of a set of BC-related elements (BCE) that we identified, e.g.
+However, we are also interested in change patterns that involve the borrow checker (BC), so we also add a dimension reflecting the presence of a set of BC-related elements (\texttt{BCE}) that we identified, e.g.
 \texttt{clone}, \texttt{Rc}, \texttt{Box}.
-The full list of \verb+BCE+ elements can be found in Appendix~\ref{sec:appendix}.
+The full list of \texttt{BCE} elements can be found in Appendix~\ref{sec:appendix}.
 
@@
 Our tool\todo{do we have a name?
-It's important to have a name} looks for changes within each root-level scope.
+It's important to have a name} looks for changes within each root-level scope (Rust book calls these root-level scopes the items of a crate\footnote{\url{https://doc.rust-lang.org/reference/items.html}}).
 That is, our assumption is that each change pattern occurs within one root-level scope.
@@
 touching $\mathtt{NT} \cup \mathtt{BCE}$).
-\todo{if you mathtt BCE here, you should do it everywhere, i.e. in the paragraph at the start of 2.1.1.}
+% \todo{if you mathtt BCE here, you should do it everywhere, i.e. in the paragraph at the start of 2.1.1.}
 
@@
 While we can implement a specific workaround to distinguish nested ifs (and we have done so), the general point about the large dimensionality of fully order-sensitive embedding still holds.
-\todo[inline]{Worse than that, isn't $d$ unbounded?}
+\todo[inline]{Worse than that, isn't $d$ unbounded? ``yes it is, but doesn't what we've said above still hold?''}
 
@@
 We propose a novel representation.
-Similar to the first\todo{which?
-the naive one?} approach, we order-insensitively define the columns of our dataset as the set of items $\mathtt{NT} \cup \mathtt{BCE}$ we collected from Syn.
+Similar to the naive approach, we order-insensitively define the columns of our dataset as the set of items $\mathtt{NT} \cup \mathtt{BCE}$ we collected from Syn.
 However, to simulate order-sensitiveness, we add two more steps.
 First, we record the number of occurrences of items within the paths at each scope.
-The reason behind this decision is because a different order of program elements yields a different number of occurrences of underlying dimension items \todo{is there an example?}.
+The reason behind this decision is because a different order of program elements yields a different number of occurrences of underlying dimension items\todo{is there an example?
+``Fig 2?''}.
 Second, we semi-automatically design a weighting scheme to prioritize non terminals that we think are more salient for Rust bugs.
@@
 The two orange bars in the right show similar data but for the borrow-checker $D_{b}$ database.
-In each subplot's legend, $C_G$ and $C_B$ show the number of clusters obtained from $D_g$ and $D_b$, respectively}
+In each subplot's legend, $C_g$ and $C_b$ show the number of clusters obtained from $D_g$ and $D_b$, respectively}
 \end{figure}
-
-\todo{figure caption: why the capital G and B in the subscript?}
 
@@
 
-\todo{that is why what? huh? be more explicit and repeat yourself a bit}
-That is why we carried out a manual analysis of clusters.
+Since both large and small values of $\epsilon$ would steer us away from obtaining meaningful clusters, we felt the need to carry out a manual analysis of clusters for better parameter tuning.
 In the manual analysis, we randomly pick 50 clusters; from each cluster we randomly choose 10 datapoints.
-The person analyzing the cluster \todo{who?
-be more specific about how we did this} then looks at the code and specifies the clusters that contained more than five datapoints showing similar patterns.
-For the manual analysis, we used $\epsilon=0.0001, Z=5$ and $\epsilon=0.001, Z=5$, which were the best parameters for $D_g$ and $D_b$, resulting in 577 and 102 clusters, respectively.
+The person analyzing the cluster (The first author) then looks at the code and specifies the clusters that contained more than five datapoints showing similar patterns.
+After conducting the manual analysis, we decided to use $\epsilon=0.0001, Z=5$ and $\epsilon=0.001, Z=5$, as our clustering parameters for $D_g$ and $D_b$, resulting in 577 and 102 clusters, respectively.
 As a final step, we exclude clusters that contained datapoints from fewer than 3 projects.
@@
 
-\todo{can we say more-specific rather than concrete?}
 In this work, we were only interested in the first two groups and ignored the clusters for refactoring changes.
-Also, we prefer reporting clusters that manifest concrete patterns.
-An abstract pattern as opposed to a concrete pattern can account for many changes in the programs.
-For instance, we deem the pattern `Changing a clone of a variable to a borrowing of it in a function argument' more concrete than `Adding a new statement to a function's body':
+Also, we prefer reporting clusters that manifest more-specific patterns.
+An abstract pattern as opposed to a more-specific pattern can account for many changes in the programs.
+For instance, we deem the pattern `Changing a clone of a variable to a borrowing of it in a function argument' more specific than `Adding a new statement to a function's body':
 

### 2022-11-02 discussion.tex
fixed discussion, results todos, added two G patterns, dropped one BC pattern

--- 2022-10-11 sections/discussion.tex
@@
-In this work, our pipeline successfully mined 19 cross project bug fix patterns.
-However, our pipeline fails to detect complex pattern due to space limitation and their innate complicated features.
-A complicated pattern is a pattern that includes many changes in many places and it makes it easier for two datapoints to be dissimilar.
+In this work, our pipeline successfully mined 20 cross project bug fix patterns.
+We introduced these patterns in two groups: 12 general patterns and 8 BC-related patterns.
+Clippy is able to detect four of the BC-related patterns.
+To the best of our knowledge, the rest of the patterns cannot be reported by the current linting tools.
+That is why, we believe our work introduces a new direction for IDE tools to help Rust developers.
 
-Also, we expected to obtain certain patterns regarding some ertain Rust related features, such as unsafe Rust, but our results did not include those patterns.
+In each group, we categorized the bug fix patterns based on the underlying program element.
+The patterns encompass a wide variety of program elements both in the general and BC-related patterns.
+This shows that our weighting scheme is acceptably unbiased.
+Also, the introduced patterns are all cross-project fix patterns.
+Each fix pattern has been seen at least in three different projects, indicating their commonality.
+
+Furthermore, we expected to obtain certain patterns regarding some certain Rust related features, such as unsafe Rust, but our results did not include those patterns.
 Our speculation is that, since Rust has a strong compiler check before creating the program executable, the programmers usually get many errors, and they only push their code once it is free of compile errors.

### 2022-11-02 results.tex
fixed discussion, results todos, added two G patterns, dropped one BC pattern

@@
 
-\todo[inline]{Can we justify the 50, both here and in the previous section?}
+\todo[inline]{Can we justify the 50, both here and in the previous section? ``It was a random number not so large and not too small, how can I justify it?''}
 
@@
 
-\todo[inline]{no, you have to compute the intersection. when is our approach effective and when is it not? more detailed stats would probably be good too}
+Although reporting most important element within the top three elements in 80\% of the cases is an accpectable outcome, we examined our pipeline for the other 20\%.
+The analysis shows that our pipeline fails in change patterns where the change involves many elements (e.g.
+adding a whole new \verb+impl+ block and implementing new functions in it).
+It is worth mentioning that these changes tend to be less cross-project as having many changed elements make it easier for datapoints to be dissimlar.
+
+\todo[inline]{no, you have to compute the intersection.
+``The numbers reported are the intersection'' when is our approach effective and when is it not?
+``added above'' more detailed stats would probably be good too ``what kind of stats do you have in mind?''}
 
@@
 We used $\epsilon=0.0001$ and $Z=5$ as our DBSCAN parameters and that resulted in 577 clusters.
-After manual analysis, as described in~\ref{sec:manual_analysis_cluster_selection}, we ended up with 10 general patterns, which we introduce below.
+After manual analysis, as described in~\ref{sec:manual_analysis_cluster_selection}, we ended up with 10 general patterns, which we describe in Table~\ref{table:general}.
+As it is shown in the table, we have separated the patterns based underlying program element (Category column).
+We have given each pattern a specific ID (ID column) and determined the number of instances manifesting that fix pattern (\# Instances column), as well as the number of projects in which the pattern was seen (\# Project column).
+Now we provide detailed description for each of these patterns:
 
@@
 & G.type.tuple & 6 & 6                                         \\\cline{1-4}
-\multirow{1}{*}{Traits} & G.trait.bound & 6 & 5 \\ 
+\multirow{1}{*}{Traits} & G.trait.bound & 6 & 5 \\\cline{1-4}
+\multirow{2}{*}{Match} & G.match.pattern & 41 & 12 \\
+& G.match.code & 8 & 5 \\
 \end{tabular}
@@
 \end{table}
-\todo[inline]{There cannot be a table without reference to the table in the text.
-The reference to the table should also summarize what we've learned from the table.
-Also do we compare table 2 to what occurs in previous work?}
+\todo[inline]{Also do we compare table 2 to what occurs in previous work? ``What previous work?''}
 
@@
 As with \textit{\textbf{G.attr.struct}}, this modification also happens in three different ways.
-Using attributes for struct fields, apart from being able to store different meta-information for different struct fields, for some use cases, the programmers can control the action they want to apply on different fields.
-\todo{this last sentence immediately prior to here has too many moving parts}
+Using these attributes, the developers can control different actions they wish to apply on struct fields, in addition to providing meta-information.
 
@@
 pub struct Manifest {
-  <@\textcolor{red}{--}@> pub features: Vec<String>,
-  <@\textcolor{green}{++}@> pub features: HashSet<String>,
+  <@\textcolor{red}{--}@> pub substitutions: HashMap<String, &'a str>,
+  <@\textcolor{green}{++}@> pub substitutions: IndexMap<String, &'a str>,
 }
@@
 Also, it can happen for refactoring purposes or performance enhancement.
+
+For instance, in the Rust project starship/starship (a cross-shell prompt), in order to preserve the insertion order of \verb+substitutions+ field, the developer changes its type from \verb+HashMap+ to \verb+IndexMap+\footnote{\url{https://github.com/starship/starship/commit/4de9e43cff46c834bd340d24a02fc95d85310a33}}.
 \\
-\todo{maybe discuss the specific example a bit?}
 
@@
 \subsubsection{Traits}
-\noindent\textit{\textbf{G.trait.bound} TraitBound Change}
+\noindent\textit{\textbf{G.trait.bound} TraitBound change}
 
@@
 This can happen as part of bug fixes or for refactoring.
+\\
+
+\subsubsection{Match}
+
+In Rust, match control flow construct is used for comparing a value against multiple patterns, and executing the code associated with the first pattern that matches with our value.
+A pattern and its associated code are called a match arm.\\
+
+\noindent\textit{\textbf{G.match.pattern} Change in a match arm's pattern}
+
+\begin{lstlisting}[language=Rust, style=colouredRust]
+<@\textcolor{red}{--}@> Value::Primitive(p) => {
+<@\textcolor{green}{++}@> UntaggedValue::Primitive(p) => {
+    let _ = builder.add_empty_child(p.format(None));
+}
+\end{lstlisting}
+
+\noindent\textbf{Description:} In this pattern, a change occurs in an arm's pattern.
+Like the code snippet provided above, it can be a change in pattern's type.
+It also can be matching to a different enum variant or string.
+The fix pattern can happen both for refactoring or debugging purposes.
+\\
+
+\noindent\textit{\textbf{G.match.code} Change in a match arm's code}
+
+\begin{lstlisting}[language=Rust, style=colouredRust]
+Operation::Insert(insert) => {
+    <@\textcolor{red}{--}@> inverted.delete(insert.count_of_utf16_code_units());
+    <@\textcolor{green}{++}@> inverted.delete(insert.utf16_size());
+}
+\end{lstlisting}
+
+\noindent\textbf{Description:} The block of code associated with a pattern in match arm can contain multiple statements.
+In this fix pattern, a change happens in at least one of these statements.
+Similar to the previous fix pattern, this also can be due to refactoring or debugging purposes.
 
@@
 We used $\epsilon=0.001$ and $Z=5$ as our DBSCAN parameters and that resulted in 102 clusters.
-After manual analysis, as described in~\ref{sec:manual_analysis_cluster_selection}, we ended up with 9 borrow-checker related fix patterns, which we introduce here:
+After manual analysis, as described in~\ref{sec:manual_analysis_cluster_selection}, we ended up with 9 borrow-checker related fix patterns, which we, similar to the previous table, describe in Table~\ref{table:bc}.
+Here we provide detailed description for each of these patterns:
 
@@
 \\\cline{1-4}
-\multirow{3}{*}{Ref and Deref} & BC.ref.add & 17 & 7 \\
-& BC.ref.drop & 145 & 14 \\
-& BC.deref.to\_string  & 7 & 4                                            \\\cline{1-4}
+\multirow{2}{*}{Ref and Deref} & BC.ref.add & 17 & 7 \\
+& BC.deref.to\_string.add  & 4 & 3                                            \\\cline{1-4}
 \multirow{2}{*}{Mut} & BC.mut.add  & 11 & 6 \\
@@
 could the developer have used the variable itself or not.
-This change does not modify the program behaviour, but it can boost the performance significantly if the program was repeatedly cloning large variables.
+If the variable clone is indeed redundant, then removing it does not modify the program behaviour, however it can boost the performance significantly if the program was repeatedly cloning large variables.
 \\
@@
 For instance after you make this change the program may no longer compile.
-Is it true that, if it compiles, behaviour is preserved?}
+Is it true that, if it compiles, behaviour is preserved?
+``Yes the assumption is that if the presence of clone() is unnecessary, then removing it would not change the program behaviour''}
 
@@
 
-\noindent\textit{\textbf{BC.ref.drop} Removing Borrowing}
-
-\begin{lstlisting}[language=Rust, style=colouredRust]
-enums.iter().for_each(|e| {
-    let mut enum_template = EnumTemplate::new();
-    <@\textcolor{red}{--}@> enum_template.set_message_enum(&e);
-    <@\textcolor{green}{++}@> enum_template.set_message_enum(e);
-    // Other statements
-});
-\end{lstlisting}
-
-\noindent\textbf{Description:} In this pattern, an unnecessary borrow is removed.
-Clippy can detect the needless borrows, and also can automatically remove them.
-This does not change the behaviour of the program, however, it does improve the performance, as the borrow checker need not to return the ownership after the execution of the function.
-\\
-\todo[inline]{I don't see why this improves performance, it should be a 0 cost operation at runtime to return ownership or not.}
-
-\noindent\textit{\textbf{BC.deref.to\_string()} Adding or removing a dereference operation before calling to\_string()}
+\noindent\textit{\textbf{BC.deref.to\_string().add} Adding a dereference operation before calling to\_string()}
 
@@
 
-\todo[inline]{Can you clarify ``has been detected by Clippy''? Did you run Clippy? You didn't talk about that in your methodology.}
-\noindent\textbf{Description:} This is a change that in all instances has been detected by Clippy.
+\noindent\textbf{Description:} This is a change that in all instances has been detected by Clippy (commit messages indicated the use of Clippy for this fix).
 The change applies on an object prior to calling \verb+to_string()+ on that object.
@@
 When adding a dereference, the reason behind the change is that the object is a reference type of \verb+T+, while \verb+T+ directly implements \verb+to_string()+.
-It is faster and more performance efficient to dereference the object and then call \verb+to_string()+ on it.
-The deletion case indicates a redundant dereference, thus it is removed.\todo{I don't understand the adding change.}
+Adding the dereference makes the compiler use the specialized implementation of \verb+to_string()+ and not go through slower string formatting methods.
 

### 2022-11-02 conclusion.tex
flattened everything ;((
added Ruxanne as the name

--- 2022-10-16 sections/conclusion.tex
@@
-In this project, we mined the top 18 most starred Rust projects in Github to discover common bug fix patterns lied within their code changes.
-We proposed a novel program modelling approach to encode most important aspects of a code change in a fixed sized datapoint.
+In this project, we developed Ruxanne, a tool for mining bug fix patterns in Rust.
+Using this tool, we mined the top 18 most starred Rust projects in Github to discover common bug fix patterns lied within their code changes.
+At the heart of Ruxanne, we used a novel code embedding approach to embed most important aspects of a code change in a fixed sized datapoint.
 We stored 87,726 datapoints drawn from 57,214 commits across these 18 projects.
-After using the DBSCAN clustering algorithm, and a subsequent manual analysis, we obtained 10 clusters of general fix patterns and 9 clusters of borrow-checker related fix patterns.
+After using the DBSCAN clustering algorithm, and a subsequent manual analysis, we obtained 12 clusters of general fix patterns and 8 clusters of borrow-checker related fix patterns.

### 2022-11-02 introduction.tex
comment about rq1, and rephrase
added Ruxanne as the name

@@
-Producing bugs has been an unavoidable part of computer programming.
-Researchers have been trying to help software engineers to be more aware of the bug producing practices in development and how to step towards creating bug-free software.
-One useful step towards that goal would be studying common bug patterns (and their corresponding fixes) that occur across multiple software systems.
-Research efforts for categorizing common bug patterns date back to 1975 where Edres tried to categorize various bug patterns which occurred in operating system implementations.
+Bugs have long been an unavoidable part of computer programming, predating Grace Hopper's literal debugging session in 1947; even today, bug-free software remains an aspirational aim for software engineering research.
+One step towards that aim has been studying common bug patterns (and their corresponding fixes) across multiple software systems.
+Research efforts categorizing common bug patterns date back to at least 1975, where Endres tried to categorize bug patterns in operating system implementations~\cite{endres1975analysis}.
 
-In this work, we empirically analyze the most popular Rust projects in GitHub, aiming to discover common bug patterns that these projects share.
-Our interest comes from a desire to develop program repair tools, but we believe these insights can also be of broader service.
-Automated program repair tools try to locate bug locations in the source code (using a fault localization module) and then generate patches to fix them, possibly without human intervention, so that the modified program meets desired specifications (implicit or explicit).
-A powerful automated program repair system has many useful applications and could significantly simplify debugging, thus reducing software development cost.
-A set of common bug fix patterns can be a useful input to an APR system: it directly provides information about fixing similar buggy structures, which can significantly reduce the search space for patch generation.
+In this work, with the help of our tool, Ruxanne, we empirically analyze the most popular Rust projects in GitHub.
+Our goal is to discover common bug patterns that afflict these projects.
+Although we are motivated by a desire to develop automated program repair tools, we believe these insights can also be of broader service.
+Automated program repair~\cite{goues2019automated,liu2018survey} tools try to find bug locations in the source code (using a fault localization module) and then generate patches to fix them, possibly without human intervention, so that the modified program meets desired specifications (implicit or explicit).
+A powerful automated program repair system has many useful applications and could significantly simplify debugging, thus reducing software development cost~\cite{le2012systematic}.
+A set of common bug fix patterns can be a useful input to an APR system: it tells how to fix recurring buggy structures, thus reducing the search space for patch generation~\cite{jeffrey2009bugfix}.
 
-Similarly, if we are observing undesired behaviour from our program, that would indicate that there lies a fault within our program.
+If we observe undesired behaviour from our program, then there is likely a fault within our program\footnote{At the extreme, a fault could be caused by moths or other hardware malfunctions.}.
 But where is the fault?
-Fault localization is the process of automatically finding the faulty statement(s) in our program.
-If a fault localization tool knows about the common bug patterns that can appear in the underlying programming language, it can prioritize code segment candidates based on their potentials of producing bugs.
-That would result in a more robust tool that leverages historical data.
+Fault localization~\cite{wong2016survey} is the process of automatically finding faulty statements in a program.
+If a fault localization tool knows about the common bug patterns that can appear in the underlying programming language, it can prioritize candidates based on their bug-producing potential.
+That would result in a more robust tool---one that leverages historical data.
 
-Moreover, knowing which code patterns in a programming language are more susceptible to bugs can help programmers to be more cautious while working with similar code structures.
-This knowledge can make the developers avoid certain mistakes during production which ultimately would result in reducing the frequent bug patterns within the program.
-As Monperrus describes, the fewer frequent bug patterns we keep in our software, the less cost we would have to pay for software maintenance.
+Moreover, knowing that certain code patterns in a programming language are more susceptible to bugs enables projects to set policies that avoid such patterns.
+For instance, in the context of C and C++, Cannon et al~\cite{Cannon1997RecommendedCS} recommend that, if at least one of the if/else sections is a compound statement, requiring braces, then both sections should have braces (i.e.
+they should be fully bracketed).
+This recommendation presumably comes from qualitative experience, as it pre-dates large-scale empirical studies.
+A project can choose to require that contributions follow such recommendations.
+Research such as ours can help formulate data-driven recommendations that reduce the number of patterns empirically linked to bugs.
+Per Monperrus, the fewer frequent bug patterns we keep in our software, the lower cost we can expect to pay for software maintenance~\cite{monperrus2014critical}.
 
-Moving further afield, researchers have been exploiting deep learning techniques to learn about the semantics of computer programs and create program modifier models that have a specific goal (e.g.
-bug fix).
-However, most learning tools get fixed-size vectors as input and map them to outputs.
-Unlike image data, it is challenging to compress programs into vectors without any semantic information loss.
-Current approaches try to define the vector dimensions in a way that accounts for all possible abstract syntax tree forms.
+Moving further afield, researchers have been exploiting deep learning techniques to create program modifier models, which can then be used for specific goals (e.g.
+bug fixes)~\cite{alon2019code2vec,alon2018general,raychev2016learning,bielik2016phog}.
+However, most extant learning tools take fixed-size vectors as inputs.
+The process of compressing a variably-sized program to a fixed-sized vector is called code embedding~\cite{chen2019literature}.
+Unlike with image data, it is challenging to compress programs into vectors without any semantic information loss.
+Current approaches try to define the dimensions of the fixed-sized vectors in a way that accounts for all possible abstract syntax tree forms.
 That results in a large, sparse space for the inputs.
-Assuming that the deep learning model we are trying to develop is associated with bug fixing tasks, then knowledge about common bug fix patterns can help prune the input space towards containing fewer dimensions.
-That can help with creating more efficient models.
+Assuming that the deep learning model we are trying to develop is used for bug fixing, then knowledge about common bug fix patterns can help prune the dimensionality of the input space, which can help create more efficient embeddings.
 
-Many studies have investigated detecting and categorizing bug fix patterns in general, or for a specific programming language.
+Many studies have investigated detecting and categorizing bug fix patterns in general~\cite{islam2021changes,madeiral2018towards,pan2009toward}, or for a specific programming language~\cite{yang2022mining,hanam2016discovering,campos2019discovering}.
 In this work, we study common bug patterns that appear in Rust projects.
-Rust is a relatively new language that allows programmers to write programs that have low-level control over resources, as well as high-level safety guarantees.
+Rust is a relatively new programming language~\cite{klabnik2019rust} which allows programmers to write programs that have low-level control over resources, while still ensuring high-level safety guarantees (for programs written in safe Rust).
 What distinguishes Rust from the other systems programming languages, e.g.
@@
 The ownership framework keeps track of references to different memory locations, and will free them once they go out of scope.
-If a program passes the compiler’s checks, it must meet several safety properties; e.g.
+If a program passes the compiler’s checks, it meets several safety properties by construction; e.g.
 it has no dangling references or double-frees.
-These problems had to be manually ensured in languages like C/C++.
-Moreover, not only does the ownership concept remove the need for a garbage collector in the runtime environment, it does not require the programmer to explicitly do certain safety tasks.
-Due to the memory and thread safety guarantees that Rust provides, many software companies adopted Rust in production.
-For example, Dropbox decided to migrate to Rust for developing their storage system, since they were not completely satisfied with the features provided by AWS’s S3 or Golang.
+Such properties had to be manually ensured (or not!) in languages like C/C++.
+Moreover, not only does ownership remove the need for a garbage collector in the runtime environment, it frees the programmer from the need to explicitly do certain safety-related tasks~\cite{qin2020understanding}.
+Due to the memory and thread safety guarantees that Rust provides, many software companies have adopted Rust in production.
+For example, Dropbox decided to migrate to Rust for developing their storage system, since they were not completely satisfied with the features provided by AWS’s S3 or Golang~\cite{dropbox}.
 
-In Rust, the so-called Borrow Checker component enforces ownership rules.
-Because of the Borrow Checker's importance in Rust, we split patterns into two groups: (1) non-borrow-checker related patterns (General Patterns), and (2) borrow-checker related patterns (BC-Related Patterns).
-We present our discovered insights of our empirical study in the following four research
-questions (RQs):
+The Rust compiler's Borrow-Checker component enforces ownership rules and hence many of Rust's safety properties.
+Because of the Borrow-Checker's importance in Rust, we split the patterns that we present into two groups: (1) non-borrow-checker related patterns (General Patterns), and (2) borrow-checker related patterns (BC-Related Patterns).
+\\
 
-\textbf{RQ1. How successful is our data modelling approach in representing program change?}
+\noindent \textbf{Research Questions} We formulated three research questions (RQs) which guided our empirical study about Rust bug patterns:
 
-\textbf{RQ2. What are the general fix patterns in Rust, and how often do they apply?}
+\begin{itemize}
+    \item \textbf{RQ1. Does our data modelling approach capture the most important aspects of bug-fixing program changes?} 
+    \item \textbf{RQ2. What are the general fix patterns in Rust, and how often do they apply?} 
+    \item \textbf{RQ3. What are borrow-checker related fix patterns in Rust, and how often do they apply?}
+\end{itemize}
 
-\textbf{RQ3. What are borrow-checker related fix patterns in Rust, and how often do they apply?}
+\noindent \textbf{Contributions} This paper makes the following primary contributions:
+
+\begin{itemize}
+    \item To the best of our knowledge, this work is the first to exploits a code analysis pipeline to automatically mine Rust open source projects and extract common cross-project bug fix patterns.
+    \item We propose a novel method to encode key AST information in fixed size datapoints using a semi-automatically derived weighting scheme.
+    \item To respect the key role of the borrow checker in Rust, we report and discuss bug fix patterns that are related to the ownership framework in Rust separately from general patterns.
+\end{itemize}

### 2022-11-02 main.tex
shuffle package order
flattened everything ;((
methodology comments; probably need to move things around
added Ruxanne as the name

@@
 \lstset{escapeinside={<@}{@>}}
-\usepackage{hyperref}
 \usepackage{tabularx}
@@
 % \addbibresource{bibtex.bib}
+\usepackage{todonotes}
+\usepackage{hyperref}
 \usepackage{subfiles} % Best loaded last in the preamble
@@
 Yet, to the best of our knowledge, no research has sought to identify the most pervasive bug fix patterns within Rust programs.
-In this project, we developed a novel embedding of Rust code into fixed-sized vectors.
-Using this embedding, we mined the top 18 most-starred Rust projects in Github to discover the most common bug fix patterns committed to their repositories.
+In this project, we introduce Ruxanne, a tool for analyzing and extracting fix patterns in Rust.
+Ruxanne uses a novel embedding of Rust code into fixed-sized vectors.
+Through Ruxanne, we mined the top 18 most-starred Rust projects in Github to discover the most common bug fix patterns committed to their repositories.
 We analyzed 87,726 code changes drawn from 57,214 commits across these 18 projects.
-After clustering the code changes, and conducting a manual analysis, we discovered 19 groups of cross-project bug fix patterns, which we categorize as (1) general patterns and (2) borrow-checker-related patterns.
+After clustering the code changes, and conducting a manual analysis, we discovered 20 groups of cross-project bug fix patterns, which we categorize as (1) general patterns and (2) borrow-checker-related patterns.
 We describe each of these patterns and their implications to automated program repair.
@@
 \section{Introduction}
-% rubber: depend sections/introduction.tex
-\subfile{sections/introduction}
+% rubber: depend introduction.tex
+\subfile{introduction}
 
 \section{Methodology}
-% rubber: depend sections/methodology.tex
-\subfile{sections/methodology}
+% rubber: depend methodology.tex
+\subfile{methodology}
 
 \section{Results}
-% rubber: depend sections/results.tex
-\subfile{sections/results}
+% rubber: depend results.tex
+\subfile{results}
 
 \section{Discussion}
-% rubber: depend sections/discussion.tex
-\subfile{sections/discussion}
+% rubber: depend discussion.tex
+\subfile{discussion}
 
 \section{Related Work}
-% rubber: depend sections/related.tex
-\subfile{sections/related}
+% rubber: depend related.tex
+\subfile{related}
 
 \section{Conclusion}
-% rubber: depend sections/conclusion.tex
-\subfile{sections/conclusion}
+% rubber: depend conclusion.tex
+\subfile{conclusion}
 
@@
 \section{Appendix}
-\subfile{sections/appendix}
+\subfile{appendix}
 

### 2022-11-02 methodology.tex
fixed discussion, results todos, added two G patterns, dropped one BC pattern
added Ruxanne as the name

@@
 In this section, we describe our methodology, providing full details in later subsections.
-Our pipeline for finding classes of pervasive patterns in Rust programs parses program versions, computes differences between them, embeds them in fixed-size datapoints, and then clusters these datapoints.
+Ruxanne uses a pipeline for finding classes of pervasive patterns in Rust programs which parses program versions, computes differences between them, embeds them in fixed-size datapoints, and then clusters these datapoints.
 After conducting a manual analysis on the obtained clusters, we refine them towards manifesting bug fix patterns.
@@
 
-We decided to implement our pipeline completely in Python.
+We decided to implement Ruxanne completely in Python.
 Our target repositories were the top 18 most-starred open source Rust projects on GitHub as of August 2021.
@@
     \draw[->]  (c1) -- (d1);
+    \draw[->]  (d1) -- (d2);
 
@@
 The third part is the content of the change---terminals and nonterminals.
-Our tool\todo{do we have a name?
-It's important to have a name} looks for changes within each root-level scope (Rust book calls these root-level scopes the items of a crate\footnote{\url{https://doc.rust-lang.org/reference/items.html}}).
+Ruxanne looks for changes within each root-level scope (Rust book calls these root-level scopes the items of a crate\footnote{\url{https://doc.rust-lang.org/reference/items.html}}).
 That is, our assumption is that each change pattern occurs within one root-level scope.
@@
         \draw[->]  (c1) -- (d1);
+        \draw[->]  (d1) -- (d2);
 

### 2022-11-03 discussion.tex
added Ruxanne as the name
added minor changes

@@
-In this work, our pipeline successfully mined 20 cross project bug fix patterns.
+In this work, through Ruxanne we successfully mined 20 cross project bug fix patterns.
 We introduced these patterns in two groups: 12 general patterns and 8 BC-related patterns.
@@
 Also, the introduced patterns are all cross-project fix patterns.
-Each fix pattern has been seen at least in three different projects, indicating their commonality.
+Each fix pattern has been seen at least in three different projects, indicating its commonality.
 

### 2022-11-03 results.tex
added minor changes

@@
 We used $\epsilon=0.0001$ and $Z=5$ as our DBSCAN parameters and that resulted in 577 clusters.
-After manual analysis, as described in~\ref{sec:manual_analysis_cluster_selection}, we ended up with 10 general patterns, which we describe in Table~\ref{table:general}.
+After manual analysis, as described in~\ref{sec:manual_analysis_cluster_selection}, we ended up with 12 general patterns, which we describe in Table~\ref{table:general}.
 As it is shown in the table, we have separated the patterns based underlying program element (Category column).
@@
 We used $\epsilon=0.001$ and $Z=5$ as our DBSCAN parameters and that resulted in 102 clusters.
-After manual analysis, as described in~\ref{sec:manual_analysis_cluster_selection}, we ended up with 9 borrow-checker related fix patterns, which we, similar to the previous table, describe in Table~\ref{table:bc}.
+After manual analysis, as described in~\ref{sec:manual_analysis_cluster_selection}, we ended up with 8 borrow-checker related fix patterns, which we, similar to the previous table, describe in Table~\ref{table:bc}.
 Here we provide detailed description for each of these patterns:
@@
 
-let displayed_attributes = schema.clone().map(|s| {
-    s.displayed_name()
-        .iter()
-<@\textcolor{red}{--}@>      .map(|s| (*s).to_string())
-<@\textcolor{green}{++}@>      .map(|s| s.to_string())
-        .collect::<HashSet<String>>()
-});
 \end{lstlisting}

### 2022-11-03 discussion.tex
changes to be made

@@
 
-Furthermore, we expected to obtain certain patterns regarding some certain Rust related features, such as unsafe Rust, but our results did not include those patterns.
+Furthermore, we expected to obtain certain patterns regarding some certain Rust related features, such as \verb+move+, but our results did not include those patterns.
 Our speculation is that, since Rust has a strong compiler check before creating the program executable, the programmers usually get many errors, and they only push their code once it is free of compile errors.

### 2022-11-03 results.tex
changes to be made

@@
 
-\todo[inline]{Can we justify the 50, both here and in the previous section? ``It was a random number not so large and not too small, how can I justify it?''}
+\todo[inline]{Can we justify the 50, both here and in the previous section?
+``It was a random number not so large and not too small, how can I justify it?'' Ask about calculating error, give yourself more credit for 50 before}
 
@@
 ``The numbers reported are the intersection'' when is our approach effective and when is it not?
-``added above'' more detailed stats would probably be good too ``what kind of stats do you have in mind?''}
+``added above'' more detailed stats would probably be good too ``what kind of stats do you have in mind?'' add a table for describing why it didn't work}
 
@@
 \end{table}
-\todo[inline]{Also do we compare table 2 to what occurs in previous work? ``What previous work?''}
+\todo[inline]{Also do we compare table 2 to what occurs in previous work? ``What previous work?'' add languages per desc}
 
@@
 Is it true that, if it compiles, behaviour is preserved?
-``Yes the assumption is that if the presence of clone() is unnecessary, then removing it would not change the program behaviour''}
+``Yes the assumption is that if the presence of clone() is unnecessary, then removing it would not change the program behaviour'' rephrase tautology}
 

### 2022-11-08 introduction.tex
some changes

@@
 A set of common bug fix patterns can be a useful input to an APR system: it tells how to fix recurring buggy structures, thus reducing the search space for patch generation~\cite{jeffrey2009bugfix}.
+Yet, to the best of our knowledge, no research has sought to identify the most pervasive bug fix patterns within Rust programs.
+In this project, we introduce Ruxanne, a tool for analyzing and extracting fix patterns in Rust.
+Ruxanne uses a novel embedding of Rust code into fixed-sized vectors.
+Through Ruxanne, we mined the top 18 most-starred Rust projects in Github to discover the most common bug fix patterns committed to their repositories.
+We analyzed 87,726 code changes drawn from 57,214 commits across these 18 projects.
+After clustering the code changes, and conducting a manual analysis, we discovered 20 groups of cross-project bug fix patterns, which we categorize as (1) general patterns and (2) borrow-checker-related patterns.
+We describe each of these patterns and their implications to automated program repair.
 

### 2022-11-08 methodology.tex
changes to be made
some changes

@@
 Our target repositories were the top 18 most-starred open source Rust projects on GitHub as of August 2021.
-We mined (using Pydriller) all the bug related commits and ran them through our pipeline; Section~\ref{sec:data_modelling} describes how we extracted fixed-size datapoints from the code changes.
-Section~\ref{sec:mining} describes our mining methodology in depth.
+We mined (using Pydriller) all the bug related commits and ran them through our pipeline; Section~\ref{sec:data_modelling} describes how we extracted fixed-size datapoints from the code changes, and Section~\ref{sec:mining} describes our mining methodology in depth.
 Then, we applied the DBSCAN clustering algorithm~\cite{ester1996density} on the resulting datapoints.
@@
 While we can implement a specific workaround to distinguish nested ifs (and we have done so), the general point about the large dimensionality of fully order-sensitive embedding still holds.
-\todo[inline]{Worse than that, isn't $d$ unbounded? ``yes it is, but doesn't what we've said above still hold?''}
+\todo[inline]{Worse than that, isn't $d$ unbounded? ``yes it is, but doesn't what we've said above still hold?'' combinatorial explosion}
 
@@
 The reason behind this decision is because a different order of program elements yields a different number of occurrences of underlying dimension items\todo{is there an example?
-``Fig 2?''}.
+``Fig 2?'' introduction to the next section}.
 Second, we semi-automatically design a weighting scheme to prioritize non terminals that we think are more salient for Rust bugs.

### 2022-11-09 methodology.tex
changes added before results

@@
 The third part is the content of the change---terminals and nonterminals.
-Ruxanne looks for changes within each root-level scope (Rust book calls these root-level scopes the items of a crate\footnote{\url{https://doc.rust-lang.org/reference/items.html}}).
+Ruxanne looks for changes within each root-level scope (Rust book calls these root-level scopes the \textit{items} of a crate\footnote{\url{https://doc.rust-lang.org/reference/items.html}}).
 That is, our assumption is that each change pattern occurs within one root-level scope.
@@
 In such an embedding, for instance, there is no difference between two nested if statements and two if statements beside each other.
-On the other hand, theoretically, for full order-sensitiveness, we would need to embed all possible combinations of items from $\mathtt{NT} \cup \mathtt{BCE}$ as our dimensions, which yields an infeasible $O(n^d)$ dimensions ($n=|\mathtt{NT} \cup \mathtt{BCE}|$ and $d$ is the depth of nested structures); this could potentially be reduced somewhat, but is still impractical.
+On the other hand, theoretically, for full order-sensitiveness, we would need to embed all possible combinations of items from $\mathtt{NT} \cup \mathtt{BCE}$ as our dimensions, which yields a combinatorial explosion; this could potentially be reduced somewhat, but is still impractical.
 While we can implement a specific workaround to distinguish nested ifs (and we have done so), the general point about the large dimensionality of fully order-sensitive embedding still holds.
-\todo[inline]{Worse than that, isn't $d$ unbounded? ``yes it is, but doesn't what we've said above still hold?'' combinatorial explosion}
 
@@
 First, we record the number of occurrences of items within the paths at each scope.
-The reason behind this decision is because a different order of program elements yields a different number of occurrences of underlying dimension items\todo{is there an example?
-``Fig 2?'' introduction to the next section}.
+The reason behind this decision is because a different order of program elements yields a different number of occurrences of underlying dimension.
 Second, we semi-automatically design a weighting scheme to prioritize non terminals that we think are more salient for Rust bugs.
+In Section~\ref{sec:weighting_scheme}, we provide and example on how this approach works.
 
@@
 We did try k-means and its more powerful fuzzy variant Fuzzy C-means~\cite{dunn1973fuzzy} and did not get useful clusters.
+\todo[inline]{Wouldn't the reviewers complain why you haven't put k-means data in the paper?}
 
@@
 In each subplot, the two blue bars in the left respectively indicate the number of clustered ($G_c$) and noise ($G_n)$ datapoints (the number is shown above each bar) after applying clustering on general database $D_g$.
-The two orange bars in the right show similar data but for the borrow-checker $D_{b}$ database.
+The two orange bars in the right show similar data but for the borrow-checker database $D_{b}$.
 In each subplot's legend, $C_g$ and $C_b$ show the number of clusters obtained from $D_g$ and $D_b$, respectively}
@@
 An abstract pattern as opposed to a more-specific pattern can account for many changes in the programs.
-For instance, we deem the pattern `Changing a clone of a variable to a borrowing of it in a function argument' more specific than `Adding a new statement to a function's body':
+For instance, we deem the pattern `Changing a clone of a variable to a borrowing of it in a function argument' more specific than `Changing a statement in a function's body':
 
@@
 fn foo2() {
-    stmt1;
+--  stmt1;
 ++  stmt2;
@@
 If the cluster had fewer than 50 datapoints, we would analyze all of the datapoints in the cluster.
-Next, we read the code and the bug reports related to all the datapoints (if they existed).
+Regarding general patterns, we obtained 577 clusters, out of which 428 included cross-project patterns.
+Out of these 428 clusters, 35 clusters included more than 50 points.
+That is, we did not do random sampling for 393 clusters, we just manually analyzed all the datapoints in them.
+Regarding BC-related patterns, out of 102 obtained clusters, 49 clusters included cross-project pattern.
+Only 4 of these 49 clusters included more than 50 datapoints, therefore, we only did a random sampling for 4 clusters.
+For each cluster, we read the code and the bug reports of the datapoints (if they existed).
 If we judged that the members of a cluster fit into a pattern, we wrote a natural language description of that cluster.
@@
 We considered these candidates and carried out appropriate ones.
-Finally, we selected 19 of the remaining clusters---the ones that manifested concrete changes.
+Finally, we selected 20 of the remaining clusters---the ones that manifested more-specific changes.
 Sections~\ref{sec:common_patterns}~and~\ref{sec:bc_patterns} present our clusters.

### 2022-11-10 results.tex
revised results section a bit

--- 2022-11-03 results.tex
@@
 
-We choose 50 random samples from each of our general and borrow-checker databases.
-These random samples are drawn from the clustered datapoints.
+We collected random samples 1000 times.
+In each iteration, we collected 50 sample datapoints from our general and borrow-checker databases (25 each).
+For each database, out of the 1000 sample sets, we picked the one that minimized the sampling error.
+The sampling error is defined as the absolute difference between the sample mean and the total mean.
+Also, these random samples are drawn from the clustered datapoints.
 The reason behind this decision is the same as why we collected a lot of noise points, that is, the majority of the commits contain multi-hunk complex changes and the key changing elements are difficult to determine.
 
-\todo[inline]{Can we justify the 50, both here and in the previous section? ``It was a random number not so large and not too small, how can I justify it?''}
+We carried out a manual experiment for evaluating our code embedding.
+In the experiment, the analyzer should rank the most important element based on the provided circlepack figure and the change's source code.
+Based on these ranks, then we calculate Top-$n$ Accuracy for $n=1$ up to $n=5$.
+Both authors carried out the experiment independently.
+We report the results as the portion of the changes which both authors agreed upon its embedding correctness.
+The final results are shown in Figure~\ref{fig:rq1}.
 
-We carried out two manual experiments for evaluating our embedding mechanism.
-In the first experiment, the analyzer investigates whether the most important element has the largest radius or not.
-In the second experiment, the analyzer investigates whether the most important element is included in three largest circles or not.
-The analyzer should give a binary answer to these questions.
-Both authors carried out these experiments independently.
-We report the results as the portion of the changes which both authors agreed upon its embedding correctness.
-The final results are 54\% and 80\% correctness in the first and second experiments, respectively.
+\begin{figure}[h]
+\centering
 
-Although reporting most important element within the top three elements in 80\% of the cases is an accpectable outcome, we examined our pipeline for the other 20\%.
+\begin{tikzpicture}
+\begin{axis}[
+    xlabel={$n$},
+    ylabel={Top-$n$ Accuracy (\%)},
+    xmin=1, xmax=5,
+    ymin=60, ymax=100,
+    xtick={1,2,3,4,5},
+    ytick={60,70,80,90,100},
+    legend pos=south east,
+    ymajorgrids=true,
+    grid style=dashed,
+]
+
+\addplot[
+    color=green,
+    mark=o,
+    ]
+    coordinates { 
+    (1,72)
+    (2,84)
+    (3,92)
+    (4,96)
+    (5,96)
+    };
+    \addlegendentry{General}
+
+\addplot[
+    color=red,
+    mark=square,
+    ]
+    coordinates { 
+    (1,68)
+    (2,80)
+    (3,88)
+    (4,96)
+    (5,100)
+    };
+    \addlegendentry{Borrow-Checker}
+
+\addplot[
+    color=blue,
+    mark=triangle,
+    ]
+    coordinates { 
+    (1,70)
+    (2,82)
+    (3,90)
+    (4,96)
+    (5,98)
+    };
+    \addlegendentry{Total}
+
+\end{axis}
+\end{tikzpicture}   
+\caption{\label{fig:rq1} Top-$n$ Accuracy of the proposed code embedding approach.
+The plot shows Top-$n$ Accuracy where $n\in\{1,2,3,4,5\}$ for both general and borrow-checker related datapoints.
+As shown in the plot, in more than 95\% of the datapoints, the most important element is seen at least in the first four ranks.}
+
+\end{figure}
+
+Although, on average, reporting most important element within the top three elements in 90\% of the cases is an accpectable outcome, we examined our pipeline for the other 10\%.
 The analysis shows that our pipeline fails in change patterns where the change involves many elements (e.g.
@@
 It is worth mentioning that these changes tend to be less cross-project as having many changed elements make it easier for datapoints to be dissimlar.
-
-\todo[inline]{no, you have to compute the intersection.
-``The numbers reported are the intersection'' when is our approach effective and when is it not?
-``added above'' more detailed stats would probably be good too ``what kind of stats do you have in mind?''}
 
@@
 \end{table}
-\todo[inline]{Also do we compare table 2 to what occurs in previous work? ``What previous work?''}
 
@@
 could the developer have used the variable itself or not.
-If the variable clone is indeed redundant, then removing it does not modify the program behaviour, however it can boost the performance significantly if the program was repeatedly cloning large variables.
+It is evident that removing a redundant clone does not modify the program behaviour, however it can boost the performance significantly if the program was repeatedly cloning large variables.
 \\
-
-\todo[inline]{I think you should talk more about ``does not modify program behaviour''.
-For instance after you make this change the program may no longer compile.
-Is it true that, if it compiles, behaviour is preserved?
-``Yes the assumption is that if the presence of clone() is unnecessary, then removing it would not change the program behaviour''}
 

### 2022-11-10 results.tex
fixed results

@@
 As discussed in the previous section, we used an embedding mechanism to incorporate AST information in fixed-sized datapoints which then could be used in the DBSCAN clustering algorithm.
-We used heuristics (a weighting scheme) while creating our datapoints to make the most important program elements stand out from all the non terminals seen in the AST (along with some BC-related elements).
+We used heuristics (a weighting scheme) while creating our datapoints to make the most impor tant program elements stand out from all the non terminals seen in the AST (along with some BC-related elements).
 Here we evaluate the effectiveness of our code embedding.
@@
 \subsubsection{Struct}
+
+Similar to other languages, such as C, a struct makes it possible for the users to define a custom type that is comprised of different types.
+The patterns that we describe in this section can also be seen in other languages that use the same concept.
+
 \noindent\textit{\textbf{G.struct.field} Adding/Removing a struct field}
@@
 \subsubsection{Types}
+
+Rust is a statically-typed language, and similar to the other statically-typed languages, some of the change patterns are associated with changes in types used in different program structures.
+Here, we introduce four bug fix patterns that are associated with types in Rust:
+
 \vspace{3mm}
@@
 \noindent\textbf{Description:} The trait bounds are the functionalities that we require from our parametric types.
+This concept might be referred to as interfaces in other languages, although there might be small differences.
 This change pattern relates to modifications in trait bounds.
@@
 In Rust, match control flow construct is used for comparing a value against multiple patterns, and executing the code associated with the first pattern that matches with our value.
-A pattern and its associated code are called a match arm.\\
+A pattern and its associated code are called a match arm.
+Match statements can be similar to switch statements in other languages, like C, although match statements in Rust are able to match more complicated patterns.\\
 

### 2022-11-20 results.tex
updated RQ1

@@
     coordinates { 
-    (1,72)
-    (2,84)
-    (3,92)
-    (4,96)
-    (5,96)
+    (1,64)
+    (2,72)
+    (3,84)
+    (4,88)
+    (5,92)
     };
@@
     (1,68)
-    (2,80)
+    (2,84)
     (3,88)
-    (4,96)
-    (5,100)
+    (4,92)
+    (5,96)
     };
@@
     coordinates { 
-    (1,70)
-    (2,82)
-    (3,90)
-    (4,96)
-    (5,98)
+    (1,66)
+    (2,78)
+    (3,86)
+    (4,90)
+    (5,94)
     };
@@
 The plot shows Top-$n$ Accuracy where $n\in\{1,2,3,4,5\}$ for both general and borrow-checker related datapoints.
-As shown in the plot, in more than 95\% of the datapoints, the most important element is seen at least in the first four ranks.}
+As shown in the plot, in more than 85\% of the datapoints, the most important element is seen at least in the first three ranks.}
 
@@
 
-Although, on average, reporting most important element within the top three elements in 90\% of the cases is an accpectable outcome, we examined our pipeline for the other 10\%.
+Although, on average, reporting most important element within the top four elements in 90\% of the cases is an accpectable outcome, we examined our pipeline for the other 10\%.
 The analysis shows that our pipeline fails in change patterns where the change involves many elements (e.g.

### 2022-11-22 introduction.tex
adopted apa style

@@
 One step towards that aim has been studying common bug patterns (and their corresponding fixes) across multiple software systems.
-Research efforts categorizing common bug patterns date back to at least 1975, where Endres tried to categorize bug patterns in operating system implementations~\cite{endres1975analysis}.
+Research efforts categorizing common bug patterns date back to at least 1975, where Endres tried to categorize bug patterns in operating system implementations~\citep{endres1975analysis}.
 
@@
 Although we are motivated by a desire to develop automated program repair tools, we believe these insights can also be of broader service.
-Automated program repair~\cite{goues2019automated,liu2018survey} tools try to find bug locations in the source code (using a fault localization module) and then generate patches to fix them, possibly without human intervention, so that the modified program meets desired specifications (implicit or explicit).
-A powerful automated program repair system has many useful applications and could significantly simplify debugging, thus reducing software development cost~\cite{le2012systematic}.
-A set of common bug fix patterns can be a useful input to an APR system: it tells how to fix recurring buggy structures, thus reducing the search space for patch generation~\cite{jeffrey2009bugfix}.
+Automated program repair~\citep{goues2019automated,liu2018survey} tools try to find bug locations in the source code (using a fault localization module) and then generate patches to fix them, possibly without human intervention, so that the modified program meets desired specifications (implicit or explicit).
+A powerful automated program repair system has many useful applications and could significantly simplify debugging, thus reducing software development cost~\citep{le2012systematic}.
+A set of common bug fix patterns can be a useful input to an APR system: it tells how to fix recurring buggy structures, thus reducing the search space for patch generation~\citep{jeffrey2009bugfix}.
 Yet, to the best of our knowledge, no research has sought to identify the most pervasive bug fix patterns within Rust programs.
@@
 But where is the fault?
-Fault localization~\cite{wong2016survey} is the process of automatically finding faulty statements in a program.
+Fault localization~\citep{wong2016survey} is the process of automatically finding faulty statements in a program.
 If a fault localization tool knows about the common bug patterns that can appear in the underlying programming language, it can prioritize candidates based on their bug-producing potential.
@@
 Moreover, knowing that certain code patterns in a programming language are more susceptible to bugs enables projects to set policies that avoid such patterns.
-For instance, in the context of C and C++, Cannon et al~\cite{Cannon1997RecommendedCS} recommend that, if at least one of the if/else sections is a compound statement, requiring braces, then both sections should have braces (i.e.
+For instance, in the context of C and C++, \cite{Cannon1997RecommendedCS} recommend that, if at least one of the if/else sections is a compound statement, requiring braces, then both sections should have braces (i.e.
 they should be fully bracketed).
@@
 Research such as ours can help formulate data-driven recommendations that reduce the number of patterns empirically linked to bugs.
-Per Monperrus, the fewer frequent bug patterns we keep in our software, the lower cost we can expect to pay for software maintenance~\cite{monperrus2014critical}.
+Per Monperrus, the fewer frequent bug patterns we keep in our software, the lower cost we can expect to pay for software maintenance~\citep{monperrus2014critical}.
 
 Moving further afield, researchers have been exploiting deep learning techniques to create program modifier models, which can then be used for specific goals (e.g.
-bug fixes)~\cite{alon2019code2vec,alon2018general,raychev2016learning,bielik2016phog}.
+bug fixes)~\citep{alon2019code2vec,alon2018general,raychev2016learning,bielik2016phog}.
 However, most extant learning tools take fixed-size vectors as inputs.
-The process of compressing a variably-sized program to a fixed-sized vector is called code embedding~\cite{chen2019literature}.
+The process of compressing a variably-sized program to a fixed-sized vector is called code embedding~\citep{chen2019literature}.
 Unlike with image data, it is challenging to compress programs into vectors without any semantic information loss.
@@
 
-Many studies have investigated detecting and categorizing bug fix patterns in general~\cite{islam2021changes,madeiral2018towards,pan2009toward}, or for a specific programming language~\cite{yang2022mining,hanam2016discovering,campos2019discovering}.
+Many studies have investigated detecting and categorizing bug fix patterns in general~\citep{islam2021changes,madeiral2018towards,pan2009toward}, or for a specific programming language~\citep{yang2022mining,hanam2016discovering,campos2019discovering}.
 In this work, we study common bug patterns that appear in Rust projects.
-Rust is a relatively new programming language~\cite{klabnik2019rust} which allows programmers to write programs that have low-level control over resources, while still ensuring high-level safety guarantees (for programs written in safe Rust).
+Rust is a relatively new programming language~\citep{klabnik2019rust} which allows programmers to write programs that have low-level control over resources, while still ensuring high-level safety guarantees (for programs written in safe Rust).
 What distinguishes Rust from the other systems programming languages, e.g.
@@
 Such properties had to be manually ensured (or not!) in languages like C/C++.
-Moreover, not only does ownership remove the need for a garbage collector in the runtime environment, it frees the programmer from the need to explicitly do certain safety-related tasks~\cite{qin2020understanding}.
+Moreover, not only does ownership remove the need for a garbage collector in the runtime environment, it frees the programmer from the need to explicitly do certain safety-related tasks~\citep{qin2020understanding}.
 Due to the memory and thread safety guarantees that Rust provides, many software companies have adopted Rust in production.
-For example, Dropbox decided to migrate to Rust for developing their storage system, since they were not completely satisfied with the features provided by AWS’s S3 or Golang~\cite{dropbox}.
+For example, Dropbox decided to migrate to Rust for developing their storage system, since they were not completely satisfied with the features provided by AWS’s S3 or Golang~\citep{dropbox}.
 

### 2022-11-22 main.tex
added minor changes
some changes
revised results section a bit
adopted apa style

@@
 \usepackage{tikz}
+\usepackage{pgfplots}
 \usetikzlibrary{graphs} 
@@
 \usepackage{multirow}
-% \usepackage[backend=bibtex, style=authoryear]{biblatex}
+% \usepackage[backend=bibtex, style=authoryear]{apa}
+\usepackage{natbib}
 % \addbibresource{bibtex.bib}
@@
 We describe each of these patterns and their implications to automated program repair.
-\keywords{First keyword \and Second keyword \and More}
+\todo[inline]{Do we need this last sentence?}
+\keywords{Bug patterns \and Pattern mining \and Bug fix changes \and Rust}
 % \PACS{PACS code1 \and PACS code2 \and more}
-\subclass{MSC code1 \and MSC code2 \and more}
+\subclass{68N01 \and 68N15} %% Computer Science : General and Programming Languages
 \end{abstract}
@@
 % BibTeX users please use one of
-\bibliographystyle{plain}      % basic style, author-year citations
+\bibliographystyle{apa}      % basic style, author-year citations
 % \bibliographystyle{spmpsci}      % mathematics and physical sciences

### 2022-11-22 methodology.tex
adopted apa style

@@
 Our approach for automatically finding pervasive code patterns aims to find clusters of similar code patterns, using existing implementations of appropriate clustering algorithms.
-Work in this vein generally makes the assumption that clusters corresponds to classes of change patterns~\cite{hanam2016discovering,campos2019discovering,yang2022mining}.
+Work in this vein generally makes the assumption that clusters corresponds to classes of change patterns~\citep{hanam2016discovering,campos2019discovering,yang2022mining}.
 We thus want to associate code changes with datapoints; one of our contributions is a fixed-size embedding which highlights the important parts of a code change.
@@
 We mined (using Pydriller) all the bug related commits and ran them through our pipeline; Section~\ref{sec:data_modelling} describes how we extracted fixed-size datapoints from the code changes, and Section~\ref{sec:mining} describes our mining methodology in depth.
-Then, we applied the DBSCAN clustering algorithm~\cite{ester1996density} on the resulting datapoints.
+Then, we applied the DBSCAN clustering algorithm~\citep{ester1996density} on the resulting datapoints.
 Section~\ref{sec:clustering_data} describes clustering in detail, justifies our choice of DBSCAN for clustering, and explains how we tuned it to improve cluster quality.
@@
 First, we need to define a fixed set of columns.
-We need fixed-size datapoints, as they are a requirement for well-known centroid-based and density-based clustering algorithms~\cite{xu2005survey}.
+We need fixed-size datapoints, as they are a requirement for well-known centroid-based and density-based clustering algorithms~\citep{xu2005survey}.
 The number of columns controls the amount of information we can embed in the datapoints.
@@
 A reasonable work around would be to reduce $n$ from the full set of non terminals to a smaller set of categories (e.g.
-a category for larger entites like class or function definitions, and a category for smaller entities like statements and expressions), and then to record all possible combinations of these categories (similar to~\cite{hanam2016discovering}).
+a category for larger entites like class or function definitions, and a category for smaller entities like statements and expressions), and then to record all possible combinations of these categories (similar to \cite{hanam2016discovering}).
 That approach reduces the number of dimensions and provides more reasonable and syntactically-correct combinations.
@@
 
-We used the Pydriller~\cite{spadini2018pydriller} library to mine software repositories.
+We used the Pydriller~\citep{spadini2018pydriller} library to mine software repositories.
 Pydriller provides APIs to extract commits from a Git repository and to search through different revisions of files.
@@
 For each repository, we search all commits that include bug-fixing related keywords within their commit messages.
-Here, we chose bug-fixing related keywords based on the set introduced in~\cite{zhang2018empirical}, excluding `nan' and `inf'.
+Here, we chose bug-fixing related keywords based on the set introduced in \cite{zhang2018empirical}, excluding `nan' and `inf'.
 
@@
 Having collected our datapoints, our next task is to cluster them so that we can categorize bug patterns.
-Like~\cite{hanam2016discovering}, we use the DBSCAN clustering algorithm for two main reasons.
+Like \cite{hanam2016discovering}, we use the DBSCAN clustering algorithm for two main reasons.
 Firstly, in contrast to k-means, it does not require the number of clusters in advance as an input to the algorithm.
 Secondly, it is a density-based clustering method, which means that it detects arbitrarily shaped clusters; centroid-based methods like k-means can't detect such clusters.
-We did try k-means and its more powerful fuzzy variant Fuzzy C-means~\cite{dunn1973fuzzy} and did not get useful clusters.
+We did try k-means and its more powerful fuzzy variant Fuzzy C-means~\citep{dunn1973fuzzy} and did not get useful clusters.
 \todo[inline]{Wouldn't the reviewers complain why you haven't put k-means data in the paper?}
@@
 
-Like Yang et al~\cite{yang2022mining}, and following the definitions of Cotroneo et al.~\cite{cotroneo2019analyzing}, we divide the clusters in three different groups:
+Like \cite{yang2022mining}, and following the definitions of \cite{cotroneo2019analyzing}, we divide the clusters in three different groups:
 

### 2022-11-22 related.tex
adopted apa style

@@
 Before a repair tool modifies code, it will typically use a fault localization module to find fault locations.
-A fault localization module ranks program statements based on suspiciousness; one example is Tarantula by Jones and Harrold~\cite{jones2005empirical}.
-Suspiciousness is calculated using a statistical model which relies on the observation that buggy statements are executed mostly by failed test cases~\cite{naish2009spectral,xie2013theoretical}.
+A fault localization module ranks program statements based on suspiciousness; one example is Tarantula by \cite{jones2005empirical}.
+Suspiciousness is calculated using a statistical model which relies on the observation that buggy statements are executed mostly by failed test cases~\citep{naish2009spectral,xie2013theoretical}.
 
 Researchers have proposed different methods for code repair, each of which can be said to follow a different mindset.
-Following the taxonomy propsed by~\cite{liu2018survey}, we briefly present two mindsets, and will discuss how having a set of predefined fix patterns can aid repair.
+Following the taxonomy propsed by \cite{liu2018survey}, we briefly present two mindsets, and will discuss how having a set of predefined fix patterns can aid repair.
 
 \subsection{Search-Based Program Repair}
-One line of research in program repair follows the competent programmer hypothesis~\cite{gopinath2014mutant}: syntactically, a faulty program is not that far away from its correct version.
+One line of research in program repair follows the competent programmer hypothesis~\citep{gopinath2014mutant}: syntactically, a faulty program is not that far away from its correct version.
 This hypothesis suggests the search-based program repair mindset.
@@
 
-Use of mutation operators in program repair became popular when Weimer et al.~\cite{forrest2009genetic,nguyen2009using} developed GenProg, a program repair tool that uses genetic programming.
+Use of mutation operators in program repair became popular when \cite{forrest2009genetic} and \cite{nguyen2009using} developed GenProg, a program repair tool that uses genetic programming.
 The main idea behind this tool is that statements follow certain patterns in a codebase.
@@
 
-However, Arcuri and Briand~\cite{arcuri2011practical} observed that GenProg mostly found the patch while carrying out random initialization---that is, before GenProg's evolution begins.
-Their tools TrpAutoRepair~\cite{qi2013efficient} and RSRepair~\cite{qi2014strength} advocate that test case prioritization is necessary while using genetic programming, as fitness evaluation is an expensive part of the repair pipeline.
+However, \cite{arcuri2011practical} observed that GenProg mostly found the patch while carrying out random initialization---that is, before GenProg's evolution begins.
+Their tools TrpAutoRepair~\citep{qi2013efficient} and RSRepair~\citep{qi2014strength} advocate that test case prioritization is necessary while using genetic programming, as fitness evaluation is an expensive part of the repair pipeline.
 
-Tan and Rovehoudhury~\cite{tan2015relifix} used mutation repairing for fixing regression bugs.
+\cite{tan2015relifix} used mutation repairing for fixing regression bugs.
 They categorized common code changes in real-world regressions after studying 73 program evolution benchmarks; our approach also yields a set of common code changes, for Rust, using a different set of changes.
@@
 Thus, having a collection of common bug fix patterns is a crucial step for developing such repair tools.
-Pan et al.~\cite{pan2009toward} analyzed seven large-scale widely used Java projects and obtained 27 common bug fix patterns.
+\cite{pan2009toward} analyzed seven large-scale widely used Java projects and obtained 27 common bug fix patterns.
 They used a bug fix pattern extractor tool to automatically parse and detect bug patterns within the bug fix hunks with less than seven statements.
@@
 Though they did the analysis on Java projects, their reported patterns were not specific to Java.
-Martinez and Monperrus~\cite{martinez2015mining,martinez2012mining} exploited the frequency of observed patterns to introduce a heuristic patch searching method.
+\cite{martinez2015mining,martinez2012mining} exploited the frequency of observed patterns to introduce a heuristic patch searching method.
 In their empirical evaluation, they concluded that choosing repair actions probabilistically (weighted by frequency) can help with reducing the search space, hence creating a more effective repair tool.
 
-Hanam et al.~\cite{hanam2016discovering} conducted similar research, but for finding pervasive bug fix patterns in JavaScript.
+\cite{hanam2016discovering} conducted similar research, but for finding pervasive bug fix patterns in JavaScript.
 They performed a large-scale study of bug fix patterns by mining 105K commits from 134 server-side JavaScript projects.
@@
 
-Yang et al.~\cite{yang2022mining} proposed a mining approach to detect Python bug fix patterns by studying fine-grained fixing code changes.
+\cite{yang2022mining} proposed a mining approach to detect Python bug fix patterns by studying fine-grained fixing code changes.
 They also examined how many bugs could be fixed using automated bug fixing approaches.
@@
 
-Qin et al.~\cite{qin2020understanding} conducted the first empirical study on real-world Rust program behaviours.
+\cite{qin2020understanding} conducted the first empirical study on real-world Rust program behaviours.
 They manually inspected 850 unsafe code usages and 17 bugs across five open-source Rust projects, five Rust libraries, two online security databases, and the Rust standard library.
@@
 
-Li et al.~\cite{li2021mirchecker} present MirChecker, a fully automated bug detection framework for Rust.
+\cite{li2021mirchecker} present MirChecker, a fully automated bug detection framework for Rust.
 This framework works by carrying out static analysis on Rust's Mid-level Intermediate Representation (MIR).

### 2022-11-22 results.tex
updated RQ1 - fixed
adopted apa style

@@
 The analyzer compares the actual change in the code and compares it to the visual representation of the respective data point.
-We used circle pack figures to visualize each data point~\cite{collins2003circle}, generated with the packcircles python library\footnote{\url{https://github.com/mhtchan/packcircles}}.
+We used circle pack figures to visualize each data point~\citep{collins2003circle}, generated with the packcircles python library\footnote{\url{https://github.com/mhtchan/packcircles}}.
 In these figures, the radius of the circle of each non terminal shows its importance in our embedding (as seen in Figure~\ref{fig:essence}).
@@
 Although, on average, reporting most important element within the top four elements in 90\% of the cases is an accpectable outcome, we examined our pipeline for the other 10\%.
-The analysis shows that our pipeline fails in change patterns where the change involves many elements (e.g.
+An analysis shows that our pipeline fails in change patterns where the change involves many elements (e.g.
 adding a whole new \verb+impl+ block and implementing new functions in it).

### 2022-11-22 results.tex
applied txt fixes for results and consecutive sections

@@
 In this section, we present the clusters of fix patterns that we found in our data set.
-\RQBC{} is about the borrow-checker related patterns, while \RQG{} is about general patterns.
+\RQG{} is about the borrow-checker related patterns, while \RQBC{} is about general patterns.
 However, our technique fundamentally relies on our data modelling approach successfully capturing the most important aspects of program changes.
@@
 As discussed in the previous section, we used an embedding mechanism to incorporate AST information in fixed-sized datapoints which then could be used in the DBSCAN clustering algorithm.
-We used heuristics (a weighting scheme) while creating our datapoints to make the most impor tant program elements stand out from all the non terminals seen in the AST (along with some BC-related elements).
+We used heuristics (a weighting scheme) while creating our datapoints to make the most important program elements stand out from all the non terminals seen in the AST (along with some BC-related elements).
 Here we evaluate the effectiveness of our code embedding.
@@
 Also, these random samples are drawn from the clustered datapoints.
-The reason behind this decision is the same as why we collected a lot of noise points, that is, the majority of the commits contain multi-hunk complex changes and the key changing elements are difficult to determine.
+The reason behind this decision is the same as why we collected a lot of noise points, that is, the majority of the commits involve changes in many program elements and the key changing elements are difficult to determine.
 
@@
 After manual analysis, as described in~\ref{sec:manual_analysis_cluster_selection}, we ended up with 12 general patterns, which we describe in Table~\ref{table:general}.
-As it is shown in the table, we have separated the patterns based underlying program element (Category column).
+As it is shown in the table, we have separated the patterns based on the underlying program element (Category column).
 We have given each pattern a specific ID (ID column) and determined the number of instances manifesting that fix pattern (\# Instances column), as well as the number of projects in which the pattern was seen (\# Project column).
@@
 \noindent\textbf{Description:} In this pattern, the attribute set of a struct changes.
-This modification can happen in three different ways: (G.attr.struct.add) adding attributes to the struct; (G.attr.struct.drop) removing attributes from the struct; and (G.attr.struct.change) changing the content of an existing attribute.
-These attributes are Outer Attributes as we can find struct related non-terminals within the ASTDiff of the instances of this pattern.
+This modification can happen in three different ways: Adding attributes to the struct (G.attr.struct.add); Removing attributes from the struct (G.attr.struct.drop); Changing the content of an existing attribute (G.attr.struct.change).
+These attributes are Outer Attributes as we can find struct-related non terminals within the ASTDiff of the instances of this pattern.
 \\
@@
 The patterns that we describe in this section can also be seen in other languages that use the same concept.
+\\
 
@@
 // G.struct.field.add: Adding a new field to a struct
+struct SnapshotService<U, R> {
+    uuid_resolver_handle: R,
+    <@\textcolor{green}{++}@> db_name: String,
+}
+// G.struct.field.drop: Removing a field from a struct
 struct InnerListeners {
@@
 
-// G.struct.field.drop: Removing a field from a struct
-struct SnapshotService<U, R> {
-    uuid_resolver_handle: R,
-    <@\textcolor{green}{++}@> db_name: String,
-}
 \end{lstlisting}
@@
 \begin{lstlisting}[language=Rust, style=colouredRust]
-
 // G.struct.field.pub.add: Adding pub to a field
@@
 \noindent\textbf{Description:} Adding \texttt{pub} to a struct field makes it possible to access that field from external modules.
-In the instances of this cluster, adding the \texttt{pub} access modifier happened in a bug fixing change, where the developer needed to access a field which had not been marked for public access.
+In the instances of this cluster, adding \texttt{pub} access modifier happened in bug fixing changes, where the developer needed to access a field which had not been marked for public access.
 On the other hand, the public access might be revoked if the developer realizes there are no external accesses to the field, and that such accesses are not desirable, e.g.
@@
 \begin{lstlisting}[language=Rust, style=colouredRust]
-
 #[ast_node("MediaRule")]
@@
 The change applies on an object prior to calling \verb+to_string()+ on that object.
-The change might add or remove a dereference, as shown above.
-When adding a dereference, the reason behind the change is that the object is a reference type of \verb+T+, while \verb+T+ directly implements \verb+to_string()+.
+In this change, a dereference operator is added to that object.
+The reason behind the change is that the object is a reference type of \verb+T+, while \verb+T+ directly implements \verb+to_string()+.
 Adding the dereference makes the compiler use the specialized implementation of \verb+to_string()+ and not go through slower string formatting methods.
@@
 
-\noindent\textbf{Description:} This change happens when the developer needs to mutate a variable which was previously immutable in a scope---they thus need to add the mut keyword.
+\noindent\textbf{Description:} This change happens when the developer needs to mutate a variable which was previously immutable in a scope---they thus need to add the \verb+mut+ keyword.
 We never observed this pattern in refactoring changes; it always accompanied a change in program behaviour.

### 2022-11-23 discussion.tex
applied txt fixes for results and consecutive sections
added changes

@@
 We do not consider the bugs that occur during a commit and also fixed during the same commit.
-(3) Our data modelling approach is based on the frequency of observed program elements in ASTs, and we use DBSCAN as our clustering algorithm.
-Other program data modelling methods and clustering algorithms (e.g.SLINK) might output new clusters that our pipeline is unable to find.
+(3) Our code embedding approach is based on the frequency of observed program elements in ASTs, and we use DBSCAN as our clustering algorithm.
+Other program code embedding methods and clustering algorithms (e.g.
+SLINK) might output new clusters that our pipeline is unable to find.
 

### 2022-11-23 introduction.tex
added changes

@@
 One step towards that aim has been studying common bug patterns (and their corresponding fixes) across multiple software systems.
-Research efforts categorizing common bug patterns date back to at least 1975, where Endres tried to categorize bug patterns in operating system implementations~\citep{endres1975analysis}.
+Research efforts categorizing common bug patterns date back to at least 1975, where \cite{endres1975analysis} tried to categorize bug patterns in operating system implementations.
 
@@
 A set of common bug fix patterns can be a useful input to an APR system: it tells how to fix recurring buggy structures, thus reducing the search space for patch generation~\citep{jeffrey2009bugfix}.
-Yet, to the best of our knowledge, no research has sought to identify the most pervasive bug fix patterns within Rust programs.
+Yet, to the best of our knowledge, no research has sought to identify the most pervasive bug fix patterns in Rust programs.
 In this project, we introduce Ruxanne, a tool for analyzing and extracting fix patterns in Rust.
@@
 Such properties had to be manually ensured (or not!) in languages like C/C++.
-Moreover, not only does ownership remove the need for a garbage collector in the runtime environment, it frees the programmer from the need to explicitly do certain safety-related tasks~\citep{qin2020understanding}.
+Moreover, not only does ownership remove the need for a garbage collector in the runtime environment, it frees the programmer of the responsibility to ensure certain safety-related properties~\citep{qin2020understanding}.
 Due to the memory and thread safety guarantees that Rust provides, many software companies have adopted Rust in production.
@@
 \begin{itemize}
-    \item \textbf{RQ1. Does our data modelling approach capture the most important aspects of bug-fixing program changes?} 
+    \item \textbf{RQ1. Does our code embedding approach capture the most important aspects of bug-fixing program changes?} 
     \item \textbf{RQ2. What are the general fix patterns in Rust, and how often do they apply?} 
@@
 \begin{itemize}
-    \item To the best of our knowledge, this work is the first to exploits a code analysis pipeline to automatically mine Rust open source projects and extract common cross-project bug fix patterns.
-    \item We propose a novel method to encode key AST information in fixed size datapoints using a semi-automatically derived weighting scheme.
-    \item To respect the key role of the borrow checker in Rust, we report and discuss bug fix patterns that are related to the ownership framework in Rust separately from general patterns.
+    \item To the best of our knowledge, this work is the first to exploit a code analysis pipeline to automatically mine Rust open source projects and extract common cross-project bug fix patterns.
+    \item This work proposes a novel method to encode key AST information in fixed size datapoints using a semi-automatically derived weighting scheme.
+    \item Keeping in mind the key role of the borrow checker in Rust, this work specifically identifies bug fix patterns that are related to Rust's ownership framework in addition to general bug fix patterns for Rust.
 \end{itemize}

### 2022-11-23 main.tex
added changes

@@
 
-\title{A Study on Common Bug Fix Patterns in Rust%\thanks{Grants or other notes
+\title{A Study of Common Bug Fix Patterns in Rust%\thanks{Grants or other notes
 %about the article that should go on the front page should be
@@
 
-\institute{F. Author \at
-              first address \\
-              Tel.: +123-45-678910\\
-              Fax: +123-45-678910\\
-              \email{fauthor@example.com}           %  \\
+\institute{Mohammad Robati Shirzad \at
+              200 University Ave W, Waterloo, ON N2L 3G1 \\
+              \email{mrobatis@uwaterloo.ca}           %  \\
 %             \emph{Present address:} of F. Author  %  if needed
            \and
-           S. Author \at
-              second address
+           Patrick Lam \at
+                200 University Ave W, Waterloo, ON N2L 3G1 \\
+                \email{patrick.lam@uwaterloo.ca}
 }
@@
 \begin{abstract}
-    Rust is a relatively new programming language which allows programmers to write programs that have low-level control over resources, while still ensuring high-level safety guarantees (for programs written in safe Rust).
+    Rust is a relatively new programming language which allows programmers to write programs that have low-level control over resources while still ensuring high-level safety guarantees (for programs written in safe Rust).
 Rust's ownership framework enables programs to meet these two seemingly-contradictory goals.
-The Rust compiler's Borrow-Checker component enforces the ownership framework requirements which ensure Rust's safety guarantees.
+The Rust compiler's Borrow-Checker component enforces the ownership framework requirements that ensure Rust's safety guarantees.
 
@@
 In this project, we introduce Ruxanne, a tool for analyzing and extracting fix patterns in Rust.
-Ruxanne uses a novel embedding of Rust code into fixed-sized vectors.
-Through Ruxanne, we mined the top 18 most-starred Rust projects in Github to discover the most common bug fix patterns committed to their repositories.
+Ruxanne implements a novel embedding of Rust code into fixed-sized vectors.
+Using Ruxanne, we mined the top 18 most-starred Rust projects in Github to discover the most common bug fix patterns committed to their repositories.
 We analyzed 87,726 code changes drawn from 57,214 commits across these 18 projects.
@@
 We describe each of these patterns and their implications to automated program repair.
-\todo[inline]{Do we need this last sentence?}
 \keywords{Bug patterns \and Pattern mining \and Bug fix changes \and Rust}

### 2022-11-23 methodology.tex
added changes

@@
-% \todo{introductory para saying that we're going to describe our methodology}
-
 In this section, we describe our methodology, providing full details in later subsections.
-Ruxanne uses a pipeline for finding classes of pervasive patterns in Rust programs which parses program versions, computes differences between them, embeds them in fixed-size datapoints, and then clusters these datapoints.
-After conducting a manual analysis on the obtained clusters, we refine them towards manifesting bug fix patterns.
+Ruxanne uses a pipeline for finding classes of pervasive patterns in Rust programs.
+This pipeline parses program versions, computes differences between them, embeds them in fixed-size datapoints, and then clusters these datapoints.
+After conducting a manual analysis on the obtained clusters, we refine them to obtain our proposed bug fix patterns.
 
@@
 A change pattern is a set of code changes that serves a specific purpose.
-This purpose can be improving code readability or maintainability (refactoring patterns), or rectifying program behaviour with respect to a specific functional or non-functional requirement (bug-fixing patterns).
+Purposes include rectifying program behaviour with respect to a specific functional or non-functional requirement (bug-fixing patterns) as well as improving code readability or maintainability (refactoring patterns).
 In this work, we are interested in bug-fixing patterns.
@@
 
-% \todo[inline]{I think we only need to talk about this next bit in a subsection.
-It's an implementation detail, not fundamental to our approach: ``We decided to implement our pipeline completely in Python.
-To make this work, we need a way to get an AST as a Python dictionary.''}
-
-% \todo[inline]{I wonder if this should really go here, or if there should be a shorter version.}
-% To compute the contents of a change, we need two code revisions: the revision before the commit, and the revision after it.
-After parsing these two revisions, we will have two ASTs.
-Tree diff algorithms can compute the differences between two arbitrary trees.
-When the trees are programs' abstract syntax trees, we call their diff an ASTDiff.
-% , encoded in two Python dictionaries; We use dictdiffer to obtain ASTDiffs.
-
-We decided to implement Ruxanne completely in Python.
+We implemented Ruxanne completely in Python.
 Our target repositories were the top 18 most-starred open source Rust projects on GitHub as of August 2021.
-We mined (using Pydriller) all the bug related commits and ran them through our pipeline; Section~\ref{sec:data_modelling} describes how we extracted fixed-size datapoints from the code changes, and Section~\ref{sec:mining} describes our mining methodology in depth.
+We mined (using Pydriller) all the bug related commits and ran them through our pipeline; Section~\ref{sec:code_embedding} describes how we extracted fixed-size datapoints from the code changes, and Section~\ref{sec:mining} describes our mining methodology in depth.
 Then, we applied the DBSCAN clustering algorithm~\citep{ester1996density} on the resulting datapoints.
@@
 
-\subsection{\label{sec:data_modelling}Data modelling}
+\subsection{\label{sec:code_embedding}Code Embedding}
 
-% \todo[inline]{Maybe this next paragraph is what you want at the top of 2.1.
-Then in the next next paragraph you put a shorter description of the pipeline, in a sentence before ``we mined'', and you reference 2.1 there.
-Anyway think about organization and try to have something that makes sense.}
-To compute the contents of a change, we need two code revisions: the revision before the commit, and the revision after it.
+To compute the contents of a change, we need two code revisions: the revision before the change, and the revision after it.
 After parsing these two revisions, we will have two ASTs.
@@
 Sections~\ref{sec:path_extraction}~and~\ref{sec:weighting_scheme} provide a detailed explanation of how we select the most important semantic information out of an ASTDiff, which then allowed us to obtain clusters that contained similar datapoints.
-
-% \todo[inline]{should always be text between subsection and subsubsection. This is the per-change workflow, right?}
 
@@
 
-The output of dictdiffer is a list of diffs, where each diff comprises three parts. The first part specifies the modification type:
+The output of dictdiffer is a list of diffs, where each diff has three parts. The first part specifies the modification type:
 
@@
 The third part is the content of the change---terminals and nonterminals.
-Ruxanne looks for changes within each root-level scope (Rust book calls these root-level scopes the \textit{items} of a crate\footnote{\url{https://doc.rust-lang.org/reference/items.html}}).
+Ruxanne looks for changes within each root-level scope (Rust book\footnote{\url{https://doc.rust-lang.org/reference/items.html}} calls these root-level scopes the \textit{items} of a crate).
 That is, our assumption is that each change pattern occurs within one root-level scope.
@@
 touching $\mathtt{NT} \cup \mathtt{BCE}$).
-% \todo{if you mathtt BCE here, you should do it everywhere, i.e. in the paragraph at the start of 2.1.1.}
 
@@
 The change replaces \verb+tail+ with \verb+&tail+ and \verb+head+ with \verb+head.clone()+.
-That is, after the change, the second argument passes a borrow of \verb+tail+ rather than sending the ownership of \verb+tail+ to the callee; and, in the third argument, instead of sending \verb+head+ to the callee, it sends a clone of it.
+That is, after the change, the second argument passes a borrow of \verb+tail+ rather than sending the ownership of \verb+tail+ to the callee; and, the third argument sends a clone of head to the callee instead of sending head itself.
 A DFS on this tree yields three different paths, which we show in red, green, and purple.
 A path represents a sequence of involved non terminals.
-We found it crucial to record the order of nodes in the path, as this allows us to record that order in our datapoint embedding.
+We found it crucial to record the order of nodes in the path, so that we could record that order in our datapoint embedding.
 
 \paragraph{Code Embedding.} A key question is how to transform these paths to a fixed size datapoint.
+We need fixed-size datapoints, as they are a requirement for well-known centroid-based and density-based clustering algorithms~\citep{xu2005survey}.
 First, we need to define a fixed set of columns.
-We need fixed-size datapoints, as they are a requirement for well-known centroid-based and density-based clustering algorithms~\citep{xu2005survey}.
 The number of columns controls the amount of information we can embed in the datapoints.
@@
 Second, we semi-automatically design a weighting scheme to prioritize non terminals that we think are more salient for Rust bugs.
-In Section~\ref{sec:weighting_scheme}, we provide and example on how this approach works.
+Section~\ref{sec:weighting_scheme}, immediately below, illustrates this approach on an example..
 
@@
 
-    \caption{\label{fig:essence}Multiplying the number of occurrences of items observed in the tree by their respective weights, resulting in the essence of change (visualized with a circle pack figure at bottom right).
+    \caption{\label{fig:essence}Multiplying the number of occurrences of items observed in the tree by their respective weights results in the essence of change, visualized with a circle pack figure at bottom right.
 Vectors show (above) the number of occurrences and (below) weighted occurrence counts.}
@@
 The last description is the most useful one, and we designed our weighting scheme to embed this value judgment.
-
 We made two observations about changes.
-First, we saw that an item (non terminal or BC-related element) that occurs closer to AST leaves tends to be more semantically important.
-Relatedly, items closer to the root tend to be repeated in all the paths within one ASTDiff (e.g.
+First, we saw that items (non terminals or BC-related elements) that occurs closer to AST leaves tend to be more semantically important.
+Second, and relatedly, items closer to the root tend to be repeated in all the paths within one ASTDiff (e.g.
 ItemFn, Local, ExprCall in Figure~\ref{fig:extraction}).
@@
 To design our weighting scheme, we collected empirical data about locations of item occurrences in a corpus.
-Specifically, we mined the last 20 bug related commits of the last 20 most-starred Rust projects on GitHub, and ran them through our pipeline.
+Specifically, we mined 20 of the most recent bug related commits of the top 20 most-starred Rust projects on GitHub, and ran them through our pipeline.
 We recorded the total number of occurrences of each item.
@@
 \begin{table}
-\caption{\label{table:repos} Target repositories for mining}
+\caption{\label{table:repos} We chose the 18 most-starred Rust projects on GitHub as target repositories for mining.
+}
 \begin{tabular}{l r r r r}
@@
 Secondly, it is a density-based clustering method, which means that it detects arbitrarily shaped clusters; centroid-based methods like k-means can't detect such clusters.
-We did try k-means and its more powerful fuzzy variant Fuzzy C-means~\citep{dunn1973fuzzy} and did not get useful clusters.
-\todo[inline]{Wouldn't the reviewers complain why you haven't put k-means data in the paper?}
 
@@
 
-% TODO fix this caption
 \begin{figure}[h]
@@
 Since both large and small values of $\epsilon$ would steer us away from obtaining meaningful clusters, we felt the need to carry out a manual analysis of clusters for better parameter tuning.
-In the manual analysis, we randomly pick 50 clusters; from each cluster we randomly choose 10 datapoints.
-The person analyzing the cluster (The first author) then looks at the code and specifies the clusters that contained more than five datapoints showing similar patterns.
+In the manual analysis, we randomly picked 50 clusters; from each cluster we randomly chose 10 datapoints.
+The person analyzing the cluster (the first author) then looked at the code and specified the clusters that contained more than five datapoints showing similar patterns.
 After conducting the manual analysis, we decided to use $\epsilon=0.0001, Z=5$ and $\epsilon=0.001, Z=5$, as our clustering parameters for $D_g$ and $D_b$, resulting in 577 and 102 clusters, respectively.
-As a final step, we exclude clusters that contained datapoints from fewer than 3 projects.
+As a final step, we excluded clusters that contained datapoints from fewer than 3 projects.
 That is because we wanted our clusters to contain cross-project (e.g.
@@
 
-In this work, we were only interested in the first two groups and ignored the clusters for refactoring changes.
+In this work, we were only interested in the first two groups and ignored clusters for refactoring changes.
 Also, we prefer reporting clusters that manifest more-specific patterns.
@@
 In our manual analysis, we randomly picked 50 datapoints of each cluster.
-If the cluster had fewer than 50 datapoints, we would analyze all of the datapoints in the cluster.
+If the cluster had fewer than 50 datapoints, we analyzed all of the datapoints in the cluster.
 Regarding general patterns, we obtained 577 clusters, out of which 428 included cross-project patterns.
-Out of these 428 clusters, 35 clusters included more than 50 points.
+Of these 428 clusters, 35 clusters included more than 50 points.
 That is, we did not do random sampling for 393 clusters, we just manually analyzed all the datapoints in them.
-Regarding BC-related patterns, out of 102 obtained clusters, 49 clusters included cross-project pattern.
+Regarding BC-related patterns, out of 102 obtained clusters, 49 clusters included cross-project patterns.
 Only 4 of these 49 clusters included more than 50 datapoints, therefore, we only did a random sampling for 4 clusters.

### 2022-11-23 results.tex
added changes

@@
 In this section, we present the clusters of fix patterns that we found in our data set.
-\RQG{} is about the borrow-checker related patterns, while \RQBC{} is about general patterns.
-However, our technique fundamentally relies on our data modelling approach successfully capturing the most important aspects of program changes.
-We thus first evaluate the data modelling approach as our RQ1, before continuing with \RQBC{} and \RQG{}.
+RQ2 is about general patterns, while RQ3 is about borrow-checker related patterns.
+However, our technique fundamentally relies on our code embedding approach successfully capturing the most important aspects of program changes.
+We thus first evaluate the code embedding approach as our RQ1, before continuing with RQ2 and RQ3.
 
-\subsection{Evaluating the data modelling approach (RQ1)}
+\subsection{Evaluating the code embedding approach (RQ1)}
 
-\textbf{Does our data modelling approach capture the most important aspect of program changes?}
+\textbf{Does our code embedding approach capture the most important aspect of program changes?}
 
 As discussed in the previous section, we used an embedding mechanism to incorporate AST information in fixed-sized datapoints which then could be used in the DBSCAN clustering algorithm.
-We used heuristics (a weighting scheme) while creating our datapoints to make the most important program elements stand out from all the non terminals seen in the AST (along with some BC-related elements).
-Here we evaluate the effectiveness of our code embedding.
+We used heuristics (a weighting scheme) while creating our datapoints to make the program elements representing the most important non terminals stand out from the complete set of non terminals seen in the AST.
+In addition to that, we also incorporated the BC-related elements in our datapoints.
+Here we evaluate the effectiveness of our embedding.
 
-The general idea for our evaluation is determining whether the most important program elements get high values in our embedding, i.e.
+Our evaluation aims to determine whether the most important program elements get high values in our embedding, i.e.
 are recognized as important.
-The analyzer compares the actual change in the code and compares it to the visual representation of the respective data point.
+A human analyzer compares the actual change in the code and compares it to the visual representation of the respective data point.
 We used circle pack figures to visualize each data point~\citep{collins2003circle}, generated with the packcircles python library\footnote{\url{https://github.com/mhtchan/packcircles}}.
-In these figures, the radius of the circle of each non terminal shows its importance in our embedding (as seen in Figure~\ref{fig:essence}).
+In these figures, the radius of the circle for each non terminal shows its importance in our embedding (Figure~\ref{fig:essence} presented an example of the circle pack figures).
 
@@
 
+The human looks at the program output and assigns a rank to it
+
 We carried out a manual experiment for evaluating our code embedding.
-In the experiment, the analyzer should rank the most important element based on the provided circlepack figure and the change's source code.
+In the experiment, the human analyzer looks at the provided circlepack figure and the change's source code and assigns a rank to it.
+This rank reflects the rank of the most important element in the circle pack figure according to the analyzer's view.
 Based on these ranks, then we calculate Top-$n$ Accuracy for $n=1$ up to $n=5$.
-Both authors carried out the experiment independently.
-We report the results as the portion of the changes which both authors agreed upon its embedding correctness.
+Both authors acted as analyzer and carried out the experiment independently.
+We report the results as the fraction of the changes on which both authors agreed upon its embedding correctness.
 The final results are shown in Figure~\ref{fig:rq1}.
@@
 
-Although, on average, reporting most important element within the top four elements in 90\% of the cases is an accpectable outcome, we examined our pipeline for the other 10\%.
-An analysis shows that our pipeline fails in change patterns where the change involves many elements (e.g.
+Although we found that our results were acceptable---the result reported by our embedding was top-4 as ranked by humans 90\% of the time---we investigated the other 10\% as well.
+We found that our embedding tended to fail in change patterns where the change involves many elements (e.g.
 adding a whole new \verb+impl+ block and implementing new functions in it).
-It is worth mentioning that these changes tend to be less cross-project as having many changed elements make it easier for datapoints to be dissimlar.
+It is worth mentioning that these change patterns tend to apply across projects less often, as having many changed elements make it easier for datapoints to be dissimlar.
 
@@
 After manual analysis, as described in~\ref{sec:manual_analysis_cluster_selection}, we ended up with 12 general patterns, which we describe in Table~\ref{table:general}.
-As it is shown in the table, we have separated the patterns based on the underlying program element (Category column).
+As shown in the table, we have separated the patterns based on the underlying program element (Category column).
 We have given each pattern a specific ID (ID column) and determined the number of instances manifesting that fix pattern (\# Instances column), as well as the number of projects in which the pattern was seen (\# Project column).
-Now we provide detailed description for each of these patterns:
+Now we provide a detailed description for each of these patterns:
 
@@
 \noindent\textbf{Description:} Adding \texttt{pub} to a struct field makes it possible to access that field from external modules.
-In the instances of this cluster, adding \texttt{pub} access modifier happened in bug fixing changes, where the developer needed to access a field which had not been marked for public access.
+In the instances of this cluster, adding the \texttt{pub} access modifier happened in bug fixing changes, where the developer needed to access a field which had not been marked for public access.
 On the other hand, the public access might be revoked if the developer realizes there are no external accesses to the field, and that such accesses are not desirable, e.g.
@@
 After manual analysis, as described in~\ref{sec:manual_analysis_cluster_selection}, we ended up with 8 borrow-checker related fix patterns, which we, similar to the previous table, describe in Table~\ref{table:bc}.
-Here we provide detailed description for each of these patterns:
+Here we provide a detailed description for each of these patterns:
 

### 2022-11-24 conclusion.tex
added changes

@@
 In this project, we developed Ruxanne, a tool for mining bug fix patterns in Rust.
-Using this tool, we mined the top 18 most starred Rust projects in Github to discover common bug fix patterns lied within their code changes.
-At the heart of Ruxanne, we used a novel code embedding approach to embed most important aspects of a code change in a fixed sized datapoint.
+Using this tool, we mined the top 18 most starred Rust projects in Github to discover common bug fix patterns within their code changes.
+At the heart of Ruxanne, we used a novel code embedding approach to embed the most important aspects of a code change in a fixed sized datapoint.
 We stored 87,726 datapoints drawn from 57,214 commits across these 18 projects.

### 2022-11-24 discussion.tex
added changes

@@
-In this work, through Ruxanne we successfully mined 20 cross project bug fix patterns.
+In this work, using Ruxanne we successfully mined 20 cross project bug fix patterns.
 We introduced these patterns in two groups: 12 general patterns and 8 BC-related patterns.
@@
 To the best of our knowledge, the rest of the patterns cannot be reported by the current linting tools.
-That is why, we believe our work introduces a new direction for IDE tools to help Rust developers.
+We therefore believe that our work introduces a new direction for IDE tools to help Rust developers.
 
@@
 The patterns encompass a wide variety of program elements both in the general and BC-related patterns.
-This shows that our weighting scheme is acceptably unbiased.
-Also, the introduced patterns are all cross-project fix patterns.
-Each fix pattern has been seen at least in three different projects, indicating its commonality.
+This shows that our weighting scheme is not biased towards specific program elements.
+Also, our patterns are all cross-project fix patterns: each fix pattern has been seen at least in three different projects.
 
-Furthermore, we expected to obtain certain patterns regarding some certain Rust related features, such as \verb+move+, but our results did not include those patterns.
+We expected to obtain certain patterns regarding some certain Rust related features, such as \verb+move+, but our results did not include those patterns.
 Our speculation is that, since Rust has a strong compiler check before creating the program executable, the programmers usually get many errors, and they only push their code once it is free of compile errors.
@@
 In our work there are three main threats to internal validity: (1) We use a weighting scheme which is based on a heuristic to give more values to tree elements closer to the leaves.
-In addition to that, we also manually adjusted the weights of borrow checker related elements to respect ownership framework's role in Rust, though we recorded their results separate from general results.
-As this approach is based on an intuition, it does not guarantee achieving optimal clusters.
+In addition to that, we also manually adjusted the weights of borrow checker related elements to respect the ownership framework's role in Rust, though we recorded their results separate from general results.
+As this approach is based on intuition, it does not guarantee achieving optimal clusters.
 (2) The collected commits are the ones that fix a bug in a previously pushed commit.
@@
 
-The threat to external validity can be the developers reporting the commits as bug fixing commits while they are not really fixing a functionality.
+A threat to external validity can be the developers reporting the commits as bug fixing commits while they are not really fixing a functionality.
 Similarly, a commit message might not contain our target keywords while the commit is associated with bug fixing changes.
+\todo[inline]{Are internal and external used properly?}

### 2022-11-24 methodology.tex
added changes

@@
 \texttt{clone}, \texttt{Rc}, \texttt{Box}.
-The full list of \texttt{BCE} elements can be found in Appendix~\ref{sec:appendix}.
+The full list of \texttt{BCE} elements can be found in our replication package.
 
@@
 We propose a novel representation.
-Similar to the naive approach, we order-insensitively define the columns of our dataset as the set of items $\mathtt{NT} \cup \mathtt{BCE}$ we collected from Syn.
+Similar to the naive approach, we order-insensitively define the columns of our dataset as the set of items $\mathtt{NT} \cup \mathtt{BCE}$ we collected from Syn (the columns are constant across all Rust programs).
 However, to simulate order-sensitiveness, we add two more steps.
@@
 For instance, the purple (right-most) path in Figure~\ref{fig:extraction} could be described as a change in a local variable declaration; a change in a function call; calling a method of one of the function arguments; or calling clone() on one of the function arguments.
-The last description is the most useful one, and we designed our weighting scheme to embed this value judgment.
+The last description, in our opinion, is the most useful one, and we designed our weighting scheme to embed this value judgment.
 We made two observations about changes.
@@
 The manual adjustment is why we characterize our weighting scheme as semi-automatically designed.
+The final weights can also be found in our replication package.
 
@@
 For BC-related code changes, we also store the detected BC-related keywords.
+In summary, a datapoint is a tuple containing essence values for each element in $\mathtt{NT} \cup \mathtt{BCE}$ plus metadata about the change (e.g.
+commit hash, filename, etc.).
 Now our databases are ready for clustering and categorization.
@@
 An abstract pattern as opposed to a more-specific pattern can account for many changes in the programs.
+We use our judgment to prioritize patterns that we deem more specific.
 For instance, we deem the pattern `Changing a clone of a variable to a borrowing of it in a function argument' more specific than `Changing a statement in a function's body':

### 2022-11-24 related.tex
added changes

@@
 In our case there were general and language-specific patterns; for Rust, the language-specific patterns were related to the borrow checker (BC patterns).
-However, in oppose to our paper, in~\cite{yang2022mining}, they did a much more elaborate manual analysis, as in their data collection, they only collected general information about the single hunk changes (e.g.
+However, as opposed to our paper, \cite{yang2022mining} needed to do a much more elaborate manual analysis, as they had only collected general information about the single hunk changes (e.g.
 the number of variables, or arguments).

### 2022-11-24 results.tex
added changes

@@
 We collected random samples 1000 times.
-In each iteration, we collected 50 sample datapoints from our general and borrow-checker databases (25 each).
+In each iteration, 50 datapoints are sampled from our general and borrow-checker databases (25 each).
 For each database, out of the 1000 sample sets, we picked the one that minimized the sampling error.
-The sampling error is defined as the absolute difference between the sample mean and the total mean.
+The sampling error is defined as the average of the absolute differences between the sample set mean and the total mean per column~\citep{degroot2012probability}.
 Also, these random samples are drawn from the clustered datapoints.
@@
 
-The human looks at the program output and assigns a rank to it
-
 We carried out a manual experiment for evaluating our code embedding.
-In the experiment, the human analyzer looks at the provided circlepack figure and the change's source code and assigns a rank to it.
-This rank reflects the rank of the most important element in the circle pack figure according to the analyzer's view.
-Based on these ranks, then we calculate Top-$n$ Accuracy for $n=1$ up to $n=5$.
+In the experiment, the human analyzer looks at the change's source code and finds the most important element that appears in the circle pack figure, recording its rank.
+A value of 1 indicates that the human analyzer concurs with the algorithm, while 2 means that the human-identified most important element was ranked 2 by the algorithm.
+After collecting all the ranks, then we calculate Top-$n$ Accuracy for $n=1$ up to $n=5$.
 Both authors acted as analyzer and carried out the experiment independently.
@@
 \noindent\textbf{Description:} In this pattern, the attribute set of a struct field changes.
-As with \textit{\textbf{G.attr.struct}}, this modification also happens in three different ways.
+As with \textit{\textbf{G.attr.struct}}, there are also three subtypes of this change pattern.
 Using these attributes, the developers can control different actions they wish to apply on struct fields, in addition to providing meta-information.
+For instance, in the code snippet provided above, Serde (A serialization/deserialization tool for Rust) calls a different function (\verb+is_empty+) to determine whether to skip serializing the \verb+attributes+ field.
 
@@
 
-Similar to other languages, such as C, a struct makes it possible for the users to define a custom type that is comprised of different types.
+Like in other languages, such as C, a struct makes it possible for users to define a custom type that is comprised of different types.
 The patterns that we describe in this section can also be seen in other languages that use the same concept.
@@
 \noindent\textbf{Description:} Rust is a null-safe language, meaning that object references cannot take on null values.
-If a developer wants a variable to contain null, they would have to use the \texttt{Option} type.
+If a developer wants a variable to contain the equivalent of null, they would have to use the \texttt{Option} type.
 \texttt{Option} is essentially an \texttt{enum} with two variants, \texttt{Some}, which indicates that a variable has a value, and \texttt{None}, which indicates the absence of any value.
@@
 For instance, in the Rust project swc-project/swc (a fast TypeScript/JavaScript compiler), the developer needed to modify the CSS parser source code to account for empty @media queries\footnote{\scriptsize \url{https://github.com/swc-project/swc/commit/75a14f98b7370226115ee24eec6eb8c802bd4837}}.
-As shown in the snippet above, they changed type \texttt{MediaQueryList} to \verb+Option<MediaQueryList>+, and modified the other parts of the source code, accordingly.
+As shown in the snippet above, they changed type \texttt{MediaQueryList} to \verb+Option<MediaQueryList>+, and modified the other parts of the source code accordingly.
 
@@
 
-Rust is a statically-typed language, and similar to the other statically-typed languages, some of the change patterns are associated with changes in types used in different program structures.
+Rust is a statically-typed language, and like in the other statically-typed languages, some change patterns are associated with changes in types used in different program structures.
 Here, we introduce four bug fix patterns that are associated with types in Rust:
@@
 For instance, the developer may be implementing a new feature or fixing the behaviour of the program.
-Also, it can happen for refactoring purposes or performance enhancement.
+Also, it can happen for refactoring purposes or for performance enhancement.
 
-For instance, in the Rust project starship/starship (a cross-shell prompt), in order to preserve the insertion order of \verb+substitutions+ field, the developer changes its type from \verb+HashMap+ to \verb+IndexMap+\footnote{\url{https://github.com/starship/starship/commit/4de9e43cff46c834bd340d24a02fc95d85310a33}}.
+For instance, in the Rust project starship/starship (a cross-shell prompt), to preserve the insertion order of the \verb+substitutions+ field, the developer changes its type from \verb+HashMap+ to \verb+IndexMap+\footnote{\url{https://github.com/starship/starship/commit/4de9e43cff46c834bd340d24a02fc95d85310a33}}.
 \\
@@
 We used $\epsilon=0.001$ and $Z=5$ as our DBSCAN parameters and that resulted in 102 clusters.
-After manual analysis, as described in~\ref{sec:manual_analysis_cluster_selection}, we ended up with 8 borrow-checker related fix patterns, which we, similar to the previous table, describe in Table~\ref{table:bc}.
-Here we provide a detailed description for each of these patterns:
+After manual analysis, as described in~\ref{sec:manual_analysis_cluster_selection}, we ended up with 8 borrow-checker related fix patterns.
+Similar to the previous table, we describe BC-related patterns in Table~\ref{table:bc}.
+Now, we provide a detailed description of each of these patterns:
 
@@
 
-\noindent\textbf{Description:} This is a change that in all instances has been detected by Clippy (commit messages indicated the use of Clippy for this fix).
+\noindent\textbf{Description:} This is a change that in all instances (in our benchmarks) has been proposed by Clippy---commit messages indicated the use of Clippy for this fix.
 The change applies on an object prior to calling \verb+to_string()+ on that object.
@@
 This change can be applied using Clippy.
+A static lifetime of a variable indicates that the variable lives for the entire lifetime of the running program, and is not bound to a specific scope.
 If the presence of static is not required, it is better that to omit it, as keeping it might create complicated types in the program.

### 2022-11-24 discussion.tex
added changes

@@
 We introduced these patterns in two groups: 12 general patterns and 8 BC-related patterns.
-Clippy is able to detect four of the BC-related patterns.
+Clippy is only able to detect four of the BC-related patterns.
+Fundamentally, Clippy cannot detect some of them.
+For instance, we proposed the pattern \textit{Adding mutability} which Clippy cannot detect, simply because it cannot predict the next change that the developer is going to make.
 To the best of our knowledge, the rest of the patterns cannot be reported by the current linting tools.

### 2022-11-24 results.tex
added changes

@@
 Using these attributes, the developers can control different actions they wish to apply on struct fields, in addition to providing meta-information.
-For instance, in the code snippet provided above, Serde (A serialization/deserialization tool for Rust) calls a different function (\verb+is_empty+) to determine whether to skip serializing the \verb+attributes+ field.
+For instance, in the code snippet provided above, the developers chose to make Serde (a serialization/deserialization tool for Rust) call a different function (\verb+is_empty+ instead of \verb+Option::is_none+) to determine whether to skip serializing the \verb+attributes+ field.
 

### 2022-11-25 results.tex
little change

@@
 We collected random samples 1000 times.
-In each iteration, 50 datapoints are sampled from our general and borrow-checker databases (25 each).
+In each iteration, we sampled 50 datapoints from our general and borrow-checker databases (25 each).
 For each database, out of the 1000 sample sets, we picked the one that minimized the sampling error.
@@
 Both authors acted as analyzer and carried out the experiment independently.
-We report the results as the fraction of the changes on which both authors agreed upon its embedding correctness.
+To avoid inflating our final results, after comparing the ranks that the authors gave to each datapoint, we consider the larger (worse) rank in our final results.
+That is, if the first author gave rank 2 to datapoint $D$ and the second author gave rank 3 to that datapoint, in the final results we gave datapoint $D$ a rank of 3.
 The final results are shown in Figure~\ref{fig:rq1}.

### 2022-11-25 results.tex
redundant clone

@@
 There are more than 500 lints included in Clippy.
-With regards to this pattern, Clippy detects whether a variable clone is necessary, i.e.
-could the developer have used the variable itself or not.
-It is evident that removing a redundant clone does not modify the program behaviour, however it can boost the performance significantly if the program was repeatedly cloning large variables.
+Since detecting whether a clone is redundant is undecidable, Clippy would have to conservatively estimate redundancy.
+By definition, removing a redundant clone does not modify functional aspects of the program behaviour.
+However, such a removal can greatly improve performance, if the program was formerly cloning large objects.
 \\

### 2022-11-25 results.tex
static lifetime

@@
 This change can be applied using Clippy.
-A static lifetime of a variable indicates that the variable lives for the entire lifetime of the running program, and is not bound to a specific scope.
+A static lifetime associated with a variable indicates that the variable lives for the entire lifetime of the running program, and is not bound to a specific scope.
 If the presence of static is not required, it is better that to omit it, as keeping it might create complicated types in the program.

### 2022-11-25 related.tex
added more to related

@@
 Like~\cite{hanam2016discovering}, we used DBSCAN to cluster fixes.
-However, our novel program embedding approach differs from what they did.
-In~\cite{hanam2016discovering}, they introduced Feature Properties table, which is a categorization for program elements.
-Embedding programs based on this table resulted in a order-sensitive embedding, however, it made the datapoints sparse.
-Moreover, like~\cite{yang2022mining}, we introduced patterns in two different categories.
+Moreover, similar to~\cite{yang2022mining}, we introduced patterns in two different categories.
 In our case there were general and language-specific patterns; for Rust, the language-specific patterns were related to the borrow checker (BC patterns).
-However, as opposed to our paper, \cite{yang2022mining} needed to do a much more elaborate manual analysis, as they had only collected general information about the single hunk changes (e.g.
+However, unlike our paper, \cite{yang2022mining} needed to do a much more elaborate manual analysis, as they had only collected general information about the single hunk changes (e.g.
 the number of variables, or arguments).
+
+\subsection{Code Embedding}
+
+Recent advances in deep learning has motivated researchers to build models for various types of data.
+A common step for building these machine learning models is to create a numerical representation of data, so that it can be fed into the model.
+To realize this mapping in natural languages, the researchers have proposed methods to embed word information in fixed size vectors.
+This process is known as word embedding.
+As programming languages are not that different from natural languages~\citep{hindle2016naturalness}, similar mappings have been suggested for computer programs~\citep{chen2019literature}.
+
+\cite{alon2019code2vec} presented code2vec, a nueral model for representing code snippets as fixed size vectors.
+In their approach, they extracted flattened paths from the AST, and stored all the leaf-to-leaf paths.
+The intuition behind this decision was that leaf-to-leaf paths tend to encode more semantic information than root-to-leaf paths.
+They associated a fixed size vector to each path and fed them as inputs to a neural network that learns how to aggregate all these paths to a single embedding.
+Our code embedding approach is simpler than code2vec, as we seek a less sophisticated goal than them---bug patterns clustering instead of method name prediction.
+Moreover, unlike us, code2vec does not offer a general code embedding methods, since the final embedding depends on the output layer of the neural network.
+That is, if we change the goal of method name prediction, the embeddings would be completely different.
+Also, our code embedding has actual interpretability as our columns are associated with non terminals.
+In code2vec, the dimensions of the final vector is a parameter of the system and to no extent is related to the underlying programming language.
+
+Our work tends to be more similar to that of \cite{hanam2016discovering}.
+However, our novel program embedding approach might differ from what they did on certain key decisions.
+In~\cite{hanam2016discovering}, they introduced the Feature Properties table, which is a categorization for program elements.
+Embedding programs based on this table resulted in an order-sensitive embedding, however, it made the datapoints sparse.
 

### 2022-11-25 related.tex
code embedding

@@
 
-Recent advances in deep learning has motivated researchers to build models for various types of data.
-A common step for building these machine learning models is to create a numerical representation of data, so that it can be fed into the model.
-To realize this mapping in natural languages, the researchers have proposed methods to embed word information in fixed size vectors.
-This process is known as word embedding.
+Recent advances in deep learning have motivated researchers to build models for various types of data.
+Machine learning models require a numerical representation of the data to feed into the model.
+To realize this mapping for natural language text, researchers have proposed \emph{word embedding} to embed word information in fixed size vectors.
 As programming languages are not that different from natural languages~\citep{hindle2016naturalness}, similar mappings have been suggested for computer programs~\citep{chen2019literature}.
 
-\cite{alon2019code2vec} presented code2vec, a nueral model for representing code snippets as fixed size vectors.
+\cite{alon2019code2vec} presented code2vec, a neural model for representing code snippets as fixed size vectors.
 In their approach, they extracted flattened paths from the AST, and stored all the leaf-to-leaf paths.
 The intuition behind this decision was that leaf-to-leaf paths tend to encode more semantic information than root-to-leaf paths.
-They associated a fixed size vector to each path and fed them as inputs to a neural network that learns how to aggregate all these paths to a single embedding.
-Our code embedding approach is simpler than code2vec, as we seek a less sophisticated goal than them---bug patterns clustering instead of method name prediction.
-Moreover, unlike us, code2vec does not offer a general code embedding methods, since the final embedding depends on the output layer of the neural network.
+They associated a fixed size vector with each path, and fed these as inputs to a neural network that learns how to aggregate all these paths to a single embedding.
+Our code embedding approach is simpler than code2vec's, as we have a specific (and simpler) goal---we aim to cluster bug patterns, not predict method names.
+Moreover, unlike us, code2vec does not offer a general code embedding method, since the final embedding depends on the output layer of the neural network.
 That is, if we change the goal of method name prediction, the embeddings would be completely different.
-Also, our code embedding has actual interpretability as our columns are associated with non terminals.
-In code2vec, the dimensions of the final vector is a parameter of the system and to no extent is related to the underlying programming language.
+Also, unlike many embeddings, our code embedding is interpretable: our columns are associated with non terminals.
+We have used the interpretability in our experiments.
+In code2vec, the layout of the final vector is a parameter of the system and  is not related to the underlying programming language.
 
-Our work tends to be more similar to that of \cite{hanam2016discovering}.
-However, our novel program embedding approach might differ from what they did on certain key decisions.
+Our work tends to be more similar to that of~\cite{hanam2016discovering}.
+However, our novel program embedding approach differs from what they did on certain key decisions.
 In~\cite{hanam2016discovering}, they introduced the Feature Properties table, which is a categorization for program elements.
 Embedding programs based on this table resulted in an order-sensitive embedding, however, it made the datapoints sparse.
+Our datapoints are more dense than theirs.
+\todo[inline]{quantify?}
 

### 2022-11-25 discussion.tex
move fixed

@@
 We introduced these patterns in two groups: 12 general patterns and 8 BC-related patterns.
-Clippy is only able to detect four of the BC-related patterns.
+Clippy is able to detect five of the BC-related patterns.
 Fundamentally, Clippy cannot detect some of them.
@@
 
-We expected to obtain certain patterns regarding some certain Rust related features, such as \verb+move+, but our results did not include those patterns.
+Furthermore, we expected to obtain some patterns associated with certain Rust-related features but our results did not include those patterns.
 Our speculation is that, since Rust has a strong compiler check before creating the program executable, the programmers usually get many errors, and they only push their code once it is free of compile errors.
-That is why, we think there are some patterns that were hidden from our sight.
+For instnace, we did not observe any patterns associated with \verb+move+ operator in Rust.
+This operator is used to transfer the ownership of values to a closure.
+In some cases (e.g.
+multithreading) not moving the ownership can create invalid references, something that Rust's compiler prevents.
+That is why we believe due to Rust's strong compiler check some patterns were hidden from our sight.
 

### 2022-11-25 discussion.tex
discussion extended

@@
 
-We believe our results can help researchers to create Rust tools for repair, fault localization, IDE linters, and etc.
-Also, these results can help program embedding methods to focus on more important aspects of a program and shrink down information more effectively.
+We believe our results can help researchers to create IDE tools for Rust.
+For instance, we discussed that \textit{Modifying the attributes of structs} is a common pattern that is observed in bug fix code changes.
+Assuming that the correct usage of attributes can be found other locations in the codebase~\citep{forrest2009genetic}, a code linter tool can be crafted to look for correct attribute list.
+
+Similarly, we presented the pattern \textit{Dropping clone and adding borrowing}.
+As discussed, Clippy does not detect this pattern, and repetitive clones can be computationally expensive and might lead to hogging the CPU.
+Using state-of-the-art method for program repair, a tool can be designed to recognize this pattern and change variable cloning to a simple borrowing.
+
+A general approach for code embedding is to extract paths with semantic information that serves better to our embedding goal.
+For instance, to design an embedding method for repairing Rust programs, the designer might want to disregard certain elements, such as moving ownerships to closure body---as they did not appear in the bug patterns we presented.
+
+All of our parsing, path extraction, weighting scheme and clustering modules in addition to our final results can be found in our replication package for verification, resuability and further extension.
+The weighting can be modified to make the embedding focus on a specific set of elements for building different code embeddings.
 

### 2022-11-27 discussion.tex
threats to validity fixed

@@
 In our work there are three main threats to internal validity: (1) We use a weighting scheme which is based on a heuristic to give more values to tree elements closer to the leaves.
-In addition to that, we also manually adjusted the weights of borrow checker related elements to respect the ownership framework's role in Rust, though we recorded their results separate from general results.
-As this approach is based on intuition, it does not guarantee achieving optimal clusters.
+In addition to that, we also manually adjusted the weights based on our knowledge of Rust.
+As this approach is based on our own judgement, it does not guarantee achieving optimal clusters.
+Nevertheless, our weighting scheme can be readjusted to find different patterns, which is why we made our pipeline publicly available.
 (2) The collected commits are the ones that fix a bug in a previously pushed commit.
@@
 
-A threat to external validity can be the developers reporting the commits as bug fixing commits while they are not really fixing a functionality.
+Threats to external validity can be described as: (1) The developers reporting the commits as bug fixing commits while they are not really fixing a functionality.
 Similarly, a commit message might not contain our target keywords while the commit is associated with bug fixing changes.
-\todo[inline]{Are internal and external used properly?}
+(2) The selected projects might cause selection bias and therefore we might not have presented patterns that may exist in other projects.

### 2022-11-27 methodology.tex
double period elimination
fixes added

@@
 Second, we semi-automatically design a weighting scheme to prioritize non terminals that we think are more salient for Rust bugs.
-Section~\ref{sec:weighting_scheme}, immediately below, illustrates this approach on an example..
+Section~\ref{sec:weighting_scheme}, immediately below, illustrates this approach on an example.
 
@@
 To design our weighting scheme, we collected empirical data about locations of item occurrences in a corpus.
-Specifically, we mined 20 of the most recent bug related commits of the top 20 most-starred Rust projects on GitHub, and ran them through our pipeline.
+Specifically, we mined 20 of the most recent bug related commits of the same projects described in Table~\ref{table:repos}, and ran them through our pipeline.
 We recorded the total number of occurrences of each item.

### 2022-11-27 results.tex
fixes added

@@
 That is, if the first author gave rank 2 to datapoint $D$ and the second author gave rank 3 to that datapoint, in the final results we gave datapoint $D$ a rank of 3.
+The authors' answers were the same for 45 of the 50 datapoints.
 The final results are shown in Figure~\ref{fig:rq1}.
@@
     coordinates { 
-    (1,64)
-    (2,72)
-    (3,84)
+    (1,72)
+    (2,76)
+    (3,88)
     (4,88)
@@
     (2,84)
-    (3,88)
-    (4,92)
+    (3,84)
+    (4,88)
     (5,96)
@@
     coordinates { 
-    (1,66)
-    (2,78)
+    (1,70)
+    (2,80)
     (3,86)
-    (4,90)
+    (4,88)
     (5,94)
@@
 
-Although we found that our results were acceptable---the result reported by our embedding was top-4 as ranked by humans 90\% of the time---we investigated the other 10\% as well.
+Although we found that our results were acceptable---the result reported by our embedding was top-4 as ranked by humans 88\% of the time---we investigated the other 12\% as well.
 We found that our embedding tended to fail in change patterns where the change involves many elements (e.g.

### 2022-11-27 discussion.tex
ttv fixed

@@
 
-In our work there are three main threats to internal validity: (1) We use a weighting scheme which is based on a heuristic to give more values to tree elements closer to the leaves.
-In addition to that, we also manually adjusted the weights based on our knowledge of Rust.
-As this approach is based on our own judgement, it does not guarantee achieving optimal clusters.
+In our work there are two main threats to internal validity: (1) A threat to internal validity is confounding, where changes to what shows up in the embedding are not due to changes in the code being embedded.
+In this case, the weighting scheme may contribute to confounding because we manually adjusted the weights.
 Nevertheless, our weighting scheme can be readjusted to find different patterns, which is why we made our pipeline publicly available.
-(2) The collected commits are the ones that fix a bug in a previously pushed commit.
-We do not consider the bugs that occur during a commit and also fixed during the same commit.
-(3) Our code embedding approach is based on the frequency of observed program elements in ASTs, and we use DBSCAN as our clustering algorithm.
-Other program code embedding methods and clustering algorithms (e.g.
+(2) Our code embedding approach is based on the frequency of observed program elements in ASTs, and we use DBSCAN as our clustering algorithm.
+Other code embedding methods and clustering algorithms (e.g.
 SLINK) might output new clusters that our pipeline is unable to find.
-
-Threats to external validity can be described as: (1) The developers reporting the commits as bug fixing commits while they are not really fixing a functionality.
+A threat to external validity is selection bias, where the benchmarks differ systematically from the actual world.
+The selected projects might cause selection bias and therefore we might not have presented patterns that may exist in other projects.
+Another threat to external validity can be the developers reporting the commits as bug fixing commits while they are not really fixing a functionality.
 Similarly, a commit message might not contain our target keywords while the commit is associated with bug fixing changes.
-(2) The selected projects might cause selection bias and therefore we might not have presented patterns that may exist in other projects.

### 2022-11-27 discussion.tex
discussion

@@
-In this work, using Ruxanne we successfully mined 20 cross project bug fix patterns.
-We introduced these patterns in two groups: 12 general patterns and 8 BC-related patterns.
+In this work, we used Ruxanne to successfully mine 20 cross-project bug fix patterns.
+We presented these patterns in two groups: 12 general patterns and 8 BC-related patterns.
 Clippy is able to detect five of the BC-related patterns.
-Fundamentally, Clippy cannot detect some of them.
-For instance, we proposed the pattern \textit{Adding mutability} which Clippy cannot detect, simply because it cannot predict the next change that the developer is going to make.
+Fundamentally, Clippy cannot detect some of the remaining patterns.
+For instance, we proposed the pattern \textit{Adding mutability}.
+In the Rust context, this change anticipates a future change to the code which will perform a mutation of a previously immutable value.
+Predicting the future is beyond Clippy's remit.
 To the best of our knowledge, the rest of the patterns cannot be reported by the current linting tools.
-We therefore believe that our work introduces a new direction for IDE tools to help Rust developers.
 
-In each group, we categorized the bug fix patterns based on the underlying program element.
-The patterns encompass a wide variety of program elements both in the general and BC-related patterns.
+In each group, we categorized the bug fix patterns based on their underlying program element.
+The patterns encompass a wide variety of program elements both in the general and BC-related groups.
 This shows that our weighting scheme is not biased towards specific program elements.
@@
 
-Furthermore, we expected to obtain some patterns associated with certain Rust-related features but our results did not include those patterns.
-Our speculation is that, since Rust has a strong compiler check before creating the program executable, the programmers usually get many errors, and they only push their code once it is free of compile errors.
-For instnace, we did not observe any patterns associated with \verb+move+ operator in Rust.
-This operator is used to transfer the ownership of values to a closure.
+Furthermore, we expected to obtain some patterns associated with certain Rust-related features.
+However, our results did not include those patterns.
+For instance, we did not observe any patterns associated with the Rust \verb+move+ operator, which has been known to trouble many beginning Rust programmers.
+\verb+move+ is used to transfer the ownership of values to a closure.
 In some cases (e.g.
-multithreading) not moving the ownership can create invalid references, something that Rust's compiler prevents.
-That is why we believe due to Rust's strong compiler check some patterns were hidden from our sight.
+multithreading), not moving the ownership could create invalid references.
+Rust compiler checks prevent such errors.
+Our speculation is that, since Rust enforces strict compiler checks, programmers usually get many errors, and they only push their code once it is free of compile errors.
+We thus would not observe patterns which don't make it to the repository.
 
@@
 For instance, we discussed that \textit{Modifying the attributes of structs} is a common pattern that is observed in bug fix code changes.
-Assuming that the correct usage of attributes can be found other locations in the codebase~\citep{forrest2009genetic}, a code linter tool can be crafted to look for correct attribute list.
+Assuming that the correct usage of attributes can be found other locations in the codebase~\citep{forrest2009genetic}, a code linter tool can be crafted to look for correct attribute lists.
 
@@
 As discussed, Clippy does not detect this pattern, and repetitive clones can be computationally expensive and might lead to hogging the CPU.
-Using state-of-the-art method for program repair, a tool can be designed to recognize this pattern and change variable cloning to a simple borrowing.
+Using state-of-the-art methods for program repair, one could design a tool to recognize this pattern and change variable cloning to simple borrowing.
 
-A general approach for code embedding is to extract paths with semantic information that serves better to our embedding goal.
-For instance, to design an embedding method for repairing Rust programs, the designer might want to disregard certain elements, such as moving ownerships to closure body---as they did not appear in the bug patterns we presented.
+A general approach for code embedding is to extract paths with semantic information that better serve a given embedding goal.
+For instance, to design an embedding method for repairing Rust programs, the designer might want to disregard certain elements, such as moving ownership to closure bodies.
+Such elements may be irrelevant for repair because they did not appear in the bug patterns that we presented.
 
-All of our parsing, path extraction, weighting scheme and clustering modules in addition to our final results can be found in our replication package for verification, resuability and further extension.
-The weighting can be modified to make the embedding focus on a specific set of elements for building different code embeddings.
+All of our parsing, path extraction, weighting scheme, and clustering modules, along with our final results can be found in our replication package, which we provide for verification, reusability and further extension.
+Weightings can be modified to make the embedding focus on a specific set of elements for building different code embeddings.
 

### 2022-11-28 discussion.tex
linters became more explicit
minor edit

@@
 Predicting the future is beyond Clippy's remit.
-To the best of our knowledge, the rest of the patterns cannot be reported by the current linting tools.
+To the best of our knowledge, the rest of the patterns (3 BC-related patterns and 12 general patterns) cannot be reported by the current linting tools.
 
@@
 The patterns encompass a wide variety of program elements both in the general and BC-related groups.
-This shows that our weighting scheme is not biased towards specific program elements.
-Also, our patterns are all cross-project fix patterns: each fix pattern has been seen at least in three different projects.
+This supports a conclusion that our weighting scheme is not biased towards specific program elements.
+Also, our patterns are all cross-project fix patterns: each reported fix pattern has been seen at least in three different projects.
 
@@
 
-All of our parsing, path extraction, weighting scheme, and clustering modules, along with our final results can be found in our replication package, which we provide for verification, reusability and further extension.
+All of our parsing, path extraction, weighting scheme, and clustering modules, along with our final results, can be found in our replication package, which we provide for verification, reusability and further extension.
 Weightings can be modified to make the embedding focus on a specific set of elements for building different code embeddings.

### 2022-11-28 results.tex
minor edit

@@
 As discussed in the previous section, we used an embedding mechanism to incorporate AST information in fixed-sized datapoints which then could be used in the DBSCAN clustering algorithm.
-We used heuristics (a weighting scheme) while creating our datapoints to make the program elements representing the most important non terminals stand out from the complete set of non terminals seen in the AST.
+We used heuristics (in the form of a weighting scheme) while creating our datapoints to make the program elements representing the most important non terminals stand out from the complete set of non terminals seen in the AST.
 In addition to that, we also incorporated the BC-related elements in our datapoints.
@@
 are recognized as important.
-A human analyzer compares the actual change in the code and compares it to the visual representation of the respective data point.
+A human analyzer (paper author) compares the actual change in the code and compares it to the visual representation of the respective data point.
 We used circle pack figures to visualize each data point~\citep{collins2003circle}, generated with the packcircles python library\footnote{\url{https://github.com/mhtchan/packcircles}}.
@@
 In each iteration, we sampled 50 datapoints from our general and borrow-checker databases (25 each).
-For each database, out of the 1000 sample sets, we picked the one that minimized the sampling error.
+For each database, out of the 1000 sample sets, we picked the one that minimized the sampling error (i.e.
+we are aiming to choose the sample that looks most like the entire dataset).
 The sampling error is defined as the average of the absolute differences between the sample set mean and the total mean per column~\citep{degroot2012probability}.
 Also, these random samples are drawn from the clustered datapoints.
-The reason behind this decision is the same as why we collected a lot of noise points, that is, the majority of the commits involve changes in many program elements and the key changing elements are difficult to determine.
+The reason behind this decision is the same as why we collected a lot of noise points: the majority of the commits involve changes in many program elements and the key changed elements are difficult to determine.
 
@@
 The authors' answers were the same for 45 of the 50 datapoints.
-The final results are shown in Figure~\ref{fig:rq1}.
 
@@
 \end{tikzpicture}   
-\caption{\label{fig:rq1} Top-$n$ Accuracy of the proposed code embedding approach.
+\caption{\label{fig:rq1} Top-$n$ Accuracy of our proposed code embedding approach.
 The plot shows Top-$n$ Accuracy where $n\in\{1,2,3,4,5\}$ for both general and borrow-checker related datapoints.
-As shown in the plot, in more than 85\% of the datapoints, the most important element is seen at least in the first three ranks.}
+As shown in the plot, in more than 85\% of the datapoints, the most important element is seen at least in the first three ranks, rising to about 95\% for the first five ranks.}
 
@@
 
+Figure~\ref{fig:rq1} shows the final results.
 Although we found that our results were acceptable---the result reported by our embedding was top-4 as ranked by humans 88\% of the time---we investigated the other 12\% as well.

### 2022-11-28 methodology.tex
capitalizations etc

@@
 The third part is the content of the change---terminals and nonterminals.
-Ruxanne looks for changes within each root-level scope (Rust book\footnote{\url{https://doc.rust-lang.org/reference/items.html}} calls these root-level scopes the \textit{items} of a crate).
+Ruxanne looks for changes within each root-level scope (the Rust book\footnote{\url{https://doc.rust-lang.org/reference/items.html}} calls these root-level scopes the \textit{items} of a crate).
 That is, our assumption is that each change pattern occurs within one root-level scope.

### 2022-11-30 cover.tex
cover letter added

\documentclass{letter}
%use letter format
\setlength{\paperheight}{11in}
\setlength{\paperwidth}{8.5in}
\setlength{\pdfpageheight}{11in}
\setlength{\pdfpagewidth}{8.5in}

\usepackage{cite}
\usepackage{letterbib}

\setlength{\topmargin}{-.5in}
\setlength{\headheight}{0in}
\setlength{\textheight}{8.5in}

\begin{document}
\begin{letter}{}
\opening{Dear Editor,}

We are submitting the attached paper, "A Study of Common Bug Fix Patterns in Rust"
for publication in EMSE. This paper is an empirical study of the common bug fix patterns that we extracted from the most popular open-source Rust projects. Our primary contributions are as follows:

\begin{itemize}
    \item We introduce Ruxanne, our pipeline for mining bug fix patterns in Rust. Ruxanne consists of different modules for parsing Rust programs, embedding the obtained ASTs after parsing in fixed sized datapoints, and clustering these fixed sized datapoints into different groups. 
    \item We propose a novel code embedding method, that is simple, highly interpretable, and effective in capturing the most important aspects of a program change. Our code embedding approach is based on a flexible weighting scheme, that can be modified to serve to different purposes. Using this weighting scheme, we are able to give more weight to the elements that we deem more important. % We conducted an experiment to manual evaluate our embedding method. The results show that after using our tool, in more than 85\% of the datapoints, the most important element is reported at least in the first three ranks.
    \item Using Ruxanne, we mined the top 18 most starred Rust projects in Github to discover common bug fix patterns within their code changes. We stored 87,726 datapoints drawn from 57,214 commits across these 18 projects. After using the DBSCAN clustering algorithm, and a subsequent manual analysis, we obtained 20 cross-project bug fix patterns. To respect the crucial role of the Borrow-Checker module in Rust's compiler, we introduce these patterns in two groups: 8 Borrow-Checker related and 12 general bug fix patterns.
    \item In the paper, we discuss how our findings can be useful for developing program repair and fault localization tools in Rust. Also, the patterns that we introduced can open up new linting opportunities for Rust code linters. 
    \item We believe our work resembles~\cite{hanam2016discovering}~and~\cite{yang2022mining}, which studied common bug patterns in Javascript and Python, respectively.  
\end{itemize}

Thank you for your consideration.

Mohammad Robati Shirzad and Patrick Lam

\small
\bibliographystyle{plain}
\bibliography{refs}

\end{letter}

\end{document}

### 2022-11-30 run.sh

@@ -11,7 +11,11 @@ rm *.log ;
 rm *.out ;
 rm *.synctex.gz ;
 pdflatex main.tex
+pdflatex cover.tex
 bibtex main
+bibtex cover
 pdflatex main.tex
+pdflatex cover.tex
 sleep 2;
 pdflatex main.tex
+pdflatex cover.tex

### 2022-12-01 cover.tex
cover edits

@@
 
+\signature{Mohammad Robati Shirzad and Patrick Lam}
+
 \begin{document}
 \begin{letter}{}
-\opening{Dear Editor,}
+\opening{Dear Prof. Feldt,}
 
-We are submitting the attached paper, "A Study of Common Bug Fix Patterns in Rust"
-for publication in EMSE. This paper is an empirical study of the common bug fix patterns that we extracted from the most popular open-source Rust projects. Our primary contributions are as follows:
+We are submitting the attached paper, ``A Study of Common Bug Fix Patterns in Rust,'' for publication in EMSE.
+This paper is an empirical study of the common bug fix patterns that we extracted from the most popular open-source Rust projects.
+Our primary contributions are as follows:
 
@@
 Ruxanne consists of different modules for parsing Rust programs, embedding the obtained ASTs after parsing in fixed sized datapoints, and clustering these fixed sized datapoints into different groups.
-    \item We propose a novel code embedding method, that is simple, highly interpretable, and effective in capturing the most important aspects of a program change.
-Our code embedding approach is based on a flexible weighting scheme, that can be modified to serve to different purposes.
+    \item We propose a novel code embedding method.
+Our method is simple, highly interpretable, and effective in capturing the most important aspects of a program change.
+Our code embedding approach is based on a flexible weighting scheme, which can be modified to serve to different purposes.
 Using this weighting scheme, we are able to give more weight to the elements that we deem more important.
@@
 After using the DBSCAN clustering algorithm, and a subsequent manual analysis, we obtained 20 cross-project bug fix patterns.
-To respect the crucial role of the Borrow-Checker module in Rust's compiler, we introduce these patterns in two groups: 8 Borrow-Checker related and 12 general bug fix patterns.
+Given the crucial role of the Borrow-Checker module in Rust's compiler, we introduce these patterns in two groups: 8 Borrow-Checker related and 12 general bug fix patterns.
     \item In the paper, we discuss how our findings can be useful for developing program repair and fault localization tools in Rust.
 Also, the patterns that we introduced can open up new linting opportunities for Rust code linters.
-    \item We believe our work resembles~\cite{hanam2016discovering}~and~\cite{yang2022mining}, which studied common bug patterns in Javascript and Python, respectively.  
+    \item Our work is most closely related to~\cite{hanam2016discovering}~and~\cite{yang2022mining}, which studied common bug patterns in JavaScript and Python; we instead consider Rust bug patterns.
+Rust greatly differs from JavaScript and Python in being statically typed and having strict static type checking, including ownership checking via the borrow checker.
+We believe that studying the question of bug patterns in a novel context is a valuable contribution.
 \end{itemize}
 
-Thank you for your consideration.
+We have not previously submitted any related work for publication in any venue.
 
-Mohammad Robati Shirzad and Patrick Lam
+\closing{Thank you for your consideration.}
 

### 2022-12-02 main.tex
added changes
cover letter added
wrapping up

@@
 \usepackage{subfiles} % Best loaded last in the preamble
+\usepackage{xcolor}
+\usepackage{fontawesome5}
 \input{defs}
@@
 
-\title{A Study of Common Bug Fix Patterns in Rust%\thanks{Grants or other notes
-%about the article that should go on the front page should be
-%placed here. General acknowledgments should be placed at the end of the article.}
+\title{A Study of Common Bug Fix Patterns in Rust
+\thanks{Competing Interests: We have no competing interests and are funded by a Discovery Grant from Canada's Natural Science and Engineering Research Council.}
 }
 \subtitle{}
+\definecolor{orcidlogocol}{rgb}{0.65, 0.807, 0.223}
+\newcommand{\orcid}[1]{$\,$\href{https://orcid.org/#1}{\textcolor{orcidlogocol}{\faOrcid}}}
 
 %\titlerunning{Short form of title}        % if too long for running head
-
-\author{Mohammad Robati Shirzad         \and
-        Patrick~Lam %etc.
+\author{Mohammad Robati Shirzad \orcid{0000-0002-6297-8229}       \and
+        Patrick~Lam \orcid{0000-0001-8278-5400} %etc.
 }
@@
 %             \emph{Present address:} of F. Author  %  if needed
-           \and
-           Patrick Lam \at
-                200 University Ave W, Waterloo, ON N2L 3G1 \\
-                \email{patrick.lam@uwaterloo.ca}
 }
@@
 
-\section{Appendix}
-\subfile{appendix}
-
 \end{document}

### 2022-12-03 methodology.tex
wrapping up
final commit before submission

@@
 \texttt{clone}, \texttt{Rc}, \texttt{Box}.
-The full list of \texttt{BCE} elements can be found in our replication package.
+The full list of \texttt{BCE} elements can be found in our replication package\footnote{\url{https://zenodo.org/record/7388618}}.
 
@@
 Because we are interested in changes, we computed (using process metrics provided by our mining library) the average code churn of the files within the projects.
-denoland/deno and swc-project/swc are the projects with largest number of LOCs and highest average code churn, respectively.
-Unsurprisingly, we captured a lot of instances from these two projects in our final clusters.
+swc-project/swc, nushell/nushell, and denoland/deno are the projects with largest number of LOCs.
+Unsurprisingly, we captured a lot of instances from these projects in our final clusters.
 

### 2022-12-05 methodology.tex
ready for resubmit

@@
 
-A reasonable work around would be to reduce $n$ from the full set of non terminals to a smaller set of categories (e.g.
+A reasonable work around would be to pass from the full set of non terminals to a smaller set of categories (e.g.
 a category for larger entites like class or function definitions, and a category for smaller entities like statements and expressions), and then to record all possible combinations of these categories (similar to \cite{hanam2016discovering}).
-That approach reduces the number of dimensions and provides more reasonable and syntactically-correct combinations.
+That ad-hoc approach reduces the number of dimensions and provides more reasonable and syntactically-correct combinations.
 However, it would still yield a sparse dataset.

### 2023-05-17 methodology.tex
reviewer2 - minor revision#2 - text bug
reviewer1 - revision#2 - link patch to vec

@@
 A path represents a sequence of involved non terminals.
-We found it crucial to record the order of nodes in the path, so that we could record that order in our datapoint embedding.
+It is crucial to record the order of nodes in the path, so that we could record that order in our datapoint embedding.
 
@@
 
-        \node[draw, minimum height=0.5, fill=yellow!20] at (4.29,1.73) {Occurrences data};
+        \node[draw, minimum height=0.5, fill=yellow!20] at (3.66,1.72) {Occurrences data};
 
@@
         {
-        Element & ItemFn  & Local & ExprCall & ExprReference & Path & Ident & ExprMethodCall & clone\\
-        \# & 1 & 1 & 1 & 1 & 2 & 3 & 1 & 1 \\
+        Element & ItemFn  & Local & ExprCall & ExprReference & Path & Ident & ExprMethodCall & clone & other\\
+        \# & 1 & 1 & 1 & 1 & 2 & 3 & 1 & 1 & 0\\
         };
 
-        \node[draw, minimum height=0.5, fill=yellow!20] at (4.12,-0.25) {Weighted data};
+        \node[draw, minimum height=0.5, fill=yellow!20] at (3.48,-0.25) {Weighted data};
 
@@
         {
-        Element & ItemFn  & Local & ExprCall & ExprReference & Path & Ident & ExprMethodCall & clone\\
-        weight & 0.0004 & 2.21e-05 & 8.68e-05 & 1 & 2.60e-05 & 2.76e-05 & 4.75e-05 & 1 \\
+        Element & ItemFn  & Local & ExprCall & ExprReference & Path & Ident & ExprMethodCall & clone & other \\
+        weight & 0.0004 & 2.21e-05 & 8.68e-05 & 1 & 2.60e-05 & 2.76e-05 & 4.75e-05 & 1 & ... \\
         };
 
-        \node[draw=none] (circlepack) at (10.5,-6.2)
+        \node[draw, minimum height=0.5, fill=yellow!20] at (3.7,-2.25) {Essence of change};
+
+        \node[draw=none] at (8.2,-1.95) {Result};
+        \draw [-stealth,double,line width=1,color=blue!40](7,-1.6) -- (7,-2.4);
+
+        \matrix (table3) [table,text width=6em,nodes={scale=0.45,transform shape}] at (8.5,-3)
+        {
+        Element & ItemFn  & Local & ExprCall & ExprReference & Path & Ident & ExprMethodCall & clone & other \\
+        value & 0.0004 & 2.21e-05 & 8.68e-05 & 1 & 5.20e-05 & 8.28e-05 & 4.75e-05 & 1 & 0\\
+        };
+
+        \node[draw=none] (circlepack) at (10.5,-8.2)
             {\includegraphics[width=0.7\textwidth]{circlepack.png}};
 
-        \node[draw=none] at (9,-2.3) {Visualizing Essence};
-        \draw [-stealth,double,line width=1,color=blue!40](7,-1.6) -- (7.5,-3.4);
+        \node[draw=none] at (9,-4.3) {Visualizing Essence};
+        \draw [-stealth,double,line width=1,color=blue!40](7,-3.6) -- (7.5,-5.4);
 
@@
 
-    \caption{\label{fig:essence}Multiplying the number of occurrences of items observed in the tree by their respective weights results in the essence of change, visualized with a circle pack figure at bottom right.
-Vectors show (above) the number of occurrences and (below) weighted occurrence counts.}
+    \caption{\label{fig:essence}Multiplying the number of occurrences of items observed in the tree by their respective weights results in the essence of change, visualized with a circle pack figure at bottom right.}
 \end{figure}
@@
 The radius of each circle corresponds to the importance of the item inside it.
+The essence of change vector essentially serves as our embedding vector, transforming the patch illustrated in Figure~\ref{fig:extraction} into a fixed-sized vector as depicted in Figure~\ref{fig:essence}.
 

### 2023-05-17 related.tex
todo resolved
reviewer1 - revision#6 - more related work

@@
 
+\subsection{Automated Patch Correctness Assessment}
+
+Overfitting poses a substantial challenge for automated program repair (APR) systems, where synthesized patches may pass test cases but fail to be correct.
+Automated patch correctness assessment (APCA) aims to inspect the correctness of generated patches using dynamic and static analyses.
+While our project has a different objective than APCA, we share similarities in design decisions with APCA systems, particularly in feature extraction.
+\cite{ye2021automated} propose ODS, an overfitting patch detection system, which assumes that code features capturing universal correctness properties can be utilized to classify overfitting patches across program repair systems and software projects.
+ODS extracts 202 manually crafted code features by comparing fixed and buggy code and employs supervised learning to build a distributional model for classifying unseen patches, achieving over 70\% accuracy in evaluations on Defects4J, Bugs.jar, and Bears benchmarks.
+Similarly, our feature extraction involves manual crafting of features.
+In another study, \cite{tian2022change} present an enhanced APCA model that transforms the problem into a question-answering scenario and incorporates a neural architecture to learn semantic correlations between bug reports and commit messages.
+Additionally, \cite{lin2022context} introduce \textsc{Cache}, a neural-based context-aware AST embedding method that captures both the change part and the unchanged part (context code) of patches.
+Evaluation results demonstrate that context knowledge significantly enhances the performance and accuracy of APCA systems.
+
 \subsection{Code Embedding}
@@
 In code2vec, the layout of the final vector is a parameter of the system and  is not related to the underlying programming language.
+Building on the same ideas, as a subsequent work, the same authors also introduced code2seq~\citep{alon2019code2seq}, an approach for transforming a code snippet to a sequence encoding.
+They evaluated their work on three seq2seq use cases: (1) method name prediction, (2) code captioning, and (3) code documentation.
+code2seq proved to have a better understanding of syntactical structure of the code and outperformed previous neural machine translation systems.
+
+\cite{hoang2020cc2vec} introduced CC2Vec, a neural network model that utilizes log messages accompanying code changes to learn representations capturing the semantic information.
+By addressing challenges in distinguishing between added and removed code and incorporating an attention mechanism, CC2Vec outperforms previous techniques for code change representation.
+The model employs a hierarchical attention network, which incorporates a multi-level bidirectional GRU recurrent neural network, to encode information about changed code tokens, lines, and hunks into vectors representing added and removed changes.
+These vectors are then combined and fed into a hidden layer for predicting the target function.
+Evaluation across log message generation, bug fixing patch identification, and just-in-time defect prediction tasks demonstrates the superior performance of CC2Vec compared to state-of-the-art approaches in all three tasks.
 
@@
 Our datapoints are more dense than theirs.
-\todo[inline]{quantify?}
+More specifically, the number of our dimensions is more than 11x smaller.
 

### 2023-05-17 methodology.tex
reviewer2 - major revisions -- revision#3 -- better explanation

@@
 It is crucial to record the order of nodes in the path, so that we could record that order in our datapoint embedding.
+Note that the tree depicted in Figure~\ref{fig:extraction} is not an AST; rather, it represents an ASTDiff.
+The number of paths in the ASTDiff does not directly correspond to the number of changed elements in the source code.
+Instead, it is determined by the structural differences between the two ASTs of the code versions.
+The paths in the ASTDiff capture the changes in the code structure between the two versions, allowing us to analyze and extract relevant information for further processing.
 
@@
 We propose a novel representation.
-Similar to the naive approach, we order-insensitively define the columns of our dataset as the set of items $\mathtt{NT} \cup \mathtt{BCE}$ we collected from Syn (the columns are constant across all Rust programs).
-However, to simulate order-sensitiveness, we add two more steps.
-First, we record the number of occurrences of items within the paths at each scope.
-The reason behind this decision is because a different order of program elements yields a different number of occurrences of underlying dimension.
-Second, we semi-automatically design a weighting scheme to prioritize non terminals that we think are more salient for Rust bugs.
-Section~\ref{sec:weighting_scheme}, immediately below, illustrates this approach on an example.
+Similar to the naive approach, we define the columns of our dataset as the set of items $\mathtt{NT} \cup \mathtt{BCE}$ collected from Syn.
+However, to account for order-sensitivity, we introduce two additional steps.
+Firstly, we record the number of occurrences of items from $\mathtt{NT} \cup \mathtt{BCE}$ within each root-level scope.
+This captures the varying occurrence counts resulting from different element orders in the program.
+Secondly, we incorporate a weighting scheme, designed semi-automatically, to prioritize non-terminals that we consider more salient for Rust bugs.
+Section~\ref{sec:weighting_scheme}, immediately below,
+illustrates this approach on an example.
 

### 2023-05-17 results.tex
reviewer2 - minor revision#4 - compares --> inspects
reviewer2 - major revisions -- revision#3 -- better explanation

@@
 are recognized as important.
-A human analyzer (paper author) compares the actual change in the code and compares it to the visual representation of the respective data point.
+A human analyzer (paper author) inspects the actual change in the code and compares it to the visual representation of the respective data point.
 We used circle pack figures to visualize each data point~\citep{collins2003circle}, generated with the packcircles python library\footnote{\url{https://github.com/mhtchan/packcircles}}.
@@
 The reason behind the change is that the object is a reference type of \verb+T+, while \verb+T+ directly implements \verb+to_string()+.
-Adding the dereference makes the compiler use the specialized implementation of \verb+to_string()+ and not go through slower string formatting methods.
+Adding the dereference makes the compiler use the specialized implementation of \verb+to_string()+ and not go through slower string formatting methods\footnote{\url{https://rust-lang.github.io/rust-clippy/master/index.html\#inefficient\_to\_string}}.
 

### 2023-05-18 results.tex
added actionability paragraphs to results section

@@
 We have given each pattern a specific ID (ID column) and determined the number of instances manifesting that fix pattern (\# Instances column), as well as the number of projects in which the pattern was seen (\# Project column).
+Since these patterns are the general Rust patterns, it is worth mentioning that similar patterns can also be observed in other programming languages.
+To make this more clear, we will highlight these similarities in each pattern's description.
+We will also discuss the actionability of some of the patterns.
+Moreover, irrespective of their actionability, the empirical frequency numbers associated with each pattern can serve as valuable guidance for program repair tools.
+These frequency numbers provide insights into the importance and prevalence of the bugs, enabling repair tools to prioritize and address them accordingly.
 Now we provide a detailed description for each of these patterns:
@@
 Inner Attributes apply to the program element encapsulating them, while Outer Attributes apply to the program element coming after the attribute.
+The concept of attributes in Rust is similar to Java's annotations and Python's decorators.
+While each language may have its own syntax and specific use cases, the fundamental idea of attaching additional information to code elements is shared across these languages.
 \\
@@
 For instance, in the code snippet provided above, the developers chose to make Serde (a serialization/deserialization tool for Rust) call a different function (\verb+is_empty+ instead of \verb+Option::is_none+) to determine whether to skip serializing the \verb+attributes+ field.
+\\
+
+\noindent\textbf{Actionability:} Bug fix patterns related to attributes in Rust, such as adding or removing attributes can be valuable in designing program repair tools.
+For instance, by detecting and removing attributes that are no longer needed, program repair tools can optimize the codebase and eliminate unnecessary overhead.
+On the other hand, when adding an attribute, these patterns offer insights into how repair tools can manipulate attributes and modify other parts of the code accordingly to make it adhere to the user's specifications.
 
@@
 for encapsulation purposes.
+\\
+
+\noindent\textbf{Actionability:} The patterns related to access modifiers offer an opportunity for repair tools to make sure that the reachability of program entities follows the user's specifications, enhancing security and maintainability.
+Additionally, the empirical numbers revealing the frequency of struct field modifications in Rust shed light on the significance of these bug fix patterns and emphasize the need for further exploration on how program repair tools can effectively modify code based on changes in struct fields to repair program behavior.
+These patterns provide valuable insights for program repair tools to automatically analyze and modify the codebase, ensuring it aligns with the desired behavior.
 
@@
 In this change pattern, the developer changes the type of a struct field to an \texttt{Option} of that type, hence enabling them to store no value in that field.
-
 For instance, in the Rust project swc-project/swc (a fast TypeScript/JavaScript compiler), the developer needed to modify the CSS parser source code to account for empty @media queries\footnote{\scriptsize \url{https://github.com/swc-project/swc/commit/75a14f98b7370226115ee24eec6eb8c802bd4837}}.
 As shown in the snippet above, they changed type \texttt{MediaQueryList} to \verb+Option<MediaQueryList>+, and modified the other parts of the source code accordingly.
+\\
+
+\noindent\textbf{Actionability:} When the user's specification requires a variable to be able to assume \texttt{None} values, these patterns indicate the need to modify the program accordingly.
+A program repair tool can leverage these patterns to automatically update the variable's type to \texttt{Option<T>}, ensuring that it can now hold \texttt{None} as a valid value.
+Furthermore, such a tool should also systematically update other occurrences of that variable throughout the program to account for the change and ensure program correctness.
 
@@
 Similar to the previous table, we describe BC-related patterns in Table~\ref{table:bc}.
-Now, we provide a detailed description of each of these patterns:
+It is important to note that due to the unique nature of the Borrow-Checker as an exclusive component of Rust, these patterns may not have direct counterparts or equivalents in other languages.
+Now we provide a detailed description of each of these patterns:
 
@@
 
+\noindent\textbf{Actionability:} A program repair tool can efficiently implement this modification by examining the necessity of the clone operation and considering the user's specifications regarding variable cloning.
+If the clone is deemed unnecessary, the tool can effortlessly replace it with a reference to reduce the overhead of unnecessary copying, as discussed in the example above.
+\\
+
 \subsubsection{Ref and Deref}
@@
 This object can then be used in a subsequent function invocation (\verb+remove_dir_all(repo_dir)+), which fixes a bug.
+\\
+
+\noindent\textbf{Actionability:} This pattern often occurs as part of a larger change, making it challenging to create a generalized repair tool that covers all possible scenarios.
+Typically, when a reference is passed instead of transferring ownership, there is subsequent usage of the variable after the function call.
+Although creating a comprehensive repair tool for such cases is complex, empirical results underscore the significance and prevalence of this fix, warranting further investigation.
 \\
@@
 We never observed this pattern in refactoring changes; it always accompanied a change in program behaviour.
+\\
+
+\noindent\textbf{Actionability:} A repair tool can utilize this pattern to meet the user's specifications by introducing mutability to the target variable and modifying other parts of the program accordingly.
+For instance, if the target variable has associated \texttt{impl} blocks, the change may involve modifying method annotations to ensure that the calling object is mutable, aligning with the desired specifications.
 \\

### 2023-05-18 discussion.tex
reorganized discussion

@@
 
+\subsection{Usefulness of Rust Bug Patterns}
+
 We believe our results can help researchers to create IDE tools for Rust.
-For instance, we discussed that \textit{Modifying the attributes of structs} is a common pattern that is observed in bug fix code changes.
-Assuming that the correct usage of attributes can be found other locations in the codebase~\citep{forrest2009genetic}, a code linter tool can be crafted to look for correct attribute lists.
-
+We have discussed the actionability of some of the general patterns and three BC-related patterns that are not detected by Clippy.
+For instance, we discussed considering the common pattern of \textit{Modifying the attributes of structs}, a code linter tool could be created to search for correct attribute lists within the codebase~\citep{forrest2009genetic}, leveraging the assumption that proper attribute usage can be found elsewhere.
+Also, the linter can analyze the usages of the struct to determine the necessity of adding the attribute.
+Removing unnecessary attributes not only optimizes the codebase but also eliminates unnecessary overhead and reduces the binary size.
 Similarly, we presented the pattern \textit{Dropping clone and adding borrowing}.
@@
 Using state-of-the-art methods for program repair, one could design a tool to recognize this pattern and change variable cloning to simple borrowing.
+These findings and insights can greatly aid researchers in creating effective IDE tools tailored for Rust development.
+Moreover, the empirical frequency numbers associated with each pattern, highlighting its prevalence, can provide valuable guidance for the development of program repair tools.
 

### 2023-05-21 methodology.tex
large clusters

@@
 
-In our manual analysis, we randomly picked 50 datapoints of each cluster.
-If the cluster had fewer than 50 datapoints, we analyzed all of the datapoints in the cluster.
-Regarding general patterns, we obtained 577 clusters, out of which 428 included cross-project patterns.
-Of these 428 clusters, 35 clusters included more than 50 points.
-That is, we did not do random sampling for 393 clusters, we just manually analyzed all the datapoints in them.
-Regarding BC-related patterns, out of 102 obtained clusters, 49 clusters included cross-project patterns.
-Only 4 of these 49 clusters included more than 50 datapoints, therefore, we only did a random sampling for 4 clusters.
+For our manual analysis, we analyzed all datapoints in clusters with fewer than 50 datapoints, and randomly picked 50 datapoints from clusters with more than 50 datapoints.
+Such large clusters account for 35/428 cross-project general patterns (out of 577 general pattern clusters) and 4/49 BC-related patterns (out of 102 BC-related pattern clusters).
+We exhaustively analyzed all datapoints in the 393 small cross-project general patterns and 44 BC-related clusters.
+
 For each cluster, we read the code and the bug reports of the datapoints (if they existed).

### 2023-05-22 results.tex
text changes

@@
 
-\noindent\textbf{Actionability:} Bug fix patterns related to attributes in Rust, such as adding or removing attributes can be valuable in designing program repair tools.
-For instance, by detecting and removing attributes that are no longer needed, program repair tools can optimize the codebase and eliminate unnecessary overhead.
-On the other hand, when adding an attribute, these patterns offer insights into how repair tools can manipulate attributes and modify other parts of the code accordingly to make it adhere to the user's specifications.
+\noindent\textbf{Actionability:} Bug fix patterns related to attributes in Rust, such as removing or adding attributes can be valuable in designing program repair tools.
+For instance, by recognizing situations where attributes are no longer needed, program repair tools can propose changes that optimize the codebase and eliminate unnecessary overhead.
+On the other hand, a repair tool could also recognize situations where attributes need to be added, enabling the code to better adhere to the user's specifications.
 

### 2023-05-23 results.tex
rq2

@@
 adding a whole new \verb+impl+ block and implementing new functions in it).
-It is worth mentioning that these change patterns tend to apply across projects less often, as having many changed elements make it easier for datapoints to be dissimlar.
+Such failing change patterns tend to apply across projects less often, as having many changed elements make it easier for datapoints to be dissimlar.
 
@@
 We collected 60583 datapoints for general fix patterns using a server with a Xeon Gold 5120 processor (Q3-2017, 14-core 2.2 to 3.2GHz) and the data collection took around 26 hours to compute both $D_g$ and $D_b$.
-We used $\epsilon=0.0001$ and $Z=5$ as our DBSCAN parameters and that resulted in 577 clusters.
+We used $\epsilon=0.0001$ and $Z=5$ as our DBSCAN parameters, resulting in 577 clusters.
 After manual analysis, as described in~\ref{sec:manual_analysis_cluster_selection}, we ended up with 12 general patterns, which we describe in Table~\ref{table:general}.
@@
 We have given each pattern a specific ID (ID column) and determined the number of instances manifesting that fix pattern (\# Instances column), as well as the number of projects in which the pattern was seen (\# Project column).
-Since these patterns are the general Rust patterns, it is worth mentioning that similar patterns can also be observed in other programming languages.
-To make this more clear, we will highlight these similarities in each pattern's description.
+Since these patterns are the general Rust patterns, analogous patterns can also occur in other programming languages, and we will discuss such patterns when describing each pattern.
 We will also discuss the actionability of some of the patterns.
-Moreover, irrespective of their actionability, the empirical frequency numbers associated with each pattern can serve as valuable guidance for program repair tools.
-These frequency numbers provide insights into the importance and prevalence of the bugs, enabling repair tools to prioritize and address them accordingly.
-Now we provide a detailed description for each of these patterns:
+Moreover, no matter whether we can find actionable outcomes for the patterns, other researchers can use the empirical pattern frequency numbers that we report to build their own program repair tools, prioritizing more-frequently-repaired patterns.
+We next provide a detailed description of each of these patterns:
 

### 2023-05-23 results.tex
new examples for attr patterns

@@
 // G.attr.struct.add: Adding attributes to a struct
-<@\textcolor{green}{++}@> #[derive(Debug, Clone)]
-pub struct Call {
-/// body 
+<@\textcolor{red}{--}@> #[derive(Clone, Properties)]
+<@\textcolor{green}{++}@> #[derive(Clone, Properties, PartialEq)]
+pub struct CounterProps {
+    pub destroy_callback: Callback<()>,
 }
@@
 //G.attr.struct.drop: Removing attributes from a struct
-<@\textcolor{red}{--}@> #[derive(Debug, Clone)]
-pub struct Call {
-/// body 
+<@\textcolor{red}{--}@> #[derive(Serialize, Deserialize, Clone, PartialEq, Eq, Debug)]
+<@\textcolor{green}{++}@> #[derive(Serialize, Deserialize, Clone, PartialEq, Eq)]
+pub struct Subset {
+    /// Invariant, maintained by `SubsetBuilder`: all `Segment`s have non-zero
+    /// length, and no `Segment` has the same count as the one before it.
+    segments: Vec<Segment>,
 }
@@
 // G.attr.struct.change: Changing an existing attribute of a struct
-<@\textcolor{red}{--}@> #[derive(Debug)]
-<@\textcolor{green}{++}@> #[derive(Debug, Clone)]
-pub struct Call {
-/// body 
+<@\textcolor{red}{--}@> #[derive(Deserialize, Debug, Copy, Clone, PartialEq, Eq)]
+<@\textcolor{green}{++}@> #[derive(ConfigDeserialize, Debug, Copy, Clone, PartialEq, Eq)]
+pub enum SearchAction {
+    SearchFocusNext,
+    SearchFocusPrevious,
+    SearchConfirm,
+    SearchCancel,
+    SearchClear,
+    SearchDeleteWord,
+    SearchHistoryPrevious,
+    SearchHistoryNext,
 }
@@
 This modification can happen in three different ways: Adding attributes to the struct (G.attr.struct.add); Removing attributes from the struct (G.attr.struct.drop); Changing the content of an existing attribute (G.attr.struct.change).
-These attributes are Outer Attributes as we can find struct-related non terminals within the ASTDiff of the instances of this pattern.
-\\
+These attributes are Outer Attributes as we can find struct-related non terminals within the ASTDiff of the instances of this pattern.\\
 
@@
 pub struct Retain {
-    <@\textcolor{green}{++}@> #[serde(skip_serializing_if = "is_empty")]
+    <@\textcolor{green}{++}@> #[serde(skip_serializing_if = "Option::is_none")]
     pub attributes: Option<Attributes>,
@@
 // G.attr.struct.field.drop: Removing attributes from struct fields
-#[derive(Serialize)]
-pub struct Retain {
-    <@\textcolor{red}{--}@> #[serde(skip_serializing_if = "Option::is_none")]
-    pub attributes: Option<Attributes>,
+pub struct ExpandContext<'context> {
+    #[get = "pub(crate)"]
+    pub registry: Box<dyn SignatureRegistry>,
+    <@\textcolor{red}{--}@> #[get = "pub(crate)"]
+    pub source: &'context Text,
+    pub homedir: Option<PathBuf>,
 }

### 2023-05-23 results.tex
attributes

@@
 
-Attributes in Rust enable developers to add metadata to program elements.
+The most common bug fix pattern in our corpus modifies attributes.
+In Rust, attributes enable developers to add metadata to program elements.
+Some of the attributes---presumably all of the ones involved in bug fixes---affect the semantics of the program; for instance, libraries can generate different code based on the attributes that the developer specifies.
+
 Developers can add attributes to structs, struct fields, enum variants, match expression arms, etc.
@@
 Inner Attributes apply to the program element encapsulating them, while Outer Attributes apply to the program element coming after the attribute.
-The concept of attributes in Rust is similar to Java's annotations and Python's decorators.
+
+The concept of attributes in Rust is similar to Python's decorators and Java's annotations (though Java annotations may be less likely to have runtime effect; for instance, Java @Override adds additional compile-time constraints).
 While each language may have its own syntax and specific use cases, the fundamental idea of attaching additional information to code elements is shared across these languages.
@@
 
-\noindent\textbf{Description:} In this pattern, the attribute set of a struct changes.
-This modification can happen in three different ways: Adding attributes to the struct (G.attr.struct.add); Removing attributes from the struct (G.attr.struct.drop); Changing the content of an existing attribute (G.attr.struct.change).
-These attributes are Outer Attributes as we can find struct-related non terminals within the ASTDiff of the instances of this pattern.\\
+\noindent\textbf{Description:} In this pattern, the developer changes the attribute set of a struct.
+There are three variants of this pattern: Adding attributes to the struct (G.attr.struct.add); Removing attributes from the struct (G.attr.struct.drop); Changing the content of an existing attribute (G.attr.struct.change).
+These attributes are Outer Attributes as we can find struct-related non terminals within the ASTDiff of the instances of this pattern.
+
+[discussion on how many of the changes are of each variant: over 90\% are modify]
+
+In the specific cases above, the Add change requests that additional boilerplate code be generated to evaluate partial equivalence on \texttt{CounterProps} (new functionality); the Remove change causes debug code to be removed (performance fix); and the Change change requests different deserialization code be generated.
+\\
 
@@
 
-\noindent\textbf{Actionability:} Bug fix patterns related to attributes in Rust, such as removing or adding attributes can be valuable in designing program repair tools.
-For instance, by recognizing situations where attributes are no longer needed, program repair tools can propose changes that optimize the codebase and eliminate unnecessary overhead.
-On the other hand, a repair tool could also recognize situations where attributes need to be added, enabling the code to better adhere to the user's specifications.
+\noindent\textbf{Actionability:} Because attributes have semantic effects, program repair tools can effect repairs by adjusting the set of attributes included.
+Because so many of our bug fixes change attributes, we believe that IDEs ought to be able to propose relevant attribute changes, as witnessed in our corpus.
+The example in the text about changing Deserialize to ConfigDeserialize is specific to that program, but other changes would apply more generally.
+\todo{which?} Also, we demonstrated a case where the developer removed a \texttt{Debug} attribute.
+A novel type of performance program repair tool working with a run-time profiler could identify unnecessary code and remove attributes that generate it.
 

### 2023-05-23 discussion.tex
top 10

@@
 
-Furthermore, we expected to obtain some patterns associated with certain Rust-related features.
-However, our results did not include those patterns.
+All of our parsing, path extraction, weighting scheme, and clustering modules, along with our final results, can be found in our replication package, which we provide for verification, reusability and further extension.
+Weightings can be modified to make the embedding focus on a specific set of elements for building different code embeddings.
+
+\subsection{Patterns Present and Missing}
+
+When presenting each pattern, we discussed actionability for the most important actions. Here, we discuss the implications of our most important patterns.
+
+The two most common patterns were general patterns modifying struct fields and attributes, with 179 and 172 instances respectively.
+Unfortunately, struct field modifications are neither interesting nor actionable; they are simply used to enable other changes.
+Attribute modifications are far more interesting, and account for the second- and fifth-most popular pattern.
+From a language design perspective, the fact that there are so many attribute changes marked as bug fixes illustrates the semantic importance of attributes in Rust, and implies that program repair must understand attributes.
+A further challenge is that different attributes have different, and often library-dependent, semantic meanings.
+Proposing specific repair techniques that reason about attributes is beyond the scope of this work, but our results suggest that they are important to develop.
+
+The most common borrow-checker-related pattern and third-most-common pattern overall is dropping \texttt{clone()}.
+This pattern is performance-related; when correctly applied, it has no semantic impact on the program, but reduces resource consumption.
+Rust is used in situations where performance is important; the number of instances of this pattern confirms that developers do work on performance issues.
+This implies that program repair tools would benefit from being performance-aware, rather than limiting themselves strictly to functional properties of the code.
+
+The fourth- and sixth-most popular patterns, changing the type of a struct field or of a generic type parameter, are difficult to make general statements about; there are many unrelated idiosyncratic changes in these clusters.
+Such changes look the same, but have different purposes.
+One example we showed changed a \texttt{HashMap} to an \texttt{IndexMap}, preserving iteration order.
+That specific nondeterminism bug could be automatically repaired upon direction from the developer.
+Another change modified \texttt{usize} to \texttt{isize}, perhaps reflecting a requirements change to allow signed integers.
+IDEs could support refactorings that modify types.
+We find it difficult to propose further general work in this direction.
+
+The seventh-most-popular pattern is the Rust-specific change of dropping mutability (i.e.
+making data immutable).
+Tools like ReImInfer~\cite{HuangMDE2012} for a Java extension, and techniques like the one proposed by Eyolfson~\cite{EyolfsonAbstractImmutability} for C++, aim to infer that data is immutable.
+In the Rust context, it is easier to make data immutable: the semantics for immutability are clear.
+Immutability is part of the core Rust language.
+mmutability escape hatches in C++ complicate issues, but in Rust, the escape hatch is the clearly-marked \texttt{unsafe} block.
+This pattern thus suggests that tools for inferring Rust immutability could be viable.
+
+We are unable to say anything meaningful about the eighth- and ninth-most popular patterns, \texttt{G.match.pattern} and \texttt{G.type.enum.variant}.
+They do not seem particularly helpful for automatic program repair.
+
+The tenth-most-popular pattern, making a struct field \texttt{pub} and hence visible to external modules, suggests that tools to help developers manage visibility would be helpful.
+Public accessibility of methods and fields very much affects library compatibility and usability.
+
+Conversely, we were surprised to find that Ruxanne reported no patterns associated with certain Rust-related features.
 For instance, we did not observe any patterns associated with the Rust \verb+move+ operator, which has been known to trouble many beginning Rust programmers.
@@
 Rust compiler checks prevent such errors.
-Our speculation is that, since Rust enforces strict compiler checks, programmers usually get many errors, and they only push their code once it is free of compile errors.
-We thus would not observe patterns which don't make it to the repository.
+Our speculation is that, since Rust enforces strict compiler checks, programmers usually get many compile-time errors, but they only push their code once it is free of compile errors.
+We thus would not observe patterns which do not make it to the repository.
 
@@
 
-We believe our results can help researchers to create IDE tools for Rust.
+We believe our results can help with the development of IDE and program repair tools for Rust.
 We have discussed the actionability of some of the general patterns and three BC-related patterns that are not detected by Clippy.
-For instance, we discussed considering the common pattern of \textit{Modifying the attributes of structs}, a code linter tool could be created to search for correct attribute lists within the codebase~\citep{forrest2009genetic}, leveraging the assumption that proper attribute usage can be found elsewhere.
+For instance, as we have discussed, for the common pattern of \textit{Modifying the attributes of structs}, a code linter tool could search for correct attribute lists within the codebase~\citep{forrest2009genetic}, leveraging the assumption that proper attribute usage can be found elsewhere.
 Also, the linter can analyze the usages of the struct to determine the necessity of adding the attribute.
-Removing unnecessary attributes not only optimizes the codebase but also eliminates unnecessary overhead and reduces the binary size.
+Removing unnecessary attributes not only optimizes the codebase but also eliminates unnecessary overhead and reduces binary size.
 Similarly, we presented the pattern \textit{Dropping clone and adding borrowing}.
-As discussed, Clippy does not detect this pattern, and repetitive clones can be computationally expensive and might lead to hogging the CPU.
+As discussed, Clippy does not detect this pattern, and repetitive clones can be computationally expensive.
 Using state-of-the-art methods for program repair, one could design a tool to recognize this pattern and change variable cloning to simple borrowing.
@@
 Such elements may be irrelevant for repair because they did not appear in the bug patterns that we presented.
-
-All of our parsing, path extraction, weighting scheme, and clustering modules, along with our final results, can be found in our replication package, which we provide for verification, reusability and further extension.
-Weightings can be modified to make the embedding focus on a specific set of elements for building different code embeddings.
 

### 2023-05-23 results.tex
top 10

@@
 
+\todo{please confirm: G.attr.struct.field is not a subset of G.attr.struct.}
+
 \noindent\textbf{Actionability:} Because attributes have semantic effects, program repair tools can effect repairs by adjusting the set of attributes included.
@@
 
-\noindent\textbf{Actionability:} The patterns related to access modifiers offer an opportunity for repair tools to make sure that the reachability of program entities follows the user's specifications, enhancing security and maintainability.
+\noindent\textbf{Actionability:} The patterns related to access modifiers suggest an opportunity for repair tools to make sure that the accessability of program entities follows the user's specifications, enhancing encapsulation, security, and maintainability.
 Additionally, the empirical numbers revealing the frequency of struct field modifications in Rust shed light on the significance of these bug fix patterns and emphasize the need for further exploration on how program repair tools can effectively modify code based on changes in struct fields to repair program behavior.
-These patterns provide valuable insights for program repair tools to automatically analyze and modify the codebase, ensuring it aligns with the desired behavior.
 
@@
 \noindent\textbf{Description:} In this pattern, a change occurs in an arm's pattern.
-Like the code snippet provided above, it can be a change in pattern's type.
+Like the code snippet provided above, it can be a change in a match pattern's type.
 It also can be matching to a different enum variant or string.
@@
 There are more than 500 lints included in Clippy.
-Since detecting whether a clone is redundant is undecidable, Clippy would have to conservatively estimate redundancy.
-By definition, removing a redundant clone does not modify functional aspects of the program behaviour.
-However, such a removal can greatly improve performance, if the program was formerly cloning large objects.
+Since detecting whether a clone is redundant is undecidable, Clippy conservatively estimates redundancy.
+By definition, removing a clone that is indeed redundant does not modify functional aspects of the program behaviour.
+(Removing a non-redundant clone introduces a compile error or, worse yet, a bug.) However, such a removal can greatly improve performance, if the program was formerly cloning large objects.
 \\
@@
 
-\noindent\textbf{Actionability:} A program repair tool can efficiently implement this modification by examining the necessity of the clone operation and considering the user's specifications regarding variable cloning.
-If the clone is deemed unnecessary, the tool can effortlessly replace it with a reference to reduce the overhead of unnecessary copying, as discussed in the example above.
+\noindent\textbf{Actionability:} A program repair tool can implement clone removal by examining the necessity of the clone operation and considering the user's specifications regarding variable cloning.
+If the clone is unnecessary, the tool can replace it with a reference to reduce the overhead of unnecessary copying, as discussed in the example above.
 \\

### 2023-05-27 discussion.tex
add two todos

@@
 making data immutable).
-Tools like ReImInfer~\cite{HuangMDE2012} for a Java extension, and techniques like the one proposed by Eyolfson~\cite{EyolfsonAbstractImmutability} for C++, aim to infer that data is immutable.
+Tools like ReImInfer by~\cite{HuangMDE2012} (for a Java extension), and techniques like the one proposed in~\cite{EyolfsonAbstractImmutability} for C++, aim to infer that data is immutable.
 In the Rust context, it is easier to make data immutable: the semantics for immutability are clear.

### 2023-05-27 main.tex
final commit before submission
add two todos

@@
 
-\institute{Mohammad Robati Shirzad \at
-              200 University Ave W, Waterloo, ON N2L 3G1 \\
-              \email{mrobatis@uwaterloo.ca}           %  \\
+\institute{Mohammad Robati Shirzad --- \email{mrobatis@uwaterloo.ca} \\
+           Patrick Lam --- \email{patrick.lam@uwaterloo.ca} \at
+           University of Waterloo --- Address: 200 University Ave W, Waterloo, ON N2L 3G1 \\
 %             \emph{Present address:} of F. Author  %  if needed
@@
 After clustering the code changes, and conducting a manual analysis, we discovered 20 groups of cross-project bug fix patterns, which we categorize as (1) general patterns and (2) borrow-checker-related patterns.
-We describe each of these patterns and their implications to automated program repair.
+We describe each of these patterns and their implications to automated program repair.\todo{talk about the most popular pattern}
 \keywords{Bug patterns \and Pattern mining \and Bug fix changes \and Rust}

### 2023-05-27 discussion.tex
rework most useful patterns and add more todos

@@
 We presented these patterns in two groups: 12 general patterns and 8 BC-related patterns.
-Clippy is able to detect five of the BC-related patterns.
+Clippy is able to detect 5 of the 8 BC-related patterns.
 Fundamentally, Clippy cannot detect some of the remaining patterns.
-For instance, we proposed the pattern \textit{Adding mutability}.
+For instance, we proposed pattern \textit{Adding mutability}.
 In the Rust context, this change anticipates a future change to the code which will perform a mutation of a previously immutable value.
 Predicting the future is beyond Clippy's remit.
-To the best of our knowledge, the rest of the patterns (3 BC-related patterns and 12 general patterns) cannot be reported by the current linting tools.
+To the best of our knowledge, the remaining patterns (3 BC-related patterns and 12 general patterns) cannot be reported by the current linting tools.
 
@@
 \subsection{Patterns Present and Missing}
+\label{subsec:patterns}
 
-When presenting each pattern, we discussed actionability for the most important actions. Here, we discuss the implications of our most important patterns.
+When presenting each pattern, we discussed actionability for the most important actions.
+Here, we discuss the implications of our most important patterns.
+In summary, program repair tools must understand semantics of attributes and should be performance-aware.
+It may also be useful for tools to be able to automatically infer immutability.
 
@@
 Unfortunately, struct field modifications are neither interesting nor actionable; they are simply used to enable other changes.
-Attribute modifications are far more interesting, and account for the second- and fifth-most popular pattern.
-From a language design perspective, the fact that there are so many attribute changes marked as bug fixes illustrates the semantic importance of attributes in Rust, and implies that program repair must understand attributes.
-A further challenge is that different attributes have different, and often library-dependent, semantic meanings.
+Attribute modifications are far more interesting, and account for the second- and fifth-most popular patterns.
+From a language design perspective, the fact that there are so many attribute changes marked as bug fixes illustrates the semantic importance of attributes in Rust, and \emph{implies that program repair must understand the effects of attributes}.
+Yet, different attributes have different, and often library-dependent, semantic meanings.
 Proposing specific repair techniques that reason about attributes is beyond the scope of this work, but our results suggest that they are important to develop.
@@
 Rust is used in situations where performance is important; the number of instances of this pattern confirms that developers do work on performance issues.
-This implies that program repair tools would benefit from being performance-aware, rather than limiting themselves strictly to functional properties of the code.
+This implies that \emph{program repair tools would benefit from being performance-aware}, rather than limiting themselves strictly to functional properties of the code.
 
@@
 Immutability is part of the core Rust language.
-mmutability escape hatches in C++ complicate issues, but in Rust, the escape hatch is the clearly-marked \texttt{unsafe} block.
-This pattern thus suggests that tools for inferring Rust immutability could be viable.
+Immutability escape hatches in C++ complicate issues, but in Rust, the escape hatch is the clearly-marked \texttt{unsafe} block.
+This pattern thus suggests that \emph{tools for inferring Rust immutability could be viable}.
 
@@
 Conversely, we were surprised to find that Ruxanne reported no patterns associated with certain Rust-related features.
-For instance, we did not observe any patterns associated with the Rust \verb+move+ operator, which has been known to trouble many beginning Rust programmers.
+For instance, we did not observe any patterns associated with the Rust \verb+move+ operator, which has been known to plague many beginning Rust programmers.
 \verb+move+ is used to transfer the ownership of values to a closure.
@@
 
-We believe our results can help with the development of IDE and program repair tools for Rust.
+Having discussed the most frequent patterns and their implications on repair tools, we continue by discussing potential features for repair and refactoring tools in more detail.
+We believe our results can help with the development of program repair and refactoring tools for Rust.
 We have discussed the actionability of some of the general patterns and three BC-related patterns that are not detected by Clippy.
+
 For instance, as we have discussed, for the common pattern of \textit{Modifying the attributes of structs}, a code linter tool could search for correct attribute lists within the codebase~\citep{forrest2009genetic}, leveraging the assumption that proper attribute usage can be found elsewhere.
@@
 Removing unnecessary attributes not only optimizes the codebase but also eliminates unnecessary overhead and reduces binary size.
+
 Similarly, we presented the pattern \textit{Dropping clone and adding borrowing}.
@@
 Using state-of-the-art methods for program repair, one could design a tool to recognize this pattern and change variable cloning to simple borrowing.
-These findings and insights can greatly aid researchers in creating effective IDE tools tailored for Rust development.
+
+We believe that our insights can greatly aid researchers in creating effective IDE tools tailored for Rust development.
 Moreover, the empirical frequency numbers associated with each pattern, highlighting its prevalence, can provide valuable guidance for the development of program repair tools.
 
+\todo{this next paragraph doesn't seem to fit here, do we need it?}
 A general approach for code embedding is to extract paths with semantic information that better serve a given embedding goal.
@@
 
-In our work there are two main threats to internal validity: (1) A threat to internal validity is confounding, where changes to what shows up in the embedding are not due to changes in the code being embedded.
+There are two main threats to the internal validity of our work: (1) A threat to internal validity is confounding, where changes to what shows up in the embedding are not due to changes in the code being embedded.
 In this case, the weighting scheme may contribute to confounding because we manually adjusted the weights.
@@
 SLINK) might output new clusters that our pipeline is unable to find.
+
 A threat to external validity is selection bias, where the benchmarks differ systematically from the actual world.

### 2023-05-27 results.tex
rework most useful patterns and add more todos

@@
 \textbf{What are the general fix patterns in Rust, and how often do they apply?} 
+We collected general fixes, asked DBSCAN to create clusters, and manually analyzed them.
+In this subsection, we will discuss each pattern, as well as its actionability---possible usages of the pattern for bug repair.
+Section~\ref{subsec:patterns} contains a global discussion of our patterns and their value.
 
-We collected 60583 datapoints for general fix patterns using a server with a Xeon Gold 5120 processor (Q3-2017, 14-core 2.2 to 3.2GHz) and the data collection took around 26 hours to compute both $D_g$ and $D_b$.
-We used $\epsilon=0.0001$ and $Z=5$ as our DBSCAN parameters, resulting in 577 clusters.
+Specifically, our dataset includes 60583 datapoints for general fix patterns.
+We processed our dataset on a server with a Xeon Gold 5120 processor (Q3-2017, 14-core 2.2 to 3.2GHz).
+The data collection took around 26 hours to populate the general and BC-related databases $D_g$ and $D_b$.
+We used DBSCAN with parameters $\epsilon=0.0001$ and $Z=5$, resulting in 577 clusters.
 After manual analysis, as described in~\ref{sec:manual_analysis_cluster_selection}, we ended up with 12 general patterns, which we describe in Table~\ref{table:general}.
@@
 We have given each pattern a specific ID (ID column) and determined the number of instances manifesting that fix pattern (\# Instances column), as well as the number of projects in which the pattern was seen (\# Project column).
-Since these patterns are the general Rust patterns, analogous patterns can also occur in other programming languages, and we will discuss such patterns when describing each pattern.
+
+Since these patterns are general Rust patterns, analogous patterns can also occur in other programming languages, and we will discuss such analogous patterns when describing each pattern.
 We will also discuss the actionability of some of the patterns.
-Moreover, no matter whether we can find actionable outcomes for the patterns, other researchers can use the empirical pattern frequency numbers that we report to build their own program repair tools, prioritizing more-frequently-repaired patterns.
-We next provide a detailed description of each of these patterns:
+Moreover, other researchers can use the empirical pattern frequency numbers that we report to build their own program repair tools, prioritizing more-frequently-repaired patterns; they are not limited to our suggested actionability.
+We next provide a detailed description of each of these patterns.
 
@@
 Developers can add attributes to structs, struct fields, enum variants, match expression arms, etc.
-There are two types of attributes: Inner Attributes and Outer Attributes.
+There are two types of attributes\todo{how important is inner versus outer?}: Inner Attributes and Outer Attributes.
 Inner Attributes apply to the program element encapsulating them, while Outer Attributes apply to the program element coming after the attribute.
 
-The concept of attributes in Rust is similar to Python's decorators and Java's annotations (though Java annotations may be less likely to have runtime effect; for instance, Java @Override adds additional compile-time constraints).
+The concept of attributes in Rust is similar to Python's decorators\todo{are python decorator changes important in the Python dataset?} and Java's annotations (though Java annotations may be less likely to have runtime effect; for instance, Java @Override adds additional compile-time constraints).
 While each language may have its own syntax and specific use cases, the fundamental idea of attaching additional information to code elements is shared across these languages.
@@
 
-[discussion on how many of the changes are of each variant: over 90\% are modify]
+\todo{discussion on how many of the changes are of each variant: over 90\% are modify, but get the actual number}
 
@@
 \todo{please confirm: G.attr.struct.field is not a subset of G.attr.struct.}
+\todo{we should also get the count of subtypes}
 
 \noindent\textbf{Actionability:} Because attributes have semantic effects, program repair tools can effect repairs by adjusting the set of attributes included.
-Because so many of our bug fixes change attributes, we believe that IDEs ought to be able to propose relevant attribute changes, as witnessed in our corpus.
+Starting from the observation that so many of our bug fixes change attributes, we believe that IDEs need to be able to propose relevant attribute changes, as witnessed in our corpus.
 The example in the text about changing Deserialize to ConfigDeserialize is specific to that program, but other changes would apply more generally.
@@
 In the instances of this cluster, adding the \texttt{pub} access modifier happened in bug fixing changes, where the developer needed to access a field which had not been marked for public access.
-On the other hand, the public access might be revoked if the developer realizes there are no external accesses to the field, and that such accesses are not desirable, e.g.
+On the other hand, public access might be revoked if the developer realizes there are no external accesses to the field, and that such accesses are not desirable, e.g.
 for encapsulation purposes.
@@
 \noindent\textbf{Actionability:} When the user's specification requires a variable to be able to assume \texttt{None} values, these patterns indicate the need to modify the program accordingly.
-A program repair tool can leverage these patterns to automatically update the variable's type to \texttt{Option<T>}, ensuring that it can now hold \texttt{None} as a valid value.
+A program repair\todo{is this really a repair tool or a refactoring tool?
+I feel like many of our changes are for refactorings not repairs, despite our PR} tool can leverage these patterns to automatically update the variable's type to \texttt{Option<T>}, ensuring that it can now hold \texttt{None} as a valid value.
 Furthermore, such a tool should also systematically update other occurrences of that variable throughout the program to account for the change and ensure program correctness.
@@
 \noindent\textbf{Description:} The trait bounds are the functionalities that we require from our parametric types.
-This concept might be referred to as interfaces in other languages, although there might be small differences.
-This change pattern relates to modifications in trait bounds.
-This can happen as part of bug fixes or for refactoring.
+This concept more-or-less corresponds to interfaces in other languages, although the mapping is not exact (e.g.
+it is possible to implement traits for types that one did not initially define).
+This change pattern relates to modifications in trait bounds, which can happen as part of bug fixes or for refactoring.
 \\
@@
 
-In Rust, match control flow construct is used for comparing a value against multiple patterns, and executing the code associated with the first pattern that matches with our value.
+In Rust, the match control flow construct is used for comparing a value against multiple patterns, and executing the code associated with the first pattern that matches with our value.
 A pattern and its associated code are called a match arm.
@@
 We collected 27143 datapoints for borrow-checker related fix patterns.
-We used $\epsilon=0.001$ and $Z=5$ as our DBSCAN parameters and that resulted in 102 clusters.
-After manual analysis, as described in~\ref{sec:manual_analysis_cluster_selection}, we ended up with 8 borrow-checker related fix patterns.
+We applied DBSCAN with parameters $\epsilon=0.001$ and $Z=5$, resulting in 102 clusters.
+After manual analysis (Section~\ref{sec:manual_analysis_cluster_selection}), we ended up with 8 borrow-checker related fix patterns.
 Similar to the previous table, we describe BC-related patterns in Table~\ref{table:bc}.
-It is important to note that due to the unique nature of the Borrow-Checker as an exclusive component of Rust, these patterns may not have direct counterparts or equivalents in other languages.
-Now we provide a detailed description of each of these patterns:
+Since the Borrow-Checker is unique to Rust, patterns in this section generally do not have direct counterparts in other languages.
+We next provide a detailed description of each of these patterns.
 
@@
 By definition, removing a clone that is indeed redundant does not modify functional aspects of the program behaviour.
-(Removing a non-redundant clone introduces a compile error or, worse yet, a bug.) However, such a removal can greatly improve performance, if the program was formerly cloning large objects.
+(Removing a non-redundant clone introduces a compile error or, worse yet, a bug.) However, such a removal can greatly improve the nonfunctional property of performance, if the program was formerly cloning large objects.
 \\
@@
 
-\noindent\textbf{Actionability:} A program repair tool can implement clone removal by examining the necessity of the clone operation and considering the user's specifications regarding variable cloning.
-If the clone is unnecessary, the tool can replace it with a reference to reduce the overhead of unnecessary copying, as discussed in the example above.
+\noindent\textbf{Actionability:} Like Clippy, a program repair tool can implement clone removal by reasoning about whether the clone operation was necessary.
+If the clone is unnecessary, the tool can replace it with a reference (as seen above), thus reducing the overhead of unnecessary copying.
 \\
@@
 
-To give an example, the commit in starship/starship excerpted above\footnote{\scriptsize \url{https://github.com/starship/starship/commit/56d475578ea508631275772127f49a6949fea6b0}} shows that the developer now borrows the value of \verb+repo_dir+ to keep the ownership of the object in the current scope.
+As an example, the commit in starship/starship excerpted above\footnote{\scriptsize \url{https://github.com/starship/starship/commit/56d475578ea508631275772127f49a6949fea6b0}} shows that the developer now borrows the value of \verb+repo_dir+ to keep the ownership of the object in the current scope.
 This object can then be used in a subsequent function invocation (\verb+remove_dir_all(repo_dir)+), which fixes a bug.
@@
 \noindent\textbf{Actionability:} This pattern often occurs as part of a larger change, making it challenging to create a generalized repair tool that covers all possible scenarios.
-Typically, when a reference is passed instead of transferring ownership, there is subsequent usage of the variable after the function call.
+Typically, when a reference is passed instead of transferring ownership, the variable is used again after returning from the call.
 Although creating a comprehensive repair tool for such cases is complex, empirical results underscore the significance and prevalence of this fix, warranting further investigation.
@@
 \noindent\textbf{Description:} This change happens when the developer needs to mutate a variable which was previously immutable in a scope---they thus need to add the \verb+mut+ keyword.
-We never observed this pattern in refactoring changes; it always accompanied a change in program behaviour.
+We never observed this pattern in refactoring-only changes; it always accompanied a change in program behaviour.
 \\
 
-\noindent\textbf{Actionability:} A repair tool can utilize this pattern to meet the user's specifications by introducing mutability to the target variable and modifying other parts of the program accordingly.
-For instance, if the target variable has associated \texttt{impl} blocks, the change may involve modifying method annotations to ensure that the calling object is mutable, aligning with the desired specifications.
-\\
+\noindent\textbf{Actionability:} A repair tool can use this pattern to introduce mutability to the target variable and modify other parts of the program accordingly.
+For instance, if the target variable has associated \texttt{impl} blocks, the change may involve modifying method annotations to ensure that the calling object is mutable, aligning with the desired specifications.\todo{how would the specs be provided?
+how could this be done as part of repair vs as part of refactoring?!} \\
 

### 2023-06-04 discussion.tex
resolved all but 3 of the todos

@@
 
-\todo{this next paragraph doesn't seem to fit here, do we need it?}
-A general approach for code embedding is to extract paths with semantic information that better serve a given embedding goal.
+% A general approach for code embedding is to extract paths with semantic information that better serve a given embedding goal.
 For instance, to design an embedding method for repairing Rust programs, the designer might want to disregard certain elements, such as moving ownership to closure bodies.

### 2023-06-04 introduction.tex
ready for resubmit
reviewer2 - minor revision#1 - put Rust intro in the beginning
add two todos
resolved all but 3 of the todos

@@
+Rust is a relatively new programming language~\citep{klabnik2019rust} which allows programmers to write programs that have low-level control over resources, while still ensuring high-level safety guarantees (for programs written in safe Rust).
+What distinguishes Rust from the other systems programming languages, e.g.
+C/C++, is its ownership framework, which is an integral part of Rust.
+The ownership framework keeps track of references to different memory locations, and will free them once they go out of scope.
+If a program passes the compiler’s checks, it meets several safety properties by construction; e.g.
+it has no dangling references or double-frees.
+Such properties had to be manually ensured (or not!) in languages like C/C++.
+Moreover, not only does ownership remove the need for a garbage collector in the runtime environment, it frees the programmer of the responsibility to ensure certain safety-related properties~\citep{qin2020understanding}.
+Due to the memory and thread safety guarantees that Rust provides, many software companies have adopted Rust in production.
+For example, Dropbox decided to migrate to Rust for developing their storage system, since they were not completely satisfied with the features provided by AWS’s S3 or Golang~\citep{dropbox}.
+
 Bugs have long been an unavoidable part of computer programming, predating Grace Hopper's literal debugging session in 1947; even today, bug-free software remains an aspirational aim for software engineering research.
@@
 In this work, we study common bug patterns that appear in Rust projects.
-Rust is a relatively new programming language~\citep{klabnik2019rust} which allows programmers to write programs that have low-level control over resources, while still ensuring high-level safety guarantees (for programs written in safe Rust).
-What distinguishes Rust from the other systems programming languages, e.g.
-C/C++, is its ownership framework, which is an integral part of Rust.
-The ownership framework keeps track of references to different memory locations, and will free them once they go out of scope.
-If a program passes the compiler’s checks, it meets several safety properties by construction; e.g.
-it has no dangling references or double-frees.
-Such properties had to be manually ensured (or not!) in languages like C/C++.
-Moreover, not only does ownership remove the need for a garbage collector in the runtime environment, it frees the programmer of the responsibility to ensure certain safety-related properties~\citep{qin2020understanding}.
-Due to the memory and thread safety guarantees that Rust provides, many software companies have adopted Rust in production.
-For example, Dropbox decided to migrate to Rust for developing their storage system, since they were not completely satisfied with the features provided by AWS’s S3 or Golang~\citep{dropbox}.
-
 The Rust compiler's Borrow-Checker component enforces ownership rules and hence many of Rust's safety properties.
@@
     \item Keeping in mind the key role of the borrow checker in Rust, this work specifically identifies bug fix patterns that are related to Rust's ownership framework in addition to general bug fix patterns for Rust.
+    \item We identified that the most frequently observed general pattern is the addition or removal of struct fields.
+Among the BC-related patterns, we found that the most common pattern is the removal of a clone() call.
 \end{itemize}
+
+\noindent \textbf{Data Availability Statement} The datasets generated and analyzed during the current study are available in the Zenodo repository, \url{https://zenodo.org/record/7388618}.

### 2023-06-04 main.tex
resolved all but 3 of the todos

@@
 After clustering the code changes, and conducting a manual analysis, we discovered 20 groups of cross-project bug fix patterns, which we categorize as (1) general patterns and (2) borrow-checker-related patterns.
-We describe each of these patterns and their implications to automated program repair.\todo{talk about the most popular pattern}
+Among the general patterns, the most frequently observed pattern is when the user either adds or removes one of the struct fields.
+In the case of borrow-checker-related patterns, the most common pattern we encountered is when the user removes a clone() call.
+We describe all the patterns and their implications to automated program repair.
 \keywords{Bug patterns \and Pattern mining \and Bug fix changes \and Rust}

### 2023-06-04 related.tex
add todo about refactoring tools
resolved all but 3 of the todos

@@
 
+\subsection{Refactoring tools for Rust}
+Clippy is a widely known as the best code refactoring tool for Rust that can identify over 600 common mistakes and offer automated fixes.
+It provides users with the flexibility to adjust the linting level, allowing them to control the granularity of the lint suggestions they receive.
+As discussed in Section~\ref{sec:bc_patterns}, we found that Clippy can detect and address five of our borrow-checker-related patterns.
+\cite{sam2017automated} introduced a refactoring tool that uses the Rust compiler infrastructure to perform various refactoring operations.
+These operations encompass renaming variables, arguments, fields, functions, methods, structs, and enumerations, as well as inlining local variables, and reifying and eliding lifetime parameters.
+In a more advanced use case, \cite{ling2022rust} proposed CRustS, a tool designed for refactoring C2Rust output.
+C2Rust is a C to Rust transpiler that converts C code into Rust syntax but retains the unsafe semantics of C.
+By relaxing the constraints of preserving semantics during transformations, CRustS increases the proportion of transformed code that successfully passes the safety checks of the Rust compiler.
+
 \subsection{Code Patterns in Rust}

### 2023-06-04 results.tex
add todo about manual trim
resolved all but 3 of the todos

@@
 \hline
-\multirow{2}{*}{Attributes} & G.attr.struct & 172 & 16                                  \\
-& G.attr.struct.field & 62 & 16                     \\\cline{1-4}
+\multirow{2}{*}{Attributes} & G.attr.struct & 81 & 16                                  \\
+& G.attr.struct.field & 34 & 9                     \\\cline{1-4}
 \multirow{2}{*}{Struct} & G.struct.field  & 179 & 17\\
@@
 \subsubsection{Attributes}
-
 The most common bug fix pattern in our corpus modifies attributes.
@@
 
-The concept of attributes in Rust is similar to Python's decorators\todo{are python decorator changes important in the Python dataset?} and Java's annotations (though Java annotations may be less likely to have runtime effect; for instance, Java @Override adds additional compile-time constraints).
+The concept of attributes in Rust is similar to Python's decorators and Java's annotations (though Java annotations may be less likely to have runtime effect; for instance, Java @Override adds additional compile-time constraints).
 While each language may have its own syntax and specific use cases, the fundamental idea of attaching additional information to code elements is shared across these languages.
+We describe bug patterns that involve modification in Attributes.
+In the case of Python, according to the findings of \cite{yang2022mining}, no bug patterns involving Python's decorators were reported in their study.
 \\
@@
 
-//G.attr.struct.drop: Removing attributes from a struct
+// G.attr.struct.drop: Removing attributes from a struct
 <@\textcolor{red}{--}@> #[derive(Serialize, Deserialize, Clone, PartialEq, Eq, Debug)]
@@
 pub struct Subset {
-    /// Invariant, maintained by `SubsetBuilder`: all `Segment`s have non-zero
-    /// length, and no `Segment` has the same count as the one before it.
     segments: Vec<Segment>,
@@
 There are three variants of this pattern: Adding attributes to the struct (G.attr.struct.add); Removing attributes from the struct (G.attr.struct.drop); Changing the content of an existing attribute (G.attr.struct.change).
-These attributes are Outer Attributes as we can find struct-related non terminals within the ASTDiff of the instances of this pattern.
-
-\todo{discussion on how many of the changes are of each variant: over 90\% are modify, but get the actual number}
+These variants account for 58, 10, and 13 out of 81 datapoints in the cluster, respectively.
+Thus, adding attributes is the most frequent variant.
+Alse, these attributes are Outer Attributes as we can find struct-related non terminals within the ASTDiff of the instances of this pattern.
 
@@
 As with \textit{\textbf{G.attr.struct}}, there are also three subtypes of this change pattern.
+Out of 34 datapoints in this cluster, 32 involve changes to a field attribute, while one datapoint each corresponds to the addition and removal of a field attribute.
 Using these attributes, the developers can control different actions they wish to apply on struct fields, in addition to providing meta-information.
 For instance, in the code snippet provided above, the developers chose to make Serde (a serialization/deserialization tool for Rust) call a different function (\verb+is_empty+ instead of \verb+Option::is_none+) to determine whether to skip serializing the \verb+attributes+ field.
+Note that \textit{\textbf{G.attr.struct.field}} is not a subset of \textit{\textbf{G.attr.struct}}.
 \\
-
-\todo{please confirm: G.attr.struct.field is not a subset of G.attr.struct.}
-\todo{we should also get the count of subtypes}
 
@@
 Starting from the observation that so many of our bug fixes change attributes, we believe that IDEs need to be able to propose relevant attribute changes, as witnessed in our corpus.
-The example in the text about changing Deserialize to ConfigDeserialize is specific to that program, but other changes would apply more generally.
-\todo{which?} Also, we demonstrated a case where the developer removed a \texttt{Debug} attribute.
+The example for \textit{\textbf{G.attr.struct.change}} about changing \verb+Deserialize+ to \verb+ConfigDeserialize+ is specific to that program, but other changes would apply more generally.
+Also, as for \textit{\textbf{G.attr.struct.drop}}, we demonstrated a case where the developer removed a \verb+Debug+ attribute.
 A novel type of performance program repair tool working with a run-time profiler could identify unnecessary code and remove attributes that generate it.

### 2023-06-04 related.tex
grammar

--- 2023-05-17 related.tex
@@
 
+\subsection{Refactoring tools for Rust}
+Clippy is a widely known as the best code refactoring tool for Rust.
+It can identify over 600 common mistakes and offer automated fixes.
+It provides users with the flexibility to adjust the linting level, allowing them to control the number of lint suggestions they receive.
+As discussed in Section~\ref{sec:bc_patterns}, we found that Clippy can detect and address five of our borrow-checker-related patterns.
+\cite{sam2017automated} introduced a refactoring tool that uses the Rust compiler infrastructure to perform various refactoring operations, including renaming variables, arguments, fields, functions, methods, structs, and enumerations, as well as inlining local variables, and reifying and eliding lifetime parameters.
+In a more advanced use case, \cite{ling2022rust} proposed CRustS, a tool designed for refactoring C2Rust output.
+C2Rust is a C to Rust transpiler that converts C code into Rust syntax but retains the unsafe semantics of C.
+By relaxing the constraint of preserving semantics during transformations, CRustS increases the proportion of transformed code that successfully passes the safety checks of the Rust compiler.
+
 \subsection{Code Patterns in Rust}

### 2023-06-04 results.tex
grammar

@@
 We describe bug patterns that involve modification in Attributes.
-In the case of Python, according to the findings of \cite{yang2022mining}, no bug patterns involving Python's decorators were reported in their study.
-\\
+\cite{yang2022mining} found that their Python dataset included no bug patterns involving Python's decorators, in contrast to our findings about Rust changes.
 

### 2023-06-05 discussion.tex
reread for usefulness

@@
 Predicting the future is beyond Clippy's remit.
-To the best of our knowledge, the remaining patterns (3 BC-related patterns and 12 general patterns) cannot be reported by the current linting tools.
+To our knowledge, the remaining patterns (3 BC-related patterns and 12 general patterns) cannot be reported by current linting tools.
 
@@
 Here, we discuss the implications of our most important patterns.
-In summary, program repair tools must understand semantics of attributes and should be performance-aware.
-It may also be useful for tools to be able to automatically infer immutability.
+\emph{In summary, program repair tools must understand semantics of attributes and should be performance-aware.} It may also be useful for tools to be able to automatically infer immutability.
 
-The two most common patterns were general patterns modifying struct fields and attributes, with 179 and 172 instances respectively.
+\todo{is the most common ordering still the same after trimming?}
+The two most common patterns were general patterns modifying struct fields and attributes, with 179 and 172 \todo{are those numbers stale?} instances respectively.
 Unfortunately, struct field modifications are neither interesting nor actionable; they are simply used to enable other changes.
@@
 
+\todo{recap the italicized stuff from above}
+\vspace*{1em}
+\fbox{\parbox{0.9\textwidth}{We summarize learnings from our bug patterns as follows:
+\begin{itemize}
+\item one
+\item two
+\end{itemize}
+}}
+
 \subsection{Usefulness of Rust Bug Patterns}
@@
 
-For instance, as we have discussed, for the common pattern of \textit{Modifying the attributes of structs}, a code linter tool could search for correct attribute lists within the codebase~\citep{forrest2009genetic}, leveraging the assumption that proper attribute usage can be found elsewhere.
+For instance, as we have discussed, for the common pattern of \textit{Modifying the attributes of structs}, a code linter tool could search for correct attribute lists within the codebase~\citep{forrest2009genetic}, leveraging the assumption that proper attribute usage can be found elsewhere in the code.
 Also, the linter can analyze the usages of the struct to determine the necessity of adding the attribute.

### 2023-06-05 introduction.tex
reread for usefulness

--- 2022-11-23 introduction.tex
@@
+Rust is a relatively new programming language~\citep{klabnik2019rust} which allows programmers to write programs that have low-level control over resources, while still ensuring high-level safety guarantees (for programs written in safe Rust).
+What distinguishes Rust from other systems programming languages (e.g.
+C/C++) is its ownership framework, which is an integral part of Rust.
+The ownership framework keeps track of references to different memory locations, and will free them once they go out of scope.
+If a program passes the compiler’s checks, it meets several safety properties by construction; e.g.
+it has no dangling references or double-frees.
+Such properties had to be manually ensured (or not!) in languages like C/C++.
+Moreover, not only does ownership remove the need for a garbage collector in the runtime environment, it frees the programmer from the responsibility to ensure certain safety-related properties~\citep{qin2020understanding}.
+Due to the memory and thread safety guarantees that Rust provides, many software companies have adopted Rust in production.
+For example, Dropbox decided to migrate to Rust for developing their storage system, since they were not completely satisfied with the features provided by AWS’s S3 or Golang~\citep{dropbox}.
+
 Bugs have long been an unavoidable part of computer programming, predating Grace Hopper's literal debugging session in 1947; even today, bug-free software remains an aspirational aim for software engineering research.
@@
 Yet, to the best of our knowledge, no research has sought to identify the most pervasive bug fix patterns in Rust programs.
-In this project, we introduce Ruxanne, a tool for analyzing and extracting fix patterns in Rust.
+
+In this work, we introduce Ruxanne, a tool for analyzing and extracting fix patterns in Rust.
 Ruxanne uses a novel embedding of Rust code into fixed-sized vectors.
@@
 After clustering the code changes, and conducting a manual analysis, we discovered 20 groups of cross-project bug fix patterns, which we categorize as (1) general patterns and (2) borrow-checker-related patterns.
-We describe each of these patterns and their implications to automated program repair.
+We describe each of these patterns.
+The most common general pattern is addition or removal of struct fields, while the most common borrow-checker-related bug fix pattern is removing a call to the clone() function.
+Our patterns can serve as knowledge to be embedded into the design of novel Rust automated program repair tools; we present some specific potential applications of these patterns below.
 
-If we observe undesired behaviour from our program, then there is likely a fault within our program\footnote{At the extreme, a fault could be caused by moths or other hardware malfunctions.}.
+If we observe undesired behaviour from a program, then there is likely a fault within the program\footnote{At the extreme, a fault could be caused by moths or other hardware malfunctions.}.
 But where is the fault?
@@
 Research such as ours can help formulate data-driven recommendations that reduce the number of patterns empirically linked to bugs.
-Per Monperrus, the fewer frequent bug patterns we keep in our software, the lower cost we can expect to pay for software maintenance~\citep{monperrus2014critical}.
+Monperrus espouses the claim that the fewer frequent bug patterns we keep in our software, the lower cost we can expect to pay for software maintenance~\citep{monperrus2014critical}.
 
@@
     \item Keeping in mind the key role of the borrow checker in Rust, this work specifically identifies bug fix patterns that are related to Rust's ownership framework in addition to general bug fix patterns for Rust.
+    \item This work finds that the most frequently observed general pattern is the addition or removal of struct fields and that the most frequently observed BC-related pattern is the removal of a clone() call.
 \end{itemize}
+
+\noindent \textbf{Data Availability Statement} The datasets generated and analyzed during the current study are available in the Zenodo repository, \url{https://zenodo.org/record/7388618}.

### 2023-06-05 main.tex
grammar
reread for usefulness

@@
 
-Yet, to the best of our knowledge, no research has sought to identify the most pervasive bug fix patterns within Rust programs.
+Yet, to our knowledge, no research has sought to identify the most pervasive bug fix patterns within Rust programs.
 In this project, we introduce Ruxanne, a tool for analyzing and extracting fix patterns in Rust.
@@
 After clustering the code changes, and conducting a manual analysis, we discovered 20 groups of cross-project bug fix patterns, which we categorize as (1) general patterns and (2) borrow-checker-related patterns.
-Among the general patterns, the most frequently observed pattern is when the user either adds or removes one of the struct fields.
+Among the general patterns, the most frequently observed pattern is when the user either adds or removes struct fields.
 In the case of borrow-checker-related patterns, the most common pattern we encountered is when the user removes a clone() call.
-We describe all the patterns and their implications to automated program repair.
+We describe all detected patterns and their implications to automated program repair.
 \keywords{Bug patterns \and Pattern mining \and Bug fix changes \and Rust}

### 2023-06-05 results.tex
reread for usefulness

--- 2023-06-04 results.tex
@@
 We collected general fixes, asked DBSCAN to create clusters, and manually analyzed them.
-In this subsection, we will discuss each pattern, as well as its actionability---possible usages of the pattern for bug repair.
+In this subsection, we will discuss each pattern, as well as one aspect of its value---its actionability.
+By actionability, we mean the possible usages of the pattern for bug repair.
+Other researchers can also use the empirical pattern frequency numbers that we report to build their own program repair tools, prioritizing more-frequently-repaired patterns; they are not limited to our suggested actionability.
 Section~\ref{subsec:patterns} contains a global discussion of our patterns and their value.
 
-Specifically, our dataset includes 60583 datapoints for general fix patterns.
+Our dataset includes 60583 datapoints for general fix patterns.
 We processed our dataset on a server with a Xeon Gold 5120 processor (Q3-2017, 14-core 2.2 to 3.2GHz).
@@
 We used DBSCAN with parameters $\epsilon=0.0001$ and $Z=5$, resulting in 577 clusters.
-After manual analysis, as described in~\ref{sec:manual_analysis_cluster_selection}, we ended up with 12 general patterns, which we describe in Table~\ref{table:general}.
+After manual analysis, as described in Section~\ref{sec:manual_analysis_cluster_selection}, we ended up with 12 general patterns, which we describe in Table~\ref{table:general}.
 As shown in the table, we have separated the patterns based on the underlying program element (Category column).
@@
 Since these patterns are general Rust patterns, analogous patterns can also occur in other programming languages, and we will discuss such analogous patterns when describing each pattern.
-We will also discuss the actionability of some of the patterns.
-Moreover, other researchers can use the empirical pattern frequency numbers that we report to build their own program repair tools, prioritizing more-frequently-repaired patterns; they are not limited to our suggested actionability.
 We next provide a detailed description of each of these patterns.
@@
 In Rust, attributes enable developers to add metadata to program elements.
-Some of the attributes---presumably all of the ones involved in bug fixes---affect the semantics of the program; for instance, libraries can generate different code based on the attributes that the developer specifies.
+Some of the attributes---presumably the vast majority of those involved in bug fixes---affect the semantics of the program; for instance, libraries generate different code based on the attributes that the developer specifies.
 
 Developers can add attributes to structs, struct fields, enum variants, match expression arms, etc.
-There are two types of attributes\todo{how important is inner versus outer?}: Inner Attributes and Outer Attributes.
+%I don't think we're well served to talk about inner vs outer attributes, not relevant to readers
+%There are two types of attributes\todo{how important is inner versus outer?}: Inner Attributes and Outer Attributes.
 Inner Attributes apply to the program element encapsulating them, while Outer Attributes apply to the program element coming after the attribute.
-
-The concept of attributes in Rust is similar to Python's decorators and Java's annotations (though Java annotations may be less likely to have runtime effect; for instance, Java @Override adds additional compile-time constraints).
-While each language may have its own syntax and specific use cases, the fundamental idea of attaching additional information to code elements is shared across these languages.
-We describe bug patterns that involve modification in Attributes.
-In the case of Python, according to the findings of \cite{yang2022mining}, no bug patterns involving Python's decorators were reported in their study.
-\\
+The concept of attributes in Rust is similar to Python's decorators and Java's annotations.
+While each language may have its own syntax and specific use cases, the fundamental idea of attaching metadata to code elements is shared across these languages.
+Some attributes, decorators, and annotations have no run-time effect and thus could not cause bugs (if we accept a purely operational definition of ``bug'').
+\cite{yang2022mining}~reported no bug patterns involving Python's decorators in their Python dataset.
+As we alluded to above, however, Rust has many attributes that definitely have run-time effects.
+We will see some of these attributes in the following examples.
 
@@
 Thus, adding attributes is the most frequent variant.
-Alse, these attributes are Outer Attributes as we can find struct-related non terminals within the ASTDiff of the instances of this pattern.
+%Also, these attributes are Outer Attributes as we can find struct-related non terminals within the ASTDiff of the instances of this pattern.
 
 In the specific cases above, the Add change requests that additional boilerplate code be generated to evaluate partial equivalence on \texttt{CounterProps} (new functionality); the Remove change causes debug code to be removed (performance fix); and the Change change requests different deserialization code be generated.
+In our example, adding an attribute causes additional code to be generated for a struct and could thus, for example, fix a compile-time error in a dependency.
 \\
@@
 As with \textit{\textbf{G.attr.struct}}, there are also three subtypes of this change pattern.
-Out of 34 datapoints in this cluster, 32 involve changes to a field attribute, while one datapoint each corresponds to the addition and removal of a field attribute.
+Out of 34 datapoints in this cluster, 32 involve changes to a field attribute, while two datapoints correspond to the addition and removal of a field attribute (one for addition, one for removal).
 Using these attributes, the developers can control different actions they wish to apply on struct fields, in addition to providing meta-information.
@@
 
-\noindent\textbf{Actionability:} Because attributes have semantic effects, program repair tools can effect repairs by adjusting the set of attributes included.
-Starting from the observation that so many of our bug fixes change attributes, we believe that IDEs need to be able to propose relevant attribute changes, as witnessed in our corpus.
+\noindent\textbf{Actionability of attribute change bug fixes:} Because attributes have semantic effects, program repair tools can effect repairs by adjusting the set of attributes included.
+Given the observation that so many of our bug fixes change attributes, we propose that IDEs need to be able to propose relevant attribute changes.
 The example for \textit{\textbf{G.attr.struct.change}} about changing \verb+Deserialize+ to \verb+ConfigDeserialize+ is specific to that program, but other changes would apply more generally.
-Also, as for \textit{\textbf{G.attr.struct.drop}}, we demonstrated a case where the developer removed a \verb+Debug+ attribute.
+Also, for \textit{\textbf{G.attr.struct.drop}}, we demonstrated a case where the developer removed a \verb+Debug+ attribute.
 A novel type of performance program repair tool working with a run-time profiler could identify unnecessary code and remove attributes that generate it.
@@
 
-Like in other languages, such as C, a struct makes it possible for users to define a custom type that is comprised of different types.
-The patterns that we describe in this section can also be seen in other languages that use the same concept.
+A struct enables users to define a custom type that is comprised of different types.
+This is a common construct across languages, and the patterns that we describe in this section also exist in other languages.
 \\
@@
 \noindent\textbf{Description:} Adding \texttt{pub} to a struct field makes it possible to access that field from external modules.
-In the instances of this cluster, adding the \texttt{pub} access modifier happened in bug fixing changes, where the developer needed to access a field which had not been marked for public access.
+In the instances of this cluster, adding the \texttt{pub} access modifier happened in bug fixing changes where the developer needed to access a field which had not been marked for public access.
 On the other hand, public access might be revoked if the developer realizes there are no external accesses to the field, and that such accesses are not desirable, e.g.
@@
 
-\noindent\textbf{Actionability:} The patterns related to access modifiers suggest an opportunity for repair tools to make sure that the accessability of program entities follows the user's specifications, enhancing encapsulation, security, and maintainability.
+\noindent\textbf{Actionability of struct changes:} The patterns related to access modifiers suggest an opportunity for repair tools to make sure that the accessibility of program entities follows the user's specifications, enhancing encapsulation, security, and maintainability.
 Additionally, the empirical numbers revealing the frequency of struct field modifications in Rust shed light on the significance of these bug fix patterns and emphasize the need for further exploration on how program repair tools can effectively modify code based on changes in struct fields to repair program behavior.
@@
 If a developer wants a variable to contain the equivalent of null, they would have to use the \texttt{Option} type.
-\texttt{Option} is essentially an \texttt{enum} with two variants, \texttt{Some}, which indicates that a variable has a value, and \texttt{None}, which indicates the absence of any value.
+\texttt{Option} is essentially an \texttt{enum} with two variants: \texttt{Some}, which indicates that a variable has a value; and \texttt{None}, which indicates the absence of any value.
 In this change pattern, the developer changes the type of a struct field to an \texttt{Option} of that type, hence enabling them to store no value in that field.
@@
 
-\noindent\textbf{Actionability:} When the user's specification requires a variable to be able to assume \texttt{None} values, these patterns indicate the need to modify the program accordingly.
+\noindent\textbf{Actionability of adding Option to a type:} When the user's specification requires a variable to be able to assume \texttt{None} values, these patterns indicate the need to modify the program accordingly.
 A program repair\todo{is this really a repair tool or a refactoring tool?
@@
 
-Rust is a statically-typed language, and like in the other statically-typed languages, some change patterns are associated with changes in types used in different program structures.
+Rust is a statically-typed language, and like in other statically-typed languages, some change patterns are associated with changes in types used in different program structures.
 Here, we introduce four bug fix patterns that are associated with types in Rust:
@@
 After manual analysis (Section~\ref{sec:manual_analysis_cluster_selection}), we ended up with 8 borrow-checker related fix patterns.
-Similar to the previous table, we describe BC-related patterns in Table~\ref{table:bc}.
+Table~\ref{table:bc} describes BC-related patterns, analogously to Table~\ref{table:general} for general patterns.
 Since the Borrow-Checker is unique to Rust, patterns in this section generally do not have direct counterparts in other languages.
@@
 
-\noindent\textbf{Actionability:} Like Clippy, a program repair tool can implement clone removal by reasoning about whether the clone operation was necessary.
+\noindent\textbf{Actionability of clone removal:} Like Clippy, a program repair tool can implement clone removal by reasoning about whether the clone operation was necessary.
 If the clone is unnecessary, the tool can replace it with a reference (as seen above), thus reducing the overhead of unnecessary copying.
@@
 
-\noindent\textbf{Actionability:} This pattern often occurs as part of a larger change, making it challenging to create a generalized repair tool that covers all possible scenarios.
+\noindent\textbf{Actionability of adding borrowing:} This pattern often occurs as part of a larger change, making it challenging to create a generalized repair tool that covers all possible scenarios.
 Typically, when a reference is passed instead of transferring ownership, the variable is used again after returning from the call.
-Although creating a comprehensive repair tool for such cases is complex, empirical results underscore the significance and prevalence of this fix, warranting further investigation.
+Although it is not obvious how to create a comprehensive repair tool for such cases, empirical results underscore the significance and prevalence of this fix, warranting further investigation.
 \\
 
-\noindent\textit{\textbf{BC.deref.to\_string().add} Adding a dereference operation before calling to\_string()}
+\noindent\textit{\textbf{BC.deref.to\_string().add} Adding a dereference before calling to\_string()}
 
@@
 
-\noindent\textbf{Description:} This is a change that in all instances (in our benchmarks) has been proposed by Clippy---commit messages indicated the use of Clippy for this fix.
+\noindent\textbf{Description:} This is a performance-related \todo{is it only perf?} change that in all instances (in our benchmarks) has been proposed by Clippy---commit messages indicated the use of Clippy for this fix.
 The change applies on an object prior to calling \verb+to_string()+ on that object.
@@
 
-\noindent\textbf{Actionability:} A repair tool can use this pattern to introduce mutability to the target variable and modify other parts of the program accordingly.
-For instance, if the target variable has associated \texttt{impl} blocks, the change may involve modifying method annotations to ensure that the calling object is mutable, aligning with the desired specifications.\todo{how would the specs be provided?
-how could this be done as part of repair vs as part of refactoring?!} \\
+\noindent\textbf{Actionability of adding mutability:} A repair tool can use this pattern to introduce mutability to the target variable and modify other parts of the program accordingly.
+For instance, if the target variable has associated \texttt{impl} blocks, the change may involve modifying method annotations to ensure that the calling object is mutable, aligning with the desired behaviour.\\
+%specifications.\todo{how would the specs be provided? how could this be done as part of repair vs as part of refactoring?!} \\
 
@@
 This redundant mutability is reported by Clippy and can also be removed by it.
-This change may help future-proof the code against unintended future changes.
+Such a change may help future-proof the code against unintended future changes.
 

### 2023-06-06 discussion.tex
resolved todos except one

@@
 \emph{In summary, program repair tools must understand semantics of attributes and should be performance-aware.} It may also be useful for tools to be able to automatically infer immutability.
+Here, we discuss the top ten most popular patterns, 8 of which fall into the category of general patterns, and the remaining 2 are BC-related patterns.
 
-\todo{is the most common ordering still the same after trimming?}
-The two most common patterns were general patterns modifying struct fields and attributes, with 179 and 172 \todo{are those numbers stale?} instances respectively.
+The most common general pattern and the most popular pattern overall is modifying struct fields with 179 datapoints.
 Unfortunately, struct field modifications are neither interesting nor actionable; they are simply used to enable other changes.
-Attribute modifications are far more interesting, and account for the second- and fifth-most popular patterns.
+
+The most common borrow-checker-related pattern and second-most-common pattern overall is dropping \texttt{clone()}.
+This pattern is performance-related; when correctly applied, it has no semantic impact on the program, but reduces resource consumption.
+Rust is used in situations where performance is important; the number of instances of this pattern confirms that developers do work on performance issues.
+This implies that \emph{program repair tools would benefit from being performance-aware}, rather than limiting themselves strictly to functional properties of the code.
+
+Attribute modifications are interesting changes that account for the third- and ninth-most popular patterns.
 From a language design perspective, the fact that there are so many attribute changes marked as bug fixes illustrates the semantic importance of attributes in Rust, and \emph{implies that program repair must understand the effects of attributes}.
@@
 
-The most common borrow-checker-related pattern and third-most-common pattern overall is dropping \texttt{clone()}.
-This pattern is performance-related; when correctly applied, it has no semantic impact on the program, but reduces resource consumption.
-Rust is used in situations where performance is important; the number of instances of this pattern confirms that developers do work on performance issues.
-This implies that \emph{program repair tools would benefit from being performance-aware}, rather than limiting themselves strictly to functional properties of the code.
-
-The fourth- and sixth-most popular patterns, changing the type of a struct field or of a generic type parameter, are difficult to make general statements about; there are many unrelated idiosyncratic changes in these clusters.
+The fourth- and fifth-most popular patterns, changing the type of a struct field or of a generic type parameter, are difficult to make general statements about; there are many unrelated idiosyncratic changes in these clusters.
 Such changes look the same, but have different purposes.
@@
 
-The seventh-most-popular pattern is the Rust-specific change of dropping mutability (i.e.
+The sixth-most-popular pattern is the Rust-specific change of dropping mutability (i.e.
 making data immutable).
@@
 
-We are unable to say anything meaningful about the eighth- and ninth-most popular patterns, \texttt{G.match.pattern} and \texttt{G.type.enum.variant}.
+We are unable to say anything meaningful about the seventh- and eighth-most popular patterns, \texttt{G.match.pattern} and \texttt{G.type.enum.variant}.
 They do not seem particularly helpful for automatic program repair.
@@
 \begin{itemize}
-\item one
-\item two
+\item Being performance-awareness would be advantageous for program repair tools.
+\item Understanding the effects of attributes can guide program repair.
+\item Tools for inferring Rust immutability could be viable.
 \end{itemize}

### 2023-06-06 methodology.tex
reread for usefulness
resolved todos except one

@@
-In this section, we describe our methodology, providing full details in later subsections.
+In this section, we provide an overview of our methodology, providing full details in later subsections.
 Ruxanne uses a pipeline for finding classes of pervasive patterns in Rust programs.
@@
 % 3 types, refactoring dropped
-
 Like \cite{yang2022mining}, and following the definitions of \cite{cotroneo2019analyzing}, we divide the clusters in three different groups:
@@
 If we judged that the members of a cluster fit into a pattern, we wrote a natural language description of that cluster.
-Considering the descriptions, we linked similar clusters together as possible candidates for merging.
+Considering the descriptions, we linked similar clusters together as possible candidates for merging, and removed the datapoints that did not fit into the first two groups listed above.
 We considered these candidates and carried out appropriate ones.

### 2023-06-06 results.tex
resolved todos except one

@@
 
-\noindent\textbf{Description:} This is a performance-related \todo{is it only perf?} change that in all instances (in our benchmarks) has been proposed by Clippy---commit messages indicated the use of Clippy for this fix.
+\noindent\textbf{Description:} This is a change that in all instances (in our benchmarks) has been proposed by Clippy---commit messages indicated the use of Clippy for this fix.
 The change applies on an object prior to calling \verb+to_string()+ on that object.
@@
 Adding the dereference makes the compiler use the specialized implementation of \verb+to_string()+ and not go through slower string formatting methods\footnote{\url{https://rust-lang.github.io/rust-clippy/master/index.html\#inefficient\_to\_string}}.
+This change is aimed at improving performance, as the functionality remains unchanged.
 

### 2023-06-06 results.tex
resolved the one remaining todo

@@
 \noindent\textbf{Actionability of adding Option to a type:} When the user's specification requires a variable to be able to assume \texttt{None} values, these patterns indicate the need to modify the program accordingly.
-A program repair\todo{is this really a repair tool or a refactoring tool?
-I feel like many of our changes are for refactorings not repairs, despite our PR} tool can leverage these patterns to automatically update the variable's type to \texttt{Option<T>}, ensuring that it can now hold \texttt{None} as a valid value.
-Furthermore, such a tool should also systematically update other occurrences of that variable throughout the program to account for the change and ensure program correctness.
+A program repair tool can leverage these patterns to automatically update the variable's type to \texttt{Option<T>}, ensuring that it can now hold \texttt{None} as a valid value.
+This pattern is a part of a \textit{fix-induced} change, thus, a repair tool should also systematically update other occurrences of that variable throughout the program to account for the change and ensure program correctness.
 

### 2023-06-06 discussion.tex
wording

@@
 
-\todo{recap the italicized stuff from above}
 \vspace*{1em}
@@
 \begin{itemize}
-\item Being performance-awareness would be advantageous for program repair tools.
-\item Understanding the effects of attributes can guide program repair.
+\item Rust program repair tools must understand the effects of attributes.
+\item Program repair tools should be performance-aware.
 \item Tools for inferring Rust immutability could be viable.

### 2023-06-06 methodology.tex
reedit

@@
 If we judged that the members of a cluster fit into a pattern, we wrote a natural language description of that cluster.
-Considering the descriptions, we linked similar clusters together as possible candidates for merging, and removed the datapoints that did not fit into the first two groups listed above.
+Considering the descriptions, we linked similar clusters together as possible candidates for merging, and removed the datapoints that were not either bug-fix or fix-inducing.
 We considered these candidates and carried out appropriate ones.

### 2023-06-06 results.tex
reedit

@@
 
-\noindent\textbf{Actionability of adding Option to a type:} When the user's specification requires a variable to be able to assume \texttt{None} values, these patterns indicate the need to modify the program accordingly.
-A program repair tool can leverage these patterns to automatically update the variable's type to \texttt{Option<T>}, ensuring that it can now hold \texttt{None} as a valid value.
-This pattern is a part of a \textit{fix-induced} change, thus, a repair tool should also systematically update other occurrences of that variable throughout the program to account for the change and ensure program correctness.
+\noindent\textbf{Actionability of adding Option to a type:} This pattern occurs as a \textit{fix-induced} change---a repair tool might be carrying out a broader change and need to make smaller changes like this one in the process.
+Our empirical results suggest that this type of change does occur in practice and will need to be part of a program repair tool's arsenal.
 

### 2023-06-07 results.tex
minor

@@
 As discussed in the previous section, we used an embedding mechanism to incorporate AST information in fixed-sized datapoints which then could be used in the DBSCAN clustering algorithm.
-We used heuristics (in the form of a weighting scheme) while creating our datapoints to make the program elements representing the most important non terminals stand out from the complete set of non terminals seen in the AST.
-In addition to that, we also incorporated the BC-related elements in our datapoints.
+We used heuristics (i.e.
+a weighting scheme) while creating our datapoints to make the program elements representing the most important non terminals stand out from the complete set of non terminals seen in the AST.
+In addition, we also incorporated BC-related elements in our datapoints.
 Here we evaluate the effectiveness of our embedding.
@@
 A value of 1 indicates that the human analyzer concurs with the algorithm, while 2 means that the human-identified most important element was ranked 2 by the algorithm.
-After collecting all the ranks, then we calculate Top-$n$ Accuracy for $n=1$ up to $n=5$.
+After collecting all the ranks, we then calculate Top-$n$ Accuracy for $n=1$ up to $n=5$.
 Both authors acted as analyzer and carried out the experiment independently.

### 2023-06-08 methodology.tex
two types

@@
 A change pattern is a set of code changes that serves a specific purpose.
-Purposes include rectifying program behaviour with respect to a specific functional or non-functional requirement (bug-fixing patterns) as well as improving code readability or maintainability (refactoring patterns).
-In this work, we are interested in bug-fixing patterns.
-Thus, from now on, whenever we refer to code patterns, we mean bug-fixing patterns.
+Purposes include rectifying program behaviour with respect to a specific functional or non-functional requirement (bug-fixing patterns) and changes that support bug-fixing patterns (fix-induced patterns) as well as improving code readability or maintainability (refactoring patterns).
+In this work, we are interested in bug-fixing and fix-induced patterns.
+Thus, from now on, whenever we refer to code patterns, we mean bug-fixing and fix-induced patterns.
 

### 2023-06-13 methodology.tex
more explicit

@@
 
-Figure~\ref{fig:extraction} illustrates path extraction on a simple change.
-The change replaces \verb+tail+ with \verb+&tail+ and \verb+head+ with \verb+head.clone()+.
-That is, after the change, the second argument passes a borrow of \verb+tail+ rather than sending the ownership of \verb+tail+ to the callee; and, the third argument sends a clone of head to the callee instead of sending head itself.
+\paragraph{Example: path extraction.} Figure~\ref{fig:extraction} illustrates path extraction on a simple change, starting from the ASTDiff.
+The change (shown in the top right corner of the Figure) replaces \verb+tail+ with \verb+&tail+ and \verb+head+ with \verb+head.clone()+.
+After the change, the second argument passes a borrow of \verb+tail+ rather than sending the ownership of \verb+tail+ to the callee; and, the third argument sends a clone of \verb+head+ to the callee instead of sending \verb+head+ itself.
 A DFS on this tree yields three different paths, which we show in red, green, and purple.
 A path represents a sequence of involved non terminals.
-It is crucial to record the order of nodes in the path, so that we could record that order in our datapoint embedding.
+It is crucial to record the order of nodes in the path, so that we can record that order in our datapoint embedding.
+
 Note that the tree depicted in Figure~\ref{fig:extraction} is not an AST; rather, it represents an ASTDiff.
@@
 However, to account for order-sensitivity, we introduce two additional steps.
-Firstly, we record the number of occurrences of items from $\mathtt{NT} \cup \mathtt{BCE}$ within each root-level scope.
+Firstly, we record the number of occurrences of items from $\mathtt{NT} \cup \mathtt{BCE}$ within each root-level scope (e.g.
+structs, functions, impl blocks, etc).
 This captures the varying occurrence counts resulting from different element orders in the program.
 Secondly, we incorporate a weighting scheme, designed semi-automatically, to prioritize non-terminals that we consider more salient for Rust bugs.
-Section~\ref{sec:weighting_scheme}, immediately below,
-illustrates this approach on an example.
+
+\paragraph{Example: computing fixed-sized vectors.} Figure~\ref{fig:essence} depicts the fixed-sized vectors, as described in this subsection, for the change from Figure~\ref{fig:extraction}.
+It also depicts the essence of change, to be described immediately below.
 
@@
 In this vector, the value for each item shows its importance in the change.
-Figure~\ref{fig:essence} shows the process of computing the essence of the code change that we saw in Figure~\ref{fig:extraction}.
-We have visualized this essence vector using a circle pack figure.
+
+\paragraph{Example: essence of change.} Figure~\ref{fig:essence} shows a complete example for computing the essence of change.
+The essence of change vector essentially serves as our embedding vector.
+The Figure shows the process for transforming the patch illustrated in Figure~\ref{fig:extraction} from fixed-size datapoints (computed as described in the previous subsection) into the essence of change vector.
+To better illustrate the essence vector, Figure~\ref{fig:essence} also graphically presents it using a circle pack figure.
 The radius of each circle corresponds to the importance of the item inside it.
-The essence of change vector essentially serves as our embedding vector, transforming the patch illustrated in Figure~\ref{fig:extraction} into a fixed-sized vector as depicted in Figure~\ref{fig:essence}.
 

### 2023-06-13 results.tex
more explicit

@@
 In this change, a dereference operator is added to that object.
-The reason behind the change is that the object is a reference type of \verb+T+, while \verb+T+ directly implements \verb+to_string()+.
-Adding the dereference makes the compiler use the specialized implementation of \verb+to_string()+ and not go through slower string formatting methods\footnote{\url{https://rust-lang.github.io/rust-clippy/master/index.html\#inefficient\_to\_string}}.
-This change is aimed at improving performance, as the functionality remains unchanged.
+The reason behind the change is that the object is a reference type of \verb+T+, where \verb+T+ directly implements \verb+to_string()+.
+Adding the dereference makes the compiler instead use the specialized implementation of \verb+to_string()+ and not go through slower string formatting methods\footnote{\url{https://rust-lang.github.io/rust-clippy/master/index.html\#inefficient\_to\_string}}.
+This change is aimed at improving performance; the functionality remains unchanged.
 

### 2023-06-14 methodology.tex
Path --> Ident in fig
why 3 paths but 2 modified args, explanation added

@@
     \draw (13.5,-5.5) node[fill=purple!20](g4){ExprMethodCall};
-    \draw (13.5,-6.5) node[fill=purple!20](g5){Path};
+    \draw (13.5,-6.5) node[fill=purple!20](g5){Ident};
     \draw (13.5,-7.5) node[fill=purple!20](g6){clone};
@@
 Instead, it is determined by the structural differences between the two ASTs of the code versions.
+For example, in Figure~\ref{fig:extraction} we add a \verb+&+ borrow operator to the variable \verb+tail+ and it introduces a new parent node (\verb+ExprReference+) for \verb+tail+.
+We also add a \verb+clone()+ method call, which modifies the \verb+head+ node and also adds a new sibling node for \verb+head+ to represent the \verb+clone()+.
+Among the terminals, we only keep \verb+clone+ in our paths, as it is a BC-related keyword.
+This is why we observe three paths in the ASTDiff, even though only two arguments have been modified.
 The paths in the ASTDiff capture the changes in the code structure between the two versions, allowing us to analyze and extract relevant information for further processing.

### 2023-06-14 methodology.tex
reorg and explain more about why 3

--- 2023-06-13 methodology.tex
@@
 Note that the tree depicted in Figure~\ref{fig:extraction} is not an AST; rather, it represents an ASTDiff.
+The paths in the ASTDiff capture the changes in the code structure between the two versions, allowing us to analyze and extract relevant information for further processing.
 The number of paths in the ASTDiff does not directly correspond to the number of changed elements in the source code.
 Instead, it is determined by the structural differences between the two ASTs of the code versions.
-The paths in the ASTDiff capture the changes in the code structure between the two versions, allowing us to analyze and extract relevant information for further processing.
+
+In our example, we add a \verb+&+ borrow operator to the variable \verb+tail+, which gives rise to an ASTDiff path introducing a new parent node (\verb+ExprReference+) for \verb+tail+.
+We also add a \verb+clone()+ method call.
+This addition gives rise to two ASTDiff paths: one indicating that the \verb+head+ node is touched and one indicating the new sibling node for \verb+head+ representing the \verb+clone()+.
+(Among the terminals, we only keep \verb+clone+ in our paths, as it is a BC-related keyword.) One path for the \verb+&+ borrow operator and two for the call to \verb+clone()+ result in three paths in the ASTDiff, even though only two arguments have been modified.
 

### 2023-06-14 methodology.tex
even more explicit paths

--- 2023-06-13 methodology.tex
@@
 Note that the tree depicted in Figure~\ref{fig:extraction} is not an AST; rather, it represents an ASTDiff.
+The paths in the ASTDiff capture the changes in the code structure between the two versions, allowing us to analyze and extract relevant information for further processing.
 The number of paths in the ASTDiff does not directly correspond to the number of changed elements in the source code.
 Instead, it is determined by the structural differences between the two ASTs of the code versions.
-The paths in the ASTDiff capture the changes in the code structure between the two versions, allowing us to analyze and extract relevant information for further processing.
+
+In our example, we add a \verb+&+ borrow operator to the variable \verb+tail+, which gives rise to an ASTDiff path introducing a new parent node (\verb+ExprReference+) for \verb+tail+.
+We also add a \verb+clone()+ method call.
+This addition gives rise to two ASTDiff paths: one indicating that the \verb+head+ node is now part of a \verb+ExprMethodCall+, and one indicating the new sibling node for \verb+head+ representing the \verb+clone()+.
+(Among the terminals, we only keep \verb+clone+ in our paths, as it is a BC-related keyword.) One path for the \verb+&+ borrow operator and two for the call to \verb+clone()+ result in three paths in the ASTDiff, even if it only looks like we had made two modifications.
 

### 2023-06-14 methodology.tex
bad break

@@
 In our example, we add a \verb+&+ borrow operator to the variable \verb+tail+, which gives rise to an ASTDiff path introducing a new parent node (\verb+ExprReference+) for \verb+tail+.
-We also add a \verb+clone()+ method call.
-This addition gives rise to two ASTDiff paths: one indicating that the \verb+head+ node is now part of a \verb+ExprMethodCall+, and one indicating the new sibling node for \verb+head+ representing the \verb+clone()+.
+We also add a call to \verb+clone()+.
+This addition gives rise to two ASTDiff paths: one showing that the \verb+head+ node now belongs to an \verb+ExprMethodCall+, and one indicating the new sibling node for \verb+head+ representing the \verb+clone()+.
 (Among the terminals, we only keep \verb+clone+ in our paths, as it is a BC-related keyword.) One path for the \verb+&+ borrow operator and two for the call to \verb+clone()+ result in three paths in the ASTDiff, even if it only looks like we had made two modifications.

### 2023-06-14 methodology.tex
root level scopes in picture

@@
 However, to account for order-sensitivity, we introduce two additional steps.
-Firstly, we record the number of occurrences of items from $\mathtt{NT} \cup \mathtt{BCE}$ within each root-level scope (e.g.
+Firstly, we record the number of occurrences of items from $\mathtt{NT} \cup \mathtt{BCE}$ within each distinct root-level scope (e.g.
 structs, functions, impl blocks, etc).
@@
 \paragraph{Example: computing fixed-sized vectors.} Figure~\ref{fig:essence} depicts the fixed-sized vectors, as described in this subsection, for the change from Figure~\ref{fig:extraction}.
+We have omitted scope information from this Figure for space reasons; the relevant scope is \verb+ItemFn+, indicating that these non-terminals occur within a function.
 It also depicts the essence of change, to be described immediately below.
@@
 
-    \caption{\label{fig:essence}Multiplying the number of occurrences of items observed in the tree by their respective weights results in the essence of change, visualized with a circle pack figure at bottom right.}
+    \caption{\label{fig:essence}Multiplying the number of occurrences of items observed in the tree by their respective weights results in the essence of change, visualized with a circle pack figure at bottom right.
+Scope information (ItemFn) omitted from column headings for space purposes.}
 \end{figure}

### 2023-06-14 methodology.tex
reword scope

@@
 \paragraph{Example: computing fixed-sized vectors.} Figure~\ref{fig:essence} depicts the fixed-sized vectors, as described in this subsection, for the change from Figure~\ref{fig:extraction}.
-We have omitted scope information from this Figure for space reasons; the relevant scope is \verb+ItemFn+, indicating that these non-terminals occur within a function.
-It also depicts the essence of change, to be described immediately below.
+All of the columns in this vector belong to scope \verb+ItemFn+---that is, these non-terminals occur within a function.
+The Figure also depicts the essence of change, to be described immediately below.
 
@@
     \caption{\label{fig:essence}Multiplying the number of occurrences of items observed in the tree by their respective weights results in the essence of change, visualized with a circle pack figure at bottom right.
-Scope information (ItemFn) omitted from column headings for space purposes.}
+All non terminals here occur within scope {\texttt ItemFn}.}
 \end{figure}

### 2023-06-14 introduction.tex
capitalization

@@
 Ruxanne uses a novel embedding of Rust code into fixed-sized vectors.
-Through Ruxanne, we mined the top 18 most-starred Rust projects in Github to discover the most common bug fix patterns committed to their repositories.
+Through Ruxanne, we mined the top 18 most-starred Rust projects in GitHub to discover the most common bug fix patterns committed to their repositories.
 We analyzed 87,726 code changes drawn from 57,214 commits across these 18 projects.

### 2023-06-14 conclusion.tex
capitalization
last read through

@@
 In this project, we developed Ruxanne, a tool for mining bug fix patterns in Rust.
-Using this tool, we mined the top 18 most starred Rust projects in Github to discover common bug fix patterns within their code changes.
+Using this tool, we mined the top 18 most starred Rust projects in GitHub to discover common bug fix patterns within their code changes.
 At the heart of Ruxanne, we used a novel code embedding approach to embed the most important aspects of a code change in a fixed sized datapoint.
-We stored 87,726 datapoints drawn from 57,214 commits across these 18 projects.
+We processed 87,726 datapoints drawn from 57,214 commits across these 18 projects.
 After using the DBSCAN clustering algorithm, and a subsequent manual analysis, we obtained 12 clusters of general fix patterns and 8 clusters of borrow-checker related fix patterns.
+We found that the most common general pattern was adding or removing fields, while the most common borrow-checker related pattern was removing a \texttt{clone()} call.

### 2023-06-14 discussion.tex
reedit
last read through

@@
 \emph{In summary, program repair tools must understand semantics of attributes and should be performance-aware.} It may also be useful for tools to be able to automatically infer immutability.
-Here, we discuss the top ten most popular patterns, 8 of which fall into the category of general patterns, and the remaining 2 are BC-related patterns.
+Here, we discuss the top ten most popular patterns (8 general patterns and 2 BC-related patterns).
 
@@
 This pattern is performance-related; when correctly applied, it has no semantic impact on the program, but reduces resource consumption.
-Rust is used in situations where performance is important; the number of instances of this pattern confirms that developers do work on performance issues.
+Rust is used in situations where performance is important, and the number of instances of this pattern confirms that developers do work on performance issues.
 This implies that \emph{program repair tools would benefit from being performance-aware}, rather than limiting themselves strictly to functional properties of the code.
@@
 \vspace*{1em}
-\fbox{\parbox{0.9\textwidth}{We summarize learnings from our bug patterns as follows:
+\fbox{\parbox{0.9\textwidth}{We summarize some learnings from our bug patterns:
 \begin{itemize}
@@
 A threat to external validity is selection bias, where the benchmarks differ systematically from the actual world.
-The selected projects might cause selection bias and therefore we might not have presented patterns that may exist in other projects.
-Another threat to external validity can be the developers reporting the commits as bug fixing commits while they are not really fixing a functionality.
+Our selection of projects might be biased and therefore we might not have presented patterns that may exist in other projects.
+Another threat to external validity can be the developers reporting the commits as bug fixing commits while they are not really fixing functionality.
 Similarly, a commit message might not contain our target keywords while the commit is associated with bug fixing changes.

### 2023-06-14 methodology.tex
last read through

@@
 A change pattern is a set of code changes that serves a specific purpose.
-Purposes include rectifying program behaviour with respect to a specific functional or non-functional requirement (bug-fixing patterns) and changes that support bug-fixing patterns (fix-induced patterns) as well as improving code readability or maintainability (refactoring patterns).
+Purposes include rectifying program behaviour with respect to a specific functional or non-functional requirement (bug-fixing patterns) and changes that support bug-fixes (fix-induced patterns) as well as improving code readability or maintainability (refactoring patterns).
 In this work, we are interested in bug-fixing and fix-induced patterns.
@@
 Our approach for automatically finding pervasive code patterns aims to find clusters of similar code patterns, using existing implementations of appropriate clustering algorithms.
-Work in this vein generally makes the assumption that clusters corresponds to classes of change patterns~\citep{hanam2016discovering,campos2019discovering,yang2022mining}.
+Work in this vein generally makes the assumption that clusters correspond to classes of change patterns~\citep{hanam2016discovering,campos2019discovering,yang2022mining}.
 We thus want to associate code changes with datapoints; one of our contributions is a fixed-size embedding which highlights the important parts of a code change.
@@
 Syn handles all of Rust.
-Each of our datapoints specifies whether any Syn non terminal (\texttt{NT}) is present or absent, i.e.
-containing one dimension per non terminal.
-However, we are also interested in change patterns that involve the borrow checker (BC), so we also add a dimension reflecting the presence of a set of BC-related elements (\texttt{BCE}) that we identified, e.g.
+Each of our datapoints specifies whether any Syn non~terminal (\texttt{NT}) is present or absent, i.e.
+containing one dimension per non~terminal.
+However, we are also interested in change patterns that involve the borrow checker (BC), so we add a dimension reflecting the presence of a set of BC-related elements (\texttt{BCE}) that we identified, e.g.
 \texttt{clone}, \texttt{Rc}, \texttt{Box}.
@@
 
-The second part identifies the context of the modification, that is, the path from the root of the tree to the subtree in which the modification has occured.
-The third part is the content of the change---terminals and nonterminals.
+The second part identifies the context of the modification, that is, the path from the root of the tree to the subtree in which the modification has occurred.
+The third part is the content of the change---terminals and non~terminals.
 Ruxanne looks for changes within each root-level scope (the Rust book\footnote{\url{https://doc.rust-lang.org/reference/items.html}} calls these root-level scopes the \textit{items} of a crate).
@@
 
-\paragraph{Example: path extraction.} Figure~\ref{fig:extraction} illustrates path extraction on a simple change, starting from the ASTDiff.
+\paragraph{Running example: path extraction.} Figure~\ref{fig:extraction} illustrates path extraction on a simple change, starting from the ASTDiff.
 The change (shown in the top right corner of the Figure) replaces \verb+tail+ with \verb+&tail+ and \verb+head+ with \verb+head.clone()+.
@@
 A DFS on this tree yields three different paths, which we show in red, green, and purple.
-A path represents a sequence of involved non terminals.
+A path represents a sequence of involved non~terminals.
 It is crucial to record the order of nodes in the path, so that we can record that order in our datapoint embedding.
@@
 
-The most naive approach would be to only report the observed non terminals within the diff.
+The most naive approach would be to only report the observed non~terminals within the diff.
 That yields a fully order-insensitive representation.
@@
 
-A reasonable work around would be to pass from the full set of non terminals to a smaller set of categories (e.g.
-a category for larger entites like class or function definitions, and a category for smaller entities like statements and expressions), and then to record all possible combinations of these categories (similar to \cite{hanam2016discovering}).
+A reasonable work around would be to pass from the full set of non~terminals to a smaller set of categories (e.g.
+a category for larger entities like class or function definitions, and a category for smaller entities like statements and expressions), and then to record all possible combinations of these categories (similar to \cite{hanam2016discovering}).
 That ad-hoc approach reduces the number of dimensions and provides more reasonable and syntactically-correct combinations.
@@
 This captures the varying occurrence counts resulting from different element orders in the program.
-Secondly, we incorporate a weighting scheme, designed semi-automatically, to prioritize non-terminals that we consider more salient for Rust bugs.
+Secondly, we incorporate a weighting scheme, designed semi-automatically, to prioritize non~terminals that we consider more salient for Rust bugs.
 
-\paragraph{Example: computing fixed-sized vectors.} Figure~\ref{fig:essence} depicts the fixed-sized vectors, as described in this subsection, for the change from Figure~\ref{fig:extraction}.
-All of the columns in this vector belong to scope \verb+ItemFn+---that is, these non-terminals occur within a function.
+\paragraph{Running example: computing fixed-sized vectors.} Figure~\ref{fig:essence} depicts the fixed-sized vectors, as described in this subsection, for the change from Figure~\ref{fig:extraction}.
+All of the columns in this vector belong to scope \verb+ItemFn+---that is, these non~terminals occur within a function.
 The Figure also depicts the essence of change, to be described immediately below.
@@
     \caption{\label{fig:essence}Multiplying the number of occurrences of items observed in the tree by their respective weights results in the essence of change, visualized with a circle pack figure at bottom right.
-All non terminals here occur within scope {\texttt ItemFn}.}
+All non~terminals here occur within scope {\texttt ItemFn}.}
 \end{figure}
@@
 We made two observations about changes.
-First, we saw that items (non terminals or BC-related elements) that occurs closer to AST leaves tend to be more semantically important.
+First, we saw that items (non~terminals or BC-related elements) that occur closer to AST leaves tend to be more semantically important.
 Second, and relatedly, items closer to the root tend to be repeated in all the paths within one ASTDiff (e.g.
@@
 We recorded the total number of occurrences of each item.
-Then, if $\# i$ is the number of occurences of $i$, we gave $i$ a weight of $1/\# i$.
+Then, if $\# i$ is the number of occurrences of $i$, we gave $i$ a weight of $1/\# i$.
 Because items that occur closer to the root also occur more often, per our observation, this gives higher weights to items that occur infrequently, i.e.
@@
 
-\paragraph{Example: essence of change.} Figure~\ref{fig:essence} shows a complete example for computing the essence of change.
+\paragraph{Running example: essence of change.} Figure~\ref{fig:essence} shows a complete example for computing the essence of change.
 The essence of change vector essentially serves as our embedding vector.
@@
 
-\subsubsection{\label{sec:manual_analysis_parameter_tuning}Manual Analysis for parameter tuning}
+\subsubsection{\label{sec:manual_analysis_parameter_tuning}Manual analysis for parameter tuning}
 
@@
 
-\subsubsection{\label{sec:manual_analysis_cluster_selection}Manual Analysis for cluster selection}
+\subsubsection{\label{sec:manual_analysis_cluster_selection}Manual analysis for cluster selection}
 % 3 types, refactoring dropped
@@
 \begin{lstlisting}[language=Rust, style=colouredRust]
-// First Pattern
+// First Pattern (more important): change clone to borrow
 -- foo1(arg.clone);
@@
 
-// Second Pattern
+// Second Pattern (less important): change a statement in an fn
 fn foo2() {

### 2023-06-14 related.tex
last read through

@@
 Researchers have proposed different methods for code repair, each of which can be said to follow a different mindset.
-Following the taxonomy propsed by \cite{liu2018survey}, we briefly present two mindsets, and will discuss how having a set of predefined fix patterns can aid repair.
+Following the taxonomy proposed by \cite{liu2018survey}, we briefly present two mindsets, and will discuss how having a set of predefined fix patterns can aid repair.
 
@@
 \cite{ye2021automated} propose ODS, an overfitting patch detection system, which assumes that code features capturing universal correctness properties can be utilized to classify overfitting patches across program repair systems and software projects.
-ODS extracts 202 manually crafted code features by comparing fixed and buggy code and employs supervised learning to build a distributional model for classifying unseen patches, achieving over 70\% accuracy in evaluations on Defects4J, Bugs.jar, and Bears benchmarks.
+ODS extracts 202 manually crafted code features by comparing fixed and buggy code, and employs supervised learning to build a distributional model for classifying unseen patches, achieving over 70\% accuracy in evaluations on Defects4J, Bugs.jar, and Bears benchmarks.
 Similarly, our feature extraction involves manual crafting of features.
 In another study, \cite{tian2022change} present an enhanced APCA model that transforms the problem into a question-answering scenario and incorporates a neural architecture to learn semantic correlations between bug reports and commit messages.
-Additionally, \cite{lin2022context} introduce \textsc{Cache}, a neural-based context-aware AST embedding method that captures both the change part and the unchanged part (context code) of patches.
-Evaluation results demonstrate that context knowledge significantly enhances the performance and accuracy of APCA systems.
+Additionally, \cite{lin2022context} introduce \textsc{Cache}, a neural-based context-aware AST embedding method that captures both the changed part and the unchanged part (context code) of patches.
+Evaluation results demonstrate that context knowledge significantly enhances the performance and accuracy of systems that attempt to automatically assess patch correctness (which is not our goal).
 
@@
 
-\cite{hoang2020cc2vec} introduced CC2Vec, a neural network model that utilizes log messages accompanying code changes to learn representations capturing the semantic information.
+\cite{hoang2020cc2vec} introduced CC2Vec, a neural network model that utilizes log messages accompanying code changes to learn representations capturing semantic information.
 By addressing challenges in distinguishing between added and removed code and incorporating an attention mechanism, CC2Vec outperforms previous techniques for code change representation.
@@
 In~\cite{hanam2016discovering}, they introduced the Feature Properties table, which is a categorization for program elements.
-Embedding programs based on this table resulted in an order-sensitive embedding, however, it made the datapoints sparse.
-Our datapoints are more dense than theirs.
-More specifically, the number of our dimensions is more than 11x smaller.
+Embedding programs based on this table resulted in an order-sensitive embedding.
+However, it made the datapoints sparse, and our datapoints are more dense than theirs.
+More specifically, the number of our dimensions is more than 11$\times$ smaller.
 
 \subsection{Refactoring tools for Rust}
-Clippy is a widely known as the best code refactoring tool for Rust.
+Clippy is widely known as the best code refactoring tool for Rust.
 It can identify over 600 common mistakes and offer automated fixes.
@@
 
-To our knowledge, our work is the first to use an automated mining and code analysis pipeline to find pervasive patterns in Rust open source projects, even if previous studies have investigated common bug patterns in Rust.
+To our knowledge, our work is the first to use an automated mining and code analysis pipeline to find pervasive patterns in Rust open source projects, even if previous studies have (manually) investigated common bug patterns in Rust.
 
@@
 Using the results of their manual analysis, they designed two bug detectors, and provided recommendations for developing more bug detectors in the future.
+Our automated approach to finding bug patterns has similar implications to their work but can operate at a far larger scale.
 
@@
 This framework works by carrying out static analysis on Rust's Mid-level Intermediate Representation (MIR).
-The tool exploits the insights obtained from observing existing bugs (by studying reported CVEs) in Rust code bases.
+The tool exploits the insights obtained from manually observing existing bugs (by studying reported CVEs) in Rust code bases.
 Using both numerical (e.g.

### 2023-06-14 results.tex
last read through

@@
 adding a whole new \verb+impl+ block and implementing new functions in it).
-Such failing change patterns tend to apply across projects less often, as having many changed elements make it easier for datapoints to be dissimlar.
+Such failing change patterns tend to apply across projects less often, as having many changed elements make it easier for datapoints to be dissimilar.
 
@@
 \subsubsection{Attributes}
-The most common bug fix pattern in our corpus modifies attributes.
+A quite common bug fix pattern in our corpus modifies attributes.
 In Rust, attributes enable developers to add metadata to program elements.
@@
 We will see some of these attributes in the following examples.
+\\
 

### 2023-06-18 introduction.tex
Zenodo link updated

@@
 
-\noindent \textbf{Data Availability Statement} The datasets generated and analyzed during the current study are available in the Zenodo repository, \url{https://zenodo.org/record/7388618}.
+\noindent \textbf{Data Availability Statement} The datasets generated and analyzed during the current study are available in the Zenodo repository, \url{https://zenodo.org/record/8052940}.

### 2023-06-18 introduction.tex
updated zenodo link again, small bug

@@
 
-\noindent \textbf{Data Availability Statement} The datasets generated and analyzed during the current study are available in the Zenodo repository, \url{https://zenodo.org/record/8052940}.
+\noindent \textbf{Data Availability Statement} The datasets generated and analyzed during the current study are available in the Zenodo repository, \url{https://zenodo.org/record/8052979}.

### 2023-06-19 response.tex
preparing response doc

\documentclass{article}

\usepackage{cite}
\usepackage{graphicx}
\usepackage{hyperref}

\title{Response to Reviewer Comments}
\date{2023 June}
\author{Mohammad Robati Shirzad and Patrick Lam}
\begin{document}

\maketitle

We present our detailed responses to major revisions received on 27th March 2023 for the paper titled ``A Study of Common Bug Fix Patterns in Rust''.

\section{Introduction}

We thank the reviewers for the thoughtful reviews. We first provide summaries here, followed by detailed responses below.

\subsection{Methodology \& Experiments}

We further explain our methodology and discuss why we made the choices we did. We summarize here, but invite the reviewers to read our full responses. We hope that the reviewers can better understand why we believe our choices to be appropriate.

Reviewer 1 \#1 order information: our dataset is not large enough to meaningfully incorporate it. \#2: including information about terminals would have led to dimensionality in the millions, which would be infeasible in terms of computational complexity. \#3: we have included qualitative comparisons of our approach with previous approaches and discussed how our approach is superior to those approaches.

Reviewer 2 \#1a context: we believe our results show that context is not required to get meaningful results; \#1b leaf nodes and penalizing large diffs: any clustering approach is going to have trouble clustering together large diffs, which have more ways to differ from each other than small diffs. \#2 picking 50 datapoints: e.g. we do not have the resources to evaluate the cluster with 618 datapoints manually. \#3: samples are used for answering RQ1.

In summary, we believe that we have made the best possible design decisions given limitations in datasets and resources available to us, and that our results are strong enough that they support our decisions.

\subsection{Discussion}

We wholeheartedly agree that it is important to show that the identified patterns are valuable. We have added Section 4.1, “Patterns Present and Missing” (specifically discussing the value of the top 10 patterns and summarizing our learnings) and Section 4.2 “Usefulness of Rust Bug Patterns” (discussing the general usefulness of patterns), as well as paragraphs on actionability to 6 of the groups of patterns.

\subsection{Clarity}

Thanks for the suggestions! We have added many of the suggested changes to the text of the paper. We have explained some of the reviewer questions in this response. If a reviewer thinks that those responses should be in the paper itself, or requires further clarification, we are happy to oblige.

\section{Detailed Responses}

\subsection{Reviewer 1}

\textit{This paper conducts a study on bug-fix patterns of the language Rust. Rust is a system-level programming language that has attracted huge attention among developers. Rust is specially designed to help developers to avoid illegal memory operations which are easily introduced in C/C++ programs. It has several particular language features, so the bugs in Rust may also have subtle differences from other programming languages. Thus it is important to study the bug's unique features and trends in Rust.
To perform the study, the authors collect bugs and their corresponding patches from GitHub. Then they embed the patches to vectors that have the same length, and cluster the vectors. At last, they manually analyze the groups and extract fix patterns.
This paper is very well-written and organized. The scale of this study is sufficient and the workflow is overall reasonable. In addition, this paper firstly studies Rust fix patterns, and it proposes a novel embedding method for Rust program diffs (i.e., patches). To this end, I think the significance of this paper is guaranteed.However, I would like to mark this paper as a major revision, with the following concerns.}

\subsubsection{The order information and ignoring the terminal tokens}

\textit{\textbf{Revision:} The embedding lacks order information, which could be simply appended a vector with a sequence representing the occurrence list of the AST nodes.}

\textbf{Response:} Yes, our embedding lacks order information in the dimensions. We do not believe that order information would be useful here. Our application is bug categorization, and previous works on bug categorization obtained reasonable results without ordering. If we wanted to include ordering, we would need a neural model, but our dataset is not large enough to train a neural model that can fully understand the structures of Rust programs.

To elaborate: simply appending a vector representing the occurrence list of AST nodes wouldn't be compatible with DBSCAN, which requires fixed-sized inputs for clustering. AST paths are inherently variable-sized.

code2vec does manage to create an order-sensitive embedding. However, code2vec uses neural networks to generate embeddings by training them on a 30GB dataset. Its goal is to predict certain code elements, such as method names. That is not our goal.

To apply a code2vec-style approach, we would need to create a dictionary of all possible paths within an AST, and then provide the neural network with observed paths from our target programming language.

We instead chose to follow previous bug categorization work (for JavaScript, FSE 2016; and for Python, EMSE 2022), and simply use a feature vector to represent code differences, rather than neural-based embeddings. Earlier researchers chose feature vectors for similar reasons as we did: the nature of the bug categorization task, and dataset limitations. \\ \\

\textit{\textbf{Revision:} In addition, ignoring the terminal tokens has advantages and disadvantages, and it is better to evaluate by experiments. }

\textbf{Response:} We appreciate the value of empirically validating design decisions when possible. However, in our case, not ignoring terminal tokens (variable names and constant values) would have been impractical, due to the high dimensionality (millions) of the resulting embedding vectors. The cost of processing such data would be prohibitive in both time and space. We judged that the likelihood of improvements was unlikely to justify the significant engineering effort required to get DBSCAN (which is $O(nlogn)$) to work with that magnitude of sparse data.

code2vec-style neural embedding could mitigate the high dimensionality; we could create a separate dictionary specifically for variable names (terminals) and generate embeddings by concatenating values from both the variable names dictionary and the paths dictionary. This way, the resulting embedding would incorporate knowledge from both nonterminal and terminal elements of the AST. However, as discussed above, we chose a feature vector rather than a neural embedding, following previous work, because a neural embedding would be overkill for bug categorization and because of dataset limitations. \\ \\

\subsubsection{Illustrating patch to vector transformation}

\textit{\textbf{Revision:} 2. It is better to add an example to illustrate the procedure of transforming a patch to a vector, i.e., use an example patch in Section 2 to link all the processing procedures.}

\textbf{Response:} Thanks. We have made the example in Section 2 much more explicit and added some previously-missing steps to Figure 2 (namely the essence of change vector) as well as text linking Figures 1 and 2. Section 2 now contains a complete running example of transforming a patch to a vector. This example is in Sections 2.1.2 and 2.1.3. We describe the example in paragraphs labelled “Running example” throughout these subsections. We are happy to pull the motivating example into its own subsection if the reviewer thinks that is most appropriate. \\ \\

\subsubsection{Conclusion of the patterns}

\textit{\textbf{Revision:} 3. The highlighted conclusion of the pattern is missing. I expect to know which part is more likely to be buggy and want to know several typical examples.}

\textbf{Response:} The new Section 4.1 (``Patterns Present and Missing'') summarizes the patterns that we found and their overall usefulness. We also added an “actionability” section to eight of the patterns to discuss them in further detail and how they can aid a repair tool. \\ \\

\subsubsection{Comparing the discovered patterns to other languages}

\textit{\textbf{Revision:} 4. Also, I expect to see the bug patterns that are unique in Rust programs, and further discussions about the differences with other languages like C, C++, Python and Java. }

\textbf{Response:} The BC-related patterns are unique to Rust programs because they concern Rust's unique borrow checker. On the other hand, 11 of the 12 general patterns (excluding G.option) have counterparts in other widely-used programming languages, and we discuss differences in the text.

Specifically, we have added content to Section 3.2.1 about similarities and differences between the Rust attribute bug pattern and Python decorators/Java annotations. Section 3.2.5 relates Rust traits to Java interfaces and shows an example of how traits are more complicated. Section 3.2.6 explains how Rust's match generalizes C-style switch statements. \\ \\

\subsubsection{Concerns about RQ1}

\textit{\textbf{Revision:} 5. The answer to RQ1 is not persuasive. Although the most important AST node can be somehow regarded as the most representative part of a patch, the relationship is not that strong to get the answer. It is better to compare with existing approaches and design a better metric to measure the performance.}

\textbf{Response:} We understand your concern regarding the persuasiveness of the answer to RQ1. We've acknowledged this limitation in our study and have addressed it in our threats to validity section. While it is true that the relationship between the most important AST node and the most representative part of a patch may not be exceptionally strong, we designed our approach to prioritize the non-terminals that are important, and we sought to examine the connection between those important non-terminals and actual code changes.

We'd like to point out that previous work, such as the bug pattern studies for JavaScript (FSE 2016) and Python (EMSE 2022), each only introduced one embedding method to compute code change vectors that produced meaningful clusters, and there was no comparison between different embedding methods in those works.

In contrast to the existing embedding approach proposed for JavaScript (FSE 2016), where they introduced the Feature Properties table leading to an order-sensitive embedding but sparse datapoints, our approach results in 11x denser datapoints (Section 5.4), which implies smaller storage requirements and better DBSCAN performance. The feature vector proposed for Python (EMSE 2022) focuses only on single-hunk bugs and includes information such as the number of variables, lines, and function arguments. Our approach is more general: it handles multiple-hunk commits.

In our RQ1 experiment, we opted for a manual inspection process where the two authors, who possess Rust knowledge, reviewed the vector and its corresponding code change to verify representativeness. While this manual approach may not be as statistically rigorous as an automated inspection (because it inspects smaller numbers of points), we considered it to be an intuitive and straightforward method.

Of course, an automated inspection could examine more code examples. Such an inspection would also be subject to the validity threat of embedding assumptions about what to inspect. And, developing an automated code comprehension tool specifically for Rust, capable of interpreting code changes and providing explanations for the reasons behind those changes, poses a significant challenge. Such a task would itself require a separate and comprehensive study. We believe that developing an automated inspection tool would be a distraction and would not contribute to answering the core questions about Rust bug patterns behind this work.

We further emphasize that our code embedding approach is only the initial step in our bug categorization pipeline. We had to choose a specific embedding method to create our code diff vectors: different embedding approaches would lead to entirely different datasets and possibly different clustering outcomes. Each embedding approach would require separate manual analysis of the resulting clusters. Consequently, we made the decision to select one embedding method and to validate it.

To additionally justify our choice of embedding approach, we evaluated whether the clusters generated by our approach contained dissimilar data points, as this was a critical factor in determining its suitability. Our bug categories demonstrate that our clusters successfully encompass cross-project bugs, and this can be verified by examining the introduced code changes in our replication package. \\ \\

\subsubsection{Adding more bibliography}

\textit{\textbf{Revision:} 6. Several patch/code/text embedding works should be discussed in the section of related work.
[1] Automated Classification of Overfitting Patches with Statically Extracted Code Features. (TSE 2019)
[2] Is this Change the Answer to that Problem? Correlating Descriptions of Bug and Code Changes for Evaluating Patch Correctness. (ASE 2019)
[3] Context-Aware Code Change Embedding for Better Patch Correctness Assessment. (TOSEM 2022)
[4] CC2Vec: Distributed Representations of Code Changes. (ICSE 2020)
[5] code2seq: Generating Sequences from Structured Representations of Code. (ICLR 2019)
[6] BERT: Pre-training of Deep Bidirectional Transformers for Language Understanding.}

\textbf{Response:} Thanks. We have added 5 of these works to our related work section.

\subsection{Reviewer 2}

\textit{In this paper, the authors aim to mine and study bug-fix patterns in Rust programs. They discuss a new encoding method for Rust patches, which changes a patch into a vector. Then they apply a clustering algorithm to identify common patch patterns. In the end, they discuss 20 identified patterns.
Rust is a new, popular programming language, and studying its bug-fix patterns is important. However, I have some concerns about the methodology and the value of the identified patterns. I discuss my concerns as follows.} \\ \\

\subsubsection{The value of the identified patterns}

\textit{\textbf{Revision:} The value of the identified patterns is unclear to me. The authors claim that they want to study bug-fix patterns in the title and in the methodology section. However, it is difficult for me to correlate the discussed patterns with any particular bug types. I cannot figure out any actions (e.g., how to build bug detection tools, how to inspect my Rust code) to take after reading the paper.}

\textbf{Response:} The new Section 4.1 (“Patterns Present and Missing”) summarizes the patterns that we found and what one could do with them. We also added an “actionability” section to some of the patterns to discuss them in further detail and how they can aid a repair tool. We found that many of the patterns are interesting but still require further work to distill into a repair tool. \\ \\

\subsubsection{Refactoring, not bug fixing}

\textit{\textbf{Revision:} Moreover, many studied patterns are actually for refactoring purposes and compiler errors (not for fixing bugs).}

\textbf{Response:} In Section 2.3.2 we discuss three types of changes: bug-fix, fix-induced, and refactoring, following the definitions of Cotroneo et al (2019). Our intention is to only consider bug-fixing and fix-induced commits, and we aimed to omit changes that were purely refactoring. However, some of the changes that we included may appear to be refactoring changes but actually have performance implications. We chose to specifically include those. For instance, BC.clone.ref does not alter the program's behaviour (no additional test cases pass) but can reduce CPU usage. Other changes, like adding Option to a type, can be considered refactorings, but are in support of bug fixes, and hence are fix-induced. \\ \\

\subsubsection{Committing code with compiler errors}

\textit{\textbf{Revision:} It is very rare for a programmer to submit a commit that makes a Rust program uncompiled.}

\textbf{Response:} We agree, and Section 4.1 in our discussion acknowledges this point. \\ \\

\textit{\textbf{Revision:} Why does the first pattern in Section 3.3.3 show up in the study? The code before the patch in the first figure of Section 3.3.3 cannot be compiled.}

\textbf{Response:} Whether the code compiles or not depends on the method signature of \verb+emit()+.

If \verb+emit()+ is like this:
    \verb+fn emit(&self, db: &DiagnosticBuilder);+
Then \verb+mut+  is not necessary.

If \verb+emit()+ is like this:
    \verb+fn emit(&mut self, db: &DiagnosticBuilder);+
Then the presence of \verb+mut+ is mandatory in the closure.

In the example mentioned here, the presence of \verb+mut+ in the closure parameter \verb+|mut e|+ becomes necessary following a modification of \verb+emit()+ to require a mutable reference (\verb+&mut self+). The patch that we are classifying simultaneously changes \verb+emit()+ and the reference to add \verb+mut+ in both places. \\ \\

\subsubsection{Our embedding method}

\textit{\textbf{Revision:} The method for encoding patches seems problematic to me. Giving leaf elements more weights prioritizes small patches. In section 3, all discussed patches are small patches without any information about what is the bug or the program context. My understanding is that almost all code in Section 3 before the patch is good code. They become buggy because they are used in particular contexts. Combining the patches together with the program contexts is more valuable.}

\textbf{Response:} Let's first talk about the issue of context. We agree that a bug is only a bug in a particular context. We understand that the reviewer is saying: had we considered the statements around a bug fix, then we would have gotten better results than just considering the bug fix itself in isolation.

We point out that Ruxanne considers changes that developers have labelled as bug-fixing changes (or fix-induced changes). Part of our thesis is that, given a change labelled as a bug-fixing change, the AST diff is enough to permit classification of that bug, and that context is not required. We believe that our results show that it is indeed enough: we identified 12 general patterns and 8 borrow-checker patterns based on the information we collected.  Section 3 provides links to the changes, so that the reader can examine the changes themselves.

We'd ask the reviewer to consider our results on their own merits, rather than thinking about a possible other approach that also could have worked. It is possible that more information, as the reviewer suggests, gives better results, but it also may overfit, for instance, and computations are harder with more data. In our response to Reviewer 1, we discussed the embedding approaches taken by the Python and JavaScript related works.

Having said that, Ruxanne does consider the parents in the AST when reasoning about a change. It does not specifically consider siblings of a change in the AST, which is the context that we believe the reviewer is talking about. For instance, in Figure 1, the head.clone() change paths include the addition of the method call node to the path containing the nonterminal head, as well as the addition of the path containing the clone() call. We consider both of those paths from root to leaf, shown as Paths 2 and 3 in the Figure.

As for the comment about small patches and leaf nodes, we'd like to clarify the role of the weights. We performed a preliminary step to determine fixed weights for different types of AST nodes. We then used these weights, uniformly across our data set, to compute essence-of-change vectors (as depicted in Figure 2). Both small and large patches have their occurrences vectors multiplied by the same weights, giving essence-of-change vectors. RQ1 validates this essence-of-change computation. In subsequent steps, we then cluster the essence-of-change vectors to find bug patterns.

We acknowledge that it is possible that a large patch may be harder to cluster, because it has to look like other patches.  Additionally, we believe that any clustering-based approach is likely going to have large diffs be further apart for the same reason that ours does: there's just more going on. But there is nothing inherently penalizing a larger diff. Also, we point out that the related work by Pan et al specifically excludes diffs with seven statements or more, while we allow them. \\ \\

\subsubsection{AST Transformation}

\textit{The implementation described in Section 2.1 is difficult to follow.}

\textit{\textbf{Revision:} First, I don't understand how an AST is transformed to a dictionary in Section 2.1.1.}

\textbf{Response:} In the paper, we tried to avoid explaining technical details of how a Syn tree which is written in Rust Debugging format is transformed to JSON format. JSON format and Python dictionary are basically the same format. Two images below show an AST in Rust Debugging format and in JSON (Python Dictionary) format. As you can see, the difference is only in quotation marks and curly/square brackets.

\begin{figure}[h]
    \centering
    \includegraphics[width=1\textwidth]{astpics}
    \caption{Syn tree in Rust Debugging format (left) and Python dictionary format (right)}
\end{figure}

\subsubsection{Inconsistency between number of paths and the number of changed arguments}

\textit{\textbf{Revision:} Second, I cannot understand Figure 1. What do the three paths represent? If they represent the changed parameters, why are there three paths, but there are only two parameters changed? If they represent parameters, how can they represent changes from the last version?}

\textbf{Response:} This is a really good question. We added a few more sentences to make sure this doesn't confuse the readers. To clarify: the tree depicted in Figure 1 is not the actual AST; rather, it represents the ASTDiff (now written on the label on top left). In our approach, we consider two versions of the code: the buggy code and the fixed code. Consequently, we have two ASTs. The ASTDiff, shown in Figure 1, represents the difference between these two ASTs in Python dictionary format. The number of paths in the ASTDiff does not directly correspond to the number of changed elements in the source code. Instead, it is determined by the structural differences between the two ASTs of the code versions. The paths in the ASTDiff capture the changes in the code structure between the two versions, allowing us to analyze and extract relevant information for further processing. We have also added more text to Section 2.1.2 explaining this.

\subsubsection{Inconsistency between number of paths and the number of changed arguments}

\textit{\textbf{Revision:} Third, I don't understand the following sentence on page 7: "we record the number of occurrences of items within the paths at each scope." Overall, the implementation can be benefited from a carefully described example.}

\textbf{Response:} We revised the whole paragraph (now on page 8) and made the running example in Section 2 more explicit, as also requested by Reviewer 1. We meant that we record the number of occurrences of elements of \verb+NT+ \verb+BCE+ within each distinct root level scope. Root level scopes can be structs, functions, impl blocks, …. Note that Figure 2 does not explicitly show the single root level scope for this example, which is ItemFn. Other diffs could contribute items in different root level scopes. \\

\subsubsection{Minor issues}

\textit{\textbf{Revision:} 1. The introduction is organized in a very strange way. The authors should introduce Rust first, and then discuss studying common bug/fix patterns is important for Rust. However, the current writing does not mention Rust until the second page.}

\textbf{Response:} Thanks. Moved Rust introduction to the beginning. \\ \\

\textit{\textbf{Revision:} 2. Line 15 on page 7: "We found it crucial to record" => it is crucial.}

\textbf{Response:} fixed \\ \\

\textit{\textbf{Revision:} 3. The authors picked 50 datapoints for each cluster. Are there any particular reasons for choosing 50? There are very few clusters with more than 50 datapoints. A more reasonable method is to inspect all datapoints.}

\textbf{Response:} Our analysis revealed many clusters (39 in total) with more than 50 datapoints. More specifically, one cluster has 618 datapoints, four clusters more than 300 datapoints, and 15 clusters more than 100 datapoints. With our number of datapoints, manual inspection would be overwhelmingly time-consuming.

We believe that the choice of 50 datapoints per cluster successfully balances ensuring that the selected datapoints truly represent the same change and making manual inspection feasible. Our sample size allows us to get meaningful insights into the characteristics of each cluster. \\ \\

\textit{\textbf{Revision:} 4. Line 6 on page 14: "A human analyzer (paper author) compares the actual change in the code and compares it to the visual representation of the respective data point." => A human analyzer (paper author) inspects the actual change … and compares it with the visual representation …}

\textbf{Response:} fixed \\ \\

\textit{\textbf{Revision:} 5. What are the collected samples in the third paragraph of page 14 used for?}

\textbf{Response:} We proposed a method to embed key AST information in a fixed-sized vector using a semi-automatically derived weighting scheme. RQ1 asks how successful we've been in terms of giving more values to the important elements in our ASTs. For instance, if in a code patch, the only change is adding \verb+.clone()+, then we should expect the element \verb+clone+ to have the highest value. In RQ1, we sample 25 datapoints from General datapoints and 25 datapoints from BC-related datapoints (50 in total) to manually analyze whether the values in their code vector actually respect the important elements in the AST. Note that we selected 50 datapoints that minimize the sampling error, i.e. it is a low bias sampling. \\ \\

\textit{\textbf{Revision:} 6. I don't understand the sentence at line 25 of page 23. Particularly, what string formatting methods are gone through if using p instead of *p?}

\textbf{Response:} fixed, added Clippy reference in the paper.

This is a common Clippy warning. You can find information about it through this link: \url{https://rust-lang.github.io/rust-clippy/master/index.html#inefficient_to_string} \\

In the link the example is: \\

\noindent\verb+// Generic implementation for `T: Display` is used (slow)+ \\
\verb+["foo", "bar"].iter().map(|s| s.to_string());+\\

\noindent\verb+// OK, the specialized impl is used+\\
\verb+["foo", "bar"].iter().map(|&s| s.to_string());+\\

It is the same as:

\noindent\verb+// Generic implementation for `T: Display` is used (slow)+\\
\verb+["foo", "bar"].iter().map(|s| s.to_string());+\\

\noindent\verb+// OK, the specialized impl is used+\\
\verb+["foo", "bar"].iter().map(|s| (*s).to_string());+\\

In summary, when \verb+p+ is used and it's not clear for Clippy that \verb+p+ cannot be of type \verb+&&T+ where \verb+T+ implements the \verb+ToString+ trait, it suggests dereferencing \verb+p+. The reason is that \verb+p.to_string()+ will use the \verb+format!()+ macro under the hood. The \verb+format!()+ macro is less efficient than direct casting of \verb+&T+ to String (probably because it allocates extra memory first then populates it based on formatting annotations). \verb+(*p).to_string()+ will not use \verb+format!()+.

% \small
% \bibliographystyle{plain}
% \bibliography{refs}

\end{document}

### 2023-06-21 response.tex
response doc is ready

@@
 We hope that the reviewers can better understand why we believe our choices to be appropriate.
+\\
 
-Reviewer 1 \#1 order information: our dataset is not large enough to meaningfully incorporate it.
-\#2: including information about terminals would have led to dimensionality in the millions, which would be infeasible in terms of computational complexity.
-\#3: we have included qualitative comparisons of our approach with previous approaches and discussed how our approach is superior to those approaches.
+Reviewer 1:
+\begin{itemize}
+    \item The order information: our dataset is not large enough to meaningfully incorporate it (\ref{rev:1:order}).
+    \item Ignoring terminal tokens: including information about terminals would have led to dimensionality in the millions, which would be infeasible in terms of computational complexity (\ref{rev:1:terminal}).
+    \item RQ1: We have included qualitative comparisons of our approach with previous approaches and discussed how our approach is superior to those approaches. (\ref{rev:1:rq1}).
+\end{itemize}
 
-Reviewer 2 \#1a context: we believe our results show that context is not required to get meaningful results; \#1b leaf nodes and penalizing large diffs: any clustering approach is going to have trouble clustering together large diffs, which have more ways to differ from each other than small diffs.
-\#2 picking 50 datapoints: e.g.
-we do not have the resources to evaluate the cluster with 618 datapoints manually.
-\#3: samples are used for answering RQ1.
+Reviewer 2:
+\begin{itemize}
+    \item Context: we believe our results show that context is not required to get meaningful results (\ref{rev:2:embedding}).
+    \item Leaf nodes and penalizing large diffs: any clustering approach is going to have trouble clustering together large diffs, which have more ways to differ from each other than small diffs (\ref{rev:2:embedding}).
+    \item Picking 50 datapoints: e.g. we do not have the resources to evaluate the cluster with 618 datapoints manually (\ref{rev:2:minor}).
+\end{itemize}
 
@@
 
-\subsubsection{The order information and ignoring the terminal tokens}
+\subsubsection{\label{rev:1:order}The order information}
 
@@
 
-code2vec does manage to create an order-sensitive embedding.
+code2vec~\cite{alon2019code2vec} does manage to create an order-sensitive embedding.
 However, code2vec uses neural networks to generate embeddings by training them on a 30GB dataset.
@@
 
-We instead chose to follow previous bug categorization work (for JavaScript, FSE 2016; and for Python, EMSE 2022), and simply use a feature vector to represent code differences, rather than neural-based embeddings.
+We instead chose to follow previous bug categorization work (for JavaScript~\cite{hanam2016discovering} and Python~\cite{yang2022mining}), and simply use a feature vector to represent code differences, rather than neural-based embeddings.
 Earlier researchers chose feature vectors for similar reasons as we did: the nature of the bug categorization task, and dataset limitations.
 \\ \\
+
+\subsubsection{\label{rev:1:terminal}Ignoring the terminal tokens}
 
@@
 
-\subsubsection{Illustrating patch to vector transformation}
+\subsubsection{\label{rev:1:patch}Illustrating patch to vector transformation}
 
-\textit{\textbf{Revision:} 2.
-It is better to add an example to illustrate the procedure of transforming a patch to a vector, i.e., use an example patch in Section 2 to link all the processing procedures.}
+\textit{\textbf{Revision:} It is better to add an example to illustrate the procedure of transforming a patch to a vector, i.e., use an example patch in Section 2 to link all the processing procedures.}
 
@@
 
-\subsubsection{Conclusion of the patterns}
+\subsubsection{\label{rev:1:conclusion}Conclusion of the patterns}
 
-\textit{\textbf{Revision:} 3. The highlighted conclusion of the pattern is missing. I expect to know which part is more likely to be buggy and want to know several typical examples.}
+\textit{\textbf{Revision:} The highlighted conclusion of the pattern is missing. I expect to know which part is more likely to be buggy and want to know several typical examples.}
 
@@
 
-\subsubsection{Comparing the discovered patterns to other languages}
+\subsubsection{\label{rev:1:compare}Comparing the discovered patterns to other languages}
 
-\textit{\textbf{Revision:} 4.
-Also, I expect to see the bug patterns that are unique in Rust programs, and further discussions about the differences with other languages like C, C++, Python and Java.
-}
+\textit{\textbf{Revision:} Also, I expect to see the bug patterns that are unique in Rust programs, and further discussions about the differences with other languages like C, C++, Python and Java. }
 
@@
 
-\subsubsection{Concerns about RQ1}
+\subsubsection{\label{rev:1:rq1}Concerns about RQ1}
 
-\textit{\textbf{Revision:} 5.
-The answer to RQ1 is not persuasive.
+\textit{\textbf{Revision:} The answer to RQ1 is not persuasive.
 Although the most important AST node can be somehow regarded as the most representative part of a patch, the relationship is not that strong to get the answer.
@@
 
-We'd like to point out that previous work, such as the bug pattern studies for JavaScript (FSE 2016) and Python (EMSE 2022), each only introduced one embedding method to compute code change vectors that produced meaningful clusters, and there was no comparison between different embedding methods in those works.
+We'd like to point out that previous work, such as the bug pattern studies for JavaScript~\cite{hanam2016discovering} and Python~\cite{yang2022mining}, each only introduced one embedding method to compute code change vectors that produced meaningful clusters, and there was no comparison between different embedding methods in those works.
 
-In contrast to the existing embedding approach proposed for JavaScript (FSE 2016), where they introduced the Feature Properties table leading to an order-sensitive embedding but sparse datapoints, our approach results in 11x denser datapoints (Section 5.4), which implies smaller storage requirements and better DBSCAN performance.
-The feature vector proposed for Python (EMSE 2022) focuses only on single-hunk bugs and includes information such as the number of variables, lines, and function arguments.
+In contrast to the existing embedding approach proposed for Hanam et al.~\cite{hanam2016discovering}, where they introduced the Feature Properties table leading to an order-sensitive embedding but sparse datapoints, our approach results in 11x denser datapoints (Section 5.4), which implies smaller storage requirements and better DBSCAN performance.
+The feature vector proposed for Yang et al.~\cite{yang2022mining} focuses only on single-hunk bugs and includes information such as the number of variables, lines, and function arguments.
 Our approach is more general: it handles multiple-hunk commits.
@@
 
-\subsubsection{Adding more bibliography}
+\subsubsection{\label{rev:1:bib}Adding more bibliography}
 
-\textit{\textbf{Revision:} 6. Several patch/code/text embedding works should be discussed in the section of related work.
-[1] Automated Classification of Overfitting Patches with Statically Extracted Code Features. (TSE 2019)
-[2] Is this Change the Answer to that Problem? Correlating Descriptions of Bug and Code Changes for Evaluating Patch Correctness. (ASE 2019)
-[3] Context-Aware Code Change Embedding for Better Patch Correctness Assessment. (TOSEM 2022)
-[4] CC2Vec: Distributed Representations of Code Changes. (ICSE 2020)
-[5] code2seq: Generating Sequences from Structured Representations of Code. (ICLR 2019)
-[6] BERT: Pre-training of Deep Bidirectional Transformers for Language Understanding.}
+\textit{\textbf{Revision:} Several patch/code/text embedding works should be discussed in the section of related work.}
+\begin{enumerate}
+    \item Automated Classification of Overfitting Patches with Statically Extracted Code Features. (TSE 2019)
+    \item Is this Change the Answer to that Problem? Correlating Descriptions of Bug and Code Changes for Evaluating Patch Correctness. (ASE 2019)
+    \item Context-Aware Code Change Embedding for Better Patch Correctness Assessment. (TOSEM 2022)
+    \item CC2Vec: Distributed Representations of Code Changes. (ICSE 2020)
+    \item code2seq: Generating Sequences from Structured Representations of Code. (ICLR 2019)
+    \item BERT: Pre-training of Deep Bidirectional Transformers for Language Understanding.
+\end{enumerate}
 
@@
 
-\subsubsection{The value of the identified patterns}
+\subsubsection{\label{rev:2:value}The value of the identified patterns}
 
@@
 
-\subsubsection{Refactoring, not bug fixing}
+\subsubsection{\label{rev:2:refactoring}Refactoring, not bug fixing}
 
@@
 We chose to specifically include those.
-For instance, BC.clone.ref does not alter the program's behaviour (no additional test cases pass) but can reduce CPU usage.
+For instance, \verb+BC.clone.ref+ does not alter the program's behaviour (no additional test cases pass) but can reduce CPU usage.
 Other changes, like adding Option to a type, can be considered refactorings, but are in support of bug fixes, and hence are fix-induced.
@@
 
-\subsubsection{Committing code with compiler errors}
+\subsubsection{\label{rev:2:compile}Committing code with compiler errors}
 
@@
 
-\subsubsection{Our embedding method}
+\subsubsection{\label{rev:2:embedding}Our embedding method}
 
@@
 It is possible that more information, as the reviewer suggests, gives better results, but it also may overfit, for instance, and computations are harder with more data.
-In our response to Reviewer 1, we discussed the embedding approaches taken by the Python and JavaScript related works.
+In our response to Reviewer 1, we discussed the embedding approaches taken by the Python~\cite{yang2022mining} and JavaScript~\cite{hanam2016discovering} related works.
 
@@
 It does not specifically consider siblings of a change in the AST, which is the context that we believe the reviewer is talking about.
-For instance, in Figure 1, the head.clone() change paths include the addition of the method call node to the path containing the nonterminal head, as well as the addition of the path containing the clone() call.
+For instance, in Figure 1, the \verb+head.clone()+ change paths include the addition of the method call node to the path containing the nonterminal \verb+head+, as well as the addition of the path containing the \verb+clone()+ call.
 We consider both of those paths from root to leaf, shown as Paths 2 and 3 in the Figure.
@@
 
-\subsubsection{AST Transformation}
+\subsubsection{\label{rev:2:ast}AST Transformation}
 
@@
 
-\subsubsection{Inconsistency between number of paths and the number of changed arguments}
+\subsubsection{\label{rev:2:path}Inconsistency between number of paths and the number of changed arguments}
 
@@
 
-\subsubsection{Inconsistency between number of paths and the number of changed arguments}
-
 \textit{\textbf{Revision:} Third, I don't understand the following sentence on page 7: "we record the number of occurrences of items within the paths at each scope." Overall, the implementation can be benefited from a carefully described example.}
@@
 \textbf{Response:} We revised the whole paragraph (now on page 8) and made the running example in Section 2 more explicit, as also requested by Reviewer 1.
-We meant that we record the number of occurrences of elements of \verb+NT+ \verb+BCE+ within each distinct root level scope.
+We meant that we record the number of occurrences of elements of \verb+NT+ $\cup$ \verb+BCE+ within each distinct root level scope.
 Root level scopes can be structs, functions, impl blocks, ….
@@
 
-\subsubsection{Minor issues}
+\subsubsection{\label{rev:2:minor}Minor issues}
 
@@
 
-\textit{\textbf{Revision:} 2. Line 15 on page 7: "We found it crucial to record" => it is crucial.}
+\textit{\textbf{Revision:} 2. Line 15 on page 7: "We found it crucial to record" $\rightarrow$ it is crucial.}
 
@@
 \textit{\textbf{Revision:} 4.
-Line 6 on page 14: "A human analyzer (paper author) compares the actual change in the code and compares it to the visual representation of the respective data point." => A human analyzer (paper author) inspects the actual change … and compares it with the visual representation …}
+Line 6 on page 14: "A human analyzer (paper author) compares the actual change in the code and compares it to the visual representation of the respective data point." $\rightarrow$ A human analyzer (paper author) inspects the actual change … and compares it with the visual representation …}
 
@@
 
-% \small
-% \bibliographystyle{plain}
-% \bibliography{refs}
+\small
+\bibliographystyle{plain}
+\bibliography{refs}
 

### 2023-06-22 response.tex
formatting fixes

@@
 
-We present our detailed responses to major revisions received on 27th March 2023 for the paper titled ``A Study of Common Bug Fix Patterns in Rust''.
+Please find below our response to the reviews for our EMSE submission ``A Study of Common Bug Fix Patterns in Rust''.
 
@@
 
-We further explain our methodology and discuss why we made the choices we did.
-We summarize here, but invite the reviewers to read our full responses.
+In response to reviewer questions about methodology and experiments, we further explain our methodology and discuss why we made the choices we did.
+We summarize here, but invite the reviewers to read our full responses in Section~\ref{sec:detailed} of this response.
 We hope that the reviewers can better understand why we believe our choices to be appropriate.
-\\
+(Hyperlinked references in this document refer to parts of the response, while un-linked references are to the paper).\\
 
@@
 \begin{itemize}
-    \item The order information: our dataset is not large enough to meaningfully incorporate it (\ref{rev:1:order}).
+    \item Order information: our dataset is not large enough to incorporate it (\ref{rev:1:order}).
     \item Ignoring terminal tokens: including information about terminals would have led to dimensionality in the millions, which would be infeasible in terms of computational complexity (\ref{rev:1:terminal}).
-    \item RQ1: We have included qualitative comparisons of our approach with previous approaches and discussed how our approach is superior to those approaches. (\ref{rev:1:rq1}).
+    \item RQ1: we have included qualitative comparisons of our approach with previous approaches and discussed how our approach is superior to those approaches. (\ref{rev:1:rq1}).
 \end{itemize}
@@
     \item Leaf nodes and penalizing large diffs: any clustering approach is going to have trouble clustering together large diffs, which have more ways to differ from each other than small diffs (\ref{rev:2:embedding}).
-    \item Picking 50 datapoints: e.g. we do not have the resources to evaluate the cluster with 618 datapoints manually (\ref{rev:2:minor}).
+    \item Picking 50 datapoints: manually evaluating large clusters is infeasible---e.g. we do not have the resources to evaluate the cluster with 618 datapoints manually (\ref{rev:2:minor}).
 \end{itemize}
@@
 \section{Detailed Responses}
+\label{sec:detailed}
+We have reproduced all of the reviewer comments below (in italics) and discuss them in detail. Once again, thanks for the detailed reviews.
 
@@
 
-\textit{\textbf{Revision:} The embedding lacks order information, which could be simply appended a vector with a sequence representing the occurrence list of the AST nodes.}
+\noindent \textit{\textbf{R1:} The embedding lacks order information, which could be simply appended a vector with a sequence representing the occurrence list of the AST nodes.}
 
-\textbf{Response:} Yes, our embedding lacks order information in the dimensions.
+\vspace*{1em} \noindent \textbf{Response:} Yes, our embedding lacks order information in the dimensions.
 We do not believe that order information would be useful here.
 Our application is bug categorization, and previous works on bug categorization obtained reasonable results without ordering.
-If we wanted to include ordering, we would need a neural model, but our dataset is not large enough to train a neural model that can fully understand the structures of Rust programs.
+If we wanted to include ordering in the embedding, we would need a neural model, but our dataset is not large enough to train a neural model that can fully understand the structures of Rust programs.
 
@@
 
-\textit{\textbf{Revision:} In addition, ignoring the terminal tokens has advantages and disadvantages, and it is better to evaluate by experiments. }
+\textit{\textbf{R1:} In addition, ignoring the terminal tokens has advantages and disadvantages, and it is better to evaluate by experiments. }
 
-\textbf{Response:} We appreciate the value of empirically validating design decisions when possible.
+\vspace*{1em} \noindent \textbf{Response:} We appreciate the value of empirically validating design decisions when possible.
 However, in our case, not ignoring terminal tokens (variable names and constant values) would have been impractical, due to the high dimensionality (millions) of the resulting embedding vectors.
 The cost of processing such data would be prohibitive in both time and space.
-We judged that the likelihood of improvements was unlikely to justify the significant engineering effort required to get DBSCAN (which is $O(nlogn)$) to work with that magnitude of sparse data.
+We judged that the likelihood of improvements was unlikely to justify the significant engineering effort required to get DBSCAN (which is $O(n~\mbox{log}~n)$) to work with that magnitude of sparse data.
 
@@
 
-\textit{\textbf{Revision:} It is better to add an example to illustrate the procedure of transforming a patch to a vector, i.e., use an example patch in Section 2 to link all the processing procedures.}
+\textit{\textbf{R1:} It is better to add an example to illustrate the procedure of transforming a patch to a vector, i.e., use an example patch in Section 2 to link all the processing procedures.}
 
-\textbf{Response:} Thanks.
+\vspace*{1em} \noindent \textbf{Response:} Thanks.
 We have made the example in Section 2 much more explicit and added some previously-missing steps to Figure 2 (namely the essence of change vector) as well as text linking Figures 1 and 2.
@@
 
-\textit{\textbf{Revision:} The highlighted conclusion of the pattern is missing. I expect to know which part is more likely to be buggy and want to know several typical examples.}
+\textit{\textbf{R1:} The highlighted conclusion of the pattern is missing. I expect to know which part is more likely to be buggy and want to know several typical examples.}
 
-\textbf{Response:} The new Section 4.1 (``Patterns Present and Missing'') summarizes the patterns that we found and their overall usefulness.
+\vspace*{1em} \noindent \textbf{Response:} The new Section 4.1 (``Patterns Present and Missing'') summarizes the patterns that we found and their overall usefulness.
 We also added an “actionability” section to eight of the patterns to discuss them in further detail and how they can aid a repair tool.
@@
 
-\textit{\textbf{Revision:} Also, I expect to see the bug patterns that are unique in Rust programs, and further discussions about the differences with other languages like C, C++, Python and Java. }
+\textit{\textbf{R1:} Also, I expect to see the bug patterns that are unique in Rust programs, and further discussions about the differences with other languages like C, C++, Python and Java. }
 
-\textbf{Response:} The BC-related patterns are unique to Rust programs because they concern Rust's unique borrow checker.
+\vspace*{1em} \noindent \textbf{Response:} The BC-related patterns are unique to Rust programs because they concern Rust's unique borrow checker.
 On the other hand, 11 of the 12 general patterns (excluding G.option) have counterparts in other widely-used programming languages, and we discuss differences in the text.
@@
 
-\textit{\textbf{Revision:} The answer to RQ1 is not persuasive.
+\textit{\textbf{R1:} The answer to RQ1 is not persuasive.
 Although the most important AST node can be somehow regarded as the most representative part of a patch, the relationship is not that strong to get the answer.
@@
 
-\textbf{Response:} We understand your concern regarding the persuasiveness of the answer to RQ1.
-We've acknowledged this limitation in our study and have addressed it in our threats to validity section.
+\vspace*{1em} \noindent \textbf{Response:} We understand your concern regarding the persuasiveness of the answer to RQ1.
+We've acknowledged this limitation in our study and our threats to validity section mentions it up front (``our code embedding approach\ldots'').
 While it is true that the relationship between the most important AST node and the most representative part of a patch may not be exceptionally strong, we designed our approach to prioritize the non-terminals that are important, and we sought to examine the connection between those important non-terminals and actual code changes.
@@
 
-In contrast to the existing embedding approach proposed for Hanam et al.~\cite{hanam2016discovering}, where they introduced the Feature Properties table leading to an order-sensitive embedding but sparse datapoints, our approach results in 11x denser datapoints (Section 5.4), which implies smaller storage requirements and better DBSCAN performance.
+In contrast to the existing embedding approach proposed by Hanam et al.~\cite{hanam2016discovering}, where they introduced the Feature Properties table leading to an order-sensitive embedding but sparse datapoints, our approach results in 11x denser datapoints (Section 5.4), which implies smaller storage requirements and better DBSCAN performance.
 The feature vector proposed for Yang et al.~\cite{yang2022mining} focuses only on single-hunk bugs and includes information such as the number of variables, lines, and function arguments.
@@
 
-\textit{\textbf{Revision:} Several patch/code/text embedding works should be discussed in the section of related work.}
+\textit{\textbf{R1:} Several patch/code/text embedding works should be discussed in the section of related work.}
 \begin{enumerate}
@@
 
-\textbf{Response:} Thanks. We have added 5 of these works to our related work section.
+\vspace*{1em} \noindent \textbf{Response:} Thanks. We have added 5 of these works to our related work section.
 
@@
 
-\textit{\textbf{Revision:} The value of the identified patterns is unclear to me.
+\textit{\textbf{R2:} The value of the identified patterns is unclear to me.
 The authors claim that they want to study bug-fix patterns in the title and in the methodology section.
@@
 
-\textbf{Response:} The new Section 4.1 (“Patterns Present and Missing”) summarizes the patterns that we found and what one could do with them.
+\vspace*{1em} \noindent \textbf{Response:} The new Section 4.1 (“Patterns Present and Missing”) summarizes the patterns that we found and what one could do with them.
 We also added an “actionability” section to some of the patterns to discuss them in further detail and how they can aid a repair tool.
@@
 
-\textit{\textbf{Revision:} Moreover, many studied patterns are actually for refactoring purposes and compiler errors (not for fixing bugs).}
+\textit{\textbf{R2:} Moreover, many studied patterns are actually for refactoring purposes and compiler errors (not for fixing bugs).}
 
-\textbf{Response:} In Section 2.3.2 we discuss three types of changes: bug-fix, fix-induced, and refactoring, following the definitions of Cotroneo et al (2019).
+\vspace*{1em} \noindent \textbf{Response:} In Section 2.3.2 we discuss three types of changes: bug-fix, fix-induced, and refactoring, following the definitions of Cotroneo et al (2019).
 Our intention is to only consider bug-fixing and fix-induced commits, and we aimed to omit changes that were purely refactoring.
@@
 
-\textit{\textbf{Revision:} It is very rare for a programmer to submit a commit that makes a Rust program uncompiled.}
+\textit{\textbf{R2:} It is very rare for a programmer to submit a commit that makes a Rust program uncompiled.}
 
-\textbf{Response:} We agree, and Section 4.1 in our discussion acknowledges this point. \\ \\
+\vspace*{1em} \noindent \textbf{Response:} We agree, and Section 4.1 in our discussion acknowledges this point. \\ \\
 
-\textit{\textbf{Revision:} Why does the first pattern in Section 3.3.3 show up in the study? The code before the patch in the first figure of Section 3.3.3 cannot be compiled.}
+\vspace*{1em} \noindent \textit{\textbf{R2:} Why does the first pattern in Section 3.3.3 show up in the study? The code before the patch in the first figure of Section 3.3.3 cannot be compiled.}
 
-\textbf{Response:} Whether the code compiles or not depends on the method signature of \verb+emit()+.
+\vspace*{1em} \noindent \textbf{Response:} Whether the code compiles or not depends on the method signature of \verb+emit()+.
 
@@
 
-\textit{\textbf{Revision:} The method for encoding patches seems problematic to me.
+\textit{\textbf{R2:} The method for encoding patches seems problematic to me.
 Giving leaf elements more weights prioritizes small patches.
@@
 
-\textbf{Response:} Let's first talk about the issue of context.
+\vspace*{1em} \noindent \textbf{Response:} Let's first talk about the issue of context.
 We agree that a bug is only a bug in a particular context.
@@
 
-\textit{The implementation described in Section 2.1 is difficult to follow.}
+\vspace*{1em} \noindent \textit{The implementation described in Section 2.1 is difficult to follow.}
 
-\textit{\textbf{Revision:} First, I don't understand how an AST is transformed to a dictionary in Section 2.1.1.}
+\vspace*{1em} \noindent \textit{\textbf{R2:} First, I don't understand how an AST is transformed to a dictionary in Section 2.1.1.}
 
-\textbf{Response:} In the paper, we tried to avoid explaining technical details of how a Syn tree which is written in Rust Debugging format is transformed to JSON format.
+\vspace*{1em} \noindent \textbf{Response:} In the paper, we tried to avoid explaining technical details of how a Syn tree which is written in Rust Debugging format is transformed to JSON format.
 JSON format and Python dictionary are basically the same format.
@@
 
-\textit{\textbf{Revision:} Second, I cannot understand Figure 1.
+\textit{\textbf{R2:} Second, I cannot understand Figure 1.
 What do the three paths represent?
@@
 
-\textbf{Response:} This is a really good question.
+\vspace*{1em} \noindent \textbf{Response:} This is a really good question.
 We added a few more sentences to make sure this doesn't confuse the readers.
@@
 
-\textit{\textbf{Revision:} Third, I don't understand the following sentence on page 7: "we record the number of occurrences of items within the paths at each scope." Overall, the implementation can be benefited from a carefully described example.}
+\vspace*{1em} \noindent \textit{\textbf{R2:} Third, I don't understand the following sentence on page 7: "we record the number of occurrences of items within the paths at each scope." Overall, the implementation can be benefited from a carefully described example.}
 
-\textbf{Response:} We revised the whole paragraph (now on page 8) and made the running example in Section 2 more explicit, as also requested by Reviewer 1.
+\vspace*{1em} \noindent \textbf{Response:} We revised the whole paragraph (now on page 8) and made the running example in Section 2 more explicit, as also requested by Reviewer 1.
 We meant that we record the number of occurrences of elements of \verb+NT+ $\cup$ \verb+BCE+ within each distinct root level scope.
@@
 
-\textit{\textbf{Revision:} 1.
+\textit{\textbf{R2:} 1.
 The introduction is organized in a very strange way.
@@
 
-\textbf{Response:} Thanks. Moved Rust introduction to the beginning. \\ \\
+\vspace*{1em} \noindent \textbf{Response:} Thanks. Moved Rust introduction to the beginning. \\ \\
 
-\textit{\textbf{Revision:} 2. Line 15 on page 7: "We found it crucial to record" $\rightarrow$ it is crucial.}
+\vspace*{1em} \noindent \textit{\textbf{R2:} 2. Line 15 on page 7: "We found it crucial to record" $\rightarrow$ it is crucial.}
 
-\textbf{Response:} fixed \\ \\
+\vspace*{1em} \noindent \textbf{Response:} fixed \\ \\
 
-\textit{\textbf{Revision:} 3.
+\vspace*{1em} \noindent \textit{\textbf{R2:} 3.
 The authors picked 50 datapoints for each cluster.
@@
 
-\textbf{Response:} Our analysis revealed many clusters (39 in total) with more than 50 datapoints.
+\vspace*{1em} \noindent \textbf{Response:} Our analysis revealed many clusters (39 in total) with more than 50 datapoints.
 More specifically, one cluster has 618 datapoints, four clusters more than 300 datapoints, and 15 clusters more than 100 datapoints.
@@
 
-\textit{\textbf{Revision:} 4.
+\vspace*{1em} \noindent \textit{\textbf{R2:} 4.
 Line 6 on page 14: "A human analyzer (paper author) compares the actual change in the code and compares it to the visual representation of the respective data point." $\rightarrow$ A human analyzer (paper author) inspects the actual change … and compares it with the visual representation …}
 
-\textbf{Response:} fixed \\ \\
+\vspace*{1em} \noindent \textbf{Response:} fixed \\ \\
 
-\textit{\textbf{Revision:} 5. What are the collected samples in the third paragraph of page 14 used for?}
+\vspace*{1em} \noindent \textit{\textbf{R2:} 5. What are the collected samples in the third paragraph of page 14 used for?}
 
-\textbf{Response:} We proposed a method to embed key AST information in a fixed-sized vector using a semi-automatically derived weighting scheme.
+\vspace*{1em} \noindent \textbf{Response:} We proposed a method to embed key AST information in a fixed-sized vector using a semi-automatically derived weighting scheme.
 RQ1 asks how successful we've been in terms of giving more values to the important elements in our ASTs.
@@
 
-\textit{\textbf{Revision:} 6. I don't understand the sentence at line 25 of page 23. Particularly, what string formatting methods are gone through if using p instead of *p?}
+\vspace*{1em} \noindent \textit{\textbf{R2:} 6. I don't understand the sentence at line 25 of page 23. Particularly, what string formatting methods are gone through if using p instead of *p?}
 
-\textbf{Response:} fixed, added Clippy reference in the paper.
+\vspace*{1em} \noindent \textbf{Response:} fixed, added Clippy reference in the paper.
 

### 2023-06-22 response.tex
tweaks and cover letter

@@
     \item Leaf nodes and penalizing large diffs: any clustering approach is going to have trouble clustering together large diffs, which have more ways to differ from each other than small diffs (\ref{rev:2:embedding}).
-    \item Picking 50 datapoints: manually evaluating large clusters is infeasible---e.g. we do not have the resources to evaluate the cluster with 618 datapoints manually (\ref{rev:2:minor}).
+    \item Picking 50 datapoints: manually evaluating large clusters is infeasible---e.g. we do not have the resources to evaluate clusters with order-of-618 datapoints manually (\ref{rev:2:minor}).
 \end{itemize}
@@
 We wholeheartedly agree that it is important to show that the identified patterns are valuable.
-We have added Section 4.1, “Patterns Present and Missing” (specifically discussing the value of the top 10 patterns and summarizing our learnings) and Section 4.2 “Usefulness of Rust Bug Patterns” (discussing the general usefulness of patterns), as well as paragraphs on actionability to 6 of the groups of patterns.
+We have added an all-new Section 4.1, “Patterns Present and Missing” (specifically discussing the value of the top 10 patterns and summarizing our learnings) and mostly-new content in Section 4.2 “Usefulness of Rust Bug Patterns” (discussing the general usefulness of patterns), as well as paragraphs on actionability to 6 of the groups of patterns.
 

### 2023-10-31 response.tex
EMSE revision#2 - reviewer comments added to response.tex, todos added, two of the raised minor issues addressed and fixed

--- 2023-06-21 response.tex
@@
 \usepackage{hyperref}
+\usepackage{todonotes}
 
 \title{Response to Reviewer Comments}
-\date{2023 June}
+\date{2023 November}
 \author{Mohammad Robati Shirzad and Patrick Lam}
@@
 
-We present our detailed responses to major revisions received on 27th March 2023 for the paper titled ``A Study of Common Bug Fix Patterns in Rust''.
+Please find below our response to the reviews for our EMSE submission ``A Study of Common Bug Fix Patterns in Rust''.
 
-\section{Introduction}
+\section{Reviewer 1}
 
-We thank the reviewers for the thoughtful reviews. We first provide summaries here, followed by detailed responses below.
+\textit{I appreciate the authors' efforts in improving the paper, and most of my concerns have been addressed.}
 
-\subsection{Methodology \& Experiments}
+\subsection{Reorganizing related work}
 
-We further explain our methodology and discuss why we made the choices we did.
-We summarize here, but invite the reviewers to read our full responses.
-We hope that the reviewers can better understand why we believe our choices to be appropriate.
-\\
+\textit{\textbf{R1:} I have observed that the related work now includes several subsections focusing on APR and patch correctness prediction.
+It would be beneficial to reorganize and consolidate these subsections into a single coherent section.
+Furthermore, the section on related work should encompass studies related to bug classification in programming languages with a longer history.}
 
-Reviewer 1:
-\begin{itemize}
-    \item The order information: our dataset is not large enough to meaningfully incorporate it (\ref{rev:1:order}).
-    \item Ignoring terminal tokens: including information about terminals would have led to dimensionality in the millions, which would be infeasible in terms of computational complexity (\ref{rev:1:terminal}).
-    \item RQ1: We have included qualitative comparisons of our approach with previous approaches and discussed how our approach is superior to those approaches. (\ref{rev:1:rq1}).
-\end{itemize}
+\vspace*{1em} \noindent \textbf{Response:} \todo{Probably no need for argument, let's just do what he asks.}
 
-Reviewer 2:
-\begin{itemize}
-    \item Context: we believe our results show that context is not required to get meaningful results (\ref{rev:2:embedding}).
-    \item Leaf nodes and penalizing large diffs: any clustering approach is going to have trouble clustering together large diffs, which have more ways to differ from each other than small diffs (\ref{rev:2:embedding}).
-    \item Picking 50 datapoints: e.g. we do not have the resources to evaluate the cluster with 618 datapoints manually (\ref{rev:2:minor}).
-\end{itemize}
+\subsection{More discussion on the impact of the evolving Rust specification}
 
-In summary, we believe that we have made the best possible design decisions given limitations in datasets and resources available to us, and that our results are strong enough that they support our decisions.
+\textit{\textbf{R1:} Additionally, I would appreciate it if the authors could discuss the potential impact of the evolving Rust specification.
+Given that Rust is a novel programming language, it is subject to rapid grammatical changes, which in turn could influence bug characteristics and symptoms.}
 
-\subsection{Discussion}
+\vspace*{1em} \noindent \textbf{Response:} \todo{Probably no need for argument, let's just do what he asks.}
 
-We wholeheartedly agree that it is important to show that the identified patterns are valuable.
-We have added Section 4.1, “Patterns Present and Missing” (specifically discussing the value of the top 10 patterns and summarizing our learnings) and Section 4.2 “Usefulness of Rust Bug Patterns” (discussing the general usefulness of patterns), as well as paragraphs on actionability to 6 of the groups of patterns.
+\section{Reviewer 2}
 
-\subsection{Clarity}
+\textit{Thanks a lot for the information and clarification in the rebuttal and the new paper version. I would suggest the authors make the following changes in the next paper version.}
 
-Thanks for the suggestions!
-We have added many of the suggested changes to the text of the paper.
-We have explained some of the reviewer questions in this response.
-If a reviewer thinks that those responses should be in the paper itself, or requires further clarification, we are happy to oblige.
+\subsection{More discussion on the size and type of the patches}
 
-\section{Detailed Responses}
+\textit{\textbf{R2:} Inspired by the last paragraph of Section 4.1, can authors discuss which of the following is more important for Rust in the next paper version: fixing bugs and fixing compilation errors?}
 
-\subsection{Reviewer 1}
+\vspace*{1em} \noindent \textbf{Response:} \todo{Need some argument here and some changes in the text.}
 
-\textit{This paper conducts a study on bug-fix patterns of the language Rust.
-Rust is a system-level programming language that has attracted huge attention among developers.
-Rust is specially designed to help developers to avoid illegal memory operations which are easily introduced in C/C++ programs.
-It has several particular language features, so the bugs in Rust may also have subtle differences from other programming languages.
-Thus it is important to study the bug's unique features and trends in Rust.
-To perform the study, the authors collect bugs and their corresponding patches from GitHub.
-Then they embed the patches to vectors that have the same length, and cluster the vectors.
-At last, they manually analyze the groups and extract fix patterns.
-This paper is very well-written and organized.
-The scale of this study is sufficient and the workflow is overall reasonable.
-In addition, this paper firstly studies Rust fix patterns, and it proposes a novel embedding method for Rust program diffs (i.e., patches).
-To this end, I think the significance of this paper is guaranteed.However, I would like to mark this paper as a major revision, with the following concerns.}
+\vspace*{1em} \noindent \textit{\textbf{R2:} Please discuss why identified patches are all small in the next paper version.
+Basically, please add the last paragraph of Section 2.2.4 of the rebuttal to the next paper version.}
 
-\subsubsection{\label{rev:1:order}The order information}
+\vspace*{1em} \noindent \textbf{Response:} \todo{Probably just do what he asks.}
 
-\textit{\textbf{Revision:} The embedding lacks order information, which could be simply appended a vector with a sequence representing the occurrence list of the AST nodes.}
+\subsection{Paraphrasing and providing clearer explanation}
 
-\textbf{Response:} Yes, our embedding lacks order information in the dimensions.
-We do not believe that order information would be useful here.
-Our application is bug categorization, and previous works on bug categorization obtained reasonable results without ordering.
-If we wanted to include ordering, we would need a neural model, but our dataset is not large enough to train a neural model that can fully understand the structures of Rust programs.
+\textit{\textbf{R2:} In Section 2.3.2, the authors use sentences from an existing paper directly (although quotes are added).
+I don't think this is a good way to write an academic paper.
+Please paraphrase these sentences.}
 
-To elaborate: simply appending a vector representing the occurrence list of AST nodes wouldn't be compatible with DBSCAN, which requires fixed-sized inputs for clustering.
-AST paths are inherently variable-sized.
+\vspace*{1em} \noindent \textbf{Response:} \todo{No argument, just paraphrase.}
 
-code2vec~\cite{alon2019code2vec} does manage to create an order-sensitive embedding.
-However, code2vec uses neural networks to generate embeddings by training them on a 30GB dataset.
-Its goal is to predict certain code elements, such as method names.
-That is not our goal.
+\vspace*{1em} \noindent \textit{\textbf{R2: } I still do not understand Figure 1. I suggest the authors label ASTDiff nodes with code to ease the explanation.}
 
-To apply a code2vec-style approach, we would need to create a dictionary of all possible paths within an AST, and then provide the neural network with observed paths from our target programming language.
+\vspace*{1em} \noindent \textbf{Response:} \todo{First, we need some changes in paper text to make it clearer. Then, need some explanation here how we've changed it.}
 
-We instead chose to follow previous bug categorization work (for JavaScript~\cite{hanam2016discovering} and Python~\cite{yang2022mining}), and simply use a feature vector to represent code differences, rather than neural-based embeddings.
-Earlier researchers chose feature vectors for similar reasons as we did: the nature of the bug categorization task, and dataset limitations.
-\\ \\
+\subsection{Minor issues}
 
-\subsubsection{\label{rev:1:terminal}Ignoring the terminal tokens}
+\textit{\textbf{R2:} Line 21 on page 8: work around -> workaround}
 
-\textit{\textbf{Revision:} In addition, ignoring the terminal tokens has advantages and disadvantages, and it is better to evaluate by experiments. }
+\vspace*{1em} \noindent \textbf{Response:} Fixed.
 
-\textbf{Response:} We appreciate the value of empirically validating design decisions when possible.
-However, in our case, not ignoring terminal tokens (variable names and constant values) would have been impractical, due to the high dimensionality (millions) of the resulting embedding vectors.
-The cost of processing such data would be prohibitive in both time and space.
-We judged that the likelihood of improvements was unlikely to justify the significant engineering effort required to get DBSCAN (which is $O(nlogn)$) to work with that magnitude of sparse data.
+\vspace*{1em} \noindent \textit{\textbf{R2:} Line 49 on page 8: which path is colored in purple in figure 1?}
 
-code2vec-style neural embedding could mitigate the high dimensionality; we could create a separate dictionary specifically for variable names (terminals) and generate embeddings by concatenating values from both the variable names dictionary and the paths dictionary.
-This way, the resulting embedding would incorporate knowledge from both nonterminal and terminal elements of the AST.
-However, as discussed above, we chose a feature vector rather than a neural embedding, following previous work, because a neural embedding would be overkill for bug categorization and because of dataset limitations.
-\\ \\
-
-\subsubsection{\label{rev:1:patch}Illustrating patch to vector transformation}
-
-\textit{\textbf{Revision:} It is better to add an example to illustrate the procedure of transforming a patch to a vector, i.e., use an example patch in Section 2 to link all the processing procedures.}
-
-\textbf{Response:} Thanks.
-We have made the example in Section 2 much more explicit and added some previously-missing steps to Figure 2 (namely the essence of change vector) as well as text linking Figures 1 and 2.
-Section 2 now contains a complete running example of transforming a patch to a vector.
-This example is in Sections 2.1.2 and 2.1.3.
-We describe the example in paragraphs labelled “Running example” throughout these subsections.
-We are happy to pull the motivating example into its own subsection if the reviewer thinks that is most appropriate.
-\\ \\
-
-\subsubsection{\label{rev:1:conclusion}Conclusion of the patterns}
-
-\textit{\textbf{Revision:} The highlighted conclusion of the pattern is missing. I expect to know which part is more likely to be buggy and want to know several typical examples.}
-
-\textbf{Response:} The new Section 4.1 (``Patterns Present and Missing'') summarizes the patterns that we found and their overall usefulness.
-We also added an “actionability” section to eight of the patterns to discuss them in further detail and how they can aid a repair tool.
-\\ \\
-
-\subsubsection{\label{rev:1:compare}Comparing the discovered patterns to other languages}
-
-\textit{\textbf{Revision:} Also, I expect to see the bug patterns that are unique in Rust programs, and further discussions about the differences with other languages like C, C++, Python and Java. }
-
-\textbf{Response:} The BC-related patterns are unique to Rust programs because they concern Rust's unique borrow checker.
-On the other hand, 11 of the 12 general patterns (excluding G.option) have counterparts in other widely-used programming languages, and we discuss differences in the text.
-
-Specifically, we have added content to Section 3.2.1 about similarities and differences between the Rust attribute bug pattern and Python decorators/Java annotations.
-Section 3.2.5 relates Rust traits to Java interfaces and shows an example of how traits are more complicated.
-Section 3.2.6 explains how Rust's match generalizes C-style switch statements.
-\\ \\
-
-\subsubsection{\label{rev:1:rq1}Concerns about RQ1}
-
-\textit{\textbf{Revision:} The answer to RQ1 is not persuasive.
-Although the most important AST node can be somehow regarded as the most representative part of a patch, the relationship is not that strong to get the answer.
-It is better to compare with existing approaches and design a better metric to measure the performance.}
-
-\textbf{Response:} We understand your concern regarding the persuasiveness of the answer to RQ1.
-We've acknowledged this limitation in our study and have addressed it in our threats to validity section.
-While it is true that the relationship between the most important AST node and the most representative part of a patch may not be exceptionally strong, we designed our approach to prioritize the non-terminals that are important, and we sought to examine the connection between those important non-terminals and actual code changes.
-
-We'd like to point out that previous work, such as the bug pattern studies for JavaScript~\cite{hanam2016discovering} and Python~\cite{yang2022mining}, each only introduced one embedding method to compute code change vectors that produced meaningful clusters, and there was no comparison between different embedding methods in those works.
-
-In contrast to the existing embedding approach proposed for Hanam et al.~\cite{hanam2016discovering}, where they introduced the Feature Properties table leading to an order-sensitive embedding but sparse datapoints, our approach results in 11x denser datapoints (Section 5.4), which implies smaller storage requirements and better DBSCAN performance.
-The feature vector proposed for Yang et al.~\cite{yang2022mining} focuses only on single-hunk bugs and includes information such as the number of variables, lines, and function arguments.
-Our approach is more general: it handles multiple-hunk commits.
-
-In our RQ1 experiment, we opted for a manual inspection process where the two authors, who possess Rust knowledge, reviewed the vector and its corresponding code change to verify representativeness.
-While this manual approach may not be as statistically rigorous as an automated inspection (because it inspects smaller numbers of points), we considered it to be an intuitive and straightforward method.
-
-Of course, an automated inspection could examine more code examples.
-Such an inspection would also be subject to the validity threat of embedding assumptions about what to inspect.
-And, developing an automated code comprehension tool specifically for Rust, capable of interpreting code changes and providing explanations for the reasons behind those changes, poses a significant challenge.
-Such a task would itself require a separate and comprehensive study.
-We believe that developing an automated inspection tool would be a distraction and would not contribute to answering the core questions about Rust bug patterns behind this work.
-
-We further emphasize that our code embedding approach is only the initial step in our bug categorization pipeline.
-We had to choose a specific embedding method to create our code diff vectors: different embedding approaches would lead to entirely different datasets and possibly different clustering outcomes.
-Each embedding approach would require separate manual analysis of the resulting clusters.
-Consequently, we made the decision to select one embedding method and to validate it.
-
-To additionally justify our choice of embedding approach, we evaluated whether the clusters generated by our approach contained dissimilar data points, as this was a critical factor in determining its suitability.
-Our bug categories demonstrate that our clusters successfully encompass cross-project bugs, and this can be verified by examining the introduced code changes in our replication package.
-\\ \\
-
-\subsubsection{\label{rev:1:bib}Adding more bibliography}
-
-\textit{\textbf{Revision:} Several patch/code/text embedding works should be discussed in the section of related work.}
-\begin{enumerate}
-    \item Automated Classification of Overfitting Patches with Statically Extracted Code Features. (TSE 2019)
-    \item Is this Change the Answer to that Problem? Correlating Descriptions of Bug and Code Changes for Evaluating Patch Correctness. (ASE 2019)
-    \item Context-Aware Code Change Embedding for Better Patch Correctness Assessment. (TOSEM 2022)
-    \item CC2Vec: Distributed Representations of Code Changes. (ICSE 2020)
-    \item code2seq: Generating Sequences from Structured Representations of Code. (ICLR 2019)
-    \item BERT: Pre-training of Deep Bidirectional Transformers for Language Understanding.
-\end{enumerate}
-
-\textbf{Response:} Thanks. We have added 5 of these works to our related work section.
-
-\subsection{Reviewer 2}
-
-\textit{In this paper, the authors aim to mine and study bug-fix patterns in Rust programs.
-They discuss a new encoding method for Rust patches, which changes a patch into a vector.
-Then they apply a clustering algorithm to identify common patch patterns.
-In the end, they discuss 20 identified patterns.
-Rust is a new, popular programming language, and studying its bug-fix patterns is important.
-However, I have some concerns about the methodology and the value of the identified patterns.
-I discuss my concerns as follows.} \\ \\
-
-\subsubsection{\label{rev:2:value}The value of the identified patterns}
-
-\textit{\textbf{Revision:} The value of the identified patterns is unclear to me.
-The authors claim that they want to study bug-fix patterns in the title and in the methodology section.
-However, it is difficult for me to correlate the discussed patterns with any particular bug types.
-I cannot figure out any actions (e.g., how to build bug detection tools, how to inspect my Rust code) to take after reading the paper.}
-
-\textbf{Response:} The new Section 4.1 (“Patterns Present and Missing”) summarizes the patterns that we found and what one could do with them.
-We also added an “actionability” section to some of the patterns to discuss them in further detail and how they can aid a repair tool.
-We found that many of the patterns are interesting but still require further work to distill into a repair tool.
-\\ \\
-
-\subsubsection{\label{rev:2:refactoring}Refactoring, not bug fixing}
-
-\textit{\textbf{Revision:} Moreover, many studied patterns are actually for refactoring purposes and compiler errors (not for fixing bugs).}
-
-\textbf{Response:} In Section 2.3.2 we discuss three types of changes: bug-fix, fix-induced, and refactoring, following the definitions of Cotroneo et al (2019).
-Our intention is to only consider bug-fixing and fix-induced commits, and we aimed to omit changes that were purely refactoring.
-However, some of the changes that we included may appear to be refactoring changes but actually have performance implications.
-We chose to specifically include those.
-For instance, \verb+BC.clone.ref+ does not alter the program's behaviour (no additional test cases pass) but can reduce CPU usage.
-Other changes, like adding Option to a type, can be considered refactorings, but are in support of bug fixes, and hence are fix-induced.
-\\ \\
-
-\subsubsection{\label{rev:2:compile}Committing code with compiler errors}
-
-\textit{\textbf{Revision:} It is very rare for a programmer to submit a commit that makes a Rust program uncompiled.}
-
-\textbf{Response:} We agree, and Section 4.1 in our discussion acknowledges this point. \\ \\
-
-\textit{\textbf{Revision:} Why does the first pattern in Section 3.3.3 show up in the study? The code before the patch in the first figure of Section 3.3.3 cannot be compiled.}
-
-\textbf{Response:} Whether the code compiles or not depends on the method signature of \verb+emit()+.
-
-If \verb+emit()+ is like this:
-    \verb+fn emit(&self, db: &DiagnosticBuilder);+
-Then \verb+mut+  is not necessary.
-
-If \verb+emit()+ is like this:
-    \verb+fn emit(&mut self, db: &DiagnosticBuilder);+
-Then the presence of \verb+mut+ is mandatory in the closure.
-
-In the example mentioned here, the presence of \verb+mut+ in the closure parameter \verb+|mut e|+ becomes necessary following a modification of \verb+emit()+ to require a mutable reference (\verb+&mut self+).
-The patch that we are classifying simultaneously changes \verb+emit()+ and the reference to add \verb+mut+ in both places.
-\\ \\
-
-\subsubsection{\label{rev:2:embedding}Our embedding method}
-
-\textit{\textbf{Revision:} The method for encoding patches seems problematic to me.
-Giving leaf elements more weights prioritizes small patches.
-In section 3, all discussed patches are small patches without any information about what is the bug or the program context.
-My understanding is that almost all code in Section 3 before the patch is good code.
-They become buggy because they are used in particular contexts.
-Combining the patches together with the program contexts is more valuable.}
-
-\textbf{Response:} Let's first talk about the issue of context.
-We agree that a bug is only a bug in a particular context.
-We understand that the reviewer is saying: had we considered the statements around a bug fix, then we would have gotten better results than just considering the bug fix itself in isolation.
-
-We point out that Ruxanne considers changes that developers have labelled as bug-fixing changes (or fix-induced changes).
-Part of our thesis is that, given a change labelled as a bug-fixing change, the AST diff is enough to permit classification of that bug, and that context is not required.
-We believe that our results show that it is indeed enough: we identified 12 general patterns and 8 borrow-checker patterns based on the information we collected.
-Section 3 provides links to the changes, so that the reader can examine the changes themselves.
-
-We'd ask the reviewer to consider our results on their own merits, rather than thinking about a possible other approach that also could have worked.
-It is possible that more information, as the reviewer suggests, gives better results, but it also may overfit, for instance, and computations are harder with more data.
-In our response to Reviewer 1, we discussed the embedding approaches taken by the Python~\cite{yang2022mining} and JavaScript~\cite{hanam2016discovering} related works.
-
-Having said that, Ruxanne does consider the parents in the AST when reasoning about a change.
-It does not specifically consider siblings of a change in the AST, which is the context that we believe the reviewer is talking about.
-For instance, in Figure 1, the \verb+head.clone()+ change paths include the addition of the method call node to the path containing the nonterminal \verb+head+, as well as the addition of the path containing the \verb+clone()+ call.
-We consider both of those paths from root to leaf, shown as Paths 2 and 3 in the Figure.
-
-As for the comment about small patches and leaf nodes, we'd like to clarify the role of the weights.
-We performed a preliminary step to determine fixed weights for different types of AST nodes.
-We then used these weights, uniformly across our data set, to compute essence-of-change vectors (as depicted in Figure 2).
-Both small and large patches have their occurrences vectors multiplied by the same weights, giving essence-of-change vectors.
-RQ1 validates this essence-of-change computation.
-In subsequent steps, we then cluster the essence-of-change vectors to find bug patterns.
-
-We acknowledge that it is possible that a large patch may be harder to cluster, because it has to look like other patches.
-Additionally, we believe that any clustering-based approach is likely going to have large diffs be further apart for the same reason that ours does: there's just more going on.
-But there is nothing inherently penalizing a larger diff.
-Also, we point out that the related work by Pan et al specifically excludes diffs with seven statements or more, while we allow them.
-\\ \\
-
-\subsubsection{\label{rev:2:ast}AST Transformation}
-
-\textit{The implementation described in Section 2.1 is difficult to follow.}
-
-\textit{\textbf{Revision:} First, I don't understand how an AST is transformed to a dictionary in Section 2.1.1.}
-
-\textbf{Response:} In the paper, we tried to avoid explaining technical details of how a Syn tree which is written in Rust Debugging format is transformed to JSON format.
-JSON format and Python dictionary are basically the same format.
-Two images below show an AST in Rust Debugging format and in JSON (Python Dictionary) format.
-As you can see, the difference is only in quotation marks and curly/square brackets.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=1\textwidth]{astpics}
-    \caption{Syn tree in Rust Debugging format (left) and Python dictionary format (right)}
-\end{figure}
-
-\subsubsection{\label{rev:2:path}Inconsistency between number of paths and the number of changed arguments}
-
-\textit{\textbf{Revision:} Second, I cannot understand Figure 1.
-What do the three paths represent?
-If they represent the changed parameters, why are there three paths, but there are only two parameters changed?
-If they represent parameters, how can they represent changes from the last version?}
-
-\textbf{Response:} This is a really good question.
-We added a few more sentences to make sure this doesn't confuse the readers.
-To clarify: the tree depicted in Figure 1 is not the actual AST; rather, it represents the ASTDiff (now written on the label on top left).
-In our approach, we consider two versions of the code: the buggy code and the fixed code.
-Consequently, we have two ASTs.
-The ASTDiff, shown in Figure 1, represents the difference between these two ASTs in Python dictionary format.
-The number of paths in the ASTDiff does not directly correspond to the number of changed elements in the source code.
-Instead, it is determined by the structural differences between the two ASTs of the code versions.
-The paths in the ASTDiff capture the changes in the code structure between the two versions, allowing us to analyze and extract relevant information for further processing.
-We have also added more text to Section 2.1.2 explaining this.
-
-\textit{\textbf{Revision:} Third, I don't understand the following sentence on page 7: "we record the number of occurrences of items within the paths at each scope." Overall, the implementation can be benefited from a carefully described example.}
-
-\textbf{Response:} We revised the whole paragraph (now on page 8) and made the running example in Section 2 more explicit, as also requested by Reviewer 1.
-We meant that we record the number of occurrences of elements of \verb+NT+ $\cup$ \verb+BCE+ within each distinct root level scope.
-Root level scopes can be structs, functions, impl blocks, ….
-Note that Figure 2 does not explicitly show the single root level scope for this example, which is ItemFn.
-Other diffs could contribute items in different root level scopes.
-\\
-
-\subsubsection{\label{rev:2:minor}Minor issues}
-
-\textit{\textbf{Revision:} 1.
-The introduction is organized in a very strange way.
-The authors should introduce Rust first, and then discuss studying common bug/fix patterns is important for Rust.
-However, the current writing does not mention Rust until the second page.}
-
-\textbf{Response:} Thanks. Moved Rust introduction to the beginning. \\ \\
-
-\textit{\textbf{Revision:} 2. Line 15 on page 7: "We found it crucial to record" $\rightarrow$ it is crucial.}
-
-\textbf{Response:} fixed \\ \\
-
-\textit{\textbf{Revision:} 3.
-The authors picked 50 datapoints for each cluster.
-Are there any particular reasons for choosing 50?
-There are very few clusters with more than 50 datapoints.
-A more reasonable method is to inspect all datapoints.}
-
-\textbf{Response:} Our analysis revealed many clusters (39 in total) with more than 50 datapoints.
-More specifically, one cluster has 618 datapoints, four clusters more than 300 datapoints, and 15 clusters more than 100 datapoints.
-With our number of datapoints, manual inspection would be overwhelmingly time-consuming.
-
-We believe that the choice of 50 datapoints per cluster successfully balances ensuring that the selected datapoints truly represent the same change and making manual inspection feasible.
-Our sample size allows us to get meaningful insights into the characteristics of each cluster.
-\\ \\
-
-\textit{\textbf{Revision:} 4.
-Line 6 on page 14: "A human analyzer (paper author) compares the actual change in the code and compares it to the visual representation of the respective data point." $\rightarrow$ A human analyzer (paper author) inspects the actual change … and compares it with the visual representation …}
-
-\textbf{Response:} fixed \\ \\
-
-\textit{\textbf{Revision:} 5. What are the collected samples in the third paragraph of page 14 used for?}
-
-\textbf{Response:} We proposed a method to embed key AST information in a fixed-sized vector using a semi-automatically derived weighting scheme.
-RQ1 asks how successful we've been in terms of giving more values to the important elements in our ASTs.
-For instance, if in a code patch, the only change is adding \verb+.clone()+, then we should expect the element \verb+clone+ to have the highest value.
-In RQ1, we sample 25 datapoints from General datapoints and 25 datapoints from BC-related datapoints (50 in total) to manually analyze whether the values in their code vector actually respect the important elements in the AST.
-Note that we selected 50 datapoints that minimize the sampling error, i.e.
-it is a low bias sampling.
-\\ \\
-
-\textit{\textbf{Revision:} 6. I don't understand the sentence at line 25 of page 23. Particularly, what string formatting methods are gone through if using p instead of *p?}
-
-\textbf{Response:} fixed, added Clippy reference in the paper.
-
-This is a common Clippy warning. You can find information about it through this link: \url{https://rust-lang.github.io/rust-clippy/master/index.html#inefficient_to_string} \\
-
-In the link the example is: \\
-
-\noindent\verb+// Generic implementation for `T: Display` is used (slow)+ \\
-\verb+["foo", "bar"].iter().map(|s| s.to_string());+\\
-
-\noindent\verb+// OK, the specialized impl is used+\\
-\verb+["foo", "bar"].iter().map(|&s| s.to_string());+\\
-
-It is the same as:
-
-\noindent\verb+// Generic implementation for `T: Display` is used (slow)+\\
-\verb+["foo", "bar"].iter().map(|s| s.to_string());+\\
-
-\noindent\verb+// OK, the specialized impl is used+\\
-\verb+["foo", "bar"].iter().map(|s| (*s).to_string());+\\
-
-In summary, when \verb+p+ is used and it's not clear for Clippy that \verb+p+ cannot be of type \verb+&&T+ where \verb+T+ implements the \verb+ToString+ trait, it suggests dereferencing \verb+p+.
-The reason is that \verb+p.to_string()+ will use the \verb+format!()+ macro under the hood.
-The \verb+format!()+ macro is less efficient than direct casting of \verb+&T+ to String (probably because it allocates extra memory first then populates it based on formatting annotations).
-\verb+(*p).to_string()+ will not use \verb+format!()+.
+\vspace*{1em} \noindent \textbf{Response:} It is the rightmost path. We changed the color to blue so it is clearer.
 

### 2023-10-31 methodology.tex
EMSE revision#2 - reviewer comments added to response.tex, todos added, two of the raised minor issues addressed and fixed

@@
 
-    \draw [->, purple] plot [smooth, tension=1] coordinates { (3,1) (3.5,-3) (5.5,-5) (5.5,-8.5)};
+    \draw [->, cyan] plot [smooth, tension=1] coordinates { (3,1) (3.5,-3) (5.5,-5) (5.5,-8.5)};
 
@@
 
-    \draw (13.5,-2.5) node[fill=purple!20](g1){ItemFn};
-    \draw (13.5,-3.5) node[fill=purple!20](g2){Local};
-    \draw (13.5,-4.5) node[fill=purple!20](g3){ExprCall};
-    \draw (13.5,-5.5) node[fill=purple!20](g4){ExprMethodCall};
-    \draw (13.5,-6.5) node[fill=purple!20](g5){Ident};
-    \draw (13.5,-7.5) node[fill=purple!20](g6){clone};
-    \node[draw=none, color=purple] at (13.5,-8.3) {Path 3};
+    \draw (13.5,-2.5) node[fill=cyan!20](g1){ItemFn};
+    \draw (13.5,-3.5) node[fill=cyan!20](g2){Local};
+    \draw (13.5,-4.5) node[fill=cyan!20](g3){ExprCall};
+    \draw (13.5,-5.5) node[fill=cyan!20](g4){ExprMethodCall};
+    \draw (13.5,-6.5) node[fill=cyan!20](g5){Ident};
+    \draw (13.5,-7.5) node[fill=cyan!20](g6){clone};
+    \node[draw=none, color=cyan] at (13.5,-8.3) {Path 3};
 
@@
 After the change, the second argument passes a borrow of \verb+tail+ rather than sending the ownership of \verb+tail+ to the callee; and, the third argument sends a clone of \verb+head+ to the callee instead of sending \verb+head+ itself.
-A DFS on this tree yields three different paths, which we show in red, green, and purple.
+A DFS on this tree yields three different paths, which we show in red, green, and blue.
 A path represents a sequence of involved non~terminals.
@@
 
-A reasonable work around would be to pass from the full set of non~terminals to a smaller set of categories (e.g.
+A reasonable workaround would be to pass from the full set of non~terminals to a smaller set of categories (e.g.
 a category for larger entities like class or function definitions, and a category for smaller entities like statements and expressions), and then to record all possible combinations of these categories (similar to \cite{hanam2016discovering}).
@@
 Our main requirement is that our datapoints summarize the key changes in ASTDiffs---a datapoint should foreground the most important change in a diff.
-For instance, the purple (right-most) path in Figure~\ref{fig:extraction} could be described as a change in a local variable declaration; a change in a function call; calling a method of one of the function arguments; or calling clone() on one of the function arguments.
+For instance, the blue (right-most) path in Figure~\ref{fig:extraction} could be described as a change in a local variable declaration; a change in a function call; calling a method of one of the function arguments; or calling clone() on one of the function arguments.
 The last description, in our opinion, is the most useful one, and we designed our weighting scheme to embed this value judgment.

### 2023-10-31 run.sh

@@ -11,11 +11,11 @@ rm *.log ;
 rm *.out ;
 rm *.synctex.gz ;
 pdflatex main.tex
-pdflatex cover.tex
+pdflatex response.tex
 bibtex main
 bibtex cover
 pdflatex main.tex
-pdflatex cover.tex
+pdflatex response.tex
 sleep 2;
 pdflatex main.tex
-pdflatex cover.tex
+pdflatex response.tex

### 2023-11-01 response.tex
two todos from reviewer 2 resolved

@@
 
-\vspace*{1em} \noindent \textbf{Response:} \todo{Need some argument here and some changes in the text.}
+\vspace*{1em} \noindent \textbf{Response:} Added more discussion regarding this to the last paragraph of Section 4.1.
+We discuss that we consider bugs as potentially more critical, because an unresolved bug can remain dormant within the system and unintentionally get deployed, whereas a program with compilation errors cannot be shipped by developers.
 
@@
 
-\vspace*{1em} \noindent \textbf{Response:} \todo{No argument, just paraphrase.}
+\vspace*{1em} \noindent \textbf{Response:} Paraphrased the three definition that were directly quoted.
 

### 2023-11-01 discussion.tex
two todos from reviewer 2 resolved

@@
 We thus would not observe patterns which do not make it to the repository.
+While comparing the importance of bugs and compilation errors can be challenging, we tend to consider bugs as potentially more critical.
+This is because a bug, if left unresolved, can remain dormant within the system and inadvertently get deployed, whereas a program with compilation errors cannot be shipped by developers.
 

### 2023-11-01 methodology.tex
two todos from reviewer 2 resolved

@@
 \begin{itemize}
-    \item bug-fix: ``The changed code actually fixes the behavior of software, which can represent fixes for a bug type''.
-We include here changes that improve program performance, as they fix the behaviour of the software.
+    \item bug-fix: The code change rectifies the undesired behavior in the software system, providing fixes for a specific bug type.
+We also include here changes that improve program performance, as they fix the behaviour of the software.
 
-    \item fix-induced: ``The changed code is a group of bug-fixing code changes but not representing an actual bug fix, e.g., adding new input parameters to a method, the method signature and method call must be changed correspondingly''.
+    \item fix-induced: The code change is a part of a bug-fixing code and does not constitute the whole bug fix.
+For instance, introducing additional input parameters to a method also requires adjustments in the corresponding method invocations and does not represent the complete bug fix on its own.
 
-    \item refactoring: ``the changed code does not modify the software behavior, e.g., for better readability or encapsulation''.
+    \item refactoring: The software behavior remains unchanged after the code change. These changes are made for imporving codebase's maintainability, better readability or encapsulation.
 \end{itemize}

### 2023-11-09 response.tex
reviewer 2 finished for now

@@
 
-\vspace*{1em} \noindent \textbf{Response:} \todo{Probably just do what he asks.}
+\vspace*{1em} \noindent \textbf{Response:} Added the mentioned paragraph to the Discussion seciton of the manuscript.
 
@@
 
-\vspace*{1em} \noindent \textbf{Response:} Paraphrased the three definition that were directly quoted.
+\vspace*{1em} \noindent \textbf{Response:} Paraphrased the three definitions that were directly quoted.
 
@@
 
-\vspace*{1em} \noindent \textbf{Response:} \todo{First, we need some changes in paper text to make it clearer. Then, need some explanation here how we've changed it.}
+\vspace*{1em} \noindent \textbf{Response:} Labeled ASTDiff nodes with code. \todo{Should we add more explanation?}
 

### 2023-11-09 discussion.tex
reviewer 2 finished for now

@@
 Also, our patterns are all cross-project fix patterns: each reported fix pattern has been seen at least in three different projects.
+
+While we introduce patterns that reflect changes at different locations in the code, such as \textit{BC.ref.add}, which by borrowing a variable instead of taking ownership enables users to use the variable in subsequent locations, we acknowledge that it is possible that a large patch may be harder to cluster, because it has to look like other patches.
+Additionally, we believe that any clustering-based approach is likely going to have large diffs be further apart for the same reason that ours does: there is just more going on.
+Nevertheless, there is nothing inherently penalizing a larger diff.
+Also, we point out that the related work by~\cite{pan2009toward} specifically excludes diffs with seven statements or more, while we allow them.
 

### 2023-11-09 discussion.tex
reviewer 1, todo 2 fixed

@@
 Similarly, a commit message might not contain our target keywords while the commit is associated with bug fixing changes.
+Lastly, Rust is a relatively new programming language, and its syntax is prone to evolving over time.
+Changes in the Rust's grammar, such as the introduction of new keywords like \verb+async/await+ in version 1.39, have the potential to introduce new bugs or modify existing ones.
+Thus, it is important to consider the language evolution as a potential external threat to the validity of our research findings.

### 2023-11-10 related.tex
text added for reviewer 1 comment 1

@@
 Thus, having a collection of common bug fix patterns is a crucial step for developing such repair tools.
-\cite{pan2009toward} analyzed seven large-scale widely used Java projects and obtained 27 common bug fix patterns.
+Studying bug fix patterns can be traced back to 1975 when Enders et al introduced a classification of bugs that occur in DOS/VS operating system.
+This work contends that categorizing bug patterns can aid in uncovering error causes, fixing them, and preventing their recurrence.
+Flanagan et al.
+proposed MrSpidey, an interactive static debugger for Scheme (a dialect of Lisp).
+MrSpidey allows the programmer to browse program invariants and their derivations over error prone patterns (unsafe operators).
+The user can determine whether the error will actually occur, or if obtaining the correctness proof is unattainable due to the tool's limitations.
+
+More recently, \cite{pan2009toward} analyzed seven large-scale widely used Java projects and obtained 27 common bug fix patterns.
 They used a bug fix pattern extractor tool to automatically parse and detect bug patterns within the bug fix hunks with less than seven statements.

### 2023-11-10 response.tex
reviewer 1, todo 2 fixed
added cites for the previous commit

@@
 
-\vspace*{1em} \noindent \textbf{Response:} \todo{Probably no need for argument, let's just do what he asks.}
+\vspace*{1em} \noindent \textbf{Response:} Included two research studies in the list of related works.
+The first is a 1975 study conducted by Enders et al.
+\cite{endres1975analysis}, which examines common error patterns in the DOS/VS operating system.
+The second is a study by Flanagan et al.
+\cite{flanagan1998new}, where they introduce a static debugger for Scheme, a Lisp dialect, centered around the error-prone operators in the language.
+\todo{How should we consolidate those subsections?}
 
@@
 
-\vspace*{1em} \noindent \textbf{Response:} \todo{Probably no need for argument, let's just do what he asks.}
+\vspace*{1em} \noindent \textbf{Response:} A brief discussion added to the threats to external validity list.
 

### 2023-11-10 related.tex
added cites for the previous commit

@@
 Thus, having a collection of common bug fix patterns is a crucial step for developing such repair tools.
-Studying bug fix patterns can be traced back to 1975 when Enders et al introduced a classification of bugs that occur in DOS/VS operating system.
+Studying bug fix patterns can be traced back to 1975 when~\cite{endres1975analysis} introduced a classification of bugs that occur in DOS/VS operating system.
 This work contends that categorizing bug patterns can aid in uncovering error causes, fixing them, and preventing their recurrence.
-Flanagan et al.
-proposed MrSpidey, an interactive static debugger for Scheme (a dialect of Lisp).
+\cite{flanagan1998new} proposed MrSpidey, an interactive static debugger for Scheme (a dialect of Lisp).
 MrSpidey allows the programmer to browse program invariants and their derivations over error prone patterns (unsafe operators).

### 2023-11-10 run.sh

@@ -13,7 +13,7 @@ rm *.synctex.gz ;
 pdflatex main.tex
 pdflatex response.tex
 bibtex main
-bibtex cover
+bibtex response
 pdflatex main.tex
 pdflatex response.tex
 sleep 2;

### 2023-11-11 response.tex
edit moh's changes

--- 2023-11-09 response.tex
@@
 
+\vspace*{1em}
+Thanks!
+
 \subsection{Reorganizing related work}
@@
 
-\vspace*{1em} \noindent \textbf{Response:} \todo{Probably no need for argument, let's just do what he asks.}
+\vspace*{1em} \noindent \textbf{Response:} We have reorganized the former Sections 5.1 through 5.3 into the new Section 5.1, adding a roadmap of that subsection at its beginning.
+
+We also added a paragraph on bug classification at the beginning of Section 5.1.
+It adds three new references.
+The first is a 1975 study conducted by Enders et al.
+\cite{endres1975analysis}, which examines common error patterns in the DOS/VS operating system (written in the DOS Macro Assembler langage).
+The second study is by Knuth~\cite{knuth89:_error_tex}, surveying errors that he added in the development of \TeX.
+Finally, we added a 1998 study by Flanagan et al.
+\cite{flanagan1998new}, where they introduce a static debugger for Scheme (a Lisp dialect), centered around the error-prone operators in the language.
+Additionally, we moved the initial citation of Pan et al~\cite{pan2009toward} up to this paragraph; it studies bugs in seven large-scale Java programs.
 
@@
 
-\vspace*{1em} \noindent \textbf{Response:} \todo{Probably no need for argument, let's just do what he asks.}
+\vspace*{1em} \noindent \textbf{Response:} Yes, the 2022 Rust roadmap\footnote{\url{https://blog.rust-lang.org/inside-rust/2022/04/04/lang-roadmap-2024.html}} includes future proposed changes to Rust.
+It is indeed a threat to external validity, and we have now discussed it in our threats to external validity (Section 4.3); it could be mitigated by refreshing the analysis results once new versions of Rust are broadly used by extant packages.
 
@@
 
-\vspace*{1em} \noindent \textbf{Response:} Added more discussion regarding this to the last paragraph of Section 4.1.
-We discuss that we consider bugs as potentially more critical, because an unresolved bug can remain dormant within the system and unintentionally get deployed, whereas a program with compilation errors cannot be shipped by developers.
+\vspace*{1em} \noindent \textbf{Response:} We added a new paragraph at the end of Section 4.1.
+We consider bugs as potentially more critical, because an unresolved bug can remain dormant within the system and unintentionally get deployed, whereas a program with compilation errors cannot be shipped by developers.
 
@@
 
-\vspace*{1em} \noindent \textbf{Response:} Added the mentioned paragraph to the Discussion seciton of the manuscript.
+\vspace*{1em} \noindent \textbf{Response:} Added the mentioned paragraph to the Discussion section of the manuscript (third paragraph of Section 4).
 
@@
 
-\vspace*{1em} \noindent \textbf{Response:} Labeled ASTDiff nodes with code. \todo{Should we add more explanation?}
+\vspace*{1em} \noindent \textbf{Response:} We have added the labels {\tt \&tail} and {\tt head.clone()} at the relevant ASTDiff nodes on the tree on the left.
+The nonterminals themselves don't directly correspond to code---it would be the terminals which are not present in the paths.
+We have also added more text to the caption of Figure 1 and point at the ExprReference in Path 1 in the text.
 
@@
 
-\vspace*{1em} \noindent \textbf{Response:} It is the rightmost path. We changed the color to blue so it is clearer.
+\vspace*{1em} \noindent \textbf{Response:} It is the rightmost path. We changed the color to blue so it is clearer and added a reference to ``Path 3'' as in the Figure.
 

### 2023-11-11 discussion.tex
edit moh's changes

--- 2023-11-01 discussion.tex
@@
 Also, our patterns are all cross-project fix patterns: each reported fix pattern has been seen at least in three different projects.
+
+While we introduce patterns that reflect changes at different locations in the code, such as \textit{BC.ref.add}, which by borrowing a variable instead of taking ownership enables users to use the variable in subsequent locations, we acknowledge that it is possible that a large patch may be harder to cluster, because it has to resemble other patches.
+Additionally, we believe that any clustering-based approach will likely place larger diffs further apart for the same reason that ours does: there is just more going on.
+Nevertheless, there is nothing inherently penalizing a larger diff.
+Also, we point out that the related work by~\cite{pan2009toward} specifically excludes diffs with seven statements or more, while we allow larger diffs.
 
@@
 We thus would not observe patterns which do not make it to the repository.
+
 While comparing the importance of bugs and compilation errors can be challenging, we tend to consider bugs as potentially more critical.
@@
 
-A threat to external validity is selection bias, where the benchmarks differ systematically from the actual world.
+We discuss three threats to external validity: selection bias, incorrect commit/bug information reported by developers, and changes to the Rust programming language.
+Selection bias is an issue because our benchmarks may differ systematically from the set of Rust programs in the world.
 Our selection of projects might be biased and therefore we might not have presented patterns that may exist in other projects.
-Another threat to external validity can be the developers reporting the commits as bug fixing commits while they are not really fixing functionality.
+Another threat to external validity can be the developers reporting commits as bug fixing commits while they are not really fixing functionality.
 Similarly, a commit message might not contain our target keywords while the commit is associated with bug fixing changes.
+A third threat is that Rust is a relatively new programming language, and its syntax is prone to evolving over time.
+Changes in Rust's grammar, such as the introduction of new keywords like \verb+async/await+ in version 1.39, have the potential to introduce new bug categories or modify existing ones.
+Thus, one must consider language evolution as a potential external threat to the validity of our research findings; this threat can be mitigated by redoing our analysis in the future using our methodology with updated versions of our published artifacts.

### 2023-11-11 methodology.tex
reviewer 2 finished for now
edit moh's changes

@@
 
+    \draw (0.8,-4.45) node[fill=yellow!20, minimum height=0.2cm](b0){\verb+&tail+};
     \draw (0,-5) node[](b1){ExprReference};
@@
 
+    \draw (4,-4.45) node[fill=yellow!20, minimum height=0.2cm](b0){\verb+head.clone()+};
     \draw (3,-5) node[](c1){ExprMethodCall};
@@
 
-    \draw [->, green] plot [smooth, tension=1] coordinates { (2.5,1) (3,-3) (5,-5) (4,-6.5) (3.7,-8.5)};
+    \draw [->, green] plot [smooth, tension=1] coordinates { (2.5,1) (3,-3) (5.2,-4.2) (4,-6.5) (3.7,-8.5)};
 
-    \draw [->, cyan] plot [smooth, tension=1] coordinates { (3,1) (3.5,-3) (5.5,-5) (5.5,-8.5)};
+    \draw [->, cyan] plot [smooth, tension=1] coordinates { (3,1) (3.5,-3) (5.5,-4.2) (5.5,-8.5)};
 
@@
 
-    \caption{\label{fig:extraction}The ASTDiff (left) corresponds to the code change in the yellow box (top right). A DFS will identify three paths, shown in three different colors at right.}
+    \caption{\label{fig:extraction}The ASTDiff (left) corresponds to the code change in the yellow box (top right).
+A DFS of the ASTDiff finds three paths of nonterminals (Paths 1--3), shown in three different colors at right.
+The first path summarizes the addition of the ExprReference {\tt \&} before tail, while the second and third paths indicate the change from a read of \texttt{head} to a method call, and a call to \texttt{clone()}, respectively.}
 
@@
 After the change, the second argument passes a borrow of \verb+tail+ rather than sending the ownership of \verb+tail+ to the callee; and, the third argument sends a clone of \verb+head+ to the callee instead of sending \verb+head+ itself.
-A DFS on this tree yields three different paths, which we show in red, green, and blue.
-A path represents a sequence of involved non~terminals.
+A DFS on this tree yields three different paths, which we label ``Path 1'' through ``Path 3'' and show in red, green, and blue.
+A path represents a sequence of involved non~terminals, e.g.
+the one starting with ExprReference in Path 1.
 It is crucial to record the order of nodes in the path, so that we can record that order in our datapoint embedding.
@@
 The number of paths in the ASTDiff does not directly correspond to the number of changed elements in the source code.
-Instead, it is determined by the structural differences between the two ASTs of the code versions.
+Instead, it is determined by the structural differences between the ASTs of the two code versions.
 
@@
 Our main requirement is that our datapoints summarize the key changes in ASTDiffs---a datapoint should foreground the most important change in a diff.
-For instance, the blue (right-most) path in Figure~\ref{fig:extraction} could be described as a change in a local variable declaration; a change in a function call; calling a method of one of the function arguments; or calling clone() on one of the function arguments.
+For instance, the blue path (Path 3) in Figure~\ref{fig:extraction} could be described as a change in a local variable declaration; a change in a function call; calling a method of one of the function arguments; or calling clone() on one of the function arguments.
 The last description, in our opinion, is the most useful one, and we designed our weighting scheme to embed this value judgment.
@@
 
-    \item refactoring: The software behavior remains unchanged after the code change. These changes are made for imporving codebase's maintainability, better readability or encapsulation.
+    \item refactoring: The software behavior remains unchanged after the code change. These changes are made for improving the codebase's maintainability, readability, or encapsulation.
 \end{itemize}

### 2023-11-11 related.tex
edit moh's changes

--- 2023-06-14 related.tex
@@
-Automated program repair aims to debug faulty source code without human involvement, sometimes using test cases to guide the repair.
+Automated program repair (APR) aims to debug faulty source code without human involvement, sometimes using test cases to guide the repair.
 Before a repair tool modifies code, it will typically use a fault localization module to find fault locations.
@@
 
-Researchers have proposed different methods for code repair, each of which can be said to follow a different mindset.
-Following the taxonomy proposed by \cite{liu2018survey}, we briefly present two mindsets, and will discuss how having a set of predefined fix patterns can aid repair.
+\subsection{Automated Program Repair}
+We next discuss a number of approaches for automated program repair.
+First, however, we discuss previous work that aimed to classify bugs---program repair must have a model of what it is attempting to repair in the first place.
+We then present two mindsets for program repair: search-based and pattern-based.
+Finally, we survey related work in automated patch correctness assessment, which aims to ensure that the generated changes are valid.
 
-\subsection{Search-Based Program Repair}
-One line of research in program repair follows the competent programmer hypothesis~\citep{gopinath2014mutant}: syntactically, a faulty program is not that far away from its correct version.
+\paragraph{Bug Classification.}
+A foundational bug study dates back to~\cite{endres1975analysis}, which classified the bugs that were fixed in a particular release of IBM's DOS/VS operating system, written in the DOS Macro Assembler language.
+Endres contended that categorizing bugs can aid in uncovering error causes, fixing them, and preventing their recurrence.
+The Group B of bugs in that work include, among others, bugs that could be prevented by the use of a different programming language, e.g.
+``assignment, loading, or saving of address registers forgotten,'' which is somewhat similar to our Rust-specific bug patterns.
+Later on, \cite{knuth89:_error_tex} analyzed the errors that he made in the development of \TeX, written in the WEB literate programming language.
+This was perhaps one of the first works in this domain where there is a publicly-available artifact (the \TeX~source code) supporting the data (i.e.
+the bugs identified), unlike the proprietary DOS/VS code.
+Knuth includes 15 categories of errors; the most relevant ones to this work are ``language liabilities,'' ``mismatches between modules,'' ``reinforcements of robustness,'' and ``surprising scenarios.'' \cite{flanagan1998new} described MrSpidey, an interactive static debugger for Scheme (a dialect of Lisp).
+MrSpidey is a front-end to a ESC-style static verification engine that aims to detect bug patterns where operators are passed invalid arguments.
+The engine relies on program invariants and presents information about attempted proofs of the invariants to the developer.
+Most recently, \cite{pan2009toward} analyzed seven large-scale widely used Java projects and manually extracted 27 common bug fix patterns.
+
+Moving from bug classification to our application, program repair, we follow the taxonomy of~\cite{liu2018survey} and propose two mindsets: search-based program repair and pattern-based program repair.
+Our work applies to both mindsets, but is most directly applicable to pattern-based program repair.
+
+\paragraph{Search-Based Program Repair.}
+One line of research in program repair follows the ``competent programmer'' hypothesis~\citep{gopinath2014mutant}: syntactically, a faulty program is not that far away from its correct version.
 This hypothesis suggests the search-based program repair mindset.
@@
 
-Use of mutation operators in program repair became popular when \cite{forrest2009genetic} and \cite{nguyen2009using} developed GenProg, a program repair tool that uses genetic programming.
+Use of mutation operators in program repair became popular with GenProg~(\cite{forrest2009genetic,nguyen2009using}), a program repair tool that uses genetic programming.
 The main idea behind this tool is that statements follow certain patterns in a codebase.
@@
 
-However, \cite{arcuri2011practical} observed that GenProg mostly found the patch while carrying out random initialization---that is, before GenProg's evolution begins.
+However, \cite{arcuri2011practical} observed that GenProg mostly found patches while carrying out random initialization---that is, before GenProg's evolution begins.
 Their tools TrpAutoRepair~\citep{qi2013efficient} and RSRepair~\citep{qi2014strength} advocate that test case prioritization is necessary while using genetic programming, as fitness evaluation is an expensive part of the repair pipeline.
@@
 
-\subsection{Pattern-based Program Repair and Mining Bug Fix Patterns}
-
+\paragraph{Pattern-based Program Repair and Mining Bug Fix Patterns.}
 The main intuition in pattern-based program repair methods is that bug fixes follow certain patterns.
 Thus, having a collection of common bug fix patterns is a crucial step for developing such repair tools.
-\cite{pan2009toward} analyzed seven large-scale widely used Java projects and obtained 27 common bug fix patterns.
-They used a bug fix pattern extractor tool to automatically parse and detect bug patterns within the bug fix hunks with less than seven statements.
-They ignored larger bug fix hunks as they tended to exhibit random changes and did not have observable meaningful patterns.
+For this mindset, the search is for a program fix (i.e.
+a delta) rather than for a correct program.
+
+As previously cited, \cite{pan2009toward}, used a bug fix pattern extractor tool to automatically parse and detect bug patterns within the bug fix hunks with less than seven statements.
+They ignored larger bug fix hunks as such hunks tended to exhibit random changes and did not have observable meaningful patterns.
 Though they did the analysis on Java projects, their reported patterns were not specific to Java.
@@
 
-We would characterize our work as the Rust version of what was done in~\cite{hanam2016discovering} and~\cite{yang2022mining}.
+We would characterize our work as the Rust version of what was done in~\cite{hanam2016discovering} and~\cite{yang2022mining} for JavaScript and Python, respectively.
 Like~\cite{hanam2016discovering}, we used DBSCAN to cluster fixes.
 Moreover, similar to~\cite{yang2022mining}, we introduced patterns in two different categories.
-In our case there were general and language-specific patterns; for Rust, the language-specific patterns were related to the borrow checker (BC patterns).
+In our case, there were general and language-specific patterns.
+The Rust language-specific patterns were related to the borrow checker (BC patterns).
 However, unlike our paper, \cite{yang2022mining} needed to do a much more elaborate manual analysis, as they had only collected general information about the single hunk changes (e.g.
@@
 
-\subsection{Automated Patch Correctness Assessment}
-
-Overfitting poses a substantial challenge for automated program repair (APR) systems, where synthesized patches may pass test cases but fail to be correct.
+\paragraph{Automated Patch Correctness Assessment.}
+The final step in automated program repair is to validate the proposed changes.
+One substantial challenge is overfitting, where synthesized patches may pass test cases but fail to be correct.
 Automated patch correctness assessment (APCA) aims to inspect the correctness of generated patches using dynamic and static analyses.
@@
 
-\cite{hoang2020cc2vec} introduced CC2Vec, a neural network model that utilizes log messages accompanying code changes to learn representations capturing semantic information.
+\cite{hoang2020cc2vec} introduced CC2Vec, a neural network model that uses log messages accompanying code changes to learn representations capturing semantic information.
 By addressing challenges in distinguishing between added and removed code and incorporating an attention mechanism, CC2Vec outperforms previous techniques for code change representation.
@@
 These vectors are then combined and fed into a hidden layer for predicting the target function.
-Evaluation across log message generation, bug fixing patch identification, and just-in-time defect prediction tasks demonstrates the superior performance of CC2Vec compared to state-of-the-art approaches in all three tasks.
+Evaluation across log message generation, bug fixing patch identification, and just-in-time defect prediction tasks demonstrates the superior performance of CC2Vec compared to the previous state-of-the-art approach in all three tasks.
 
-Our work tends to be more similar to that of~\cite{hanam2016discovering}.
-However, our novel program embedding approach differs from what they did on certain key decisions.
+Our work is more like that of~\cite{hanam2016discovering}.
+However, our novel program embedding approach differs from theirs on certain key decisions.
 In~\cite{hanam2016discovering}, they introduced the Feature Properties table, which is a categorization for program elements.
 Embedding programs based on this table resulted in an order-sensitive embedding.
-However, it made the datapoints sparse, and our datapoints are more dense than theirs.
+However, it made the datapoints sparse; our datapoints are more dense than theirs.
 More specifically, the number of our dimensions is more than 11$\times$ smaller.
@@
 \subsection{Refactoring tools for Rust}
-Clippy is widely known as the best code refactoring tool for Rust.
-It can identify over 600 common mistakes and offer automated fixes.
-It provides users with the flexibility to adjust the linting level, allowing them to control the number of lint suggestions they receive.
+Clippy is widely known as the best code refactoring tool for Rust. It can identify over 600 common mistakes and offer automated fixes.
+% true but doesn't add anything to our point
+%It provides users with the flexibility to adjust the linting level, allowing them to control the number of lint suggestions they receive.
 As discussed in Section~\ref{sec:bc_patterns}, we found that Clippy can detect and address five of our borrow-checker-related patterns.
-\cite{sam2017automated} introduced a refactoring tool that uses the Rust compiler infrastructure to perform various refactoring operations, including renaming variables, arguments, fields, functions, methods, structs, and enumerations, as well as inlining local variables, and reifying and eliding lifetime parameters.
+Beyond Clippy, \cite{sam2017automated} introduced a refactoring tool that uses the Rust compiler infrastructure to perform various refactoring operations, including renaming variables, arguments, fields, functions, methods, structs, and enumerations, as well as inlining local variables, and reifying and eliding lifetime parameters.
 In a more advanced use case, \cite{ling2022rust} proposed CRustS, a tool designed for refactoring C2Rust output.
@@
 Using the results of their manual analysis, they designed two bug detectors, and provided recommendations for developing more bug detectors in the future.
-Our automated approach to finding bug patterns has similar implications to their work but can operate at a far larger scale.
+Our approach to finding bug patterns has similar implications to their work but, due to automation, can operate at a far larger scale.
 
@@
 The tool exploits the insights obtained from manually observing existing bugs (by studying reported CVEs) in Rust code bases.
-Using both numerical (e.g.
+Using custom abstract domains that consider both numerical (e.g.
 integer bounds) and symbolic (e.g.
-modelling memory) information, the framework detects errors by using constraint solving techniques.
+modelling memory) information, the framework detects errors using constraint solving techniques.
 MirChecker detected 33 new bugs, including 16 memory safety faults, across 12 Rust crates.
+While MirChecker automatically detects bugs, the bug classes that it detects must be manually programmed into the system, and our work is complementary to MirChecker in potentially proposing novel classes of bugs that need to be fixed.

### 2023-11-12 response.tex
reread the edits

@@
 \vspace*{1em} \noindent \textbf{Response:} Yes, the 2022 Rust roadmap\footnote{\url{https://blog.rust-lang.org/inside-rust/2022/04/04/lang-roadmap-2024.html}} includes future proposed changes to Rust.
-It is indeed a threat to external validity, and we have now discussed it in our threats to external validity (Section 4.3); it could be mitigated by refreshing the analysis results once new versions of Rust are broadly used by extant packages.
+It is indeed a threat to external validity, and we have now discussed it in our threats to external validity (Section 4.3); this threat could be mitigated by refreshing the analysis results once new versions of Rust are broadly used by extant packages.
 
@@
 
-\vspace*{1em} \noindent \textbf{Response:} We added a new paragraph at the end of Section 4.1.
-We consider bugs as potentially more critical, because an unresolved bug can remain dormant within the system and unintentionally get deployed, whereas a program with compilation errors cannot be shipped by developers.
+\vspace*{1em} \noindent \textbf{Response:} We added a new paragraph at the end of Section 4.1: we consider bugs as potentially more critical.
+This is because a bug, if left unresolved, can remain dormant within the system and then inadvertently get deployed, whereas a program with compilation errors cannot be shipped by developers.
 
@@
 The nonterminals themselves don't directly correspond to code---it would be the terminals which are not present in the paths.
-We have also added more text to the caption of Figure 1 and point at the ExprReference in Path 1 in the text.
+We have also added more text to the caption of Figure 1 and point at the ExprReference in Path 1 in the text just below the Figure.
 
@@
 
-\textit{\textbf{R2:} Line 21 on page 8: work around -> workaround}
+\textit{\textbf{R2:} Line 21 on page 8: work around $\rightarrow$ workaround}
 
@@
 
-\vspace*{1em} \noindent \textbf{Response:} It is the rightmost path. We changed the color to blue so it is clearer and added a reference to ``Path 3'' as in the Figure.
+\vspace*{1em} \noindent \textbf{Response:} It is the rightmost path. We changed the color to blue so it is clearer and added a reference to ``Path 3'' to the text, to match with the Figure.
 

### 2023-11-12 discussion.tex
reread the edits

@@
 
-While comparing the importance of bugs and compilation errors can be challenging, we tend to consider bugs as potentially more critical.
-This is because a bug, if left unresolved, can remain dormant within the system and inadvertently get deployed, whereas a program with compilation errors cannot be shipped by developers.
+To compare the importance of bugs versus compilation errors: we consider bugs as potentially more critical.
+This is because a bug, if left unresolved, can remain dormant within the system and then inadvertently get deployed, whereas a program with compilation errors cannot be shipped by developers.
 

### 2023-11-12 methodology.tex
reread the edits

@@
 A DFS of the ASTDiff finds three paths of nonterminals (Paths 1--3), shown in three different colors at right.
-The first path summarizes the addition of the ExprReference {\tt \&} before tail, while the second and third paths indicate the change from a read of \texttt{head} to a method call, and a call to \texttt{clone()}, respectively.}
+The paths summarize (1) the addition of the ExprReference {\tt \&} before tail; (2) the change from a read of \texttt{head} to a method call (represented by ExprMethodCall), and (3) specifically a call to distinguished method \texttt{clone()}.}
 
@@
 This addition gives rise to two ASTDiff paths: one showing that the \verb+head+ node now belongs to an \verb+ExprMethodCall+, and one indicating the new sibling node for \verb+head+ representing the \verb+clone()+.
-(Among the terminals, we only keep \verb+clone+ in our paths, as it is a BC-related keyword.) One path for the \verb+&+ borrow operator and two for the call to \verb+clone()+ result in three paths in the ASTDiff, even if it only looks like we had made two modifications.
+(Among the terminals, we only keep the distinguished terminal \verb+clone+ in our paths, as it is a BC-related keyword.) One path for the \verb+&+ borrow operator and two for the call to \verb+clone()+ result in three paths in the ASTDiff, even if it only looks like we had made two modifications.
 
@@
 \begin{itemize}
-    \item bug-fix: The code change rectifies the undesired behavior in the software system, providing fixes for a specific bug type.
+    \item bug-fix: The code change rectifies undesired behavior in the software system, providing fixes for a specific bug type.
 We also include here changes that improve program performance, as they fix the behaviour of the software.

### 2023-11-12 related.tex
reread the edits

@@
 Before a repair tool modifies code, it will typically use a fault localization module to find fault locations.
-A fault localization module ranks program statements based on suspiciousness; one example is Tarantula by \cite{jones2005empirical}.
-Suspiciousness is calculated using a statistical model which relies on the observation that buggy statements are executed mostly by failed test cases~\citep{naish2009spectral,xie2013theoretical}.
+The fault localization modules that we are aware of rank program statements based on suspiciousness; one example is Tarantula by \cite{jones2005empirical}.
+Often, suspiciousness is calculated using a statistical model which relies on the observation that buggy statements are executed mostly by failed test cases~\citep{naish2009spectral,xie2013theoretical}.
 
@@
 We then present two mindsets for program repair: search-based and pattern-based.
-Finally, we survey related work in automated patch correctness assessment, which aims to ensure that the generated changes are valid.
+Finally, we survey related work in automated patch correctness assessment, which aims to ensure that changes by automated repair tools are valid.
 
@@
 Knuth includes 15 categories of errors; the most relevant ones to this work are ``language liabilities,'' ``mismatches between modules,'' ``reinforcements of robustness,'' and ``surprising scenarios.'' \cite{flanagan1998new} described MrSpidey, an interactive static debugger for Scheme (a dialect of Lisp).
-MrSpidey is a front-end to a ESC-style static verification engine that aims to detect bug patterns where operators are passed invalid arguments.
+MrSpidey is a front-end to an ESC-style static verification engine which aims to detect bug patterns where operators are passed invalid arguments.
 The engine relies on program invariants and presents information about attempted proofs of the invariants to the developer.
@@
 
-As previously cited, \cite{pan2009toward}, used a bug fix pattern extractor tool to automatically parse and detect bug patterns within the bug fix hunks with less than seven statements.
-They ignored larger bug fix hunks as such hunks tended to exhibit random changes and did not have observable meaningful patterns.
-Though they did the analysis on Java projects, their reported patterns were not specific to Java.
+\cite{pan2009toward} used a bug fix pattern extractor tool to automatically parse and detect bug patterns within bug fix hunks with fewer than seven statements.
+They ignored larger bug fix hunks---such hunks tended to exhibit random changes rather than having patterns that they could derive meaning from.
+Though they analyzed Java projects, their reported patterns were not specific to Java.
 \cite{martinez2015mining,martinez2012mining} exploited the frequency of observed patterns to introduce a heuristic patch searching method.

### 2023-11-12 discussion.tex
final pass

@@
 We discuss three threats to external validity: selection bias, incorrect commit/bug information reported by developers, and changes to the Rust programming language.
-Selection bias is an issue because our benchmarks may differ systematically from the set of Rust programs in the world.
+(1) Selection bias is an issue because our benchmarks may differ systematically from the set of Rust programs in the world.
 Our selection of projects might be biased and therefore we might not have presented patterns that may exist in other projects.
-Another threat to external validity can be the developers reporting commits as bug fixing commits while they are not really fixing functionality.
+We aimed to mitigate this threat by choosing all of the most-starred projects, but this biases towards popular projects by definition.
+(2) Another threat to external validity can be the developers reporting commits as bug fixing commits while they are not really fixing functionality.
 Similarly, a commit message might not contain our target keywords while the commit is associated with bug fixing changes.
-A third threat is that Rust is a relatively new programming language, and its syntax is prone to evolving over time.
+This is simply an underlying assumption of our work and may cause us to miss some bugs; we do not believe it should be a systematic threat, and is shared by much other work on understanding bug fixes.
+(3) A third threat is that Rust is a relatively new programming language, and its syntax is prone to evolving over time.
 Changes in Rust's grammar, such as the introduction of new keywords like \verb+async/await+ in version 1.39, have the potential to introduce new bug categories or modify existing ones.

### 2023-11-12 methodology.tex
final pass

@@
-In this section, we provide an overview of our methodology, providing full details in later subsections.
+In this section, we provide an overview of our methodology, with full details to appear in later subsections.
 Ruxanne uses a pipeline for finding classes of pervasive patterns in Rust programs.
@@
 Because we are interested in changes, we computed (using process metrics provided by our mining library) the average code churn of the files within the projects.
-swc-project/swc, nushell/nushell, and denoland/deno are the projects with largest number of LOCs.
+swc-project/swc, nushell/nushell, and denoland/deno are the projects with the most LOCs.
 Unsurprisingly, we captured a lot of instances from these projects in our final clusters.

### 2023-11-12 related.tex
final pass

@@
 They ignored larger bug fix hunks---such hunks tended to exhibit random changes rather than having patterns that they could derive meaning from.
-Though they analyzed Java projects, their reported patterns were not specific to Java.
+Though they analyzed Java projects, their reported patterns were not specific to Java, in the same way that our general patterns are not specific to Rust.
 \cite{martinez2015mining,martinez2012mining} exploited the frequency of observed patterns to introduce a heuristic patch searching method.
@@
 \cite{ye2021automated} propose ODS, an overfitting patch detection system, which assumes that code features capturing universal correctness properties can be utilized to classify overfitting patches across program repair systems and software projects.
-ODS extracts 202 manually crafted code features by comparing fixed and buggy code, and employs supervised learning to build a distributional model for classifying unseen patches, achieving over 70\% accuracy in evaluations on Defects4J, Bugs.jar, and Bears benchmarks.
+ODS extracts 202 manually crafted code features by comparing fixed and buggy code, and employs supervised learning to build a distributional model for classifying unseen patches, achieving over 70\% accuracy in evaluations on Defects4J, Bugs.jar, and Bears suites.
 Similarly, our feature extraction involves manual crafting of features.
@@
 However, our novel program embedding approach differs from theirs on certain key decisions.
-In~\cite{hanam2016discovering}, they introduced the Feature Properties table, which is a categorization for program elements.
+Their work introduced the Feature Properties table, which is a categorization for program elements.
 Embedding programs based on this table resulted in an order-sensitive embedding.
 However, it made the datapoints sparse; our datapoints are more dense than theirs.
-More specifically, the number of our dimensions is more than 11$\times$ smaller.
+Specifically, we have over 11$\times$ fewer dimensions than they do.
 

### 2023-11-12 results.tex
final pass

@@
 \noindent\textbf{Description:} Rust is a null-safe language, meaning that object references cannot take on null values.
-If a developer wants a variable to contain the equivalent of null, they would have to use the \texttt{Option} type.
+If a developer wants a variable to contain the equivalent of null, they must use the \texttt{Option} type.
 \texttt{Option} is essentially an \texttt{enum} with two variants: \texttt{Some}, which indicates that a variable has a value; and \texttt{None}, which indicates the absence of any value.
@@
 
-In Rust, the match control flow construct is used for comparing a value against multiple patterns, and executing the code associated with the first pattern that matches with our value.
+In Rust, the match control flow construct is used for comparing a value against multiple patterns, and executing the code associated with the first pattern that matches.
 A pattern and its associated code are called a match arm.
@@
 It also can be matching to a different enum variant or string.
-The fix pattern can happen both for refactoring or debugging purposes.
+The fix pattern can happen both for refactoring or bug-fixing purposes.
 \\
@@
 By definition, removing a clone that is indeed redundant does not modify functional aspects of the program behaviour.
-(Removing a non-redundant clone introduces a compile error or, worse yet, a bug.) However, such a removal can greatly improve the nonfunctional property of performance, if the program was formerly cloning large objects.
+(Removing a non-redundant clone introduces a compile error or, worse yet, a bug.) However, a removal of a redundant clone can greatly improve the nonfunctional property of performance, if the program was formerly cloning large objects.
 \\
@@
 Like the previous pattern, this change does not affect the program behaviour; it is done for performance purposes.
-In one of the commits\footnote{\scriptsize \url{https://github.com/tauri-apps/tauri/commit/a280ee90af0749ce18d6d0b00939b06473717bc9}} of the project tauri-apps/tauri, changing a clone to borrowing significantly improved CPU usage.
-In the commit history, developers discussed how the cloning of types such as HashMap was expensive.
+In one of the commits\footnote{\scriptsize \url{https://github.com/tauri-apps/tauri/commit/a280ee90af0749ce18d6d0b00939b06473717bc9}} of the project tauri-apps/tauri, changing a clone to borrowing significantly improved CPU usage: in the commit history, developers discussed how the cloning of types such as HashMap was expensive.
 \\

### 2023-12-12 discussion.tex
final edits

@@
 Our speculation is that, since Rust enforces strict compiler checks, programmers usually get many compile-time errors, but they only push their code once it is free of compile errors.
-We thus would not observe patterns which do not make it to the repository.
+We would not observe patterns which do not make it to the repository.
 
@@
 Also, the linter can analyze the usages of the struct to determine the necessity of adding the attribute.
-Removing unnecessary attributes not only optimizes the codebase but also eliminates unnecessary overhead and reduces binary size.
+Assuming that attributes cause additional code to be generated, removing unnecessary attributes not only optimizes the codebase but also eliminates unnecessary overhead and reduces binary size.
 

### 2023-12-12 introduction.tex
final edits

@@
 it has no dangling references or double-frees.
-Such properties had to be manually ensured (or not!) in languages like C/C++.
+Such properties must be manually ensured (or not!) in languages like C/C++.
 Moreover, not only does ownership remove the need for a garbage collector in the runtime environment, it frees the programmer from the responsibility to ensure certain safety-related properties~\citep{qin2020understanding}.
@@
 Although we are motivated by a desire to develop automated program repair tools, we believe these insights can also be of broader service.
-Automated program repair~\citep{goues2019automated,liu2018survey} tools try to find bug locations in the source code (using a fault localization module) and then generate patches to fix them, possibly without human intervention, so that the modified program meets desired specifications (implicit or explicit).
+Automated program repair (APR)~\citep{goues2019automated,liu2018survey} tools try to find bug locations in the source code (using a fault localization module) and then generate patches to fix them, possibly without human intervention, so that the modified program meets desired specifications (implicit or explicit).
 A powerful automated program repair system has many useful applications and could significantly simplify debugging, thus reducing software development cost~\citep{le2012systematic}.

### 2023-12-12 main.tex
capitalization
final edits

@@
 Ruxanne implements a novel embedding of Rust code into fixed-sized vectors.
-Using Ruxanne, we mined the top 18 most-starred Rust projects in Github to discover the most common bug fix patterns committed to their repositories.
+Using Ruxanne, we mined the top 18 most-starred Rust projects in GitHub to discover the most common bug fix patterns committed to their repositories.
 We analyzed 87,726 code changes drawn from 57,214 commits across these 18 projects.
-After clustering the code changes, and conducting a manual analysis, we discovered 20 groups of cross-project bug fix patterns, which we categorize as (1) general patterns and (2) borrow-checker-related patterns.
+After clustering the code changes, and conducting a manual analysis, we identified 20 groups of cross-project bug fix patterns, which we categorize as (1) general patterns and (2) borrow-checker-related patterns.
 Among the general patterns, the most frequently observed pattern is when the user either adds or removes struct fields.

### 2023-12-12 methodology.tex
final edits

@@
 
-In our example, we add a \verb+&+ borrow operator to the variable \verb+tail+, which gives rise to an ASTDiff path introducing a new parent node (\verb+ExprReference+) for \verb+tail+.
-We also add a call to \verb+clone()+.
+In our example, the change adds a \verb+&+ borrow operator to the variable \verb+tail+, which gives rise to an ASTDiff path introducing a new parent node (\verb+ExprReference+) for \verb+tail+.
+The change also adds a call to \verb+clone()+.
 This addition gives rise to two ASTDiff paths: one showing that the \verb+head+ node now belongs to an \verb+ExprMethodCall+, and one indicating the new sibling node for \verb+head+ representing the \verb+clone()+.
-(Among the terminals, we only keep the distinguished terminal \verb+clone+ in our paths, as it is a BC-related keyword.) One path for the \verb+&+ borrow operator and two for the call to \verb+clone()+ result in three paths in the ASTDiff, even if it only looks like we had made two modifications.
+(Among the terminals, we only keep the distinguished terminal \verb+clone+ in our paths, as it is a BC-related keyword.) One path for the \verb+&+ borrow operator and two for the call to \verb+clone()+ result in three paths in the ASTDiff, even if the change only appears to include two modifications.
 
@@
 
-For each cluster, we read the code and the bug reports of the datapoints (if they existed).
+For each cluster, we read the code and the bug reports of the datapoints (if available).
 If we judged that the members of a cluster fit into a pattern, we wrote a natural language description of that cluster.
 Considering the descriptions, we linked similar clusters together as possible candidates for merging, and removed the datapoints that were not either bug-fix or fix-inducing.
-We considered these candidates and carried out appropriate ones.
+We re-considered merging candidates and carried out appropriate merges.
 Finally, we selected 20 of the remaining clusters---the ones that manifested more-specific changes.

### 2023-12-12 related.tex
final edits

@@
 \cite{pan2009toward} used a bug fix pattern extractor tool to automatically parse and detect bug patterns within bug fix hunks with fewer than seven statements.
-They ignored larger bug fix hunks---such hunks tended to exhibit random changes rather than having patterns that they could derive meaning from.
+They ignored larger bug fix hunks---they claimed that such hunks tended to exhibit random changes rather than having patterns that they could derive meaning from.
 Though they analyzed Java projects, their reported patterns were not specific to Java, in the same way that our general patterns are not specific to Rust.
@@
 
-We would characterize our work as the Rust version of what was done in~\cite{hanam2016discovering} and~\cite{yang2022mining} for JavaScript and Python, respectively.
+We would characterize our work as the Rust version of what was done in~\cite{hanam2016discovering} and~\cite{yang2022mining} for JavaScript and Python, respectively, but with additional infrastructure to create a representation of the Rust code using fixed-size vectors.
 Like~\cite{hanam2016discovering}, we used DBSCAN to cluster fixes.

### 2023-12-12 results.tex
final edits

@@
 Figure~\ref{fig:rq1} shows the final results.
-Although we found that our results were acceptable---the result reported by our embedding was top-4 as ranked by humans 88\% of the time---we investigated the other 12\% as well.
+Although we judged that our results were acceptable---the result reported by our embedding was top-4 as ranked by humans 88\% of the time---we investigated the other 12\% as well.
 We found that our embedding tended to fail in change patterns where the change involves many elements (e.g.
@@
 We collected general fixes, asked DBSCAN to create clusters, and manually analyzed them.
-In this subsection, we will discuss each pattern, as well as one aspect of its value---its actionability.
+In this subsection, we will discuss each pattern, as well as an important aspect of its value: its actionability.
 By actionability, we mean the possible usages of the pattern for bug repair.
@@
 This concept more-or-less corresponds to interfaces in other languages, although the mapping is not exact (e.g.
-it is possible to implement traits for types that one did not initially define).
+it is possible to implement traits for others' types).
 This change pattern relates to modifications in trait bounds, which can happen as part of bug fixes or for refactoring.
@@
 A pattern and its associated code are called a match arm.
-Match statements can be similar to switch statements in other languages, like C, although match statements in Rust are able to match more complicated patterns.\\
+Match statements are similar to switch statements in other languages, like C, although Rust match statements are able to match more complicated patterns.\\
 
@@
 We collected 27143 datapoints for borrow-checker related fix patterns.
-We applied DBSCAN with parameters $\epsilon=0.001$ and $Z=5$, resulting in 102 clusters.
+We then applied DBSCAN with parameters $\epsilon=0.001$ and $Z=5$, resulting in 102 clusters.
 After manual analysis (Section~\ref{sec:manual_analysis_cluster_selection}), we ended up with 8 borrow-checker related fix patterns.
@@
 
-\noindent\textbf{Description:} The developer drops the mut keyword before a variable, as they do not mutate the variable.
+\noindent\textbf{Description:} The developer drops the mut keyword before a variable, as they do not mutate the variable (which is statically checkable).
 This redundant mutability is reported by Clippy and can also be removed by it.
-Such a change may help future-proof the code against unintended future changes.
+Such a change may help future-proof the code against unintended future changes: mutability must be explicitly added back before the variable can be mutated.
 
@@
 \noindent\textbf{Description:} In a struct field definition, or in the types of a function's formal parameters, type \verb+&Vec<T>+ changes to slice \verb+&[T]+.
-Clippy can detect the opportunity to apply this pattern.
+Clippy can detect opportunities to apply this pattern.
 This simplifies the code; slice types \verb+&[T]+ or \verb+&str+ are sufficient for most use cases.

### 2024-01-05 main.tex
boilerplate about preprint

@@
 
-\date{Received: date / Accepted: date}
+\date{Accepted: 10 December 2023}
 % The correct dates will be entered by the editor
@@
 \end{abstract}
+
+\vspace*{1em}
+\paragraph{Preprint.} This preprint has not undergone peer review (when applicable) or any
+post-submission improvements or corrections. The Version of Record of
+this article is published in \emph{Empirical Software Engineering}, and is available
+online at \url{https://doi.org/10.1007/s10664-023-10437-1}.
+
+To our knowledge, the publisher has not contributed any post-submission improvements or corrections, aside from some edits that we reverted.
 

### 2024-01-05 main.tex
change boilerplate

--- 2023-12-12 main.tex
@@
 \end{abstract}
+
+\vspace*{1em}
+\paragraph{Note.}
+This version of the article has been accepted for publication, after peer
+review (when applicable) but is not the Version of Record and does not reflect post-acceptance
+improvements, or any corrections. The Version of Record is available online at:
+\url{https://doi.org/10.1007/s10664-023-10437-1}. Use of this Accepted Version is subject to the publisher’s Accepted
+Manuscript terms of use https://www.springernature.com/gp/open-research/policies/accepted-
+manuscript-terms.
+
+To our knowledge, the publisher has not contributed any post-acceptance improvements or corrections, aside from some edits that we reverted.
 
